### 2019-09-09 cavity.tex
latest

\subsection{Cavity Flow}
The next numerical example considers hyper-reduced ROMs of a viscous, compressible flow in a two-dimensional cavity. 
The flow is described by the two-dimensional compressible Navier-Stokes equations,
\begin{equation}\label{eq:compressible_ns}
\frac{\partial \mathbf{u}}{\partial t} + \nabla \cdot \big( \mathbf{F}(\mathbf{u} ) - \mathbf{F}_v (\mathbf{u},\nabla \mathbf{u} )      \big) =\bz,
\end{equation}
where $\mathbf{u}: [0,T] \times \Omega \rightarrow \RR{4}$ comprise the density, x and y momentum, and total energy. The terms $\mathbf{F}$ and $ \mathbf{F}_v$ are the inviscid and viscous fluxes, respectively. For a two-dimensional flow, the state vector and inviscid fluxes are
$$
\mathbf{u} = \begin{Bmatrix}
\rho \\ \rho u_1 \\ \rho u_2 \\ \rho E \end{Bmatrix}, \qquad \mathbf{F}_{1} = \begin{Bmatrix} \rho u_1 \\ \rho u_1^2 +      p \\ \rho u_1 u_2 \\ u_1(E + p) \end{Bmatrix}, 
\qquad \mathbf{F}_{2} = \begin{Bmatrix} \rho u_2 \\ \rho u_1 u_2  \\ \rho u_2^2 + p \\ u_2(E + p) \end{Bmatrix}.
$$
The viscous fluxes are given by
$$
\qquad \mathbf{F}_{v_1} = \begin{Bmatrix} 0 \\ \tau_{11} \\ \tau_{12}  \\ u_j \tau_{j1} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_1}  \end{Bmatrix}, 
\qquad \mathbf{F}_{v_2} = \begin{Bmatrix} 0 \\ \tau_{21} \\ \tau_{22}  \\ u_j \tau_{j2} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_2}  \end{Bmatrix}.
$$
We assume a Newtonian fluid, which leads to a viscous stress tensor of the form,
\begin{equation*}
\tau_{ij} = 2\mu S_{ij},
\end{equation*}
where,
\begin{equation*}
 S_{ij} = \frac{1}{2} \big( \frac{\partial u_i}{\partial \mathsf{x}_j} + \frac{\partial u_j}{\partial \mathsf{x}_i} \big) - \frac{1}{3} \frac{\partial      u_k}{\partial \mathsf{x}_i} \delta_{ij}.
\end{equation*}
The Navier-Stokes equations are closed with a constitutive relationship for a calorically perfect gas,
$$p = (\gamma - 1)( \rho E - \frac{1}{2} \rho u_1^2 - \frac{1}{2} \rho u_2^2 \big),$$
where $\gamma = 1.4$ is the heat-capacity ratio.

Figure~\ref{fig:cav_fig} depicts the domain $\Omega$ and flow conditions. Free-stream boundary conditions 
are used at the inlet, outlet, and top wall of the cavity. No-slip boundary conditions are enforced 
on the bottom wall of the cavity.

\subsubsection{Description of FOM and Generation of Trial Space}
The full-order model comprises a discontinuous-Galerkin discretization. The discretization 
is obtained by partitioning the domain into $100$ elements in the flow direction and $40$ elements 
in the wall-normal direction. The solution is represented to third order over each element using tensor product polynomials of order $p=2$, 
resulting in $36000$ unknowns for each conserved variable. Spatial discretization via the discontinuous Galerkin method yields a dynamical system 
of the form,
$$\frac{d \state}{dt} = \mass_{\text{DG}} \velocity_{\text{DG}} (\state),$$
where $\mass_{\text{DG}} \in \RR{N \times N}$ is the DG mass matrix and $\velocity_{\text{DG}}: \RR{N} \rightarrow \RR{N}$ is the DG velocity operator containing 
surface and volume integrals. By the definition of the FOM~\eqref{eq:FOM}, the dynamical system velocity is $\velocity = \mass_{\text{DG}} \velocity_{\text{DG}}$. 
Figure~\ref{fig:cav_mesh} shows the computational mesh. The 
DG method leverages the Rusanov flux at the cell interfaces~\cite{rusanov} and uses the first form of Bassi and Rebay~\cite{br1} for the viscous fluxes. Time integration 
is performed via a third-order strong stability preserving Runge-Kutta method~\cite{ssp_rk3} with a time-step size of $\Delta t = 0.001$.

The reduced-order models leverage POD to construct the trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to construct the weighting 
matrix $\stweightingMatArg{n}$. The process used to construct the initial conditions and trial spaces for the ROMs are as follows:
\begin{enumerate}
\item Initialize the FOM with uniform free-stream conditions.
\item Evolve the FOM for $t \in [0,400]$.
\item Reset the time-coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 100$ and $K = \{136,193\}$ over $t \in [0,100]$ to construct two trial spaces comprising $\approx 95\%$ and $\approx 97\%$ of the snapshot energy, respectively. 
\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 100$, $n_s = 129$ to obtain the sampling point matrix of dimension $\stweightingMatOneArg{} \in \{0,1\}^{1603 \times \fomdim }$. 
\end{enumerate}
Figure~\ref{fig:cav_sampmesh} shows the resulting sample mesh used in the ROM simulations. To depict the nature of the flow, Figure~\ref{fig:fom_sols_cav} shows snapshots of the vorticity field generated by the FOM for several time instances used in training.  
\begin{table}[]
\begin{centering}
\begin{tabular}{c c c c}
\hline
Basis \# & Trial Basis Dimension ($K$) &  Energy Criterion & Sample Points ($n_S$) \\
\hline
%1    & $136$ & $95.0\%$ & $580$ \\
%2    & $193$ & $97.0\%$ & $580$ \\
1    & $136$ & $95.0\%$ & $1603$ \\
2    & $193$ & $97.0\%$ & $1603$ \\
\hline
\end{tabular}
\caption{Summary of the various basis sizes employed in the cavity flow example.}
\label{tab:rom_basis_details}
\end{centering}
\end{table}

\subsubsection{Description of Reduced-Order Models}
Hyper-reduced ROMs based on the Galerkin, collocated Least--Squares Petrov--Galerkin, and \methodAcronym\ approaches are 
again considered. Details on the implementation of the methods is as follows:
\begin{itemize}
\item \textbf{Galerkin ROM with Collocation}: The Galerkin ROM with collocation is obtained through Galerkin projection of the FOM. Unless 
otherwise noted, the  Galerkin ROM is evolved in time with the Crank--Nicolson time scheme at a time step of $\Delta t =0.1$.

\item \textbf{LSPG ROM with Collocation}: The LSPG ROM with collocation is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
discretization. Unless otherwise noted, the time step for the LSPG ROM is $\Delta t  = 0.1$. As in the previous example, 
the Gauss--Newton method is used to solve the non-linear least-squares problem arising at each time instance. The Gauss--Newton 
algorithm is deemed converged when the gradient norm is less than $1e-4$. All Jacobians are again computed via automatic differentiation.

\item \textbf{\methodAcronymROMs\ with Collocation}: \methodAcronymROMs\ are solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time-step size of 
$\Delta t = 0.1$. The Gauss--Newton algorithm is deemed converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation. Uniform quadrature 
weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.

%\begin{itemize}
%\item Direct Method: \methodAcronym ROMs solved via the direct method are again implemented with the same Crank--Nicolson time discretization with a nominal 
%time-step size of $\Delta t = 0.1$. 
%\end{itemize}
\end{itemize}

\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\includegraphics[trim={2cm 7cm 4cm 6cm},clip,width=0.95\linewidth]{cav_fig.pdf}
%\end{subfigure}
\caption{Figure depicting geometry and flow conditions of the cavity flow problem.} 
\label{fig:cav_fig}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[trim={0cm 14cm 0cm 14cm},clip,width=1.\linewidth]{figs/cavity/grid.png}
\caption{Computational mesh}
\label{fig:cav_mesh}
\end{center}
\end{figure}

\begin{figure} 
\begin{center}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.65\linewidth]{figs/cavity/hyper_grid.png}
\caption{Close up of computational mesh with highlighted collocation cells} 
\label{fig:cav_sampmesh}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0000.png}
\caption{$t=0.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0040.png}
\caption{$t=4.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0080.png}
\caption{$t=8.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0120.png}
\caption{$t=12.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0160.png}
\caption{$t=16.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0200.png}
\caption{$t=20.0$}
\end{subfigure}
\caption{Vorticity snapshots from the FOM simulation at various time instances.} 
\label{fig:fom_sols_cav}
\end{center}
\end{figure}

\subsubsection{Numerical Results}
We first assess the performance of \methodAcronymROMs\ with varying window sizes. To this end, we consider \methodAcronymROMs\ with uniform window 
sizes of $\Delta T^n \equiv \Delta T = 0.2,0.5,1.0,$ and $2.0$, along with the Galerkin and LSPG ROMs. Results are first considered for basis \#1 as described in 
Table~\ref{tab:rom_basis_details}. For all ROMs, we evolve 
the solution for $t \in [0,100]$. This comprises the same time interval used to construct the trial space. First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$ state error of the various reduced-order models. Both the hyper-reduced Galerkin and LSPG ROMs blow up/fail to converge within the first several time units. The \methodAcronymROM\ minimizing the residual over a window of size $\Delta T = 0.2$ also fails to converge. The \methodAcronymROMs\ that minimize the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are again less than $10\%$. The most notable discrepancy between the \methodAcronymROM\ and FOM solutions is a slight phase difference.

\begin{figure}
\begin{center}

\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.0cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/pressure_basis2.pdf}
\caption{Pressure} 
\label{fig:cav_results1a}
\end{subfigure}

\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 2.8cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/error_basis2.pdf}
\caption{Normalized $\elltwo$ error}
\label{fig:cav_results1b}
\end{subfigure}

\end{center}
\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various hyper-reduced ROMs to the full-order model solution.}
\label{fig:cav_results1}
\end{figure}

Figure~\ref{fig:cav_results2} shows the integrated error and objective function of the stable \methodAcronymROMs. It is seen that growing the window size over 
which the residual is minimized leads to a lower space--time residual, but not necessarily a lower $\elltwo$ error. This result is consistent with the previous numerical example. Next, Figure~\ref{fig:cav_wallclock}
shows the wall-clock times of the \methodAcronymROMs\ for $t \in [0,10]$ as compared to the LSPG ROM\footnote{Note that LSPG blew up at $t \approx 16.0$, so we focus on the first ten time units}. As expected, increasing the time-slab size again leads to an increase in computational cost. Minimizing the residual over a time-slab comprising 20 time-steps leads to a 2.5x increase in cost over LSPG.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window_basis2.pdf}
\caption{Integrated $\elltwo$ error}
\label{fig:cav_results2a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/objective_vs_window_basis2.pdf}
\caption{Objective function} 
\label{fig:cav_results2b}
\end{subfigure}
\end{center}
\caption{Integrated error (left) and objective function (right) as a function of window size.}
\label{fig:cav_results2}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.49\linewidth]{figs/cavity/walltime_vs_window_compare_basis2.pdf}
\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM}
\label{fig:cav_wallclock}
\end{center}
\end{figure}

Figure~\ref{fig:cav_snapshots} shows vorticity fields for the FOM, LSPG ROM, and \methodAcronymROM\ with $\Delta T = 0.5$ for the time instance $t = 5.0$. LSPG is observed to exhibit artificial oscillations; the Gauss-Newton method fails to converge at $t \approx 16.0$. The \methodAcronymROM\ at $\Delta T = 0.5$ is able to capture the important features of the flow, including the points of flow separation at the start and end of the ramp, and remains stable for the entire time interval.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_fom_t5_basis2.pdf}
\caption{FOM} 
\end{subfigure}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c2_t10.pdf}
%\caption{\methodAcronym, $\Delta T = 0.2$} 
%\end{subfigure}
\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_lspg_t5_basis2.pdf}
\caption{LSPG} 
\end{subfigure}
\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c5_t5_basis2.pdf}
\caption{\methodAcronym, $\Delta T = 0.5$} 
\end{subfigure}
%\begin{subfigure}[t]{0.95\textwidth}
%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c10_t10.pdf}
%\caption{\methodAcronym, $\Delta T = 2.0$} 
%\end{subfigure}
\caption{Field solutions for the vorticity at $t = 5.0$}
\label{fig:cav_snapshots}
\end{center}
\end{figure}

Next, we assess the performance of the various ROMs for basis \#2 as described in Table~\ref{tab:rom_basis_details}, which comprises a richer spatial basis.
 Figure~\ref{fig:cav_results1_basis1} shows the same pressure and error profiles as Figure~\ref{fig:cav_results1}, but for the enriched basis. The LSPG and Galerkin ROMs blow up faster as compared to Figure~\ref{fig:cav_results1}; LSPG fails to converge around $t \approx 8$ (opposed to $t \approx 16$), while Galerkin blows up almost immediately. The 
\methodAcronymROMs\ again yield improved performance: \methodAcronymROMs\ minimizing the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are less than $5\%$. The \methodAcronymROMs\ with $\Delta T \ge 0.5$ leveraging the enriched basis yield more accurate results than \methodAcronymROMs\ using the coarse basis.

\begin{figure}
\begin{center}

\begin{subfigure}[t]{0.95\textwidth}
\includegraphics[trim={0cm 2.5cm 0cm 2.5cm},clip,width=1.\linewidth]{figs/cavity/pressure.pdf}
\caption{Pressure} 
\label{fig:cav_results1a_basis1}
\end{subfigure}

\begin{subfigure}[t]{1.0\textwidth}
\includegraphics[trim={0cm 2.5cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/error.pdf}
\caption{Normalized $\elltwo$ error}
\label{fig:cav_results1b_basis1}
\end{subfigure}
\end{center}
\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various hyper-reduced ROMs to the full-order model solution.}
\label{fig:cav_results1_basis1}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window.pdf}
\caption{Integrated $\elltwo$ error}
\label{fig:cav_results3a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/objective_vs_window.pdf}
\caption{Objective function} 
\label{fig:cav_results3b}
\end{subfigure}
\end{center}
\caption{Integrated error (left) and objective function (right) as a function of window size.}
\label{fig:cav_results3}
\end{figure}

Finally, Figure~\ref{fig:cav_wallclock} shows the wall-clock times of the \methodAcronymROMs\ as compared to the LSPG ROMs for basis \#2. Increasing the window size again leads to an increase in computational cost. Minimizing the residual over a window comprising 20 time-steps leads to a 2.5x increase in cost over LSPG. This increase in computational cost is particularly reasonable in this example as LSPG yielded an unstable solution. 
\begin{figure}
\begin{center}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.49\linewidth]{figs/cavity/walltime_vs_window_compare.pdf}
\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM.}
\label{fig:cav_wallclock}
\end{center}
\end{figure}

%\begin{figure}
%\begin{center}
%\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window.pdf}
%\caption{Integrated $L^2$ error}
%\label{fig:cav_results2a_basis1}
%\end{subfigure}
%\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/objective_vs_window.pdf}
%\caption{Objective function} 
%\label{fig:cav_results2b_basis1}
%\end{subfigure}
%\end{center}
%\caption{Integrated error (left) and objective function (right) as a function of window size.}
%\label{fig:cav_results2_basis1}
%\end{figure}

### 2019-09-09 main.tex
latest

%\documentclass{article}
\documentclass[3p,computermodern,10pt]{elsarticle}
\input{packages}
\input{commands}
\input{commands_cwst}
%\input{title}
%\begin{frontmatter}
%\title{Continuous Least-Squares Petrov-Galerkin for Model Reduction}
%\author{Eric J. Parish}
\begin{document}
\begin{frontmatter}

%\title{Model Reduction via \methodName}
\title{Windowed Least--Squares Reduced-Order Models for Dynamical Systems}

\author[a]{Eric J. Parish and Kevin T. Carlberg}
%\ead{ejparis@sandia.gov}

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a windowed least-squares (\methodAcronym) reduced-order modeling formulation for dynamical systems. The proposed approach 
sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time windows.
Trial spaces with spatial and space--time dimension reduction are considered. 
The stationary conditions for the resulting minimization problems are derived. Of particular interest, the stationary conditions for trial spaces with spatial dimension reduction are obtained via the Euler--Lagrange equations and comprise a coupled two-point Hamiltonian boundary value problem containing a forward and backward system. %The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model. The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
%For spatial trial spaces, the stationary conditions for the time-continuous minimization problems are obtained 
%by deriving the associated Euler--Lagrange equations. 
Two differing methods are examined to solve the minimization problems: direct (discretize then minimize) and indirect (minimize and then discretize). Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov--Galerkin, and Space--Time Least-Squares Petrov--Galerkin. 
%By formulating the least--squares problem from the time-continuous level, the \methodAcronym\ formulation overcomes 
%the time-step sensitivity and time-scheme dependence that LSPG is subject to. 
%The stationary points of the minimization problems are obtained by solving the associated Euler-Lagrange equations. The Euler-Lagrange equations consist of a set of coupled two point boundary value problems defined on each time-slab containing a forward primal system and a backward adjoint system. %The approach is conceptually similar to the classic Least-Squares Petrov-Galerkin approach, but displays several advantages. First, by formulating the residual minimization problem at the continuous level, the sensitivity of the resulting method to the underlying time-discretization scheme is greatly reduced. 
Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.

%This work proposes a least-squares residual minimization approach for reduced-order models of dynamical systems. The proposed approach, which comprises a general least-squares reduced-order modeling (LSROM) formulation for dynamical systems, sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time slabs. The stationary conditions for the time-continuous minimization problems are obtained 
%by deriving the associated Euler--Lagrange equations. The time-continuous formulation exposes two differing methods that can be used to solve the minimization problem, namely direct (discretize then minimize) and indirect (minimize and then discretize). The proposed approach displays commonalities with optimal control problems and is a generalization of the popular 
%Least-Squares Petrov--Galerkin (LSPG) method. By formulating the residual minimization problem from the time-continuous level, the \methodAcronym\ approach overcomes 
%the time-step sensitivity and time-scheme dependence that LSPG is subject to. 
%The stationary points of the minimization problems are obtained by solving the associated Euler-Lagrange equations. The Euler-Lagrange equations consist of a set of coupled two point boundary value problems defined on each time-slab containing a forward primal system and a backward adjoint system. %The approach is conceptually similar to the classic Least-Squares Petrov-Galerkin approach, but displays several advantages. First, by formulating the residual minimization problem at the continuous level, the sensitivity of the resulting method to the underlying time-discretization scheme is greatly reduced. 
%Numerical experiments on the compressible Euler equations demonstrate that the proposed approach leads to more accurate and physically relevant solutions than existing model reduction approaches.

\end{abstract}
\end{frontmatter}

%\maketitle
\section{Introduction}
The simulation of dynamical systems plays an important role in a variety of areas in science and engineering; such examples include engineering design and control, financial analysis, and epidemiology. It is often the case, however, that (numerically) executing a single dynamical system solve is a computationally intensive process; e.g., this is the case when the state-space dimension of the dynamical system is prohibitively large and/or a wide range of time-scales need to be resolved. In such circumstances, it is common for analysts to rely on approximate models that generate approximations to the system of interest at a greatly reduced cost.

Projection-based reduced-order models (ROMs) have proved to be a promising tool for developing low-cost approximate solutions. These ROMs leverage ``snapshot" data from high-fidelity simulations and/or numerical experiments to identify a low-dimensional ``trial" space on which the state variables can be well approximated. ROM methods then seek a solution to the governing equations that lies on this low-dimensional trial space via projection or residual minimization. Efforts by a variety of research groups have led to significant advancements in ROM technologies over the past several decades, and we refer to Ref.~\cite{wilcox_benner_rev} for a review.

The construction of reduced-order models for nonlinear dynamical systems has proved challenging. The principle challenge encountered when trying to construct reduced-order models for such problems is that most existing methods lack \textit{a priori} guarantees of stability, accuracy, and convergence. %The standard Galerkin and least-squares Petrov-Galerkin methods, for instance, are both subject to exponentially growing error bounds~\cite{carlberg_lspg_v_galerkin}. 
A great deal of effort has been devoted towards developing methods that allow for more robust and accurate reduced-order models of dynamical systems. These efforts include residual-norm minimization~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis}, space--time methods~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st},  ``energy-based" inner products~\cite{rowley_pod_energyproj,Kalashnikova_sand2014}, symmetry transformations~\cite{sirovich_symmetry_trans}, basis adaptation~\cite{carlberg_hadaptation,adeim_peherstorfer}, projection subspace rotations~\cite{basis_rotation}, and subgrid-scale modeling~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}. Many of these developed techniques are not mutually exclusive. Residual minimization is most relevant to the current work and is the focus of the following review.

%~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in Refs~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}. Specifically, Refs~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}

%\begin{itemize}
%\item Galerkin is inaccurate/unstable for non-symmetric static systems -- motivates residual minimization
%\item Residual minimization (stems from FEM): minimizes the residual of the full-order model over the range of the trial space -- originally developed for static systems
%\item Can guarantee stability for linear static systems 
%\item Extension to time-varying systems is challenging 
%\begin{itemize}
%\item What is the definition of the ODE residual?
%\item Over what time-interval is the residual being minmized?
%\end{itemize}
%\end{itemize} 
Least-squares residual minimization as a numerical solution technique for (partial) differential equations was pioneered in the finite element community and led to what is known as the least-squares finite element method (LSFEM)~\cite{ls_review,bochev_leastsquares}. Motivated by the fact that the Galerkin method can be inaccurate and unstable for nonlinear and non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic functional comprising the residual norm of the partial differential equation of interest and (2.) leveraging variational principles to derive the stationary conditions of the quadratic functional; i.e, LSFEM can be thought of as finding the solution within a finite element trial space the minimizes the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous advantages (e.g., least-squares system are symmetric and positive definite) over the more standard Galerkin discretization.

Residual minimization methods were extended-to and developed-for reduced-order models by a variety of researchers. Similar to LSFEM, residual minimization approaches for reduced-order models comprise finding the solution within a low-dimensional trial space that minimizes the residual norm of the full-order model. Residual minimization approaches were first applied to static (i.e., no time-dependence) reduced-order models in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}. Specifically, Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg} formulated the residual minimization problem for dynamical systems by sequentially minimizing the \textit{fully discrete} residual (i.e., the residual after temporal discretization) at each time-step instance. This formulation is most commonly known as the \textit{Least--Squares Petrov--Galerkin} (LSPG) method. Ref.~\cite{carlberg_lspg_v_galerkin} provided numerous analyses on this formulation and examined connections with the standard Galerkin method and Petrov--Galerkin methods. Critically, it was shown that (1.) for various time-stepping methods LSPG has an equivalent continuous form that corresponds to a Petrov--Galerkin projection and (2.) the standard Galerkin method minimizes the time-instantaneous residual of the full-order model. Numerous numerical experiments have demonstrated that LSPG yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}. The common physical explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.

The LSPG approach only guarantees residual minimization over a single time-step, and as such is not associated with residual minimization of the full dynamical system.  
Indeed, LSPG is not equipped with \textit{a priori} guarantees of stability or convergence, even for linear time-invariant systems~\cite{bui_thesis}. Further, as it is defined at the fully discrete level, the performance of LSPG is subject to the time-step size: changing the time-step size changes the window over which the residual is minimized \textit{as well as} the temporal discretization error. As a result, LSPG provides optimal results at an intermediary time-step size. For too small a time-step, LSPG yields results similar to the Galerkin approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin. For too-large a time-step, the accuracy of LSPG degrades. It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over. The LSPG approach is additionally sensitive to the time discretization scheme. When used with explicit temporal discretization, for instance, LSPG again reverts to Galerkin. To gain the enhanced stability associated with LSPG, the underlying time-discretization scheme must be implicit with an appropriate time-step. This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.

%The thesis of least--squares residual minimization was extended to construct reduced-order models of static systems by several authors, with two such instances being~\cite{rovas_thesis} and~\cite{bui_resmin}. Similar to the finite-element case, numerous theoretical analyses and numerical experiments have demonstrated that residual minimization methods yield more accurate and stable solutions than the Galerkin method. The extension of least--squares residual minimization methods to dynamical systems, however, has proved challenging. The least-squares Petrov-Galerkin method is arguably the most commonly adapted residual minimization method for dynamical systems. LSPG extends the idea of least--squares residual minimization to dynamical systems by sequentially minimizing the \textit{fully discrete} residual (i.e., after temporal discretization) in a weighted norm (nominally $\elltwo$) over each time step. Numerous numerical experiments have demonstrated that the LSPG approach yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}. The common explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.

%The LSPG approach, however, is still subject to exponentially 
%growing error bounds and, in general, lacks \textit{a priori} guarantees of stability and accuracy. Further, as it is defined at the fully discrete level, LSPG contains several nuances. First, the LSPG approach is sensitive to the time-step. In Ref.~\cite{carlberg_lspg_v_galerkin}, LSPG was shown to provide optimal results at an intermediary time-step. For too small a time-step, the LSPG approach is prone to the same stability challenges as the Galerkin approach. In the limiting case of $\Delta t \rightarrow 0$, the LSPG approach reverts to a Galerkin approach. For too-large a time-step, on the other hand, the accuracy of the LSPG method degrades. It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over. The LSPG approach is additionally sensitive to the time discretization scheme. When used with explicit temporal discretization, for instance, the LSPG approach reverts to a Galerkin approach. To gain the enhanced stability associated with the LSPG method, the underlying time-discretization scheme must be implicit with an appropriate time-step. This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.

%Residual minimization methods are the most relevant to the current work. Residual minimization methods seek to minimize the residual of the full-order model over the trial space defined by the POD basis. Two of the most popular reduced-order modeling approaches are (or can be interpreted as) residual minimization approaches. The Galerkin method, which is arguably the most common reduced-order modeling approach, minimizes the instantaneous FOM residual over the range of the trial basis~\cite{carlberg_lspg_v_galerkin}. %The Galerkin method is arguably the most common reduced-order modeling approach. In Ref~\cite{carlberg_lspg_v_galerkin}, Carlberg and Barone demonstrated that the Galerkin method corresponds to finding a velocity that minimizies the instananeous FOM residual over the range of the trial basis. 
%Numerical experiments and theoretical analyses have shown, however, that the Galerkin method can be inappropriate for model-order reduction of general nonlinear dynamical systems~\cite{carlberg_gnat,carlberg_thesis,huang_combustion_roms}; error analysis has demonstrated that the Galerkin method is subject to exponentially growing error bounds while a variety of numerical experiments have demonstrated the lack of stability and inaccuracy of the method in practice. Indeed, it is well known in numerical analysis that Galerkin method performs poorly for non-symmetric systems. To address these issues, Carlberg et al. developed the least-squares Petrov-Galerkin (LSPG) method. LSPG sequentially minimizes the \textit{fully discrete} residual (i.e., after temporal discretization) in a weighted norm over each time-step. In this sense, the LSPG method is typically referred to as being \textit{discretely-optimal}. Numerous numerical experiments have demonstrated that the LSPG approach yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}. The common explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method. The LSPG approach, however, is still subject to exponentially 
%growing error bounds and, in general, lacks \textit{a priori} guarantees of stability and accuracy. Further, as it is defined at the fully discrete level, LSPG contains several nuances. First, the LSPG approach is sensitive to the time-step. In Ref.~\cite{carlberg_lspg_v_galerkin}, LSPG was shown to provide optimal results at an intermediary time-step. For too small a time-step, the LSPG approach is prone to the same stability challenges as the Galerkin approach. In the limiting case of $\Delta t \rightarrow 0$, the LSPG approach reverts to a Galerkin approach. For too-large a time-step, on the other hand, the accuracy of the LSPG method degrades. It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over. The LSPG approach is additionally sensitive to the time discretization scheme. When used with explicit temporal discretization, for instance, the LSPG approach reverts to a Galerkin approach. To gain the enhanced stability associated with the LSPG method, the underlying time-discretization scheme must be implicit with an appropriate time-step. This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.

%The nuances associated with the LSPG approach lead to practical challenges. For instance, minimal analysis on the optimal \textit{a priori} time-step exists, and as such specifying an appropriate time-step for LSPG \textit{a priori} is challenging. The limitation to implicit time-schemes is also restrictive, especially considering the fact that ROMs are often much less stiff than the corresponding FOM~\cite{Bach2018StabilityCF} and thus explicit time-stepping schemes with large time-steps can sometimes be used. It is emphasized that, despite these challenges, LSPG is arguably the most robust model-order reduction 
%formulation for general nonlinear dynamical systems and has gained wide-spread use~\cite{narmal,cheng,4,5,6}.\EP{Finish with good Refs using LSPG} 
%Time-discrete space-time residual minimization~\cite{choi_stlspg,constantine_strom} is the third approach that has been proposed. Reference~\cite{choi_stlspg} extends the LSPG method to space-time formulations and develops the ST-LSPG method. %ST-LPSG seeks a space-time solution within a space-time trial space that is a minimizer of the fully discrete residual. The space-time trial spaces
%ST-LSPG seeks to find the minimizer of the \textit{fully discrete} residual over the entire space-time domain within a given space-time trial space. These space-time trial spaces are constructed through higher-order singular value decompositions (SVD), such as the truncated high-order SVD.
%\item How can a finite-time residual minimization approach be developed that is independent of the underlying temporal discretization?
%\item Does minimizing the residual over a larger time-window improve the model prediction?
%\end{enumerate}

%Recently, LSROMs for dynamical systems have been extended to the space--time setting~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC}. These space--time LSROMs (ST-LSROMs), which are typically motivated from the need to reduce the spatial \textit{and} temporal dimension of a dynamical system model, comprise finding a space--time solution that minimizes the space--time residual within a low-dimensional space--time trial space. In contrast to standard ROMs for dynamical systems, in where the generalized coordinates are \textit{(discretized) time-dependent functions}, space--time methods ROMs represent the temporal dependence of the state via a finite dimensional temporal basis and thus the generalized coordinates are a stationary vector. ST-LSROMs thus differ from their standard LSROM counterparts in that they 1.) minimize the full space--time residual and 2.) represent the temporal dependence of the state via a prescribed trial space. ST-LSROMs thus comprise algebraic systems, while LSROMs comprise dynamical systems.

%ST-LSROMs involve a monolithic space--time least--squares solve, and are typically equipped with more robust error bounds than their non-space--time counterparts. To best describe how these existing methods fit into the present work, they must be viewed in a two dimensional space: one dimension corresponding the space--time trial space and one dimension corresponding to the space--time residual minimization problem.  With the exception of Ref.~\cite{bui_thesis}, existing space--time methods leverage reduced spatial and temporal trial spaces. In Ref.~\ref{choi_stlspg}, these trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. 
%Ref.~\cite{choi_stlspg} develops a space--time formulation of the least-squares Petrov-Galerkin approach. The formulation minimizes the \textit{fully discrete} residual over the entire space--time domain within a given space--time trial space. These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. 
%Another piece of relevant work is that performed by Constantine and Wang~\cite{constantine_strom}, in which a space--time residual minimization formulation for model interpolation is pursued. The present work again distinguishes itself in that it seeks to minimize the time-continuous residual over space--time slabs;~\cite{constantine_strom} considers the solution of a fully discrete optimization problem over the entire space--time domain. It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work.

This work seeks to overcome the practical challenges associated with existing residual minimization approaches for reduced-order models. To this end, we develop a reduced-order modeling approach that minimizes the FOM ODE residual over a series of time windows. The key novelties of the work are two fold. First, the residual minimization problem is formulated to minimize the \textit{time-continuous} residual. By working from a continuous viewpoint, the present formulation decouples the underlying temporal discretization scheme from the residual minimization problem; thus addressing a key deficiency of LSPG. The continuous viewpoint additionally exposes two differing methods that can be used to solve the 
minimization problems, namely \textit{discretize and then optimize} (direct) methods and \textit{optimize and then discretize} (indirect) methods. The second novelty is that the residual is minimized over \textit{time windows}, as opposed to over a single time step or the entire time domain. This allows for maximum flexibility in terms of computational cost and numerical accuracy. We demonstrate that this formulation recovers the Galerkin, LSPG, and space--time LSPG methods. Figure~\ref{fig:flowchart} depicts how the proposed method interfaces with existing approaches.

The resulting approach is best viewed as a hybrid space--time method. A variety of research has examined the formulation of space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis} and the present approach displays commonalities with several such efforts. First,~\cite{bui_thesis} briefly formulated a space--time least-squares ROM and connected this formulation with optimal control; specifically the thesis of optimize then discretize vs discretize then optimize was mentioned. This formulation, however, was not fully developed and was eschewed for the time-local residual minimization method (i.e., LSPG). The present work thus formally develops (and extends) several of the concepts put forth in~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a space--time residual minimization formulation for model interpolation. 
The present work distinguishes itself from Ref.~\cite{constantine_strom} in that (1.) in addition to trial spaces with space--time dimension reduction, we consider trial spaces with spatial dimension reduction only (i.e., a classic spatial basis is employed, opposed to a space--time basis) -- in which case the generalized coordinates comprise a vector of functions, and (2.) we minimize the time-continuous residual over space--time windows;~\cite{constantine_strom} considers the solution of a fully discrete optimization problem over the entire space--time domain. It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work. Lastly, Ref.~\cite{choi_stlspg} develops a space--time formulation of the least-squares Petrov--Galerkin approach. The formulation minimizes the \textit{fully discrete} residual over the entire space--time domain within a given space--time trial space. %These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. 
The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that (1.) we perform model reduction of the time-continuous system, (2.) we consider trial spaces with both space--time dimension reduction and spatial dimension reduction, and (3.) residual minimization is performed over time windows; Ref.~\cite{choi_stlspg} considers space--time bases and minimizes the fully-discrete residual over the entire domain.

%Ref.~\cite{choi_stlspg} develops a space--time formulation of the least-squares Petrov-Galerkin approach. The formulation minimizes the \textit{fully discrete} residual over the entire space--time domain within a given space--time trial space. These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that it seeks to minimize the time-continuous residual over time slabs, as opposed to the fully-discrete residual over the entire domain. Another point of departure from Ref.~\cite{choi_stlspg} is that the present work does not consider temporal reduction; this will be a subject of future work. Instead, the temporal basis is taken to have full support. Another piece of relevant work is that performed by Constantine and Wang~\cite{constantine_strom}, in which a space--time residual minimization formulation for model interpolation is pursued. The present work again distinguishes itself in that it seeks to minimize the time-continuous residual over space--time slabs;~\cite{constantine_strom} considers the solution of a fully discrete optimization problem over the entire space--time domain. It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work.

\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
\caption{Flow chart of \methodAcronym.} 
\label{fig:flowchart} 
\end{centering} 
\end{figure}

 The novel contributions and findings of the present work are:
\begin{enumerate}
%\item The development of a time-continuous residual minimization framework for reduced-order models.
\item The development of the \methodName\ (\methodAcronym) approach for dynamical systems. The approach sequentially minimizes the time-continuous full-order model residual over time windows.
\item The derivation of the stationary conditions for \methodAcronymROMs\ leveraging spatial and space--time dimension reduction trial spaces. The case of spatial dimension reduction trial spaces is of particular interest: the stationary conditions are derived via the Euler--Lagrange equations and comprise a coupled 
two-point Hamiltonian boundary value problem containing a forward and backward system. The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the ROM. The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate. 
\item An exploration of the use of both direct (discretize then optimize) and indirect (optimize then discretize) solution techniques. Conditions under which \methodAcronym\ recovers Galerkin, LSPG, and space--time LSPG are presented.
\item Error analysis of the \methodAcronym\ approach for trial spaces with spatial dimension reduction. In particular, it is shown that the error in \methodAcronymROMs\ is bounded \textit{a priori} by a combination of the error at the start of the window and the integrated FOM ODE residual evaluated at the FOM state projected onto the trial space. 
\item Numerical experiments for trial spaces with spatial dimension reduction demonstrating two key findings:
\begin{itemize}
\item Residual minimization over larger time windows leads to more stable and physically relevant solutions as well as lower space--time residuals. 
\item Minimizing the residual over a larger time window does not necessarily lead to a more accurate solution in the $\elltwo$ norm. Residual 
minimization over an intermediary-sized time-window leads to the lowest error in the $\elltwo$ norm.
\end{itemize}
\end{enumerate}

The layout of the rest of this paper is as follows: Section~\ref{sec:math} outlines the problem statement of interest, along with the Galerkin, LSPG, and space--time LSPG ROMs. Section~\ref{sec:tclspg} outlines the \methodAcronym\ formulation proposed in this work. Section~\ref{sec:numerical_techniques} outlines numerical techniques for solving \methodAcronymROMs. Both direct and indirect methods are discussed. Next, Section~\ref{sec:analysis} provides equivalency conditions and error analysis for \methodAcronymROMs. Section~\ref{sec:numerical_experiments} presents numerical experiments. Conclusions and perspectives are provided in Section~\ref{sec:conclude}.
The notation used is as follows: We denote vector functionals in stylized bold text, i.e., $\boldsymbol x$. We denote vectors in standard bold text, i.e., $\mathbf{x}$. Matrices are denoted with capital bold letters, i.e., $\mathbf{X} \equiv \begin{bmatrix} \mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$. Vector and functional spaces are denoted with calligraphic fonts, i.e., $\mathcal{X}$.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical Setting}\label{sec:math}
We consider the system of nonlinear ordinary differential equations (ODEs),
\begin{equation}\label{eq:FOM}
 \frac{d}{dt}\stateFOMArg{}{t} = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) = \stateFOMIC, \qquad t \in [0,T],
\end{equation}
where $T \in \mathbb{R}^+$ is the final time, $\stateFOM \in \RR{\fomdim} \otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$ denotes the state implicitly defined to be the solution to~\eqref{eq:FOM}, $\timeSpace$ is the set of functions acting on $[0,T]$, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ are the initial conditions, and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow \mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,t) \mapsto \velocity(\stateyDiscrete,t)$ is the velocity. Equation~\eqref{eq:FOM} will be referred to as the ``full-order model" (FOM) ODE.

Directly solving~\eqref{eq:FOM} can be computationally expensive due to both the high dimensionality of the state and the range of time-scales that needs to be resolved. For time-critical applications, it is important to devise computationally efficient approximations to~\eqref{eq:FOM}. Projection-based ROMs are 
one such promising approach.

\subsection{Projection-Based Reduced-Order Models}
Projection-based ROMs generate approximate solutions to the FOM ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial space. For dynamical 
systems, two distinct types of trial spaces are used: space--time trial spaces with spatial dimension reduction (\spatialAcronym) and space--time trial spaces with space--time dimension reduction (\spaceTimeAcronym). \spatialAcronym\ trial spaces are most commonly employed and are used in, e.g., the Galerkin and Least-Squares Petrov--Galerkin (LSPG) methods. \spaceTimeAcronym\ trial spaces are typically 
employed in space--time ROMs. The following subsections outline these two trial spaces, along with the Galerkin, LSPG, and 
space--time LSPG method.

\subsection{\spatialAcronym\ Trial Spaces}
\spatialAcronym\ trial spaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial space, 
\begin{align}\label{eq:spatial_subspace}
& \stspaceS \defeq \trialspace \otimes \timeSpace,\\ 
& \trialspace \defeq \text{Range}(\basismat) +  \stateIntercept,  \label{eq:spatial_subspace2}  \\
&\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, \label{eq:spatial_subspace3} 
\end{align}
where we refer to $\basismat \in \RRStar{\romdim}{\fomdim}$, with $\RRStar{\romdim}{\fomdim}$ denoting the set  $\{ \mathbf{X} \in \RR{\fomdim \times \romdim} \big| \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$, as the spatial trial basis matrix, $\trialspace$ as the spatial trial space, and $\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state. 
%where $\basisvec_i \in \mathbb{R}^N$ are orthonormal basis vectors and $\stateIntercept \in \mathbb{R}^{\fomdim}$ the ``intercept''.
%We refer to $\basismat \in \RRStar{\fomdim \times \romdim}$, with $\RRStar{p \times q}$ denoting the set of orthogonal $q$ frames in $p$, as the spatial trial basis, and $\trialspace$ as the spatial trial space. 
The basis vectors $\basisvec_i$ are typically constructed through, e.g., POD~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}, etc. 
ROMs leveraging \spatialAcronym\ trial spaces approximate the state as,
\begin{equation}\label{eq:affine_trialspace}
\stateFOMArg{}{t} \approx \approxstate(t) \defeq \basisspace \genstateArg{}{t} + \stateIntercept,
\end{equation}
where $\genstate \in \RR{\romdim} \otimes \timeSpace$, with $\genstate : [0,T] \rightarrow \mathbb{R}^{\romdim}$, are the generalized coordinates. Substituting the approximation given in~\eqref{eq:affine_trialspace} into~\eqref{eq:FOM} and performing orthogonal $\elltwo$ projection of the initial conditions onto the trial space, one obtains,
\begin{equation}\label{eq:g_truncation}
\basisspace \frac{d }{dt}\genstateArg{}{t} = \velocity(\basisspace
\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
\end{equation}
Equation~\eqref{eq:g_truncation} provides an over-determined system of $N$ equations for the $K$ generalized coordinates. Dimension reduction is predominantly performed with either the \textit{Galerkin method} or the \textit{Least-Squares Petrov--Galerkin method} (LSPG). We now outline these two methods.

\subsubsection{The Galerkin Method}
The Galerkin method is (traditionally) obtained by the following steps: (1.) The approximation~\eqref{eq:affine_trialspace} is substituted into~\eqref{eq:FOM}, and (2.) the resulting residual is enforced to be $\stweightingMatArg{}$-orthogonal to the trial space, with $\stweightingMatArg{ } \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ }$ being a positive (semi-)definite weighting matrix (commonly taken to be $\mathbf{I}$). This projection process results in what is known as the Galerkin method,
\begin{equation*}\label{eq:g_truncation2}
\frac{d}{dt}\genstateGalerkinArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
\end{equation*}
where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$ is a mass matrix. 
Model reduction via the Galerkin method is seen to result in a $K$-dimensional dynamical system for the 
generalized coordinates. 
The Galerkin method can alternatively be viewed as a residual minimization method~\cite{carlberg_lspg_v_galerkin}. In this setting, the Galerkin ROM can be 
written as,
$$
\frac{d}{dt}\genstateGalerkinArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}  }{\text{arg\,min}} || \basisspace \hat{\stateyDiscrete} -  \velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t) ||_{\stweightingMatArg{n}}^2. 
%\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
$$
%where
%$$\hat{ \mathbf{v}} = \underset{\hat{ \stateyDiscrete}  }{\text{arg min}} || \basisspace \hat{\stateyDiscrete} -  \velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept) ||_{\stweightingMatArg{n}}^2.$$
%The minimization problem has the solution,
%$$\hat{\mathbf{v}} = \massArgnt{n}^{-1}\basismat^T \stweightingMatArg{n} \velocity(\basismat \genstateGalerkinArg{}{t} + \stateIntercept ).$$
%Galerkin method is interpreted as finding the velocity that minimizes the FOM ODE residual~\cite{carlberg_lspg_v_galerkin},
% The velocity that minimize the FOM ODE residual is given as the solution to the optimization problem, %To demonstrate this, we define the FOM residual, %The LSPG residual corresponds to minimizing the FOM O$\Delta$ E residual at each time-instance. obtained either by projecting Eq.~\ref{eq:g_truncation} via a test basis or by residual minimization. Equivalence conditions for these two approaches are established in Ref.~\cite{carlberg_lspg_v_galerkin}. This work considers dimension reduction by residual minimization.
%The most common way to perform dimension reduction is by projecting Eq.~\ref{eq:g_truncation} via a test basis. An alternative philosophy that is pursued here is to perform dimension reduction through residual minimization. 
%Define the residual,
%\begin{align*}
%\resid & \vcentcolon \state \mapsto \frac{d \state}{dt} - \velocity(\state), \\
%& \vcentcolon \RR{\fomdim} \times [0,T] \rightarrow \RR{\fomdim} \times [0,T].
%\end{align*}
%The Galerkin method can be interpreted as finding the velocity such that the instantaneous time residual is minimized,%Minimization of the residual over the trial space leads to what is known as the Galerkin method,
%$$\mathbf{v} = \underset{\stateyDiscrete \in \text{Range}(\basismat)  }{\text{arg min}}|| \stateyDiscrete -  \velocity(\state(t)) ||_2^2.$$
%As $\mathbf{v} \in \text{Range}(\basismat)$, the minimization problem can be written as,%and hence $\boldsymbol v = \basismat \hat{\boldsymbol v} $,
%$$\hat{ \mathbf{v}} = \underset{\hat{ \stateyDiscrete}  }{\text{arg min}} || \basisspace \hat{\stateyDiscrete} -  \velocity(\state(t)) ||_2^2,$$
%where $\hat{\mathbf{v}} \in \RR{K}$ are the generalized coordinates of $\boldsymbol v$ and $\boldsymbol v = \basisspace \hat{\boldsymbol v}$. 
%The minimization problem has the solution~\cite{carlberg_lspg_v_galerkin},
%$$\hat{\mathbf{v}} = \basismat^T \velocity(\basismat \genstate(t) + \stateIntercept ).$$
%The Galerkin ROM is thus given by,
%\begin{equation*}\label{eq:g_truncation2}
%\frac{d}{dt}\genstateGalerkinArg{}{t} = \mass^{-1}\basisspace^T \stweightingMatArg{n} \velocity(\basisspace
%\genstateGalerkinArg{}{t} + \stateIntercept).
%\end{equation*}
The Galerkin ROM corresponds to \textit{time-instantaneous} minimization of the FOM ODE residual.
%To perform dimension reduction, 
%To perform dimension reduction, the governing equations are projected by a ``test" basis,
%\begin{equation}\label{eq:g_truncation}
%\testbasis^T \basisspace \frac{d \genstate}{dt} = \testbasis^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
%The most common choice is to set the test space and trial space to be equivalent, which results in the Galerkin reduced-order model,
%\begin{equation}\label{eq:g_truncation}
%\frac{d \genstate}{dt} = \basisspace^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
\subsubsection{Least-Squares Petrov--Galerkin}
The Galerkin method can be unstable and inaccurate for non-self-adjoint and nonlinear problems. 
An alternative model-reduction approach that enjoys several advantageous over the Galerkin method is the Least-Squares Petrov--Galerkin (LSPG) method~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}. LSPG is defined at the fully discrete level and minimizes the fully discrete FOM O$\Delta$E residual (i.e., the FOM ODE residual after temporal discretization) at each time-step instance.  We outline the LSPG method for linear multistep methods as a demonstrative example. Defining a discretization of the continuous time domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time step $\Delta t = T/N_t$, temporal discretization via a linear multistep method yields the FOM O$\Delta$E,
\begin{equation}\label{eq:lms}
\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, \qquad n=1,\ldots,N_t, 
\end{equation}
where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance and 
$\residLMSArg{n}$ is the discrete linear multistep residual at the $n$th time step,
\begin{align*}
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
\\
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim}. 
\end{align*} 
In the above, $k^n$ is the width of the multistep scheme at the $n$th time step and $\alpha^n_j,\beta^n_j$ are coefficients that define the type of multistep scheme at the $n$th time step.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.
The FOM O$\Delta$E solution will exactly satisfy~\eqref{eq:lms}; the ROM solution will not. The LSPG method minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial space over each time step. %We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
%\begin{align*}
%\resid_{\text{IE}} & \vcentcolon (\stateyDiscrete;\stateFOMDiscreteArg{n-1}) \mapsto \frac{\stateyDiscrete - \stateFOMDiscreteArg{n-1}  }{\Delta t} -  \velocity(\stateyDiscrete), \\
%& \vcentcolon \RR{\fomdim} \times \RR{\fomdim} \rightarrow \RR{\fomdim}. 
%\end{align*}
The LSPG method is defined as follows,
\begin{equation*}
\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,2,\hdots,\ntimeSteps,
\end{equation*}
where $\approxstateLSPG^n \approx \stateFOMDiscreteArg{n}$ denotes the LSPG approximation to the state at $n$th time-step instance and 
$\lspgWeightingArg{\cdot} \in \RR{\nsamples \times \fomdim}$, with $\romdim \le \nsamples \le \fomdim$, is a weighting matrix that can be used, e.g., for hyper-reduction. 
Although illustrated for linear multistep methods, LSPG can be 
formulated for \textit{any} type of time-marching scheme (see Ref.~\cite{carlberg_lspg_v_galerkin} for, e.g., Runge--Kutta schemes).

In general, the LSPG approach has been observed to be more stable and accurate than the 
Galerkin approach~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat}. The common interpretation of the improved performance of the LSPG method over the Galerkin method is that LSPG minimizes the residual over a finite-sized time window, while Galerkin minimizes only the instantaneous residual~\cite{carlberg_lspg_v_galerkin}.  
However, for general non-self-adjoint and nonlinear problems, LSPG, lacks \textit{a priori} guarantees of stability, accuracy, and convergence. Further, 
LSPG is sensitive to the time-step size: changing the time-step size changes the window over which the residual is minimized 
\textit{as well as} the temporal discretization error. As a result, optimal results are obtained at an intermediary time-step. 
For too small a time-step, the accuracy and stability of LSPG is similar to Galerkin; in the 
limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin~\cite{carlberg_lspg_v_galerkin}. 
For too large a time-step, the accuracy of LSPG degrades~\cite{carlberg_lspg_v_galerkin,parish_apg}. The 
precise reason for this degradation in accuracy is difficult to discern as increasing the time-step changes both the error incurred by 
temporal discretization and the window the residual is being minimized over. Lastly, LSPG is sensitive to the time-marching scheme: changing the 
time-marching scheme modifies the minimization problem. For example, LSPG reverts to a Galerkin approach 
when used with an explicit time-marching scheme. It is emphasized that, despite these challenges, 
LSPG is arguably the most popular reduced-order modeling technique for general nonlinear systems.

\subsection{\spaceTimeAcronym\ Trial Spaces and Space--Time ROMs}
Recent work has investigated space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}. Space--time methods 
are advantageous in that they allow for both spatial \textit{and} temporal dimension reduction. Further, space--time 
methods are often equipped with more favorable error bounds than non-space--time methods. The approach proposed in this work displays commonalities 
with space--time methods and, as such, we briefly outline \spaceTimeAcronym\ trial spaces and space--time LSPG.

Space--time methods leverage the \spaceTimeAcronym\ trial space $\stspaceST$,
\begin{equation}\label{eq:sttrialspace_def}
 \approxstateArgnt{} \in  \stspaceST \subseteq \RR{N} \otimes \timeSpace,
\end{equation}
and the space--time trial basis \textit{functions},
\begin{align*}
\stbasisArgnt{} &: t \mapsto \stbasis(t) \\
 &: [0,T] \rightarrow  \RR{N \times \stdim} ,%\spaceTrialSpace , \\
%&  \in \RR{N \times \stdim} \otimes \timeTrialSpaceArg{},  
\end{align*}
with $\text{Range}(\stbasis) + \stateInterceptArg{} \equiv \stspaceST$ and where $\stdim$ is the dimension of the space--time generalized coordinates.
%with $\mathbb{V}_{\stdim}( \RR{N} )$ denoting the set of orthogonal $\stdim$ frames in $\RR{N}$ (e.g., an $N \times \stdim$ matrix with orthogonal column vectors).
The state is approximated by,
\begin{equation}\label{eq:stapprox1}
 \stateFOMArg{}{t} \approx \approxstateArg{}{t}  = \stbasisArg{}{t} \stgenstate + \stateInterceptArg{},
\end{equation}
%where $\stbasisArgnt{} : \RR{\stdim} \rightarrow \spaceTrialSpace \otimes \timeTrialSpaceArg{}$ with $ \stbasisArgnt{n} \in \RR{N \times \stdim} \otimes \timeTrialSpaceArg{n}$ is the space--time trial basis, 
where $\stgenstateArg{} \in \RR{\stdim}$ are the space--time generalized coordinates. Critically, the 
space--time basis matrix comprises time-dependent \textit{functions}, while the generalized coordinates are a stationary vector. This is fundamentally different than  
\spatialAcronym\ trial spaces, in where the generalized coordinates are time-dependent functions while the basis matrix is stationary.

Substituting the approximation~\eqref{eq:stapprox1} into the FOM ODE~\eqref{eq:FOM},
\begin{equation}\label{eq:st_fom}
\frac{d }{dt} \stbasisArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
\end{equation}
where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial space (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG} 
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{\text{LMS}} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{LMS}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{LMS}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.

ST-LSPG is appealing in that it (1.) allows for both spatial and temporal dimension reduction and (2.) minimizes the residual over the \textit{entire} space--time domain (as opposed to a single time-step instance as in LSPG). As a result, ST-LSPG can potentially be more efficient than standard LSPG and is equipped with more favorable error bounds~\cite{choi_stlspg}. ST-LSPG, however, is subject to several practical challenges.  Most importantly, ST-LSPG can be computationally 
intensive as it involves (1.) computing space--time basis vectors and (2.) minimizing the entire space--time residual. For high-dimensional systems, these tasks may be intractable. Another challenge associated with ST-LSPG is that, as it requires a pre-defined temporal basis, it is unclear how it can be used, e.g., to perform future state prediction.

\subsection{Outstanding Challenges}
This work seeks to overcome the limitations of existing reduced-order modeling approaches. Specifically, 
we look to overcome LSPG's time-step and time-scheme dependence, the stability challenges of both Galerkin and LSPG, and the potential computational cost of ST-LSPG. To this end, we now outline the \methodName\ formulation. 
%\end{comment}
%\input{formulation_3}

%\input{space_time}
%\input{timecont_lspg}
\input{section3/tclsrm_formulation}
\input{section4/numerical_methods}
\input{analysis}
\input{numerical_experiments}

\section{Conclusions}\label{sec:conclude}
This paper introduced a windowed least-squares reduced-order modeling (\methodAcronym) formulation for dynamical systems. The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time windows. The formulation is applicable to both spatial and space--time dimension reduction trial spaces. The Galerkin, Least-Squares Petrov--Galerkin (LSPG) and space--time LSPG methods were shown to be specific instances of the proposed method

Two differing techniques to solve the least-squares minimization problems were considered: direct (discretize then minimize) and indirect (minimize then discretize). 
The case of \spatialAcronym\ trial spaces proved particularly interesting: it was shown that indirect methods comprise solving a coupled two-point 
Hamiltonian boundary value problem. The forward problem evolves the standard Galerkin ROM forced by a ``costate" variable, while a backward problem evolves the costate and is forced by the ROM residual.

Numerical experiments of the compressible Euler and Navier-Stokes equations demonstrated the utility in the proposed approach. The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded smoother and more physically relevant solutions. Further, increasing the window size over which the residual was minimized led to a monotonic decrease in the total space--time residual. Increasing the window size over which the residual was minimized, however, did not necessarily decrease the $\elltwo$ solution error; this was observed to occur over an intermediary window size. It was 
additionally shown that the \methodAcronymROMs\ overcome the time-step sensitivity challenges that LSPG is subject to. The second numerical experiment, which examined hyper-reduced ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow. In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up. Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower $\elltwo$ error.

%In the case of spatial dimension-reduction-only, \methodAcronym\ displays improved performance over LSPG at the expense of an increased computational cost. This increase in computational cost is due to the fact that forming and/or solving the nonlinear minimization problem becomes more difficult as the time-slab size grows. To minimize this increase in cost, we intend to pursue the development of new solution techniques tailored to \methodAcronym.

The principle challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to, in general, a higher computational cost. This increased cost is due to the fact that forming and solving the nonlinear least-squares problem becomes more expensive and difficult as the window size grows. While numerical experiments demonstrated that this increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.

%The second subspace we consider arrises when the decoder, $\decoder$, is linear and time-varying, while the generalized coordinates are time-stationary. This is the case in space-time discretizations. In this case one has,
%$$\state(t) \approx \approxstate(t)  =  \decoder(\genstate(t),t) = \basisst(t) \genstate + \state_0 $$
%where $\genstate \in \mathbb{R}^{\romstdim}$ are the time-stationary (space-time) generalized coordinates and the time-continuous basis is given by $\basisst \in \mathbb{R}^{romdim \times \romstdim} \times [0,T]$. The Euler-Lagrange equations are given by,
%\begin{equation}\label{eq:euler_lagrange_st}
% \bigg[ \dot{\basisst}^T \stweightingMat^T  - \basisst^T \big[\frac{\partial \velocity}{\partial \state} \big]^T \stweightingMat^T \bigg]  \bigg( \dot{\basisst} \genstate -    \velocity \bigg) = \mathbf{0}, \qquad \state(t=0) = \state_0\end{equation}
%Note that the end-point boundary condition in the Euler-Lagrange equations is automatically satisfied.

%\subsubsection{Solution techniques}
%Equation~\ref{eq:euler_lagrange_st} consists of a nonlinear algebraic system to be solved at each time-instance, $t$. To obtain solutions, we must employ a temporal discretization. Here, we propose using a variational discretization with tensor products, e.g. finite elements. Let,
%$$\basisst \equiv \basisspace \otimes \basistime \subseteq \basisspace \otimes \mathcal{T},$$
%where $\mathcal{T}$ is the finite elements space. Similarly, let
%$$\genstate = \genstate_s \otimes \genstate_t$$

\section{Acknowledgments}
The authors thank Yukiko Shimizo and Patrick Blonigan for numerous conversations from which this work benefited. 
E.\ Parish acknowledges an appointment to the Sandia National Laboratories'
John von Neumann Postdoctoral Research Fellowship in Computational Science. This paper
describes objective technical results and analysis. Any subjective views or
opinions that might be expressed in the paper do not necessarily represent the
views of the U.S.\ Department of Energy or the United States Government.
Sandia National Laboratories is a multimission laboratory managed and operated
by National Technology and Engineering Solutions of Sandia, LLC., a wholly
owned subsidiary of Honeywell International, Inc., for the U.S.\  Department of
Energy's National Nuclear Security Administration under contract
DE-NA-0003525.

\begin{appendices}
\section{Proper orthogonal decomposition}
Algorithm~\ref{alg:pod} presents the algorithm for computing the trial basis via proper orthogonal decomposition.
\begin{algorithm}
\caption{Algorithm for generating POD Basis.}
\label{alg:pod}
Input: Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$ \; 
%\newline
Output: POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
%\newline
Steps:
\begin{enumerate}
    \item Solve FOM O$\Delta$E and collect solutions into snapshot matrix
$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} - \stateInterceptArg{n} & \stateFOMDiscreteArg{N_{\text{skip}}} - \stateInterceptArg{n} & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} - \stateInterceptArg{n} \end{bmatrix}$$
    \item Compute the thin singular value decomposition, $$\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_K \end{bmatrix}$
\end{enumerate}

\end{algorithm}

\section{Selection of Sampling Points}
To construct the sampling point matrix used for hyper-reduction in the second numerical experiment, we employ q-sampling~\cite{qdeim_drmac} and the sample 
mesh concept~\cite{carlberg_gnat}. Algorithm~\ref{alg:qdeim} outlines the steps used in the second numerical experiment to compute the sampling points.

\begin{algorithm}
\caption{Algorithm for generating the sampling matrix through q-sampling.}
\label{alg:qdeim}
Input: Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$ \; 
%\newline
Output: Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneArg{}^T \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
%\newline
Steps:
\begin{enumerate}
    \item Solve FOM O$\Delta$E and collect velocity snapshots 
$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \velocity (\stateFOMDiscreteArg{0})  & \velocity( \stateFOMDiscreteArg{N_{\text{skip}}}) & \cdots & \velocity( \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} ) \end{bmatrix}$$
    \item Compute the thin singular value decomposition, $$\mathbf{F} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
    \item Compute the QR factorization of  $\mathbf{U}^T$ with column pivoting,
    \begin{equation*}
        \mathbf{U}^T \mathbf{P}^* = \mathbf{QR}
    \end{equation*}
    with $\mathbf{P}^* \equiv \begin{bmatrix} \mathbf{p}_1 & \cdots & \mathbf{p}_{\text{floor}(N_t/N_{\text{skip}})} \end{bmatrix}$, $\mathbf{p}_i \in \{0,1\}^N$. 
    \item Select the first $n_s$ columns of $\mathbf{P}^*$ to form the sampling point matrix $\stweightingMatOneArg{}^T \in \{0,1\}^{N \times n_s}$. 
    \item Augment the sampling point matrix, $\stweightingMatOneArg{}$, with additional columns such that all unknowns are computed at the mesh cells selected by Step 3. For the second numerical experiment, these additional unknowns correspond to each conserved variable and quadrature point in the selected cells. 
\end{enumerate}

\end{algorithm}

\section{Derivation of the Euler--Lagrange Equations}\label{appendix:eulerlagrange}
This section details the derivation of the Euler Lagrange equations. To this end, we consider the generic functional of the form,
\begin{equation}\label{eq:el_gen}
\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_a^b \integrand(\statey(t),\stateyDot(t),t)  dt,
\end{equation}
where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RR{}$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,t) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,t)$.%, $\statez: [a,b] \rightarrow \RR{M}$ and $\statezDot: [a,b] \rightarrow \RR{M}$.
 We now introduce the function $\statez: [a,b] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(a) = \statez_a$. %To this end, we start by supposing that $\statez$ is a stationary point of $\mathcal{J}$ that satisfies the boundary conditions. 
We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$ with the boundary condition $\variationArg{a} = \bz$ and define a variation from the extremal by,
\begin{align*}
\statezBar  &: (\tau;\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
&: \RR{} \times [a,b] \rightarrow \RR{M}.
\end{align*}
We note that $\statezBar$ obeys the same boundary conditions as $\statez$ since $\variationArg{a} = \bz$.
Evaluating the functional~\eqref{eq:el_gen} at $\statezBar(t,\epsilon)$,
$$
\mathcal{J}(\epsilon) = \int_a^b \integrand(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t)  dt.
$$
The objective is to now find $\overline{\statez}$ that makes $\mathcal{J}(\epsilon)$ stationary. This can be done by differentiating with respect to $\epsilon$, and setting the result to zero,% Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
$$
\frac{d}{d\epsilon} \bigg[ \mathcal{J}\big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) \big)\bigg] =  0,
$$
or equivalently,
$$
\int_a^b \frac{d}{d\epsilon} \bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big)\bigg] dt=  0.
$$
%As $\statez$ is a stationary point of $\mathcal{J}$, it then directly follows that,
%\begin{equation}\label{eq:el1_appendix}
%\bigg[ \frac{d \mathcal{J}}{d \epsilon} (\statezBar(t;0),\statezDotBar(t;0) \bigg]= 0.
%\end{equation}
%This fact can be used to derive the Euler--Lagrange equations. Evaluating Eq.~\eqref{eq:el1_appendix},
%$$ \int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) \big) \bigg] = 0. $$
We first evaluate the derivative. First using the chain rule,
$$ \int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t\big) \bigg] dt = \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t;\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t;\epsilon)\bigg]dt . $$
%$$\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \statezBar } \frac{\partial \statezBar }{\partial \epsilon}\big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) \big)  + \frac{\partial \integrand}{\partial \statezDotBar }\frac{\partial \statezDotBar}{\partial \epsilon} \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) \big)\bigg]dt = 0. $$
%Setting $\epsilon = 0$,
%$$\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \statezBar } \frac{\partial \statezBar }{\partial \epsilon}\big(\statezBar(t,0),\statezDotBar(t,0) \big)  + \frac{\partial \integrand}{\partial \statezDotBar }\frac{\partial \statezDotBar}{\partial \epsilon} \big(\statezBar(t,0),\statezDotBar(t,0) \big)\bigg]dt = 0. $$
Noting that,
$$\frac{\partial \statezBar}{\partial \delta} (\cdot,t)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(\cdot,t)= \variationDot(t),$$
we have,
$$\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt = \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variationDot(t) \bigg]dt. $$
%$$\int_a^b \bigg[\frac{\partial \integrand  }{\partial \overline{\genstate}} \variation + \frac{\partial \integrand}{\partial {\overline{\genstateDot}}}\variationDot \bigg]_{\epsilon = 0}dt = 0.$$
We integrate the second term by parts,
\begin{multline*}
\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt =  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon) ,b \big) \variation(b)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(a;\epsilon),\statezDotBar(a;\epsilon) ,a\big) \variation(a) .
\end{multline*}
Noting that $\variationArg{a} = \bz$,
\begin{multline*}
\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg] dt =  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) , \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon) ,b \big) \variation(b) . 
\end{multline*}
Setting equal to zero to find the stationary points,
$$
  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon),b \big) \variation(b)  = 0. 
$$
By definition, we know $\statez(\cdot) \equiv \statezBar(\cdot;0)$ is a stationary point. Thus, setting $\epsilon = 0$,
 $$
  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b) ,b\big) \variation(b)  = 0. 
$$
%$$  \int_a^b \bigg[\frac{\partial \integrand  }{\partial \overline{\genstate}} \variation  - \frac{d}{dt}\frac{\partial \integrand}{\partial {\overline{\genstateDot}}}\variation \bigg]_{\epsilon = 0}dt + \bigg[\bigg[ \frac{\partial \integrand}{\partial {\overline{\genstateDot}}} \variation \bigg]_a^b\bigg]_{\epsilon = 0} = 0. $$
%At $\epsilon = 0$ we have $\overline{\genstate} = \genstate$ and, as $\variationArg{a} = 0$, we obtain,
%$$ \int_a^b \bigg(\frac{\partial \integrand  }{\partial \genstate} \variation  - \frac{\partial \integrand}{\partial {\genstateDot}}\variation \bigg)dt +  \frac{\partial \integrand}{\partial {\genstateDot}} \variation \bigg|_{t=b} = 0.$$
As $\variation$ is an arbitrary function, in order for the equality to hold we must have,
%\begin{equation}\label{eq:euler_lagrange_appendix}
%\bigg[ \frac{\partial \integrand  }{\partial \genstate} \bigg]^T  - \bigg[\frac{\partial \integrand}{\partial {\genstateDot}}\bigg]^T = \boldsymbol 0, 
%\end{equation}
\begin{equation}\label{eq:euler_lagrange_appendix}
\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t) \big) \bigg) \bigg]^T = \bz  
\end{equation}
subject to,
$$
\statez(a) = \statez_a, \qquad \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \variation(b) \bigg]^T = \bz.
%\genstate(a) = \genstate_a, \qquad \bigg[ \frac{\partial \integrand}{\partial {\genstateDot}} \bigg]^T_{t=b} = \bz.
$$
Equation~\eqref{eq:euler_lagrange_appendix} is known as the Euler--Lagrange equation. It states that, if $\statez$ is an extremal of $\mathcal{J}$, then it must satisfy~\eqref{eq:euler_lagrange_appendix}. It is emphasized that~\eqref{eq:euler_lagrange_appendix} is a necessary condition on $\statez$ to make $\mathcal{J}$ stationary, but it is not a sufficient condition. It is additionally noted that~\eqref{eq:euler_lagrange_appendix} provides a stationary point of $\mathcal{J}$, but the resulting stationary point can be either a maximum or a minimum.

\section{Evaluation of Gradients in the Euler--Lagrange Equations for \methodAcronym}\label{appendix:vector_calc} 
To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:el1} for the integrand,
\begin{align*}\label{eq:integrand_apx}
 \minintegrand & \vcentcolon
(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,t) \mapsto \frac{1}{2} \big[
\basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
+ \stateIntercept,t) \big]^T \stweightingMatArg{} \big[
\basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
\stateIntercept,t) \big], \\ & 
\vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
 \rightarrow \RR{} .  
\end{align*}
To evaluate the gradients, we first expand $\minintegrand$:
\begin{align*}
 \minintegrand(\genstateyDiscrete,\dot{\genstateyDiscrete},t)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,t) \big]^T \stweightingMatArg{n} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\ 
 &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{n}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big].
\end{align*}
Since $\stweightingMat$ is symmetric,
\begin{equation}\label{eq:int_expand}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,t)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{n}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big].
\end{equation}
For notational purposes, we write the above as,
$$  \minintegrand(\genstateyDiscrete,\dot{\genstateyDiscrete},t) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,t) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,t) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,t),$$
where
\begin{align*}
&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,t) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{n}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,t) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big], \\
&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,t) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] . 
\end{align*}
Constructing the Euler--Lagrange equations for this functional $\minintegrand$ requires evaluating the derivatives $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot}$. We start by evaluating $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and go term by term.

Starting with $\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot)$, we see,
$$\frac{\partial \minintegrand_1 }{\partial \genstateyDiscrete} = \boldsymbol 0,$$
where it is noted that $\mathcal{I}_1$ only depends on $\genstateyDiscreteDot$. Working with the second term:
\begin{align*}
\frac{\partial \minintegrand_2}{\partial \genstateyDiscrete}  &= -\frac{\partial}{\partial \genstateyDiscrete} \bigg( \big[ \basisspace \genstateyDiscreteDot \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] \bigg) \\ 
&= - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \frac{\partial}{\partial \genstateyDiscreteArgnt{}} \bigg( \big[ \velocity(\veloargsromy) \big]  \bigg)\\
 &= -\big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n}   \frac{\partial \velocity}{\partial \stateyDiscrete}\frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
& = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace \\
&= -[\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete}  \basisspace.
\end{align*}
For $\minintegrand_3$,
\begin{align}\label{eq:i3sol}
\frac{\partial  \minintegrand_3}{\partial \genstateyDiscrete}  &= \frac{1}{2} \frac{\partial }{\partial \genstateyDiscrete} \bigg( \big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n}\big[ \velocity(\veloargsromy) \big] \bigg) \\  
&=  [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
 &=   [\velocity(\veloargsromy) ]^T  \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
\end{align}
This gives the final expression,
$$
 \frac{\partial \minintegrand}{\partial \boldsymbol \genstateyDiscrete} = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace +  [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
$$%
We now evaluate $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDotArgnt{}}$ and again go term by term. Starting with $\minintegrand_1$,
\begin{align*}
\frac{\partial \minintegrand_1}{\partial \genstateyDiscreteDotArgnt{}} &=
\frac{1}{2} \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} }\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \stweightingMatArg{n}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \bigg) \\
&= [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{n} \basisspace.
\end{align*}
Now working with the second term:
\begin{align*}
\frac{\partial \minintegrand_2}{\partial \genstateyDiscreteDotArgnt{} } &=
\frac{\partial}{\partial \genstateyDiscreteDotArgnt{}  } \bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \bigg)\\
 &= \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} } \bigg( [\genstateyDiscreteDotArgnt{}]^T \bigg) \basisspace^T  \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] \\ 
 &= \bigg[  \basisspace^T  \stweightingMatArg{n}  \velocity(\veloargsromy )   \bigg]^T \\
 &= [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace.
\end{align*}
Finally, for the last term,
\begin{align*}
\frac{\partial \minintegrand_3}{\partial \genstateyDiscreteDotArgnt{} } &= \boldsymbol 0,
\end{align*}
where it is noted that $\mathcal{I}_3$ only depends on $\genstateyDiscrete$. We thus have,
$$\frac{\partial \minintegrand}{\partial \dot{ \genstateyDiscrete} } =   [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{n} \basisspace -  [\velocity(\veloargsromy ) ]^T \stweightingMatArg{n} \basisspace.$$
Combining all terms, the Euler--Lagrange equation reads,
\begin{multline*}
 \frac{\partial \minintegrand}{\partial \genstateyDiscrete}(\genstateArg{n}{t},\genstateDotArg{n}{t},t)  - \frac{d}{dt} \bigg[ \frac{\partial \minintegrand}{\partial \genstateyDiscreteDotArgnt{}} (\genstateArg{n}{t},\genstateDotArg{n}{t},t) \bigg] =  - \big[ \basisspace \genstateDotArg{n}{t}  \big]^T \stweightingMatArg{n} \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \big] \basisspace \\ + [\velocity(\veloargsromn) ]^T \stweightingMatArg{n} \big[ \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \big] \basisspace - \frac{d}{dt} \bigg[  [\genstateDotArg{n}{t} ]^T  \basisspace^T \stweightingMatArg{n} \basisspace - [\velocity(\veloargsromn) ]^T \stweightingMatArg{n} \basisspace \bigg]  = \boldsymbol 0.
\end{multline*}
To put this in a more recognizable form, we can pull out the common factor in the first two terms,
$$ - \bigg( \big[ \basisspace \genstateDotArg{n}{t} \big]^T  -  [\velocity(\veloargsromn)] ^T \bigg) \stweightingMatArg{n}
 \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisspace - \frac{d}{dt} \bigg[  [\genstateDotArg{n}{t}]^T \basisspace^T \stweightingMatArg{n} \basisspace -  
[\velocity(\veloargsromn) ]^T \stweightingMatArg{n} \basisspace  \bigg] = \boldsymbol 0.$$
Taking the transpose to put into the common column major format,
$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)  \big]^T \stweightingMatArg{n} \bigg( \big[ \basisspace \genstateDotArg{n}{t}\big]  -  \velocity(\veloargsromn)  \bigg) -  \frac{d}{dt} \bigg[  \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArg{n}{t}  - \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn)   \bigg] = \bz.
 $$
We now factor the second term,
$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)  \big]^T \stweightingMatArg{n} \bigg(  \basisspace \genstateDotArg{n}{t}  -  \velocity(\veloargsromn) \bigg) -  \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg[   \basisspace \genstateDotArg{n}{t}  - \velocity(\veloargsromn) \bigg] = \bz. $$
Gathering terms and multiplying by negative one, the final form of the Euler Lagrange equations are obtained,
$$ \bigg[\basisspace^T \big[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \big]^T \stweightingMatArg{n} + \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{n}{t} -  \velocity(\veloargsromn) \bigg) = \bz. $$
This is the \methodAcronym-ROM. Note that this is a second order equation and can be written as two separate first order equations. Equation~\ref{eq:clspg_2ord} can be written equivalently as two first order equations by introducing an auxiliary ``costate" variable. Defining the costate as,
\begin{equation*}
%\adjoint^n(t) \defeq \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArg{n}{t}  -  \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) ,
\adjoint^n(t) \defeq  \genstateDotArg{n}{t}  -  \mass^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) ,
\end{equation*}
we can manipulate Eq.~\ref{eq:clspg_2ord} as follows: %First, to enable hyper-reduction later on, we leverage the fact that $\stweightingMat^2 = \stweightingMat$,
%$$
% \bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)\bigg]^T \stweightingMat^n \stweightingMat^n + \basisspace^T \stweightingMat^n \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}^n   -  \velocity(\veloargsromn) \bigg) = \bz.
%$$
First, we add and subtract the first term multiplied by $\basisspace [\massArgnt{n}]^{-1}\basisspace^T \stweightingMat$, 
$$
 \bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1} \basisspace^T\stweightingMatArg{n} + \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  + \basisspace^T \stweightingMatArg{n}  \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}^n   -  \velocity(\veloargsromn) \bigg) = \bz.
$$
Pulling out the term multiplied by the positive portion of $\basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}$,
\begin{multline*} 
\bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1}  \basisspace^T  \stweightingMatArg{n} \bigg)  + \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  \basisspace^T \stweightingMatArg{n} +   \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg) = \bz.
\end{multline*}
Splitting into two separate terms, 
\begin{multline*}
\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1}  \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg)  + \\  
\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} +   \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{n}{t}  -  \velocity(\veloargsromn) \bigg) = \bz.
\end{multline*}
Pulling $\mass^{-1} \basisspace^T \stweightingMatArg{n}$ inside the parenthesis on the second term,
\begin{multline*}
\basisspace^T \bigg[\frac{\partial \velocity}{\partial \state^n}(\veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg)  + \\  
\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace  +   \mass \frac{d}{dt} \bigg] \bigg( \genstateDotArg{n}{t}  - \mass^{-1} \basisspace^T \stweightingMatArg{n}  \velocity(\veloargsromn) \bigg) = \bz.
\end{multline*}
By definition, the term inside the parenthesis of the second term is $\adjointArg{n}{t}$,
\begin{multline*}
\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}    -  \velocity(\veloargsromn) \bigg)  +   \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace \adjointArg{n}{t}  +  \massArgnt{n} \frac{d}{dt} \adjointArg{n}{t}= \bz.
\end{multline*}
Re-arranging,
\begin{multline*}
\mass \frac{d }{dt}\adjointArg{n}{t} + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsromn ) \bigg]^T \stweightingMatArg{n} \basisspace  \adjointArg{n}{t}  \\
= - \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)\bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg).
\end{multline*}
We thus get the splitting,
%Eq.~\ref{eq:clspg_2ord} can be split as,
%\begin{equation}
\begin{align*}\label{eq:lspg_continuous_appendix}
& \mass \frac{d}{dt}\genstateArg{n}{t}  -  \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) =  \mass \adjointArg{n}{t},\\
&\mass \frac{d}{dt} \adjointArg{n}{t}  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace \adjointArg{n}{t} = \\
& -\basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \big] ^T \stweightingMatArg{n}  \bigg( \mathbf{I} -   \basisspace [\massArgnt{n}]^{-1} \basisspace^T   \stweightingMatArg{n} \bigg)\bigg( \basisspace \genstateDotArg{n}{t}   -   \velocity(\veloargsromn) \bigg) . 
\end{align*}
\section{Evaluation of Gradients for Optimal Control Formulation}\label{appendix:optimal_control}
When formulated as an optimal control problem of Lagrange type, the gradients of the Hamiltonian with respect to the state, controller, and costate 
need to be evaluated. This section details this evaluation.

The Pontryagin Maximum Principle leverages the following Hamiltonian,
\begin{align*}
\hamiltonian^n \; &: \;  (\genstateyDiscreteArgnt{},\adjointDiscreteDumArgnt{},\controllerDiscreteDumArgnt{},t) \mapsto 
 \adjointDiscreteDumArgnt{T} \bigg[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \bigg] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t) \\
&: \; \RR{\romdim} \times \RR{\romdim} \times \RR{\romdim} \times [0,T] \rightarrow \RR{},
\end{align*} 
where,
\begin{align*}
 \objectiveControlArg{} &:  (\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t)
\mapsto \frac{1}{2} \bigg[ \basisspace \bigg(  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \bigg) -
\velocity(\veloargsromy) \bigg]^T
\stweightingMatArg{n}  \\ & \hspace{1.5 in}\bigg[ \basisspace \bigg(
[\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{}
\bigg) - \velocity( \veloargsromy ) \bigg]
, \nonumber \\ & : \RR{\romdim} \times \RR{\romdim} \times [0,T] \rightarrow \RR{}.
\end{align*} 
As described in Section~\ref{sec:optimal_control}, to derive the stationary conditions of the \methodAcronym\ objective function, we require evaluating the following gradients,
$ \frac{\partial \hamiltonianArg{n}}{\partial \adjointDiscreteDumArgnt{}{}}, \;   \frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}}$, and $\frac{\partial \hamiltonianArg{n}}{\partial \controllerDiscreteDumArgnt{}{}}.$  Starting with $\frac{\partial \hamiltonianArg{n}}{\partial \adjointDiscreteDumArgnt{}{}}$, we have,
$$
\bigg[\frac{\partial \hamiltonianArg{n}}{\partial \adjointDiscreteDumArgnt{}{}}\bigg]^T = [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} .
$$
Next, we address $ \frac{\partial \hamiltonianArg{n}}{\partial \genstateyArgnt{}{}}$. 
\begin{align*}
  \frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}} &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
 &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy
) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
 &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg( \velocity(\veloargsromy)\bigg) \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
 &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscreteArgnt{}{}}{\partial  \genstateyDiscreteArgnt{}{} } \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
 &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg]. 
\end{align*}
To evaluate $\frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg]$, we leverage the previous result~\eqref{eq:int_expand} and insert $\genstateyDiscreteDotArgnt{} =  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} .$ This leads to the expression for the expanded Lagrangian,
\begin{multline*}
 \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) = \frac{1}{2}\big[\basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T  \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t}\big] \\  -  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big].
\end{multline*}
Again for notational purposes, we split this into three terms,
$$
\objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg)  = 
\objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg)  + 
\objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg)  + 
\objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) ,$$
where
\begin{align*}
& \objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \equiv  \frac{1}{2}\big[\basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t}  \big]^T  \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy)   +  \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} \big],\\
& \objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \defeq -  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{n}]^{-1} \controllerDiscreteDumArgnt{n}\big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] ,\\
& \objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \defeq \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] .
\end{align*}
Evaluating the first term,
\begin{align*}
\frac{\partial  \objectiveControlArg{}_1 }{\partial \genstateyDiscreteArgnt{}{}} &= 
\frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) +  \\
& \qquad \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{n}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) + \\
& \qquad \frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \controllerDiscreteDumArgnt{}\bigg) 
,\\ 
&= 
[\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace + 
[\controllerDumArgnt{n}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \statey} \basisspace , \\ 
%&= 
%[\velocity(\basisspace \genstateyArgnt{} + \stateIntercept)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  
% \frac{\partial \velocity}{\partial \statey} \basisspace + 
%[\controllerArgnt{n}]^T [\massArgnt{n}]^{-1} \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
%\stweightingMatArg{n} \frac{\partial \velocity}{\partial \statey} \basisspace , \\  
&= 
\bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \bigg)\basisspace^T  \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
&= 
[\genstateyDiscreteDotArgnt{} ]^T \basisspace^T  \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace , \\
&= 
[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace .
\end{align*}
Now evaluating the second term,
\begin{align*}
\frac{\partial  \objectiveControlArg{}_2 }{\partial \genstateyDiscreteArgnt{}{}} &= 
-\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}   \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\&= 
-\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}    [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T 
 \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  - 
  [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace 
\\
&= 
- 2[\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  
 \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace - 
 [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
&= - \bigg[ 2[\velocity(\veloargsromy)]^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T  [\massArgnt{n}]^{-1} \bigg] \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
&= -\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \bigg) \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace. 
\end{align*}
Next, we can use the result~\eqref{eq:i3sol} to have,
$$ \frac{\partial  \objectiveControlArg{}_3 }{\partial \genstateyDiscreteArgnt{}{}} =
  [\velocity(\veloargsromy) ]^T  \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
$$
Thus we have,
\begin{align*}
\frac{\partial \objectiveControlArg{}}{\partial \genstateyDiscreteArgnt{}} &= 
[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)]^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T  \bigg) \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace + 
 [\velocity(\veloargsromy) ]^T  \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
&= 
\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{n}
\frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace \\ 
&= 
\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{n} \basisspace  [\massArgnt{n}]^{-1}\basisspace^T - \mathbf{I} \bigg)
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, 
\end{align*}
such that,
\begin{equation*}
\frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}} = 
  \adjointDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big]  + 
\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{n} \basisspace  [\massArgnt{n}]^{-1}\basisspace^T - \mathbf{I} \bigg)
\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
\end{equation*}
Or, equivalently,
\begin{equation*}
\bigg[ \frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}} \bigg]^T =  \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T \stweightingMatArg{n}  \basisspace [\massArgnt{n}]^{-1}  \adjointDiscreteDumArgnt{} + 
 \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T 
  \stweightingMatArg{n} \bigg(  \basisspace  [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n}- \mathbf{I} \bigg)
 \bigg( \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \bigg) 
\end{equation*}
Next, we evaluate $\frac{\partial \hamiltonianArg{n}}{\partial \controllerDumArgnt{}}$:
\begin{align}
  \frac{\partial \hamiltonianArg{n}}{\partial \controllerDumArgnt{}{}} &= \frac{\partial}{\partial \controllerDumArgnt{}{}} \bigg[  \adjointDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t) \bigg] \\
 &= \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
 &= [\adjointDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} + \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t) \bigg]. 
\end{align}
To evaluate $\frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg]$, we again 
go term by term. Starting with the first term,
\begin{align*}
\frac{\partial  \objectiveControlArg{}_1 }{\partial \genstateyDiscreteArgnt{}{}} &= 
\frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) + \\
& \qquad \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{n}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) + \\
& \qquad \frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \controllerDiscreteDumArgnt{}\bigg) \\
&= \bigg(  [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T                      
\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg)^T +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \\
&= \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}. 
\end{align*}
Moving on to the second term,
\begin{align*}
\frac{\partial  \objectiveControlArg{}_2 }{\partial \controllerDiscreteDumArgnt{}{}} &= 
-\frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}}   \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
\stweightingMatArg{n}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\
&=- \bigg[ [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace \genstateyDiscrete + \stateIntercept) \bigg]^T \\
&= - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}. 
\end{align*}
For the third term we have simply,
$$ \frac{\partial  \objectiveControlArg{}_3 }{\partial \controllerDiscreteDumArgnt{}{}} = \bz .$$
Thus,
\begin{align*}
\frac{\partial \hamiltonianArg{n}}{\partial \controllerDiscreteDumArgnt{}} &= 
[\adjointDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1}  +  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} + \\
& \qquad   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}, \\
& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace \bigg) [\massArgnt{n}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  \\
& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace \bigg) [\massArgnt{n}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{n}]^{-1}  \\
& =  [\adjointDiscreteDumArgnt{}]^T  [\massArgnt{n}]^{-1}  +  [\controllerDiscreteDumArgnt{}]^T   [\massArgnt{n}]^{-1}  
\end{align*}
Or, equivalently,
\begin{equation*}
\frac{\partial \hamiltonianArg{n}}{\partial \controllerDiscreteDumArgnt{}} = 
  [\massArgnt{n}]^{-1} [\adjointDiscreteDumArgnt{} +  \controllerDiscreteDumArgnt{}]. 
\end{equation*}
Thus, the conditions stated via the Pontryagin Maximum Principle are,
\begin{align*}%\label{eq:hamiltonian_sysa}
&\frac{d}{dt}\genstateArg{n}{t} =  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromn) + [\massArgnt{n}]^{-1}\controllerArg{n}{t}\\ 
&\frac{d }{dt} \adjointArg{n}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n}  \basisspace [\massArgnt{n}]^{-1}  \adjointArg{n}{t} = 
 - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n} 
\bigg(  \basisspace  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n} - \mathbf{I} \bigg)
 \bigg( \basisspace \genstateDotArg{n}{t} - \velocity(\basisspace \genstateArg{n}{t} + \stateIntercept) \bigg) \\ 
& \adjointArg{n}{t}= -\controllerArg{n}{t}.
\end{align*}
Equivalently,
\begin{align*}%\label{eq:hamiltonian_sys}
&\frac{d}{dt}\genstateArg{n}{t} =  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\basisspace \genstateArg{n}{t} + \stateIntercept) + [\massArgnt{n}]^{-1}\controllerArg{n}{t} \\ 
&\frac{d}{dt} \controllerArg{n}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n}  \basisspace [\massArgnt{n}]^{-1}  \controllerArg{n}{t} = 
 - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n} 
\bigg( \mathbf{I} -   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T \bigg)
\stweightingMatArg{n} \bigg( \basisspace \genstateDotArg{n}{t} - \velocity(\basisspace \genstateArg{n}{t} + \stateIntercept) \bigg) 
\end{align*}

\begin{comment}

====================
To do this, first write the above as,
$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T + \basisspace \basisspace^T \bigg)  \stweightingMat^T + \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0. $$
Re-arranging gives,
$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg]  +  \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T  \basisspace   +  \frac{d}{dt} \bigg] \bigg( \basisspace^T \stweightingMat \basisspace \dot{\genstate}   -  \basisspace^T \stweightingMat  \velocity \bigg) = 0. $$
Defining the adjoint as,
\begin{equation}
\basisspace^T \stweightingMat \basisspace \frac{d \genstate}{dt}  -  \basisspace^T \stweightingMat \velocity =  \boldsymbol \lambda , \qquad \genstate(t=0) = \genstate_0,
%\basisspace \frac{d \genstate}{dt}  -   \velocity =  \adjoint  , \qquad \genstate(t=0) = \genstate_0
\end{equation}
we then get the auxiliary adjoint equation,
\begin{equation}
%\basisspace^T \stweightingMat^T \frac{d}{dt} \adjoint  + \basisspace^T \frac{\partial \velocity}{\partial \genstate}^T \stweightingMat^T \adjoint = 0, \qquad \basisspace^T \stweightingMat \adjoint(t=T) = 0
 \frac{d}{dt} \adjoint  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \state} \bigg]^T \basisspace \adjoint = -\bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg] , \qquad \adjoint(t=T) = 0.
\end{equation}
\end{comment}
\begin{comment}
=================================

%We start $ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate}$. We go term by term and note the useful link: \url{https://en.wikipedia.org/wiki/Matrix_calculus#Layout_conventions}
%First term
\begin{enumerate}
\item First term: Only depends on $\dot{\genstate}$, term is zero
\item Second term: $$\frac{\partial}{\partial \genstate} \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)=  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \frac{\partial}{\partial \genstate} \bigg( \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg)$$
$$ = \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat   \frac{\partial \boldsymbol f(\state)}{\partial \state}\frac{\partial \state}{\partial \genstate} $$
$$ =  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace $$
%$${\color{red} \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big] \bigg) = \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \bigg) \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \basisspace.$$
%$${\color{red} ?  = \big[ \frac{\partial \velocity}{\partial \state}\frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
%$${\color{red} ?  = \big[ \basisspace \frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
\item Third term: 
$$\frac{1}{2} \frac{\partial }{\partial \genstate} \bigg( \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)  =  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
$$ =   \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisspace$$
\end{enumerate}
This gives us,
$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} = - \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace +  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace $$

Now we work on $\frac{\partial \minintegrand}{\partial \dot{ \genstate } }$:
\begin{enumerate}
\item First term:
$$\frac{1}{2} \frac{\partial}{\partial \dot{\genstate}}\bigg( \big[\dot{\genstate}^T  \basisspace^T \big]  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  \bigg) =  \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace$$
\item Second term:
$$\frac{\partial}{\partial \dot{\genstate} } \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg) = \frac{\partial}{\partial \dot{\genstate} } \bigg( \dot{\genstate}^T \bigg) \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  $$
$$ = \bigg[  \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg]^T $$
$$ = \velocity^T \stweightingMat^T \basisspace$$
\end{enumerate}
This gives,
$$\frac{\partial \minintegrand}{\partial \dot{ \genstate } } =   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace -  \velocity^T \stweightingMat^T \basisspace.$$
%\end{comment}

\begin{comment}
Let $\stweightingMat = \mathbf{I}$,
$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} - \frac{d}{dt} \bigg[ \frac{\partial \boldsymbol F}{\partial \dot{\genstate}} \bigg] =   \bigg( -\big[ \basisspace \dot{\genstate} \big]^T  +  \velocity^T \bigg) \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T  - \velocity^T \basisspace \bigg] $$
Transposing,
$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[   \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[ \basisspace^T \basisspace  \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
$$ =    -\basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( \basisspace \dot{\genstate}  -  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
$$ = -\bigg[ \basisspace^T \frac{d}{dt} + \basisspace^T \big[ \frac{\partial \velocity}{\partial \state} \big]^T  \bigg] \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$ 
Adjoint style
$$ =   \basisspace^T  \frac{d }{dt}\lambda  +   \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \lambda  $$
$$ \basisspace  \dot{\genstate}  -\velocity = \lambda$$
%Gathering terms

\end{comment}
%\begin{comment}

\begin{comment}

\subsection{Time-varying spacetime subspace}
We have $\state(t)= \basisst(t) \genstate + \state_0$. Note that, in this case, the generalized coordinates are stationary in time, $\dot{\genstate} = \mathbf{0}$. Thus we only need to compute $\frac{\partial \minintegrand}{\partial \genstate}$. Going term by term:

\begin{enumerate}
\item First term: 
$$\frac{1}{2} \frac{\partial}{\partial \genstate} \bigg(\big[\frac{\partial \decoder}{\partial t}  \big]^T \stweightingMat  \big[\frac{\partial \decoder}{\partial t} \big]\bigg) = 
\frac{\partial}{\partial \genstate} \bigg(\big[\frac{\partial \basisst}{\partial t} \genstate \big]^T \stweightingMat  \big[\frac{\partial \basisst}{\partial t} \genstate \big]\bigg) $$
$$
= \genstate^T \dot{\basisst}^T \stweightingMat \dot{ \basisst }
$$
\item Second term: $$\frac{\partial}{\partial \genstate} \bigg( \big[ \frac{\partial \decoder}{\partial t} \big]^T \stweightingMat \big[ \velocity(\decoder,t;\param) \big] \bigg)= \frac{\partial}{\partial \genstate} \bigg( \big[ \frac{\partial \basisst}{\partial t} \genstate  \big]^T \stweightingMat  \big[ \velocity(\decoder,t;\param) \big]  \bigg)$$
$$ = \big[ \frac{\partial \basisst}{\partial t} \genstate  \big]^T \stweightingMat \frac{\partial}{\partial \genstate}\bigg( \big[ \velocity(\decoder,t;\param) \big] \bigg) +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T \frac{\partial}{\partial \genstate} \bigg( \big[ \frac{\partial \basisst}{\partial t} \genstate  \big] \bigg)$$
$$ = \big[ \dot{\basisst} \genstate  \big]^T \stweightingMat \frac{\partial \velocity}{\partial \decoder} \frac{\partial \decoder}{\partial \genstate} +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T  \dot{ \basisst}$$
$$ = \big[ \dot{\basisst} \genstate  \big]^T \stweightingMat \frac{\partial \velocity}{\partial \decoder} \basisst +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T  \dot{ \basisst}$$
\item Third term: 
$$\frac{1}{2} \frac{\partial }{\partial \genstate} \bigg( \big[\velocity(\decoder,t;\param) \big]^T \stweightingMat \big[ \velocity(\decoder,t;\param) \big] \bigg)  =  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
$$ =   \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisst$$
Putting it all together we get,
$$ \genstate^T \dot{\basisst}^T \stweightingMat \dot{\basisst} - \bigg(  \big[ \dot{\basisst} \genstate  \big]^T \stweightingMat \frac{\partial \velocity}{\partial \decoder} \basisst +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T  \dot{ \basisst} \bigg) + \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisst = \mathbf{0}$$
Transposing, we obtain the Euler--Lagrange equations for a space-time basis
$$ \basisst^T \stweightingMat^T \dot{\basisst} \genstate - \bigg( \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \dot{\basisst} \genstate + \dot{\basisst}^T \stweightingMat \velocity  \bigg) + \basisst^T  \big[ \frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \velocity  = \mathbf{0}, \qquad \state(t=0) = \state_0.$$
Note that the additional boundary condition is automatically satisfied.
$$ \dot{\basisst}^T \stweightingMat^T \bigg( \dot{\basisst} \genstate -    \velocity \bigg)   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \dot{\basisst} \genstate  + \basisst^T  \big[ \frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \velocity  = \mathbf{0}, \qquad \state(t=0) = \state_0.$$
$$ \dot{\basisst}^T \stweightingMat^T \bigg( \dot{\basisst} \genstate -    \velocity \bigg)   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \bigg( \dot{\basisst} \genstate  -  \velocity \bigg) = \mathbf{0} $$
$$ \bigg[ \dot{\basisst}^T \stweightingMat^T  - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \bigg]  \bigg( \dot{\basisst} \genstate -    \velocity \bigg) = \mathbf{0}$$

Note that we can write this as two systems. To put it in the best form, write the above as,
$$ \dot{\basisst}^T \bigg( \mathbf{I} -  \basisst \basisst^T  +  \basisst \basisst^T \bigg)  \stweightingMat^T \bigg( \dot{\basisst} \genstate -    \velocity \bigg)   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \bigg( \mathbf{I} -  \basisst \basisst^T  +  \basisst \basisst^T \bigg)\stweightingMat^T \bigg( \dot{\basisst} \genstate  -  \velocity \bigg) = \mathbf{0} $$
Defining,
$$  \basisst^T \stweightingMat^T \dot{\basisst} \genstate  -  \basisst^T \stweightingMat^T \velocity = \boldsymbol \lambda,$$
we have,
$$ \bigg[ \dot{\basisst}^T   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \bigg] \basisst \adjoint  = -\bigg[\dot{\basisst}^T   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \bigg]  \bigg( \mathbf{I} -  \basisst \basisst^T  \bigg)\stweightingMat^T \bigg( \dot{\basisst} \genstate  -  \velocity \bigg)$$

%The above can be written as two systems,
%$$  \dot{\basisst}^T \stweightingMat^T \dot{\basisst} \genstate - \dot{\basisst}^T \stweightingMat \velocity  = \boldsymbol \lambda $$
%$$  \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \dot{\basisst} \genstate  - \basisst^T  \big[ \frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \velocity  = \lambda$$

\end{enumerate}
\end{comment}

\begin{comment}
\textbf{Linear subspace with full temporal support}:
For a linear subspace with full temporal support, we have $\state = \basisspace \genstate + \state_0$
$ \frac{\partial \state}{\partial \genstate} = \basisspace.$
We have,
$$\minintegrand = \frac{1}{2} \big[ \basisspace \dot{\genstate} \big]^T  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  -  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] 
  + \frac{1}{2} \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]$$
Write this as,
$$\minintegrand = \minintegrand_1 + \minintegrand_2 + \minintegrand_3 $$

%other form
%$${\color{red} = \big[ \basisspace \dot{\genstate} \big]^T  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  - 2 \big[ \velocity(\state(\genstate),t;\param) \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big] + \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] }$$

First we compute $ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate}$. We go term by term and note the useful link: \url{https://en.wikipedia.org/wiki/Matrix_calculus#Layout_conventions}
%First term
\begin{enumerate}
\item First term: Only depends on $\dot{\genstate}$, term is zero
\item Second term: $$\frac{\partial}{\partial \genstate} \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)=  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \frac{\partial}{\partial \genstate} \bigg( \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg)$$
$$ = \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat   \frac{\partial \boldsymbol f(\state)}{\partial \state}\frac{\partial \state}{\partial \genstate} $$
$$ =  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace $$
%$${\color{red} \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big] \bigg) = \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \bigg) \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \basisspace.$$
%$${\color{red} ?  = \big[ \frac{\partial \velocity}{\partial \state}\frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
%$${\color{red} ?  = \big[ \basisspace \frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
\item Third term: 
$$\frac{1}{2} \frac{\partial }{\partial \genstate} \bigg( \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)  =  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
$$ =   \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisspace$$
\end{enumerate}
This gives us,
$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} = - \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace +  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace $$
Now we work on $\frac{\partial \minintegrand}{\partial \dot{ \genstate } }$:
\begin{enumerate}
\item First term:
$$\frac{1}{2} \frac{\partial}{\partial \dot{\genstate}}\bigg( \big[\dot{\genstate}^T  \basisspace^T \big]  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  \bigg) =  \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace$$
\item Second term:
$$\frac{\partial}{\partial \dot{\genstate} } \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg) = \frac{\partial}{\partial \dot{\genstate} } \bigg( \dot{\genstate}^T \bigg) \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  $$
$$ = \bigg[  \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg]^T $$
$$ = \velocity^T \stweightingMat^T \basisspace$$
\end{enumerate}
This gives,
$$\frac{\partial \minintegrand}{\partial \dot{ \genstate } } =   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace -  \velocity^T \stweightingMat^T \basisspace.$$
Putting it all together, the Euler--Lagrange equation reads,
$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} - \frac{d}{dt} \bigg[ \frac{\partial \minintegrand}{\partial \dot{\genstate}} \bigg] =  - \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace +  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace - \velocity^T \stweightingMat^T \basisspace \bigg]  = 0$$
Factoring the first term,
$$ - \bigg( \big[ \basisspace \dot{\genstate} \big]^T  -  \velocity^T \bigg) \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace -  \velocity^T \stweightingMat^T \basisspace  \bigg] = 0$$
Taking the transpose,
$$ -  \basisspace^T \frac{\partial \velocity}{\partial \state}^T \stweightingMat^T \bigg( \big[ \basisspace \dot{\genstate} \big]  -  \velocity \bigg) -  \frac{d}{dt} \bigg[  \basisspace^T \stweightingMat^T \basisspace \dot{\genstate} - \basisspace^T \stweightingMat \velocity  \bigg] = 0.
 $$
Noting that $\stweightingMat = \stweightingMat^T$, we factor the second term,
$$ -  \basisspace^T \frac{\partial \velocity}{\partial \state}^T \stweightingMat^T \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) -  \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg[   \basisspace \dot{\genstate} - \velocity  \bigg] = 0. $$
Gathering terms and multiplying by negative one,
$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \stweightingMat^T + \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0. $$
%Equivalently,
%$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T  \basisspace \basisspace^T \stweightingMat^T + \basisspace^T  \basisspace \basisspace^T \stweightingMat^T  \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0. $$
%Pulling in the $\stweightingMat^T \basisspace^T$,
%$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \basisspace  +  \frac{d}{dt} \bigg] \bigg(  \basisspace^T \stweightingMat \basisspace \dot{\genstate}   -  \basisspace^T \stweightingMat \velocity \bigg) = 0. $$

This is the time-continuous LSPG ROM. Note that this is a second order equation and can be written as two separate first order equations. To do this, first write the above as,
$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T + \basisspace \basisspace^T \bigg)  \stweightingMat^T + \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0. $$
Re-arranging gives,
$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg]  +  \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T  \basisspace   +  \frac{d}{dt} \bigg] \bigg( \basisspace^T \stweightingMat \basisspace \dot{\genstate}   -  \basisspace^T \stweightingMat  \velocity \bigg) = 0. $$
Defining the adjoint as,
\begin{equation}
\basisspace^T \stweightingMat \basisspace \frac{d \genstate}{dt}  -  \basisspace^T \stweightingMat \velocity =  \boldsymbol \lambda , \qquad \genstate(t=0) = \genstate_0,
%\basisspace \frac{d \genstate}{dt}  -   \velocity =  \adjoint  , \qquad \genstate(t=0) = \genstate_0
\end{equation}
we then get the auxiliary adjoint equation,
\begin{equation}
%\basisspace^T \stweightingMat^T \frac{d}{dt} \adjoint  + \basisspace^T \frac{\partial \velocity}{\partial \genstate}^T \stweightingMat^T \adjoint = 0, \qquad \basisspace^T \stweightingMat \adjoint(t=T) = 0
 \frac{d}{dt} \adjoint  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \state} \bigg]^T \basisspace \adjoint = -\bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg] , \qquad \adjoint(t=T) = 0.
\end{equation}

\end{comment}

\begin{comment}
Let $\stweightingMat = \mathbf{I}$,
$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} - \frac{d}{dt} \bigg[ \frac{\partial \boldsymbol F}{\partial \dot{\genstate}} \bigg] =   \bigg( -\big[ \basisspace \dot{\genstate} \big]^T  +  \velocity^T \bigg) \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T  - \velocity^T \basisspace \bigg] $$
Transposing,
$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[   \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[ \basisspace^T \basisspace  \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
$$ =    -\basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( \basisspace \dot{\genstate}  -  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
$$ = -\bigg[ \basisspace^T \frac{d}{dt} + \basisspace^T \big[ \frac{\partial \velocity}{\partial \state} \big]^T  \bigg] \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$ 
Adjoint style
$$ =   \basisspace^T  \frac{d }{dt}\lambda  +   \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \lambda  $$
$$ \basisspace  \dot{\genstate}  -\velocity = \lambda$$
%Gathering terms
\end{comment}

\end{appendices}

%\section*{References}
\bibliographystyle{aiaa}
\bibliography{refs}{}
\end{document}

### 2019-09-09 numerical_experiments.tex
latest

\section{Numerical Experiments}\label{sec:numerical_experiments}
This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial spaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity. In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs. Additionally, we assess the impact of the window size, time-step size, and time scheme on \methodAcronym. It is noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs. Future state prediction and/or interpolation to new flow conditions is not considered as these problems introduce confounding factors (e.g., accuracy of the basis).

\subsection{Sod Shock Tube}
We first consider reduced-order models of the Sod shock tube problem. The system is governed by the compressible Euler equations in one dimension, 
\begin{equation}\label{eq:euler_1D}
    \frac{\partial \boldsymbol u}{\partial t} + \frac{\partial \boldsymbol F}{\partial \mathsf{x}} = 0, \quad
    \boldsymbol u= 
    \begin{Bmatrix} \rho \\ \rho u \\ \rho E \end{Bmatrix}, \quad 
    \boldsymbol F = \begin{Bmatrix} \rho u \\ \rho u^2 + p \\  u(\rho E + p) \end{Bmatrix},
\end{equation}
where $\boldsymbol u : \Omega \times [0,T] \rightarrow \RR{3}$ comprise the density, x-momentum, and total energy, $\mathsf{x} \in \Omega \defeq  [0,1]$ is the spatial domain, and 
$T = 1$ the final time. 
The problem setup is given by the initial conditions,

\begin{equation*}
\rho = 
\begin{cases} 
      1 & \mathsf{x}\leq 0.5 \\
      0.125 & \mathsf{x} > 0.5 
   \end{cases},
\qquad
p = 
\begin{cases} 
      1 & \mathsf{x}\leq 0.5 \\
      0.1 & \mathsf{x} > 0.5 
   \end{cases},
\qquad
u = 
\begin{cases} 
      0 & \mathsf{x}\leq 0.5 \\
      0 & \mathsf{x} > 0.5 
   \end{cases},
\end{equation*}
along with reflecting boundary conditions at $\mathsf{x}=0$ and $\mathsf{x}=1$.

\subsubsection{Description of FOM and Generation of Trial Space}
The 1D compressible Euler equations are solved using a finite volume method. The domain is partitioned into 500 cells of uniform width. The finite volume method uses the Rusanov flux~\cite{rusanov} at the cell interfaces. The FOM uses the Crank--Nicolson (CN) scheme, which is a linear multistep method defined by the coefficients $\alpha_0 = 1,\alpha_1 = -1, \beta_0 = \beta_1 = 1/2$, for temporal integration. The FOM is evolved for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$. The \spatialAcronym\ trial space is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $. The resulting trial space comprises $99.99\%$ of the snapshot matrix energy.
%  The solution is saved every other time-step, resulting in 500 solutions snapshots. These snapshots are used to generate a trial basis for the reduced-order models through POD. The trial basis is selected to be of dimension $K = 46$, which corresponds to a $99.99\%$ energy criterion. No affine transformation is used in the construction of the trial space, i.e., $\stateIntercept = \bz$.

\subsubsection{Description of Reduced-Order Models}
Reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approaches are considered. No hyper-reduction is considered 
in this example, i.e., $\stweightingMatArg{} = \mathbf{I}$. Details on the implementation of the 
different reduced-order models is as follows:
\begin{itemize}
\item \textbf{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
a constant time-step size of $ \Delta t = 0.002$.

\item \textbf{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time-scheme as previously described. Unless noted otherwise, the LSPG ROM uses a constant time-step size of $\Delta t = 0.002$. The nonlinear least-squares problem arising at each time-step instance is solved via the Gauss--Newton method. All Jacobians are computed via automatic differentiation~\cite{adolc}. The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$. 
\item \textbf{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
\begin{itemize}
\item Direct Method: \methodAcronymROMs\ solved via the direct method are implemented for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time-step size of $\Delta t = 0.0005$. The nonlinear least-squares problem 
arising over each window is solved via the Gauss--Newton method. All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$. Critically, it is noted that the Jacobian 
matrix over a window is constructed by assembling local (dense) Jacobians. The global Jacobian over a single window is then stored in a compressed sparse row format. Uniform quadrature weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.

\item Indirect Method: Two \methodAcronymROMs\ solved via the indirect method are implemented. The first uses the same CN discretization at at time-step size of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time-step size of $\Delta t = 0.0005$. The coupled two-point boundary 
problem is solved via the Forward Backward Sweep Method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. 
\end{itemize}
\end{itemize}

%We explore the evaluation of the \methodAcronym\ ROM using both indirect and direct solution methodologies. 
%\subsubsection{Indirect Method}
%The indirect \methodAcronym-ROM solves the coupled boundary value problem using the Forward-Backward Sweep Method (FBSM) described in Section~\ref{sec:fbsm}. A secondd-order collocation method with a time-step size of $\Delta t = 0.002$ and with collocation points placed at the Lobatto points is used for discretization of both the forward and backwards problems. The action of the Jacobian transpose on the residual and adjoint state required for the backwards problem is computed through automatic differentiation~\cite{adolc,pyadolc}. 
%
%\subsubsection{Direct Method}
%The direct \methodAcronym-ROM uses the same second-order collocation method with uniformly distributed collocation points for temporal discretization of both the state 
%and integrand. The resulting least-squares problem is solved using the Gauss-Newton method. The inner Gauss-Newton iterations are solved using the conjugate gradient 
%method preconditioned by the inverse approximate Hessian about the initial guess.

\subsubsection{Numerical Results}
The impact of the window size on \methodAcronymROMs\ is first assessed. A set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$ are considered. The standard Galerkin and LSPG ROMs are additionally considered. Results are shown for the direct method with the CN time-discretization scheme; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section. 
First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = \{0.5,1.0\}$. Figure~\ref{fig:sod_xt} shows 
$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, it is observed that the LSPG and Galerkin ROMs accurately characterize 
the system: they correctly track the shock location, expansion waves, etc. Both predictions, however, are observed to by highly oscillatory. These oscillations are 
not physical and can lead to numerical blowup; e.g., due to negative pressure. The \methodAcronymROMs\ produce less oscillatory solutions than both 
the Galerkin and LSPG ROMs. Critically, it is seen that, as the window size over which the residual is minimized grows, the solution becomes less oscillatory. 
When the residual is minimized over the entire space--time domain, the solution displays no oscillations.

%The impact of the window-size over which the residual is minimized on the ROM solution is first examined. To this end, reduced-order models that minimize the residual over windows of size $\Delta \tau = \Delta t \times [2,5,10,20,50,100,200,1000,2000]$ are considered. Figure~\ref{fig:xtrho} summarizes the performance of the various reduced-order models. Figure~\ref{fig:sod_error_a} shows the integrated $L^2$ state error relative to the Galerkin ROM. Figure~\ref{fig:sod_error_b} shows the space-time residual of the various reduced-order models as measured realitive to the full-order model. In Figure~\ref{fig:sod_error_a}, it is seen that the optimal ROM as measured by the integrated $L^2$ state error occurs at an intermediate window size. This result is perhaps somewhat surprising as one may expect the optimal $L^2$ error to occur when the window size spans the entire temporal domain, in which case the total space-time residual is minimized. As is evident from Figure~\ref{fig:sod_error_a}, however, a lower space-time residual does not necesarily imply a lower error. Figure~\ref{fig:sod_error_b} shows the total integrated space-time residual of the various ROMs as measured with respect to the projected full-order model solution. Increasing the window-size over which the residual is minimized leads to a lower over-all space-time residual. Further, when the residual is minimized over the entire space-time domain ($\Delta \tau =2000$), it is lower than that of the projected full-order model. This numerical evidence supports the theoretical error analysis presented in Section~\ref{sec:error}. In this example, the global space-time residual is lower than that of the projected full-order model for $\Delta \tau > 0.01$.  
%
%Figure~\ref{fig:sod_density} shows density profiles predicted by the CWST-LSPG ROM at two different time instances along with integrated total variation of the solution as measured with respect to the full-order model. It is observed that, as the window size over which the residual is being minimized grows, the solution becomes more smooth. This is measured quantitatively by the total variation of the solution, which decreases monotonically as the window size grows. Next, Figure~\ref{fig:xtrho} shows $x-t$ diagrams for the density variable for the various reduced-order models. It is again seen that, as $\Delta \tau$ grows, the solution becomes more smooth. The increased smoothness comes at the tradeoff that the shock profiles become more smeared. 
%
%Lastly, Figure~\ref{fig:sod_walltime} shows the wall-time of the CWST-LSPG ROM as a function of window-size. The cost is measured with respect to the Galerkin ROM. Two main observations are made. First, the cost of the CWST-LSPG ROM increases as the window size grows. This increased cost is because, as the window size grows, more iterations of the forward and backwards problems are required to converge the solution. While a more sophisticated solution methodology may help mitigate this increased cost, the trends are expected to say the same. The second observation is that the CWST-LSPG ROM is significantly more expensive than the standard Galerkin ROM. As discussed in Section~\ref{sec:implement}, this increased cost is due to the need to iteratively solve the two-point boundary value problem. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/rho_t05_compare.pdf}
\caption{$t=0.5$}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/rho_t1_compare.pdf}
\caption{$t=1$}
\end{subfigure}
\caption{Density profiles at various time instances.}
\label{fig:sod_density}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_fom.pdf}
\caption{Full-order model}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_grom.pdf}
\caption{G-ROM}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_lspg.pdf}
\caption{LSPG}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_w1.pdf}
\caption{\methodAcronym: $\DeltaSlabArg{}= 0.002$}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_w50.pdf}
\caption{\methodAcronym: $\DeltaSlabArg{} = 0.1$}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_w500.pdf}
\caption{\methodAcronym: $\DeltaSlabArg{} = 1.0$}
\end{subfigure}
\caption{$x-t$ diagrams for the density fields as predicted by the FOM, G-ROM, LSPG-ROM, and \methodAcronymROMs.}
\label{fig:sod_xt}
\end{center}
\end{figure}

Next, Figure~\ref{fig:sod_error} shows integrated state errors and the objective function (i.e., the residual norm) over $t \in [0,1]$ for the various  ROMs. Most notably, it is observed 
that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the integrated $\elltwo$  
error. Increasing the window size does, however, lead to a monotonic decrease in the total integrated residual. It is additionally noted that, although the integrated
$\elltwo$ error of 
the projected FOM solution is significantly lower than that of the various ROM solutions, the integrated residual norm of the projected FOM solution is \textit{higher} 
than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$ norm, it does not satisfy the governing equations as well the ROM solutions.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_compare.pdf}
\caption{Integrated $\elltwo$ error of various ROMs.}
\label{fig:sod_error_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_compare.pdf}
\caption{Integrated residual norm of various ROMs.} 
\label{fig:sod_error_b}
\end{subfigure}
\caption{Integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.  Note that the Galerkin and LSPG ROMs do not depend on the window size.} 
\label{fig:sod_error}
\end{center}
\end{figure}

We now examine the comparative performance of the direct and indirect solution techniques for the \methodAcronymROMs\ using various time discretization techniques. 
Figure~\ref{fig:sod_error_methods} 
shows the same integrated $\elltwo$ state errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, it is observed that, although some 
deviations are observed, the \methodAcronym\ method is relatively insensitive to the solution technique (direct vs indirect) and underlying discretization scheme.
In particular, ROMs using the second order explicit AB2 scheme with a time-step size of $\Delta t = 0.0005$ provide similar results to the 
ROMs using the second-order CN scheme at a time-step size of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal integrated 
$\elltwo$ state error occurs when the window size is $\DeltaSlabArg{} = 0.1$ and the residual decreases monotonically as the window size grows. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window.pdf}
\caption{Integrated $\elltwo$ error of the \methodAcronymROMs.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window.pdf}
\caption{Integrated residual norm of the \methodAcronymROMs.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Integrated performance metrics of various \methodAcronymROMs.} 
\label{fig:sod_error_methods}
\end{center}
\end{figure}

Next, Figure~\ref{fig:sod_walltime} provides the CPU wall-clock times for the various \methodAcronymROMs. The computational cost of all methods is seen to grow 
as the window size is increased. For indirect methods, this increase in cost is due to the fact that, as the window size 
grows, more iterations of the FBSM are required for convergence. For direct methods, the increase in cost is due to (1.) the cost 
associated with forming and solving the normal equations at each Gauss--Newton iteration and (2.) the increased number of 
Gauss--Newton iterations required for convergence. The cost of the FBSM method is observed to increase more rapidly 
than the direct methods. Encouragingly, \methodAcronymROMs\ based on the CN discretization
minimizing the full space--time residual (comprising 
500 time instances) only  
cost approximately one order of magnitude more than the case where the residual is minimized over a single time-step (i.e., LSPG). Also of interest is the 
fact that the direct method utilizing the AB2 time-discretization scheme costs only slightly more per a given window size than the direct method 
utilizing the CN time discretization. This is despite the fact that AB2 is evolved at a time step of $\Delta t = 0.0005$, while CN uses $\Delta t = 0.002$; thus 
AB2 contains 4 times more temporal degrees of freedom than CN.  Finally, it is emphasized that the 
results presented here are presented for standard algorithms (e.g., Guass--Newton and the FBSM). As mentioned in 
Remarks~\ref{remark:gaussnewton} and~\ref{remark:fbsm}, it is expected the computational cost 
of both the indirect methods and direct methods can be decreased through the use and/or development of 
algorithms tailored to the minimization problem.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[width=1.\linewidth]{/Users/ejparis/PyROM/pyrom/sod/postProcess_wv/error_vs_window.pdf}
\includegraphics[width=1.\linewidth]{figs/sod/walltime_vs_window.pdf}
\label{fig:sod_error_a}
\end{subfigure}
\caption{Comparison of wall-clock times of the direct and indirect \methodAcronymROMs\ as a function of window size.} 
\label{fig:sod_walltime}
\end{center}
\end{figure}

Finally, we study the impact of the time-step size on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
window size of $\Delta T = 0.1$, with time-step sizes of $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
$100$, $50$, $20$, $10$, $2$, and $1$ time-steps per window). We additionally 
consider LSPG ROMs leveraging the same set of time steps. Figure~\ref{fig:time_step_study} shows the integrated $\elltwo$ error 
of the various ROMs as well as the integrated residual norm. We observe that both the $\elltwo$ state error and integrated residual 
of the \methodAcronymROMs\ decrease and converge as the time-step size decreases. This is in direct contrast to the LSPG ROMs, in where both the $\elltwo$ 
error and integrated residual norm increase as the time-step size decreases. This result 
demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that, for all time-step sizes considered, the \methodAcronym-ROMs 
are less than 6x the cost LSPG; despite the window sizes comprising up to 100 time steps. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
\caption{Integrated $\elltwo$ error of the \methodAcronymROM.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
\caption{Integrated residual norm of the \methodAcronymROM.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs} 
\label{fig:time_step_study}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[width=1.\linewidth]{/Users/ejparis/PyROM/pyrom/sod/postProcess_wv/error_vs_window.pdf}
\includegraphics[width=1.\linewidth]{figs/sod/walltime_vs_window_dtvar.pdf}
\label{fig:sod_error_a}
\end{subfigure}
\caption{Comparison of wall-clock times of the \methodAcronymROMs\ to LSPG ROMs as a function of time-step size.} 
\label{fig:walltime_dtvar}
\end{center}
\end{figure}

\subsubsection{Summary of Numerical Results for the Sod Shock Tube}
The key observations from the results of the first numerical example are: 
\begin{enumerate}
\item Increasing the window size over which the residual is minimized lead to more physically relevant solutions. Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ lead to less oscillatory solutions.
\item Increasing the window size over which the residual is minimized does not necessary lead to a more accurate solution in the $\elltwo$ sense. We observed that minimizing the residual over an intermediary window size led to the most accurate results as measured by the $\elltwo$ norm.
\item The \methodAcronym\ displays time-step convergence: both the integrated $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time-step size decreased. This is in contrast to LSPG.
\item For all examples considered, in where the windows comprised up to 2000 time-steps, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
direct method was observed to be, in general, slightly more efficient and robust than the indirect method.
\end{enumerate}

\input{cavity}

### 2019-09-09 packages.tex
latest

\usepackage{fullpage}
%\usepackage{algorithm,algorithmic}
\usepackage[ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}[section] 
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage[titletoc,title]{appendix}
%======
\usepackage{float}
%\usepackage{wrapfig,blindtext}
\usepackage{graphicx} % to include images
\usepackage{epstopdf} % to convert eps to pdf
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pdfpages}
\usepackage{comment}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\urlstyle{same}
%\setlength{\belowcaptionskip}{-5pt}
%\usepackage{sfmath}

\usepackage{tikz}
\usetikzlibrary{positioning, fit, arrows.meta, shapes}
\newcommand{\empt}[2]{$#1^{\langle #2 \rangle}$}
\usetikzlibrary{
  arrows.meta, % for Straight Barb arrow tip
  fit, % to fit the group box around the central neurons
  positioning, % for relative positioning of the neurons
}
\tikzset{
  neuron/.style={ % style for each neuron
    circle,draw,thick, % drawn as a thick circle
    inner sep=0pt, % no built-in padding between the text and the circle shape
    minimum size=3.5em, % make each neuron the same size regardless of the text inside
    node distance=1ex and 2em, % spacing between neurons (y and x)
  },
  group/.style={ % style for the groups of neurons
    rectangle,draw,thick, % drawn as a thick rectangle
    inner sep=0pt, % no padding between the node contents and the rectangle shape
  },
  io/.style={ % style for the inputs/outputs
    neuron, % inherit the neuron style
    fill=gray!15, % add a fill color
  },
  conn/.style={ % style for the connections
    -{Straight Barb[angle=60:2pt 3]}, % simple barbed arrow tip
    thick, % draw in a thick weight to match other drawing elements
  },
}

### 2019-09-09 section3/general_form.tex
latest

\section{Windowed Least-Squares Formulation}\label{sec:tclspg} This section
outlines the proposed \methodName\ 
formulation (\methodAcronym). In contrast to (1.) the Galerkin method, which minimizes the 
time-instantaneous residual, (2.) the LSPG method, which sequentially minimizes the discrete 
residual over a single time step, and (3.) the ST-LSPG method, which minimizes the discrete 
space--time residual, the proposed \methodAcronym\ method sequentially minimizes the 
time-continuous FOM ODE residual over a series of
\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial spaces. 
In this section, we start by outlining the \methodAcronym\ formulation  
for a general trial space. We then examine the \spatialAcronym\ 
trial space, followed by the \spaceTimeAcronym\ trial space. In each case, we derive the stationary conditions 
associated with the minimization problems.
 %The proposed method can be formulated equivalently from
%two viewpoints. The first formulation corresponds to a sequence of
%minimization problems for the generalized coordinates that minimize the FOM
%ODE residual over each time slab. We refer to this as the 
%\textit{residual-minimization viewpoint}. The second viewpoint corresponds to
%formulating an optimal control problem for the Galerkin ROM. In this
%viewpoint, the \methodAcronym\ method is formulated as an optimal control
%problem in where the objective is to find a ``controller" that minimizes the
%residual of the Galerkin ROM. We refer to this viewpoint as the
%\textit{optimal control viewpoint.}  This section outlines the derivation of the \methodAcronym\ method
%from both the residual minimization and optimal control viewpoints.

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Residual Minimization Formulation} 
We start by partitioning the
temporal domain into a set of $\nslabs$ windows of size $\DeltaSlabArg{n}$. This partitioning is depicted in Fig.~\ref{fig:slab_fig}.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} \caption{Decomposition of time
domain into time windows.} \label{fig:slab_fig} \end{centering} 
\end{figure}
Over each time window, we introduce the space--time trial space,
$$\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,$$ 
where $\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
\RR{\fomdim}$
is the \methodAcronym\ approximation to the state over the $n$th time window, $\timeSpaceArg{n}$ is the set of functions acting on $[\timeStartArg{n},\timeEndArg{n}]$, and  
$\timeStartArg{n}$ and $\timeEndArg{n}$ are the beginning and end of the $n$th
time window, respectively. Over each window, define the objective,
%\begin{equation}\label{eq:obj} \mathcal{J}^n(\stateArg{}{t}) = \frac{1}{2}
%\int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\state}(\tau) -
%\velocity(\stateArg{}{\tau}) \big]^T \stweightingMatArg{n}(\tau) \big[
%\dot{\state}(\tau) - \velocity(\stateArg{}{\tau}) \big] d \tau,
%\end{equation}
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RR{}, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ } \in \RR{\fomdim \times \fomdim}$ is a
symmetric positive semi-definite matrix that, e.g., enables hyper-reduction and/or modifies the norm. For simplicity, 
it is assumed that the weighting matrix is the same for all time windows.

\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
%      &\text{subject to } \;  \stateyArg{}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}

%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
\begin{equation}\label{eq:tclsrm}
\begin{split}
       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
\end{split}
\end{equation}
where $\mathbb{P}^n: \RR{N} \otimes \timeSpaceArg{n} \rightarrow \stspaceArg{n}$ is an orthogonal projection onto the 
trial space (e.g., $\elltwo$).
%\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing
%the FOM ODE residual over the trial space for each time slab.
%\methodAcronym\ is defined as follows,
%\begin{equation}\label{eq:tclsrm}
%\approxstate^n= \underset{\statey \in \stspace^n}{\text{arg\,min } }
%\mathcal{J}^n(\statey), \qquad n = 1,2,...,\nslabs,
%\end{equation}
% subject to
%the boundary condition, \begin{equation*} \approxstate^n(\timeStartArg{n}) =
%\begin{cases}\mathbb{P}^n \approxstate^{n-1}(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n \stateFOMIC & n=1, \end{cases} \end{equation*}
%where $\mathbb{P}^n: \RR{N} \rightarrow \stspaceArg{n}$ projects onto the 
%trial space (e.g., an $L^2$ projector).
\begin{remark}
It is possible to construct the trial spaces such that they have continuous support between time windows through the 
use of, e.g., affine transformations.
\end{remark}
%and 
%$\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
%\RR{\fomdim}$
%is the approximate solution over the $n$th time slab.% and $\approxstate_0 \in
%\RR{\fomdim}$ is the FOM initial condition
%projected onto the trial space; (e.g., via for $L^2$ projection).
We now define the trial spaces $\stspaceArg{n}$. We consider \spatialAcronym\ trial spaces and \spaceTimeAcronym\ trial spaces. 
We leave further investigation into other trial spaces, such as nonlinear manifolds~\cite{leeCarlberg}, as a subject for future work.
%A variety of trial spaces have been examined in the model-order reduction community.
%The most popular trial space comprises a linear, spatial-projection-only trial space (e.g., Eqns.~\eqref{eq:spatial_subspace}-\eqref{eq:spatial_subspace3}), although recently 
%nonlinear trial spaces~\cite{leeCarlberg} and space--time trial spaces~\cite{choi_stlspg,constantine_strom,2,3} have gained attention. This work focuses on the classic 
%spatial-projection-only trial space, and also briefly considers space--time trial spaces.
%We leave further investigation into other trial spaces as future work.

### 2019-09-09 section3/space_time.tex
latest

\subsection{\spaceTimeAcronym\ Trial Spaces}
We now briefly consider \spaceTimeAcronym\ trial spaces. Over 
each window, we introduce the space--time trial space
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
with $\text{Range}(\stbasisArgnt{n}) + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
the \spaceTimeAcronym\ trial space computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
  \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMArgnt{1})(0)
 & n=1. \end{cases}
%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary Conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial spaces is as follows: in \spaceTimeAcronym\ trial spaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial spaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be derived using standard calculus. Differentiating the objective with respect to $\stgenstatey$ and setting equal to zero yields,
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial spaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.

### 2019-09-09 section3/tclsrm_formulation.tex
latest

\input{section3/general_form}
\input{section3/spatial_projection}
\input{section3/space_time}

\subsection{\methodAcronym\ Summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial space for a series of time windows:
\begin{equation*}
\approxstate^n= \underset{\statey \in \stspace^n}{\text{arg\,min } }
\mathcal{J}^n(\statey), \qquad n = 1,2,...,\nslabs,
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RR{},
\end{align*}
subject to the boundary and continuity conditions. For the \spatialAcronym\ trial spaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\mass \frac{d}{dt} \genstateArg{n}{t}  -  \basisspace^T
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \mass \adjointArg{n}{t} , \\
 &\mass \frac{d}{dt} \adjointArg{n}{t}  + \basisspace^T \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspace
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspace^T \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
\basisspace [\massArgt{n}{t}]^{-1} \basisspace^T \stweightingMatArgt{n}{t} \bigg)
 \bigg( \basisspace \frac{d}{dt}\genstateArg{n}{t}  -
\velocity(\veloargsromn) \bigg),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
trial spaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property.

\spaceTimeAcronym\ trial spaces were additionally considered. For \spaceTimeAcronym\ trial spaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.

### 2019-09-09 section4/numerical_methods.tex
latest

\section{Numerical Solution Techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged. This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system. We focus primarily on direct and indirect methods for \spatialAcronym\ trial spaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial spaces. 
%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems. For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$. The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates. 
%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).

\subsection{\spatialAcronym: Direct and Indirect Methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". For both direct and indirect methods, a variety of
discretization/solution approaches are then possible. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial spaces. We consider:

\begin{itemize} \item Discretize then Optimize (Direct Methods): 
\begin{itemize}
\item Discretize and then optimize approaches are outlined for linear multistep methods. 
\end{itemize} 
\item Optimize then Discretize (Indirect Methods): 
\begin{itemize} 
\item An indirect method leveraging the forward-backward sweep algorithm is outlined 
to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
 \end{itemize} 
\end{itemize} 
It
is emphasized that a variety of other approaches exist. A detailed exploration
of additional approaches will be a subject of future work. Section~\ref{sec:direct} 
will outline direct solution approaches based on linear multistep methods, 
while Section~\ref{sec:indirect} will outline the indirect solution approach.

\input{section4/spatial_numerical}
\input{section4/spacetime_numerical}

### 2019-09-09 section4/spacetime_numerical.tex
latest

\subsection{\spaceTimeAcronym\ Trial Spaces: Direct and Indirect Methods}
We now briefly consider \spaceTimeAcronym\ trial spaces. 
For \spaceTimeAcronym\ trial spaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
dimensional representation of the objective function~\eqref{eq:obj_gen_slab}. In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE. We associate direct methods as those 
that work with the O$\Delta$E and indirect methods with those that work directly with the continuous ODE.

\subsection{\spaceTimeAcronym\ Trial Spaces: Direct Solution Technique}
The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods.

%The definition of the fully discrete objective functions~\eqref{eq:obj_lms} as outlined in Section~\ref{sec:direct}, can be leveraged for this purpose. For simplicity, we only 
%outline the case for linear multistep schemes.

Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis is given by, 
 \begin{equation}\label{eq:obj_gen_lms_final_st}
\begin{split}
&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMArgnt{1})(0) &
n=1. \end{cases}\end{split} 
\end{equation}

The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss-Newton method.

\begin{remark}
For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial space recovers ST-LSPG.
\end{remark}

\subsection{\spaceTimeAcronym\ Trial Spaces: Indirect Solution Technique}
As opposed to the direct solution technique, the indirect solution technique directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime},
\begin{equation}\label{eq:obj_gen_slab2}
\begin{split}
 &\stgenstate^n =
\underset{\stgenstatey}{\text{arg\,min }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) \\  
&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMArgnt{1})(0) &
n=1. \end{cases} 
\end{split} 
\end{equation}
%The first order optimality conditions are given by,
%\begin{equation}\label{eq:st_stationary}
% \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
%\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
%\stateInterceptArg{n})\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n}) \bigg) dt = \bz.\end{equation}
Numerically solving the minimization problem requires the introduction of a quadrature rule for 
discretization of the integral. We introduce $\ncollocSTArg{n} \ge \text{ciel}(\stdim / N )$ collocation points over the $n$th window, 
$$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \ldots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$
Leveraging these collocation points, the indirect method for \methodAcronym\ with a \spaceTimeAcronym\ trial space yields,
\begin{equation}\label{eq:obj_gen_slab2} 
\begin{split}&\stgenstate^n =
\underset{\stgenstatey}{\text{arg\,min }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \\
&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMArgnt{1})(0) &
n=1,
\end{cases}
\end{split} 
\end{equation}
where the discrete objective function is given by, 
$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptArg{n},\collocPointSTArg{n,i}) \bigg]^T 
\stweightingMatArg{n} 
\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptArg{n},\collocPointSTArg{n,i}) \bigg].
$$
The nonlinear least-squares problem can again be solved, e.g., via the Gauss-Newton method. 
%The stationary points are given by,
%\begin{multline*}\label{eq:st_stationary_discrete}
%\frac{\partial \objectiveDiscreteSTArg{n}}{\partial \stgenstatey}(\stgenstateArg{n}) = 
%\sum_{i=0}^{\ncollocSTArg{n}} \gamma_i \bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}}^T  - \stbasisArg{n}{\collocPointSTArg{n}{i}}^T \bigg[\frac{\partial
%\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} +                    
%\stateInterceptArg{n})\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} + \stateInterceptArg{n}) \bigg),\end{multline*}
%%subject to the boundary conditions 
%\begin{equation}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\approxstateIC &
%n=1. \end{cases} \end{equation}

\subsection{\spaceTimeAcronym\ Trial Space: Summary}
As was just shown, \spaceTimeAcronym\ trial spaces yield a series of space--time algebraic systems over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial spaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
space--time trial space yields a series of dense (non)linear systems to be solved over each window.

### 2019-09-09 section4/spatial_numerical.tex
latest

\subsection{\spatialAcronym: Direct Solution Approach}\label{sec:direct} 
Direct approaches solve~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective function in time.
The minimization problem is then reformulated as a nonlinear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, orthogonal polynomial (spectral  method) 
approaches, and genetic algorithms.

In the present context, the most straightforward direct solution approach consists of the 
following steps:
\begin{enumerate}
\item Numerically discretize the FOM ODE (and hence the \textit{integrand} of~\eqref{eq:obj_gen_slab}).
\item Select a numerical quadrature rule to evaluate the \textit{integral}
in~\eqref{eq:obj_gen_slab}.
\end{enumerate}
To achieve this agenda, first define a discrete temporal grid 
by dividing each time window $[\timeStartArg{n},\timeEndArg{n}]$ into
$\nstepsArg{n}$ strictly increasing time-step instances described by
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with  
\begin{align}
&\timeWindowArg{n}{0} \le \ldots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
\timeWindowArg{n}{0} \defeq \timeStartArg{n},\;
\timeWindowArg{n}{\nstepsArg{n}} \defeq \timeEndArg{n} . 
%\timeWindowArg{n}{0} = \timeStartArg{n}, \timeWindowArg{n}{\nstepsArg{n}} =
%\timeEndArg{n}
\end{align}
This discretization is depicted in Figure~\ref{fig:slab_fig2}.
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-stepping methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
\caption{Decomposition of time windows into time-step instances.} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.

%\subsubsection{Temporal Discretization of the FOM ODE}
%To develop the direct solution approach, we first outline several 
%commonly employed methods used to discretize the FOM ODE, and then 
%subsequently demonstrate how these techniques can be used to discretize 
%and solve Eq.~\ref{eq:obj_gen_slab}. To develop the direct solution approach, 
%we start by dividing each time slab $[\timeStartArg{n},\timeEndArg{n}]$ into
%$\nstepsArg{n}$ strictly increasing time-step instances described by,
%$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$, with  \begin{align*}
%&\timeWindowArg{n}{0} \le \ldots \le \timeWindowArg{n}{\nstepsArg{n}} , \\
%&\timeWindowArg{n}{0} \defeq \timeStartArg{n}, \\
%&\timeWindowArg{n}{\nstepsArg{n}} \defeq \timeEndArg{n}. 
%%\timeWindowArg{n}{0} = \timeStartArg{n}, \timeWindowArg{n}{\nstepsArg{n}} =
%%\timeEndArg{n}
%\end{align*}
\subsubsection{Linear Multistep Schemes}
Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time-instances.  
%Discretizing the FOM ODE~\eqref{eq:FOM} with a linear $k$-step scheme at the $i$th time-instance over the $n$th time slab yields,
%\begin{equation}\label{eq:kstep}
%\sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha_j \stateFOMDiscreteArg{n,i-j} = \Delta t^{n,i} \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta_j \velocity(\stateFOMDiscreteArg{n,i-j}),
%\end{equation}
Employing a linear multistep method for temporal discretization of the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
\begin{align*}
&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i},\ldots,\stateFOMDiscreteArg{n,i-k^n(i)}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
&\stateArgnt{n,0} = \begin{cases}
\stateFOMDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
\stateFOMIC & n=1, \end{cases}
\end{align*}
where $\stateFOMDiscreteArg{n,i} \approx \stateFOMArg{n}{\timeWindowArg{n}{i}} \in \RR{\fomdim}$ is the discrete approximation to the FOM state at the $i$th time instance over the $n$th window and $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time instance of the $n$th window,
\begin{align*}
\residArg{n,i} &: (\stateyDiscreteArgnt{n,i},\ldots,\stateyDiscreteArgnt{n,i-k(i)}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{n,i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{n,i-j},\timeWindowArg{n}{i-j},\tau^{n,i-j}) \\
               &: \RR{\fomdim} \otimes \RR{k(i)+1} \rightarrow \RR{\fomdim}. 
\end{align*} 
Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time-step size, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that 
define the specific type of multistep scheme. Examples of linear multistep schemes include Adams-Bashforth schemes and  
Adams-Moulton schemes. 
\begin{remark}\label{remark:LMS}
For simplicity, we assume the linear multistep method employs time-step instances from only within the current time window. In general, the linear multistep method at the $n$th window could be designed to employ multiple time-step instances from previous time windows.
\end{remark}
%Employing a linear multistep method for temporal discretization of the FOM ODE leads to the FOM O$\Delta$E over the $n$th slab,
%\begin{align*}
%&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i},\ldots,\stateFOMDiscreteArg{n,i-k^n(i)}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
%&\stateArgnt{n,0} = \begin{cases}
%\stateFOMDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
%\stateFOMIC & n=1, \end{cases}
%\end{align*}
%where $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time-step instance of the $n$th slab,
%\begin{align*}
%\residArg{n,i} &: (\stateyDiscreteArgnt{n,i},\ldots,\stateyDiscreteArgnt{n,i-k(i)}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha_j \stateyDiscreteArgnt{n,i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta_j \velocity(\stateyDiscreteArgnt{n,i-j}),
%\\
%               &: \RR{\fomdim} \otimes \RR{k(i)} \rightarrow \RR{\fomdim}. 
%\end{align*} 
Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with an objective \textit{function},
\begin{equation}\label{eq:obj_lms}
\begin{split} 
\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{n,i},\ldots, \stateyDiscreteArgnt{n,i-k(n)})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{n,i},\ldots, \stateyDiscreteArgnt{n,i-k(n)}), \\
&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}+1} \rightarrow
\RR{}, 
\end{split}
\end{equation}
%where 
%\begin{align*}
%\residArg{n,i} &: (\stateyArgnt{n,i},\ldots,\stateyArgnt{n,i-k(i)}) \mapsto  \sum_{j=0}^k \alpha_j \stateyArgnt{n,i-j} = \Delta t^{n,i} \sum_{j=0}^k \beta_j \velocity(\stateyArgnt{n,i-j}),
%\\
%               &: \RR{\fomdim} \otimes \RR{k(i)} \rightarrow \RR{\fomdim}, 
%\end{align*} 
%is the discrete residual resulting from the linear multistep scheme and 
where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. \methodAcronym\ solves the minimization problem over each time window, 
\begin{align*}
&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
\approxstateIC & n=1. \end{cases} \end{align*}
Defining $\genstateDiscreteArg{n,i} \in \RR{\romdim}$ to be the discrete approximation to the generalized coordinates at the $i$th time-step instance over the $n$th window, 
the minimization problem can be written equivalently as,
\begin{equation}\label{eq:obj_gen_lms_final}
\begin{split}
& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }
\objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
& \text{subject to} \hspace{0.2 in}
\genstateDiscreteArg{n,0} =
\begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
\basisspace^T(\stateFOMIC - \stateIntercept)& n=1. \end{cases} 
\end{split}
\end{equation}
The optimization problem takes the form of a \textit{weighted nonlinear least-squares problem}, 
in where the weights correspond the quadrature weights.  
Before proceeding, it is worth mentioning that~\eqref{eq:obj_gen_lms_final} corresponds to a least-squares problem due
to the reduction in \textit{spatial} complexity, as opposed to the reduction
in temporal complexity (the same  number of temporal constraints and temporal 
degrees of freedom have been introduced).

The minimization problem~\eqref{eq:obj_gen_lms_final} requires specification of the quadrature weights (and hence the integration scheme used to discretize 
the objective functional). Typically, the same integration scheme used to discretize the FOM ODE is employed for consistency~\cite{colloc_review}; e.g., if a  
backward Euler method is used to discretize the FOM ODE, then a backward Euler method is the used to numerically integrate the objective functional.

\begin{remark}
For the limiting case where the window size is equivalent to the time-step size, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
quadrature weights are used, \methodAcronym\ recovers LSPG.
\end{remark} 
%%%%%%%%%%%%%%%%COLLOCATION%%%%%%%
\begin{comment}
\subsubsection{Collocation (Runge--Kutta) Methods}
Collocation methods comprise another popular technique for temporal discretization. 
Collocation methods represent the FOM state over each time-step instance $[\timeWindowArg{n}{i-1},\timeWindowArg{n}{i}]$ as,
\begin{equation}
\stateFOMArgnt{n}(t) \approx  \collocMatArg{n}{i} \collocBasisArg{n}{i}(t), \qquad t \in [\timeWindowArg{n}{i-1}, \timeWindowArg{n}{i} ],
\end{equation}
with the
continuity and boundary constraints, 
\begin{equation}\label{eq:colloc_constraint}
\collocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
\begin{cases}
\collocMatArg{n}{i-1}\collocBasisArg{n}{i-1}(\timeWindowArg{n}{i-1}) & n = 1,\ldots,\nslabs, \; i=2,\ldots,\nstepsArg{n},\\ 
\collocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1,  \\
\stateFOMIC & n=1, \;  i=1.
\end{cases}
\end{equation}
%\begin{equation}\label{eq:continuity_constraint}
%\genCollocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
%%\genstatecollocArg{n}{i}{\timeWindowArg{n}{i-1}} =
%%\genstatecollocArg{n}{i-1}{\timeWindowArg{n}{i-1}} , \qquad  i =
%%2,\ldots,\nstepsArg{n}. 
%\begin{cases} \genstatecollocArg{n}{i-1}{\timeWindowArg{n}{i-1}} & i =
%2,\ldots,\nstepsArg{n} \\
%\genstatecollocArg{n-1}{\nstepsArg{n}}{\timeWindowArg{n}{i-1}} & i= 1. \\
%\end{cases} 
%\end{equation} 
In the above, $\collocMatArg{n}{i}  \in \RR{\fomdim \times \collocOrderArg{n}{i}}$ are \textit{space--time coordinates}
over the $i$th time-step instance of the $n$th window, 
%$\genstatecollocArgnt{n}{i} : [\timeWindowArg{n}{i-1},\timeWindowArg{n}{i}] \mapsto \RR{\fomdim}$ is the collocated 
%approximation to the state over the $i$th time-step instance within the $n$th slab
and $\collocBasisArg{n}{i} :
[\timeWindowArg{n}{i-1},\timeWindowArg{n}{i}] \mapsto \RR{\collocOrderArg{n}{i}}$ comprise the
$\collocOrderArg{n}{i}$ 
\textit{polynomial basis functions} over the $i$th time-step instance of the $nth$ window. 
%While a variety of temporal basis functions are possible,
%polynomial basis functions are the most commonly employed and as such we consider these here.

The collocation approach generates an approximation to~\eqref{eq:FOM} by enforcing the FOM ODE at a set of \textit{collocation points}. To this end, over the $i$th time-step instance of the $n$th window, we define $\ncollocArg{n}{i} \defeq \collocOrderArg{n}{i}-1$ strictly increasing \textit{collocation points},
\begin{align*}
&\collocPointVecArg{n}{i} \defeq \begin{bmatrix} \collocPointArg{n}{i}{1}  &  \collocPointArg{n}{i}{2} \hdots & \collocPointArg{n}{i}{\ncollocArg{n}{i}} \end{bmatrix}, \\
& \timeWindowArg{n}{i-1} \le \collocPointArg{n}{i}{1} \le \hdots \le \collocPointArg{n}{i}{\ncollocArg{n}{i}} \le \timeWindowArg{n}{i}.
\end{align*} 
Enforcing~\eqref{eq:FOM} at the collocation points replaces the FOM ODE system over the $i$th time-step instance of the 
$n$th window with the FOM O$\Delta$E,
\begin{align*}
&\residCollocArg{n,i} ( \collocMatArg{n}{i} ) = \boldsymbol  0, 
\end{align*}
where $\collocMatArg{n}{i}$ is subject to the continuity and boundary constraints~\eqref{eq:colloc_constraint} and 
$\residCollocArg{n,i}$ is the discrete collocation residual over the $i$th time-step instance of the $n$th window,
\begin{align*}
\residCollocArg{n,i} &: ( \collocMatyArg{n}{i}) \mapsto \begin{bmatrix} \big[ \collocMatyArg{n}{i} \collocBasisDotArg{n}{i}(\collocPointArg{n}{i}{1}) - 
\velocity\big(\collocMatyArg{n}{i} \collocBasisArg{n}{i}(\collocPointArg{n}{i}{1})  \big) \big]^T& 
\cdots &
\big[ \collocMatyArg{n}{i} \collocBasisDotArg{n}{i}(\collocPointArg{n}{i}{\ncollocArg{n}{i}}) - 
\velocity\big(\collocMatyArg{n}{i} \collocBasisArg{n}{i}(\collocPointArg{n}{i}{\ncollocArg{n}{i}})) \big]^T 
\end{bmatrix}^T, \\
&:  \RR{\fomdim \times \collocOrderArg{n}{i}} \rightarrow \RR{ \fomdim  \ncollocArg{n}{i} }.
\end{align*}
Employing a collocation method thus replaces the objective functional~\eqref{eq:obj} with the objective function,
\begin{align}\label{eq:obj_colloc}
\objectiveArgC{n} &: (\collocMatyArg{n}{1},\ldots,\collocMatyArg{n}{\nstepsArg{n}} ) \mapsto 
\frac{1}{2} \sum_{i = 1}^{\nstepsArg{n}} [ \residCollocArg{n,i} ( \collocMatyArg{n}{i} ) \circ  \sqrt{\quadWeightsVecArg{n,i}}]^T   \stweightingMatCollocArg{n}{i}  [  \residCollocArg{n,i}  ( \collocMatyArg{n}{i} ) \circ \sqrt{\quadWeightsVecArg{n,i }}] \\
&: \RR{\fomdim} \otimes \RR{\collocOrderArg{n}{i} \nstepsArg{n}} \rightarrow \RR{} ,
\end{align}
where $\quadWeightsVecArg{n,i} \in \RR{\fomdim \ncollocArg{n}{i}}$ are quadrature weights of the form,
$$\quadWeightsVecArg{n,i} = \begin{bmatrix} \quadWeightsScalarArg{n}{i}{1} & \quadWeightsScalarArg{n}{i}{1} & \ldots & &  \quadWeightsScalarArg{n}{i}{\ncollocArg{n}{i} } & \quadWeightsScalarArg{n}{i}{\ncollocArg{n}{i}} \end{bmatrix},$$
and $\circ$ represents the element-wise (Hadamard) product. Additionally,  
\begin{equation*}
\stweightingMatCollocArg{n}{i} \defeq \begin{bmatrix}
\stweightingMatArgt{n}{\collocPointArg{n}{i}{1}} &  & & \\
& \stweightingMatArgt{n}{\collocPointArg{n}{i}{2}} & & \\ 
& &  \ddots & \\
& & &  \stweightingMatArgt{n}{\collocPointArg{n}{i}{\ncollocArg{n}{i}}}
\end{bmatrix}
\in  \RR{\fomdim \ncollocArg{n}{i} \times \fomdim \ncollocArg{n}{i}}.
\end{equation*}
The \methodAcronym-ROM then  solves the minimization 
problem,
\begin{equation}\label{eq:min_colloc}
\collocMatArg{n}{1},\ldots,\collocMatArg{n}{\nstepsArg{n}} 
 = \underset{\collocMatyArg{n}{1},\ldots,\collocMatyArg{n}{\nstepsArg{n}}  \in \trialspace}{\text{arg\,min } }
\objectiveArgC{n} (\collocMatyArg{n}{1},\ldots,\collocMatyArg{n}{\nstepsArg{n}} ) \qquad \text{ for } n = 1,2,...,\nslabs,
\end{equation}
subject to the boundary conditions,
\begin{equation*}
\collocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
\begin{cases}
\collocMatArg{n}{i-1}\collocBasisArg{n}{i-1}(\timeWindowArg{n}{i-1}) & n = 2,\ldots,\nslabs, \; i=2,\ldots,\nstepsArg{n} ,\\ 
\collocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1 , \\
\approxstateIC & n=1, \;  i=1.
\end{cases}
\end{equation*}
Defining  $\genCollocMatArg{n}{i} \in \RR{\romdim \times \collocOrderArg{n}{i}}$ to be the generalized space--time coordinates over the $i$th time-step instance of the $n$th window,
the objective function can be equivalently written as,
 \begin{equation*}
\genCollocMatArg{n}{1},\ldots,\genCollocMatArg{n}{\nstepsArg{n}} 
 = \underset{\genCollocMatyArg{n}{1},\ldots,\genCollocMatyArg{n}{\nstepsArg{n}} }{\text{argmin } }
\objectiveArgC{n} (\basisspace \genCollocMatyArg{n}{1} + \stateInterceptMat,\ldots,\basisspace \genCollocMatyArg{n}{\nstepsArg{n}} + \stateInterceptMat) \qquad \text{ for } n = 1,2,...,\nslabs,
\end{equation*}
subject to the boundary and continuity conditions,
\begin{equation*}
 \genCollocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
\begin{cases}
\genCollocMatArg{n}{i-1}\collocBasisArg{n}{i-1}(\timeWindowArg{n}{i-1}) & n = 2,\ldots,\nslabs, \; i=2,\ldots,\nstepsArg{n} \\ 
\genCollocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1  \\
\basisspace^T (\approxstateIC - \stateIntercept)& n=1, \;  i=1,
\end{cases}
\end{equation*}
with $\stateInterceptMat \defeq \begin{bmatrix} \stateIntercept & \ldots & \stateIntercept \end{bmatrix} \in \RR{\fomdim \times \collocOrderArg{n}{i} }$.
The optimization problem again corresponds to a weighted nonlinear least-squares minimization problem. As before, the specification of a numerical 
integration scheme is required to specify the quadrature weights. For the consistent case of a collocated quadrature rule, in where the integral is approximated 
via the Lagrange polynomials at the collocation points, the quadrature weights are given by, $\quadWeightsScalarArg{n}{i}{j} = \intSlabArg{n}\lagrangeBasisScalarArg{n}{i}{j}(t)dt,$ where $\lagrangeBasisScalarArg{n}{i}{j}$ are the Lagrange interpolating polynomials,
$$\lagrangeBasisScalarArg{n}{i}{j}(t) = \underset{\underset{m \ne j}{1 \le m \le \ncollocArg{n}{i}}}{\Pi} \frac{t - \collocPointArg{n}{i}{m}}{\collocPointArg{n}{i}{j}- \collocPointArg{n}{i}{m}}.$$

\begin{remark} %The presented collocation formulation corresponds to minimizing the
%weighted O$\Delta$E residual over the time slab, in where the O$\Delta$E
%results from temporal
%discretization of the FOM ODE according to the temporal basis and collocation points. For 
%the case where the temporal basis functions are
%polynomials, the resulting temporal discretization schemes belongs to the family of Runge--Kutta
%schemes~\cite{1}. 
Collocation methods belong to the family of Runge--Kutta schemes.
Alternative temporal discretization schemes, such as
linear multistep methods, can be obtained by modifying the representation of
the state and the collocation points~\cite{ajams2015362}.
\end{remark}
\end{comment}
\subsubsection{Solution to the Nonlinear Least-Squares Problem through the Gauss--Newton Method}
Discretization through linear multistep methods (as well as other techniques) 
results in a nonlinear least-squares problem.
A variety of algorithms exist for solving nonlinear  
least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
The numerical experiments presented in this work consider the Gauss--Newton method, and as such we outline this approach here.

To describe the Gauss--Newton method, we first define a ``vectorization" function that takes a tuple of time-local solutions at each time-step instance and outputs 
a concatenated vector, 
\begin{align*}
 \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm},
\end{align*}
for arbitrary $p$ and $m$. The stacked vector of generalized coordinates over the $n$th time window are then defined as, 
\begin{equation*}
\genstatecollocMatSlabArg{n} \defeq 
\unroll (\genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
\end{equation*}
Now, define the weighted space--time residual over the entire window as
\begin{equation*}
\residLMSSlabArg{n} : \genstatecollocMatySlabArg{n} \mapsto \begin{bmatrix}
 \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n,1} \residLMSArg{n,1}( \basisspace \genstateyDiscreteArgnt{n,1} + \stateIntercept, \ldots,\basisspace \genstateyDiscreteArgnt{n,1 - k^n(1) }  + \stateIntercept) \\
\vdots \\
 \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n,\nstepsArg{n}} \residLMSArg{n,\nstepsArg{n}}( \basisspace \genstateyDiscreteArgnt{n,\nstepsArg{n}} + \stateIntercept, \ldots,\basisspace \genstateyDiscreteArgnt{n,\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateIntercept) \\
\end{bmatrix},
\end{equation*}
where $\genstatecollocMatySlabArg{n} \defeq \unroll (\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}} ).$
It is worth noting that, by design,
\begin{equation*}
\objectiveArgLMS{n} \bigg( \basisspace \genstateDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateDiscreteArg{n,\nstepsArg{n}} + \stateIntercept \ \bigg) 
\equiv
\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
\end{equation*} 
Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method. The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the time window residual-Jacobian and form the normal equations, and (3) solve the normal equations and update the state.

The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the time window residual-Jacobians. In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the time window residual Jacobian will (almost) be a block-diagonal 
sparse matrix with the sparsity pattern,
\begin{equation*}
\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{n}}(\genstatecollocMatSlabArg{n})  =
 \begin{bmatrix*}[l]
\matshapeb & \\%[-5pt]
 \matshapea & \matshapea & \\%[-5pt]
 & \matshapea  & \matshapea & \\%[-5pt]
&  & \ddots & \\%[-5pt]
 & &  & \matshapea &  \matshapea 
\end{bmatrix*}.
% \begin{bmatrix*}[c]
%\matshapeb & \\[-5pt]
% & \hspace{-12pt}\matshapea & \\[-5pt]
% &  & \hspace{-12pt}\matshapea & \\[-5pt]
%&  & \hspace{5pt} \ddots & \\[-5pt]
% & &  & & \hspace{-15 pt} \matshapea 
%\end{bmatrix*}.
\end{equation*}
This sparsity pattern can be leveraged to assemble the time window residual-Jacobian from local residual Jacobians. 
Further, the sparsity pattern can be 
leveraged to efficiently compute the Jacobian matrix-matrix product in the normal equations.
It is also worth noting that the 
normal equations will also (almost) be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.

\begin{remark}\label{remark:gaussnewton}\textit{(Acceleration of the Gauss--Newton Solve)}\\
The principle cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
reducing this computational burden exist; Jacobian-free Newton-Krylov methods~\cite{jfnk}, Broyden's method~\cite{broyden} (as explored in Ref.~\cite{carlberg_thesis}, Appendix A), and frozen Jacobian approximations are several such examples. Further, the space--time formulation exposes an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM. The investigation of these additional, potentially more efficient, solution algorithms will be a topic of future work. 
\end{remark}
%\subsubsection{Jacobian-Free Implementation}
%The dominant cost of the Gauss Newton algorithm is the computatation of the action of the Jacobian matrix on the trial basis. To alleviate this burden, the following Jacobian-free method is proposed:
\begin{algorithm}
\caption{\spatialAcronym\ trial space: algorithm for the direct solution technique with the Gauss--Newton method and a linear multistep method over the $n$th window}
\label{alg:colloc_gn}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $tol$; initial guess, $\genstateGuessDiscreteArg{n,0}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$}
\Output{Solution to least squares problem, $\genstatecollocMatSlabArg{n}$} 
\textbf{Online Steps}: \\
%\While{ $tol \le \frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \sum_{j=1}^{\ncollocArg{n}{i}} \big[ \dot{\state}
%(\collocPointArg{n}{i}{j})  - \velocity(\stateArg{}{\collocPointArg{n}{i}{j}}) \big]^T \stweightingMatArg{n}(\collocPointArg{n}{i}{j}) \big[ \dot{\state}(\collocPointArg{n}{i}{j}) - \velocity(\stateArg{}{\collocPointArg{n}{i}{j}}) \big] \int_{\timeStartArg{n}}^{\timeEndArg{n}}   \lagrangeBasisScalarArg{n}{i}{j}(t) d \tau$}
$\text{converged} \leftarrow \text{false}$ \Comment{Set convergence checker} \\
$k \leftarrow 0$ \Comment{Set counter}\\
$\genstatecollocMatSlabArg{n}_k \leftarrow \unroll(\genstateGuessDiscreteArg{n,0}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
\While{\text{converged} == \text{false}}
{
%\For{$i=1,\hdots,\nstepsArg{n}$}{
%  \For{$j=1,\hdots,\ncollocArg{n}{i}$}{
%Compute: $\approxstateArgnt{n,i}  =  \basisspace  \genstateDiscreteArg{n,i} + \stateIntercept$ 
%\Comment{Compute state} \\
%Compute: $\velocity(\basisspace \genstateDiscreteArg{n,i} + \stateIntercept ) $ \Comment{Compute velocity at each time-step instance}\\
%Compute: $\residLMSArg{n,i}(\basisspace \genstateDiscreteArg{n,i} + \stateIntercept, \ldots , \basisspace \genstateDiscreteArg{n,i - k^n(i)} + \stateIntercept) $  \Comment{Compute residual} \\
%}
$\mathbf{r} \leftarrow \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{Compute weighted residual over window} \\
$\mathbf{J} \leftarrow  
\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{n}}(\genstatecollocMatSlabArg{n}_k) 
$ \Comment{Compute weighted residual-Jacobian over window} \\
%Compute: $[\jacobianSlabArg{n}]^T \jacobianSlabArg{n}$ \Comment{Compute system matrix for the normal equations} \\
%Compute: $[\jacobianSlabArg{n}]^T \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{compute RHS for normal equations} \\
$\Delta \genstatecollocMatSlabArg{n} \solveAlgo [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
\If{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \text{tol}$ }{
{\text{converged} $\leftarrow$ \text{true}}  \Comment{Check and set convergence based on gradient norm} \\
Return: $\genstatecollocMatSlabArg{n} = \genstatecollocMatSlabArg{n}_{k+1} $ \Comment{Return converged solution}\\
}
$k\leftarrow k+1$
}
\end{algorithm}

%\input{direct_collocation}

\subsection{Spatial Trial Space: Indirect Solution Approach}\label{sec:indirect}
In contrast to the direct approach,
indirect methods ``indirectly" solve the minimization
problem~\eqref{eq:tclsrm} by solving the Euler-Lagrange equations. The
system given by the Euler Lagrange equations~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_adjoint} comprises a coupled two-point boundary value
problem. Similar to the direct solution approach, a variety of techniques have
been devised to solve two-point boundary value problems of this type. These
techniques include shooting methods, multiple shooting
methods~\cite{multiple_shooting}, and the forward-backward sweep
method~\cite{fbs} (FBSM).  This work explores using the FBSM.

%Solving this coupled problem is significantly more challenging and
%computationally expensive than the standard Galerkin and LSPG methods as it
%requires converging both the forward and backward solve. It is emphasized,
%however, that the entire forward-backward system is compatable with
%hyper-reduction techniques and thus is entirely independent of the full-order
%model size. In addition, as the approach is minimizing the entire space-time
%residual, we expect it to be capable of providing stable and accurate
%solutions in cases where the standard Galerkin and LSPG methods can not.
%Obtaining numerical solutions to Eq.~\ref{eq:lspg_continuous} requires three
%ingrediants: \begin{enumerate} \item Solution strategy for solving the
%coupled forward and backwards problems \item Time discretization schemes for
%the forward and backward problems \item Efficient strategy for evaluating the
%action of the Jacobian transpose on a vector \end{enumerate}

\subsubsection{The Forward Backward Sweep Method (FBSM)}\label{sec:FBSM}
%\subsubsection{Solution Strategy} Solving Eq.~\ref{eq:lspg_continuous} is
%made challenging by the fact that it is a coupled two-point boundary value
%problem. The forward problem is coupled to the backwards adjoint problem,
%while the backwards adjoint sytem is coupled to the forward problem. The
%problem is thus inherently implicit. The most popular techniques to solve
%these types of two-point boundary value problems are shooting methods,
%multiple shooting methods, fully-implicit methods, and the forward backward
%sweep (FBS) method. This work explores using the FBS method.

%The most straightforward approach to solving Eq.~\ref{eq:lspg_continuous} is
%to 1.) discretized both the forward and backwards problems in time and 2.)
%form and solve the resulting (implicit) nonlinear space-time system. In
%practice, however, this technique may not be practical for larger
%time-windows due to the size of the resuling nonlinear problem.

The FBSM is an iterative approach that can be used to
solve the two-point boundary value problem. The general process of the FBSM is
as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
time using an initial guess to the costate state to obtain an initial guess to
the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
approximation to the generalized coordinates. This gives a new estimate to the
costate, which is then used to again solve the forward problem to obtain
a new estimate of the generalized coordinates. This process is continued until
convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
contains three parameters: the damping factor $\rho$, the growth factor
$\fbsmGrowth$, and the decay factor $\fbsmDecay$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
closer $\rho$ is to unity, the faster the resulting algorithm will converge.
For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
A proper value of $\rho$ can be
obtained with a line search. The line search presented in~\eqref{eq:lspg_continuous} adapts the damping factor
according to the objective. Convergence properties of the FBSM method are
presented in Ref.~\cite{McAsey2012ConvergenceOT}. It is shown that, for a small 
enough value of $\rho$, the algorithm will converge.

\begin{algorithm} \caption{\spatialAcronym\ trial space: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $tol$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
factor, $\fbsmDecay$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\genstateArgnt{n}_1 \solveAlgo \basisspace^T \stweightingMatArg{n}
\basisspace \genstateDotArgnt{n}_1  -  \basisspace^T \stweightingMatArg{n}
\velocity(\veloargsromntArg{1}) =  \controllerArgnt{n}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous} with initial guess to
controller}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$tol \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\adjointArgnt{n} \underset{\text{solve}}{\leftarrow} 
\frac{d}{dt} \adjoint^n  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \basisspace \mass^{-1} \adjoint^n = \\ -\bigg[\basisspace^T \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspace \mass^{-1} \basisspace^T \bigg)  \stweightingMatArg{n} \bigg( \basisspace \dot{\genstate}_i^n   -   \velocity( \basisspace \genstate_i^n  +\stateIntercept,t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate state} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update Counter}\\
$\genstateArgnt{n}_i \solveAlgo \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArgnt{n}_i   -  \basisspace^T \stweightingMatArg{n} \velocity(\basisspace \genstate^n_i + \stateIntercept,t) =  \controllerArgnt{n} $
\Comment{Solve~\ref{eq:lspg_continuous} with updated guess to costate}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\genstate_i^n(t)}) \le \objectiveArg{n}({\genstate_{i-1}^n(t)})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical Considerations for the Forward and Backwards Problems}
The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1.) temporal discretization schemes for the forward and backward problems and (2.) 
an efficient method for computing action of the Jacobian transpose on a vector.

This work examines the use of linear multistep methods for the purpose of temporal discretization of the forward and 
backwards problems. As described in  
the direct solution approach, temporal discretization is achieved by dividing each time window into $\nstepsArg{n}$ time-step instances~\eqref{eq:timegrid1}. Linear multistep methods then use this time-grid for temporal discretization of~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}.

The second ingredient, namely an efficient method for computing the action of the Jacobian transpose on a vector, can be challenging. 
In the case one does not have access to the full-order model Jacobian (or it is too costly to compute), the evaluation of this term can be both challenging and expensive. Here, two methods are discussed that can be used to evaluate the action of the Jacobian transpose on a vector as seen in~\eqref{eq:lspg_adjoint}:
\begin{enumerate}
\item Jacobian-free Derivative Approximation: A non-intrusive way to evaluate the Jacobian transpose in~\eqref{eq:lspg_adjoint} is to recognize that all terms including the Jacobian transpose are left multiplied by the transpose of the trial basis. One can make the manipulation,
$$\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} \bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace \bigg]^T.$$
This manipulation allows one to compute the action of the Jacobian on each column of $\basisspace$ by, e.g., a finite difference approximation,
$$\frac{\partial \velocity}{\partial \stateyDiscrete} \basisvec_i \approx \frac{1}{\epsilon}\bigg( \velocity(\stateyDiscrete + \epsilon \basisvec_i) - \velocity(\stateyDiscrete) \bigg).$$
The Jacobian transpose can be formed in $K+1$ evaluations of the velocity. For cases where the reduced-order model is low dimensional, this approach is feasible. For higher dimensional reduced-order models, this approach can be prohibitively expensive.

\item Automatic Differentiation: A more intrusive, but potentially more efficient, method of computing the action of the Jacobian transpose on a vector is through automatic differentiation (AD). AD methods comprise a class of techniques that can be used to numerically evaluate derivatives of functions (e.g., Jacobians, vector-Jacobian products, etc.) by recursively applying the chain rule. The numerical examples presented later in this work leverage AD. The principle drawback of AD is that AD methods are intrusive and may not be suitable for, e.g., legacy codes.  
\end{enumerate}

\begin{remark}\label{remark:fbsm}(Acceleration of Indirect Methods)
The FBSM is a simple iterative method for solving the coupled two-point boundary value problem. For large time windows, however, the FBSM may require many 
forward-backward iterations for convergence. More sophisticated solution techniques, such as a multiple FBSM method or multiple shooting methods, promise 
to reduce this cost. Analyses of additional solution techniques will be a subject of future work.
\end{remark}

### 2019-09-09 figs/make_grid.py

@@ -1,0 +1,38 @@
+from pylab import *
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt
+from matplotlib import rc
+rc('text', usetex=True)
+rc('text.latex', preamble='\usepackage{amsmath},\usepackage{amssymb}')
+import os
+axis_font = {'size':16,'family':'serif'}
+mpl.rc('font',family='serif')
+rc('text', usetex=True)
+axis_font = {'size':'14'}
+close("all")
+import matplotlib.patches as patches
+fig,ax = plt.subplots(1)
+matplotlib.pyplot.arrow(0.,0.,1.,0.,head_width=0.05,head_length=0.05,color='black')
+rsz = 0.02
+nslabs = 4
+dx = 1./nslabs
+slab_start = linspace(0.,1.-dx,nslabs)
+for i in range(0,np.size(slab_start)):
+  start = slab_start[i]
+  rect = patches.Rectangle((start-rsz/2.,-rsz/2.),rsz,rsz,linewidth=1,edgecolor='black',facecolor='none')
+  ax.add_patch(rect)
+  matplotlib.pyplot.text(start,rsz,r'$t_s^{}$'.format(i+1),**axis_font)
+  if (i > 0):
+    matplotlib.pyplot.text(start,rsz*5,r'$t_f^{}$'.format(i),**axis_font)
+    matplotlib.pyplot.text(start - dx/5*3,-rsz*4,r'$\Delta T^{}$'.format(i),**axis_font)
+matplotlib.pyplot.text(start + dx/5*2,-rsz*3,r'$\ldots$'.format(i),**axis_font)
+ 
+ylim([-0.2,0.2])
+xlim([-0.1,1.1])
+ax.set_aspect(aspect=1.)
+ax.axis('off')
+#ax.set_xticks([])
+#ax.set_yticks([])
+savefig('time_grid.pdf')
+show()

### 2019-09-18 section3/general_form.tex
latest

@@
 \methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+\begin{equation}\label{eq:tclsrm}
+\begin{split}
+      &\underset{\stateyArgnt{n} \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\stateyArgnt{n}), \\
+      &\text{subject to } \;  \stateyArg{n}{\timeStartArg{n}} =
+\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
+\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
+\end{split}
+\end{equation}
+%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
 %\begin{equation}\label{eq:tclsrm}
 %\begin{split}
-%      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
-%      &\text{subject to } \;  \stateyArg{}{\timeStartArg{n}} =
+%       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
+%      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
 %\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
@@
 %\end{equation}
-
-%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
-\begin{equation}\label{eq:tclsrm}
-\begin{split}
-       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
-      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
-\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
-\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
-\end{split}
-\end{equation}
 where $\mathbb{P}^n: \RR{N} \otimes \timeSpaceArg{n} \rightarrow \stspaceArg{n}$ is an orthogonal projection onto the 

### 2019-09-18 section3/tclsrm_formulation.tex
latest

@@
 This section outlined the \methodAcronym\ method for model reduction.
-In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial space for a series of time windows:
+In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial space for a series of time windows: \methodAcronym\
+computes the generalized coordinates $\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the minimization problems,
 \begin{equation*}
-\approxstate^n= \underset{\statey \in \stspace^n}{\text{arg\,min } }
-\mathcal{J}^n(\statey), \qquad n = 1,2,...,\nslabs,
+\underset{\stateyArgnt{n} \in \stspace^n}{\text{minimize } }
+\mathcal{J}^n(\stateyArgnt{n}), \qquad n = 1,2,...,\nslabs,
 \end{equation*}

### 2019-09-18 section4/spacetime_numerical.tex
latest

@@
 \begin{split}
-&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
+& \underset{\stgenstatey}{\text{minimize } }
 \objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
-&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
 \mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
@@
 \end{equation}
+
+% \begin{equation}\label{eq:obj_gen_lms_final_st}
+%\begin{split}
+%&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
+%\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
+%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%n=1. \end{cases}\end{split} 
+%\end{equation}
 
@@
 \begin{split}
- &\stgenstate^n =
-\underset{\stgenstatey}{\text{arg\,min }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) \\  
-&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+ & \underset{\stgenstatey}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) \\  
+&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
 \mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
@@
 \end{equation}
+%\begin{equation}\label{eq:obj_gen_slab2}
+%\begin{split}
+% &\stgenstate^n =
+%\underset{\stgenstatey}{\text{arg\,min }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) \\  
+%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%n=1. \end{cases} 
+%\end{split} 
+%\end{equation}
 %The first order optimality conditions are given by,
@@
 \begin{equation}\label{eq:obj_gen_slab2} 
-\begin{split}&\stgenstate^n =
-\underset{\stgenstatey}{\text{arg\,min }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \\
-&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+\begin{split}
+&\underset{\stgenstatey}{\text{minimize }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \\
+&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
 \mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
@@
 \end{equation}
+%\begin{equation}\label{eq:obj_gen_slab2} 
+%\begin{split}&\stgenstate^n =
+%\underset{\stgenstatey}{\text{arg\,min }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \\
+%&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%n=1,
+%\end{cases}
+%\end{split} 
+%\end{equation}
 where the discrete objective function is given by, 

### 2019-09-18 main.tex
update

@@
 %\title{Model Reduction via \methodName}
-\title{Windowed Least--Squares Reduced-Order Models for Dynamical Systems}
+%\title{Windowed least--squares reduced-order models for dynamical systems}
+\title{Windowed residual-minimization model reduction for dynamical systems}
 
@@
 \begin{abstract}
-This work proposes a windowed least-squares (\methodAcronym) reduced-order modeling formulation for dynamical systems. The proposed approach 
-sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time windows.
-Trial spaces with spatial and space--time dimension reduction are considered. 
-The stationary conditions for the resulting minimization problems are derived.
-Of particular interest, the stationary conditions for trial spaces with spatial dimension reduction are obtained via the Euler--Lagrange equations and comprise a coupled two-point Hamiltonian boundary value problem containing a forward and backward system.
+This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
+	formulation for dynamical systems. The proposed approach minimizes the
+	time-continuous full-order-model residual over a low-dimensional space--time
+	trial subspace across a sequence of time windows that decompose the time
+	domain of interest.  We consider two types of space--time trial subspaces
+	within this framework:
+	one that reduces the spatial dimension of the full-order model, and one that
+	reduces both the spatial and temporal dimensions of the full-order model.
+	\KTC{Why are we saying the next two sentences? These seem kind of random.
+	The abstract should explain why the work is necessary; this more reads like
+	a list of interesting factoids.} The stationary conditions for the resulting minimization problems are
+	derived.  Of particular interest, the stationary conditions for trial spaces
+	with spatial dimension reduction are obtained via the Euler--Lagrange
+	equations and comprise a coupled two-point Hamiltonian boundary value
+	problem containing a forward and backward system. 
 %The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model.
@@
 %by deriving the associated Euler--Lagrange equations. 
-Two differing methods are examined to solve the minimization problems: direct (discretize then minimize) and indirect (minimize and then discretize).
-Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov--Galerkin, and Space--Time Least-Squares Petrov--Galerkin.
+We consider two methods to solve the residual-minimization problems: direct
+	(i.e., discretize-then-optimize) and indirect (i.e.,
+	optimize-then-discretize).  Critically, we show that the proposed method
+	comprises a generalization of many existing techniques and provides a
+	unified framework to reason about existing model-redcution techniques. In
+	particular, we show that specific instances of the proposed method lead to
+	the existing reduced-order modeling techniques of Galerkin projection,
+	least-squares Petrov--Galerkin (LSPG) projection, and space--time
+	least-squares Petrov--Galerkin (ST-LSPG) projection. 
 %By formulating the least--squares problem from the time-continuous level, the \methodAcronym\ formulation overcomes 
@@
 \section{Introduction}
-The simulation of dynamical systems plays an important role in a variety of areas in science and engineering; such examples include engineering design and control, financial analysis, and epidemiology.
-It is often the case, however, that (numerically) executing a single dynamical system solve is a computationally intensive process; e.g., this is the case when the state-space dimension of the dynamical system is prohibitively large and/or a wide range of time-scales need to be resolved.
-In such circumstances, it is common for analysts to rely on approximate models that generate approximations to the system of interest at a greatly reduced cost.
-
-Projection-based reduced-order models (ROMs) have proved to be a promising tool for developing low-cost approximate solutions.
-These ROMs leverage ``snapshot" data from high-fidelity simulations and/or numerical experiments to identify a low-dimensional ``trial" space on which the state variables can be well approximated.
-ROM methods then seek a solution to the governing equations that lies on this low-dimensional trial space via projection or residual minimization.
-Efforts by a variety of research groups have led to significant advancements in ROM technologies over the past several decades, and we refer to Ref.~\cite{wilcox_benner_rev} for a review.
-
-The construction of reduced-order models for nonlinear dynamical systems has proved challenging.
-The principle challenge encountered when trying to construct reduced-order models for such problems is that most existing methods lack \textit{a priori} guarantees of stability, accuracy, and convergence.
-%The standard Galerkin and least-squares Petrov-Galerkin methods, for instance, are both subject to exponentially growing error bounds~\cite{carlberg_lspg_v_galerkin}.
-A great deal of effort has been devoted towards developing methods that allow for more robust and accurate reduced-order models of dynamical systems.
-These efforts include residual-norm minimization~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis}, space--time methods~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st},  ``energy-based" inner products~\cite{rowley_pod_energyproj,Kalashnikova_sand2014}, symmetry transformations~\cite{sirovich_symmetry_trans}, basis adaptation~\cite{carlberg_hadaptation,adeim_peherstorfer}, projection subspace rotations~\cite{basis_rotation}, and subgrid-scale modeling~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
-Many of these developed techniques are not mutually exclusive.
-Residual minimization is most relevant to the current work and is the focus of the following review.
+
+Simulating parameterized dynamical systems underpins many applications across
+science and engineering, including engineering design, control, financial
+analysis, and epidemiology. In many contexts, executing a dynamical-system
+simulation at a single parameter instance---which entails the numerical
+integration of a system of ordinary differential equations (ODEs)---incurs an
+extremely large computational expense.  This occurs, for example, when the
+state-space dimension is large (e.g., due to fine spatial resolution when
+discretizing a partial-differential-equations problem) and/or when the number
+of time instances is large (e.g., due to stability or accuracy requirements
+that limit the time step).  When the application is time-critical or
+many-query in nature, analysts must replace such large-scale parameterized
+dynamical-system models with a low-cost approximation that makes the
+application tractable.
+
+\KTC{I reworked this. Methods like balanced truncation or IRKA do \textit{not}
+use snapshot data!}
+
+Projection-based reduced-order models (ROMs) comprise one such approximation
+strategy. First, these techniques execute a computationally expensive
+\textit{offline} stage that computes a low-dimensional ``trial'' subspace on
+which the (parameterized) dynamical-system state can be well approximated
+(e.g., by computing state ``snapshots'' at different parameter instances, by
+solving Lyapunov equations). Second, these methods execute an inexpensive
+\textit{online} stage during which they compute approximations to the
+dynamical-system trajectory that reside on this trial subspace via a
+residual-minimization or projection process.
+
+Model reduction for linear-time-invariant systems (and other strongly
+structured dynamical systems) is quite mature \cite{wilcox_benner_rev}
+\KTC{Also cite classical refs on balanced truncation, Gugercin's IRKA,. Note
+that I moved \cite{wilcox_benner_rev} here because that review mainly deals
+with linear systems}, as system-theoretic properties (e.g., controllability,
+observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
+quantified and accounted for in the projection process, which yields
+reduced-order models that inherit such properties.  The primary challenge for
+developing reduced-order models for general nonlinear dynamical systems is
+that such properties are difficult to assess for such systems. As a result, it
+is difficult to develop projection methods that preserve important
+dynamical-system properties in the reduced-order model, which can lead to
+models that are inaccurate or unstable.  To address this, researchers have
+pursued several directions that aim to imbue reduced-order models of nonlinear
+dynamical systems with properties that can improve robustness and accuracy.
+These
+efforts include residual-minimizing
+projection that imbues the ROM solution with a notion of optimality~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis};
+space--time
+projection that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
+``energy-based'' inner
+products that ensure non-increasing entropy in the ROM solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014}; symmetry
+transformations that \KTC{what?} ~\cite{sirovich_symmetry_trans},
+basis-adaptation methods that improve the ROM's accuracy \textit{a posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer}, subspace
+rotations that \KTC{what?} ~\cite{basis_rotation}, structure-preserving
+methods that enforce conservation \cite{carlberg_conservative_rom} or
+(port-)Hamiltonian/Lagrangian structure
+\cite{carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional}
+\KTC{Can use more here. See my presentation} in the ROM; and subgrid-scale
+modeling methods that aim to improve accuracy by addressing the closure
+problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
+We note that these techniques are often not mutually exclusive.
+Residual-minimizing and space--time projection are the most relevant classes of
+methods to the current work and comprise the focus of the following review.
+\KTC{Shouldn't we review space-time also?}
+
+\KTC{HERE}
 
@@
 
-This work seeks to overcome the practical challenges associated with existing residual minimization approaches for reduced-order models.
-To this end, we develop a reduced-order modeling approach that minimizes the FOM ODE residual over a series of time windows.
-The key novelties of the work are two fold.
-First, the residual minimization problem is formulated to minimize the \textit{time-continuous} residual.
-By working from a continuous viewpoint, the present formulation decouples the underlying temporal discretization scheme from the residual minimization problem; thus addressing a key deficiency of LSPG.
-The continuous viewpoint additionally exposes two differing methods that can be used to solve the
-minimization problems, namely \textit{discretize and then optimize} (direct) methods and \textit{optimize and then discretize} (indirect) methods.
-The second novelty is that the residual is minimized over \textit{time windows}, as opposed to over a single time step or the entire time domain.
-This allows for maximum flexibility in terms of computational cost and numerical accuracy.
-We demonstrate that this formulation recovers the Galerkin, LSPG, and space--time LSPG methods.
-Figure~\ref{fig:flowchart} depicts how the proposed method interfaces with existing approaches.
-
-The resulting approach is best viewed as a hybrid space--time method.
-A variety of research has examined the formulation of space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis} and the present approach displays commonalities with several such efforts.
-First,~\cite{bui_thesis} briefly formulated a space--time least-squares ROM and connected this formulation with optimal control; specifically the thesis of optimize then discretize vs discretize then optimize was mentioned.
-This formulation, however, was not fully developed and was eschewed for the time-local residual minimization method (i.e., LSPG).
-The present work thus formally develops (and extends) several of the concepts put forth in~\cite{bui_thesis}.
-Next, Ref.~\cite{constantine_strom} developed a space--time residual minimization formulation for model interpolation.
-The present work distinguishes itself from Ref.~\cite{constantine_strom} in that (1.) in addition to trial spaces with space--time dimension reduction, we consider trial spaces with spatial dimension reduction only (i.e., a classic spatial basis is employed, opposed to a space--time basis) -- in which case the generalized coordinates comprise a vector of functions, and (2.) we minimize the time-continuous residual over space--time windows;~\cite{constantine_strom} considers the solution of a fully discrete optimization problem over the entire space--time domain.
-It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work.
-Lastly, Ref.~\cite{choi_stlspg} develops a space--time formulation of the least-squares Petrov--Galerkin approach.
-The formulation minimizes the \textit{fully discrete} residual over the entire space--time domain within a given space--time trial space.
+This work seeks to overcome the practical challenges associated with existing
+residual minimization approaches for reduced-order models. To this end, we
+develop a reduced-order modeling approach that minimizes the FOM ODE residual
+over a series of time windows. The key novelties of the work are two fold.
+First, the residual minimization problem is formulated to minimize the
+\textit{time-continuous} residual. By working from a continuous viewpoint, the
+present formulation decouples the underlying temporal discretization scheme
+from the residual minimization problem; thus addressing a key deficiency of
+LSPG. The continuous viewpoint additionally exposes two differing methods that
+can be used to solve the minimization problems, namely \textit{discretize and
+then optimize} (direct) methods and \textit{optimize and then discretize}
+(indirect) methods. The second novelty is that the residual is minimized over
+\textit{time windows}, as opposed to over a single time step or the entire
+time domain. This allows for maximum flexibility in terms of computational
+cost and numerical accuracy. We demonstrate that this formulation recovers the
+Galerkin, LSPG, and space--time LSPG methods. Figure~\ref{fig:flowchart}
+depicts how the proposed method interfaces with existing approaches.
+
+The resulting approach is best viewed as a hybrid space--time method. A
+variety of research has examined the formulation of space--time reduced-order
+models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis}
+and the present approach displays commonalities with several such efforts.
+First,~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
+and connected this formulation with optimal control; specifically the thesis
+of optimize then discretize vs discretize then optimize was mentioned. This
+formulation, however, was not fully developed and was eschewed for the
+time-local residual minimization method (i.e., LSPG). The present work thus
+formally develops (and extends) several of the concepts put forth
+in~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
+space--time residual minimization formulation for model interpolation.  The
+present work distinguishes itself from Ref.~\cite{constantine_strom} in that
+(1.) in addition to trial spaces with space--time dimension reduction, we
+consider trial spaces with spatial dimension reduction only (i.e., a classic
+spatial basis is employed, opposed to a space--time basis) -- in which case
+the generalized coordinates comprise a vector of functions, and (2.) we
+minimize the time-continuous residual over space--time
+windows;~\cite{constantine_strom} considers the solution of a fully discrete
+optimization problem over the entire space--time domain. It is noted that,
+similar to the current work, a time-continuous residual-minimization
+formulation is used as a starting point in~\cite{constantine_strom}, and as
+such thematic similarities exist between~\cite{constantine_strom} and the
+present work. Lastly, Ref.~\cite{choi_stlspg} develops a space--time
+formulation of the least-squares Petrov--Galerkin approach. The formulation
+minimizes the \textit{fully discrete} residual over the entire space--time
+domain within a given space--time trial space. 
 %These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. 
-The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that (1.) we perform model reduction of the time-continuous system, (2.) we consider trial spaces with both space--time dimension reduction and spatial dimension reduction, and (3.) residual minimization is performed over time windows; Ref.~\cite{choi_stlspg} considers space--time bases and minimizes the fully-discrete residual over the entire domain.
+The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that
+(1.) we perform model reduction of the time-continuous system, (2.) we
+consider trial spaces with both space--time dimension reduction and spatial
+dimension reduction, and (3.) residual minimization is performed over time
+windows; Ref.~\cite{choi_stlspg} considers space--time bases and minimizes the
+fully-discrete residual over the entire domain.
 

### 2019-09-19 main.tex
small progress

@@
 %\end{itemize}
-Least-squares residual minimization as a numerical solution technique for (partial) differential equations was pioneered in the finite element community and led to what is known as the least-squares finite element method (LSFEM)~\cite{ls_review,bochev_leastsquares}.
-Motivated by the fact that the Galerkin method can be inaccurate and unstable for nonlinear and non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic functional comprising the residual norm of the partial differential equation of interest and (2.) leveraging variational principles to derive the stationary conditions of the quadratic functional; i.e, LSFEM can be thought of as finding the solution within a finite element trial space the minimizes the $2$-norm of the PDE residual.
-Least-squares methods often exhibit numerous advantages (e.g., least-squares system are symmetric and positive definite) over the more standard Galerkin discretization.
-
-Residual minimization methods were extended-to and developed-for reduced-order models by a variety of researchers.
-Similar to LSFEM, residual minimization approaches for reduced-order models comprise finding the solution within a low-dimensional trial space that minimizes the residual norm of the full-order model.
-Residual minimization approaches were first applied to static (i.e., no time-dependence) reduced-order models in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
-Specifically, Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg} formulated the residual minimization problem for dynamical systems by sequentially minimizing the \textit{fully discrete} residual (i.e., the residual after temporal discretization) at each time-step instance.
-This formulation is most commonly known as the \textit{Least--Squares Petrov--Galerkin} (LSPG) method.
-Ref.~\cite{carlberg_lspg_v_galerkin} provided numerous analyses on this formulation and examined connections with the standard Galerkin method and Petrov--Galerkin methods.
-Critically, it was shown that (1.) for various time-stepping methods LSPG has an equivalent continuous form that corresponds to a Petrov--Galerkin projection and (2.) the standard Galerkin method minimizes the time-instantaneous residual of the full-order model.
-Numerous numerical experiments have demonstrated that LSPG yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
-The common physical explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.
-
-The LSPG approach only guarantees residual minimization over a single time-step, and as such is not associated with residual minimization of the full dynamical system.  
-Indeed, LSPG is not equipped with \textit{a priori} guarantees of stability or convergence, even for linear time-invariant systems~\cite{bui_thesis}.
-Further, as it is defined at the fully discrete level, the performance of LSPG is subject to the time-step size: changing the time-step size changes the window over which the residual is minimized \textit{as well as} the temporal discretization error.
-As a result, LSPG provides optimal results at an intermediary time-step size.
-For too small a time-step, LSPG yields results similar to the Galerkin approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin.
-For too-large a time-step, the accuracy of LSPG degrades.
-It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
-The LSPG approach is additionally sensitive to the time discretization scheme.
-When used with explicit temporal discretization, for instance, LSPG again reverts to Galerkin.
-To gain the enhanced stability associated with LSPG, the underlying time-discretization scheme must be implicit with an appropriate time-step.
-This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.
+
+\KTC{Why are we reviewing LSFEM? This isn't a ROM. Don't think it's relevant}
+Least-squares residual minimization as a numerical solution technique for
+partial differential equations was pioneered in the finite-element community
+and led to what is known as the least-squares finite-element method
+(LSFEM)~\cite{ls_review,bochev_leastsquares}. 
+\KTC{Specifically, LSFEM was designed to solve the inf--sup stability problem.
+Be more precise.}
+Motivated by the fact that the
+Galerkin method can be inaccurate and unstable for nonlinear and
+non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic
+functional comprising the residual norm of the partial differential equation
+of interest and (2.) leveraging variational principles to derive the
+stationary conditions of the quadratic functional; i.e, LSFEM can be thought
+of as finding the solution within a finite element trial space the minimizes
+the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous
+advantages (e.g., least-squares system are symmetric and positive definite)
+\KTC{This is not an `advantage'. It is guaranteed to be inf--sup stable.}
+over the more standard Galerkin discretization.
+
+Researchers also developed residual-minimization methods for model reduction.
+Similar to LSFEM, residual minimization in this context associates with
+computing the solution within a low-dimensional trial space that minimizes the
+$\ell^2$-norm of the full-order-model residual. \KTC{haven't defined
+`full-order model'} Researchers have developed such residual-minimizing
+model-reduction methods for both static systems (i.e., systems without
+time-dependence)
+\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
+and dynamical
+systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
+In the latter category,
+Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
+formulated the residual minimization problem for dynamical systems by
+sequentially minimizing the \textit{time-discrete} full-order-model residual
+(i.e., the residual arising after applying time discretization) at each time
+instance on the time-discretization grid. This formulation is often referred
+to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
+Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
+formulation and examined its connections with Galerkin projection. Critically,
+this work demonstrated that (1) under certain conditions, LSPG can be derived
+from a Petrov--Galerkin projection executed on the time-continuous
+full-order-model residual, and (2) LSPG and Galekrin projection are equivalent
+in the limit as the time step goes to zero (i.e., Galerkin projection
+minimizes the time-instantaneous full-order model residual). Numerous
+numerical experiments have demonstrated that LSPG projection often yields more
+accurate and stable solutions than Galerkin
+projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
+The common intuitive explanation for this improved performance is that, by
+minimizing the full-order-model residual over a finite time window (rather
+than time instantanously), LSPG projection computes solutions that are more
+accurate over a larger part the trajectory compared with Galerkin projection.
+
+\KTC{HERE}
+The LSPG approach only guarantees residual minimization over a single
+time-step, and as such is not associated with residual minimization of the
+full dynamical system.  Indeed, LSPG is not equipped with \textit{a priori}
+guarantees of stability or convergence, even for linear time-invariant
+systems~\cite{bui_thesis}. Further, as it is defined at the fully discrete
+level, the performance of LSPG is subject to the time-step size: changing the
+time-step size changes the window over which the residual is minimized
+\textit{as well as} the temporal discretization error. As a result, LSPG
+provides optimal results at an intermediary time-step size. For too small a
+time-step, LSPG yields results similar to the Galerkin approach; in the
+limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin. For
+too-large a time-step, the accuracy of LSPG degrades. It is unclear if this is
+due to the numerical error associated with enlarging the time-step size, or
+rather if it is due to the size of the window the residual is being minimized
+over. The LSPG approach is additionally sensitive to the time discretization
+scheme. When used with explicit temporal discretization, for instance, LSPG
+again reverts to Galerkin. To gain the enhanced stability associated with
+LSPG, the underlying time-discretization scheme must be implicit with an
+appropriate time-step. This can increase the cost of the reduced-order model,
+especially in the case where the optimal time-step required by LSPG is small.
 

### 2019-09-19 main.tex
latest

--- 2019-09-18 main.tex
@@
 %\title{Windowed least--squares reduced-order models for dynamical systems}
-\title{Windowed residual-minimization model reduction for dynamical systems}
+\title{Windowed Residual-Minimization Model Reduction for Dynamical Systems}
 
@@
 	formulation for dynamical systems. The proposed approach minimizes the
-	time-continuous full-order-model residual over a low-dimensional space--time
-	trial subspace across a sequence of time windows that decompose the time
-	domain of interest.  We consider two types of space--time trial subspaces
+	time-continuous full-order-model residual over a low-dimensional 
+	trial subspace across a sequence of time windows and addresses key deficiencies in existing model-reduction approaches; e.g., the time-step and time-scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG.
+We consider two types of trial subspaces
+	within the proposed framework:
+	one that reduces the spatial dimension of the full-order model, and one that
+	reduces both the spatial and temporal dimensions of the full-order model.
+For both types of trial spaces, we consider two differing solution techniques: direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
+Critically, we show that various instances of this framework yield existing model-reduction approaches, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
+Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.
+
+\begin{comment}
+This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
+	formulation for dynamical systems.  
+% In contrast to (1.) the Galerkin method, which minimizes the time-instantaneous residual, (2.) the Least-Squares Petrov--Galerkin (LSPG) method, which sequentially minimizes the discrete residual over a single time step, and (3.) the Space--Time LSPG method, which minimizes the discrete space–time residual,
+%In contrast to existing residual minimization model-reduction approaches, which generate solutions in a low-dimensional space--time trial space that either minimizes (1.) the time-continuous residual over an infinitesimal time-window (Galerkin), (2.) the time-discrete residual over a single time-step instance (Least-Squares Petrov--Galerkin), or (3.) the time-discrete residual over the entire time domain (Space--Time Least-Squares Petrov--Galerkin),
+%the proposed approach minimizes the
+%	time-continuous full-order-model residual over a low-dimensional space--time
+%	trial subspace across a sequence of time windows. Among other advantages, this process yields an approach that overcomes the 
+%time-step and time-scheme dependence that the least-squares Petrov--Galerkin method is subject to. 
+%that decompose the time
+%	domain of interest.
+The proposed approach minimizes the
+	time-continuous full-order-model residual over a low-dimensional 
+	trial subspace across a sequence of time windows. % that decompose the time
+%	domain of interest. The proposed approach overcomes the time-step and time-scheme dependence that the least-squares Petrov--Galerkin method is subject to.  
+%	\KTC{Why are we saying the next two sentences? These seem kind of random.
+%	The abstract should explain why the work is necessary; this more reads like
+%	a list of interesting factoids.} %The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model.
+The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
+%For spatial trial spaces, the stationary conditions for the time-continuous minimization problems are obtained 
+%by deriving the associated Euler--Lagrange equations. 
+We consider two types of trial subspaces
 	within this framework:
@@
 	reduces both the spatial and temporal dimensions of the full-order model. 
-	\KTC{Why are we saying the next two sentences? These seem kind of random.
-	The abstract should explain why the work is necessary; this more reads like
-	a list of interesting factoids.} The stationary conditions for the resulting minimization problems are
-	derived.  Of particular interest, the stationary conditions for trial spaces
-	with spatial dimension reduction are obtained via the Euler--Lagrange
-	equations and comprise a coupled two-point Hamiltonian boundary value
-	problem containing a forward and backward system. 
-%The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model.
-The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
-%For spatial trial spaces, the stationary conditions for the time-continuous minimization problems are obtained 
-%by deriving the associated Euler--Lagrange equations. 
+%The stationary conditions for the resulting minimization problems are derived.
+Of particular interest, the stationary conditions for trial spaces with spatial dimension reduction are obtained via the Euler–Lagrange equations and comprise a coupled two-point Hamiltonian boundary value problem containing a forward and backward system.
 We consider two methods to solve the residual-minimization problems: direct
 	(i.e., discretize-then-optimize) and indirect (i.e.,
-	optimize-then-discretize).  Critically, we show that the proposed method
-	comprises a generalization of many existing techniques and provides a
-	unified framework to reason about existing model-redcution techniques. In
-	particular, we show that specific instances of the proposed method lead to
-	the existing reduced-order modeling techniques of Galerkin projection,
-	least-squares Petrov--Galerkin (LSPG) projection, and space--time
-	least-squares Petrov--Galerkin (ST-LSPG) projection. 
+	optimize-then-discretize).
+The approach addresses key deficiencies in a variety 
+of existing model-reduction approaches; e.g., the time-step and time-scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG.
+Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
+%  We show that the proposed method
+%	comprises a generalization of many existing techniques and provides a
+%	unified framework to reason about existing model-redcution techniques. In
+%	particular, we show that specific instances of the proposed method lead to
+%	the existing reduced-order modeling techniques of Galerkin projection,
+%	least-squares Petrov--Galerkin (LSPG) projection, and space--time
+%	least-squares Petrov--Galerkin (ST-LSPG) projection. 
 %By formulating the least--squares problem from the time-continuous level, the \methodAcronym\ formulation overcomes 
@@
 Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.
-
+\end{comment}
 %This work proposes a least-squares residual minimization approach for reduced-order models of dynamical systems.
@@
 
-\KTC{I reworked this. Methods like balanced truncation or IRKA do \textit{not}
-use snapshot data!}
+%\KTC{I reworked this. Methods like balanced truncation or IRKA do \textit{not}
+%use snapshot data!}
 
@@
 Model reduction for linear-time-invariant systems (and other strongly
-structured dynamical systems) is quite mature \cite{wilcox_benner_rev}
-\KTC{Also cite classical refs on balanced truncation, Gugercin's IRKA,. Note
-that I moved \cite{wilcox_benner_rev} here because that review mainly deals
-with linear systems}, as system-theoretic properties (e.g., controllability,
+structured dynamical systems) is quite mature \cite{wilcox_benner_rev,moore,roberts,GugercinIRKA} as system-theoretic properties (e.g., controllability,
 observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
@@
 developing reduced-order models for general nonlinear dynamical systems is
-that such properties are difficult to assess for such systems. As a result, it
-is difficult to develop projection methods that preserve important
-dynamical-system properties in the reduced-order model, which can lead to
-models that are inaccurate or unstable.  To address this, researchers have
+that such properties are difficult to assess. As a result, it
+is difficult to develop reduced-order models that preserve important
+dynamical-system properties, a fact that can lead to inaccurate or unstable methods.  To address this, researchers have
 pursued several directions that aim to imbue reduced-order models of nonlinear
@@
 ``energy-based'' inner
-products that ensure non-increasing entropy in the ROM solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014}; symmetry
-transformations that \KTC{what?} ~\cite{sirovich_symmetry_trans},
-basis-adaptation methods that improve the ROM's accuracy \textit{a posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer}, subspace
-rotations that \KTC{what?} ~\cite{basis_rotation}, structure-preserving
+products that ensure non-increasing entropy in the ROM solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014};
+basis-adaptation methods that improve the ROM's accuracy \textit{a posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
 methods that enforce conservation \cite{carlberg_conservative_rom} or
 (port-)Hamiltonian/Lagrangian structure
-\cite{carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional}
-\KTC{Can use more here. See my presentation} in the ROM; and subgrid-scale
+\cite{carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional,LALL2003304} in the ROM; and subgrid-scale
 modeling methods that aim to improve accuracy by addressing the closure
@@
 We note that these techniques are often not mutually exclusive.
+%Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
 Residual-minimizing and space--time projection are the most relevant classes of

### 2019-09-19 main.tex
latest

@@
 state-space dimension is large (e.g., due to fine spatial resolution when
-discretizing a partial-differential-equations problem) and/or when the number
+discretizing a partial-differential-equation) and/or when the number
 of time instances is large (e.g., due to stability or accuracy requirements

### 2019-09-19 main.tex
latest

@@
 	time-continuous full-order-model residual over a low-dimensional 
-	trial subspace across a sequence of time windows and addresses key deficiencies in existing model-reduction approaches; e.g., the time-step and time-scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG.
+	trial subspace across a sequence of time windows and addresses key deficiencies in existing model-reduction approaches; e.g., the time step and time scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG.
 We consider two types of trial subspaces

### 2019-09-19 main.tex
latest

@@
 	reduces both the spatial and temporal dimensions of the full-order model.
-For both types of trial spaces, we consider two differing solution techniques: direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
+For both types of trial subspaces, we consider two differing solution techniques: direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
 Critically, we show that various instances of this framework yield existing model-reduction approaches, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
@@
 Residual minimization methods were extended-to and developed-for reduced-order models by a variety of researchers.
-Similar to LSFEM, residual minimization approaches for reduced-order models comprise finding the solution within a low-dimensional trial space that minimizes the residual norm of the full-order model.
+Similar to LSFEM, residual minimization approaches for reduced-order models comprise finding the solution within a low-dimensional trial subspace that minimizes the residual norm of the full-order model.
 Residual minimization approaches were first applied to static (i.e., no time-dependence) reduced-order models in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
@@
 present work distinguishes itself from Ref.~\cite{constantine_strom} in that
-(1.) in addition to trial spaces with space--time dimension reduction, we
-consider trial spaces with spatial dimension reduction only (i.e., a classic
+(1.) in addition to trial subspaces with space--time dimension reduction, we
+consider trial subspaces with spatial dimension reduction only (i.e., a classic
 spatial basis is employed, opposed to a space--time basis) -- in which case
@@
 minimizes the \textit{fully discrete} residual over the entire space--time
-domain within a given space--time trial space. 
+domain within a given space--time trial subspace. 
 %These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. 
@@
 (1.) we perform model reduction of the time-continuous system, (2.) we
-consider trial spaces with both space--time dimension reduction and spatial
+consider trial subspaces with both space--time dimension reduction and spatial
 dimension reduction, and (3.) residual minimization is performed over time
@@
 %\item The development of a time-continuous residual minimization framework for reduced-order models.
-\item The development of the \methodName\ (\methodAcronym) approach for dynamical systems. The approach sequentially minimizes the time-continuous full-order model residual over time windows.
-\item The derivation of the stationary conditions for \methodAcronymROMs\ leveraging spatial and space--time dimension reduction trial spaces.
-The case of spatial dimension reduction trial spaces is of particular interest: the stationary conditions are derived via the Euler--Lagrange equations and comprise a coupled
+\item The development of the \methodNameLower\ (\methodAcronym) approach for dynamical systems. The approach sequentially minimizes the time-continuous full-order model residual over time windows.
+\item The derivation of the stationary conditions for \methodAcronymROMs\ leveraging spatial and space--time dimension reduction trial subspaces.
+The case of spatial dimension reduction trial subspaces is of particular interest: the stationary conditions are derived via the Euler--Lagrange equations and comprise a coupled
 two-point Hamiltonian boundary value problem containing a forward and backward system.
@@
 Conditions under which \methodAcronym\ recovers Galerkin, LSPG, and space--time LSPG are presented.
-\item Error analysis of the \methodAcronym\ approach for trial spaces with spatial dimension reduction.
-In particular, it is shown that the error in \methodAcronymROMs\ is bounded \textit{a priori} by a combination of the error at the start of the window and the integrated FOM ODE residual evaluated at the FOM state projected onto the trial space.
-\item Numerical experiments for trial spaces with spatial dimension reduction demonstrating two key findings:
+\item Error analysis of the \methodAcronym\ approach for trial subspaces with spatial dimension reduction.
+In particular, it is shown that the error in \methodAcronymROMs\ is bounded \textit{a priori} by a combination of the error at the start of the window and the integrated FOM ODE residual evaluated at the FOM state projected onto the trial subspace.
+\item Numerical experiments for trial subspaces with spatial dimension reduction demonstrating two key findings:
 \begin{itemize}
@@
 \subsection{Projection-Based Reduced-Order Models}
-Projection-based ROMs generate approximate solutions to the FOM ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial space. For dynamical 
-systems, two distinct types of trial spaces are used: space--time trial spaces with spatial dimension reduction (\spatialAcronym) and space--time trial spaces with space--time dimension reduction (\spaceTimeAcronym).
-\spatialAcronym\ trial spaces are most commonly employed and are used in, e.g., the Galerkin and Least-Squares Petrov--Galerkin (LSPG) methods.
-\spaceTimeAcronym\ trial spaces are typically
-employed in space--time ROMs. The following subsections outline these two trial spaces, along with the Galerkin, LSPG, and 
+Projection-based ROMs generate approximate solutions to the FOM ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial subspace. For dynamical 
+systems, two distinct types of trial subspaces are used: space--time trial subspaces with spatial dimension reduction (\spatialAcronym) and space--time trial subspaces with space--time dimension reduction (\spaceTimeAcronym).
+\spatialAcronym\ trial subspaces are most commonly employed and are used in, e.g., the Galerkin and Least-Squares Petrov--Galerkin (LSPG) methods.
+\spaceTimeAcronym\ trial subspaces are typically
+employed in space--time ROMs. The following subsections outline these two trial subspaces, along with the Galerkin, LSPG, and 
 space--time LSPG method.
 
-\subsection{\spatialAcronym\ Trial Spaces}
-\spatialAcronym\ trial spaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial space, 
+\subsection{\spatialAcronym\ Trial Subspaces}
+\spatialAcronym\ trial subspaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial subspace, 
 \begin{align}\label{eq:spatial_subspace}
@@
 \end{align}
-where we refer to $\basismat \in \RRStar{\romdim}{\fomdim}$, with $\RRStar{\romdim}{\fomdim}$ denoting the set  $\{ \mathbf{X} \in \RR{\fomdim \times \romdim} \big| \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$, as the spatial trial basis matrix, $\trialspace$ as the spatial trial space, and $\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state.
+where we refer to $\basismat \in \RRStar{\romdim}{\fomdim}$, with $\RRStar{\romdim}{\fomdim}$ denoting the set  $\{ \mathbf{X} \in \RR{\fomdim \times \romdim} \big| \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$, as the spatial trial basis matrix, $\trialspace$ as the spatial trial subspace, and $\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state.
 %where $\basisvec_i \in \mathbb{R}^N$ are orthonormal basis vectors and $\stateIntercept \in \mathbb{R}^{\fomdim}$ the ``intercept''.
@@
 The basis vectors $\basisvec_i$ are typically constructed through, e.g., POD~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}, etc. 
-ROMs leveraging \spatialAcronym\ trial spaces approximate the state as,
+ROMs leveraging \spatialAcronym\ trial subspaces approximate the state as,
 \begin{equation}\label{eq:affine_trialspace}
@@
 where $\genstate \in \RR{\romdim} \otimes \timeSpace$, with $\genstate : [0,T] \rightarrow \mathbb{R}^{\romdim}$, are the generalized coordinates.
-Substituting the approximation given in~\eqref{eq:affine_trialspace} into~\eqref{eq:FOM} and performing orthogonal $\elltwo$ projection of the initial conditions onto the trial space, one obtains,
+Substituting the approximation given in~\eqref{eq:affine_trialspace} into~\eqref{eq:FOM} and performing orthogonal $\elltwo$ projection of the initial conditions onto the trial subspace, one obtains,
 \begin{equation}\label{eq:g_truncation}
@@
 \subsubsection{The Galerkin Method}
-The Galerkin method is (traditionally) obtained by the following steps: (1.) The approximation~\eqref{eq:affine_trialspace} is substituted into~\eqref{eq:FOM}, and (2.) the resulting residual is enforced to be $\stweightingMatArg{}$-orthogonal to the trial space, with $\stweightingMatArg{ } \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ }$ being a positive (semi-)definite weighting matrix (commonly taken to be $\mathbf{I}$).
+The Galerkin method is (traditionally) obtained by the following steps: (1.) The approximation~\eqref{eq:affine_trialspace} is substituted into~\eqref{eq:FOM}, and (2.) the resulting residual is enforced to be $\stweightingMatArg{}$-orthogonal to the trial subspace, with $\stweightingMatArg{ } \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ }$ being a positive (semi-)definite weighting matrix (commonly taken to be $\mathbf{I}$).
 This projection process results in what is known as the Galerkin method,
@@
 The FOM O$\Delta$E solution will exactly satisfy~\eqref{eq:lms}; the ROM solution will not.
-The LSPG method minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial space over each time step.
+The LSPG method minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial subspace over each time step.
 %We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
@@
 methods are often equipped with more favorable error bounds than non-space--time methods. The approach proposed in this work displays commonalities 
-with space--time methods and, as such, we briefly outline \spaceTimeAcronym\ trial spaces and space--time LSPG.
-
-Space--time methods leverage the \spaceTimeAcronym\ trial space $\stspaceST$,
+with space--time methods and, as such, we briefly outline \spaceTimeAcronym\ trial subspaces and space--time LSPG.
+
+Space--time methods leverage the \spaceTimeAcronym\ trial subspace $\stspaceST$,
 \begin{equation}\label{eq:sttrialspace_def}
@@
 space--time basis matrix comprises time-dependent \textit{functions}, while the generalized coordinates are a stationary vector. This is fundamentally different than  
-\spatialAcronym\ trial spaces, in where the generalized coordinates are time-dependent functions while the basis matrix is stationary.
+\spatialAcronym\ trial subspaces, in where the generalized coordinates are time-dependent functions while the basis matrix is stationary.
 
@@
 \end{equation}
-where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial space (e.g., $\elltwo$).
+where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
 Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
@@
 This paper introduced a windowed least-squares reduced-order modeling (\methodAcronym) formulation for dynamical systems.
-The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time windows.
-The formulation is applicable to both spatial and space--time dimension reduction trial spaces.
+The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial subspace over a series of time windows.
+The formulation is applicable to both spatial and space--time dimension reduction trial subspaces.
 The Galerkin, Least-Squares Petrov--Galerkin (LSPG) and space--time LSPG methods were shown to be specific instances of the proposed method
@@
 Two differing techniques to solve the least-squares minimization problems were considered: direct (discretize then minimize) and indirect (minimize then discretize). 
-The case of \spatialAcronym\ trial spaces proved particularly interesting: it was shown that indirect methods comprise solving a coupled two-point 
+The case of \spatialAcronym\ trial subspaces proved particularly interesting: it was shown that indirect methods comprise solving a coupled two-point 
 Hamiltonian boundary value problem.

### 2019-09-19 section3/general_form.tex
latest

@@
 \section{Windowed Least-Squares Formulation}\label{sec:tclspg} This section
-outlines the proposed \methodName\ 
+outlines the proposed \methodNameLower\ 
 formulation (\methodAcronym). In contrast to (1.) the Galerkin method, which minimizes the 
@@
 time-continuous FOM ODE residual over a series of
-\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial spaces. 
+\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
 In this section, we start by outlining the \methodAcronym\ formulation  
-for a general trial space. We then examine the \spatialAcronym\ 
-trial space, followed by the \spaceTimeAcronym\ trial space. In each case, we derive the stationary conditions 
+for a general trial subspace. We then examine the \spatialAcronym\ 
+trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
 associated with the minimization problems.
@@
 \end{figure}
-Over each time window, we introduce the space--time trial space,
+Over each time window, we introduce the space--time trial subspace,
 $$\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,$$ 
@@
 where $\mathbb{P}^n: \RR{N} \otimes \timeSpaceArg{n} \rightarrow \stspaceArg{n}$ is an orthogonal projection onto the 
-trial space (e.g., $\elltwo$).
+trial subspace (e.g., $\elltwo$).
 %\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing
@@
 \begin{remark}
-It is possible to construct the trial spaces such that they have continuous support between time windows through the 
+It is possible to construct the trial subspaces such that they have continuous support between time windows through the 
 use of, e.g., affine transformations.
@@
 %projected onto the trial space; (e.g., via for $L^2$ projection).
-We now define the trial spaces $\stspaceArg{n}$. We consider \spatialAcronym\ trial spaces and \spaceTimeAcronym\ trial spaces. 
-We leave further investigation into other trial spaces, such as nonlinear manifolds~\cite{leeCarlberg}, as a subject for future work.
+We now define the trial subspaces $\stspaceArg{n}$. We consider \spatialAcronym\ trial subspaces and \spaceTimeAcronym\ trial subspaces. 
+We leave further investigation into other trial subspaces, such as nonlinear manifolds~\cite{leeCarlberg}, as a subject for future work.
 %A variety of trial spaces have been examined in the model-order reduction community.

### 2019-09-19 section4/numerical_methods.tex
latest

@@
 This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system.
-We focus primarily on direct and indirect methods for \spatialAcronym\ trial spaces,
-and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial spaces. 
+We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces,
+and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
 %Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems.
@@
 
-In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial spaces. We consider:
+In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial subspaces. We consider:
 

### 2019-09-19 section4/spacetime_numerical.tex
latest

@@
 \subsection{\spaceTimeAcronym\ Trial Spaces: Direct and Indirect Methods}
-We now briefly consider \spaceTimeAcronym\ trial spaces. 
-For \spaceTimeAcronym\ trial spaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
+We now briefly consider \spaceTimeAcronym\ trial subspaces. 
+For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
 dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
@@
 \begin{remark}
-For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial space recovers ST-LSPG.
+For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
 \end{remark}
@@
 $$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \ldots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$
-Leveraging these collocation points, the indirect method for \methodAcronym\ with a \spaceTimeAcronym\ trial space yields,
+Leveraging these collocation points, the indirect method for \methodAcronym\ with a \spaceTimeAcronym\ trial subspace yields,
 \begin{equation}\label{eq:obj_gen_slab2} 
@@
 \subsection{\spaceTimeAcronym\ Trial Space: Summary}
-As was just shown, \spaceTimeAcronym\ trial spaces yield a series of space--time algebraic systems over each window.
-As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial spaces,
+As was just shown, \spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window.
+As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces,
 a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
-space--time trial space yields a series of dense (non)linear systems to be solved over each window.
+space--time trial subspace yields a series of dense (non)linear systems to be solved over each window.

### 2019-09-19 section4/spatial_numerical.tex
latest

@@
 %is the discrete residual resulting from the linear multistep scheme and 
-where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. \methodAcronym\ solves the minimization problem over each time window, 
+where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. 
+Over each window, \methodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
 \begin{align*}
-&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
+&\underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{minimize } }
 \objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
-& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
+& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{n,0} =
 \begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
 \approxstateIC & n=1. \end{cases} \end{align*}
+%\methodAcronym\ solves the minimization problem over each time window, 
+%\begin{align*}
+%&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
+%\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
+%& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
+%\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
+%\approxstateIC & n=1. \end{cases} \end{align*}
 Defining $\genstateDiscreteArg{n,i} \in \RR{\romdim}$ to be the discrete approximation to the generalized coordinates at the $i$th time-step instance over the $n$th window, 
@@
 \begin{split}
-& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }
+& \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{minimize } }
 \objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
 & \text{subject to} \hspace{0.2 in}
-\genstateDiscreteArg{n,0} =
+\genstateyDiscreteArg{n,0} =
 \begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
@@
 \end{equation}
+%\begin{equation}\label{eq:obj_gen_lms_final}
+%\begin{split}
+%& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }
+%\objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
+%& \text{subject to} \hspace{0.2 in}
+%\genstateDiscreteArg{n,0} =
+%\begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
+%\basisspace^T(\stateFOMIC - \stateIntercept)& n=1. \end{cases} 
+%\end{split}
+%\end{equation}
 The optimization problem takes the form of a \textit{weighted nonlinear least-squares problem}, 
@@
 \begin{algorithm}
-\caption{\spatialAcronym\ trial space: algorithm for the direct solution technique with the Gauss--Newton method and a linear multistep method over the $n$th window}
+\caption{\spatialAcronym\ trial subspace: algorithm for the direct solution technique with the Gauss--Newton method and a linear multistep method over the $n$th window}
 \label{alg:colloc_gn}
@@
 
-\begin{algorithm} \caption{\spatialAcronym\ trial space: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
+\begin{algorithm} \caption{\spatialAcronym\ trial subspace: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}

### 2019-09-23 cavity.tex

@@
 The reduced-order models leverage POD to construct the trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to construct the weighting 
-matrix $\stweightingMatArg{n}$. The process used to construct the initial conditions and trial spaces for the ROMs are as follows:
+matrix $\stweightingMatArg{n}$. The process used to construct the initial conditions and trial subspaces for the ROMs are as follows:
 \begin{enumerate}
@@
 \item Evolve the FOM for $t \in [0,400]$.
-\item Reset the time-coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 100$ and $K = \{136,193\}$ over $t \in [0,100]$ to construct two trial spaces comprising $\approx 95\%$ and $\approx 97\%$ of the snapshot energy, respectively.
+\item Reset the time-coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 100$ and $K = \{136,193\}$ over $t \in [0,100]$ to construct two trial subspaces comprising $\approx 95\%$ and $\approx 97\%$ of the snapshot energy, respectively.
 \item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 100$, $n_s = 129$ to obtain the sampling point matrix of dimension $\stweightingMatOneArg{} \in \{0,1\}^{1603 \times \fomdim }$. 
@@
 the solution for $t \in [0,100]$.
-This comprises the same time interval used to construct the trial space.
+This comprises the same time interval used to construct the trial subspace.
 First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$ state error of the various reduced-order models.

### 2019-09-23 numerical_experiments.tex

@@
 \section{Numerical Experiments}\label{sec:numerical_experiments}
-This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial spaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity.
+This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity.
 In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs.
@@
 The FOM is evolved for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$.
-The \spatialAcronym\ trial space is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $.
-The resulting trial space comprises $99.99\%$ of the snapshot matrix energy.
+The \spatialAcronym\ trial subspace is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $.
+The resulting trial subspace comprises $99.99\%$ of the snapshot matrix energy.
 %  The solution is saved every other time-step, resulting in 500 solutions snapshots.
@@
 The trial basis is selected to be of dimension $K = 46$, which corresponds to a $99.99\%$ energy criterion.
-No affine transformation is used in the construction of the trial space, i.e., $\stateIntercept = \bz$.
+No affine transformation is used in the construction of the trial subspace, i.e., $\stateIntercept = \bz$.
 

### 2019-09-23 section3/space_time.tex
latest

@@
 \subsection{\spaceTimeAcronym\ Trial Spaces}
-We now briefly consider \spaceTimeAcronym\ trial spaces. Over 
-each window, we introduce the space--time trial space
+We now briefly consider \spaceTimeAcronym\ trial subspaces. Over 
+each window, we introduce the space--time trial subspace
 $$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
@@
 Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
-the \spaceTimeAcronym\ trial space computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+the \spaceTimeAcronym\ trial subspace computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
 \begin{align}\label{eq:obj_gen_slab_spacetime}
 \begin{split}
-&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
-      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
+&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptArg{n} ), \\ 
+      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =
   \begin{cases} 
@@
 \end{align}
+
+%\begin{align}\label{eq:obj_gen_slab_spacetime}
+%\begin{split}
+%&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
+%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
+%  \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0)
+% & n=1. \end{cases}
+%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
+%%\genstateIC & n=1, \end{cases} 
+%\end{split}
+%\end{align}
 
@@
 \subsubsection{Stationary Conditions} 
-The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial spaces is as follows: in \spaceTimeAcronym\ trial spaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial spaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$.
+The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$.
 Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
@@
 %n=1. \end{cases} \end{equation}
-For \spaceTimeAcronym\ trial spaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.
+For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.

### 2019-09-23 section3/tclsrm_formulation.tex

@@
 This section outlined the \methodAcronym\ method for model reduction.
-In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial space for a series of time windows: \methodAcronym\
+In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows: \methodAcronym\
 computes the generalized coordinates $\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the minimization problems,
@@
 subject to the boundary and continuity conditions.
-For the \spatialAcronym\ trial spaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
+For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
 \begin{align*}\
@@
 subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
-trial spaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
+trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
 controller that enforces the residual minimization property.
 
-\spaceTimeAcronym\ trial spaces were additionally considered.
-For \spaceTimeAcronym\ trial spaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.
+\spaceTimeAcronym\ trial subspaces were additionally considered.
+For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.

### 2019-10-02 main.tex
eric latest

@@
 	formulation for dynamical systems. The proposed approach minimizes the
-	time-continuous full-order-model residual over a low-dimensional 
-	trial subspace across a sequence of time windows and addresses key deficiencies in existing model-reduction approaches; e.g., the time step and time scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG.
+	time-continuous full-order-model residual over a low-dimensional trial
+	subspace across a sequence of time windows and addresses key deficiencies in
+	existing model-reduction approaches, e.g., the time-step and time-scheme
+	dependence of least-squares Petrov--Galerkin (LSPG) projection, \textit{a
+	posteriori} error bounds that grow exponentially in time for Galerkin and
+	LSPG projection, and the offline cost of space--time LSPG projection.
+	\KTC{How do we address the offline cost of ST-LSPG?}
 We consider two types of trial subspaces
 	within the proposed framework:
-	one that reduces the spatial dimension of the full-order model, and one that
+	one that reduces only the spatial dimension of the full-order model, and one that
 	reduces both the spatial and temporal dimensions of the full-order model.
-For both types of trial subspaces, we consider two differing solution techniques: direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
-Critically, we show that various instances of this framework yield existing model-reduction approaches, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
-Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.
+	For each type of trial subspaces, we consider two different solution
+	techniques: direct (i.e., discretize-then-optimize) and indirect (i.e.,
+	optimize-then-discretize). Critically, we show that various instances of
+	this framework yield existing model-reduction approaches, namely Galerkin,
+	LSPG, and space–time LSPG projection. Numerical experiments on the
+	compressible Euler and Navier--Stokes equations demonstrate that, at an
+	increased computational cost, the \methodAcronym\ formulation can yield more
+	accurate and physically relevant solutions than Galerkin and
+	LSPG projection.  \KTC{So which is better? Should we emphasize that it is
+	more general?}
 
@@
 state-space dimension is large (e.g., due to fine spatial resolution when
-discretizing a partial-differential-equation) and/or when the number
-of time instances is large (e.g., due to stability or accuracy requirements
-that limit the time step).  When the application is time-critical or
-many-query in nature, analysts must replace such large-scale parameterized
+discretizing a partial differential equation) and/or when the number of time
+instances is large (e.g., due to time-step limitations incurred by stability
+or accuracy considerations).  When the application is time critical or many
+query in nature, analysts must replace such large-scale parameterized
 dynamical-system models with a low-cost approximation that makes the
@@
 \textit{offline} stage that computes a low-dimensional ``trial'' subspace on
-which the (parameterized) dynamical-system state can be well approximated
-(e.g., by computing state ``snapshots'' at different parameter instances, by
-solving Lyapunov equations). Second, these methods execute an inexpensive
+which the dynamical-system state can be well approximated (e.g., by computing
+state ``snapshots'' at different time and parameter instances, by solving
+Lyapunov equations). Second, these methods execute an inexpensive
 \textit{online} stage during which they compute approximations to the
 dynamical-system trajectory that reside on this trial subspace via a
-residual-minimization or projection process.
+projection process.
 
@@
 developing reduced-order models for general nonlinear dynamical systems is
-that such properties are difficult to assess. As a result, it
+that such properties are difficult to quantitatively assess. As a result, it
 is difficult to develop reduced-order models that preserve important
-dynamical-system properties, a fact that can lead to inaccurate or unstable methods.  To address this, researchers have
-pursued several directions that aim to imbue reduced-order models of nonlinear
+dynamical-system properties, which often results in inaccurate or unstable
+methods.  To address this, researchers have pursued several directions that
+aim to imbue reduced-order models for nonlinear
 dynamical systems with properties that can improve robustness and accuracy.
@@
 efforts include residual-minimizing
-projection that imbues the ROM solution with a notion of optimality~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis};
+projection that equips the ROM solution with a notion of optimality~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis};
 space--time
@@
 products that ensure non-increasing entropy in the ROM solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014};
-basis-adaptation methods that improve the ROM's accuracy \textit{a posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
+basis-adaptation methods that improve the ROM's accuracy \textit{a
+posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer,etter2019online}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
 methods that enforce conservation \cite{carlberg_conservative_rom} or
 (port-)Hamiltonian/Lagrangian structure
-\cite{carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional,LALL2003304} in the ROM; and subgrid-scale
+\cite{LALL2003304,carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional} in the ROM; and subgrid-scale
 modeling methods that aim to improve accuracy by addressing the closure
@@
 methods to the current work and comprise the focus of the following review.
-\KTC{Shouldn't we review space-time also?}
 
@@
 %\end{itemize}
-Least-squares residual minimization as a numerical solution technique for (partial) differential equations was pioneered in the finite element community and led to what is known as the least-squares finite element method (LSFEM)~\cite{ls_review,bochev_leastsquares}.
-Motivated by the fact that the Galerkin method can be inaccurate and unstable for nonlinear and non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic functional comprising the residual norm of the partial differential equation of interest and (2.) leveraging variational principles to derive the stationary conditions of the quadratic functional; i.e, LSFEM can be thought of as finding the solution within a finite element trial space the minimizes the $2$-norm of the PDE residual.
-Least-squares methods often exhibit numerous advantages (e.g., least-squares system are symmetric and positive definite) over the more standard Galerkin discretization.
-
-Residual minimization methods were extended-to and developed-for reduced-order models by a variety of researchers.
-Similar to LSFEM, residual minimization approaches for reduced-order models comprise finding the solution within a low-dimensional trial subspace that minimizes the residual norm of the full-order model.
-Residual minimization approaches were first applied to static (i.e., no time-dependence) reduced-order models in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
-Specifically, Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg} formulated the residual minimization problem for dynamical systems by sequentially minimizing the \textit{fully discrete} residual (i.e., the residual after temporal discretization) at each time-step instance.
-This formulation is most commonly known as the \textit{Least--Squares Petrov--Galerkin} (LSPG) method.
-Ref.~\cite{carlberg_lspg_v_galerkin} provided numerous analyses on this formulation and examined connections with the standard Galerkin method and Petrov--Galerkin methods.
-Critically, it was shown that (1.) for various time-stepping methods LSPG has an equivalent continuous form that corresponds to a Petrov--Galerkin projection and (2.) the standard Galerkin method minimizes the time-instantaneous residual of the full-order model.
-Numerous numerical experiments have demonstrated that LSPG yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
-The common physical explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.
-
-The LSPG approach only guarantees residual minimization over a single time-step, and as such is not associated with residual minimization of the full dynamical system.  
-Indeed, LSPG is not equipped with \textit{a priori} guarantees of stability or convergence, even for linear time-invariant systems~\cite{bui_thesis}.
-Further, as it is defined at the fully discrete level, the performance of LSPG is subject to the time-step size: changing the time-step size changes the window over which the residual is minimized \textit{as well as} the temporal discretization error.
-As a result, LSPG provides optimal results at an intermediary time-step size.
-For too small a time-step, LSPG yields results similar to the Galerkin approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin.
-For too-large a time-step, the accuracy of LSPG degrades.
-It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
-The LSPG approach is additionally sensitive to the time discretization scheme.
-When used with explicit temporal discretization, for instance, LSPG again reverts to Galerkin.
-To gain the enhanced stability associated with LSPG, the underlying time-discretization scheme must be implicit with an appropriate time-step.
-This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.
+
+\KTC{Why are we reviewing LSFEM? This isn't a ROM. Don't think it's relevant}
+Least-squares residual minimization as a numerical solution technique for
+partial differential equations was pioneered in the finite-element community
+and led to what is known as the least-squares finite-element method
+(LSFEM)~\cite{ls_review,bochev_leastsquares}. 
+\KTC{Specifically, LSFEM was designed to solve the inf--sup stability problem.
+Be more precise.}
+Motivated by the fact that the
+Galerkin method can be inaccurate and unstable for nonlinear and
+non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic
+functional comprising the residual norm of the partial differential equation
+of interest and (2.) leveraging variational principles to derive the
+stationary conditions of the quadratic functional; i.e, LSFEM can be thought
+of as finding the solution within a finite element trial space the minimizes
+the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous
+advantages (e.g., least-squares system are symmetric and positive definite)
+\KTC{This is not an `advantage'. It is guaranteed to be inf--sup stable.}
+over the more standard Galerkin discretization.
+
+Researchers also developed residual-minimization methods for model reduction.
+Similar to LSFEM, residual minimization in this context associates with
+computing the solution within a low-dimensional trial space that minimizes the
+$\ell^2$-norm of the full-order-model residual. \KTC{haven't defined
+`full-order model'} Researchers have developed such residual-minimizing
+model-reduction methods for both static systems (i.e., systems without
+time-dependence)
+\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
+and dynamical
+systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
+In the latter category,
+Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
+formulated the residual minimization problem for dynamical systems by
+sequentially minimizing the \textit{time-discrete} full-order-model residual
+(i.e., the residual arising after applying time discretization) at each time
+instance on the time-discretization grid. This formulation is often referred
+to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
+Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
+formulation and examined its connections with Galerkin projection. Critically,
+this work demonstrated that (1) under certain conditions, LSPG can be derived
+from a Petrov--Galerkin projection executed on the time-continuous
+full-order-model residual, and (2) LSPG and Galekrin projection are equivalent
+in the limit as the time step goes to zero (i.e., Galerkin projection
+minimizes the time-instantaneous full-order model residual). Numerous
+numerical experiments have demonstrated that LSPG projection often yields more
+accurate and stable solutions than Galerkin
+projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
+The common intuitive explanation for this improved performance is that, by
+minimizing the full-order-model residual over a finite time window (rather
+than time instantanously), LSPG projection computes solutions that are more
+accurate over a larger part the trajectory compared with Galerkin projection.
+
+\KTC{HERE}
+
+The LSPG approach sequentially performs residual minimization over a single
+time-step, and as such is not associated with residual minimization over the
+entire time interval. \KTC{This is what ST-LSPG does. Why do we ignore it? the
+full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
+guarantees of stability or convergence, even for linear time-invariant
+systems~\cite{bui_thesis}. Further, LSPG performs residual minimization over a
+time step, its performance is affected by the time-step size in two distinct
+ways: changing the time-step size modifies the window over which the residual
+is minimized \textit{as well as} the time-discretization error. As a result,
+LSPG often yields the lowest error for an intermediate value of the time step.
+For too small a time-step, LSPG yields results similar to the Galerkin
+approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers
+Galerkin. For too-large a time-step, the accuracy of LSPG degrades. It is
+unclear if this is due to the numerical error associated with enlarging the
+time-step size, or rather if it is due to the size of the window the residual
+is being minimized over. The LSPG approach is additionally sensitive to the
+time discretization scheme. When used with explicit temporal discretization,
+for instance, LSPG again reverts to Galerkin. To gain the enhanced stability
+associated with LSPG, the underlying time-discretization scheme must be
+implicit with an appropriate time-step. This can increase the cost of the
+reduced-order model, especially in the case where the optimal time-step
+required by LSPG is small.
 
@@
 windows; Ref.~\cite{choi_stlspg} considers space--time bases and minimizes the
-fully-discrete residual over the entire domain.
+fully-discrete residual over the entire time domain.
 

### 2019-10-02 main.tex

@@
 %\title{Windowed least--squares reduced-order models for dynamical systems}
-\title{Windowed Residual-Minimization Model Reduction for Dynamical Systems}
+\title{Windowed residual-minimization model reduction for dynamical systems}
 
@@
 
+Residual minimization methods were extended-to and developed-for reduced-order
+models by a variety of researchers. Similar to LSFEM, residual minimization
+approaches for reduced-order models comprise finding the solution within a
+low-dimensional trial subspace that minimizes the residual norm of the
+full-order model. Residual minimization approaches were first applied to
+static (i.e., no time-dependence) reduced-order models
+in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
+and then extended to dynamical systems
+in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
+Specifically,
+Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
+formulated the residual minimization problem for dynamical systems by
+sequentially minimizing the \textit{fully discrete} residual (i.e., the
+residual after temporal discretization) at each time-step instance. This
+formulation is most commonly known as the \textit{Least--Squares
+Petrov--Galerkin} (LSPG) method. Ref.~\cite{carlberg_lspg_v_galerkin} provided
+numerous analyses on this formulation and examined connections with the
+standard Galerkin method and Petrov--Galerkin methods. Critically, it was
+shown that (1.) for various time-stepping methods LSPG has an equivalent
+continuous form that corresponds to a Petrov--Galerkin projection and (2.) the
+standard Galerkin method minimizes the time-instantaneous residual of the
+full-order model. Numerous numerical experiments have demonstrated that LSPG
+yields more stable (although not always more accurate) solutions than the
+Galerkin
+method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
+The common physical explanation for this improved performance is that, by
+minimizing the fully discrete residual, LSPG minimizes the residual over a
+\textit{finite-sized time-window}, rather than instantaneously as is done in
+the Galerkin method.
+
+\KTC{Let's discuss. What specifically do we want to point out about
+deficiencies in LSPG in this section? Should we also look in ST-LSPG? It seems you're trying
+to say that we don't like how the time step size \textit{and} time
+discretization method dictate so much about the method (I agree with this). If
+so, we should structure this to include other discretize-then-optimize methods
+I think. Or we could just focus on the time-step aspect (i.e., not the time
+discretization method), in which case we should focus on LSPG.}
 \KTC{HERE}
@@
 guarantees of stability or convergence, even for linear time-invariant
-systems~\cite{bui_thesis}. Further, LSPG performs residual minimization over a
-time step, its performance is affected by the time-step size in two distinct
-ways: changing the time-step size modifies the window over which the residual
-is minimized \textit{as well as} the time-discretization error. As a result,
-LSPG often yields the lowest error for an intermediate value of the time step.
-For too small a time-step, LSPG yields results similar to the Galerkin
-approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers
-Galerkin. For too-large a time-step, the accuracy of LSPG degrades. It is
-unclear if this is due to the numerical error associated with enlarging the
-time-step size, or rather if it is due to the size of the window the residual
-is being minimized over. The LSPG approach is additionally sensitive to the
-time discretization scheme. When used with explicit temporal discretization,
-for instance, LSPG again reverts to Galerkin. To gain the enhanced stability
-associated with LSPG, the underlying time-discretization scheme must be
-implicit with an appropriate time-step. This can increase the cost of the
-reduced-order model, especially in the case where the optimal time-step
-required by LSPG is small.
+systems~\cite{bui_thesis}. Further, the performance of LSPG depends on the
+time step in two distinct ways: changing the time-step size modifies both the
+time window over which LSPG minimizes the residual \textit{as well as} the
+time-discretization error of the full-order model on which LSPG is based. As a
+result, LSPG often yields the smallest error for an intermediate value of the
+time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}). In
+particular, as the time step decreases, the accuracy of LSPG approaches that
+of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
+is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
+as the time step increases, the time discretization error of the full-order
+model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
+additionally sensitive to the time discretization scheme. When used with
+explicit temporal discretization, for instance, LSPG again reverts to
+Galerkin. To gain the enhanced stability associated with LSPG, the underlying
+time-discretization scheme must be implicit with an appropriate time-step.
+This can increase the cost of the reduced-order model, especially in the case
+where the optimal time-step required by LSPG is small.
 
@@
 residual minimization approaches for reduced-order models. To this end, we
-develop a reduced-order modeling approach that minimizes the FOM ODE residual
-over a series of time windows. The key novelties of the work are two fold.
+propose the \methodName\ method for model reduction, which minimizes the FOM ODE residual
+over a sequence of time windows. The key novelties of the work are two fold.
 First, the residual minimization problem is formulated to minimize the
-\textit{time-continuous} residual. By working from a continuous viewpoint, the
+\textit{time-continuous} residual. By working from a time-continuous viewpoint, the
 present formulation decouples the underlying temporal discretization scheme
 from the residual minimization problem; thus addressing a key deficiency of
-LSPG. The continuous viewpoint additionally exposes two differing methods that
-can be used to solve the minimization problems, namely \textit{discretize and
-then optimize} (direct) methods and \textit{optimize and then discretize}
-(indirect) methods. The second novelty is that the residual is minimized over
-\textit{time windows}, as opposed to over a single time step or the entire
-time domain. This allows for maximum flexibility in terms of computational
-cost and numerical accuracy. We demonstrate that this formulation recovers the
-Galerkin, LSPG, and space--time LSPG methods. Figure~\ref{fig:flowchart}
-depicts how the proposed method interfaces with existing approaches.
+LSPG \KTC{This is also true for ST-LSPG and maybe other discretization-based
+methods? Have to check}. The time-continuous viewpoint additionally exposes
+two different methods that
+can be used to solve the residual-minimization problems, namely
+a \textit{discretize-then-optimize} (i.e., direct) method and
+an \textit{optimize-then-discretize} (i.e., indirect) method. The second novelty
+is that the residual is minimized over \textit{time windows} rather than over
+a single time step or the entire time domain. This equips the method with
+additionaly flexibility that enables it
+to explore more fine-grained tradeoffs in computational cost and error. We
+demonstrate that this method can be viewed as a generalization of existing
+methods, as particular instances of this formulation recover Galerkin, LSPG,
+and space--time LSPG projection. Figure~\ref{fig:flowchart} depicts how the
+proposed method is related to existing approaches.
 
@@
 \includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
-\caption{Flow chart of \methodAcronym.} 
+\caption{Relationship diagram for the \methodAcronym\ framework for model
+	reduction.} 
 \label{fig:flowchart} 

### 2019-10-02 main.tex
fix intro

@@
 
-Residual minimization methods were extended-to and developed-for reduced-order
-models by a variety of researchers. Similar to LSFEM, residual minimization
-approaches for reduced-order models comprise finding the solution within a
-low-dimensional trial subspace that minimizes the residual norm of the
-full-order model. Residual minimization approaches were first applied to
-static (i.e., no time-dependence) reduced-order models
-in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
-and then extended to dynamical systems
-in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
-Specifically,
-Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
-formulated the residual minimization problem for dynamical systems by
-sequentially minimizing the \textit{fully discrete} residual (i.e., the
-residual after temporal discretization) at each time-step instance. This
-formulation is most commonly known as the \textit{Least--Squares
-Petrov--Galerkin} (LSPG) method. Ref.~\cite{carlberg_lspg_v_galerkin} provided
-numerous analyses on this formulation and examined connections with the
-standard Galerkin method and Petrov--Galerkin methods. Critically, it was
-shown that (1.) for various time-stepping methods LSPG has an equivalent
-continuous form that corresponds to a Petrov--Galerkin projection and (2.) the
-standard Galerkin method minimizes the time-instantaneous residual of the
-full-order model. Numerous numerical experiments have demonstrated that LSPG
-yields more stable (although not always more accurate) solutions than the
-Galerkin
-method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
-The common physical explanation for this improved performance is that, by
-minimizing the fully discrete residual, LSPG minimizes the residual over a
-\textit{finite-sized time-window}, rather than instantaneously as is done in
-the Galerkin method.
-
 \KTC{Let's discuss. What specifically do we want to point out about

### 2019-10-02 main.tex
up

--- 2019-10-02 main.tex
@@
 residual minimization approaches for reduced-order models. To this end, we
-propose the \methodName\ method for model reduction, which minimizes the FOM ODE residual
-over a sequence of time windows. The key novelties of the work are two fold.
-First, the residual minimization problem is formulated to minimize the
-\textit{time-continuous} residual. By working from a time-continuous viewpoint, the
-present formulation decouples the underlying temporal discretization scheme
-from the residual minimization problem; thus addressing a key deficiency of
-LSPG \KTC{This is also true for ST-LSPG and maybe other discretization-based
-methods? Have to check}. The time-continuous viewpoint additionally exposes
-two different methods that
-can be used to solve the residual-minimization problems, namely
-a \textit{discretize-then-optimize} (i.e., direct) method and
-an \textit{optimize-then-discretize} (i.e., indirect) method. The second novelty
-is that the residual is minimized over \textit{time windows} rather than over
-a single time step or the entire time domain. This equips the method with
-additionaly flexibility that enables it
-to explore more fine-grained tradeoffs in computational cost and error. We
+propose the \methodNameLower\ (\methodAcronym) method for model reduction,
+which minimizes the FOM ODE residual over a sequence of arbitrarily defined
+time windows. The key novelties of the work are twofold.  First, the method
+minimizes the \textit{time-continuous}  residual (i.e., that associated with
+the full-order model ODE). By adopting a time-continuous viewpoint, the
+formulation decouples the underlying temporal discretization scheme from the
+residual minimization problem, thus addressing a key deficiency of LSPG
+\KTC{This is also true for ST-LSPG and maybe other discretization-based
+methods? Have to check}. Time-continuous residual minimization also exposes
+two different solution methods,  namely a \textit{discretize-then-optimize}
+(i.e., direct) method and an \textit{optimize-then-discretize} (i.e.,
+indirect) method. The second novelty is that the method sequentially minimizes
+the residual over arbitrarily defined \textit{time windows} rather than
+sequentially minimizing the residual over time steps or over the entire time
+domain. This equips the method with additionaly flexibility that enables it to
+explore more fine-grained tradeoffs between computational cost and error. We
 demonstrate that this method can be viewed as a generalization of existing
-methods, as particular instances of this formulation recover Galerkin, LSPG,
-and space--time LSPG projection. Figure~\ref{fig:flowchart} depicts how the
-proposed method is related to existing approaches.
+methods, because particular instances of this formulation are equivalent to
+Galerkin, LSPG, and space--time LSPG projection. Figure~\ref{fig:flowchart}
+depicts how the proposed \methodAcronym\ method is related to existing
+approaches.
 

### 2019-10-02 main.tex
p 6

@@
 
-The layout of the rest of this paper is as follows: Section~\ref{sec:math} outlines the problem statement of interest, along with the Galerkin, LSPG, and space--time LSPG ROMs.
-Section~\ref{sec:tclspg} outlines the \methodAcronym\ formulation proposed in this work.
-Section~\ref{sec:numerical_techniques} outlines numerical techniques for solving \methodAcronymROMs.
-Both direct and indirect methods are discussed.
-Next, Section~\ref{sec:analysis} provides equivalency conditions and error analysis for \methodAcronymROMs.
-Section~\ref{sec:numerical_experiments} presents numerical experiments.
-Conclusions and perspectives are provided in Section~\ref{sec:conclude}.
-The notation used is as follows: We denote vector functionals in stylized bold text, i.e., $\boldsymbol x$.
-We denote vectors in standard bold text, i.e., $\mathbf{x}$.
-Matrices are denoted with capital bold letters, i.e., $\mathbf{X} \equiv \begin{bmatrix} \mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$.
-Vector and functional spaces are denoted with calligraphic fonts, i.e., $\mathcal{X}$.
+The paper proceeds as follows: Section~\ref{sec:math}
+outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
+space--time LSPG ROMs; Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
+formulation; Section~\ref{sec:numerical_techniques}
+outlines numerical techniques for solving \methodAcronymROMs, including both direct and
+indirect methods; Section~\ref{sec:analysis} provides
+equivalence conditions and error analysis for \methodAcronymROMs;
+Section~\ref{sec:numerical_experiments} presents numerical experiments;
+Section~\ref{sec:conclude} provides conclusions and perspectives.
+We denote vector-valued functions with stylized bold text (i.e., $\boldsymbol
+x$), vectors with standard bold text (i.e., $\mathbf{x}$), 
+matrices with capital bold letters (i.e., $\mathbf{X} \equiv \begin{bmatrix}
+\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
+calligraphic fonts (i.e., $\mathcal{X}$).
 %Model reduction of nonlinear dynamical systems is a growing research area.
@@
 %\begin{comment} 
-\section{Mathematical Setting}\label{sec:math}
+\section{Mathematical setting}\label{sec:math}
 We consider the system of nonlinear ordinary differential equations (ODEs),
 \begin{equation}\label{eq:FOM}
- \frac{d}{dt}\stateFOMArg{}{t} = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) = \stateFOMIC, \qquad t \in [0,T],
+ \frac{d\stateFOM}{dt} = \velocity(\stateFOM,t), \qquad \stateFOM(0) = \stateFOMIC,
 \end{equation}
-where $T \in \mathbb{R}^+$ is the final time, $\stateFOM \in \RR{\fomdim} \otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$ denotes the state implicitly defined to be the solution to~\eqref{eq:FOM}, $\timeSpace$ is the set of functions acting on $[0,T]$, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ are the initial conditions, and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow \mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,t) \mapsto \velocity(\stateyDiscrete,t)$ is the velocity.
-Equation~\eqref{eq:FOM} will be referred to as the ``full-order model" (FOM) ODE.
-
-Directly solving~\eqref{eq:FOM} can be computationally expensive due to both the high dimensionality of the state and the range of time-scales that needs to be resolved.
-For time-critical applications, it is important to devise computationally efficient approximations to~\eqref{eq:FOM}.
-Projection-based ROMs are
-one such promising approach.
-
-\subsection{Projection-Based Reduced-Order Models}
-Projection-based ROMs generate approximate solutions to the FOM ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial subspace. For dynamical 
-systems, two distinct types of trial subspaces are used: space--time trial subspaces with spatial dimension reduction (\spatialAcronym) and space--time trial subspaces with space--time dimension reduction (\spaceTimeAcronym).
-\spatialAcronym\ trial subspaces are most commonly employed and are used in, e.g., the Galerkin and Least-Squares Petrov--Galerkin (LSPG) methods.
-\spaceTimeAcronym\ trial subspaces are typically
-employed in space--time ROMs. The following subsections outline these two trial subspaces, along with the Galerkin, LSPG, and 
-space--time LSPG method.
-
-\subsection{\spatialAcronym\ Trial Subspaces}
+where $t \in [0,T]$ denotes time with 
+$T \in \mathbb{R}^+$ denoting the final time, $\stateFOM \in \RR{\fomdim}
+	\otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$
+	denotes the state implicitly defined as the solution to initial value
+	problem~\eqref{eq:FOM}, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial conditions,
+	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
+	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,t) \mapsto
+	\velocity(\stateyDiscrete,t)$ denotes the velocity, which is possibly
+	nonlinear in its first argument. We refer to Eq.~\eqref{eq:FOM} 
+	as the ``full-order model" (FOM) ODE.
+
+Directly solving initial value problem~\eqref{eq:FOM} is computationally
+	expensive if either the state-space dimension $\fomdim$ is large, or if the
+	time-interval length $T$ is large relative to the time step required to
+	numerically integrate Eq.~\eqref{eq:FOM}. For time-critical or many-query
+	applications, it is essential to employ an approximation in lieu of the FOM
+	ODE~\eqref{eq:FOM}. Projection-based ROMs constitute one such promising
+	approach.
+
+\subsection{Projection-based reduced-order models}
+Projection-based ROMs generate approximate solutions to the FOM
+	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
+	subspace. For dynamical 
+	systems, two distinct types of space--time trial subspaces are commonly used: 
+\begin{enumerate} 
+	\item \textit{\spatialAcronym\ trial subspaces}}. These 
+	trial subspaces reduce only the spatial dimension of the FOM, and are commonly used
+		in model-reduction methods based on spatial projection followed by time
+		integration (e.g., Galerkin projection, LSPG projection)
+	\item \textit{\spaceTimeAcronym\ trial subspaces}. These 
+	trial subspaces reduce both the spatial and temporal dimensions of the FOM,
+		and are commonly employed 
+		employed in model-reduction methods based on space--time projection (e.g.,
+		space--time LSPG projection).
+\end{enumerate}
+ We now describe these two types of trial subspaces and their application to 
+	Galerkin, LSPG, and space--time LSPG projection. 
+	\KTC{HERE}
+
+\subsection{\spatialAcronym\ trial subspaces}
 \spatialAcronym\ trial subspaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial subspace, 
@@
 
-\subsubsection{The Galerkin Method}
+\subsubsection{Galerkin projection}
 The Galerkin method is (traditionally) obtained by the following steps: (1.) The approximation~\eqref{eq:affine_trialspace} is substituted into~\eqref{eq:FOM}, and (2.) the resulting residual is enforced to be $\stweightingMatArg{}$-orthogonal to the trial subspace, with $\stweightingMatArg{ } \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ }$ being a positive (semi-)definite weighting matrix (commonly taken to be $\mathbf{I}$).
@@
 %\end{equation}
-\subsubsection{Least-Squares Petrov--Galerkin}
+\subsubsection{Least-Squares Petrov--Galerkin projection}
 The Galerkin method can be unstable and inaccurate for non-self-adjoint and nonlinear problems. 
@@
 
-\subsection{\spaceTimeAcronym\ Trial Spaces and Space--Time ROMs}
+\subsection{\spaceTimeAcronym\ trial spaces and space--time ROMs}
 Recent work has investigated space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}. Space--time methods 
@@
 principle, however, we do not outline it here.
-\subsubsection{Space--Time LSPG} 
+\subsubsection{Space--Time LSPG projection}  
 As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain.
@@
 
-\subsection{Outstanding Challenges}
+\subsection{Outstanding challenges}
 This work seeks to overcome the limitations of existing reduced-order modeling approaches. Specifically, 

### 2019-10-02 section4/numerical_methods.tex
p 6

@@
-\section{Numerical Solution Techniques}\label{sec:numerical_techniques}
+\section{Numerical solution techniques}\label{sec:numerical_techniques}
 The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
@@
 
-\subsection{\spatialAcronym: Direct and Indirect Methods}
+\subsection{\spatialAcronym: direct and indirect methods}
 Numerical techniques to

### 2019-10-02 section4/spacetime_numerical.tex
p 6

@@
-\subsection{\spaceTimeAcronym\ Trial Spaces: Direct and Indirect Methods}
+\subsection{\spaceTimeAcronym\ trial spaces: direct and indirect methods}
 We now briefly consider \spaceTimeAcronym\ trial subspaces. 
@@
 
-\subsection{\spaceTimeAcronym\ Trial Spaces: Direct Solution Technique}
+\subsection{\spaceTimeAcronym\ trial spaces: direct solution technique}
 The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods.
@@
 
-\subsection{\spaceTimeAcronym\ Trial Spaces: Indirect Solution Technique}
+\subsection{\spaceTimeAcronym\ trial spaces: indirect solution technique}
 As opposed to the direct solution technique, the indirect solution technique directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime},
@@
 
-\subsection{\spaceTimeAcronym\ Trial Space: Summary}
+\subsection{\spaceTimeAcronym\ trial space: summary}
 As was just shown, \spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window.

### 2019-10-02 main.tex
latest

--- 2019-10-02 main.tex
@@
 
-\KTC{Why are we reviewing LSFEM? This isn't a ROM. Don't think it's relevant}
-Least-squares residual minimization as a numerical solution technique for
-partial differential equations was pioneered in the finite-element community
-and led to what is known as the least-squares finite-element method
-(LSFEM)~\cite{ls_review,bochev_leastsquares}. 
-\KTC{Specifically, LSFEM was designed to solve the inf--sup stability problem.
-Be more precise.}
-Motivated by the fact that the
-Galerkin method can be inaccurate and unstable for nonlinear and
-non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic
-functional comprising the residual norm of the partial differential equation
-of interest and (2.) leveraging variational principles to derive the
-stationary conditions of the quadratic functional; i.e, LSFEM can be thought
-of as finding the solution within a finite element trial space the minimizes
-the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous
-advantages (e.g., least-squares system are symmetric and positive definite)
-\KTC{This is not an `advantage'. It is guaranteed to be inf--sup stable.}
-over the more standard Galerkin discretization.
-
-Researchers also developed residual-minimization methods for model reduction.
-Similar to LSFEM, residual minimization in this context associates with
-computing the solution within a low-dimensional trial space that minimizes the
-$\ell^2$-norm of the full-order-model residual. \KTC{haven't defined
-`full-order model'} Researchers have developed such residual-minimizing
+%\KTC{Why are we reviewing LSFEM? This isn't a ROM. Don't think it's relevant}
+%Least-squares residual minimization as a numerical solution technique for
+%partial differential equations was pioneered in the finite-element community
+%and led to what is known as the least-squares finite-element method
+%(LSFEM)~\cite{ls_review,bochev_leastsquares}. 
+%\KTC{Specifically, LSFEM was designed to solve the inf--sup stability problem.
+%Be more precise.}
+%Motivated by the fact that the
+%Galerkin method can be inaccurate and unstable for nonlinear and
+%non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic
+%functional comprising the residual norm of the partial differential equation
+%%of interest and (2.) leveraging variational principles to derive the
+%stationary conditions of the quadratic functional; i.e, LSFEM can be thought
+%of as finding the solution within a finite element trial space the minimizes
+%the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous
+%advantages (e.g., least-squares system are symmetric and positive definite)
+%\KTC{This is not an `advantage'. It is guaranteed to be inf--sup stable.}
+%over the more standard Galerkin discretization.
+
+Residual minimization methods in model-order reduction comprise 
+computing a solution within a low-dimensional trial space that minimizes some residual norm of the system of interest. Researchers have developed such residual-minimizing
 model-reduction methods for both static systems (i.e., systems without
@@
 than time instantanously), LSPG projection computes solutions that are more
-accurate over a larger part the trajectory compared with Galerkin projection.
+accurate over a larger part of the trajectory as compared to Galerkin projection.
 
@@
 
-The LSPG approach sequentially performs residual minimization over a single
-time-step, and as such is not associated with residual minimization over the
-entire time interval. \KTC{This is what ST-LSPG does. Why do we ignore it? the
-full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
-guarantees of stability or convergence, even for linear time-invariant
-systems~\cite{bui_thesis}. Further, the performance of LSPG depends on the
-time step in two distinct ways: changing the time-step size modifies both the
+The LSPG approach has several shortcomings. First, the performance of LSPG depends on the
+time discretization. Changing the time-step size modifies both the
 time window over which LSPG minimizes the residual \textit{as well as} the
@@
 of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
-is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
-as the time step increases, the time discretization error of the full-order
-model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
-additionally sensitive to the time discretization scheme. When used with
-explicit temporal discretization, for instance, LSPG again reverts to
-Galerkin. To gain the enhanced stability associated with LSPG, the underlying
-time-discretization scheme must be implicit with an appropriate time-step.
-This can increase the cost of the reduced-order model, especially in the case
-where the optimal time-step required by LSPG is small.
+is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.}. 
+%The accuracy of LSPG additionally degrades if the time-step size becomes too large.
+It is unclear if this is due to the discretization error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
+LSPG projection additionally is
+%additionally sensitive to the time discretization scheme. When used with
+%explicit temporal discretization, for instance, LSPG again reverts to
+%Galerkin. To gain the enhanced stability associated with LSPG, the underlying
+%time-discretization scheme must be implicit with an appropriate time-step.
+%%This can increase the cost of the reduced-order model, especially in the case
+%where the optimal time-step required by LSPG is small. 
+Second, as the LSPG approach sequentially performs residual minimization over
+time-steps, is not equipped with any notion of optimality over the entire 
+time domain of interest. As a result, LSPG is not equipped with \textit{a priori}
+guarantees of stability or convergence, even for linear time-invariant
+systems~\cite{bui_thesis}.
+
+Researchers have pursued the development of space--time residual minimization ROMs~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis} to address this last issue.
+In existing space--time methods, the spatial \textit{and} temporal dependence of the state is represented in a low-dimensional space--time trial space.
+Space--time residual minimization approaches then comprise computing a space--time solution within this low-dimensional space--time trial space that minimizes the space--time norm of the full-order model.
+Space--time methods are thus equipped with a notion of optimality over the entire time domain of interest and, as a result, contain more favorable error bounds than the Galerkin and LSPG approaches.
+Several drawbacks of existing space--time ROMs are that they (1) require minimizing the entire space--time residual (which can be a computationally intensive process) and (2) require the specification of a finite-dimensional temporal basis (computed, e.g., via POD applied to space--time snapshots).
+This latter issue is extremely relevant to the present work.
+%These space--time In these space--time ROMs, the temporal dependence of the state 
+%is represented via a finite dimensional basis representation. 
+%Space--time residual minimization methods can be prohibitively expensive, however, as they involve minimizing the entire space--time residual.
+Further, existing space--time approaches require a finite-dimensional space--time basis.
+
+This work seeks to overcome the practical challenges associated with existing
+residual minimization approaches for reduced-order models. To this end, we
+propose the \methodName\ method for model reduction, which minimizes the FOM ODE residual
+over a sequence of time windows. The key novelties of the work are two fold.
+First, the residual minimization problem is formulated to minimize the
+\textit{time-continuous} residual. By working from a time-continuous viewpoint, the
+present formulation decouples the underlying temporal discretization scheme
+from the residual minimization problem; thus addressing a key deficiency of
+LSPG \KTC{This is also true for ST-LSPG and maybe other discretization-based
+methods? Have to check}. The time-continuous viewpoint additionally exposes
+two different methods that
+can be used to solve the residual-minimization problems, namely
+a \textit{discretize-then-optimize} (i.e., direct) method and
+an \textit{optimize-then-discretize} (i.e., indirect) method. The second novelty
+is that the residual is minimized over \textit{time windows} rather than over
+a single time step or the entire time domain. This equips the method with
+additionaly flexibility that enables it
+to explore more fine-grained tradeoffs in computational cost and error. We
+demonstrate that this method can be viewed as a generalization of existing
+methods, as particular instances of this formulation recover Galerkin, LSPG,
+and space--time LSPG projection. Figure~\ref{fig:flowchart} depicts how the
+proposed method is related to existing approaches.
+
+% \KTC{This is what ST-LSPG does. Why do we ignore it? the
+%full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
+%guarantees of stability or convergence, even for linear time-invariant
+%systems~\cite{bui_thesis}. Further, the performance of LSPG depends on the
+%time step in two distinct ways: changing the time-step size modifies both the
+%time window over which LSPG minimizes the residual \textit{as well as} the
+%time-discretization error of the full-order model on which LSPG is based. As a
+%result, LSPG often yields the smallest error for an intermediate value of the
+%time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}). In
+%particular, as the time step decreases, the accuracy of LSPG approaches that
+%of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
+%is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
+%as the time step increases, the time discretization error of the full-order
+%model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
+%additionally sensitive to the time discretization scheme. When used with
+%explicit temporal discretization, for instance, LSPG again reverts to
+%Galerkin. To gain the enhanced stability associated with LSPG, the underlying
+%time-discretization scheme must be implicit with an appropriate time-step.
+%This can increase the cost of the reduced-order model, especially in the case
+%where the optimal time-step required by LSPG is small.
 
@@
 It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work.
-
-This work seeks to overcome the practical challenges associated with existing
-residual minimization approaches for reduced-order models. To this end, we
-propose the \methodNameLower\ (\methodAcronym) method for model reduction,
-which minimizes the FOM ODE residual over a sequence of arbitrarily defined
-time windows. The key novelties of the work are twofold.  First, the method
-minimizes the \textit{time-continuous}  residual (i.e., that associated with
-the full-order model ODE). By adopting a time-continuous viewpoint, the
-formulation decouples the underlying temporal discretization scheme from the
-residual minimization problem, thus addressing a key deficiency of LSPG
-\KTC{This is also true for ST-LSPG and maybe other discretization-based
-methods? Have to check}. Time-continuous residual minimization also exposes
-two different solution methods,  namely a \textit{discretize-then-optimize}
-(i.e., direct) method and an \textit{optimize-then-discretize} (i.e.,
-indirect) method. The second novelty is that the method sequentially minimizes
-the residual over arbitrarily defined \textit{time windows} rather than
-sequentially minimizing the residual over time steps or over the entire time
-domain. This equips the method with additionaly flexibility that enables it to
-explore more fine-grained tradeoffs between computational cost and error. We
-demonstrate that this method can be viewed as a generalization of existing
-methods, because particular instances of this formulation are equivalent to
-Galerkin, LSPG, and space--time LSPG projection. Figure~\ref{fig:flowchart}
-depicts how the proposed \methodAcronym\ method is related to existing
-approaches.
 

### 2019-10-03 main.tex
into 2.2.2

--- 2019-10-02 main.tex
@@
 \begin{equation}\label{eq:FOM}
- \frac{d\stateFOM}{dt} = \velocity(\stateFOM,t), \qquad \stateFOM(0) = \stateFOMIC,
+ \frac{d\stateFOM}{dt} = \velocity(\stateFOM,t), \qquad \stateFOM(0) =
+	\stateFOMIC,\qquad t \in [0,T]
 \end{equation}
-where $t \in [0,T]$ denotes time with 
-$T \in \mathbb{R}^+$ denoting the final time, $\stateFOM \in \RR{\fomdim}
+where 
+$T \in \mathbb{R}^+$ denotes the final time, $\stateFOM \in \RR{\fomdim}
 	\otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$
 	denotes the state implicitly defined as the solution to initial value
-	problem~\eqref{eq:FOM}, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial conditions,
+	problem~\eqref{eq:FOM}, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
 	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
@@
 \begin{enumerate} 
-	\item \textit{\spatialAcronym\ trial subspaces}}. These 
+	\item \textit{\spatialAcronym\ trial subspaces}. These 
 	trial subspaces reduce only the spatial dimension of the FOM, and are commonly used
@@
 \subsection{\spatialAcronym\ trial subspaces}
-\spatialAcronym\ trial subspaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial subspace, 
+\spatialAcronym\ trial subspaces approximate the FOM ODE solution
+	$\stateFOM$ with an approximation $\approxstate$ that resides in an
+	affine space--time trial subspace of dimension $K\ll\fomdim$, i.e., 
 \begin{align}\label{eq:spatial_subspace}
-& \stspaceS \defeq \trialspace \otimes \timeSpace,\\ 
-& \trialspace \defeq \text{Range}(\basismat) +  \stateIntercept,  \label{eq:spatial_subspace2}  \\
+& \stspaceS \defeq \trialspace \otimes \timeSpace +
+	\stateIntercept\otimes\onesFunction,\\ 
+& \trialspace \defeq \text{Range}(\basismat),  \label{eq:spatial_subspace2}  \\
 &\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, \label{eq:spatial_subspace3} 
 \end{align}
-where we refer to $\basismat \in \RRStar{\romdim}{\fomdim}$, with $\RRStar{\romdim}{\fomdim}$ denoting the set  $\{ \mathbf{X} \in \RR{\fomdim \times \romdim} \big| \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$, as the spatial trial basis matrix, $\trialspace$ as the spatial trial subspace, and $\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state.
+where 
+	$\timeSpace$ denotes the set of real-valued functions acting on the time
+	domain
+	$[0,T]$, $\onesFunction\in\timeSpace$ is defined as $\onesFunction:[0,T]\rightarrow\{1\}$,
+	$\basismat \in \RRStar{\romdim}{\fomdim}$ denotes the spatial trial subspace with
+	$\RRStar{\romdim}{\fomdim}$ denoting the compact Stiefel manifold (i.e.,  $
+	\RRStar{\romdim}{\fomdim}\defeq
+	\{ \mathbf{X} \in \RR{\fomdim
+	\times \romdim}\, \big|\, \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$),
+	$\trialspace\subseteq\RR{\fomdim}$ denotes the spatial trial subspace, and
+	$\stateIntercept \in \mathbb{R}^{\fomdim}$ denotes the reference state that
+	is often taken to be the initial condition (i.e., $\stateIntercept = \stateFOMIC$). 
 %where $\basisvec_i \in \mathbb{R}^N$ are orthonormal basis vectors and $\stateIntercept \in \mathbb{R}^{\fomdim}$ the ``intercept''.
 %We refer to $\basismat \in \RRStar{\fomdim \times \romdim}$, with $\RRStar{p \times q}$ denoting the set of orthogonal $q$ frames in $p$, as the spatial trial basis, and $\trialspace$ as the spatial trial space.
-The basis vectors $\basisvec_i$ are typically constructed through, e.g., POD~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}, etc. 
-ROMs leveraging \spatialAcronym\ trial subspaces approximate the state as,
+The basis vectors $\basisvec_i$, $i=1,\ldots,K$ are typically constructed
+using state snapshots, e.g., via
+proper orthogonal decomposition (POD)~\cite{berkooz_turbulence_pod} or the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}. 
+At any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\ trial subspace approximate the state as
 \begin{equation}\label{eq:affine_trialspace}
-\stateFOMArg{}{t} \approx \approxstate(t) \defeq \basisspace \genstateArg{}{t} + \stateIntercept,
+\stateFOMArg{}{t} \approx \approxstate(t) = \basisspace \genstateArg{}{t} + \stateIntercept,
 \end{equation}
-where $\genstate \in \RR{\romdim} \otimes \timeSpace$, with $\genstate : [0,T] \rightarrow \mathbb{R}^{\romdim}$, are the generalized coordinates.
-Substituting the approximation given in~\eqref{eq:affine_trialspace} into~\eqref{eq:FOM} and performing orthogonal $\elltwo$ projection of the initial conditions onto the trial subspace, one obtains,
+where $\genstate \in \RR{\romdim} \otimes \timeSpace$ denotes the generalized
+coordinates. Substituting the approximation~\eqref{eq:affine_trialspace} in
+Eq.~\eqref{eq:FOM} and performing orthogonal
+$\elltwo$-projection of the initial condition onto the trial subspace yields
+the overdetermined system of ODEs
 \begin{equation}\label{eq:g_truncation}
-\basisspace \frac{d }{dt}\genstateArg{}{t} = \velocity(\basisspace
-\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
+\basisspace \frac{d\genstate }{dt} = \velocity(\basisspace
+\genstate + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
 \end{equation}
-Equation~\eqref{eq:g_truncation} provides an over-determined system of $N$ equations for the $K$ generalized coordinates.
-Dimension reduction is predominantly performed with either the \textit{Galerkin method} or the \textit{Least-Squares Petrov--Galerkin method} (LSPG).
-We now outline these two methods.
+Because Eq.~\eqref{eq:g_truncation} is overdetermined, a solution may not
+exist. Typically, either \textit{Galerkin projection} or \textit{least-squares
+Petrov--Galerkin projection} is employed to reduce the number of equations
+such that a unique solution exists. We now describe these two methods.
 
 \subsubsection{Galerkin projection}
-The Galerkin method is (traditionally) obtained by the following steps: (1.) The approximation~\eqref{eq:affine_trialspace} is substituted into~\eqref{eq:FOM}, and (2.) the resulting residual is enforced to be $\stweightingMatArg{}$-orthogonal to the trial subspace, with $\stweightingMatArg{ } \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ }$ being a positive (semi-)definite weighting matrix (commonly taken to be $\mathbf{I}$).
-This projection process results in what is known as the Galerkin method,
+Galerkin projection reduces the number of equations in
+Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the associated
+residual to the trial subspace in the (semi-)inner product induced by the 
+positive (semi-)definite matrix $\stweightingMatArg{}\equiv \stweightingMatOneArg{ }^T
+\stweightingMatOneArg{ }$ (commonly set to
+$\stweightingMatArg{}=\mathbf{I}$), i.e.,
 \begin{equation*}\label{eq:g_truncation2}
-\frac{d}{dt}\genstateGalerkinArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
-\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
+\frac{d\genstateGalerkin}{dt} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
+\genstateGalerkin + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
 \end{equation*}
-where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$ is a mass matrix. 
-Model reduction via the Galerkin method is seen to result in a $K$-dimensional dynamical system for the 
-generalized coordinates. 
-The Galerkin method can alternatively be viewed as a residual minimization method~\cite{carlberg_lspg_v_galerkin}. In this setting, the Galerkin ROM can be 
-written as,
-$$
-\frac{d}{dt}\genstateGalerkinArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}  }{\text{arg\,min}} || \basisspace \hat{\stateyDiscrete} -  \velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t) ||_{\stweightingMatArg{n}}^2.
+where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
+denotes a mass matrix.  Note that the Galerkin ODE~\eqref{eq:g_truncation2} is
+a system of $K$ equations in $K$ unknowns.
+As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, Galerkin projection
+can be viewed alternatively as a residual-minimization
+method. In particular, the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
+to
+\begin{equation}\label{eq:GalOptimal}
+\frac{d\genstateGalerkin}{dt} = \underset{ \genstateyDiscrete \in \RR{\romdim}
+	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
+	\velocity(\basisspace \genstateGalerkin + \stateIntercept,t)
+	\|_{\stweightingMatArg{n}}^2,
 %\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
-$$
+\end{equation}
+where $\|\mathbf{x}\|_{\stweightingMatArg{n}}\equiv
+\sqrt{\mathbf{x}^T\stweightingMatArg{n}\mathbf{x}}$.
 %where
@@
 %\end{equation*}
-The Galerkin ROM corresponds to \textit{time-instantaneous} minimization of the FOM ODE residual.
+Thus, velocity $\frac{d\genstateGalerkin}{dt}(t)$ computed by Galerkin projection minimizes the
+time-instantaneous FOM ODE residual evaluated at the state $\basisspace
+\genstateGalerkin(t) + \stateIntercept$ and time $t$.
 %To perform dimension reduction, 
@@
 \subsubsection{Least-Squares Petrov--Galerkin projection}
-The Galerkin method can be unstable and inaccurate for non-self-adjoint and nonlinear problems. 
-An alternative model-reduction approach that enjoys several advantageous over the Galerkin method is the Least-Squares Petrov--Galerkin (LSPG) method~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}.
-LSPG is defined at the fully discrete level and minimizes the fully discrete FOM O$\Delta$E residual (i.e., the FOM ODE residual after temporal discretization) at each time-step instance.
-We outline the LSPG method for linear multistep methods as a demonstrative example.
-Defining a discretization of the continuous time domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time step $\Delta t = T/N_t$, temporal discretization via a linear multistep method yields the FOM O$\Delta$E,
+Despite its time-instantaneous optimality property expressed in
+Eq.~\eqref{eq:GalOptimal}, Galerkin projection can yield unstable and
+inaccurate trajectories, particularly if the velocity is not self-adjoint or
+is nonlinear. 
+Least-squares Petrov--Galerkin (LSPG)
+projection~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
+is an alternative approach that exhibits several advantages over Galerkin
+projection.
+Rather than minimize the (time-continuous) FOM ODE residual as in Galerkin
+projection, LSPG minimizes
+the residual arising at a given time step after applying time discretization to the FOM ODE; we
+refer to the residual arising at the fully discrete level as the FOM O$\Delta$E residual.  
+We now describe LSPG projection when a linear multistep method is employed for
+time discretization; Ref.~\cite{carlberg_lspg_v_galerkin} derives LSPG projection for Runge--Kutta
+schemes as well. 
+\KTC{HERE}
+Defining a discretization
+of the continuous time domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta
+t, n=1,2,\ldots,N_t$ with a constant time step $\Delta t = T/N_t$, temporal
+discretization via a linear multistep method yields the FOM O$\Delta$E,
 \begin{equation}\label{eq:lms}

### 2019-10-03 main.tex
progress

@@
 Galerkin projection reduces the number of equations in
-Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the associated
-residual to the trial subspace in the (semi-)inner product induced by the 
-positive (semi-)definite matrix $\stweightingMatArg{}\equiv \stweightingMatOneArg{ }^T
+Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
+residual to the spatial trial subspace in the (semi-)inner product induced by the 
+positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneArg{ }^T
 \stweightingMatOneArg{ }$ (commonly set to
 $\stweightingMatArg{}=\mathbf{I}$), i.e.,
-\begin{equation*}\label{eq:g_truncation2}
+\begin{equation}\label{eq:g_truncation2}
 \frac{d\genstateGalerkin}{dt} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
 \genstateGalerkin + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
-\end{equation*}
+\end{equation}
 where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
-denotes a mass matrix.  Note that the Galerkin ODE~\eqref{eq:g_truncation2} is
-a system of $K$ equations in $K$ unknowns.
+denotes the $K\times K$ positive definite mass matrix.
 As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, Galerkin projection
 can be viewed alternatively as a residual-minimization
-method. In particular, the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
+method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
 to
@@
 %\end{equation*}
-Thus, velocity $\frac{d\genstateGalerkin}{dt}(t)$ computed by Galerkin projection minimizes the
+Thus, the computed velocity $\frac{d\genstateGalerkin}{dt}(t)$ minimizes the
 time-instantaneous FOM ODE residual evaluated at the state $\basisspace
-\genstateGalerkin(t) + \stateIntercept$ and time $t$.
+\genstateGalerkin(t) + \stateIntercept$ and time $t$ over the spatial trial
+subspace $\trialspace$.
 %To perform dimension reduction, 
@@
 \subsubsection{Least-Squares Petrov--Galerkin projection}
-Despite its time-instantaneous optimality property expressed in
-Eq.~\eqref{eq:GalOptimal}, Galerkin projection can yield unstable and
-inaccurate trajectories, particularly if the velocity is not self-adjoint or
+Despite its time-instantaneous optimality property 
+\eqref{eq:GalOptimal}, Galerkin projection can yield inaccurate solutions,
+particularly if the velocity is not self-adjoint or
 is nonlinear. 
@@
 projection~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
-is an alternative approach that exhibits several advantages over Galerkin
+is an alternative projection method that exhibits several advantages over Galerkin
 projection.
 Rather than minimize the (time-continuous) FOM ODE residual as in Galerkin
-projection, LSPG minimizes
-the residual arising at a given time step after applying time discretization to the FOM ODE; we
-refer to the residual arising at the fully discrete level as the FOM O$\Delta$E residual.  
-We now describe LSPG projection when a linear multistep method is employed for
-time discretization; Ref.~\cite{carlberg_lspg_v_galerkin} derives LSPG projection for Runge--Kutta
-schemes as well. 
-\KTC{HERE}
-Defining a discretization
-of the continuous time domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta
-t, n=1,2,\ldots,N_t$ with a constant time step $\Delta t = T/N_t$, temporal
-discretization via a linear multistep method yields the FOM O$\Delta$E,
+projection, at each time step, LSPG minimizes
+the FOM O$\Delta$E residual, i.e., the residual arising after applying time discretization to the FOM ODE.  
+We now describe LSPG projection in the case of linear multistep methods;
+Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG projection for Runge--Kutta
+schemes.
+
+To begin, and without loss of generality, we introduce a uniform time
+grid characterized by the time step $\Delta t$ and time instances
+$t^n = n\Delta
+t$, $n=1,\ldots,N_t$.
+Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
+with this time grid
+yields the FOM O$\Delta$E, which computes the sequence of FOM O$\Delta$E
+solutions
+$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
+as the implicit solution to the system of algebraic equations
 \begin{equation}\label{eq:lms}
-\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, \qquad n=1,\ldots,N_t, 
+\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, 
 \end{equation}
-where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance and 
-$\residLMSArg{n}$ is the discrete linear multistep residual at the $n$th time step,
+where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
+time instance and 
+$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual, which is defined as
 \begin{align*}
@@
 \\
-&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim}. 
-\end{align*} 
-In the above, $k^n$ is the width of the multistep scheme at the $n$th time step and $\alpha^n_j,\beta^n_j$ are coefficients that define the type of multistep scheme at the $n$th time step.
+&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim},
+\end{align*} 
+Here, $\alpha^n_j$, $\beta^n_j$, $j=0,\ldots,k^n$ are coefficients
+that define the linear multistep method at the $n$th time instance.
 %\begin{equation}\label{eq:coarse_implicit_euler_0}
@@
 %where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.
-The FOM O$\Delta$E solution will exactly satisfy~\eqref{eq:lms}; the ROM solution will not.
-The LSPG method minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial subspace over each time step.
-%We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
+
+\KTC{HERE}
+The LSPG method substitutes the \spatialAcronym\ trial subspace approximation
+minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial subspace over each time step. %We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
 %\begin{align*}

### 2019-10-03 numerical_experiments.tex

@@
-\section{Numerical Experiments}\label{sec:numerical_experiments}
+\section{Numerical experiments}\label{sec:numerical_experiments}
 This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity.
@@
 
-\subsection{Sod Shock Tube}
+\subsection{Sod shock tube}
 We first consider reduced-order models of the Sod shock tube problem. The system is governed by the compressible Euler equations in one dimension, 
@@
 
-\subsubsection{Description of FOM and Generation of Trial Space}
+\subsubsection{Description of FOM and generation of trial space}
 The 1D compressible Euler equations are solved using a finite volume method.
@@
 
-\subsubsection{Description of Reduced-Order Models}
+\subsubsection{Description of reduced-order models}
 Reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approaches are considered. No hyper-reduction is considered 
@@
 
-\subsubsection{Numerical Results}
+\subsubsection{Numerical results}
 The impact of the window size on \methodAcronymROMs\ is first assessed. A set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
@@
 
-\subsubsection{Summary of Numerical Results for the Sod Shock Tube}
+\subsubsection{Summary of numerical results for the Sod shock tube}
 The key observations from the results of the first numerical example are: 

### 2019-10-03 section3/general_form.tex

@@
-\section{Windowed Least-Squares Formulation}\label{sec:tclspg} This section
+\section{Windowed residual minimization formulation}\label{sec:tclspg} This section
 outlines the proposed \methodNameLower\ 
@@
 %
-\subsection{Residual Minimization Formulation} 
+\subsection{Residual minimization formulation} 
 We start by partitioning the

### 2019-10-03 section3/space_time.tex

@@
-\subsection{\spaceTimeAcronym\ Trial Spaces}
+\subsection{\spaceTimeAcronym\ trial spaces}
 We now briefly consider \spaceTimeAcronym\ trial subspaces. Over 
@@
 % & n=1. \end{cases} \end{equation}
-\subsubsection{Stationary Conditions} 
+\subsubsection{Stationary conditions} 
 The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$.

### 2019-10-03 section4/spatial_numerical.tex

@@
-\subsection{\spatialAcronym: Direct Solution Approach}\label{sec:direct} 
+\subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
 Direct approaches solve~\eqref{eq:tclsrm} by
@@
 %\end{align*}
-\subsubsection{Linear Multistep Schemes}
+\subsubsection{Linear multistep schemes}
 Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time-instances.  
@@
 \begin{comment}
-\subsubsection{Collocation (Runge--Kutta) Methods}
+\subsubsection{Collocation (Runge--Kutta) methods}
 Collocation methods comprise another popular technique for temporal discretization. 
@@
 \end{comment}
-\subsubsection{Solution to the Nonlinear Least-Squares Problem through the Gauss--Newton Method}
+\subsubsection{Solution to the nonlinear least-squares problem through the Gauss--Newton method}
 Discretization through linear multistep methods (as well as other techniques) 
@@
 
-\subsection{Spatial Trial Space: Indirect Solution Approach}\label{sec:indirect}
+\subsection{Spatial trial space: indirect solution approach}\label{sec:indirect}
 In contrast to the direct approach,
@@
 \end{algorithm}
-\subsubsection{Numerical Considerations for the Forward and Backwards Problems}
+\subsubsection{Numerical considerations for the forward and backward problems}
 The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 

### 2019-10-03 main.tex
working on intro

@@
 of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
-is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
-as the time step increases, the time discretization error of the full-order
-model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
-additionally sensitive to the time discretization scheme. When used with
-explicit temporal discretization, for instance, LSPG again reverts to
-Galerkin. To gain the enhanced stability associated with LSPG, the underlying
-time-discretization scheme must be implicit with an appropriate time-step.
-This can increase the cost of the reduced-order model, especially in the case
-where the optimal time-step required by LSPG is small.
+is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.}. 
+%The accuracy of LSPG additionally degrades if the time-step size becomes too large.
+It is unclear if this is due to the discretization error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
+LSPG projection additionally is
+%additionally sensitive to the time discretization scheme. When used with
+%explicit temporal discretization, for instance, LSPG again reverts to
+%Galerkin. To gain the enhanced stability associated with LSPG, the underlying
+%time-discretization scheme must be implicit with an appropriate time-step.
+%%This can increase the cost of the reduced-order model, especially in the case
+%where the optimal time-step required by LSPG is small. 
+Second, as the LSPG approach sequentially performs residual minimization over
+time-steps, is not equipped with any notion of optimality over the entire 
+time domain of interest. As a result, LSPG is not equipped with \textit{a priori}
+guarantees of stability or convergence, even for linear time-invariant
+systems~\cite{bui_thesis}.
+
+Researchers have pursued the development of space--time residual minimization ROMs~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis} to address this last issue.
+In existing space--time methods, the spatial \textit{and} temporal dependence of the state is represented in a low-dimensional space--time trial space.
+Space--time residual minimization approaches then comprise computing a space--time solution within this low-dimensional space--time trial space that minimizes the space--time norm of the full-order model.
+Space--time methods are thus equipped with a notion of optimality over the entire time domain of interest and, as a result, contain more favorable error bounds than the Galerkin and LSPG approaches.
+Several drawbacks of existing space--time ROMs are that they (1) require minimizing the entire space--time residual (which can be a computationally intensive process) and (2) require the specification of a finite-dimensional temporal basis (computed, e.g., via POD applied to space--time snapshots).
+This latter issue is extremely relevant to the present work.
+%These space--time In these space--time ROMs, the temporal dependence of the state 
+%is represented via a finite dimensional basis representation. 
+%Space--time residual minimization methods can be prohibitively expensive, however, as they involve minimizing the entire space--time residual.
+Further, existing space--time approaches require a finite-dimensional space--time basis.
+
+%This work seeks to overcome the practical challenges associated with existing
+%residual minimization approaches for reduced-order models. To this end, we
+%propose the \methodName\ method for model reduction, which minimizes the FOM ODE residual
+%over a sequence of time windows. The key novelties of the work are two fold.
+%First, the residual minimization problem is formulated to minimize the
+%\textit{time-continuous} residual. By working from a time-continuous viewpoint, the
+%present formulation decouples the underlying temporal discretization scheme
+%from the residual minimization problem; thus addressing a key deficiency of
+%LSPG \KTC{This is also true for ST-LSPG and maybe other discretization-based
+%methods? Have to check}. The time-continuous viewpoint additionally exposes
+%two different methods that
+%can be used to solve the residual-minimization problems, namely
+%a \textit{discretize-then-optimize} (i.e., direct) method and
+%an \textit{optimize-then-discretize} (i.e., indirect) method. The second novelty
+%is that the residual is minimized over \textit{time windows} rather than over
+%a single time step or the entire time domain. This equips the method with
+%additionaly flexibility that enables it
+%to explore more fine-grained tradeoffs in computational cost and error. We
+%demonstrate that this method can be viewed as a generalization of existing
+%methods, as particular instances of this formulation recover Galerkin, LSPG,
+%and space--time LSPG projection. Figure~\ref{fig:flowchart} depicts how the
+%proposed method is related to existing approaches.
+
+% \KTC{This is what ST-LSPG does. Why do we ignore it? the
+%full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
+%guarantees of stability or convergence, even for linear time-invariant
+%systems~\cite{bui_thesis}. Further, the performance of LSPG depends on the
+%time step in two distinct ways: changing the time-step size modifies both the
+%time window over which LSPG minimizes the residual \textit{as well as} the
+%time-discretization error of the full-order model on which LSPG is based. As a
+%result, LSPG often yields the smallest error for an intermediate value of the
+%time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}). In
+%particular, as the time step decreases, the accuracy of LSPG approaches that
+%of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
+%is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
+%as the time step increases, the time discretization error of the full-order
+%model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
+%additionally sensitive to the time discretization scheme. When used with
+%explicit temporal discretization, for instance, LSPG again reverts to
+%Galerkin. To gain the enhanced stability associated with LSPG, the underlying
+%time-discretization scheme must be implicit with an appropriate time-step.
+%This can increase the cost of the reduced-order model, especially in the case
+%where the optimal time-step required by LSPG is small.
 
@@
 \begin{equation}\label{eq:FOM}
- \frac{d\stateFOM}{dt} = \velocity(\stateFOM,t), \qquad \stateFOM(0) =
+ \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
 	\stateFOMIC,\qquad t \in [0,T]
 \end{equation}
-where 
+where $\stateFOMDot \equiv \frac{d \stateFOM}{dt}$, 
 $T \in \mathbb{R}^+$ denotes the final time, $\stateFOM \in \RR{\fomdim}
@@
 	denotes the state implicitly defined as the solution to initial value
-	problem~\eqref{eq:FOM}, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
+	problem~\eqref{eq:FOM}, 
+	$\timeSpace$ denotes the set of real-valued functions acting on the time
+	domain, 
+ $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
 	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
@@
 \begin{enumerate} 
-	\item \textit{\spatialAcronym\ trial subspaces}. These 
+	\item \textit{Space–time trial subspaces with spatial dimension reduction (\spatialAcronym)}. These 
 	trial subspaces reduce only the spatial dimension of the FOM, and are commonly used
 		in model-reduction methods based on spatial projection followed by time
-		integration (e.g., Galerkin projection, LSPG projection)
-	\item \textit{\spaceTimeAcronym\ trial subspaces}. These 
+		integration (e.g., Galerkin projection, LSPG projection.)
+	\item \textit{Space–time trial subspaces with spatial and temporal dimension reduction (\spaceTimeAcronym)}. These 
 	trial subspaces reduce both the spatial and temporal dimensions of the FOM,
@@
 	affine space--time trial subspace of dimension $K\ll\fomdim$, i.e., 
-\begin{align}\label{eq:spatial_subspace}
+\begin{equation}\label{eq:spatial_subspace}
+\begin{split}
 & \stspaceS \defeq \trialspace \otimes \timeSpace +
 	\stateIntercept\otimes\onesFunction,\\ 
-& \trialspace \defeq \text{Range}(\basismat),  \label{eq:spatial_subspace2}  \\
-&\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, \label{eq:spatial_subspace3} 
-\end{align}
+& \trialspace \defeq \text{Range}(\basismat),  \\%\label{eq:spatial_subspace2}  \\
+&\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, %\label{eq:spatial_subspace3} 
+\end{split}
+\end{equation}
 where 
-	$\timeSpace$ denotes the set of real-valued functions acting on the time
-	domain
-	$[0,T]$, $\onesFunction\in\timeSpace$ is defined as $\onesFunction:[0,T]\rightarrow\{1\}$,
+%	$\timeSpace$ denotes the set of real-valued functions acting on the time
+%	domain
+%	$[0,T]$, 
+$\onesFunction\in\timeSpace$ is defined as $\onesFunction:[0,T]\rightarrow\{1\}$,
 	$\basismat \in \RRStar{\romdim}{\fomdim}$ denotes the spatial trial subspace with
@@
 where $\genstate \in \RR{\romdim} \otimes \timeSpace$ denotes the generalized
-coordinates. Substituting the approximation~\eqref{eq:affine_trialspace} in
-Eq.~\eqref{eq:FOM} and performing orthogonal
+coordinates. Substituting the approximation~\eqref{eq:affine_trialspace} into the FOM ODE~\eqref{eq:FOM} and performing orthogonal
 $\elltwo$-projection of the initial condition onto the trial subspace yields
@@
 \begin{equation}\label{eq:g_truncation}
-\basisspace \frac{d\genstate }{dt} = \velocity(\basisspace
-\genstate + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
+\basisspace \frac{d \genstate} {dt}(t)= \velocity(\basisspace
+\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
 \end{equation}
@@
 \begin{equation}\label{eq:g_truncation2}
-\frac{d\genstateGalerkin}{dt} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
-\genstateGalerkin + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
+\frac{d\genstateGalerkin}{dt}(t)= \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
+\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
 \end{equation}
@@
 \begin{equation}\label{eq:GalOptimal}
-\frac{d\genstateGalerkin}{dt} = \underset{ \genstateyDiscrete \in \RR{\romdim}
+\frac{d\genstateGalerkin}{dt}(t) = \underset{ \genstateyDiscrete \in \RR{\romdim}
 	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
-	\velocity(\basisspace \genstateGalerkin + \stateIntercept,t)
+	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
 	\|_{\stweightingMatArg{n}}^2,

### 2019-10-03 main.tex
latest

@@
 This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
-	formulation for dynamical systems. The proposed approach minimizes the
+	formulation for dynamical systems. The proposed approach sequentially minimizes the
 	time-continuous full-order-model residual over a low-dimensional trial
-	subspace across a sequence of time windows and addresses key deficiencies in
-	existing model-reduction approaches, e.g., the time-step and time-scheme
-	dependence of least-squares Petrov--Galerkin (LSPG) projection, \textit{a
-	posteriori} error bounds that grow exponentially in time for Galerkin and
-	LSPG projection, and the offline cost of space--time LSPG projection.
-	\KTC{How do we address the offline cost of ST-LSPG?}
+	subspace across time windows. The approach comprises a generalization 
+of existing model reduction approaches; we show that various instances of
+  the method recover the Galerkin,
+  least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. The approach further addresses key deficiencies in
+	existing model-reduction techniques, e.g., the dependence of (space--time) LSPG 
+on the time-discretization and the exponentially growing \textit{a
+	posteriori} error bounds present in Galerkin and
+	LSPG.
 We consider two types of trial subspaces
@@
 	techniques: direct (i.e., discretize-then-optimize) and indirect (i.e.,
-	optimize-then-discretize). Critically, we show that various instances of
-	this framework yield existing model-reduction approaches, namely Galerkin,
-	LSPG, and space–time LSPG projection. Numerical experiments on the
-	compressible Euler and Navier--Stokes equations demonstrate that, at an
-	increased computational cost, the \methodAcronym\ formulation can yield more
+	optimize-then-discretize). %Critically, we show that various instances of
+	%this framework yield existing model-reduction approaches, namely Galerkin,
+	%LSPG, and space–time LSPG projection. 
+  Numerical experiments are presented for trial subspaces with spatial dimension reduction, in where it is demonstrated that the \methodAcronym\ formulation can yield more
 	accurate and physically relevant solutions than Galerkin and
-	LSPG projection.  \KTC{So which is better? Should we emphasize that it is
-	more general?}
+	LSPG projection.
 
@@
 query in nature, analysts must replace such large-scale parameterized
-dynamical-system models with a low-cost approximation that makes the
+dynamical-system models (which we refer to as the full-order model) with a low-cost approximation that makes the
 application tractable.
@@
 state ``snapshots'' at different time and parameter instances, by solving
-Lyapunov equations). Second, these methods execute an inexpensive
+Lyapunov equations, etc). Second, these methods execute an inexpensive
 \textit{online} stage during which they compute approximations to the
@@
 efforts include residual-minimizing
-projection that equips the ROM solution with a notion of optimality~\cite{l1,carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis};
+projection that equips the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
 space--time
@@
 methods to the current work and comprise the focus of the following review.
-
-\KTC{HERE}
 
@@
 Residual minimization methods in model-order reduction comprise 
-computing a solution within a low-dimensional trial space that minimizes some residual norm of the system of interest. Researchers have developed such residual-minimizing
+computing a solution within a low-dimensional trial subspace that minimizes some residual norm of the system of interest. Researchers have developed such residual-minimizing
 model-reduction methods for both static systems (i.e., systems without
@@
 from a Petrov--Galerkin projection executed on the time-continuous
-full-order-model residual, and (2) LSPG and Galekrin projection are equivalent
+full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
 in the limit as the time step goes to zero (i.e., Galerkin projection
@@
 minimizing the full-order-model residual over a finite time window (rather
-than time instantanously), LSPG projection computes solutions that are more
+than time instantaneously), LSPG projection computes solutions that are more
 accurate over a larger part of the trajectory as compared to Galerkin projection.
-
-\KTC{Let's discuss. What specifically do we want to point out about
-deficiencies in LSPG in this section? Should we also look in ST-LSPG? It seems you're trying
-to say that we don't like how the time step size \textit{and} time
-discretization method dictate so much about the method (I agree with this). If
-so, we should structure this to include other discretize-then-optimize methods
-I think. Or we could just focus on the time-step aspect (i.e., not the time
-discretization method), in which case we should focus on LSPG.}
-\KTC{HERE}
-
-The LSPG approach has several shortcomings. First, the performance of LSPG depends on the
+%\KTC{Let's discuss. What specifically do we want to point out about
+%deficiencies in LSPG in this section? Should we also look in ST-LSPG? It seems you're trying
+%to say that we don't like how the time step size \textit{and} time
+%discretization method dictate so much about the method (I agree with this). If
+%so, we should structure this to include other discretize-then-optimize methods
+%I think. Or we could just focus on the time-step aspect (i.e., not the time
+%discretization method), in which case we should focus on LSPG.}
+%\KTC{HERE}
+
+The LSPG approach, however, has several shortcomings. First, the performance of LSPG depends on the
 time discretization. Changing the time-step size modifies both the
@@
 
-Researchers have pursued the development of space--time residual minimization ROMs~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis} to address this last issue.
-In existing space--time methods, the spatial \textit{and} temporal dependence of the state is represented in a low-dimensional space--time trial space.
-Space--time residual minimization approaches then comprise computing a space--time solution within this low-dimensional space--time trial space that minimizes the space--time norm of the full-order model.
+Researchers have pursued the development of residual-minimizing space--time projection~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to address this last issue of global optimality. 
+%In existing space--time methods, the spatial \textit{and} temporal dependence of the state is represented in a low-dimensional space--time trial space. 
+Space--time residual minimization approaches comprise computing a space--time solution within a low-dimensional space--time trial space that minimizes the space--time norm of the full-order model.
 Space--time methods are thus equipped with a notion of optimality over the entire time domain of interest and, as a result, contain more favorable error bounds than the Galerkin and LSPG approaches.
-Several drawbacks of existing space--time ROMs are that they (1) require minimizing the entire space--time residual (which can be a computationally intensive process) and (2) require the specification of a finite-dimensional temporal basis (computed, e.g., via POD applied to space--time snapshots).
-This latter issue is extremely relevant to the present work.
+Several drawbacks of existing space--time ROMs are that they (1) require minimizing the entire space--time residual (which can be a computationally intensive process) and (2) require the specification of temporal basis (computed, e.g., via proper orthogonal decomposition applied to space--time snapshots).
+This latter issue precludes space--time projection methods from being used for, e.g., future state prediction.
+It is further worth noting that space--time projection comprises a fundamentally different approach than spatial projection.
+In ROMs leveraging space--time projection, the generalized coordinates (e.g., the reduced state) are a stationary vector, while the temporal dependence of the state is represented via a pre-defined basis.
+In ROMs leveraging spatial projection only, the generalized coordinates comprise a vector of \textit{time dependent functions}.
+Space--time projection executed at the time-continuous level yields an algebraic system of equations, while spatial projection executed at the time-continuous level yields a dynamical system.
+%the trial basis comprises space--time \textit{functions} that map a low-dimensional vector of generalized coordinates to the space--time solution.
+In spatial projection, the trial space comprises space--time functions that map a low-dimensional vector of
 %These space--time In these space--time ROMs, the temporal dependence of the state 
@@
 residual minimization approaches for reduced-order models. To this end, we
-propose the \methodNameLower\ (\methodAcronym) method for model reduction,
-which minimizes the FOM ODE residual over a sequence of arbitrarily defined
-time windows. The key novelties of the work are twofold.  First, the method
+propose the \methodNameLower\ (\methodAcronym) method for model reduction. The 
+\methodAcronym\ approach sequentially minimizes the FOM ODE residual over a sequence of arbitrarily defined
+time windows. The key novelties of the work are threefold.  First, the method
 minimizes the \textit{time-continuous}  residual (i.e., that associated with
@@
 formulation decouples the underlying temporal discretization scheme from the
-residual minimization problem, thus addressing a key deficiency of LSPG
-\KTC{This is also true for ST-LSPG and maybe other discretization-based
-methods? Have to check}. Time-continuous residual minimization also exposes
+residual minimization problem, thus addressing a key deficiency of (space--time) LSPG. 
+Time-continuous residual minimization also exposes
 two different solution methods,  namely a \textit{discretize-then-optimize}
@@
 sequentially minimizing the residual over time steps or over the entire time
-domain. This equips the method with additionaly flexibility that enables it to
-explore more fine-grained tradeoffs between computational cost and error. We
-demonstrate that this method can be viewed as a generalization of existing
-methods, because particular instances of this formulation are equivalent to
+domain. This equips the method with additionally flexibility that enables it to
+explore more fine-grained tradeoffs between computational cost and error. Finally, \methodAcronym\ is formulated for 
+a general space--time trial space, thus allowing application to both classic spatial projection-only trial subspaces as well as  space--time projection trial subspaces; previous work on space--time residual minimization requires space-time trial subspaces obtained through space--time projection.
+We
+demonstrate that \methodAcronym\ is a generalization of existing
+methods; particular instances of the formulation are equivalent to
 Galerkin, LSPG, and space--time LSPG projection. Figure~\ref{fig:flowchart}
@@
 
-The resulting approach is best viewed as a hybrid space--time method. A
-variety of research has examined the formulation of space--time reduced-order
-models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,bui_thesis}
-and the present approach displays commonalities with several such efforts.
+The \methodAcronym\ approach is best viewed as a hybrid space--time method and displays 
+commonalities with several related efforts. 
 First,~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
 and connected this formulation with optimal control; specifically the thesis
-of optimize then discretize vs discretize then optimize was mentioned. This
-formulation, however, was not fully developed and was eschewed for the
-time-local residual minimization method (i.e., LSPG). The present work thus
+of optimize-then-discretize vs discretize-then-optimize was mentioned. This
+formulation, however, was not fully developed and was eschewed for 
+time-local residual minimization (i.e., LSPG). The present work thus
 formally develops (and extends) several of the concepts put forth
@@
 present work distinguishes itself from Ref.~\cite{constantine_strom} in that
-(1.) in addition to trial subspaces with space--time dimension reduction, we
-consider trial subspaces with spatial dimension reduction only (i.e., a classic
-spatial basis is employed, opposed to a space--time basis) -- in which case
-the generalized coordinates comprise a vector of functions, and (2.) we
+(1) in addition to trial subspaces obtained via space--time projection, we
+consider trial subspaces obtained via spatial projection only (i.e., a classic
+spatial basis is employed, opposed to a space--time basis)% -- in which case
+%the generalized coordinates comprise a vector of functions
+, and (2) we
 minimize the time-continuous residual over space--time
@@
 The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that
-(1.) we perform model reduction of the time-continuous system, (2.) we
-consider trial subspaces with both space--time dimension reduction and spatial
-dimension reduction, and (3.) residual minimization is performed over time
-windows; Ref.~\cite{choi_stlspg} considers space--time bases and minimizes the
+(1) we perform model reduction of the time-continuous system, (2) we
+consider trial subspaces obtained via both spatial projection only and space--time projection,
+ and (3) residual minimization is performed over time
+windows; Ref.~\cite{choi_stlspg} considers trial subspaces obtained via space--time projection and minimizes the
 fully-discrete residual over the entire time domain.
@@
 \begin{align*}
-\residST_{\text{LMS}} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
+\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
 %\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
@@
 \begin{align*}
-\genresidST_{\text{LMS}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
+\genresidST_{\text{}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
 %\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
@@
 \begin{equation*}
-\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{LMS}}(\stgenstatey;\stateFOMIC) ||_2^2, 
+\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
 \end{equation*}
@@
 
-ST-LSPG is appealing in that it (1.) allows for both spatial and temporal dimension reduction and (2.) minimizes the residual over the \textit{entire} space--time domain (as opposed to a single time-step instance as in LSPG).
+ST-LSPG is appealing in that it (1) allows for both spatial and temporal dimension reduction and (2) minimizes the residual over the \textit{entire} space--time domain (as opposed to a single time-step instance as in LSPG).
 As a result, ST-LSPG can potentially be more efficient than standard LSPG and is equipped with more favorable error bounds~\cite{choi_stlspg}.
@@
 we look to overcome LSPG's time-step and time-scheme dependence, the stability challenges of both Galerkin and LSPG, and the potential computational cost of ST-LSPG.
-To this end, we now outline the \methodName\ formulation.
+To this end, we now outline the \methodNameLower\ formulation.
 %\end{comment}

### 2019-10-03 section3/general_form.tex
working on intro
latest

@@
 outlines the proposed \methodNameLower\ 
-formulation (\methodAcronym). In contrast to (1.) the Galerkin method, which minimizes the 
-time-instantaneous residual, (2.) the LSPG method, which sequentially minimizes the discrete 
-residual over a single time step, and (3.) the ST-LSPG method, which minimizes the discrete 
+formulation (\methodAcronym). In contrast to (1) the Galerkin method, which minimizes the 
+time-instantaneous residual, (2) the LSPG method, which sequentially minimizes the discrete 
+residual over a single time step, and (3) the ST-LSPG method, which minimizes the discrete 
 space--time residual, the proposed \methodAcronym\ method sequentially minimizes the 
@@
 \RR{\fomdim}$
-is the \methodAcronym\ approximation to the state over the $n$th time window, $\timeSpaceArg{n}$ is the set of functions acting on $[\timeStartArg{n},\timeEndArg{n}]$, and  
+is the \methodAcronym\ approximation to the state over the $n$th time window, $\timeSpaceArg{n}$ is the set of real-valued functions acting on $[\timeStartArg{n},\timeEndArg{n}]$, and  
 $\timeStartArg{n}$ and $\timeEndArg{n}$ are the beginning and end of the $n$th

### 2019-10-05 main.tex
name change
reread intro

@@
 %\title{Windowed least--squares reduced-order models for dynamical systems}
-\title{Windowed residual-minimization model reduction for dynamical systems}
+\title{Windowed least-squares model reduction for dynamical systems}
 
@@
 	formulation for dynamical systems. The proposed approach sequentially minimizes the
-	time-continuous full-order-model residual over a low-dimensional trial
-	subspace across time windows. The approach comprises a generalization 
-of existing model reduction approaches; we show that various instances of
-  the method recover the Galerkin,
-  least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. The approach further addresses key deficiencies in
-	existing model-reduction techniques, e.g., the dependence of (space--time) LSPG 
-on the time-discretization and the exponentially growing \textit{a
-	posteriori} error bounds present in Galerkin and
-	LSPG.
-We consider two types of trial subspaces
-	within the proposed framework:
-	one that reduces only the spatial dimension of the full-order model, and one that
-	reduces both the spatial and temporal dimensions of the full-order model.
-	For each type of trial subspaces, we consider two different solution
-	techniques: direct (i.e., discretize-then-optimize) and indirect (i.e.,
-	optimize-then-discretize). %Critically, we show that various instances of
-	%this framework yield existing model-reduction approaches, namely Galerkin,
-	%LSPG, and space–time LSPG projection. 
-  Numerical experiments are presented for trial subspaces with spatial dimension reduction, in where it is demonstrated that the \methodAcronym\ formulation can yield more
-	accurate and physically relevant solutions than Galerkin and
-	LSPG projection.
+	time-continuous full-order-model residual over a low-dimensional space--time trial
+	subspace defined over an arbitrarily defined time window. The approach comprises a generalization 
+of existing model reduction approaches, as particular instances of
+  the methodology associate with Galerkin,
+	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
+	addition, the approach
+	addresses key deficiencies in existing model-reduction
+	techniques, e.g., the dependence of (space--time) LSPG projection on the
+	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
+	error bounds for Galerkin and LSPG projection.  We consider two types of
+	space--time trial
+	subspaces within the proposed framework: one that reduces only the spatial
+	dimension of the full-order model, and one that reduces both the spatial and
+	temporal dimensions of the full-order model.  For each type of trial
+	subspace, we consider two different solution techniques: direct (i.e.,
+	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
+	Numerical experiments conducted using the trial subspaces characterized by spatial
+	dimension reduction demonstrate that the \methodAcronym\
+	approach can yield more accurate and physically relevant solutions \KTC{What
+	does `physically relevant' mean? This is imprecise.} than
+	Galerkin and LSPG projection.
 
@@
 
-Simulating parameterized dynamical systems underpins many applications across
-science and engineering, including engineering design, control, financial
-analysis, and epidemiology. In many contexts, executing a dynamical-system
+Simulating parameterized dynamical systems arises in many applications across
+science and engineering, including design, control, uncertainty
+quantification, and epidemiology, for example. In many contexts, executing a dynamical-system
 simulation at a single parameter instance---which entails the numerical
 integration of a system of ordinary differential equations (ODEs)---incurs an
-extremely large computational expense.  This occurs, for example, when the
+extremely large computational cost.  This occurs, for example, when the
 state-space dimension is large (e.g., due to fine spatial resolution when
@@
 strategy. First, these techniques execute a computationally expensive
-\textit{offline} stage that computes a low-dimensional ``trial'' subspace on
+\textit{offline} stage that computes a low-dimensional \textit{trial subspace} on
 which the dynamical-system state can be well approximated (e.g., by computing
 state ``snapshots'' at different time and parameter instances, by solving
-Lyapunov equations, etc). Second, these methods execute an inexpensive
+Lyapunov equations). Second, these methods execute an inexpensive
 \textit{online} stage during which they compute approximations to the
@@
 
-Model reduction for linear-time-invariant systems (and other strongly
-structured dynamical systems) is quite mature \cite{wilcox_benner_rev,moore,roberts,GugercinIRKA} as system-theoretic properties (e.g., controllability,
+Model reduction for linear-time-invariant systems (and other well structured
+dynamical systems) is quite mature
+\cite{wilcox_benner_rev,moore,roberts,GugercinIRKA}, as system-theoretic properties (e.g., controllability,
 observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
-quantified and accounted for in the projection process, which yields
-reduced-order models that inherit such properties.  The primary challenge for
+quantified and accounted for in the projection process; this often results in
+reduced-order models that inherit such important properties.  The primary
+challenge in
 developing reduced-order models for general nonlinear dynamical systems is
-that such properties are difficult to quantitatively assess. As a result, it
+that such properties are difficult to assess quantitatively. As a result, it
 is difficult to develop reduced-order models that preserve important
-dynamical-system properties, which often results in inaccurate or unstable
-methods.  To address this, researchers have pursued several directions that
+dynamical-system properties, which often results in methods that yield
+trajectories that are
+inaccurate, unstable, or violate physical properties.  To address this, researchers have pursued several directions that
 aim to imbue reduced-order models for nonlinear
@@
 ``energy-based'' inner
-products that ensure non-increasing entropy in the ROM solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014};
+products that ensure non-increasing entropy in the ROM
+solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014,chan2019entropy};
 basis-adaptation methods that improve the ROM's accuracy \textit{a
@@
 Residual-minimizing and space--time projection are the most relevant classes of
-methods to the current work and comprise the focus of the following review.
+methods for the current work and comprise the focus of the following review.
 
@@
 
-Residual minimization methods in model-order reduction comprise 
+\KTC{HERE, COFFEE BREAK}
+Residual minimization methods in model reduction comprise 
 computing a solution within a low-dimensional trial subspace that minimizes some residual norm of the system of interest. Researchers have developed such residual-minimizing
@@
 %\section*{References}
-\bibliographystyle{aiaa}
-\bibliography{refs}{}
+\bibliographystyle{siam}
+\bibliography{refs}
 \end{document}

### 2019-10-05 main.tex
prog

@@
 	formulation for dynamical systems.  
-% In contrast to (1.) the Galerkin method, which minimizes the time-instantaneous residual, (2.) the Least-Squares Petrov--Galerkin (LSPG) method, which sequentially minimizes the discrete residual over a single time step, and (3.) the Space--Time LSPG method, which minimizes the discrete space–time residual,
-%In contrast to existing residual minimization model-reduction approaches, which generate solutions in a low-dimensional space--time trial space that either minimizes (1.) the time-continuous residual over an infinitesimal time-window (Galerkin), (2.) the time-discrete residual over a single time-step instance (Least-Squares Petrov--Galerkin), or (3.) the time-discrete residual over the entire time domain (Space--Time Least-Squares Petrov--Galerkin),
-%the proposed approach minimizes the
-%	time-continuous full-order-model residual over a low-dimensional space--time
-%	trial subspace across a sequence of time windows. Among other advantages, this process yields an approach that overcomes the 
-%time-step and time-scheme dependence that the least-squares Petrov--Galerkin method is subject to. 
-%that decompose the time
-%	domain of interest.
 The proposed approach minimizes the
@@
 	trial subspace across a sequence of time windows. % that decompose the time
-%	domain of interest. The proposed approach overcomes the time-step and time-scheme dependence that the least-squares Petrov--Galerkin method is subject to.  
-%	\KTC{Why are we saying the next two sentences? These seem kind of random.
-%	The abstract should explain why the work is necessary; this more reads like
-%	a list of interesting factoids.} %The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the reduced-order model.
-The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
-%For spatial trial spaces, the stationary conditions for the time-continuous minimization problems are obtained 
-%by deriving the associated Euler--Lagrange equations. 
 We consider two types of trial subspaces
@@
 	reduces both the spatial and temporal dimensions of the full-order model. 
-%The stationary conditions for the resulting minimization problems are derived.
-Of particular interest, the stationary conditions for trial spaces with spatial dimension reduction are obtained via the Euler–Lagrange equations and comprise a coupled two-point Hamiltonian boundary value problem containing a forward and backward system.
 We consider two methods to solve the residual-minimization problems: direct
@@
 Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
-%  We show that the proposed method
-%	comprises a generalization of many existing techniques and provides a
-%	unified framework to reason about existing model-redcution techniques. In
-%	particular, we show that specific instances of the proposed method lead to
-%	the existing reduced-order modeling techniques of Galerkin projection,
-%	least-squares Petrov--Galerkin (LSPG) projection, and space--time
-%	least-squares Petrov--Galerkin (ST-LSPG) projection. 
-%By formulating the least--squares problem from the time-continuous level, the \methodAcronym\ formulation overcomes 
-%the time-step sensitivity and time-scheme dependence that LSPG is subject to. 
-%The stationary points of the minimization problems are obtained by solving the associated Euler-Lagrange equations.
-The Euler-Lagrange equations consist of a set of coupled two point boundary value problems defined on each time-slab containing a forward primal system and a backward adjoint system.
-%The approach is conceptually similar to the classic Least-Squares Petrov-Galerkin approach, but displays several advantages.
-First, by formulating the residual minimization problem at the continuous level, the sensitivity of the resulting method to the underlying time-discretization scheme is greatly reduced.
 Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.
 \end{comment}
-%This work proposes a least-squares residual minimization approach for reduced-order models of dynamical systems.
-The proposed approach, which comprises a general least-squares reduced-order modeling (LSROM) formulation for dynamical systems, sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial space over a series of time slabs.
-The stationary conditions for the time-continuous minimization problems are obtained
-%by deriving the associated Euler--Lagrange equations.
-The time-continuous formulation exposes two differing methods that can be used to solve the minimization problem, namely direct (discretize then minimize) and indirect (minimize and then discretize).
-The proposed approach displays commonalities with optimal control problems and is a generalization of the popular
-%Least-Squares Petrov--Galerkin (LSPG) method. By formulating the residual minimization problem from the time-continuous level, the \methodAcronym\ approach overcomes 
-%the time-step sensitivity and time-scheme dependence that LSPG is subject to. 
-%The stationary points of the minimization problems are obtained by solving the associated Euler-Lagrange equations.
-The Euler-Lagrange equations consist of a set of coupled two point boundary value problems defined on each time-slab containing a forward primal system and a backward adjoint system.
-%The approach is conceptually similar to the classic Least-Squares Petrov-Galerkin approach, but displays several advantages.
-First, by formulating the residual minimization problem at the continuous level, the sensitivity of the resulting method to the underlying time-discretization scheme is greatly reduced.
-%Numerical experiments on the compressible Euler equations demonstrate that the proposed approach leads to more accurate and physically relevant solutions than existing model reduction approaches.
-
 \end{abstract}
@@
 application tractable.
-
-%\KTC{I reworked this. Methods like balanced truncation or IRKA do \textit{not}
-%use snapshot data!}
 
@@
 
-%~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in Refs~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
-Specifically, Refs~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
-
-%\begin{itemize}
-%\item Galerkin is inaccurate/unstable for non-symmetric static systems -- motivates residual minimization
-%\item Residual minimization (stems from FEM): minimizes the residual of the full-order model over the range of the trial space -- originally developed for static systems
-%\item Can guarantee stability for linear static systems 
-%\item Extension to time-varying systems is challenging 
-%\begin{itemize}
-%\item What is the definition of the ODE residual?
-%\item Over what time-interval is the residual being minmized?
-%\end{itemize}
-%\end{itemize}
-
-%\KTC{Why are we reviewing LSFEM? This isn't a ROM. Don't think it's relevant}
-%Least-squares residual minimization as a numerical solution technique for
-%partial differential equations was pioneered in the finite-element community
-%and led to what is known as the least-squares finite-element method
-%(LSFEM)~\cite{ls_review,bochev_leastsquares}. 
-%\KTC{Specifically, LSFEM was designed to solve the inf--sup stability problem.
-%Be more precise.}
-%Motivated by the fact that the
-%Galerkin method can be inaccurate and unstable for nonlinear and
-%non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic
-%functional comprising the residual norm of the partial differential equation
-%%of interest and (2.) leveraging variational principles to derive the
-%stationary conditions of the quadratic functional; i.e, LSFEM can be thought
-%of as finding the solution within a finite element trial space the minimizes
-%the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous
-%advantages (e.g., least-squares system are symmetric and positive definite)
-%\KTC{This is not an `advantage'. It is guaranteed to be inf--sup stable.}
-%over the more standard Galerkin discretization.
-
-\KTC{HERE, COFFEE BREAK}
-Residual minimization methods in model reduction comprise 
-computing a solution within a low-dimensional trial subspace that minimizes some residual norm of the system of interest. Researchers have developed such residual-minimizing
-model-reduction methods for both static systems (i.e., systems without
-time-dependence)
+Residual-minimization methods in model reduction compute the solution within a
+low-dimensional trial subspace that minimizes the full-order-model residual.
+Researchers have developed such residual-minimizing model-reduction methods
+for both static systems (i.e., systems without time-dependence)
 \cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
@@
 in the limit as the time step goes to zero (i.e., Galerkin projection
-minimizes the time-instantaneous full-order model residual). Numerous
+minimizes the time-instantaneous full-order-model residual). Numerous
 numerical experiments have demonstrated that LSPG projection often yields more
@@
 accurate over a larger part of the trajectory as compared to Galerkin projection.
-%\KTC{Let's discuss. What specifically do we want to point out about
-%deficiencies in LSPG in this section? Should we also look in ST-LSPG? It seems you're trying
-%to say that we don't like how the time step size \textit{and} time
-%discretization method dictate so much about the method (I agree with this). If
-%so, we should structure this to include other discretize-then-optimize methods
-%I think. Or we could just focus on the time-step aspect (i.e., not the time
-%discretization method), in which case we should focus on LSPG.}
-%\KTC{HERE}
-
-The LSPG approach, however, has several shortcomings. First, the performance of LSPG depends on the
-time discretization. Changing the time-step size modifies both the
-time window over which LSPG minimizes the residual \textit{as well as} the
-time-discretization error of the full-order model on which LSPG is based. As a
-result, LSPG often yields the smallest error for an intermediate value of the
-time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}). In
-particular, as the time step decreases, the accuracy of LSPG approaches that
-of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
-is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.}. 
-%The accuracy of LSPG additionally degrades if the time-step size becomes too large.
-It is unclear if this is due to the discretization error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
-LSPG projection additionally is
-%additionally sensitive to the time discretization scheme. When used with
-%explicit temporal discretization, for instance, LSPG again reverts to
-%Galerkin. To gain the enhanced stability associated with LSPG, the underlying
-%time-discretization scheme must be implicit with an appropriate time-step.
-%%This can increase the cost of the reduced-order model, especially in the case
-%where the optimal time-step required by LSPG is small. 
-Second, as the LSPG approach sequentially performs residual minimization over
-time-steps, is not equipped with any notion of optimality over the entire 
-time domain of interest. As a result, LSPG is not equipped with \textit{a priori}
-guarantees of stability or convergence, even for linear time-invariant
-systems~\cite{bui_thesis}.
-
-Researchers have pursued the development of residual-minimizing space--time projection~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to address this last issue of global optimality. 
-%In existing space--time methods, the spatial \textit{and} temporal dependence of the state is represented in a low-dimensional space--time trial space. 
-Space--time residual minimization approaches comprise computing a space--time solution within a low-dimensional space--time trial space that minimizes the space--time norm of the full-order model.
-Space--time methods are thus equipped with a notion of optimality over the entire time domain of interest and, as a result, contain more favorable error bounds than the Galerkin and LSPG approaches.
-Several drawbacks of existing space--time ROMs are that they (1) require minimizing the entire space--time residual (which can be a computationally intensive process) and (2) require the specification of temporal basis (computed, e.g., via proper orthogonal decomposition applied to space--time snapshots).
-This latter issue precludes space--time projection methods from being used for, e.g., future state prediction.
-It is further worth noting that space--time projection comprises a fundamentally different approach than spatial projection.
-In ROMs leveraging space--time projection, the generalized coordinates (e.g., the reduced state) are a stationary vector, while the temporal dependence of the state is represented via a pre-defined basis.
-In ROMs leveraging spatial projection only, the generalized coordinates comprise a vector of \textit{time dependent functions}.
-Space--time projection executed at the time-continuous level yields an algebraic system of equations, while spatial projection executed at the time-continuous level yields a dynamical system.
-%the trial basis comprises space--time \textit{functions} that map a low-dimensional vector of generalized coordinates to the space--time solution.
-In spatial projection, the trial space comprises space--time functions that map a low-dimensional vector of
-%These space--time In these space--time ROMs, the temporal dependence of the state 
-%is represented via a finite dimensional basis representation. 
-%Space--time residual minimization methods can be prohibitively expensive, however, as they involve minimizing the entire space--time residual.
-Further, existing space--time approaches require a finite-dimensional space--time basis.
-
-%This work seeks to overcome the practical challenges associated with existing
-%residual minimization approaches for reduced-order models. To this end, we
-%propose the \methodName\ method for model reduction, which minimizes the FOM ODE residual
-%over a sequence of time windows. The key novelties of the work are two fold.
-%First, the residual minimization problem is formulated to minimize the
-%\textit{time-continuous} residual. By working from a time-continuous viewpoint, the
-%present formulation decouples the underlying temporal discretization scheme
-%from the residual minimization problem; thus addressing a key deficiency of
-%LSPG \KTC{This is also true for ST-LSPG and maybe other discretization-based
-%methods? Have to check}. The time-continuous viewpoint additionally exposes
-%two different methods that
-%can be used to solve the residual-minimization problems, namely
-%a \textit{discretize-then-optimize} (i.e., direct) method and
-%an \textit{optimize-then-discretize} (i.e., indirect) method. The second novelty
-%is that the residual is minimized over \textit{time windows} rather than over
-%a single time step or the entire time domain. This equips the method with
-%additionaly flexibility that enables it
-%to explore more fine-grained tradeoffs in computational cost and error. We
-%demonstrate that this method can be viewed as a generalization of existing
-%methods, as particular instances of this formulation recover Galerkin, LSPG,
-%and space--time LSPG projection. Figure~\ref{fig:flowchart} depicts how the
-%proposed method is related to existing approaches.
-
-% \KTC{This is what ST-LSPG does. Why do we ignore it? the
-%full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
-%guarantees of stability or convergence, even for linear time-invariant
-%systems~\cite{bui_thesis}. Further, the performance of LSPG depends on the
-%time step in two distinct ways: changing the time-step size modifies both the
-%time window over which LSPG minimizes the residual \textit{as well as} the
-%time-discretization error of the full-order model on which LSPG is based. As a
-%result, LSPG often yields the smallest error for an intermediate value of the
-%time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}). In
-%particular, as the time step decreases, the accuracy of LSPG approaches that
-%of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
-%is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
-%as the time step increases, the time discretization error of the full-order
-%model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
-%additionally sensitive to the time discretization scheme. When used with
-%explicit temporal discretization, for instance, LSPG again reverts to
-%Galerkin. To gain the enhanced stability associated with LSPG, the underlying
-%time-discretization scheme must be implicit with an appropriate time-step.
-%This can increase the cost of the reduced-order model, especially in the case
-%where the optimal time-step required by LSPG is small.
-
-%The thesis of least--squares residual minimization was extended to construct reduced-order models of static systems by several authors, with two such instances being~\cite{rovas_thesis} and~\cite{bui_resmin}.
-Similar to the finite-element case, numerous theoretical analyses and numerical experiments have demonstrated that residual minimization methods yield more accurate and stable solutions than the Galerkin method.
-The extension of least--squares residual minimization methods to dynamical systems, however, has proved challenging.
-The least-squares Petrov-Galerkin method is arguably the most commonly adapted residual minimization method for dynamical systems.
-LSPG extends the idea of least--squares residual minimization to dynamical systems by sequentially minimizing the \textit{fully discrete} residual (i.e., after temporal discretization) in a weighted norm (nominally $\elltwo$) over each time step.
-Numerous numerical experiments have demonstrated that the LSPG approach yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
-The common explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.
-
-%The LSPG approach, however, is still subject to exponentially 
-%growing error bounds and, in general, lacks \textit{a priori} guarantees of stability and accuracy.
-Further, as it is defined at the fully discrete level, LSPG contains several nuances.
-First, the LSPG approach is sensitive to the time-step.
-In Ref.~\cite{carlberg_lspg_v_galerkin}, LSPG was shown to provide optimal results at an intermediary time-step.
-For too small a time-step, the LSPG approach is prone to the same stability challenges as the Galerkin approach.
-In the limiting case of $\Delta t \rightarrow 0$, the LSPG approach reverts to a Galerkin approach.
-For too-large a time-step, on the other hand, the accuracy of the LSPG method degrades.
-It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
-The LSPG approach is additionally sensitive to the time discretization scheme.
-When used with explicit temporal discretization, for instance, the LSPG approach reverts to a Galerkin approach.
-To gain the enhanced stability associated with the LSPG method, the underlying time-discretization scheme must be implicit with an appropriate time-step.
-This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.
-
-%Residual minimization methods are the most relevant to the current work.
-Residual minimization methods seek to minimize the residual of the full-order model over the trial space defined by the POD basis.
-Two of the most popular reduced-order modeling approaches are (or can be interpreted as) residual minimization approaches.
-The Galerkin method, which is arguably the most common reduced-order modeling approach, minimizes the instantaneous FOM residual over the range of the trial basis~\cite{carlberg_lspg_v_galerkin}.
-%The Galerkin method is arguably the most common reduced-order modeling approach.
-In Ref~\cite{carlberg_lspg_v_galerkin}, Carlberg and Barone demonstrated that the Galerkin method corresponds to finding a velocity that minimizies the instananeous FOM residual over the range of the trial basis.
-%Numerical experiments and theoretical analyses have shown, however, that the Galerkin method can be inappropriate for model-order reduction of general nonlinear dynamical systems~\cite{carlberg_gnat,carlberg_thesis,huang_combustion_roms}; error analysis has demonstrated that the Galerkin method is subject to exponentially growing error bounds while a variety of numerical experiments have demonstrated the lack of stability and inaccuracy of the method in practice.
-Indeed, it is well known in numerical analysis that Galerkin method performs poorly for non-symmetric systems.
-To address these issues, Carlberg et al.
-developed the least-squares Petrov-Galerkin (LSPG) method.
-LSPG sequentially minimizes the \textit{fully discrete} residual (i.e., after temporal discretization) in a weighted norm over each time-step.
-In this sense, the LSPG method is typically referred to as being \textit{discretely-optimal}.
-Numerous numerical experiments have demonstrated that the LSPG approach yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
-The common explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.
-The LSPG approach, however, is still subject to exponentially
-%growing error bounds and, in general, lacks \textit{a priori} guarantees of stability and accuracy.
-Further, as it is defined at the fully discrete level, LSPG contains several nuances.
-First, the LSPG approach is sensitive to the time-step.
-In Ref.~\cite{carlberg_lspg_v_galerkin}, LSPG was shown to provide optimal results at an intermediary time-step.
-For too small a time-step, the LSPG approach is prone to the same stability challenges as the Galerkin approach.
-In the limiting case of $\Delta t \rightarrow 0$, the LSPG approach reverts to a Galerkin approach.
-For too-large a time-step, on the other hand, the accuracy of the LSPG method degrades.
-It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over.
-The LSPG approach is additionally sensitive to the time discretization scheme.
-When used with explicit temporal discretization, for instance, the LSPG approach reverts to a Galerkin approach.
-To gain the enhanced stability associated with the LSPG method, the underlying time-discretization scheme must be implicit with an appropriate time-step.
-This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.
-
-%The nuances associated with the LSPG approach lead to practical challenges.
-For instance, minimal analysis on the optimal \textit{a priori} time-step exists, and as such specifying an appropriate time-step for LSPG \textit{a priori} is challenging.
-The limitation to implicit time-schemes is also restrictive, especially considering the fact that ROMs are often much less stiff than the corresponding FOM~\cite{Bach2018StabilityCF} and thus explicit time-stepping schemes with large time-steps can sometimes be used.
-It is emphasized that, despite these challenges, LSPG is arguably the most robust model-order reduction
-%formulation for general nonlinear dynamical systems and has gained wide-spread use~\cite{narmal,cheng,4,5,6}.\EP{Finish with good Refs using LSPG} 
-%Time-discrete space-time residual minimization~\cite{choi_stlspg,constantine_strom} is the third approach that has been proposed.
-Reference~\cite{choi_stlspg} extends the LSPG method to space-time formulations and develops the ST-LSPG method.
-%ST-LPSG seeks a space-time solution within a space-time trial space that is a minimizer of the fully discrete residual.
-The space-time trial spaces
-%ST-LSPG seeks to find the minimizer of the \textit{fully discrete} residual over the entire space-time domain within a given space-time trial space.
-These space-time trial spaces are constructed through higher-order singular value decompositions (SVD), such as the truncated high-order SVD.
-%\item How can a finite-time residual minimization approach be developed that is independent of the underlying temporal discretization?
-%\item Does minimizing the residual over a larger time-window improve the model prediction?
-%\end{enumerate}
-
-%Recently, LSROMs for dynamical systems have been extended to the space--time setting~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC}.
-These space--time LSROMs (ST-LSROMs), which are typically motivated from the need to reduce the spatial \textit{and} temporal dimension of a dynamical system model, comprise finding a space--time solution that minimizes the space--time residual within a low-dimensional space--time trial space.
-In contrast to standard ROMs for dynamical systems, in where the generalized coordinates are \textit{(discretized) time-dependent functions}, space--time methods ROMs represent the temporal dependence of the state via a finite dimensional temporal basis and thus the generalized coordinates are a stationary vector.
-ST-LSROMs thus differ from their standard LSROM counterparts in that they 1.) minimize the full space--time residual and 2.) represent the temporal dependence of the state via a prescribed trial space.
-ST-LSROMs thus comprise algebraic systems, while LSROMs comprise dynamical systems.
-
-%ST-LSROMs involve a monolithic space--time least--squares solve, and are typically equipped with more robust error bounds than their non-space--time counterparts.
-To best describe how these existing methods fit into the present work, they must be viewed in a two dimensional space: one dimension corresponding the space--time trial space and one dimension corresponding to the space--time residual minimization problem.
-With the exception of Ref.~\cite{bui_thesis}, existing space--time methods leverage reduced spatial and temporal trial spaces.
-In Ref.~\ref{choi_stlspg}, these trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD.
-%Ref.~\cite{choi_stlspg} develops a space--time formulation of the least-squares Petrov-Galerkin approach.
-The formulation minimizes the \textit{fully discrete} residual over the entire space--time domain within a given space--time trial space.
-These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD.
-%Another piece of relevant work is that performed by Constantine and Wang~\cite{constantine_strom}, in which a space--time residual minimization formulation for model interpolation is pursued.
-The present work again distinguishes itself in that it seeks to minimize the time-continuous residual over space--time slabs;~\cite{constantine_strom} considers the solution of a fully discrete optimization problem over the entire space--time domain.
-It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work.
-
-This work seeks to overcome the practical challenges associated with existing
-residual minimization approaches for reduced-order models. To this end, we
-propose the \methodNameLower\ (\methodAcronym) method for model reduction. The 
-\methodAcronym\ approach sequentially minimizes the FOM ODE residual over a sequence of arbitrarily defined
-time windows. The key novelties of the work are threefold.  First, the method
+
+However, the LSPG approach has several notable shortcomings. First, LSPG
+performance exhibits a complex dependence on the time discretization. In
+particular, changing
+the time step modifies both the
+time window over which LSPG minimizes the residual as well as the
+time-discretization error of the full-order model on which LSPG is based. 
+In particular, because LSPG and Galerkin projection are equivalent in the
+limit as the time step goes to zero, LSPG accuracy approaches the (poor)
+accuracy of Galerkin as the time step shrinks, but this also has the effect of
+reducing the time-dicretization error of the full-order model.
+As a
+consequence, LSPG often yields the smallest error for an intermediate value of the
+time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}); there
+is no known way to compute this optimal time step \textit{a priori}.
+Second, as the LSPG approach performs sequential residual minimization in
+time, its \textit{a posteriori} error bounds grow exponentially in time
+\cite{carlberg_lspg_v_galerkin}, and it is not equipped with any notion of
+optimality over the entire time domain of interest. As a result, LSPG is not
+equipped with \textit{a priori} guarantees of accuracy or stability, even for
+linear time-invariant systems~\cite{bui_thesis}.
+
+Researchers have pursued the development of space--time
+residual-minimization
+projection~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
+address the issues incurred by sequential residual minimization in time.
+Space--time residual-minimization approaches compute the space--time
+trajectory within a low-dimensional space--time trial subspace (spanned by
+a small number of space--time basis vectors) that minimizes the full-order-model
+residual over the entire time domain.
+We note that space--time residual minimization yields a fundamentlaly
+different model than sequential residual minimization in time. In particular, space--time
+residual minimization associates with an algebraic system of equations defined
+over all space and time, whose solution comprises a \text{fixed} vector of
+(space--time) generalized coordinate; on the other hand, sequential residual
+minimization in time (often) associates with a system of ODEs whose solution
+comprises a \textit{time-dependent} vector of (spatial) generalized
+coordinates.
+As a result, space--time residual minimization yields models that are equipped with \textit{a priori} error bounds that grow more slowly
+in time, and the resulting \textit{trajectory} exhibits a notion of
+optimality; this implies that the solution exhibits a notion of optimality
+over the entire time domain. However, these techniques also suffer from several
+limitations. First, the computational cost of solving the algebraic system arising from space--time
+projection scales cubically with the number of space--time degrees of freedom;
+in contrast, the computational cost incurred by standard
+spatial-projection-based ROMs is linear in the number of temporal degrees of
+freedom, as the attendant solvers can leverage the lower-block triangular
+structure of the system arising from the sequential nature of time
+evolution. As a result, solving the algebraic systems arising from space--time
+projection is generally intractable without applying hyper-reduction in time
+\cite{choi_stlspg,constantine_strom}. Second, space--time residual
+minimization precludes future
+state prediction, as these methods employ space--time basis vectors defined over the
+entire time domain of interest, which must have been including in the training
+simulations.
+
+The objectives of this work are to overcome the shortcomings of existing
+residual-minimizing methods, and to provide a unifying framework from which
+existing methods can be assessed. In essence, the proposed  \textit{\methodNameLower\
+(\methodAcronym)} method sequentially minimizes the FOM
+residual over a sequence of arbitrarily defined
+time windows. The method is characterized by three notable aspects.  First, the method
 minimizes the \textit{time-continuous}  residual (i.e., that associated with
@@
 formulation decouples the underlying temporal discretization scheme from the
-residual minimization problem, thus addressing a key deficiency of (space--time) LSPG. 
-Time-continuous residual minimization also exposes
-two different solution methods,  namely a \textit{discretize-then-optimize}
-(i.e., direct) method and an \textit{optimize-then-discretize} (i.e.,
-indirect) method. The second novelty is that the method sequentially minimizes
+residual-minimization problem, thus addressing a key deficiency of both LSPG
+and space--time LSPG projection. Critically, time-continuous residual minimization also exposes
+two different solution methods:  a \textit{discretize-then-optimize}
+(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
+indirect) method. Second, the method sequentially minimizes
 the residual over arbitrarily defined \textit{time windows} rather than
-sequentially minimizing the residual over time steps or over the entire time
-domain. This equips the method with additionally flexibility that enables it to
-explore more fine-grained tradeoffs between computational cost and error. Finally, \methodAcronym\ is formulated for 
-a general space--time trial space, thus allowing application to both classic spatial projection-only trial subspaces as well as  space--time projection trial subspaces; previous work on space--time residual minimization requires space-time trial subspaces obtained through space--time projection.
-We
-demonstrate that \methodAcronym\ is a generalization of existing
-methods; particular instances of the formulation are equivalent to
-Galerkin, LSPG, and space--time LSPG projection. Figure~\ref{fig:flowchart}
-depicts how the proposed \methodAcronym\ method is related to existing
-approaches.
-
-The \methodAcronym\ approach is best viewed as a hybrid space--time method and displays 
+sequentially minimizing the residual over time steps (as in LSPG projection)
+or over the entire time domain (as in space--time residual-minimization
+methods). This equips the method with additionally flexibility that enables it
+to explore more fine-grained tradeoffs between computational cost and error.
+Finally, \methodAcronym\ is formulated for two kinds of space--time trial
+subspaces: one that reduces only the spatial dimension of the full-order
+model (as employed by traditional spatial-projection-based methods), and one that reduces both the spatial and temporal dimensions of the
+full-order model (as employed by space--time residual-minimization methods).
+The above attributes allow that \methodAcronym\ method to be viewed as a
+generalization of existing model-reduction methods, as 
+Galerkin, LSPG, and space--time LSPG projection correspond to specific
+instances of the formulation.
+Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
+provides a unifying framework from which these existing approaches can be derived.
+
+The \methodAcronym\ approach can be viewed as a hybrid space--time method and displays 
 commonalities with several related efforts. 
-First,~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
-and connected this formulation with optimal control; specifically the thesis
-of optimize-then-discretize vs discretize-then-optimize was mentioned. This
-formulation, however, was not fully developed and was eschewed for 
-time-local residual minimization (i.e., LSPG). The present work thus
+First, Ref.~\cite{bui_thesis} tersely formulated a space--time least-squares ROM
+and connected this formulation with optimal control; specifically it mentioned the
+optimize-then-discretize vs.\ discretize-then-optimize approaches. This work
+did not fully develop this approach, eschewing it 
+for 
+sequential residual minimization in time (i.e., LSPG). \KTC{HERE} The present work thus
 formally develops (and extends) several of the concepts put forth
@@
 domain within a given space--time trial subspace. 
-%These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD. 
 The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that
@@
 fully-discrete residual over the entire time domain.
-
-%Ref.~\cite{choi_stlspg} develops a space--time formulation of the least-squares Petrov-Galerkin approach.
-The formulation minimizes the \textit{fully discrete} residual over the entire space--time domain within a given space--time trial space.
-These space--time trial spaces are constructed through higher-order singular value decompositions (SVDs), such as the truncated high-order SVD.
-The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that it seeks to minimize the time-continuous residual over time slabs, as opposed to the fully-discrete residual over the entire domain.
-Another point of departure from Ref.~\cite{choi_stlspg} is that the present work does not consider temporal reduction; this will be a subject of future work.
-Instead, the temporal basis is taken to have full support.
-Another piece of relevant work is that performed by Constantine and Wang~\cite{constantine_strom}, in which a space--time residual minimization formulation for model interpolation is pursued.
-The present work again distinguishes itself in that it seeks to minimize the time-continuous residual over space--time slabs;~\cite{constantine_strom} considers the solution of a fully discrete optimization problem over the entire space--time domain.
-It is noted that, similar to the current work, a time-continuous residual-minimization formulation is used as a starting point in~\cite{constantine_strom}, and as such thematic similarities exist between~\cite{constantine_strom} and the present work.
 
@@
 However, for general non-self-adjoint and nonlinear problems, LSPG, lacks \textit{a priori} guarantees of stability, accuracy, and convergence. Further, 
-LSPG is sensitive to the time-step size: changing the time-step size changes the window over which the residual is minimized 
+LSPG is sensitive to the time step: changing the time step changes the window over which the residual is minimized 
 \textit{as well as} the temporal discretization error. As a result, optimal results are obtained at an intermediary time-step. 

### 2019-10-06 main.tex
done for day

@@
 \title{Windowed least-squares model reduction for dynamical systems}
+%\title{The windowed least-squares framework for model reduction of dynamical systems}
 
@@
 for 
-sequential residual minimization in time (i.e., LSPG). \KTC{HERE} The present work thus
-formally develops (and extends) several of the concepts put forth
-in~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
+sequential residual minimization in time (i.e., LSPG). The present work thus
+formally develops and extends several of the concepts put forth
+in Ref.~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
 space--time residual minimization formulation for model interpolation.  The
 present work distinguishes itself from Ref.~\cite{constantine_strom} in that
-(1) in addition to trial subspaces obtained via space--time projection, we
-consider trial subspaces obtained via spatial projection only (i.e., a classic
-spatial basis is employed, opposed to a space--time basis)% -- in which case
-%the generalized coordinates comprise a vector of functions
-, and (2) we
-minimize the time-continuous residual over space--time
-windows;~\cite{constantine_strom} considers the solution of a fully discrete
-optimization problem over the entire space--time domain. It is noted that,
-similar to the current work, a time-continuous residual-minimization
-formulation is used as a starting point in~\cite{constantine_strom}, and as
-such thematic similarities exist between~\cite{constantine_strom} and the
-present work. Lastly, Ref.~\cite{choi_stlspg} develops a space--time
-formulation of the least-squares Petrov--Galerkin approach. The formulation
-minimizes the \textit{fully discrete} residual over the entire space--time
-domain within a given space--time trial subspace. 
-The present work distinguishes itself from Ref.~\cite{choi_stlspg} in that
-(1) we perform model reduction of the time-continuous system, (2) we
-consider trial subspaces obtained via both spatial projection only and space--time projection,
- and (3) residual minimization is performed over time
-windows; Ref.~\cite{choi_stlspg} considers trial subspaces obtained via space--time projection and minimizes the
-fully-discrete residual over the entire time domain.
+(1) this work also considers trial subspaces characterized by a reduction of
+the spatial dimension of the FOM only, and (2) we
+minimize the time-continuous FOM residual over arbitrary time
+windows. We note that,
+similar to the current work, Ref.~\cite{constantine_strom} employs
+minimization of the time-continuous FOM residual as a starting point; as such,
+this work shares some thematic similarities with
+Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
+space--time extension of the LSPG approach that minimizes the time-discrete
+FOM residual over the entire time domain.  The present work distinguishes
+itself from Ref.~\cite{choi_stlspg} in that (1) this work minimizes the
+time-continuous FOM residual, (2) this work minimizes this residual over
+arbitrary time windows, and (3) this work considers trial subspaces
+characterized by a reduction of the spatial dimension of the FOM only.
 
@@
 \caption{Relationship diagram for the \methodAcronym\ framework for model
-	reduction.} 
+	reduction. \KTC{Inconsistent capitalization. I recommend using (1)
+	``Limiting case'', (2) Space--time LSPG. Also, can we make this more specific? In particular, can
+	we write specific equations (e.g., ``$W = T$ in Eq.~(10)'') for the limiting
+	cases?}} 
 \label{fig:flowchart} 
@@
 
- The novel contributions and findings of the present work are:
+ In summary, specific contributions of this work include:
 \begin{enumerate}
 %\item The development of a time-continuous residual minimization framework for reduced-order models.
-\item The development of the \methodNameLower\ (\methodAcronym) approach for dynamical systems. The approach sequentially minimizes the time-continuous full-order model residual over time windows.
-\item The derivation of the stationary conditions for \methodAcronymROMs\ leveraging spatial and space--time dimension reduction trial subspaces.
-The case of spatial dimension reduction trial subspaces is of particular interest: the stationary conditions are derived via the Euler--Lagrange equations and comprise a coupled
-two-point Hamiltonian boundary value problem containing a forward and backward system.
-The forward system, which is forced by an auxiliary costate, evolves the generalized coordinates of the ROM.
-The backwards system, which is forced by the FOM residual evaluated about the ROM state, governs the dynamics of the costate.
-\item An exploration of the use of both direct (discretize then optimize) and indirect (optimize then discretize) solution techniques.
-Conditions under which \methodAcronym\ recovers Galerkin, LSPG, and space--time LSPG are presented.
-\item Error analysis of the \methodAcronym\ approach for trial subspaces with spatial dimension reduction.
-In particular, it is shown that the error in \methodAcronymROMs\ is bounded \textit{a priori} by a combination of the error at the start of the window and the integrated FOM ODE residual evaluated at the FOM state projected onto the trial subspace.
-\item Numerical experiments for trial subspaces with spatial dimension reduction demonstrating two key findings:
+\item The \methodNameLower\ (\methodAcronym) framework for dynamical-system
+	model reduction. The approach sequentially minimizes the time-continuous
+		full-order model residual over arbitrary time windows.
+\item Support of two space--time trial subspaces: one that reduces the spatial
+	dimension of the FOM only, and one that reduces both the spatial and
+		temporal dimensions of the FOM. The former case is of particular
+		interest in the \methodAcronym\  contex, as the stationary conditions are
+		derived via the Euler--Lagrange equations and comprise a coupled two-point
+		Hamiltonian boundary value problem containing a forward and backward
+		system. The forward system, which is forced by an auxiliary costate,
+		evolves the (spatial) generalized coordinates of the ROM in time. The
+		backward system, which is forced by the time-continuous FOM residual
+		evaluated about the ROM state, governs the dynamics of the costate. 
+	\item Derivation of two solution techniques:
+		\textit{discretize-then-optimize} (i.e., direct) and
+		\textit{optimize-then-discretize} (i.e., indirect). 
+	\item Derivations of conditions under which the
+		\methodAcronym\ framework recovers Galerkin, LSPG, and space--time LSPG
+		projection.
+\item Error analysis of the \methodAcronym\ framework using
+	trial subspaces characterized by a reduction of
+the spatial dimension of the FOM only.
+	This analysis demonstrates that the \methodAcronymROMs\ error 
+		is bounded \textit{a priori} by a combination of the
+		error at the start of the window and the integrated FOM ODE residual
+		evaluated at the FOM state projected onto the trial subspace. As a result,
+		the error bound exhibits exponential growth in the number of time windows
+		(not total time instances), yielding smaller error-bound growth in thime than
+		Galerkin or LSPG projection. \KTC{Check this}
+\item Numerical experiments for 
+	trial subspaces characterized by a reduction of
+the spatial dimension of the FOM only, which demonstrate two key findings:
 \begin{itemize}
-\item Residual minimization over larger time windows leads to more stable and physically relevant solutions as well as lower space--time residuals. 
-\item Minimizing the residual over a larger time window does not necessarily lead to a more accurate solution in the $\elltwo$ norm. Residual 
-minimization over an intermediary-sized time-window leads to the lowest error in the $\elltwo$ norm.
+\item Minimizing the residual over a larger time window leads to more stable
+	and physically relevant solutions \KTC{What is `physically relevant'?} as
+		well as smaller space--time residual norms. 
+\item Minimizing the residual over a larger time window does not necessarily
+	lead to a more accurate trajectory (as measured in the space--time
+		$\elltwo$-norm of the solution). Instead, minimizing the residual over an
+		intermediate-sized time window leads to the smallest trajectory error.
 \end{itemize}
@@
 outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
-space--time LSPG ROMs; Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
-formulation; Section~\ref{sec:numerical_techniques}
+space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
+framework. Section~\ref{sec:numerical_techniques}
 outlines numerical techniques for solving \methodAcronymROMs, including both direct and
-indirect methods; Section~\ref{sec:analysis} provides
-equivalence conditions and error analysis for \methodAcronymROMs;
-Section~\ref{sec:numerical_experiments} presents numerical experiments;
+indirect methods. Section~\ref{sec:analysis} provides
+equivalence conditions and error analysis for \methodAcronymROMs.
+Section~\ref{sec:numerical_experiments} presents numerical experiments.
 Section~\ref{sec:conclude} provides conclusions and perspectives.
-We denote vector-valued functions with stylized bold text (i.e., $\boldsymbol
-x$), vectors with standard bold text (i.e., $\mathbf{x}$), 
-matrices with capital bold letters (i.e., $\mathbf{X} \equiv \begin{bmatrix}
+We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
+x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
+matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
 \mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
-calligraphic fonts (i.e., $\mathcal{X}$).
+calligraphic symbols (e.g., $\mathcal{X}$).
 %Model reduction of nonlinear dynamical systems is a growing research area.
@@
 %\begin{comment} 
-\section{Mathematical setting}\label{sec:math}
+\section{Mathematical formulation}\label{sec:math}
+	\KTC{HERE}
+	We begin by providing the formulation for the full-order model,
+	followed by a description of standard model-reduction methods
+	classified according to the type of trial subspace they employ.
+	\subsection{Full-order model}\label{sec;FOM}
 We consider the system of nonlinear ordinary differential equations (ODEs),
@@
 
-\subsection{Projection-based reduced-order models}
+\subsection{Reduced-order models}
 Projection-based ROMs generate approximate solutions to the FOM
@@
 	Galerkin, LSPG, and space--time LSPG projection.
-	\KTC{HERE}
 
@@
 E.\ Parish acknowledges an appointment to the Sandia National Laboratories'
-John von Neumann Postdoctoral Research Fellowship in Computational Science. This paper
-describes objective technical results and analysis. Any subjective views or
-opinions that might be expressed in the paper do not necessarily represent the
-views of the U.S.\ Department of Energy or the United States Government.
-Sandia National Laboratories is a multimission laboratory managed and operated
-by National Technology and Engineering Solutions of Sandia, LLC., a wholly
-owned subsidiary of Honeywell International, Inc., for the U.S.\  Department of
-Energy's National Nuclear Security Administration under contract
-DE-NA-0003525.
+John von Neumann Postdoctoral Research Fellowship in Computational Science. 
+This work was partially sponsored by Sandia's Advanced Simulation and
+Computing (ASC) Verification and Validation (V\&V) Project/Task
+\#103723/05.30.02.  This paper describes objective technical results and
+analysis. Any subjective views or opinions that might be expressed in the
+paper do not necessarily represent the views of the U.S.\ Department of Energy
+or the United States Government.  Sandia National Laboratories is a
+multimission laboratory managed and operated by National Technology and
+Engineering Solutions of Sandia, LLC., a wholly owned subsidiary of Honeywell
+International, Inc., for the U.S.\  Department of Energy's National Nuclear
+Security Administration under contract DE-NA-0003525.
 

### 2019-10-06 main.tex
latest, working on notation

@@
 	dimension reduction demonstrate that the \methodAcronym\
-	approach can yield more accurate and physically relevant solutions \KTC{What
-	does `physically relevant' mean? This is imprecise.} than
+	approach can yield more accurate solutions that better satisfy the governing equations than
 	Galerkin and LSPG projection.
@@
 \mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
-calligraphic symbols (e.g., $\mathcal{X}$).
+calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
+function with respect to time as $\stateDot \equiv \frac{d \state}{dt}$.
 %Model reduction of nonlinear dynamical systems is a growing research area.
@@
 	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
-	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,t) \mapsto
-	\velocity(\stateyDiscrete,t)$ denotes the velocity, which is possibly
+	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDumArg) \mapsto
+	\velocity(\stateyDiscrete,\timeDumArg)$ denotes the velocity, which is possibly
 	nonlinear in its first argument. We refer to Eq.~\eqref{eq:FOM} 
@@
 \begin{equation}\label{eq:g_truncation}
-\basisspace \frac{d \genstate} {dt}(t)= \velocity(\basisspace
+\basisspace \genstateDotArg{}{t} = \velocity(\basisspace
 \genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
@@
 \begin{equation}\label{eq:g_truncation2}
-\frac{d\genstateGalerkin}{dt}(t)= \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
+\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
 \genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
@@
 \begin{equation}\label{eq:GalOptimal}
-\frac{d\genstateGalerkin}{dt}(t) = \underset{ \genstateyDiscrete \in \RR{\romdim}
+\genstateGalerkinDotArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}
 	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
@@
 %\end{equation*}
-Thus, the computed velocity $\frac{d\genstateGalerkin}{dt}(t)$ minimizes the
+Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
 time-instantaneous FOM ODE residual evaluated at the state $\basisspace
@@
 \end{equation}
-where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RR{}$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,t) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,t)$.%, $\statez: [a,b] \rightarrow \RR{M}$ and $\statezDot: [a,b] \rightarrow \RR{M}$.
+where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RR{}$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,t) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,t)$.
 We now introduce the function $\statez: [a,b] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(a) = \statez_a$.
-%To this end, we start by supposing that $\statez$ is a stationary point of $\mathcal{J}$ that satisfies the boundary conditions.
 We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$ with the boundary condition $\variationArg{a} = \bz$ and define a variation from the extremal by,
@@
 \statezBar  &: (\tau;\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
-&: \RR{} \times [a,b] \rightarrow \RR{M}.
+&: [a,b] \times \RR{} \rightarrow \RR{M}.
 \end{align*}

### 2019-10-06 main.tex
small

--- 2019-10-06 main.tex
@@
 \section{Mathematical formulation}\label{sec:math}
-	\KTC{HERE}
 	We begin by providing the formulation for the full-order model,
@@
 	\subsection{Full-order model}\label{sec;FOM}
-We consider the system of nonlinear ordinary differential equations (ODEs),
+We consider the full-order model to be a dynamical system expressed as a
+	system of ordinary differential equations (ODEs)
 \begin{equation}\label{eq:FOM}

### 2019-10-06 main.tex
fix eq 10
math thru 2.3.1

@@
 calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
-function with respect to time as $\stateDot \equiv \frac{d \state}{dt}$.
+function with respect to with the $\dot\ $ operator.
 %Model reduction of nonlinear dynamical systems is a growing research area.
@@
 \end{equation}
-where $\stateFOMDot \equiv \frac{d \stateFOM}{dt}$, 
-$T \in \mathbb{R}^+$ denotes the final time, $\stateFOM \in \RR{\fomdim}
-	\otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$
+where  $\stateFOM: [0,T]
+	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
+	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
 	denotes the state implicitly defined as the solution to initial value
 	problem~\eqref{eq:FOM}, 
-	$\timeSpace$ denotes the set of real-valued functions acting on the time
-	domain, 
+$T \in \mathbb{R}^+$ denotes the final time, 
  $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
 	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
-	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDumArg) \mapsto
-	\velocity(\stateyDiscrete,\timeDumArg)$ denotes the velocity, which is possibly
-	nonlinear in its first argument. We refer to Eq.~\eqref{eq:FOM} 
-	as the ``full-order model" (FOM) ODE.
-
-Directly solving initial value problem~\eqref{eq:FOM} is computationally
-	expensive if either the state-space dimension $\fomdim$ is large, or if the
-	time-interval length $T$ is large relative to the time step required to
-	numerically integrate Eq.~\eqref{eq:FOM}. For time-critical or many-query
-	applications, it is essential to employ an approximation in lieu of the FOM
-	ODE~\eqref{eq:FOM}. Projection-based ROMs constitute one such promising
-	approach.
+	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
+	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
+	nonlinear in its first argument. For subsequent exposition, we introduce
+	$\timeSpace$ to denote the set of real-valued functions acting on the time
+	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
+	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
+	\timeSpace $.  We refer to the initial value problem defined in
+	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
+	although the problem of interest described in the introduction corresponds
+	to a parameterized dynamical system, we suppress dependence of the FOM ODE
+	\eqref{eq:FOM} on such parameters for notational convenience, as this work
+	focuses on devising a projection applicable to a specific parameter instance.
+
+Directly solving the FOM ODE~\eqref{eq:FOM} is computationally expensive if
+	either the state-space dimension $\fomdim$ is large, or if the time-interval
+	length $T$ is large relative to the time step required to numerically
+	integrate Eq.~\eqref{eq:FOM}. For time-critical or many-query applications,
+	it is essential to relplace the FOM ODE~\eqref{eq:FOM} with a strategy that
+	enables an approximate trajectory to be computed at lower computational
+	cost. Projection-based ROMs constitute one such promising approach.
 
@@
 	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
-	subspace. For dynamical 
-	systems, two distinct types of space--time trial subspaces are commonly used: 
+	subspace. Two types of space--time trial subspaces are commonly used for
+	this purpose: 
 \begin{enumerate} 
-	\item \textit{Space–time trial subspaces with spatial dimension reduction (\spatialAcronym)}. These 
-	trial subspaces reduce only the spatial dimension of the FOM, and are commonly used
+	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
+		model (\spatialAcronym)}. These 
+	trial subspaces are commonly used
 		in model-reduction methods based on spatial projection followed by time
-		integration (e.g., Galerkin projection, LSPG projection.)
-	\item \textit{Space–time trial subspaces with spatial and temporal dimension reduction (\spaceTimeAcronym)}. These 
-	trial subspaces reduce both the spatial and temporal dimensions of the FOM,
-		and are commonly employed 
-		employed in model-reduction methods based on space--time projection (e.g.,
-		space--time LSPG projection).
+		integration (e.g., Galerkin projection, LSPG projection).
+	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
+		model (\spaceTimeAcronym)}. These trial subspaces are commonly employed in
+		model-reduction methods based on space--time projection (e.g., space--time
+		LSPG projection).
 \end{enumerate}
- We now describe these two types of trial subspaces and their application to 
-	Galerkin, LSPG, and space--time LSPG projection.
+ We now describe these two types of space--time trial subspaces and their
+	application to Galerkin, LSPG, and space--time LSPG projection.
 
@@
 \spatialAcronym\ trial subspaces approximate the FOM ODE solution
-	$\stateFOM$ with an approximation $\approxstate$ that resides in an
+	$\stateFOM(t)$, $t\in[0,T]$ with an approximation $\approxstate(t)$, $t\in[0,T]$ that resides in an
 	affine space--time trial subspace of dimension $K\ll\fomdim$, i.e., 
+	$\approxstate\in \stspaceS$, where
 \begin{equation}\label{eq:spatial_subspace}
@@
 & \stspaceS \defeq \trialspace \otimes \timeSpace +
-	\stateIntercept\otimes\onesFunction,\\ 
-& \trialspace \defeq \text{Range}(\basismat),  \\%\label{eq:spatial_subspace2}  \\
-&\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, %\label{eq:spatial_subspace3} 
+	\stateIntercept\otimes\onesFunction.\\ 
 \end{split}
 \end{equation}
-where 
+Here, $\trialspace\subseteq\RR{\fomdim}$ with $\dim(\trialspace) = K$ denotes
+the spatial trial subspace spanned by an orthogonal basis such that
+$ \trialspace= \Range{\basismat}$
+with 
+$ \basismat \equiv \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}
+	\in \RRStar{\romdim}{\fomdim}$, where $\RRStar{\romdim}{\fomdim}$ denotes the compact Stiefel manifold (i.e.,  $
+	\RRStar{\romdim}{\fomdim}\defeq
+	\{ \mathbf{X} \in \RR{\fomdim
+	\times \romdim}\, \big|\, \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$).
 %	$\timeSpace$ denotes the set of real-valued functions acting on the time
@@
 %	$[0,T]$, 
-$\onesFunction\in\timeSpace$ is defined as $\onesFunction:[0,T]\rightarrow\{1\}$,
-	$\basismat \in \RRStar{\romdim}{\fomdim}$ denotes the spatial trial subspace with
-	$\RRStar{\romdim}{\fomdim}$ denoting the compact Stiefel manifold (i.e.,  $
-	\RRStar{\romdim}{\fomdim}\defeq
-	\{ \mathbf{X} \in \RR{\fomdim
-	\times \romdim}\, \big|\, \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$),
-	$\trialspace\subseteq\RR{\fomdim}$ denotes the spatial trial subspace, and
-	$\stateIntercept \in \mathbb{R}^{\fomdim}$ denotes the reference state that
+We have also introduced $\onesFunction\in\timeSpace$ as the function $\onesFunction:\timeDummy\mapsto 1$
+and
+	$\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state, which
 	is often taken to be the initial condition (i.e., $\stateIntercept = \stateFOMIC$). 
@@
 using state snapshots, e.g., via
-proper orthogonal decomposition (POD)~\cite{berkooz_turbulence_pod} or the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}. 
+proper orthogonal decomposition (POD)~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}. 
 At any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\ trial subspace approximate the state as
@@
 \end{equation}
-where $\genstate \in \RR{\romdim} \otimes \timeSpace$ denotes the generalized
+where $\genstate \in \RR{\romdim} \otimes \timeSpace$ with
+$\genstate:\timeDummy\mapsto\genstate(\timeDummy)$
+denotes the generalized
 coordinates. Substituting the approximation~\eqref{eq:affine_trialspace} into the FOM ODE~\eqref{eq:FOM} and performing orthogonal
@@
 \basisspace \genstateDotArg{}{t} = \velocity(\basisspace
-\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
-\end{equation}
+\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC,
+	\qquad t \in [0,T],
+\end{equation}
+where $\genstateDot\equiv {d \genstate}/{d\tau}$.
 Because Eq.~\eqref{eq:g_truncation} is overdetermined, a solution may not
@@
 	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
-	\|_{\stweightingMatArg{n}}^2,
+	\|_{\stweightingMatArg{n}}^2, \qquad \genstate(0) = \genstateIC,
+	\qquad t \in [0,T],
 %\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
@@
 Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
-time-instantaneous FOM ODE residual evaluated at the state $\basisspace
-\genstateGalerkin(t) + \stateIntercept$ and time $t$ over the spatial trial
+FOM ODE residual evaluated at the state $\basisspace
+\genstateGalerkin(t) + \stateIntercept$ and time instance $t$ over the spatial trial
 subspace $\trialspace$.
@@
 %\end{equation}
-\subsubsection{Least-Squares Petrov--Galerkin projection}
+\subsubsection{Least-squares Petrov--Galerkin projection}
+\KTC{HERE}
 Despite its time-instantaneous optimality property 
@@
 \end{align*}
-with $\text{Range}(\stbasis) + \stateInterceptArg{} \equiv \stspaceST$ and where $\stdim$ is the dimension of the space--time generalized coordinates.
+with $\Range{\stbasis} + \stateInterceptArg{} \equiv \stspaceST$ and where $\stdim$ is the dimension of the space--time generalized coordinates.
 %with $\mathbb{V}_{\stdim}( \RR{N} )$ denoting the set of orthogonal $\stdim$ frames in $\RR{N}$ (e.g., an $N \times \stdim$ matrix with orthogonal column vectors).

### 2019-10-06 main.tex
update

@@
 \subsubsection{Least-squares Petrov--Galerkin projection}
-\KTC{HERE}
-Despite its time-instantaneous optimality property 
+Despite its time-instantaneous residual-minimization optimality property 
 \eqref{eq:GalOptimal}, Galerkin projection can yield inaccurate solutions,
@@
 projection~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
-is an alternative projection method that exhibits several advantages over Galerkin
+was developed
+as an alternative projection method that exhibits several advantages over Galerkin
 projection.
-Rather than minimize the (time-continuous) FOM ODE residual as in Galerkin
-projection, at each time step, LSPG minimizes
-the FOM O$\Delta$E residual, i.e., the residual arising after applying time discretization to the FOM ODE.  
+Rather than minimize the (time-continuous) FOM ODE residual at a time instance as in Galerkin
+projection, LSPG minimizes
+the (time-discrete) FOM O$\Delta$E residual 
+(i.e., the residual arising after applying time discretization to the FOM ODE)
+over a time step.
 We now describe LSPG projection in the case of linear multistep methods;
@@
 
-To begin, and without loss of generality, we introduce a uniform time
-grid characterized by the time step $\Delta t$ and time instances
+Without loss of generality, we introduce a uniform time
+grid characterized by time step $\Delta t$ and time instances
 $t^n = n\Delta
-t$, $n=1,\ldots,N_t$.
+t$, $n=0,\ldots,N_t$.
 Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
 with this time grid
-yields the FOM O$\Delta$E, which computes the sequence of FOM O$\Delta$E
+yields the FOM O$\Delta$E, which computes the sequence of discrete
 solutions
@@
 \begin{equation}\label{eq:lms}
-\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, 
+\residLMSArg{n}
+	(\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n})
+	= \bz,\qquad n=1,\ldots,N_t,
 \end{equation}
@@
 time instance and 
-$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual, which is defined as
+$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual defined as
 \begin{align*}
@@
 \end{align*} 
-Here, $\alpha^n_j$, $\beta^n_j$, $j=0,\ldots,k^n$ are coefficients
+Here, $\alpha^n_j,\beta^n_j\in\RR{}$, $j=0,\ldots,k^n$ are coefficients
 that define the linear multistep method at the $n$th time instance.
@@
 \begin{equation}\label{eq:st_fom}
-\frac{d }{dt} \stbasisArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
+\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
 \end{equation}

### 2019-10-06 main.tex
update

@@
 \subsection{\spatialAcronym\ trial subspaces}
+At a given time instance 
+$t\in[0,T]$,
 \spatialAcronym\ trial subspaces approximate the FOM ODE solution
-	$\stateFOM(t)$, $t\in[0,T]$ with an approximation $\approxstate(t)$, $t\in[0,T]$ that resides in an
-	affine space--time trial subspace of dimension $K\ll\fomdim$, i.e., 
-	$\approxstate\in \stspaceS$, where
-\begin{equation}\label{eq:spatial_subspace}
-\begin{split}
-& \stspaceS \defeq \trialspace \otimes \timeSpace +
-	\stateIntercept\otimes\onesFunction.\\ 
-\end{split}
-\end{equation}
-Here, $\trialspace\subseteq\RR{\fomdim}$ with $\dim(\trialspace) = K$ denotes
-the spatial trial subspace spanned by an orthogonal basis such that
+	as $\approxstate(t)\approx\stateFOM(t)$, which is enforced to reside in an
+	affine spatial trial subspace of dimension $K\ll\fomdim$ such that
+	$\approxstate(t)\in
+	\stateIntercept+\trialspace
+\subseteq\RR{\fomdim}$, where $\dim(\trialspace) = K$
+and $\stateIntercept \in \mathbb{R}^{\fomdim}$ denotes the reference state, which
+	is often taken to be the initial condition (i.e., $\stateIntercept = \stateFOMIC$).
+Here, the trial subspace
+$\trialspace$ 
+is spanned by an orthogonal basis such that
 $ \trialspace= \Range{\basismat}$
@@
 	\times \romdim}\, \big|\, \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$).
-%	$\timeSpace$ denotes the set of real-valued functions acting on the time
-%	domain
-%	$[0,T]$, 
-We have also introduced $\onesFunction\in\timeSpace$ as the function $\onesFunction:\timeDummy\mapsto 1$
-and
-	$\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state, which
-	is often taken to be the initial condition (i.e., $\stateIntercept = \stateFOMIC$). 
-%where $\basisvec_i \in \mathbb{R}^N$ are orthonormal basis vectors and $\stateIntercept \in \mathbb{R}^{\fomdim}$ the ``intercept''.
-%We refer to $\basismat \in \RRStar{\fomdim \times \romdim}$, with $\RRStar{p \times q}$ denoting the set of orthogonal $q$ frames in $p$, as the spatial trial basis, and $\trialspace$ as the spatial trial space.
 The basis vectors $\basisvec_i$, $i=1,\ldots,K$ are typically constructed
@@
 proper orthogonal decomposition (POD)~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}. 
-At any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\ trial subspace approximate the state as
+Thus, at any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\ trial subspace approximate the state as
 \begin{equation}\label{eq:affine_trialspace}
@@
 denotes the generalized
-coordinates. Substituting the approximation~\eqref{eq:affine_trialspace} into the FOM ODE~\eqref{eq:FOM} and performing orthogonal
+coordinates.
+
+	From the space--time perspective, this is equivalent to approximating the
+	FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$ with 
+	$\approxstate\in \stspaceS$, where
+\begin{equation}\label{eq:spatial_subspace}
+\begin{split}
+& \stspaceS \defeq \trialspace \otimes \timeSpace +
+	\stateIntercept\otimes\onesFunction\subseteq\RR{N}\otimes\timeSpace
+\end{split}
+\end{equation}
+with
+	$\dim(\stspaceS) = K\ntimeSteps$
+and $\onesFunction\in\timeSpace$ defined as
+$\onesFunction:\timeDummy\mapsto 1$.
+
+Substituting the approximation~\eqref{eq:affine_trialspace} into the FOM ODE~\eqref{eq:FOM} and performing orthogonal
 $\elltwo$-projection of the initial condition onto the trial subspace yields
@@
 
-\KTC{HERE}
-The LSPG method substitutes the \spatialAcronym\ trial subspace approximation
-minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial subspace over each time step. %We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
-%\begin{align*}
-%\resid_{\text{IE}} & \vcentcolon (\stateyDiscrete;\stateFOMDiscreteArg{n-1}) \mapsto \frac{\stateyDiscrete - \stateFOMDiscreteArg{n-1}  }{\Delta t} -  \velocity(\stateyDiscrete), \\
-%& \vcentcolon \RR{\fomdim} \times \RR{\fomdim} \rightarrow \RR{\fomdim}. 
-%\end{align*}
-The LSPG method is defined as follows,
+At each time instance on the time grid, the LSPG method substitutes the \spatialAcronym\ trial subspace approximation
+\eqref{eq:affine_trialspace} into the FOM O$\Delta$E~\eqref{eq:lms} and
+minimizes the residual, i.e., LSPG sequentially computes the solutions
+$\approxstateLSPG^n \approx \stateFOMDiscreteArg{n}$, $n=1,\ldots,
+\ntimeSteps$ that satisfy
 \begin{equation*}
-\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,2,\hdots,\ntimeSteps,
+\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2, \qquad n = 1,\ldots,\ntimeSteps,
 \end{equation*}
-where $\approxstateLSPG^n \approx \stateFOMDiscreteArg{n}$ denotes the LSPG approximation to the state at $n$th time-step instance and 
-$\lspgWeightingArg{\cdot} \in \RR{\nsamples \times \fomdim}$, with $\romdim \le \nsamples \le \fomdim$, is a weighting matrix that can be used, e.g., for hyper-reduction. 
-Although illustrated for linear multistep methods, LSPG can be 
-formulated for \textit{any} type of time-marching scheme (see Ref.~\cite{carlberg_lspg_v_galerkin} for, e.g., Runge--Kutta schemes).
-
-In general, the LSPG approach has been observed to be more stable and accurate than the 
-Galerkin approach~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat}.
-The common interpretation of the improved performance of the LSPG method over the Galerkin method is that LSPG minimizes the residual over a finite-sized time window, while Galerkin minimizes only the instantaneous residual~\cite{carlberg_lspg_v_galerkin}.
-However, for general non-self-adjoint and nonlinear problems, LSPG, lacks \textit{a priori} guarantees of stability, accuracy, and convergence. Further, 
-LSPG is sensitive to the time step: changing the time step changes the window over which the residual is minimized 
-\textit{as well as} the temporal discretization error. As a result, optimal results are obtained at an intermediary time-step. 
-For too small a time-step, the accuracy and stability of LSPG is similar to Galerkin; in the 
-limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin~\cite{carlberg_lspg_v_galerkin}. 
-For too large a time-step, the accuracy of LSPG degrades~\cite{carlberg_lspg_v_galerkin,parish_apg}. The 
-precise reason for this degradation in accuracy is difficult to discern as increasing the time-step changes both the error incurred by 
-temporal discretization and the window the residual is being minimized over. Lastly, LSPG is sensitive to the time-marching scheme: changing the 
-time-marching scheme modifies the minimization problem. For example, LSPG reverts to a Galerkin approach 
-when used with an explicit time-marching scheme. It is emphasized that, despite these challenges, 
-LSPG is arguably the most popular reduced-order modeling technique for general nonlinear systems.
+where 
+$\lspgWeightingArg{\cdot} \in \RR{\nsamples \times \fomdim}$, with $\romdim
+\le \nsamples \le \fomdim$, is a weighting matrix that can be used, e.g., to
+enable hyper-reduction by requiring it to have a small number of nonzero
+columns.
+
+As described in the introduction, although numerical experiments have
+demonstrated that LSPG projection often yields more accurate and stable
+solutions than Galerkin
+projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg},
+LSPG still suffers from several shortcomings, in particular its complex
+dependence on the time discretization, exponentially growing error bounds, and
+a lack of optimality for the trajectory defined over the entire time domain.
 
 \subsection{\spaceTimeAcronym\ trial spaces and space--time ROMs}
-Recent work has investigated space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}. Space--time methods 
-are advantageous in that they allow for both spatial \textit{and} temporal dimension reduction. Further, space--time 
-methods are often equipped with more favorable error bounds than non-space--time methods. The approach proposed in this work displays commonalities 
-with space--time methods and, as such, we briefly outline \spaceTimeAcronym\ trial subspaces and space--time LSPG.
-
-Space--time methods leverage the \spaceTimeAcronym\ trial subspace $\stspaceST$,
+
+Space--time projection methods that employ \spaceTimeAcronym\ trial
+spaces~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}
+aim to overcome the latter two shortcomings of LSPG projection, which are
+shared by all model-reduction methods based on spatial projection followed by
+time integration.  Because these methods employ \spaceTimeAcronym\ trial
+spaces, they reduce both the spatial and temporal dimensions of the full-order
+model; further, they yield error bounds that grow more slowly in time and
+their trajectories exhibit an optimality property over the entire time domain.
+
+\spaceTimeAcronym\ trial subspaces approximate the FOM ODE solution
+trajectory
+	$\stateFOM\in\RR{N}\otimes \timeSpace$ with an approximation that resides in an
+	affine space--time trial subspace of dimension $\stdim\ll\fomdim$, i.e., 
+	$\approxstate\in \stspaceST$ with $\dim(\stspaceST) =\stdim $, where
 \begin{equation}\label{eq:sttrialspace_def}
- \approxstateArgnt{} \in  \stspaceST \subseteq \RR{N} \otimes \timeSpace,
-\end{equation}
-and the space--time trial basis \textit{functions},
-\begin{align*}
-\stbasisArgnt{} &: t \mapsto \stbasis(t) \\
- &: [0,T] \rightarrow  \RR{N \times \stdim} ,%\spaceTrialSpace , \\
-%&  \in \RR{N \times \stdim} \otimes \timeTrialSpaceArg{},  
-\end{align*}
-with $\Range{\stbasis} + \stateInterceptArg{} \equiv \stspaceST$ and where $\stdim$ is the dimension of the space--time generalized coordinates.
-%with $\mathbb{V}_{\stdim}( \RR{N} )$ denoting the set of orthogonal $\stdim$ frames in $\RR{N}$ (e.g., an $N \times \stdim$ matrix with orthogonal column vectors).
-The state is approximated by,
+ \stspaceST \defeq 
+	\Range{\stbasis} + 
+	\stateIntercept\otimes\onesFunction
+	\subseteq \RR{N} \otimes \timeSpace
+\end{equation}
+and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
+$\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$
+denotes the space--time trial basis.
+
+Thus, at any time instance $t\in[0,T]$, ROMs that employ the
+\spaceTimeAcronym\ trial subspace approximate the state as
 \begin{equation}\label{eq:stapprox1}
@@
 \end{equation}
-%where $\stbasisArgnt{} : \RR{\stdim} \rightarrow \spaceTrialSpace \otimes \timeTrialSpaceArg{}$ with $ \stbasisArgnt{n} \in \RR{N \times \stdim} \otimes \timeTrialSpaceArg{n}$ is the space--time trial basis,
-where $\stgenstateArg{} \in \RR{\stdim}$ are the space--time generalized coordinates. Critically, the 
-space--time basis matrix comprises time-dependent \textit{functions}, while the generalized coordinates are a stationary vector. This is fundamentally different than  
-\spatialAcronym\ trial subspaces, in where the generalized coordinates are time-dependent functions while the basis matrix is stationary.
-
-Substituting the approximation~\eqref{eq:stapprox1} into the FOM ODE~\eqref{eq:FOM},
+where $\stgenstateArg{} \in \RR{\stdim}$ denote the space--time generalized coordinates. 
+Critically, comparing the approximations arising from \spatialAcronym\ and
+\spaceTimeAcronym\ subspaces in Eqs.~\eqref{eq:affine_trialspace} and \eqref{eq:stapprox1}, respectively,
+highlights that the former approximation associates with a time-dependent
+generalized coordinates, while the latter approximation associates with a
+time-dependent basis matrix.
+
+\KTC{HERE}
+Substituting the approximation~\eqref{eq:stapprox1} into the FOM
+ODE~\eqref{eq:FOM} yields
 \begin{equation}\label{eq:st_fom}
@@
 \end{equation}
-where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
+where 
+$\stbasisDot\equiv d\stbasis/d\timeDummy$,
+$\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
 Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 

### 2019-10-06 main.tex
eq 10
fix 10
updates

@@
 
-\KTC{HERE}
 Substituting the approximation~\eqref{eq:stapprox1} into the FOM
@@
 \begin{equation}\label{eq:st_fom}
-\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
+\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
+	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM(0))(0), \qquad t \in [0,T],
 \end{equation}
 where 
-$\stbasisDot\equiv d\stbasis/d\timeDummy$,
-$\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
-Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
-The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
-principle, however, we do not outline it here.
+$\stbasisDot\equiv d\stbasis/d\timeDummy$ and
+$\mathbb{P} : \RR{N} \rightarrow \stspaceST$ denotes the
+orthogonal $\ell^2$-projection onto the space--time trial subspace, i.e.,
+$$
+\mathbb{P}:\stateyDiscrete\mapsto\underset{\statey\in\stspaceST}{\text{arg\,min}}\int_0^T\|\stateyDiscrete
+- \statey( t)\|_2^2d t.
+$$
+Space--time projection methods reduce the number of equations in
+\eqref{eq:st_fom} to ensure a unique solution exists.
+We now outline one such method: space--time least-squares Petrov--Galerkin (ST-LSPG)
+projection~\cite{choi_stlspg}. 
+While space--time Galerkin projection \KTC{add citations. See our ST-LSPG work
+where we have some of these} is another alternative, it does not
+associate with any residual minimization principle, and thus we do not discuss
+it further.
 \subsubsection{Space--Time LSPG projection}  
-As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain.
-For brevity, ST-LSPG is outlined only for linear multistep methods.
-%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
-%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
-%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
-%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$.
-We define now the space--time discrete residual for the implicit Euler method as,
+Analogously to LSPG projection, space--time LSPG projection~\cite{choi_stlspg}
+minimizes the (time-discrete) FOM O$\Delta$E residual, but does so over the
+entire time domain. 
 We first define the space--time discrete residual for linear multistep methods,

### 2019-10-06 main.tex
update

@@
 \stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
-	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM(0))(0), \qquad t \in [0,T],
+	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOMIC)(0), \qquad t \in [0,T],
 \end{equation}
@@
 it further.
-\subsubsection{Space--Time LSPG projection}  
+\subsubsection{Space--time LSPG projection}  
 Analogously to LSPG projection, space--time LSPG projection~\cite{choi_stlspg}
-minimizes the (time-discrete) FOM O$\Delta$E residual, but does so over the
-entire time domain. 
-We first define the space--time discrete residual for linear multistep methods,
-\begin{align*}
-\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
+minimizes the (time-discrete) FOM O$\Delta$E residual, but does so
+using the \spaceTimeAcronym\ subspace and simultaneously minimizes this
+residual over all $N_t$ time instances.
+We first introduce the full space--time FOM O$\Delta$E residual for linear
+multistep methods as
+\begin{align*}
+\residST_{} & \vcentcolon (\stateyDiscreteArg{1}, \ldots
+	,\stateyDiscreteArg{N_t};\stateFOMIC) \mapsto \begin{bmatrix}
+\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
 %\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
-\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
+\\ 
 %\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
@@
 %\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
-\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
-\end{bmatrix}  \\
-& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
-\end{align*}
-For convenience we define the same residual with the generalized coordinates as the inputs,
+\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \end{bmatrix}  \\
+& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t},
+\end{align*}
+and define the counterpart function acting on space--time generalized
+coordinates associated with the \spaceTimeAcronym\ subspace:
 \begin{align*}
@@
 %\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
-\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\
+\residLMSArg{1}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg)
+\\ 
 \vdots \\
-\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
+\residLMSArg{N_t}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg)  \\
 \end{bmatrix}  \\

### 2019-10-06 main.tex
working on EL

--- 2019-10-06 main.tex
@@
 \begin{equation}\label{eq:st_fom}
-\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
+\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
+	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM(0)\otimes
+	\onesFunction)(0), \qquad t \in [0,T],
 \end{equation}
@@
 \begin{align*}
-\statezBar  &: (\tau;\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
+\statezBar  &: (\tau,\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
 &: [a,b] \times \RR{} \rightarrow \RR{M}.
@@
 We note that $\statezBar$ obeys the same boundary conditions as $\statez$ since $\variationArg{a} = \bz$.
-Evaluating the functional~\eqref{eq:el_gen} at $\statezBar(t,\epsilon)$,
-$$
-\mathcal{J}(\epsilon) = \int_a^b \integrand(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t)  dt.
-$$
-The objective is to now find $\overline{\statez}$ that makes $\mathcal{J}(\epsilon)$ stationary.
-This can be done by differentiating with respect to $\epsilon$, and setting the result to zero,% Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
-$$
-\frac{d}{d\epsilon} \bigg[ \mathcal{J}\big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) \big)\bigg] =  0,
-$$
-or equivalently,
-$$
-\int_a^b \frac{d}{d\epsilon} \bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big)\bigg] dt=  0.
-$$
+We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar(t,\cdot)$,
+$$
+\mathcal{J}_{\delta}: \delta \mapsto \int_a^b \integrand(\statezBar(t,\delta),\statezDotBar(t,\delta) ,t)  dt.
+$$
+
+%$$
+%\mathcal{J}(\epsilon) = \int_a^b \integrand(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t)  dt.
+%$$
+The objective is to now find $\overline{\statez}$ that makes $\mathcal{J}_{\delta}$ stationary.
+This can be done by differentiating with respect to $\delta$, and setting the result to zero,% Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
+$$
+\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)=  0.
+$$
+%or equivalently,
+%$$
+% \int_a^b \frac{d}{d \delta} \big(\integrand(\statezBar(t,\delta),\statezDotBar(t,\delta) ,t)\big)  dt = 0
+%$$
+%\int_a^b \frac{d}{d\epsilon} \bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big)\bigg] dt=  0.
+%$$
 %As $\statez$ is a stationary point of $\mathcal{J}$, it then directly follows that,
@@
 We first evaluate the derivative. First using the chain rule,
-$$ \int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t\big) \bigg] dt = \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t;\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t;\epsilon)\bigg]dt .
+$$ 
+\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)= 
+\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt .
 $$
@@
 Noting that,
-$$\frac{\partial \statezBar}{\partial \delta} (\cdot,t)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(\cdot,t)= \variationDot(t),$$
+$$\frac{\partial \statezBar}{\partial \delta} (t,\cdot)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(t,\cdot)= \variationDot(t),$$
 we have,
-$$\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt = \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variationDot(t) \bigg]dt.
+$$
+%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
+\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
+= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variationDot(t) \bigg]dt.
 $$
@@
 \begin{multline*}
-\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt =  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon) ,b \big) \variation(b)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(a;\epsilon),\statezDotBar(a;\epsilon) ,a\big) \variation(a) .
+\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
+%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
+=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(a,\epsilon),\statezDotBar(a,\epsilon) ,a\big) \variation(a) .
 \end{multline*}
+\EP{Here}
 Noting that $\variationArg{a} = \bz$,

### 2019-10-06 main.tex

@@
 \stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
-	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM(0)\otimes
-	\onesFunction)(0), \qquad t \in [0,T],
+	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOMIC)(0), \qquad t \in [0,T],
 \end{equation}
 where 
-$\stbasisDot\equiv d\stbasis/d\timeDummy$,
-$\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
-Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
-The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
-principle, however, we do not outline it here.
-\subsubsection{Space--Time LSPG projection}  
-As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain.
-For brevity, ST-LSPG is outlined only for linear multistep methods.
-%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
-%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
-%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
-%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$.
-We define now the space--time discrete residual for the implicit Euler method as,
-We first define the space--time discrete residual for linear multistep methods,
-\begin{align*}
-\residST_{} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
+$\stbasisDot\equiv d\stbasis/d\timeDummy$ and
+$\mathbb{P} : \RR{N} \rightarrow \stspaceST$ denotes the
+orthogonal $\ell^2$-projection onto the space--time trial subspace, i.e.,
+$$
+\mathbb{P}:\stateyDiscrete\mapsto\underset{\statey\in\stspaceST}{\text{arg\,min}}\int_0^T\|\stateyDiscrete
+- \statey( t)\|_2^2d t.
+$$
+Space--time projection methods reduce the number of equations in
+\eqref{eq:st_fom} to ensure a unique solution exists.
+We now outline one such method: space--time least-squares Petrov--Galerkin (ST-LSPG)
+projection~\cite{choi_stlspg}. 
+While space--time Galerkin projection \KTC{add citations. See our ST-LSPG work
+where we have some of these} is another alternative, it does not
+associate with any residual minimization principle, and thus we do not discuss
+it further.
+\subsubsection{Space--time LSPG projection}  
+Analogously to LSPG projection, space--time LSPG projection~\cite{choi_stlspg}
+minimizes the (time-discrete) FOM O$\Delta$E residual, but does so
+using the \spaceTimeAcronym\ subspace and simultaneously minimizes this
+residual over all $N_t$ time instances.
+We first introduce the full space--time FOM O$\Delta$E residual for linear
+multistep methods as
+\begin{align*}
+\residST_{} & \vcentcolon (\stateyDiscreteArg{1}, \ldots
+	,\stateyDiscreteArg{N_t};\stateFOMIC) \mapsto \begin{bmatrix}
+\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
 %\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
-\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
+\\ 
 %\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\

### 2019-10-06 main.tex
eq 10
update
up
some fixes

--- 2019-10-06 main.tex
@@
 	\Range{\stbasis} + 
-	\stateIntercept\otimes\onesFunction
+	\stateFOMIC\otimes\onesFunction
 	\subseteq \RR{N} \otimes \timeSpace
@@
 and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
-$\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$
+$\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ and $\stbasis(0) =
+\boldsymbol 0$ 
 denotes the space--time trial basis.
-
 Thus, at any time instance $t\in[0,T]$, ROMs that employ the
@@
 \begin{equation}\label{eq:stapprox1}
- \stateFOMArg{}{t} \approx \approxstateArg{}{t}  = \stbasisArg{}{t} \stgenstate + \stateInterceptArg{},
+ \stateFOMArg{}{t} \approx \approxstateArg{}{t}  = \stbasisArg{}{t} \stgenstate + \stateFOMIC,
 \end{equation}
@@
 \stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
-	\stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOMIC)(0), \qquad t \in [0,T],
+	\stateFOMIC,t) , \qquad t \in [0,T],
 \end{equation}
 where 
-$\stbasisDot\equiv d\stbasis/d\timeDummy$ and
-$\mathbb{P} : \RR{N} \rightarrow \stspaceST$ denotes the
-orthogonal $\ell^2$-projection onto the space--time trial subspace, i.e.,
-$$
-\mathbb{P}:\stateyDiscrete\mapsto\underset{\statey\in\stspaceST}{\text{arg\,min}}\int_0^T\|\stateyDiscrete
-- \statey( t)\|_2^2d t.
-$$
+$\stbasisDot\equiv d\stbasis/d\timeDummy$. We note that the initial conditions
+are automatically satisfied from the definition of the \spaceTimeAcronym\
+subspace.
+
 Space--time projection methods reduce the number of equations in
@@
 projection~\cite{choi_stlspg}. 
-While space--time Galerkin projection \KTC{add citations. See our ST-LSPG work
-where we have some of these} is another alternative, it does not
+While space--time Galerkin projection \cite{benner_st} is another alternative, it does not
 associate with any residual minimization principle, and thus we do not discuss
@@
 and define the counterpart function acting on space--time generalized
-coordinates associated with the \spaceTimeAcronym\ subspace:
+coordinates:
 \begin{align*}

### 2019-10-06 main.tex
small

@@
 \end{align*}
-ST-LSPG is defined as,
-\begin{equation*}
-\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
-\end{equation*}
-where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.
-
-ST-LSPG is appealing in that it (1) allows for both spatial and temporal dimension reduction and (2) minimizes the residual over the \textit{entire} space--time domain (as opposed to a single time-step instance as in LSPG).
-As a result, ST-LSPG can potentially be more efficient than standard LSPG and is equipped with more favorable error bounds~\cite{choi_stlspg}.
-ST-LSPG, however, is subject to several practical challenges.
-Most importantly, ST-LSPG can be computationally
-intensive as it involves (1.) computing space--time basis vectors and (2.) minimizing the entire space--time residual.
-For high-dimensional systems, these tasks may be intractable.
-Another challenge associated with ST-LSPG is that, as it requires a pre-defined temporal basis, it is unclear how it can be used, e.g., to perform future state prediction.
+ST-LSPG projection computes the space--time generalized coordinates that
+minimize the space--time FOM O$\Delta$ E residual:
+\begin{equation}\label{eq:stlspg}
+	\stgenstate_\text{ST-LSPG} = \underset{ \stgenstatey\in\RR{\stdim} }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
+\end{equation}
+where $\lspgWeightingSTArg{\cdot} \in \RR{\stdim \times N N_t}$, with $\stdim
+\le n_{st} \le N N_t$, is a space--time weighting matrix that can be chosen,
+e.g., to enable hyper-reduction.
+
+ST-LSPG overcomes two of the primary shortcomings of LSPG. In particular, it
+leads to error bounds that grow sub-quadratically in time rather than
+exponentially in time, and it generates entire trajectories that associate
+with an optimality property over the entire time domain \cite{choi_stlspg}.
+However, it is subject to several challenges. First, the computational cost of
+solving Eq.~\eqref{eq:stlspg} scales cubically with the number of space--time
+degrees of freedom $\stdim$, because---by performing projection in time---it
+leads to dense systems that do not expose any natural mechanism for exploiting
+the sequential nature of time evolution. Second, it is unclear how these
+methods can be employed for future state prediction, as the space--time trial
+basis $\stbasis$ must be defined over the entire time interval of interest
+$[0,T]$. Third, ST-LSPG is still strongly tied to the time discretization
+employed for the full-order model, as it minimizes the (time-discrete) FOM
+O$\Delta$E residual over all time instance.
 
 \subsection{Outstanding challenges}
-This work seeks to overcome the limitations of existing reduced-order modeling approaches. Specifically, 
-we look to overcome LSPG's time-step and time-scheme dependence, the stability challenges of both Galerkin and LSPG, and the potential computational cost of ST-LSPG.
-To this end, we now outline the \methodNameLower\ formulation.
+This work seeks to overcome the limitations of existing residual-minimizing
+model-reduction methods, and to provide a unifying framework from which
+exisiting methods can be assessed. Specifically, we look to overcome the
+complex dependence of LSPG projection on the time discretization,
+exponential time growth of the error bounds for Galerkin and LSPG projection,
+the cubic dependence of the computational cost of ST-LSPG on the number of
+degrees of freedom, and the lack of ability for ST-LSPG to perform predictions
+in time. We now describe the proposed \methodNameLower\ framework for this
+purpose. 
 %\end{comment}

### 2019-10-06 section3/general_form.tex
update

--- 2019-09-19 section3/general_form.tex
@@
-\section{Windowed Least-Squares Formulation}\label{sec:tclspg} This section
+\section{Windowed least-squares framework}\label{sec:tclspg} 
+This section
 outlines the proposed \methodNameLower\ 
-formulation (\methodAcronym). In contrast to (1.) the Galerkin method, which minimizes the 
-time-instantaneous residual, (2.) the LSPG method, which sequentially minimizes the discrete 
-residual over a single time step, and (3.) the ST-LSPG method, which minimizes the discrete 
-space--time residual, the proposed \methodAcronym\ method sequentially minimizes the 
-time-continuous FOM ODE residual over a series of
+(\methodAcronym) framework. In contrast to (1) Galerkin projection, 
+which minimizes (time-continuous) FOM ODE residual at a time instance, 
+(2) LSPG projection, 
+which minimizes
+the (time-discrete) FOM O$\Delta$E residual 
+over a time step, and (3) ST-LSPG projection, 
+which minimizes
+the  FOM O$\Delta$E residual 
+over the entire time interval, the proposed \methodAcronym\ method sequentially minimizes the 
+time-continuous FOM ODE residual over a series of arbitrarily defined
 \textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
 In this section, we start by outlining the \methodAcronym\ formulation  
-for a general trial subspace. We then examine the \spatialAcronym\ 
+for a general space--time trial subspace. We then examine the \spatialAcronym\ 
 trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
-associated with the minimization problems.
+associated with the associated residual-minimization problems.
  %The proposed method can be formulated equivalently from
@@
 %
-\subsection{Residual Minimization Formulation} 
-We start by partitioning the
-temporal domain into a set of $\nslabs$ windows of size $\DeltaSlabArg{n}$. This partitioning is depicted in Fig.~\ref{fig:slab_fig}.
+\subsection{Windowed least-squares for general space--time trial subspaces} 
+We begin by introducing a (nonuniform) partition of the time domain $[0,T]$
+into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
+\timeEndArg{n}]\subseteq[0,T]$ of size $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
+\timeStartArg{n}$ such that $\timeStartArg{n+1} = \timeEndArg{n}$,
+$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts this partitioning.
 \begin{figure} 
@@
 \end{figure}
-Over each time window, we introduce the space--time trial subspace,
-$$\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,$$ 
-where $\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
-\RR{\fomdim}$
-is the \methodAcronym\ approximation to the state over the $n$th time window, $\timeSpaceArg{n}$ is the set of functions acting on $[\timeStartArg{n},\timeEndArg{n}]$, and  
-$\timeStartArg{n}$ and $\timeEndArg{n}$ are the beginning and end of the $n$th
-time window, respectively. Over each window, define the objective,
-%\begin{equation}\label{eq:obj} \mathcal{J}^n(\stateArg{}{t}) = \frac{1}{2}
-%\int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\state}(\tau) -
-%\velocity(\stateArg{}{\tau}) \big]^T \stweightingMatArg{n}(\tau) \big[
-%\dot{\state}(\tau) - \velocity(\stateArg{}{\tau}) \big] d \tau,
-%\end{equation}
+Over the $n$th time window, we approximate the FOM ODE solution as 
+$\approxstate^n(t)\approx \stateFOM(t)$,
+$t\in[\timeStartArg{n},\timeEndArg{n}]$, which is enforced to reside in 
+the $n$th space--time trial subspace such that
+\begin{equation}
+\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,
+\end{equation}
+where $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on $[\timeStartArg{n},\timeEndArg{n}]$
+and
+$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. 
+\KTC{HERE}
+Over each window, define the objective,
 \begin{equation}\label{eq:obj}

### 2019-10-06 section3/general_form.tex
update

@@
 into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
-\timeEndArg{n}]\subseteq[0,T]$ of size $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
-\timeStartArg{n}$ such that $\timeStartArg{n+1} = \timeEndArg{n}$,
+\timeEndArg{n}]\subseteq[0,T]$ of length $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
+\timeStartArg{n}$, $n=1,\ldots,\nslabs$ such that 
+$\timeStartArg{1} = 0$, $\timeEndArg{\nslabs} = T$, and
+$\timeStartArg{n+1} = \timeEndArg{n}$,
 $n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts this partitioning.
@@
 \end{equation}
-where $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on $[\timeStartArg{n},\timeEndArg{n}]$
+where $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
+$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} = \{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$)
 and
 $\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$.
+
 \KTC{HERE}
-Over each window, define the objective,
+Over the $n$th window, define the objective functional
 \begin{equation}\label{eq:obj}
@@
 &\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
-\RR{}, 
+\RR{}_+, 
 \end{split}
@@
 %[0,T] \rightarrow \RR{}.  \end{align}
-where $\stweightingMatArg{n} \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ } \in \RR{\fomdim \times \fomdim}$ is a
-symmetric positive semi-definite matrix that, e.g., enables hyper-reduction and/or modifies the norm. For simplicity, 
-it is assumed that the weighting matrix is the same for all time windows.
+where $\stweightingMatArg{n} \equiv \stweightingMatOneArg{ }^T
+\stweightingMatOneArg{ } \in \RR{\fomdim \times \fomdim}$ denotes a
+symmetric positive semi-definite matrix that can enable hyper-reduction, for
+example. For notational simplicity, 
+we employ the same weighting matrix for all time windows.
 
-\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+The \methodAcronym\ framework sequentially computes approximate solutions
+$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
+$\approxstateArgnt{n}$ is the solution to the
+minimization problem
 \begin{equation}\label{eq:tclsrm}
 \begin{split}
-      &\underset{\stateyArgnt{n} \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\stateyArgnt{n}), \\
-      &\text{subject to } \;  \stateyArg{n}{\timeStartArg{n}} =
+      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
+			&\text{subject to } \;  \statey(\timeStartArg{n}) =
 \begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\

### 2019-10-07 section3/general_form.tex

@@
 			&\text{subject to } \;  \statey(\timeStartArg{n}) =
-\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
-\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
+\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs \\
+\mathbb{P}^n( \stateFOMIC)& n=1, \end{cases} 
 \end{split}
@@
 %\end{equation}
-where $\mathbb{P}^n: \RR{N} \otimes \timeSpaceArg{n} \rightarrow \stspaceArg{n}$ is an orthogonal projection onto the 
-trial subspace (e.g., $\elltwo$).
+where 
+$\mathbb{P}^n (\statey) = XXX$ \KTC{EP TODO} for \spatialAcronym\ subspaces
+and
+$\mathbb{P}^n (\statey) = \statey$ for \spaceTimeAcronym\ subspaces.
 %\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing
@@
 %trial space (e.g., an $L^2$ projector).
+	\KTC{Remove next remark?}
 \begin{remark}

### 2019-10-07 section4/numerical_methods.tex
thru p 15

@@
 The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
-As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged.
-This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system.
-We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces,
+\KTC{Passive voice:} As \methodAcronym\ displays commonalities with optimal
+control problems, a large existing body of work can be leveraged. This section
+discuses a variety of such techniques and outlines how they can be used to
+solve the \methodAcronym\ system. We focus primarily on \KTC{earlier we kept
+saying 
+direct (i.e.,
+	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
+	Let's be consistent when we re-introduce it here.}
+optimize direct and indirect methods for \spatialAcronym\ trial subspaces, 
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
@@
 \textit{direct} or \textit{indirect}
-methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
+methods~\cite{conway_optimalcontrolreview}.  Direct \KTC{Typo:} method eschew
+the calculus of variations \KTC{weird to say they `eschew' calculus. Find
+something better here}
 and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
@@
 Indirect methods, on the other hand, leverage the calculus of variations to
-derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
+derive the Euler--Lagrange equations, which comprise the first-order
+optimality conditions\KTC{use endash}~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
 functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
 solve the optimization problem
-``indirectly". For both direct and indirect methods, a variety of
+``indirectly". \KTC{Passive voice:} For both direct and indirect methods, a variety of
 discretization/solution approaches are then possible. Collocation methods,
@@
 
-In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial subspaces. We consider:
+In the present context, both direct and indirect methods \KTC{Passive voice:}
+are investigated to solve \methodAcronym\ with \spatialAcronym\ trial
+subspaces. We consider: \KTC{Weird capitalization and bulleting. Make
+consistent with earlier in the paper, e.g., introduction, and don't break into
+sub-bullet, looks ugly.}
 
@@
 \end{itemize} 
-It
+\KTC{Passive voice:} It
 is emphasized that a variety of other approaches exist. A detailed exploration
-of additional approaches will be a subject of future work. Section~\ref{sec:direct} 
-will outline direct solution approaches based on linear multistep methods, 
-while Section~\ref{sec:indirect} will outline the indirect solution approach.
+of additional approaches is the subject of future work.
+Section~\ref{sec:direct}  outlines direct solution approaches based on linear
+multistep methods, while Section~\ref{sec:indirect} outlines the indirect
+solution approach.
 

### 2019-10-07 section4/spatial_numerical.tex
thru p 15

@@
 \subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
+\KTC{Does the time grid definted in this section have any relationship to the
+time grid used for time discretization? Haven't gotten past HERE below, but
+wanted to check (maybe it's made clear later).}
 Direct approaches solve~\eqref{eq:tclsrm} by
@@
 dimensional one by discretizing the state and objective function in time.
-The minimization problem is then reformulated as a nonlinear programming, or
+\KTC{Passive voice:} The minimization problem is then reformulated as a nonlinear programming, or
 optimization, problem. A variety of direct solution approaches exist, 
-including collocation approaches, orthogonal polynomial (spectral  method) 
+including collocation approaches, orthogonal polynomial \KTC{is this a
+spectral method, or the spectral method? clarify} (spectral  method) 
 approaches, and genetic algorithms.
@@
 In the present context, the most straightforward direct solution approach consists of the 
-following steps:
+following steps: \KTC{Keep this inline. Don't see a benefit to multi-lining a
+two-step algorithm}
 \begin{enumerate}
@@
 \end{enumerate}
-To achieve this agenda, first define a discrete temporal grid 
-by dividing each time window $[\timeStartArg{n},\timeEndArg{n}]$ into
+To achieve this agenda \KTC{agenda? odd way to say it.}, first define a discrete temporal grid 
+by \KTC{you can't devide (i.e., decompose) a window into instances. you can divide a window into
+sub-windows. you can also simply define a set of time instances in the window.
+no need to emphasize `strictly increasing', as any set of points in the window
+will by necessity be ordered.} dividing each time window $[\timeStartArg{n},\timeEndArg{n}]$ into
 $\nstepsArg{n}$ strictly increasing time-step instances described by
-$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with  
+$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with  \KTC{use cdots, not
+hdots or ldots. Don't use $\defeq$ for the first and last times; these are
+just equalities, i.e., you happen to set them equal to those values, but they
+aren't identically the same:}
 \begin{align}
@@
 \end{align}
-This discretization is depicted in Figure~\ref{fig:slab_fig2}.
+\KTC{Passive voice:} This discretization is depicted in Figure~\ref{fig:slab_fig2}.
 We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
@@
 \includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
-\caption{Decomposition of time windows into time-step instances.} 
+	\caption{Decomposition of time windows into time-step instances. \KTC{You
+	can't decompose a time interval into time instances. You can define time
+	instances on that interval.} \KTC{Don't call them `time-step instances'.
+	This woudl be an instance of a time step. These are `time instances',
+	correct? If so, fix this terminology everywhere.}} 
 \label{fig:slab_fig2} 
@@
 \subsubsection{Linear multistep schemes}
+\KTC{HERE}
 Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time-instances.  

### 2019-10-07 section3/general_form.tex

@@
 
-\KTC{HERE}
 Over the $n$th window, define the objective functional
@@
 where 
-$\mathbb{P}^n (\statey) = XXX$ \KTC{EP TODO} for \spatialAcronym\ subspaces
+$\mathbb{P}^n (\stateyDiscrete) = XXX$ \KTC{EP TODO. Not trivila because
+haven't introduced how you're going to use a \spatialAcronym\ subspace in this
+context yet.} for \spatialAcronym\ subspaces
 and
-$\mathbb{P}^n (\statey) = \statey$ for \spaceTimeAcronym\ subspaces.
+$\mathbb{P}^n (\stateyDiscrete) = \stateyDiscrete$ for \spaceTimeAcronym\ subspaces.
 %\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing
@@
 %trial space (e.g., an $L^2$ projector).
-	\KTC{Remove next remark?}
-\begin{remark}
-It is possible to construct the trial subspaces such that they have continuous support between time windows through the 
-use of, e.g., affine transformations.
-\end{remark}
 %and 
@@
 %projected onto the trial space; (e.g., via for $L^2$ projection).
-We now define the trial subspaces $\stspaceArg{n}$. We consider \spatialAcronym\ trial subspaces and \spaceTimeAcronym\ trial subspaces. 
-We leave further investigation into other trial subspaces, such as nonlinear manifolds~\cite{leeCarlberg}, as a subject for future work.
+We now define the trial subspaces $\stspaceArg{n}$ considered in this work. In
+	particular, we introduce \spatialAcronym\ trial subspaces and
+	\spaceTimeAcronym\ trial subspaces tailored for this context. 
+We leave further investigation into other approaches, such as nonlinear
+	manifolds~\cite{leeCarlberg}, as a subject for future work.
 %A variety of trial spaces have been examined in the model-order reduction community.

### 2019-10-07 section3/space_time.tex

@@
 \subsection{\spaceTimeAcronym\ trial spaces}
-We now briefly consider \spaceTimeAcronym\ trial subspaces. Over 
+We now briefly consider \spaceTimeAcronym\ trial subspaces. 
+\KTC{Error in first equation. This is the approximate solution not the real
+one. Use parallel language to hwo this was done in the last section for
+S-reduction trial subspaces. Also make sure the basis is zero at the beginning
+of the interval and the affine transformation is taken to be the solution at
+the end of the last window. This will not be too trivial, as the solution on
+the previous slab enters the trial subspace definition.}
+Over 
 each window, we introduce the space--time trial subspace
@@
 \end{align*}
-with $\text{Range}(\stbasisArgnt{n}) + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
+\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
+be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window.
 For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
+\KTC{Remove all commas before equations unless it makes sense gramatically.
+Add the $n$ superscript to $x_{ref}$}
 The state over each window is approximated by,
@@
 
-Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
+\KTC{Next sentence, use parallel language to the last section. It's subtly
+different but I think written better there.} Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
 the \spaceTimeAcronym\ trial subspace computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
@@
 \end{align}
+\KTC{In prior equation, remove projection and align properly with ST-LSPG
+stuff before.}
 
@@
 Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
-respect to vector.
-As such, the first-order optimality conditions can be derived using standard calculus.
-Differentiating the objective with respect to $\stgenstatey$ and setting equal to zero yields,
+respect to vector.  As such, the first-order optimality conditions can be
+	derived using standard calculus. Differentiating the objective with respect
+	to $\stgenstatey$ and setting equal to zero yields \KTC{Make sure that xref
+	is replaced with the last solution at the previous time step}
 \begin{equation}\label{eq:st_stationary}

### 2019-10-07 section3/tclsrm_formulation.tex
p 6
small

@@
 
-\subsection{\methodAcronym\ Summary}
+\subsection{\methodAcronym\ summary}
 This section outlined the \methodAcronym\ method for model reduction.
 In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows: \methodAcronym\
-computes the generalized coordinates $\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the minimization problems,
+\KTC{Here, use the same language as earlier regarding sequential computation
+of these coordinates that satisfy...} computes the generalized coordinates
+$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
+minimization problems, \KTC{remove superscript from optimizaton variable
+everywhere}
 \begin{equation*}
@@
 \end{equation*}
+\KTC{Everywhere in the manuscript, eliminate $1,2,\ldots,n$. Make this 
+$1,\ldots,n$. Also, above you use three dots instead of ldots. Fix this
+everywhere}
 where,
@@
 \end{align*} 
-subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
+\KTC{use en-dash here to separate numbers} subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
 trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
@@
 
-\spaceTimeAcronym\ trial subspaces were additionally considered.
+\KTC{Passive voice:} \spaceTimeAcronym\ trial subspaces were additionally considered.
 For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.

### 2019-10-07 main.tex
working

@@
 	integrate Eq.~\eqref{eq:FOM}. For time-critical or many-query applications,
-	it is essential to relplace the FOM ODE~\eqref{eq:FOM} with a strategy that
+	it is essential to replace the FOM ODE~\eqref{eq:FOM} with a strategy that
 	enables an approximate trajectory to be computed at lower computational
@@
 denotes the generalized
-coordinates.
-
-	From the space--time perspective, this is equivalent to approximating the
+coordinates. From the space--time perspective, this is equivalent to approximating the
 	FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$ with 
@@
 \end{equation}
-with
-	$\dim(\stspaceS) = K\ntimeSteps$
-and $\onesFunction\in\timeSpace$ defined as
+with $\onesFunction\in\timeSpace$ defined as
 $\onesFunction:\timeDummy\mapsto 1$.
@@
 residual to the spatial trial subspace in the (semi-)inner product induced by the 
-positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneArg{ }^T
+positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneTArg{ }
 \stweightingMatOneArg{ }$ (commonly set to
@@
 	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
-	\|_{\stweightingMatArg{n}}^2, \qquad \genstate(0) = \genstateIC,
+	\|_{\stweightingMatArg{}}^2, \qquad \genstate(0) = \genstateIC,
 	\qquad t \in [0,T],
@@
 \end{equation}
-where $\|\mathbf{x}\|_{\stweightingMatArg{n}}\equiv
-\sqrt{\mathbf{x}^T\stweightingMatArg{n}\mathbf{x}}$.
+where $\|\mathbf{x}\|_{\stweightingMatArg{}}\equiv
+\sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
 %where
@@
 ST-LSPG projection computes the space--time generalized coordinates that
-minimize the space--time FOM O$\Delta$ E residual:
+minimize the space--time FOM O$\Delta$E residual:
 \begin{equation}\label{eq:stlspg}
@@
 solving Eq.~\eqref{eq:stlspg} scales cubically with the number of space--time
-degrees of freedom $\stdim$, because---by performing projection in time---it
+degrees of freedom $\stdim$, because it
 leads to dense systems that do not expose any natural mechanism for exploiting
@@
 %\newline
-Output: Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneArg{}^T \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
+Output: Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneTArg{} \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
 %\newline
@@
     with $\mathbf{P}^* \equiv \begin{bmatrix} \mathbf{p}_1 & \cdots & \mathbf{p}_{\text{floor}(N_t/N_{\text{skip}})} \end{bmatrix}$, $\mathbf{p}_i \in \{0,1\}^N$. 
-    \item Select the first $n_s$ columns of $\mathbf{P}^*$ to form the sampling point matrix $\stweightingMatOneArg{}^T \in \{0,1\}^{N \times n_s}$. 
+    \item Select the first $n_s$ columns of $\mathbf{P}^*$ to form the sampling point matrix $\stweightingMatOneTArg{} \in \{0,1\}^{N \times n_s}$. 
     \item Augment the sampling point matrix, $\stweightingMatOneArg{}$, with additional columns such that all unknowns are computed at the mesh cells selected by Step 3.

### 2019-10-07 section3/general_form.tex
working

@@
 \subsection{Windowed least-squares for general space--time trial subspaces} 
-We begin by introducing a (nonuniform) partition of the time domain $[0,T]$
+We begin by introducing a (potentially nonuniform) partition of the time domain $[0,T]$
 into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
@@
 %[0,T] \rightarrow \RR{}.  \end{align}
-where $\stweightingMatArg{n} \equiv \stweightingMatOneArg{ }^T
-\stweightingMatOneArg{ } \in \RR{\fomdim \times \fomdim}$ denotes a
+where $\stweightingMatArg{n} \equiv \stweightingMatOneTArg{n}
+\stweightingMatOneArg{n} \in \RR{\fomdim \times \fomdim}$ denotes a
 symmetric positive semi-definite matrix that can enable hyper-reduction, for
-example. For notational simplicity, 
-we employ the same weighting matrix for all time windows.
+example. 
+%For notational simplicity, we employ the same weighting matrix for all time windows.
 

### 2019-10-07 main.tex
in progress

@@
 of existing model reduction approaches, as particular instances of
-  the methodology associate with Galerkin,
-	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
+  the methodology recover the Galerkin,
+	least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. In
 	addition, the approach
 	addresses key deficiencies in existing model-reduction
-	techniques, e.g., the dependence of (space--time) LSPG projection on the
+	techniques, e.g., the dependence of (space--time) LSPG on the
 	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
-	error bounds for Galerkin and LSPG projection.  We consider two types of
+	error bounds for both Galerkin and LSPG.  We consider two types of
 	space--time trial
@@
 	approach can yield more accurate solutions that better satisfy the governing equations than
-	Galerkin and LSPG projection.
-
-\begin{comment}
-This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
-	formulation for dynamical systems.  
-The proposed approach minimizes the
-	time-continuous full-order-model residual over a low-dimensional 
-	trial subspace across a sequence of time windows. % that decompose the time
-We consider two types of trial subspaces
-	within this framework:
-	one that reduces the spatial dimension of the full-order model, and one that
-	reduces both the spatial and temporal dimensions of the full-order model. 
-We consider two methods to solve the residual-minimization problems: direct
-	(i.e., discretize-then-optimize) and indirect (i.e.,
-	optimize-then-discretize).
-The approach addresses key deficiencies in a variety 
-of existing model-reduction approaches; e.g., the time-step and time-scheme dependence of least-squares Petrov--Galerkin (LSPG), the exponentially growing error bounds present in Galerkin and LSPG, and the offline cost of space--time LSPG.
-Various limiting cases of the proposed method yield existing reduced-order modeling techniques, namely Galerkin, Least-Squares Petrov–Galerkin, and Space–Time Least-Squares Petrov– Galerkin.
-Numerical experiments on the compressible Euler and Navier--Stokes equations demonstrate that, at an increased computational cost, the \methodAcronym\ formulation can yield more accurate and physically relevant solutions than the Galerkin and Least-Squares Petrov--Galerkin approaches.
-\end{comment}
+	the Galerkin and LSPG approaches.
+
 \end{abstract}
@@
 Simulating parameterized dynamical systems arises in many applications across
-science and engineering, including design, control, uncertainty
-quantification, and epidemiology, for example. In many contexts, executing a dynamical-system
+science and engineering. In many contexts, executing a dynamical-system
 simulation at a single parameter instance---which entails the numerical
@@
 \textit{online} stage during which they compute approximations to the
-dynamical-system trajectory that reside on this trial subspace via a
-projection process.
+dynamical-system trajectory that reside on this trial subspace via, e.g., 
+residual minimization.
 
@@
 observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
-quantified and accounted for in the projection process; this often results in
+quantified and accounted for; this often results in
 reduced-order models that inherit such important properties.  The primary
@@
 These
-efforts include residual-minimizing
-projection that equips the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
-space--time
-projection that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
+efforts include residual-minimization approaches that equip the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
+space--time approaches that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
 ``energy-based'' inner
@@
 %Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
-Residual-minimizing and space--time projection are the most relevant classes of
+Residual-minimizing and space--time approaches are the most relevant classes of
 methods for the current work and comprise the focus of the following review.
@@
 Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
-formulation and examined its connections with Galerkin projection. Critically,
-this work demonstrated that (1) under certain conditions, LSPG can be derived
-from a Petrov--Galerkin projection executed on the time-continuous
-full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
-in the limit as the time step goes to zero (i.e., Galerkin projection
+formulation and examined its connections with the Galerkin approach. Critically,
+this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
+Petrov--Galerkin approach applied to the time-continuous
+full-order-model residual, and (2) the LSPG and Galerkin approaches are equivalent
+in the limit as the time step goes to zero (i.e., the Galerkin approach 
 minimizes the time-instantaneous full-order-model residual). Numerous
-numerical experiments have demonstrated that LSPG projection often yields more
+numerical experiments have demonstrated that LSPG often yields more
 accurate and stable solutions than Galerkin
-projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
+~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
 The common intuitive explanation for this improved performance is that, by
 minimizing the full-order-model residual over a finite time window (rather
-than time instantaneously), LSPG projection computes solutions that are more
-accurate over a larger part of the trajectory as compared to Galerkin projection.
-
-However, the LSPG approach has several notable shortcomings. First, LSPG
-performance exhibits a complex dependence on the time discretization. In
+than time instantaneously), LSPG computes solutions that are more
+accurate over a larger part of the trajectory as compared to Galerkin.
+
+However, LSPG has several notable shortcomings. First, LSPG
+exhibits a complex dependence on the time discretization. In
 particular, changing
@@
 time-discretization error of the full-order model on which LSPG is based. 
-In particular, because LSPG and Galerkin projection are equivalent in the
-limit as the time step goes to zero, LSPG accuracy approaches the (poor)
+In particular, because LSPG and Galerkin are equivalent in the
+limit as the time step goes to zero, the accuracy of LSPG approaches the (poor)
 accuracy of Galerkin as the time step shrinks, but this also has the effect of
@@
 Researchers have pursued the development of space--time
-residual-minimization
-projection~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
+residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
 address the issues incurred by sequential residual minimization in time.
@@
 over all space and time, whose solution comprises a \text{fixed} vector of
-(space--time) generalized coordinate; on the other hand, sequential residual
+(space--time) generalized coordinates; on the other hand, sequential residual
 minimization in time (often) associates with a system of ODEs whose solution
@@
 optimality; this implies that the solution exhibits a notion of optimality
-over the entire time domain. However, these techniques also suffer from several
+over the entire time domain. 
+\EP{This has nothing to do the basis, which was the point I was trying to get across}
+However, these techniques also suffer from several
 limitations. First, the computational cost of solving the algebraic system arising from space--time
-projection scales cubically with the number of space--time degrees of freedom;
+approaches scales cubically with the number of space--time degrees of freedom;
 in contrast, the computational cost incurred by standard
@@
 state prediction, as these methods employ space--time basis vectors defined over the
-entire time domain of interest, which must have been including in the training
+entire time domain of interest, which must have been included in the training
 simulations.
@@
 residual-minimization problem, thus addressing a key deficiency of both LSPG
-and space--time LSPG projection. Critically, time-continuous residual minimization also exposes
+and space--time LSPG. Critically, time-continuous residual minimization also exposes
 two different solution methods:  a \textit{discretize-then-optimize}
@@
 the residual over arbitrarily defined \textit{time windows} rather than
-sequentially minimizing the residual over time steps (as in LSPG projection)
+sequentially minimizing the residual over time steps (as in LSPG)
 or over the entire time domain (as in space--time residual-minimization
@@
 model (as employed by traditional spatial-projection-based methods), and one that reduces both the spatial and temporal dimensions of the
-full-order model (as employed by space--time residual-minimization methods).
+full-order model (as employed by space--time methods).
 The above attributes allow that \methodAcronym\ method to be viewed as a
 generalization of existing model-reduction methods, as 
-Galerkin, LSPG, and space--time LSPG projection correspond to specific
+Galerkin, LSPG, and space--time LSPG correspond to specific
 instances of the formulation.
@@
 commonalities with several related efforts. 
-First, Ref.~\cite{bui_thesis} tersely formulated a space--time least-squares ROM
+First, Ref.~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
 and connected this formulation with optimal control; specifically it mentioned the
@@
 time-continuous FOM residual, (2) this work minimizes this residual over
-arbitrary time windows, and (3) this work considers trial subspaces
+arbitrary time windows, and (3) this work also considers trial subspaces
 characterized by a reduction of the spatial dimension of the FOM only.
@@
 	\item Derivations of conditions under which the
-		\methodAcronym\ framework recovers Galerkin, LSPG, and space--time LSPG
-		projection.
+		\methodAcronym\ framework recovers Galerkin, LSPG, and space--time LSPG.
 \item Error analysis of the \methodAcronym\ framework using
@@
 the spatial dimension of the FOM only.
-	This analysis demonstrates that the \methodAcronymROMs\ error 
+	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
 		is bounded \textit{a priori} by a combination of the
 		error at the start of the window and the integrated FOM ODE residual
-		evaluated at the FOM state projected onto the trial subspace. As a result,
-		the error bound exhibits exponential growth in the number of time windows
-		(not total time instances), yielding smaller error-bound growth in thime than
-		Galerkin or LSPG projection. \KTC{Check this}
+		evaluated at the FOM state projected onto the trial subspace. 
+\EP{Updated this, I don't show what you said}%As a result,
+%		the error bound exhibits exponential growth in the number of time windows
+%		(not total time instances), yielding smaller error-bound growth in thime than
+%		Galerkin or LSPG projection. \KTC{Check this}
 \item Numerical experiments for 
@@
 \begin{itemize}
-\item Minimizing the residual over a larger time window leads to more stable
-	and physically relevant solutions \KTC{What is `physically relevant'?} as
-		well as smaller space--time residual norms. 
+\item Minimizing the residual over a larger time window leads to more stable solutions 
+with lower space--time residuals norms.
 \item Minimizing the residual over a larger time window does not necessarily
@@
 outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
-space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
+space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
 framework. Section~\ref{sec:numerical_techniques}
@@
 	\eqref{eq:FOM} on such parameters for notational convenience, as this work
-	focuses on devising a projection applicable to a specific parameter instance.
+	focuses on devising a model-reduction approach applicable to a specific parameter instance.
 
@@
 		in model-reduction methods based on spatial projection followed by time
-		integration (e.g., Galerkin projection, LSPG projection).
+		integration (e.g., the Galerkin and LSPG approaches).
 	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
@@
 		model-reduction methods based on space--time projection (e.g., space--time
-		LSPG projection).
+		LSPG).
 \end{enumerate}
  We now describe these two types of space--time trial subspaces and their
-	application to Galerkin, LSPG, and space--time LSPG projection.
+	application to the Galerkin, LSPG, and space--time LSPG approaches.
 
@@
 Because Eq.~\eqref{eq:g_truncation} is overdetermined, a solution may not
-exist. Typically, either \textit{Galerkin projection} or \textit{least-squares
-Petrov--Galerkin projection} is employed to reduce the number of equations
+exist. Typically, either the \textit{Galerkin} or \textit{least-squares
+Petrov--Galerkin} approach is employed to reduce the number of equations
 such that a unique solution exists. We now describe these two methods.
 
-\subsubsection{Galerkin projection}
-Galerkin projection reduces the number of equations in
+\subsubsection{Galerkin approach}
+The Galerkin approach reduces the number of equations in
 Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
@@
 \begin{equation}\label{eq:g_truncation2}
-\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace
+\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace
 \genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
@@
 denotes the $K\times K$ positive definite mass matrix.
-As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, Galerkin projection
+As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, the Galerkin approach
 can be viewed alternatively as a residual-minimization
@@
 \sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
-%where
-%$$\hat{ \mathbf{v}} = \underset{\hat{ \stateyDiscrete}  }{\text{arg min}} || \basisspace \hat{\stateyDiscrete} -  \velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept) ||_{\stweightingMatArg{n}}^2.$$
-%The minimization problem has the solution,
-%$$\hat{\mathbf{v}} = \massArgnt{n}^{-1}\basismat^T \stweightingMatArg{n} \velocity(\basismat \genstateGalerkinArg{}{t} + \stateIntercept ).$$
-%Galerkin method is interpreted as finding the velocity that minimizes the FOM ODE residual~\cite{carlberg_lspg_v_galerkin},
-% The velocity that minimize the FOM ODE residual is given as the solution to the optimization problem, %To demonstrate this, we define the FOM residual, %The LSPG residual corresponds to minimizing the FOM O$\Delta$ E residual at each time-instance.
-obtained either by projecting Eq.~\ref{eq:g_truncation} via a test basis or by residual minimization.
-Equivalence conditions for these two approaches are established in Ref.~\cite{carlberg_lspg_v_galerkin}.
-This work considers dimension reduction by residual minimization.
-%The most common way to perform dimension reduction is by projecting Eq.~\ref{eq:g_truncation} via a test basis.
-An alternative philosophy that is pursued here is to perform dimension reduction through residual minimization.
-%Define the residual,
-%\begin{align*}
-%\resid & \vcentcolon \state \mapsto \frac{d \state}{dt} - \velocity(\state), \\
-%& \vcentcolon \RR{\fomdim} \times [0,T] \rightarrow \RR{\fomdim} \times [0,T].
-%\end{align*}
-%The Galerkin method can be interpreted as finding the velocity such that the instantaneous time residual is minimized,%Minimization of the residual over the trial space leads to what is known as the Galerkin method,
-%$$\mathbf{v} = \underset{\stateyDiscrete \in \text{Range}(\basismat)  }{\text{arg min}}|| \stateyDiscrete -  \velocity(\state(t)) ||_2^2.$$
-%As $\mathbf{v} \in \text{Range}(\basismat)$, the minimization problem can be written as,%and hence $\boldsymbol v = \basismat \hat{\boldsymbol v} $,
-%$$\hat{ \mathbf{v}} = \underset{\hat{ \stateyDiscrete}  }{\text{arg min}} || \basisspace \hat{\stateyDiscrete} -  \velocity(\state(t)) ||_2^2,$$
-%where $\hat{\mathbf{v}} \in \RR{K}$ are the generalized coordinates of $\boldsymbol v$ and $\boldsymbol v = \basisspace \hat{\boldsymbol v}$. 
-%The minimization problem has the solution~\cite{carlberg_lspg_v_galerkin},
-%$$\hat{\mathbf{v}} = \basismat^T \velocity(\basismat \genstate(t) + \stateIntercept ).$$
-%The Galerkin ROM is thus given by,
-%\begin{equation*}\label{eq:g_truncation2}
-%\frac{d}{dt}\genstateGalerkinArg{}{t} = \mass^{-1}\basisspace^T \stweightingMatArg{n} \velocity(\basisspace
-%\genstateGalerkinArg{}{t} + \stateIntercept).
-%\end{equation*}
 Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
@@
 %\end{equation}
-\subsubsection{Least-squares Petrov--Galerkin projection}
+\subsubsection{The Least-squares Petrov--Galerkin approach}
 Despite its time-instantaneous residual-minimization optimality property 
-\eqref{eq:GalOptimal}, Galerkin projection can yield inaccurate solutions,
+\eqref{eq:GalOptimal}, the Galerkin approach can yield inaccurate solutions,
 particularly if the velocity is not self-adjoint or
@@
 Least-squares Petrov--Galerkin (LSPG)
-projection~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
+~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
 was developed
-as an alternative projection method that exhibits several advantages over Galerkin
-projection.
-Rather than minimize the (time-continuous) FOM ODE residual at a time instance as in Galerkin
-projection, LSPG minimizes
+as an alternative method that exhibits several advantages over the Galerkin approach.
+Rather than minimize the (time-continuous) FOM ODE residual at a time instance (as in Galerkin), LSPG minimizes
 the (time-discrete) FOM O$\Delta$E residual 
@@
 over a time step.
-We now describe LSPG projection in the case of linear multistep methods;
-Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG projection for Runge--Kutta
+We now describe the LSPG approach in the case of linear multistep methods;
+Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG for Runge--Kutta
 schemes.

### 2019-10-07 main.tex
latest

@@
 As described in the introduction, although numerical experiments have
-demonstrated that LSPG projection often yields more accurate and stable
-solutions than Galerkin
-projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg},
-LSPG still suffers from several shortcomings, in particular its complex
+demonstrated that LSPG often yields more accurate and stable
+solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg},
+LSPG still suffers from several shortcomings. In particular, LSPG suffers from its complex
 dependence on the time discretization, exponentially growing error bounds, and
-a lack of optimality for the trajectory defined over the entire time domain.
+lack of optimality for the trajectory defined over the entire time domain.
 
@@
 spaces~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}
-aim to overcome the latter two shortcomings of LSPG projection, which are
+aim to overcome the latter two shortcomings of LSPG, which are
 shared by all model-reduction methods based on spatial projection followed by
-time integration.  Because these methods employ \spaceTimeAcronym\ trial
+time integration. \EP{This is misleading, going to delete} Because these methods employ \spaceTimeAcronym\ trial
 spaces, they reduce both the spatial and temporal dimensions of the full-order
@@
 \boldsymbol 0$
-denotes the space--time trial basis.
+denotes the space--time trial basis.\footnote{For simplicity, we consider the case where the trial space is homogeneous on the boundary.} 
 Thus, at any time instance $t\in[0,T]$, ROMs that employ the
@@
 Critically, comparing the approximations arising from \spatialAcronym\ and
-\spaceTimeAcronym\ subspaces in Eqs.~\eqref{eq:affine_trialspace} and \eqref{eq:stapprox1}, respectively,
-highlights that the former approximation associates with a time-dependent
+\spaceTimeAcronym\ trial subspaces in Eqs.~\eqref{eq:affine_trialspace} and \eqref{eq:stapprox1}, respectively,
+highlights that the former approximation associates with time-dependent
 generalized coordinates, while the latter approximation associates with a
@@
 are automatically satisfied from the definition of the \spaceTimeAcronym\
-subspace.
-
-Space--time projection methods reduce the number of equations in
+trial subspace.
+
+Space--time methods reduce the number of equations in
 \eqref{eq:st_fom} to ensure a unique solution exists.

### 2019-10-07 main.tex
latest

@@
 \begin{equation*}
-\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2, \qquad n = 1,\ldots,\ntimeSteps,
+\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,\ldots,\ntimeSteps,
 \end{equation*}
@@
 We now outline one such method: space--time least-squares Petrov--Galerkin (ST-LSPG)
-projection~\cite{choi_stlspg}. 
-While space--time Galerkin projection \cite{benner_st} is another alternative, it does not
+~\cite{choi_stlspg}. 
+While the space--time Galerkin method~\cite{benner_st} is another alternative, it does not
 associate with any residual minimization principle, and thus we do not discuss
@@
 \subsubsection{Space--time LSPG projection}  
-Analogously to LSPG projection, space--time LSPG projection~\cite{choi_stlspg}
+Analogously to LSPG, space--time LSPG~\cite{choi_stlspg}
 minimizes the (time-discrete) FOM O$\Delta$E residual, but does so
@@
 \end{align*}
-ST-LSPG projection computes the space--time generalized coordinates that
+ST-LSPG computes the space--time generalized coordinates that
 minimize the space--time FOM O$\Delta$E residual:
@@
 solving Eq.~\eqref{eq:stlspg} scales cubically with the number of space--time
-degrees of freedom $\stdim$, because it
-leads to dense systems that do not expose any natural mechanism for exploiting
+degrees of freedom $\stdim$. This cost is due to the fact that ST-LSPG yields 
+a dense systems that do not expose any natural mechanism for exploiting
 the sequential nature of time evolution. Second, it is unclear how these

### 2019-10-07 section3/general_form.tex
latest

@@
 outlines the proposed \methodNameLower\ 
-(\methodAcronym) framework. In contrast to (1) Galerkin projection, 
-which minimizes (time-continuous) FOM ODE residual at a time instance, 
-(2) LSPG projection, 
+(\methodAcronym) framework. In contrast to (1) the Galerkin approach, 
+which minimizes the (time-continuous) FOM ODE residual at a time instance, 
+(2) the LSPG approach, 
 which minimizes
 the (time-discrete) FOM O$\Delta$E residual 
-over a time step, and (3) ST-LSPG projection, 
+over a time step, and (3) the ST-LSPG approach,  
 which minimizes
@@
 \begin{centering} 
-\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} \caption{Decomposition of time
-domain into time windows.} \label{fig:slab_fig} \end{centering} 
+\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} 
+\caption{Partitioning of the time
+domain into time windows.} 
+\label{fig:slab_fig} 
+\end{centering} 
 \end{figure}
@@
 \end{equation}
-where $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
+where $\stspaceArg{n}$ is the space--time trial subspace over the $n$th window, $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
 $[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} = \{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$)
 and
-$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$.
-
-Over the $n$th window, define the objective functional
+$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
+$$\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n} \subset \stspaceArg{n}, \qquad n=1,\ldots,\nslabs. 
+%, \qquad 
+%\approxstateArg{n}{\timeEndArg{n}} \in \stspaceBoundEndArg{n} \subset \stspaceArg{n}.
+$$ 
+To outline \methodAcronym, we now define the objective functional over the $n$th window:
 \begin{equation}\label{eq:obj}
@@
 %\end{equation}
-where 
-$\mathbb{P}^n (\stateyDiscrete) = XXX$ \KTC{EP TODO. Not trivila because
-haven't introduced how you're going to use a \spatialAcronym\ subspace in this
-context yet.} for \spatialAcronym\ subspaces
-and
-$\mathbb{P}^n (\stateyDiscrete) = \stateyDiscrete$ for \spaceTimeAcronym\ subspaces.
+where $\projectorArg{n} : \RR{\fomdim} \rightarrow \stspaceBoundStartArg{n}$ 
+is a (spatial) projector onto the start of the $n$th trial space (e.g., $\elltwo$).
+
+%for \spatialAcronym\ subspaces
+%and
+%$\mathbb{P}^n (\stateyDiscrete) = \stateyDiscrete$ for \spaceTimeAcronym\ subspaces.
 %\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing

### 2019-10-07 section3/space_time.tex
latest

@@
 \subsection{\spaceTimeAcronym\ trial spaces}
-We now briefly consider \spaceTimeAcronym\ trial subspaces. 
 \KTC{Error in first equation. This is the approximate solution not the real
@@
 the previous slab enters the trial subspace definition.}
-Over 
+
+We now briefly consider \spaceTimeAcronym\ trial subspaces. 
+The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
+the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
+with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
+\begin{equation}\label{eq:st_sttrialspace}
+ \stspaceSTArg{n} \defeq
+ \Range{\stbasisArgnt{n}} + \stateInterceptArg{n}\otimes \onesFunctionArg{n} .
+\end{equation}
+Here $\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}} \otimes \timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function}.
+At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
+ \begin{equation}\label{eq:stapprox}
+ \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},
+\end{equation}
+where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.
+
+\begin{comment}
 each window, we introduce the space--time trial subspace
@@
 %\end{equation}
-where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.
+%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.
 
+\end{comment}
 \KTC{Next sentence, use parallel language to the last section. It's subtly
@@
   \begin{cases} 
-\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
-\mathbb{P}^n(\stateFOMArgnt{1})(0)
+\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
+\mathbb{P}^n(\stateFOMIC)
  & n=1. \end{cases}

### 2019-10-07 section4/numerical_methods.tex
latest

--- 2019-10-02 section4/numerical_methods.tex
@@
 The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
-As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged.
-This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system.
-We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces,
+\EP{I think passive voice is fine in this instance, it reads easier}
+As \methodAcronym\ displays commonalities with optimal
+control problems, a large existing body of work can be leveraged. This section
+discuses a variety of such techniques and outlines how they can be used to
+solve the \methodAcronym\ system. We focus primarily on 
+%\KTC{earlier we kept
+%saying 
+%direct (i.e.,
+%	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
+%	Let's be consistent when we re-introduce it here.}
+direct (discretize-the-optimize) and indirect (optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces.
-%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems.
-For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$.
-The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates.
-%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).
 
@@
 \textit{direct} or \textit{indirect}
-methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
-and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
+methods~\cite{conway_optimalcontrolreview}.  Direct methods eschew the first order optimality conditions
+~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} and, instead, ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
 numerically discretizing the system and objective functional and ``transcribing"
@@
 Indirect methods, on the other hand, leverage the calculus of variations to
-derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
+derive the Euler--Lagrange equations, which comprise the first-order
+optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} associated with the objective 
 functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
 solve the optimization problem
-``indirectly". For both direct and indirect methods, a variety of
-discretization/solution approaches are then possible. Collocation methods,
+``indirectly". A variety of discretization/solution techniques are possible for both direct and indirect methods. Collocation methods,
 finite elements, spectral methods, shooting methods, etc., are all examples of
@@
 
-In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial subspaces. We consider:
-
-\begin{itemize} \item Discretize then Optimize (Direct Methods): 
-\begin{itemize}
-\item Discretize and then optimize approaches are outlined for linear multistep methods. 
-\end{itemize} 
-\item Optimize then Discretize (Indirect Methods): 
-\begin{itemize} 
-\item An indirect method leveraging the forward-backward sweep algorithm is outlined 
+In the present context, we investigate both direct and indirect methods to solve \methodAcronym\ with \spatialAcronym\ trial
+subspaces. We consider: 
+\begin{itemize} \item \textit{Direct Methods (discretize-then-optimize)}: Discretize and then optimize approaches are outlined for linear multistep methods. 
+\item \textit{Indirect methods (optimize-then-discretize)}: An indirect method leveraging the forward-backward sweep algorithm is outlined 
 to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
 \end{itemize} 
-\end{itemize} 
-It
-is emphasized that a variety of other approaches exist. A detailed exploration
-of additional approaches will be a subject of future work. Section~\ref{sec:direct} 
-will outline direct solution approaches based on linear multistep methods, 
-while Section~\ref{sec:indirect} will outline the indirect solution approach.
+We emphasize that a variety of other approaches exist. Future work will comprise a detailed exploration
+of additional approaches.
+Section~\ref{sec:direct}  outlines direct solution approaches based on linear
+multistep methods, while Section~\ref{sec:indirect} outlines the indirect
+solution approach.
 

### 2019-10-07 section4/spatial_numerical.tex
latest

--- 2019-10-03 section4/spatial_numerical.tex
@@
 \subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
+\KTC{Does the time grid definted in this section have any relationship to the
+time grid used for time discretization? Haven't gotten past HERE below, but
+wanted to check (maybe it's made clear later).}
+\EP{This time grid becomes the time discretization over each window. At this point in the paper, 
+there is no notion of a time grid for \methodAcronym}
+
 Direct approaches solve~\eqref{eq:tclsrm} by
@@
 dimensional one by discretizing the state and objective function in time.
-The minimization problem is then reformulated as a nonlinear programming, or
+\KTC{Passive voice:} \EP{I think this reads clearly} The minimization problem is then reformulated as a nonlinear programming, or
 optimization, problem. A variety of direct solution approaches exist, 
@@
 approaches, and genetic algorithms.  
+\KTC{is this a
+spectral method, or the spectral method? clarify} \EP{I don't understand what you mean. What is ``the" spectral method}
+
+\KTC{you can't devide (i.e., decompose) a window into instances. you can divide a window into
+sub-windows. you can also simply define a set of time instances in the window.
+no need to emphasize `strictly increasing', as any set of points in the window
+will by necessity be ordered.}
 
 In the present context, the most straightforward direct solution approach consists of the 
-following steps:
-\begin{enumerate}
-\item Numerically discretize the FOM ODE (and hence the \textit{integrand} of~\eqref{eq:obj_gen_slab}).
-\item Select a numerical quadrature rule to evaluate the \textit{integral}
-in~\eqref{eq:obj_gen_slab}.
-\end{enumerate}
+following steps: (1) numerically discretize the FOM ODE (and hence the \textit{integrand} of~\eqref{eq:obj_gen_slab}) and 
+(2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
 To achieve this agenda, first define a discrete temporal grid 
-by dividing each time window $[\timeStartArg{n},\timeEndArg{n}]$ into
-$\nstepsArg{n}$ strictly increasing time-step instances described by
+by defining
+$\nstepsArg{n}+1$ strictly increasing time instances described by
 $\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with
+%\KTC{use cdots, not
+%hdots or ldots. Don't use $\defeq$ for the first and last times; these are
+%just equalities, i.e., you happen to set them equal to those values, but they
+%aren't identically the same:}
 \begin{align}
 &\timeWindowArg{n}{0} \le \ldots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
-\timeWindowArg{n}{0} \defeq \timeStartArg{n},\;
-\timeWindowArg{n}{\nstepsArg{n}} \defeq \timeEndArg{n} . 
+\timeWindowArg{n}{0} = \timeStartArg{n},\;
+\timeWindowArg{n}{\nstepsArg{n}} = \timeEndArg{n} . 
 %\timeWindowArg{n}{0} = \timeStartArg{n}, \timeWindowArg{n}{\nstepsArg{n}} =
@@
 \end{align}
-This discretization is depicted in Figure~\ref{fig:slab_fig2}.
+Figure~\ref{fig:slab_fig2} depicts this discretization.
 We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
@@
 \includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
-\caption{Decomposition of time windows into time-step instances.} 
+	\caption{Partitioning of time windows into time-step instances. \KTC{Don't call them `time-step instances'.
+	This woudl be an instance of a time step. These are `time instances',
+	correct? If so, fix this terminology everywhere.}} 
 \label{fig:slab_fig2} 
@@
 \begin{split} 
-\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{n,i},\ldots, \stateyDiscreteArgnt{n,i-k(n)})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{n,i},\ldots, \stateyDiscreteArgnt{n,i-k(n)}), \\
+\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-k^n(i)})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-k^n(i)}), \\
 &\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}+1} \rightarrow
@@
 where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. 
-Over each window, \methodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+Over each window, \directMethodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
 \begin{align*}
-&\underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{minimize } }
-\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
-& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{n,0} =
-\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
-\approxstateIC & n=1. \end{cases} \end{align*}
+&\underset{\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}} \in \trialspace}{\text{minimize } }
+\objectiveArgLMS{n} (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \\
+& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
+\begin{cases} \spatialIC & n = 2,\ldots,\nslabs,\\
+\genstateICOne & n=1. \end{cases} \end{align*}
 %\methodAcronym\ solves the minimization problem over each time window, 
@@
 %\approxstateIC & n=1. \end{cases} \end{align*}
-Defining $\genstateDiscreteArg{n,i} \in \RR{\romdim}$ to be the discrete approximation to the generalized coordinates at the $i$th time-step instance over the $n$th window, 
+Defining $\genstateDiscreteArg{n,i} \in \RR{\romdim}$ to be the discrete approximation to the generalized coordinates at the $i$th time instance over the $n$th window, 
 the minimization problem can be written equivalently as,
@@
 \begin{split}
-& \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
+& \underset{\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}}}{\text{minimize } }
+\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}) \\ 
 & \text{subject to} \hspace{0.2 in}
-\genstateyDiscreteArg{n,0} =
-\begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
-\basisspace^T(\stateFOMIC - \stateIntercept)& n=1. \end{cases} 
+\genstateyDiscreteArg{0} =
+\begin{cases} \spatialIC & n = 2,\ldots,\nslabs \\
+\genstateICOne& n=1. \end{cases} 
 \end{split}
@@
 \genCollocMatArg{n}{1},\ldots,\genCollocMatArg{n}{\nstepsArg{n}} 
- = \underset{\genCollocMatyArg{n}{1},\ldots,\genCollocMatyArg{n}{\nstepsArg{n}} }{\text{argmin } }
-\objectiveArgC{n} (\basisspace \genCollocMatyArg{n}{1} + \stateInterceptMat,\ldots,\basisspace \genCollocMatyArg{n}{\nstepsArg{n}} + \stateInterceptMat) \qquad \text{ for } n = 1,2,...,\nslabs,
+ = \underset{\genCollocMatyArg{}{1},\ldots,\genCollocMatyArg{}{\nstepsArg{n}} }{\text{argmin } }
+\objectiveArgC{n} (\basisspaceArg{n} \genCollocMatyArg{}{1} + \stateInterceptMat,\ldots,\basisspaceArg{n} \genCollocMatyArg{}{\nstepsArg{n}} + \stateInterceptMat) \qquad \text{ for } n = 1,2,...,\nslabs,
 \end{equation*}
@@
 \genCollocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1  \\
-\basisspace^T (\approxstateIC - \stateIntercept)& n=1, \;  i=1,
+\basisspaceTArg{n} (\approxstateIC - \stateIntercept)& n=1, \;  i=1,
 \end{cases}
@@
 \genstatecollocMatSlabArg{n} \defeq 
-\unroll (\genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
+\unroll (\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
 \end{equation*}
@@
 \begin{equation*}
-\residLMSSlabArg{n} : \genstatecollocMatySlabArg{n} \mapsto \begin{bmatrix}
- \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n,1} \residLMSArg{n,1}( \basisspace \genstateyDiscreteArgnt{n,1} + \stateIntercept, \ldots,\basisspace \genstateyDiscreteArgnt{n,1 - k^n(1) }  + \stateIntercept) \\
+\residLMSSlabArg{n} : \genstatecollocMatySlabArg{} \mapsto \begin{bmatrix}
+ \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n,1} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n},) \\
 \vdots \\
- \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n,\nstepsArg{n}} \residLMSArg{n,\nstepsArg{n}}( \basisspace \genstateyDiscreteArgnt{n,\nstepsArg{n}} + \stateIntercept, \ldots,\basisspace \genstateyDiscreteArgnt{n,\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateIntercept) \\
+ \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n,\nstepsArg{n}} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
 \end{bmatrix},
 \end{equation*}
-where $\genstatecollocMatySlabArg{n} \defeq \unroll (\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}} ).$
+where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
 It is worth noting that, by design,
 \begin{equation*}
-\objectiveArgLMS{n} \bigg( \basisspace \genstateDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateDiscreteArg{n,\nstepsArg{n}} + \stateIntercept \ \bigg) 
+\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
 \equiv
@@
 \begin{equation*}
-\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{n}}(\genstatecollocMatSlabArg{n})  =
+\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}(\genstatecollocMatSlabArg{n})  =
  \begin{bmatrix*}[l]
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $tol$; initial guess, $\genstateGuessDiscreteArg{n,0}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$}
+\Input{tolerance, $\text{tol}$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$; initial condition $\genstateDiscreteArg{n,0}$}
 \Output{Solution to least squares problem, $\genstatecollocMatSlabArg{n}$} 
 \textbf{Online Steps}: \\
-%\While{ $tol \le \frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \sum_{j=1}^{\ncollocArg{n}{i}} \big[ \dot{\state}
-%(\collocPointArg{n}{i}{j})  - \velocity(\stateArg{}{\collocPointArg{n}{i}{j}}) \big]^T \stweightingMatArg{n}(\collocPointArg{n}{i}{j}) \big[ \dot{\state}(\collocPointArg{n}{i}{j}) - \velocity(\stateArg{}{\collocPointArg{n}{i}{j}}) \big] \int_{\timeStartArg{n}}^{\timeEndArg{n}}   \lagrangeBasisScalarArg{n}{i}{j}(t) d \tau$}
 $\text{converged} \leftarrow \text{false}$ \Comment{Set convergence checker} \\
 $k \leftarrow 0$ \Comment{Set counter}\\
-$\genstatecollocMatSlabArg{n}_k \leftarrow \unroll(\genstateGuessDiscreteArg{n,0}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
+$\genstatecollocMatSlabArg{n}_k \leftarrow \unroll(\genstateDiscreteArg{n,0},\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
 \While{\text{converged} == \text{false}}
@@
 $\mathbf{J} \leftarrow  
-\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{n}}(\genstatecollocMatSlabArg{n}_k) 
+\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}(\genstatecollocMatSlabArg{n}_k) 
 $ \Comment{Compute weighted residual-Jacobian over window} \\

### 2019-10-07 section3/space_time.tex
latest

@@
-\subsection{\spaceTimeAcronym\ trial spaces}
+\subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
 \KTC{Error in first equation. This is the approximate solution not the real
@@
 \end{equation}
-Here $\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}} \otimes \timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function}.
+Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function}.
 At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,

### 2019-10-07 section3/tclsrm_formulation.tex
latest

@@
 \begin{equation*}
-\underset{\stateyArgnt{n} \in \stspace^n}{\text{minimize } }
-\mathcal{J}^n(\stateyArgnt{n}), \qquad n = 1,2,...,\nslabs,
+\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
+\mathcal{J}^n(\stateyArgnt{}), \qquad n = 1,\ldots,\nslabs,
 \end{equation*}
-\KTC{Everywhere in the manuscript, eliminate $1,2,\ldots,n$. Make this 
-$1,\ldots,n$. Also, above you use three dots instead of ldots. Fix this
-everywhere}
 where,
@@
 \begin{align*}\
-&\mass \frac{d}{dt} \genstateArg{n}{t}  -  \basisspace^T
-\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \mass \adjointArg{n}{t} , \\
- &\mass \frac{d}{dt} \adjointArg{n}{t}  + \basisspace^T \bigg[\frac{\partial
-\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspace
+&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
+\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
+ &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
+\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
  \adjointArg{n}{t} = \nonumber \\ 
-&-\basisspace^T \big[
+&-\basisspaceTArg{n} \big[
 \frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
-\basisspace [\massArgt{n}{t}]^{-1} \basisspace^T \stweightingMatArgt{n}{t} \bigg)
- \bigg( \basisspace \frac{d}{dt}\genstateArg{n}{t}  -
+\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \bigg)
+ \bigg( \basisspace \genstateDotArg{n}{t}  -
 \velocity(\veloargsromn) \bigg),  &
@@
 \end{align*} 
-\KTC{use en-dash here to separate numbers} subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
+subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
 trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
@@
 
-\KTC{Passive voice:} \spaceTimeAcronym\ trial subspaces were additionally considered.
+Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces.
 For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.

### 2019-10-07 section4/spatial_numerical.tex
latest

@@
 dimensional one by discretizing the state and objective function in time.
-\KTC{Passive voice:} \EP{I think this reads clearly} The minimization problem is then reformulated as a nonlinear programming, or
+The minimization problem is then reformulated as a nonlinear programming, or
 optimization, problem. A variety of direct solution approaches exist, 
@@
 
-\KTC{you can't devide (i.e., decompose) a window into instances. you can divide a window into
-sub-windows. you can also simply define a set of time instances in the window.
-no need to emphasize `strictly increasing', as any set of points in the window
-will by necessity be ordered.}
-
 In the present context, the most straightforward direct solution approach consists of the 
@@
 To achieve this agenda, first define a discrete temporal grid 
-by defining
+by introducing
 $\nstepsArg{n}+1$ strictly increasing time instances described by
 $\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with
+\EP{These should be strictly increasing}
 %\KTC{use cdots, not
@@
 \includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
-	\caption{Partitioning of time windows into time-step instances. \KTC{Don't call them `time-step instances'.
-	This woudl be an instance of a time step. These are `time instances',
-	correct? If so, fix this terminology everywhere.}} 
+	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 3$ for all $n$.} 
 \label{fig:slab_fig2} 
@@
 \begin{align*}
-\residArg{n,i} &: (\stateyDiscreteArgnt{n,i},\ldots,\stateyDiscreteArgnt{n,i-k(i)}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{n,i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{n,i-j},\timeWindowArg{n}{i-j},\tau^{n,i-j}) \\
-               &: \RR{\fomdim} \otimes \RR{k(i)+1} \rightarrow \RR{\fomdim}. 
+\residArg{n,i} &: (\stateyDiscreteArgnt{i},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
+               &: \RR{\fomdim} \otimes \RR{\lmsWidthArg{n}{i}+1} \rightarrow \RR{\fomdim}. 
 \end{align*} 
@@
 where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. 
-Over each window, \directMethodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+Over each window, \methodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
 \begin{align*}
@@
 & \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
-\begin{cases} \spatialIC & n = 2,\ldots,\nslabs,\\
-\genstateICOne & n=1. \end{cases} \end{align*}
+\begin{cases} \basisspaceArg{n} \spatialIC + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
+\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} \end{align*}
 %\methodAcronym\ solves the minimization problem over each time window, 
@@
 \begin{equation*}
-\residLMSSlabArg{n} : \genstatecollocMatySlabArg{} \mapsto \begin{bmatrix}
- \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n,1} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n},) \\
+\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{};\genstateDiscreteArg{n,0}) \mapsto \begin{bmatrix}
+ \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n,1} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
 \vdots \\
@@
 \end{equation*}
-where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
+where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
 It is worth noting that, by design,
@@
 \equiv
-\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
+\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg].
 \end{equation*} 
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $\text{tol}$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$; initial condition $\genstateDiscreteArg{n,0}$}
+\Input{tolerance, $\epsilon$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$; initial condition $\genstateDiscreteArg{n,0}$}
 \Output{Solution to least squares problem, $\genstatecollocMatSlabArg{n}$} 
@@
 $k \leftarrow 0$ \Comment{Set counter}\\
-$\genstatecollocMatSlabArg{n}_k \leftarrow \unroll(\genstateDiscreteArg{n,0},\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
+$\genstatecollocMatSlabArg{n}_k \leftarrow \unroll(\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
 \While{\text{converged} == \text{false}}
@@
 %}
-$\mathbf{r} \leftarrow \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{Compute weighted residual over window} \\
+$\mathbf{r} \leftarrow \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k;\genstateDiscreteArg{n,0})$ \Comment{Compute weighted residual over window} \\
 $\mathbf{J} \leftarrow  
@@
 $\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
-\If{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \text{tol}$ }{
+\If{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \epsilon$ }{
 {\text{converged} $\leftarrow$ \text{true}}  \Comment{Check and set convergence based on gradient norm} \\
@@
 indirect methods ``indirectly" solve the minimization
-problem~\eqref{eq:tclsrm} by solving the Euler-Lagrange equations. The
-system given by the Euler Lagrange equations~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_adjoint} comprises a coupled two-point boundary value
+problem~\eqref{eq:tclsrm} by solving the Euler--Lagrange equations. The
+system given by the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_adjoint} comprises a coupled two-point boundary value
 problem. Similar to the direct solution approach, a variety of techniques have
@@
 A proper value of $\rho$ can be
-obtained with a line search. The line search presented in~\eqref{eq:lspg_continuous} adapts the damping factor
+obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the damping factor
 according to the objective. Convergence properties of the FBSM method are
@@
 guess for costate}\\ 
-$\genstateArgnt{n}_1 \solveAlgo \basisspace^T \stweightingMatArg{n}
-\basisspace \genstateDotArgnt{n}_1  -  \basisspace^T \stweightingMatArg{n}
+$\genstateArgnt{n}_1 \solveAlgo \basisspaceTArg{n} \stweightingMatArg{n}
+\basisspaceArg{n} \genstateDotArgnt{n}_1  -  \basisspaceTArg{n} \stweightingMatArg{n}
 \velocity(\veloargsromntArg{1}) =  \controllerArgnt{n}$ 
@@
 \adjointArgnt{n} \underset{\text{solve}}{\leftarrow} 
-\frac{d}{dt} \adjoint^n  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \basisspace \mass^{-1} \adjoint^n = \\ -\bigg[\basisspace^T \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspace \mass^{-1} \basisspace^T \bigg)  \stweightingMatArg{n} \bigg( \basisspace \dot{\genstate}_i^n   -   \velocity( \basisspace \genstate_i^n  +\stateIntercept,t) \bigg) \bigg]
+\frac{d}{dt} \adjoint^n  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjoint^n = \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspace [\massArg{n}]^{-1} \basisspaceTArg{n} \bigg)  \stweightingMatArg{n} \bigg( \basisspaceArg{n} \dot{\genstate}_i^n   -   \velocity( \basisspaceArg{n} \genstate_i^n  +\stateInterceptArg{n},t) \bigg) \bigg]
 \end{multline*} }
@@
 $i \leftarrow i+1$ \Comment{Update Counter}\\
-$\genstateArgnt{n}_i \solveAlgo \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArgnt{n}_i   -  \basisspace^T \stweightingMatArg{n} \velocity(\basisspace \genstate^n_i + \stateIntercept,t) =  \controllerArgnt{n} $
+$\genstateArgnt{n}_i \solveAlgo \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i + \stateInterceptArg{n},t) =  \controllerArgnt{n} $
 \Comment{Solve~\ref{eq:lspg_continuous} with updated guess to costate}\\
 %$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
-\uIf{ $\objectiveArg{n}({\genstate_i^n(t)}) \le \objectiveArg{n}({\genstate_{i-1}^n(t)})$}
+\uIf{ $\objectiveArg{n}({\genstate_i^n}) \le \objectiveArg{n}({\genstate_{i-1}^n})$}
 {
@@
 The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
-The numerical implementation of the FBSM requires two main ingredients: (1.) temporal discretization schemes for the forward and backward problems and (2.) 
+The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
 an efficient method for computing action of the Jacobian transpose on a vector.
@@
 backwards problems. As described in  
-the direct solution approach, temporal discretization is achieved by dividing each time window into $\nstepsArg{n}$ time-step instances~\eqref{eq:timegrid1}.
+the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time-step instances~\eqref{eq:timegrid1} over each time window.
 Linear multistep methods then use this time-grid for temporal discretization of~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}.
@@
 One can make the manipulation,
-$$\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} \bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace \bigg]^T.$$
+$$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T.$$
 This manipulation allows one to compute the action of the Jacobian on each column of $\basisspace$ by, e.g., a finite difference approximation,
-$$\frac{\partial \velocity}{\partial \stateyDiscrete} \basisvec_i \approx \frac{1}{\epsilon}\bigg( \velocity(\stateyDiscrete + \epsilon \basisvec_i) - \velocity(\stateyDiscrete) \bigg).$$
+$$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$
 The Jacobian transpose can be formed in $K+1$ evaluations of the velocity.

### 2019-10-08 main.tex
latest
through space--time, proofing

@@
 over the entire time domain. 
-\EP{This has nothing to do the basis, which was the point I was trying to get across}
+\EP{Come back to this, this also has to do with residual minimization}
 However, these techniques also suffer from several
@@
 spaces~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}
-aim to overcome the latter two shortcomings of LSPG, which are
-shared by all model-reduction methods based on spatial projection followed by
-time integration. \EP{This is misleading, going to delete} Because these methods employ \spaceTimeAcronym\ trial
+aim to overcome the latter two shortcomings of LSPG. Because these methods employ \spaceTimeAcronym\ trial
 spaces, they reduce both the spatial and temporal dimensions of the full-order

### 2019-10-08 section3/general_form.tex
through space--time, proofing

@@
 $\approxstateArgnt{n}$ is the solution to the
-minimization problem
+minimization problems
 \begin{equation}\label{eq:tclsrm}

### 2019-10-08 section3/space_time.tex
through space--time, proofing

@@
 \subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
-\KTC{Error in first equation. This is the approximate solution not the real
-one. Use parallel language to hwo this was done in the last section for
-S-reduction trial subspaces. Also make sure the basis is zero at the beginning
-of the interval and the affine transformation is taken to be the solution at
-the end of the last window. This will not be too trivial, as the solution on
-the previous slab enters the trial subspace definition.}
+%\KTC{Error in first equation. This is the approximate solution not the real
+%one. Use parallel language to hwo this was done in the last section for
+%S-reduction trial subspaces. Also make sure the basis is zero at the beginning
+%of the interval and the affine transformation is taken to be the solution at
+%the end of the last window. This will not be too trivial, as the solution on
+%the previous slab enters the trial subspace definition.}
 
-We now briefly consider \spaceTimeAcronym\ trial subspaces. 
+%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
 The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
@@
  \stspaceSTArg{n} \defeq
- \Range{\stbasisArgnt{n}} + \stateInterceptArg{n}\otimes \onesFunctionArg{n} .
+ \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} .
 \end{equation}
-Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function}.
+Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
+For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
 At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
  \begin{equation}\label{eq:stapprox}
- \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},
+ \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
 \end{equation}
 where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
-
 \begin{comment}
@@
 \KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
-be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window.
+be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window.
 For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
@@
 \begin{equation}\label{eq:stapprox}
- \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
+ \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
 \end{equation}
@@
 \end{comment}
-\KTC{Next sentence, use parallel language to the last section. It's subtly
-different but I think written better there.} Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
-the \spaceTimeAcronym\ trial subspace computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm} implies that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
 \begin{align}\label{eq:obj_gen_slab_spacetime}
 \begin{split}
-&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptArg{n} ), \\ 
-      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =
-  \begin{cases} 
-\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
-\mathbb{P}^n(\stateFOMIC)
- & n=1. \end{cases}
-%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
+&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} ).% \\ 
+%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =0
+%  \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMIC)
+% & n=1. \end{cases}
+%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
 %\genstateIC & n=1, \end{cases} 
@@
 \end{align}
-\KTC{In prior equation, remove projection and align properly with ST-LSPG
-stuff before.}
-
 %\begin{align}\label{eq:obj_gen_slab_spacetime}
@@
 	derived using standard calculus. Differentiating the objective with respect
-	to $\stgenstatey$ and setting equal to zero yields \KTC{Make sure that xref
-	is replaced with the last solution at the previous time step}
+	to $\stgenstatey$ and setting equal to zero yields 
 \begin{equation}\label{eq:st_stationary}
@@
 \velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
-\stateInterceptArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},t) \bigg) dt = \bz.\end{equation}
+\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
 %subject to the boundary conditions 

### 2019-10-08 section3/tclsrm_formulation.tex
through space--time, proofing

@@
 This section outlined the \methodAcronym\ method for model reduction.
-In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows: \methodAcronym\
-\KTC{Here, use the same language as earlier regarding sequential computation
-of these coordinates that satisfy...} computes the generalized coordinates
-$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
-minimization problems, \KTC{remove superscript from optimizaton variable
-everywhere}
+In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows; i.e.,
+%\methodAcronym\  
+%computes the generalized coordinates
+%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
+%minimization problems, 
+\methodAcronym\ sequentially computes approximate solutions
+$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
+$\approxstateArgnt{n}$ is the solution to the
+minimization problem
+
 \begin{equation*}
@@
 \end{align*}
-subject to the boundary and continuity conditions.
+subject to the boundary conditions.
 For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
@@
 \end{align*} 
-subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
+subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler--Lagrange equations expose the fact that, for \spatialAcronym\
 trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
 controller that enforces the residual minimization property. 
-
 Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces.

### 2019-10-08 section4/numerical_methods.tex
through space--time, proofing

@@
 The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
-\EP{I think passive voice is fine in this instance, it reads easier}
-As \methodAcronym\ displays commonalities with optimal
-control problems, a large existing body of work can be leveraged. This section
-discuses a variety of such techniques and outlines how they can be used to
-solve the \methodAcronym\ system. We focus primarily on 
+\methodAcronym\ displays commonalities with optimal
+control problems, and as such a large existing body of work can be leveraged. In this section we 
+outline a variety of such techniques. We focus primarily on 
 %\KTC{earlier we kept
@@
 \textit{direct} or \textit{indirect}
-methods~\cite{conway_optimalcontrolreview}.  Direct methods eschew the first order optimality conditions
-~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} and, instead, ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
+methods~\cite{conway_optimalcontrolreview}.
+Rather than working with the first order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
 numerically discretizing the system and objective functional and ``transcribing"
 the infinite dimensional problem into a finite-dimensional nonlinear
-programming (optimization) problem. Direct approaches thus ``discretize then optimize."
-Indirect methods, on the other hand, leverage the calculus of variations to
-derive the Euler--Lagrange equations, which comprise the first-order
-optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} associated with the objective 
-functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
-solve the optimization problem
+programming (optimization) problem. Direct approaches thus ``discretize-then-optimize."
+Indirect methods, on the other hand, compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first order optimality conditions of the objective functional).
+Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
 ``indirectly". A variety of discretization/solution techniques are possible for both direct and indirect methods. Collocation methods,
@@
 subspaces. We consider: 
-\begin{itemize} \item \textit{Direct Methods (discretize-then-optimize)}: Discretize and then optimize approaches are outlined for linear multistep methods. 
-\item \textit{Indirect methods (optimize-then-discretize)}: An indirect method leveraging the forward-backward sweep algorithm is outlined 
-to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
+\begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A direct approach leveraging linear multistep methods to directly solve the optimization problem~\eqref{eq:tclsrm}.
+\item \textit{Indirect methods (optimize-then-discretize)}: An indirect approach leveraging the forward-backward sweep algorithm to solve the Euler--Lagrange equations~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}.
+
 \end{itemize} 

### 2019-10-08 section4/spacetime_numerical.tex
through space--time, proofing

@@
 
-\subsection{\spaceTimeAcronym\ trial spaces: direct solution technique}
+\subsection{\spaceTimeAcronym: direct solution approach}
 The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods.
@@
 
-Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis is given by,
+Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
  \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 & \underset{\stgenstatey}{\text{minimize } }
-\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
-&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
-\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
-\mathbb{P}^n(\stateFOMArgnt{1})(0) &
-n=1. \end{cases}\end{split} 
+\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}).%  \\
+%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%n=1. \end{cases}
+\end{split} 
 \end{equation}
-
 % \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 %\end{equation}
-
+It is noted that the boundary conditions are automatically satisfied through the definition of the $\spaceTimeAcronym$\ trial subspace.
 The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss-Newton method.
@@
 \begin{remark}
-For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
+For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and a uniform quadrature weights are used, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
 \end{remark}
 
-\subsection{\spaceTimeAcronym\ trial spaces: indirect solution technique}
-As opposed to the direct solution technique, the indirect solution technique directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime},
+\subsection{\spaceTimeAcronym: indirect solution approach}
+As opposed to the direct approach, the indirect approach directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime},
 \begin{equation}\label{eq:obj_gen_slab2}
 \begin{split}
- & \underset{\stgenstatey}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) \\  
-&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
-\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
-\mathbb{P}^n(\stateFOMArgnt{1})(0) &
-n=1. \end{cases} 
+ & \underset{\stgenstatey}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} \bigg) .%\\  
+%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%n=1. \end{cases} 
 \end{split} 
@@
 Numerically solving the minimization problem requires the introduction of a quadrature rule for 
-discretization of the integral. We introduce $\ncollocSTArg{n} \ge \text{ciel}(\stdim / N )$ collocation points over the $n$th window, 
-$$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \ldots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$
+discretization of the integral. We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ N )$ collocation points over the $n$th window, 
+$$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$
 Leveraging these collocation points, the indirect method for \methodAcronym\ with a \spaceTimeAcronym\ trial subspace yields,
@@
 \begin{split}
-&\underset{\stgenstatey}{\text{minimize }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \\
-&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
-\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
-\mathbb{P}^n(\stateFOMArgnt{1})(0) &
-n=1,
-\end{cases}
+&\underset{\stgenstatey}{\text{minimize }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n}  \bigg) , %\\
+%&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
+%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%n=1,
+%\end{cases}
 \end{split} 
@@
 $$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
-\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptArg{n},\collocPointSTArg{n,i}) \bigg]^T 
+\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T
 \stweightingMatArg{n} 
-\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptArg{n},\collocPointSTArg{n,i}) \bigg].
+\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
 $$
-The nonlinear least-squares problem can again be solved, e.g., via the Gauss-Newton method. 
+The least-squares problem can again be solved, e.g., via the Gauss-Newton method. 
 %The stationary points are given by,
@@
 
-\subsection{\spaceTimeAcronym\ trial space: summary}
-As was just shown, \spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window.
+\subsection{\spaceTimeAcronym\: summary}
+\spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window.
 As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces,

### 2019-10-08 section4/spatial_numerical.tex
through space--time, proofing

@@
 dimensional one by discretizing the state and objective function in time.
-The minimization problem is then reformulated as a nonlinear programming, or
+The minimization problem is then reformulated as a (non)linear programming, or
 optimization, problem. A variety of direct solution approaches exist, 
@@
 \end{align*} 
-Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time-step size, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that
+Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time step, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that
 define the specific type of multistep scheme. Examples of linear multistep schemes include Adams-Bashforth schemes and  
@@
 \begin{remark}\label{remark:LMS}
-For simplicity, we assume the linear multistep method employs time-step instances from only within the current time window.
-In general, the linear multistep method at the $n$th window could be designed to employ multiple time-step instances from previous time windows.
+For notational simplicity, we assume the linear multistep method employs time instances from only within the current time window.
+In general, the linear multistep method at the $n$th window could be designed to employ multiple time instances from previous time windows.
 \end{remark}
@@
 %\end{align*} 
-Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with an objective \textit{function},
+Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
 \begin{equation}\label{eq:obj_lms}
@@
 &\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}+1} \rightarrow
-\RR{}, 
+\RR{}_+, 
 \end{split}
@@
 where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. 
-Over each window, \methodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
+\methodAcronym\ sequentially computes the approximate solutions $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, $n=1,\ldots,\nslabs$, as the solution to the minimization problems
 \begin{align*}
@@
 Defining $\genstateDiscreteArg{n,i} \in \RR{\romdim}$ to be the discrete approximation to the generalized coordinates at the $i$th time instance over the $n$th window, 
-the minimization problem can be written equivalently as,
+the minimization problem can be written equivalently as
 \begin{equation}\label{eq:obj_gen_lms_final}
@@
 %\end{equation}
-The optimization problem takes the form of a \textit{weighted nonlinear least-squares problem}, 
+The optimization problem takes the form of a \textit{weighted (non)linear least-squares problem}, 
 in where the weights correspond the quadrature weights.  
@@
 \end{remark} 
-%%%%%%%%%%%%%%%%COLLOCATION%%%%%%%
-\begin{comment}
-\subsubsection{Collocation (Runge--Kutta) methods}
-Collocation methods comprise another popular technique for temporal discretization. 
-Collocation methods represent the FOM state over each time-step instance $[\timeWindowArg{n}{i-1},\timeWindowArg{n}{i}]$ as,
-\begin{equation}
-\stateFOMArgnt{n}(t) \approx  \collocMatArg{n}{i} \collocBasisArg{n}{i}(t), \qquad t \in [\timeWindowArg{n}{i-1}, \timeWindowArg{n}{i} ],
-\end{equation}
-with the
-continuity and boundary constraints, 
-\begin{equation}\label{eq:colloc_constraint}
-\collocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
-\begin{cases}
-\collocMatArg{n}{i-1}\collocBasisArg{n}{i-1}(\timeWindowArg{n}{i-1}) & n = 1,\ldots,\nslabs, \; i=2,\ldots,\nstepsArg{n},\\ 
-\collocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1,  \\
-\stateFOMIC & n=1, \;  i=1.
-\end{cases}
-\end{equation}
-%\begin{equation}\label{eq:continuity_constraint}
-%\genCollocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
-%%\genstatecollocArg{n}{i}{\timeWindowArg{n}{i-1}} =
-%%\genstatecollocArg{n}{i-1}{\timeWindowArg{n}{i-1}} , \qquad  i =
-%%2,\ldots,\nstepsArg{n}. 
-%\begin{cases} \genstatecollocArg{n}{i-1}{\timeWindowArg{n}{i-1}} & i =
-%2,\ldots,\nstepsArg{n} \\
-%\genstatecollocArg{n-1}{\nstepsArg{n}}{\timeWindowArg{n}{i-1}} & i= 1. \\
-%\end{cases} 
-%\end{equation} 
-In the above, $\collocMatArg{n}{i}  \in \RR{\fomdim \times \collocOrderArg{n}{i}}$ are \textit{space--time coordinates}
-over the $i$th time-step instance of the $n$th window, 
-%$\genstatecollocArgnt{n}{i} : [\timeWindowArg{n}{i-1},\timeWindowArg{n}{i}] \mapsto \RR{\fomdim}$ is the collocated 
-%approximation to the state over the $i$th time-step instance within the $n$th slab
-and $\collocBasisArg{n}{i} :
-[\timeWindowArg{n}{i-1},\timeWindowArg{n}{i}] \mapsto \RR{\collocOrderArg{n}{i}}$ comprise the
-$\collocOrderArg{n}{i}$ 
-\textit{polynomial basis functions} over the $i$th time-step instance of the $nth$ window. 
-%While a variety of temporal basis functions are possible,
-%polynomial basis functions are the most commonly employed and as such we consider these here.
-
-The collocation approach generates an approximation to~\eqref{eq:FOM} by enforcing the FOM ODE at a set of \textit{collocation points}.
-To this end, over the $i$th time-step instance of the $n$th window, we define $\ncollocArg{n}{i} \defeq \collocOrderArg{n}{i}-1$ strictly increasing \textit{collocation points},
-\begin{align*}
-&\collocPointVecArg{n}{i} \defeq \begin{bmatrix} \collocPointArg{n}{i}{1}  &  \collocPointArg{n}{i}{2} \hdots & \collocPointArg{n}{i}{\ncollocArg{n}{i}} \end{bmatrix}, \\
-& \timeWindowArg{n}{i-1} \le \collocPointArg{n}{i}{1} \le \hdots \le \collocPointArg{n}{i}{\ncollocArg{n}{i}} \le \timeWindowArg{n}{i}.
-\end{align*} 
-Enforcing~\eqref{eq:FOM} at the collocation points replaces the FOM ODE system over the $i$th time-step instance of the 
-$n$th window with the FOM O$\Delta$E,
-\begin{align*}
-&\residCollocArg{n,i} ( \collocMatArg{n}{i} ) = \boldsymbol  0, 
-\end{align*}
-where $\collocMatArg{n}{i}$ is subject to the continuity and boundary constraints~\eqref{eq:colloc_constraint} and 
-$\residCollocArg{n,i}$ is the discrete collocation residual over the $i$th time-step instance of the $n$th window,
-\begin{align*}
-\residCollocArg{n,i} &: ( \collocMatyArg{n}{i}) \mapsto \begin{bmatrix} \big[ \collocMatyArg{n}{i} \collocBasisDotArg{n}{i}(\collocPointArg{n}{i}{1}) - 
-\velocity\big(\collocMatyArg{n}{i} \collocBasisArg{n}{i}(\collocPointArg{n}{i}{1})  \big) \big]^T& 
-\cdots &
-\big[ \collocMatyArg{n}{i} \collocBasisDotArg{n}{i}(\collocPointArg{n}{i}{\ncollocArg{n}{i}}) - 
-\velocity\big(\collocMatyArg{n}{i} \collocBasisArg{n}{i}(\collocPointArg{n}{i}{\ncollocArg{n}{i}})) \big]^T 
-\end{bmatrix}^T, \\
-&:  \RR{\fomdim \times \collocOrderArg{n}{i}} \rightarrow \RR{ \fomdim  \ncollocArg{n}{i} }.
-\end{align*}
-Employing a collocation method thus replaces the objective functional~\eqref{eq:obj} with the objective function,
-\begin{align}\label{eq:obj_colloc}
-\objectiveArgC{n} &: (\collocMatyArg{n}{1},\ldots,\collocMatyArg{n}{\nstepsArg{n}} ) \mapsto 
-\frac{1}{2} \sum_{i = 1}^{\nstepsArg{n}} [ \residCollocArg{n,i} ( \collocMatyArg{n}{i} ) \circ  \sqrt{\quadWeightsVecArg{n,i}}]^T   \stweightingMatCollocArg{n}{i}  [  \residCollocArg{n,i}  ( \collocMatyArg{n}{i} ) \circ \sqrt{\quadWeightsVecArg{n,i }}] \\
-&: \RR{\fomdim} \otimes \RR{\collocOrderArg{n}{i} \nstepsArg{n}} \rightarrow \RR{} ,
-\end{align}
-where $\quadWeightsVecArg{n,i} \in \RR{\fomdim \ncollocArg{n}{i}}$ are quadrature weights of the form,
-$$\quadWeightsVecArg{n,i} = \begin{bmatrix} \quadWeightsScalarArg{n}{i}{1} & \quadWeightsScalarArg{n}{i}{1} & \ldots & &  \quadWeightsScalarArg{n}{i}{\ncollocArg{n}{i} } & \quadWeightsScalarArg{n}{i}{\ncollocArg{n}{i}} \end{bmatrix},$$
-and $\circ$ represents the element-wise (Hadamard) product. Additionally,  
-\begin{equation*}
-\stweightingMatCollocArg{n}{i} \defeq \begin{bmatrix}
-\stweightingMatArgt{n}{\collocPointArg{n}{i}{1}} &  & & \\
-& \stweightingMatArgt{n}{\collocPointArg{n}{i}{2}} & & \\ 
-& &  \ddots & \\
-& & &  \stweightingMatArgt{n}{\collocPointArg{n}{i}{\ncollocArg{n}{i}}}
-\end{bmatrix}
-\in  \RR{\fomdim \ncollocArg{n}{i} \times \fomdim \ncollocArg{n}{i}}.
-\end{equation*}
-The \methodAcronym-ROM then  solves the minimization 
-problem,
-\begin{equation}\label{eq:min_colloc}
-\collocMatArg{n}{1},\ldots,\collocMatArg{n}{\nstepsArg{n}} 
- = \underset{\collocMatyArg{n}{1},\ldots,\collocMatyArg{n}{\nstepsArg{n}}  \in \trialspace}{\text{arg\,min } }
-\objectiveArgC{n} (\collocMatyArg{n}{1},\ldots,\collocMatyArg{n}{\nstepsArg{n}} ) \qquad \text{ for } n = 1,2,...,\nslabs,
-\end{equation}
-subject to the boundary conditions,
-\begin{equation*}
-\collocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
-\begin{cases}
-\collocMatArg{n}{i-1}\collocBasisArg{n}{i-1}(\timeWindowArg{n}{i-1}) & n = 2,\ldots,\nslabs, \; i=2,\ldots,\nstepsArg{n} ,\\ 
-\collocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1 , \\
-\approxstateIC & n=1, \;  i=1.
-\end{cases}
-\end{equation*}
-Defining  $\genCollocMatArg{n}{i} \in \RR{\romdim \times \collocOrderArg{n}{i}}$ to be the generalized space--time coordinates over the $i$th time-step instance of the $n$th window,
-the objective function can be equivalently written as,
- \begin{equation*}
-\genCollocMatArg{n}{1},\ldots,\genCollocMatArg{n}{\nstepsArg{n}} 
- = \underset{\genCollocMatyArg{}{1},\ldots,\genCollocMatyArg{}{\nstepsArg{n}} }{\text{argmin } }
-\objectiveArgC{n} (\basisspaceArg{n} \genCollocMatyArg{}{1} + \stateInterceptMat,\ldots,\basisspaceArg{n} \genCollocMatyArg{}{\nstepsArg{n}} + \stateInterceptMat) \qquad \text{ for } n = 1,2,...,\nslabs,
-\end{equation*}
-subject to the boundary and continuity conditions,
-\begin{equation*}
- \genCollocMatArg{n}{i} \collocBasisArg{n}{i}(\timeWindowArg{n}{i-1}) = 
-\begin{cases}
-\genCollocMatArg{n}{i-1}\collocBasisArg{n}{i-1}(\timeWindowArg{n}{i-1}) & n = 2,\ldots,\nslabs, \; i=2,\ldots,\nstepsArg{n} \\ 
-\genCollocMatArg{n-1}{\nstepsArg{n-1}}\collocBasisArg{n-1}{\nstepsArg{n-1}}(\timeWindowArg{n-1}{\nstepsArg{n-1}}) & n = 2,\ldots,\nslabs, \; i=1  \\
-\basisspaceTArg{n} (\approxstateIC - \stateIntercept)& n=1, \;  i=1,
-\end{cases}
-\end{equation*}
-with $\stateInterceptMat \defeq \begin{bmatrix} \stateIntercept & \ldots & \stateIntercept \end{bmatrix} \in \RR{\fomdim \times \collocOrderArg{n}{i} }$.
-The optimization problem again corresponds to a weighted nonlinear least-squares minimization problem. As before, the specification of a numerical 
-integration scheme is required to specify the quadrature weights. For the consistent case of a collocated quadrature rule, in where the integral is approximated 
-via the Lagrange polynomials at the collocation points, the quadrature weights are given by, $\quadWeightsScalarArg{n}{i}{j} = \intSlabArg{n}\lagrangeBasisScalarArg{n}{i}{j}(t)dt,$ where $\lagrangeBasisScalarArg{n}{i}{j}$ are the Lagrange interpolating polynomials,
-$$\lagrangeBasisScalarArg{n}{i}{j}(t) = \underset{\underset{m \ne j}{1 \le m \le \ncollocArg{n}{i}}}{\Pi} \frac{t - \collocPointArg{n}{i}{m}}{\collocPointArg{n}{i}{j}- \collocPointArg{n}{i}{m}}.$$
-
-\begin{remark} %The presented collocation formulation corresponds to minimizing the
-%weighted O$\Delta$E residual over the time slab, in where the O$\Delta$E
-%results from temporal
-%discretization of the FOM ODE according to the temporal basis and collocation points. For 
-%the case where the temporal basis functions are
-%polynomials, the resulting temporal discretization schemes belongs to the family of Runge--Kutta
-%schemes~\cite{1}. 
-Collocation methods belong to the family of Runge--Kutta schemes.
-Alternative temporal discretization schemes, such as
-linear multistep methods, can be obtained by modifying the representation of
-the state and the collocation points~\cite{ajams2015362}.
-\end{remark}
-\end{comment}
-\subsubsection{Solution to the nonlinear least-squares problem through the Gauss--Newton method}
+\subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
 Discretization through linear multistep methods (as well as other techniques) 
-results in a nonlinear least-squares problem.
-A variety of algorithms exist for solving nonlinear  
-least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
+results in a (non)linear least-squares problem.
+A variety of algorithms exist for solving least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
 The numerical experiments presented in this work consider the Gauss--Newton method, and as such we outline this approach here.
 
-To describe the Gauss--Newton method, we first define a ``vectorization" function that takes a tuple of time-local solutions at each time-step instance and outputs 
+To describe the Gauss--Newton method, we first define a ``vectorization" function that takes a tuple of time-local solutions at each time instance and outputs 
 a concatenated vector, 
@@
 
-\subsection{Spatial trial space: indirect solution approach}\label{sec:indirect}
+\subsection{\spatialAcronym: indirect solution approach}\label{sec:indirect}
 In contrast to the direct approach,
@@
 system given by the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_adjoint} comprises a coupled two-point boundary value
-problem. Similar to the direct solution approach, a variety of techniques have
+problem. A variety of techniques have
 been devised to solve two-point boundary value problems of this type. These
@@
 
-\subsubsection{The Forward Backward Sweep Method (FBSM)}\label{sec:FBSM}
+\subsubsection{Forward-backward sweep method (FBSM)}\label{sec:FBSM}
 %\subsubsection{Solution Strategy} Solving Eq.~\ref{eq:lspg_continuous} is
@@
 as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
-time using an initial guess to the costate state to obtain an initial guess to
+time using an initial guess to the costate to obtain an initial guess to
 the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
@@
 backwards problems. As described in  
-the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time-step instances~\eqref{eq:timegrid1} over each time window.
+the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time instances~\eqref{eq:timegrid1} over each time window.
 Linear multistep methods then use this time-grid for temporal discretization of~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}.
@@
 $$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T.$$
-This manipulation allows one to compute the action of the Jacobian on each column of $\basisspace$ by, e.g., a finite difference approximation,
+This manipulation allows one to compute the action of the Jacobian on each column of $\basisspaceArg{n}$ by, e.g., a finite difference approximation,
 $$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$
@@
 \item Automatic Differentiation: A more intrusive, but potentially more efficient, method of computing the action of the Jacobian transpose on a vector is through automatic differentiation (AD).
-AD methods comprise a class of techniques that can be used to numerically evaluate derivatives of functions (e.g., Jacobians, vector-Jacobian products, etc.) by recursively applying the chain rule.
+AD methods comprise a class of techniques that can be used to numerically evaluate derivatives of functions (e.g., Jacobians, vector-Jacobian products) by recursively applying the chain rule.
 The numerical examples presented later in this work leverage AD.

### 2019-10-08 main.tex
working

@@
 \begin{abstract}
-This work proposes a \methodNameLower\ (\methodAcronym) model-reduction
-	formulation for dynamical systems. The proposed approach sequentially minimizes the
-	time-continuous full-order-model residual over a low-dimensional space--time trial
-	subspace defined over an arbitrarily defined time window. The approach comprises a generalization 
+This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
+	of dynamical systems. The proposed approach sequentially minimizes the
+	time-continuous full-order-model residual within a low-dimensional space--time trial
+	subspace over time windows. The approach comprises a generalization 
 of existing model reduction approaches, as particular instances of
@@
 	addresses key deficiencies in existing model-reduction
-	techniques, e.g., the dependence of (space--time) LSPG on the
+	techniques, e.g., the dependence of the (space--time) LSPG approach on the
 	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
-	error bounds for both Galerkin and LSPG.  We consider two types of
+	error bounds for both the Galerkin and LSPG approaches.  We consider two types of
 	space--time trial
-	subspaces within the proposed framework: one that reduces only the spatial
-	dimension of the full-order model, and one that reduces both the spatial and
-	temporal dimensions of the full-order model.  For each type of trial
+	subspaces within the proposed approach: one that associates with spatial projection only, and one that 
+  associates with space--time projection. For each type of trial
 	subspace, we consider two different solution techniques: direct (i.e.,
 	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
-	Numerical experiments conducted using the trial subspaces characterized by spatial
-	dimension reduction demonstrate that the \methodAcronym\
-	approach can yield more accurate solutions that better satisfy the governing equations than
+	Numerical experiments conducted using trial subspaces characterized by spatial
+	projection only demonstrate that the \methodAcronym\
+	approach can yield more accurate solutions with lower space--time residuals than
 	the Galerkin and LSPG approaches.
@@
 particular, changing
-the time step modifies both the
+the time step ($\Delta t$) modifies both the
 time window over which LSPG minimizes the residual as well as the
 time-discretization error of the full-order model on which LSPG is based. 
-In particular, because LSPG and Galerkin are equivalent in the
-limit as the time step goes to zero, the accuracy of LSPG approaches the (poor)
-accuracy of Galerkin as the time step shrinks, but this also has the effect of
-reducing the time-dicretization error of the full-order model.
+As LSPG and Galerkin are equivalent in the
+limit of $\Delta t \rightarrow 0$, the accuracy of LSPG approaches the (sometimes poor)
+accuracy of Galerkin as the time step shrinks.
+For too-large a time step the accuracy of LSPG also degrades.
+It is unclear if this is due to the time-discretization error associated with enlarging the time step, or rather if it is due to the size of the window the residual is being minimized over.
+% but this also has the effect of
+%reducing the time-dicretization error of the full-order model.
 As a
@@
 address the issues incurred by sequential residual minimization in time.
-Space--time residual-minimization approaches compute the space--time
-trajectory within a low-dimensional space--time trial subspace (spanned by
-a small number of space--time basis vectors) that minimizes the full-order-model
-residual over the entire time domain.
-We note that space--time residual minimization yields a fundamentlaly
-different model than sequential residual minimization in time. In particular, space--time
-residual minimization associates with an algebraic system of equations defined
+Existing space--time approaches differ from the classic LSPG and Galerkin approaches in (1) the definition of the space--time trial subspace and (2) the definition of the residual minimization problem.
+First, space--time approaches leverage a space--time trial
+subspace that associates with a space--time projection; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with spatial projection only.
+Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time-step instance (LSPG).
+The result of these differences is that space--time approaches yield an algebraic system of equations defined
 over all space and time, whose solution comprises a \text{fixed} vector of
-(space--time) generalized coordinates; on the other hand, sequential residual
-minimization in time (often) associates with a system of ODEs whose solution
+(space--time) generalized coordinates; on the other hand, spatial projection only approaches 
+generally associates with a system of ODEs whose solution
 comprises a \textit{time-dependent} vector of (spatial) generalized
 coordinates.
-As a result, space--time residual minimization yields models that are equipped with \textit{a priori} error bounds that grow more slowly
-in time, and the resulting \textit{trajectory} exhibits a notion of
-optimality; this implies that the solution exhibits a notion of optimality
-over the entire time domain. 
-\EP{Come back to this, this also has to do with residual minimization}
+
+%Space--time residual-minimization approaches compute the space--time
+%trajectory within a low-dimensional space--time trial trial space that associates with space--time projection (spanned by
+%a small number of space--time basis vectors) that minimizes the full-order-model
+%residual over the entire time domain.
+%We note that methods leveraging a trial subspace that associates with space--time 
+%projection yields a fundamentally different model than methods that leverage a trial subspace that 
+%associates with spatial projection only; in particular space--time
+%projection associates with an algebraic system of equations defined
+%over all space and time, whose solution comprises a \text{fixed} vector of
+%(space--time) generalized coordinates; on the other hand, spatial projection only  
+%generally associates with a system of ODEs whose solution
+%comprises a \textit{time-dependent} vector of (spatial) generalized
+%coordinates.
+%We note that space--time residual minimization yields a fundamentlaly
+%different model than sequential residual minimization in time. In particular, space--time
+%residual minimization associates with an algebraic system of equations defined
+%over all space and time, whose solution comprises a \text{fixed} vector of
+%(space--time) generalized coordinates; on the other hand, sequential residual
+%minimization in time (often) associates with a system of ODEs whose solution
+%comprises a \textit{time-dependent} vector of (spatial) generalized
+%coordinates.
+
+Space--time residual minimization approaches minimize the FOM residual over all of space and time and, as a result, yield models that are equipped with a notion of space--time optimality and \textit{a priori} error bounds that grow more slowly
+in time.
+% and the resulting \textit{trajectory} exhibits a notion of
+%optimality; this implies that the solution exhibits a notion of optimality
+%over the entire time domain. 
+%\EP{Come back to this, this also has to do with residual minimization}
 However, these techniques also suffer from several
@@
 Finally, \methodAcronym\ is formulated for two kinds of space--time trial
-subspaces: one that reduces only the spatial dimension of the full-order
-model (as employed by traditional spatial-projection-based methods), and one that reduces both the spatial and temporal dimensions of the
-full-order model (as employed by space--time methods).
-The above attributes allow that \methodAcronym\ method to be viewed as a
+subspaces: one that associates with spatial projection only (as employed by traditional spatial-projection-based methods), and one that associates with space--time projection (as employed by space--time methods).
+The above attributes allow the \methodAcronym\ approach to be viewed as a
 generalization of existing model-reduction methods, as 
@@
 present work distinguishes itself from Ref.~\cite{constantine_strom} in that
-(1) this work also considers trial subspaces characterized by a reduction of
-the spatial dimension of the FOM only, and (2) we
+(1) this work also considers trial subspaces characterized by spatial projection only, and (2) we
 minimize the time-continuous FOM residual over arbitrary time
@@
 arbitrary time windows, and (3) this work also considers trial subspaces
-characterized by a reduction of the spatial dimension of the FOM only.
+associated with spatial projection only.
 
@@
 \includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
-\caption{Relationship diagram for the \methodAcronym\ framework for model
+\caption{Relationship diagram for the \methodAcronym\ approach for model
 	reduction. \KTC{Inconsistent capitalization. I recommend using (1)
@@
 %\item The development of a time-continuous residual minimization framework for reduced-order models.
-\item The \methodNameLower\ (\methodAcronym) framework for dynamical-system
+\item The \methodNameLower\ (\methodAcronym) approach for dynamical-system
 	model reduction. The approach sequentially minimizes the time-continuous
 		full-order model residual over arbitrary time windows.
-\item Support of two space--time trial subspaces: one that reduces the spatial
-	dimension of the FOM only, and one that reduces both the spatial and
-		temporal dimensions of the FOM. The former case is of particular
+%\item Support of two space--time trial subspaces: one that reduces the spatial
+%	dimension of the FOM only, and one that reduces both the spatial and
+%		temporal dimensions of the FOM. 
+\item Support of two space--time trial subspaces: one that associates with spatial projection only and 
+    one that associates with space--time projection.
+The former case is of particular
 		interest in the \methodAcronym\  contex, as the stationary conditions are
@@
 		\textit{optimize-then-discretize} (i.e., indirect). 
-	\item Derivations of conditions under which the
-		\methodAcronym\ framework recovers Galerkin, LSPG, and space--time LSPG.
-\item Error analysis of the \methodAcronym\ framework using
-	trial subspaces characterized by a reduction of
-the spatial dimension of the FOM only.
+	\item Remarks and derivations of conditions under which the
+		\methodAcronym\ approach recovers Galerkin, LSPG, and space--time LSPG.
+\item Error analysis of the \methodAcronym\ approach using
+	trial subspaces associated with spatial projection.
 	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
@@
 \item Numerical experiments for 
-	trial subspaces characterized by a reduction of
-the spatial dimension of the FOM only, which demonstrate two key findings:
+	trial subspaces associated with spatial projection only, which demonstrate two key findings:
 \begin{itemize}
@@
 space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
-framework. Section~\ref{sec:numerical_techniques}
+approach. Section~\ref{sec:numerical_techniques}
 outlines numerical techniques for solving \methodAcronymROMs, including both direct and
@@
 	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
-		model (\spatialAcronym)}. These 
-	trial subspaces are commonly used
-		in model-reduction methods based on spatial projection followed by time
-		integration (e.g., the Galerkin and LSPG approaches).
+		model (\spatialAcronym)}. These trial spaces are characterized by a \textit{spatial projection} process and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
+%	trial subspaces are commonly used
+%		in model-reduction methods based on spatial projection followed by time
+%		integration (e.g., the Galerkin and LSPG approaches).
 	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
-		model (\spaceTimeAcronym)}. These trial subspaces are commonly employed in
+		model (\spaceTimeAcronym)}. 
+ These trial subspaces are commonly employed in
 		model-reduction methods based on space--time projection (e.g., space--time

### 2019-10-08 main.tex
proofing

@@
 	space--time trial
-	subspaces within the proposed approach: one that associates with spatial projection only, and one that 
-  associates with space--time projection. For each type of trial
+	subspaces within the proposed approach: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model.
+For each type of trial
 	subspace, we consider two different solution techniques: direct (i.e.,
@@
 	Numerical experiments conducted using trial subspaces characterized by spatial
-	projection only demonstrate that the \methodAcronym\
+	dimension reduction demonstrate that the \methodAcronym\
 	approach can yield more accurate solutions with lower space--time residuals than
@@
 dynamical-system trajectory that reside on this trial subspace via, e.g., 
-residual minimization.
+projection of the full-order model or residual minimization.
 
@@
 calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
-function with respect to with the $\dot\ $ operator.
+function with respect to time with the $\dot\ $ operator.
 %Model reduction of nonlinear dynamical systems is a growing research area.
@@
  \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
-	\stateFOMIC,\qquad t \in [0,T]
+	\stateFOMIC,\qquad t \in [0,T],
 \end{equation}
@@
 	problem~\eqref{eq:FOM}, 
-$T \in \mathbb{R}^+$ denotes the final time, 
+$T \in \RRplus$ denotes the final time, 
  $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
@@
 	nonlinear in its first argument. For subsequent exposition, we introduce
-	$\timeSpace$ to denote the set of real-valued functions acting on the time
+	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
 	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
@@
 	subspace. Two types of space--time trial subspaces are commonly used for
-	this purpose: 
+	this purpose:\footnote{For both spatial and space--time ROMs of dynamical systems, all trial subspaces are, strictly speaking, space--time subspaces.} 
 \begin{enumerate} 
 	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
-		model (\spatialAcronym)}. These trial spaces are characterized by a \textit{spatial projection} process and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
+		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
 %	trial subspaces are commonly used
@@
 		model (\spaceTimeAcronym)}.
- These trial subspaces are commonly employed in
-		model-reduction methods based on space--time projection (e.g., space--time
-		LSPG).
+These trial subspaces are characterized by a \textit{space--time} projection and are employed in space--time 
+model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
+% These trial subspaces are commonly employed in
+%		model-reduction methods based on space--time projection (e.g., space--time
+%		LSPG).
 \end{enumerate}
@@
 %\end{equation}
-\subsubsection{The Least-squares Petrov--Galerkin approach}
+\subsubsection{The least-squares Petrov--Galerkin approach}
 Despite its time-instantaneous residual-minimization optimality property 
@@
 \\
-&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim},
+&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim}.
 \end{align*} 
@@
 \end{equation}
-and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
-$\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ and $\stbasis(0) =
+%and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
+and $\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ with $\stbasis(0) =
 \boldsymbol 0$
-denotes the space--time trial basis.\footnote{For simplicity, we consider the case where the trial space is homogeneous on the boundary.} 
+denotes the space--time trial basis. 
 Thus, at any time instance $t\in[0,T]$, ROMs that employ the
@@
 \end{equation}
-where $\stgenstateArg{} \in \RR{\stdim}$ denote the space--time generalized coordinates. 
+where $\stgenstateArg{} \in \RR{\stdim}$ denotes the space--time generalized coordinates. 
 Critically, comparing the approximations arising from \spatialAcronym\ and
@@
 it further.
-\subsubsection{Space--time LSPG projection}  
+\subsubsection{The space--time LSPG approach}  
 Analogously to LSPG, space--time LSPG~\cite{choi_stlspg}
@@
 \end{equation}
-where $\lspgWeightingSTArg{\cdot} \in \RR{\stdim \times N N_t}$, with $\stdim
+where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim
 \le n_{st} \le N N_t$, is a space--time weighting matrix that can be chosen,
@@
 degrees of freedom $\stdim$. This cost is due to the fact that ST-LSPG yields 
-a dense systems that do not expose any natural mechanism for exploiting
+dense systems that do not expose any natural mechanism for exploiting
 the sequential nature of time evolution. Second, it is unclear how these
@@
 employed for the full-order model, as it minimizes the (time-discrete) FOM
-O$\Delta$E residual over all time instance.
+O$\Delta$E residual over all time instances.
 
@@
 exisiting methods can be assessed. Specifically, we look to overcome the
-complex dependence of LSPG projection on the time discretization,
-exponential time growth of the error bounds for Galerkin and LSPG projection,
+complex dependence of LSPG on the time discretization,
+exponential time growth of the error bounds for Galerkin and LSPG,
 the cubic dependence of the computational cost of ST-LSPG on the number of
-degrees of freedom, and the lack of ability for ST-LSPG to perform predictions
-in time. We now describe the proposed \methodNameLower\ framework for this
+degrees of freedom, and the lack of ability for ST-LSPG to perform prediction
+in time. We now describe the proposed \methodNameLower\ \approachKwd\ for this
 purpose. 

### 2019-10-08 section3/general_form.tex
proofing

--- 2019-10-07 section3/general_form.tex
@@
-\section{Windowed least-squares framework}\label{sec:tclspg} 
+\section{Windowed least-squares approach}\label{sec:tclspg} 
 This section
 outlines the proposed \methodNameLower\ 
-(\methodAcronym) framework. In contrast to (1) the Galerkin approach, 
+(\methodAcronym) \approachKwd. In contrast to (1) the Galerkin approach, 
 which minimizes the (time-continuous) FOM ODE residual at a time instance, 
@@
 \end{equation}
-where $\stspaceArg{n}$ is the space--time trial subspace over the $n$th window, $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
+where $\stspaceArg{n}$ is the space--time trial subspace over the $n$th window and $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
 $[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} = \{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$)
-and
-$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
+%and
+%$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. 
+For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
 $$\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n} \subset \stspaceArg{n}, \qquad n=1,\ldots,\nslabs. 
@@
 &\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
-\RR{}_+, 
+\RRplus, 
 \end{split}

### 2019-10-08 section3/space_time.tex
proofing

@@
  \stspaceSTArg{n} \defeq
- \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} .
+ \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} \subseteq \RR{\fomdim} \otimes \timeSpaceArg{n}.
 \end{equation}
-Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
+Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
 For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
@@
 \subsubsection{Stationary conditions} 
-The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$.
+The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$.
 Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
@@
 %n=1. \end{cases} \end{equation}
-For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.
+For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of equations, as opposed to a system of differential equations.

### 2019-10-08 section3/tclsrm_formulation.tex
proofing

@@
 \methodAcronym\ sequentially computes approximate solutions
-$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
-$\approxstateArgnt{n}$ is the solution to the
-minimization problem
-
+$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that satisfy
 \begin{equation*}
 \underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
-\mathcal{J}^n(\stateyArgnt{}), \qquad n = 1,\ldots,\nslabs,
+\mathcal{J}^n(\stateyArgnt{}),
 \end{equation*}
@@
 &\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
-\RR{},
+\RRplus,
 \end{align*}
@@
 &-\basisspaceTArg{n} \big[
-\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
-\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \bigg)
- \bigg( \basisspace \genstateDotArg{n}{t}  -
-\velocity(\veloargsromn) \bigg),  &
+\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
+\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
+ \big( \basisspace \genstateDotArg{n}{t}  -
+\velocity(\veloargsromn) \big),  &
 %\genstateArg{n}{\timeStartArg{n}} = \begin{cases}

### 2019-10-08 main.tex
latest

@@
 of existing model reduction approaches, as particular instances of
-  the methodology recover the Galerkin,
-	least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. In
+  the methodology recover Galerkin,
+	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
 	addition, the approach
 	addresses key deficiencies in existing model-reduction
-	techniques, e.g., the dependence of the (space--time) LSPG approach on the
+	techniques, e.g., the dependence of (space--time) LSPG projection on the
 	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
-	error bounds for both the Galerkin and LSPG approaches.  We consider two types of
+	error bounds for both Galerkin and LSPG projection.  We consider two types of
 	space--time trial
@@
 	approach can yield more accurate solutions with lower space--time residuals than
-	the Galerkin and LSPG approaches.
+	Galerkin and LSPG projection.
 
@@
 Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
-formulation and examined its connections with the Galerkin approach. Critically,
+formulation and examined its connections with Galerkin projection. Critically,
 this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
-Petrov--Galerkin approach applied to the time-continuous
-full-order-model residual, and (2) the LSPG and Galerkin approaches are equivalent
-in the limit as the time step goes to zero (i.e., the Galerkin approach 
+Petrov--Galerkin projection applied to the time-continuous
+full-order-model residual, and (2) the LSPG and Galerkin projection are equivalent
+in the limit as the time step goes to zero (i.e., Galerkin projection 
 minimizes the time-instantaneous full-order-model residual). Numerous
@@
 time-discretization error of the full-order model on which LSPG is based. 
-As LSPG and Galerkin are equivalent in the
+As LSPG and Galerkin projection are equivalent in the
 limit of $\Delta t \rightarrow 0$, the accuracy of LSPG approaches the (sometimes poor)
@@
 First, space--time approaches leverage a space--time trial
-subspace that associates with a space--time projection; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with spatial projection only.
+subspace that associates with a space--time projection operator; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with a spatial projection operator only.
 Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time-step instance (LSPG).
@@
 existing methods can be assessed. In essence, the proposed \textit{\methodNameLower\
-(\methodAcronym)} method sequentially minimizes the FOM
+(\methodAcronym)} \approachKwd\ sequentially minimizes the FOM
 residual over a sequence of arbitrarily defined
@@
 generalization of existing model-reduction methods, as 
-Galerkin, LSPG, and space--time LSPG correspond to specific
+Galerkin, LSPG, and space--time LSPG projection correspond to specific
 instances of the formulation.
@@
 Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
-space--time extension of the LSPG approach that minimizes the time-discrete
+space--time extension of LSPG projection that minimizes the time-discrete
 FOM residual over the entire time domain.  The present work distinguishes
@@
 The paper proceeds as follows: Section~\ref{sec:math}
-outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
-space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
-approach. Section~\ref{sec:numerical_techniques}
+outlines the mathematical setting for the full-order model, along with Galerkin, LSPG, and
+space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
+\approachKwd. Section~\ref{sec:numerical_techniques}
 outlines numerical techniques for solving \methodAcronymROMs, including both direct and
@@
 	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
-		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
+		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection operator} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
 %	trial subspaces are commonly used
@@
 		model (\spaceTimeAcronym)}.
-These trial subspaces are characterized by a \textit{space--time} projection and are employed in space--time 
+These trial subspaces are characterized by a \textit{space--time projection operator} and are employed in space--time 
 model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
@@
 
-\subsubsection{Galerkin approach}
+\subsubsection{Galerkin projection}
 The Galerkin approach reduces the number of equations in
@@
 %\end{equation}
-\subsubsection{The least-squares Petrov--Galerkin approach}
+\subsubsection{Least-squares Petrov--Galerkin projection}
 Despite its time-instantaneous residual-minimization optimality property 
@@
 it further.
-\subsubsection{The space--time LSPG approach}  
+\subsubsection{Space--time LSPG projection} 
 Analogously to LSPG, space--time LSPG~\cite{choi_stlspg}

### 2019-10-08 section3/general_form.tex
latest

@@
 outlines the proposed \methodNameLower\ 
-(\methodAcronym) \approachKwd. In contrast to (1) the Galerkin approach, 
+(\methodAcronym) \approachKwd. In contrast to (1) Galerkin projection, 
 which minimizes the (time-continuous) FOM ODE residual at a time instance, 
-(2) the LSPG approach, 
+(2) LSPG projection, 
 which minimizes
 the (time-discrete) FOM O$\Delta$E residual 
-over a time step, and (3) the ST-LSPG approach,  
+over a time step, and (3) ST-LSPG projection, 
 which minimizes
 the  FOM O$\Delta$E residual 
-over the entire time interval, the proposed \methodAcronym\ method sequentially minimizes the 
+over the entire time interval, the proposed \methodAcronym\ \approachKwd\ sequentially minimizes the 
 time-continuous FOM ODE residual over a series of arbitrarily defined

### 2019-10-08 section3/tclsrm_formulation.tex
latest

@@
 subject to the boundary conditions.
-For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
+For \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
 \begin{align*}\

### 2019-10-08 section4/numerical_methods.tex
latest

@@
 of additional approaches.
-Section~\ref{sec:direct}  outlines direct solution approaches based on linear
-multistep methods, while Section~\ref{sec:indirect} outlines the indirect
-solution approach.
+Section~\ref{sec:direct} outlines a direct solution approach based on linear
+multistep methods, while Section~\ref{sec:indirect} outlines an indirect
+solution approach based on the forward-backward sweep algorithm.
 

### 2019-10-08 section4/spatial_numerical.tex
latest

@@
 optimization, problem. A variety of direct solution approaches exist, 
-including collocation approaches, orthogonal polynomial (spectral  method) 
-approaches, and genetic algorithms.  
-\KTC{is this a
-spectral method, or the spectral method? clarify} \EP{I don't understand what you mean. What is ``the" spectral method}
+including collocation approaches, spectral  methods, and genetic algorithms.
 
@@
 (2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
-To achieve this agenda, first define a discrete temporal grid 
-by introducing
+To achieve this agenda, first define a discrete time grid 
+over each window by introducing
 $\nstepsArg{n}+1$ strictly increasing time instances described by
 $\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with
-\EP{These should be strictly increasing}
 %\KTC{use cdots, not

### 2019-10-09 main.tex
latest

@@
 \end{equation}
-where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RR{}$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,t) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,t)$.
+where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RRplus$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$.
 We now introduce the function $\statez: [a,b] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(a) = \statez_a$.
@@
 \end{align*}
-We note that $\statezBar$ obeys the same boundary conditions as $\statez$ since $\variationArg{a} = \bz$.
-We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar(t,\cdot)$,
+We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{a} = \bz$.
+We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 
+$\statezDotBar \equiv d\statezBar / d\timeDummy$ in the second argument,
 $$
@@
 $$
-
 %$$
@@
 %$$ \int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) \big) \bigg] = 0. $$
-We first evaluate the derivative. First using the chain rule,
+Using the chain rule,
 $$ 
@@
 $$
-Noting that,
+Noting that
 $$\frac{\partial \statezBar}{\partial \delta} (t,\cdot)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(t,\cdot)= \variationDot(t),$$
-we have,
+where $\variationDot \equiv d \variation / d\timeDummy$, 
+we have
 $$
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \variationDot(t) \bigg]dt.
+= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variationDot(t) \bigg]dt.
 $$
@@
 \end{multline*}
-\EP{Here}
 Noting that $\variationArg{a} = \bz$,
 \begin{multline*}
-\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg] dt =  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) , \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon) ,b \big) \variation(b) .
+%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg] dt 
+\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
+=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) .
 \end{multline*}
@@
 $$
-  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b;\epsilon),\statezDotBar(b;\epsilon),b \big) \variation(b)  = 0.
+  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b;\epsilon),b \big) \variation(b)  = 0.
 $$
@@
 
-\section{Evaluation of Gradients in the Euler--Lagrange Equations for \methodAcronym}\label{appendix:vector_calc} 
+\section{Evaluation of gradients in the Euler--Lagrange Equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
+We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces.
+Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.
 To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:el1} for the integrand,
@@
  \minintegrand & \vcentcolon
-(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,t) \mapsto \frac{1}{2} \big[
+(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,\timeDummy) \mapsto \frac{1}{2} \big[
 \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
-+ \stateIntercept,t) \big]^T \stweightingMatArg{} \big[
++ \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[
 \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
@@
 \begin{align*}
- \minintegrand(\genstateyDiscrete,\dot{\genstateyDiscrete},t)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,t) \big]^T \stweightingMatArg{n} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\
- &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{n}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big].
+ \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\
+ &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
 \end{align*}
@@
 \begin{equation}\label{eq:int_expand}
- \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,t)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{n}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big].
+ \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
 \end{equation}
 For notational purposes, we write the above as,
-$$  \minintegrand(\genstateyDiscrete,\dot{\genstateyDiscrete},t) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,t) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,t) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,t),$$
+$$  \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy),$$
 where
 \begin{align*}
-&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,t) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{n}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
-&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,t) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big], \\
-&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,t) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] . 
+&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
+&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
+&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
 \end{align*}
@@
 \begin{align*}
-\frac{\partial \minintegrand_2}{\partial \genstateyDiscrete}  &= -\frac{\partial}{\partial \genstateyDiscrete} \bigg( \big[ \basisspace \genstateyDiscreteDot \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] \bigg) \\
-&= - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \frac{\partial}{\partial \genstateyDiscreteArgnt{}} \bigg( \big[ \velocity(\veloargsromy) \big]  \bigg)\\
- &= -\big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n}   \frac{\partial \velocity}{\partial \stateyDiscrete}\frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
-& = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace \\
-&= -[\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete}  \basisspace.
+\frac{\partial \minintegrand_2}{\partial \genstateyDiscrete}  &= -\frac{\partial}{\partial \genstateyDiscrete} \bigg( \big[ \basisspace \genstateyDiscreteDot \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] \bigg) \\
+&= - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \frac{\partial}{\partial \genstateyDiscreteArgnt{}} \bigg( \big[ \velocity(\veloargsromy) \big]  \bigg)\\
+ &= -\big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}   \frac{\partial \velocity}{\partial \stateyDiscrete}\frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
+& = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace \\
+&= -[\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete}  \basisspace.
 \end{align*}
@@
 \begin{align}\label{eq:i3sol}
-\frac{\partial  \minintegrand_3}{\partial \genstateyDiscrete}  &= \frac{1}{2} \frac{\partial }{\partial \genstateyDiscrete} \bigg( \big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n}\big[ \velocity(\veloargsromy) \big] \bigg) \\
-&=  [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
- &=   [\velocity(\veloargsromy) ]^T  \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
+\frac{\partial  \minintegrand_3}{\partial \genstateyDiscrete}  &= \frac{1}{2} \frac{\partial }{\partial \genstateyDiscrete} \bigg( \big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{}\big[ \velocity(\veloargsromy) \big] \bigg) \\
+&=  [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
+ &=   [\velocity(\veloargsromy) ]^T  \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
 \end{align}
@@
 $$
- \frac{\partial \minintegrand}{\partial \boldsymbol \genstateyDiscrete} = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace +  [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
+ \frac{\partial \minintegrand}{\partial \boldsymbol \genstateyDiscrete} = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace +  [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
 $$%
@@
 \frac{\partial \minintegrand_1}{\partial \genstateyDiscreteDotArgnt{}} &=
-\frac{1}{2} \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} }\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \stweightingMatArg{n}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \bigg) \\
-&= [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{n} \basisspace.
+\frac{1}{2} \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} }\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \bigg) \\
+&= [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \basisspace.
 \end{align*}
@@
 \frac{\partial \minintegrand_2}{\partial \genstateyDiscreteDotArgnt{} } &=
-\frac{\partial}{\partial \genstateyDiscreteDotArgnt{}  } \bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \bigg)\\
- &= \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} } \bigg( [\genstateyDiscreteDotArgnt{}]^T \bigg) \basisspace^T  \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] \\ 
- &= \bigg[  \basisspace^T  \stweightingMatArg{n}  \velocity(\veloargsromy )   \bigg]^T \\
- &= [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace.
+\frac{\partial}{\partial \genstateyDiscreteDotArgnt{}  } \bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \bigg)\\
+ &= \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} } \bigg( [\genstateyDiscreteDotArgnt{}]^T \bigg) \basisspace^T  \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] \\ 
+ &= \bigg[  \basisspace^T  \stweightingMatArg{}  \velocity(\veloargsromy )   \bigg]^T \\
+ &= [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace.
 \end{align*}
@@
 where it is noted that $\mathcal{I}_3$ only depends on $\genstateyDiscrete$. We thus have,
-$$\frac{\partial \minintegrand}{\partial \dot{ \genstateyDiscrete} } =   [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{n} \basisspace -  [\velocity(\veloargsromy ) ]^T \stweightingMatArg{n} \basisspace.$$
+$$\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot } =   [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \basisspace -  [\velocity(\veloargsromy ) ]^T \stweightingMatArg{} \basisspace.$$
+\EP{Here}
 Combining all terms, the Euler--Lagrange equation reads,

### 2019-10-09 cavity.tex
updated appendix

@@
-\subsection{Cavity Flow}
+\subsection{Cavity flow}
 The next numerical example considers hyper-reduced ROMs of a viscous, compressible flow in a two-dimensional cavity. 
@@
 
-\subsubsection{Description of FOM and Generation of Trial Space}
+\subsubsection{Description of FOM and generation of \spatialAcronym\ trial subspace}
 The full-order model comprises a discontinuous-Galerkin discretization. The discretization 
@@
 
-\subsubsection{Description of Reduced-Order Models}
+\subsubsection{Description of reduced-order models}
 Hyper-reduced ROMs based on the Galerkin, collocated Least--Squares Petrov--Galerkin, and \methodAcronym\ approaches are 
@@
 \begin{itemize}
-\item \textbf{Galerkin ROM with Collocation}: The Galerkin ROM with collocation is obtained through Galerkin projection of the FOM. Unless 
+\item \textit{Galerkin ROM with collocation}: The Galerkin ROM with collocation is obtained through Galerkin projection of the FOM. Unless 
 otherwise noted, the  Galerkin ROM is evolved in time with the Crank--Nicolson time scheme at a time step of $\Delta t =0.1$.
 
-\item \textbf{LSPG ROM with Collocation}: The LSPG ROM with collocation is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
+\item \textit{LSPG ROM with collocation}: The LSPG ROM with collocation is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
 discretization. Unless otherwise noted, the time step for the LSPG ROM is $\Delta t  = 0.1$. As in the previous example, 
@@
 
-\item \textbf{\methodAcronymROMs\ with Collocation}: \methodAcronymROMs\ are solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time-step size of 
+\item \textit{\methodAcronymROMs\ with collocation}: \methodAcronymROMs\ are solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time-step size of 
 $\Delta t = 0.1$. The Gauss--Newton algorithm is deemed converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation. Uniform quadrature 
@@
 
-\subsubsection{Numerical Results}
+\subsubsection{Numerical results}
 We first assess the performance of \methodAcronymROMs\ with varying window sizes. To this end, we consider \methodAcronymROMs\ with uniform window 

### 2019-10-09 main.tex
updated appendix

@@
 
-\section{Selection of Sampling Points}
+\section{Selection of sampling points}
 To construct the sampling point matrix used for hyper-reduction in the second numerical experiment, we employ q-sampling~\cite{qdeim_drmac} and the sample 
@@
 
-\section{Derivation of the Euler--Lagrange Equations}\label{appendix:eulerlagrange}
-This section details the derivation of the Euler Lagrange equations. To this end, we consider the generic functional of the form,
+\section{Derivation of the Euler--Lagrange equations}\label{appendix:eulerlagrange}
+This section details the derivation of the Euler--Lagrange equations. To this end, we consider the generic functional of the form,
 \begin{equation}\label{eq:el_gen}
-\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_a^b \integrand(\statey(t),\stateyDot(t),t)  dt,
-\end{equation}
-where $\integrand: \RR{M} \times \RR{M} \times[a,b] \rightarrow  \RRplus$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$.
-We now introduce the function $\statez: [a,b] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(a) = \statez_a$.
-We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$ with the boundary condition $\variationArg{a} = \bz$ and define a variation from the extremal by,
+\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_0^T \integrand(\statey(t),\stateyDot(t),t)  dt,
+\end{equation}
+where $\integrand: \RR{M} \otimes \timeSpaceArg{} \times \RR{M} \otimes \timeSpaceArg{} \times[0,T] \rightarrow  \RRplus$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$.
+We now introduce the function $\statez: [0,T] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(0) = \statez_0$.
+We introduce an arbitrary function $\variation : [0,T] \rightarrow \RR{M}$ with the boundary condition $\variationArg{0} = \bz$ and define a variation from the extremal by,
 \begin{align*}
@@
 \end{align*}
-We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{a} = \bz$.
+We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{0} = \bz$.
 We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)= 
-\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt .
+\int_0^T \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt .
 $$
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variationDot(t) \bigg]dt.
+= \int_0^T \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variationDot(t) \bigg]dt.
 $$
@@
 %\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
-=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(a,\epsilon),\statezDotBar(a,\epsilon) ,a\big) \variation(a) .
+=  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon) ,T \big) \variation(T)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(0,\epsilon),\statezDotBar(0,\epsilon) ,0\big) \variation(0) .
 \end{multline*}
-Noting that $\variationArg{a} = \bz$,
+Noting that $\variationArg{0} = \bz$,
 \begin{multline*}
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) .
+=  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon) ,T \big) \variation(T) .
 \end{multline*}
@@
 $$
-  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b;\epsilon),b \big) \variation(b)  = 0.
+  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon),T \big) \variation(T)  = 0.
 $$
@@
  $$
-  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b) ,b\big) \variation(b)  = 0.
+  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(T),\statezDot(T) ,T\big) \variation(T)  = 0.
 $$
@@
 \begin{equation}\label{eq:euler_lagrange_appendix}
-\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t) \big) \bigg) \bigg]^T = \bz
+\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz
 \end{equation}
@@
 
-\section{Evaluation of gradients in the Euler--Lagrange Equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
+\section{Evaluation of gradients in the Euler--Lagrange equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
 We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces.
 Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.
-To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:el1} for the integrand,
+To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand,
 \begin{align*}\label{eq:integrand_apx}
@@
 & = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace \\
-&= -[\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete}  \basisspace.
-\end{align*}
+&= -[\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete}  \basisspace,
+\end{align*}
+where we have suppressed the arguments of the Jacobian for simplicity; e.g., formally 
+$$\frac{\partial \velocity}{\partial \stateyDiscrete} : (\statewDiscrete,\timeDummy) \mapsto \frac{ \partial \velocity }{ \partial \stateyDiscrete} (\statewDiscrete, \timeDummy).$$ 
 For $\minintegrand_3$,
@@
 $$\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot } =   [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \basisspace -  [\velocity(\veloargsromy ) ]^T \stweightingMatArg{} \basisspace.$$
-\EP{Here}
-Combining all terms, the Euler--Lagrange equation reads,
+Combining all terms and evaluating at $(\genstateArg{}{t},\genstateDotArg{}{t},t)$,
 \begin{multline*}
- \frac{\partial \minintegrand}{\partial \genstateyDiscrete}(\genstateArg{n}{t},\genstateDotArg{n}{t},t)  - \frac{d}{dt} \bigg[ \frac{\partial \minintegrand}{\partial \genstateyDiscreteDotArgnt{}} (\genstateArg{n}{t},\genstateDotArg{n}{t},t) \bigg] =  - \big[ \basisspace \genstateDotArg{n}{t}  \big]^T \stweightingMatArg{n} \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \big] \basisspace \\ + [\velocity(\veloargsromn) ]^T \stweightingMatArg{n} \big[ \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \big] \basisspace - \frac{d}{dt} \bigg[  [\genstateDotArg{n}{t} ]^T  \basisspace^T \stweightingMatArg{n} \basisspace - [\velocity(\veloargsromn) ]^T \stweightingMatArg{n} \basisspace \bigg]  = \boldsymbol 0.
+ \frac{\partial \minintegrand}{\partial \genstateyDiscrete}(\genstateArg{}{t},\genstateDotArg{}{t},t)  - \frac{d}{dt} \bigg[ \frac{\partial \minintegrand}{\partial \genstateyDiscreteDotArgnt{}} (\genstateArg{}{t},\genstateDotArg{}{t},t) \bigg] =  - \big[ \basisspace \genstateDotArg{}{t}  \big]^T \stweightingMatArg{} \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \big] \basisspace + \\  [\velocity(\veloargsrom) ]^T \stweightingMatArg{} \big[ \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \big] \basisspace - \frac{d}{dt} \bigg[  [\genstateDotArg{}{t} ]^T  \basisspace^T \stweightingMatArg{} \basisspace - [\velocity(\veloargsrom) ]^T \stweightingMatArg{} \basisspace \bigg]  = \boldsymbol 0.
 \end{multline*}
 To put this in a more recognizable form, we can pull out the common factor in the first two terms,
-$$ - \bigg( \big[ \basisspace \genstateDotArg{n}{t} \big]^T  -  [\velocity(\veloargsromn)] ^T \bigg) \stweightingMatArg{n}
- \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisspace - \frac{d}{dt} \bigg[  [\genstateDotArg{n}{t}]^T \basisspace^T \stweightingMatArg{n} \basisspace -  
-[\velocity(\veloargsromn) ]^T \stweightingMatArg{n} \basisspace  \bigg] = \boldsymbol 0.$$
+\begin{multline*}
+- \bigg( \big[ \basisspace \genstateDotArg{}{t} \big]^T  -  [\velocity(\veloargsrom)] ^T \bigg) \stweightingMatArg{}
+ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \basisspace -  \\
+\frac{d}{dt} \bigg[  [\genstateDotArg{}{t}]^T \basisspace^T \stweightingMatArg{} \basisspace -  
+[\velocity(\veloargsrom) ]^T \stweightingMatArg{} \basisspace  \bigg] = \boldsymbol 0.
+\end{multline*}
 Taking the transpose to put into the common column major format,
-$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)  \big]^T \stweightingMatArg{n} \bigg( \big[ \basisspace \genstateDotArg{n}{t}\big]  -  \velocity(\veloargsromn)  \bigg) -  \frac{d}{dt} \bigg[  \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArg{n}{t}  - \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn)   \bigg] = \bz.
+$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom)  \big]^T \stweightingMatArg{} \bigg( \big[ \basisspace \genstateDotArg{}{t}\big]  -  \velocity(\veloargsrom)  \bigg) -  \frac{d}{dt} \bigg[  \basisspace^T \stweightingMatArg{} \basisspace \genstateDotArg{}{t}  - \basisspace^T \stweightingMatArg{} \velocity(\veloargsrom)   \bigg] = \bz.
  $$
 We now factor the second term,
-$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)  \big]^T \stweightingMatArg{n} \bigg(  \basisspace \genstateDotArg{n}{t}  -  \velocity(\veloargsromn) \bigg) -  \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg[   \basisspace \genstateDotArg{n}{t}  - \velocity(\veloargsromn) \bigg] = \bz.
-$$
-Gathering terms and multiplying by negative one, the final form of the Euler Lagrange equations are obtained,
-$$ \bigg[\basisspace^T \big[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \big]^T \stweightingMatArg{n} + \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{n}{t} -  \velocity(\veloargsromn) \bigg) = \bz.
-$$
-This is the \methodAcronym-ROM.
-Note that this is a second order equation and can be written as two separate first order equations.
-Equation~\ref{eq:clspg_2ord} can be written equivalently as two first order equations by introducing an auxiliary ``costate" variable.
-Defining the costate as,
+$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom)  \big]^T \stweightingMatArg{ } \bigg(  \basisspace \genstateDotArg{ }{t}  -  \velocity(\veloargsrom) \bigg) -  \basisspace^T \stweightingMatArg{ } \frac{d}{dt} \bigg[   \basisspace \genstateDotArg{}{t}  - \velocity(\veloargsrom) \bigg] = \bz.
+$$
+Gathering terms and multiplying by negative one, the final form of the Euler--Lagrange equations are obtained,
+\begin{equation}\label{eq:el_secondorder}
+ \bigg[\basisspace^T \big[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \big]^T \stweightingMatArg{} + \basisspace^T \stweightingMatArg{} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{}{t} -  \velocity(\veloargsrom) \bigg) = \bz.
+\end{equation} 
+This is the \methodAcronym-ROM. Note that this is a second order equation and can be written as two separate first order equations. Defining the ``costate" as,
 \begin{equation*}
 %\adjoint^n(t) \defeq \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArg{n}{t}  -  \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) ,
-\adjoint^n(t) \defeq  \genstateDotArg{n}{t}  -  \mass^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) ,
+\adjointArg{}{\timeDummy}  : \timeDummy \mapsto \genstateDotArg{}{\timeDummy}  -  \mass^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace \genstateArg{}{t} + \stateIntercept, \timeDummy) ,
 \end{equation*}
-we can manipulate Eq.~\ref{eq:clspg_2ord} as follows: %First, to enable hyper-reduction later on, we leverage the fact that $\stweightingMat^2 = \stweightingMat$,
+we can manipulate Eq.~\eqref{eq:el_secondorder} as follows: %First, to enable hyper-reduction later on, we leverage the fact that $\stweightingMat^2 = \stweightingMat$,
 %$$
@@
 First, we add and subtract the first term multiplied by $\basisspace [\massArgnt{n}]^{-1}\basisspace^T \stweightingMat$, 
-$$
- \bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1} \basisspace^T\stweightingMatArg{n} + \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  + \basisspace^T \stweightingMatArg{n}  \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}^n   -  \velocity(\veloargsromn) \bigg) = \bz.
-$$
-Pulling out the term multiplied by the positive portion of $\basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}$,
 \begin{multline*} 
-\bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1}  \basisspace^T  \stweightingMatArg{n} \bigg)  + \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  \basisspace^T \stweightingMatArg{n} +   \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg) = \bz.
+\bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1} \basisspace^T\stweightingMatArg{} + \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  + \basisspace^T \stweightingMatArg{}  \frac{d}{dt} \bigg] \\
+\bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg) = \bz.
+\end{multline*}
+Pulling out the term multiplied by the positive portion of $\basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}$,
+\begin{multline*} 
+\bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1}  \basisspace^T  \stweightingMatArg{} \bigg)  + \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  \basisspace^T \stweightingMatArg{} +   \basisspace^T \stweightingMatArg{} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg) = \bz.
 \end{multline*}
@@
 \begin{multline*}
-\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1}  \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg)  + \\
-\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} +   \basisspace^T \stweightingMatArg{n} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{n}{t}  -  \velocity(\veloargsromn) \bigg) = \bz.
+\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1}  \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg)  + \\
+\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} +   \basisspace^T \stweightingMatArg{} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{}{t}  -  \velocity(\veloargsrom) \bigg) = \bz.
 \end{multline*}
-Pulling $\mass^{-1} \basisspace^T \stweightingMatArg{n}$ inside the parenthesis on the second term,
+Pulling $\mass^{-1} \basisspace^T \stweightingMatArg{}$ inside the parenthesis on the second term,
 \begin{multline*}
-\basisspace^T \bigg[\frac{\partial \velocity}{\partial \state^n}(\veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg)  + \\
-\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace  +   \mass \frac{d}{dt} \bigg] \bigg( \genstateDotArg{n}{t}  - \mass^{-1} \basisspace^T \stweightingMatArg{n}  \velocity(\veloargsromn) \bigg) = \bz.
+\basisspace^T \bigg[\frac{\partial \velocity}{\partial \state^n}(\veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg)  + \\
+\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace  +   \mass \frac{d}{dt} \bigg] \bigg( \genstateDotArg{}{t}  - \mass^{-1} \basisspace^T \stweightingMatArg{}  \velocity(\veloargsrom) \bigg) = \bz.
 \end{multline*}
-By definition, the term inside the parenthesis of the second term is $\adjointArg{n}{t}$,
+By definition, the term inside the parenthesis of the second term is $\adjointArg{}{t}$,
 \begin{multline*}
-\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}    -  \velocity(\veloargsromn) \bigg)  +   \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace \adjointArg{n}{t}  +  \massArgnt{n} \frac{d}{dt} \adjointArg{n}{t}= \bz.
+\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}    -  \velocity(\veloargsrom) \bigg)  +   \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace \adjointArg{}{t}  +  \massArgnt{} \frac{d}{dt} \adjointArg{}{t}= \bz.
 \end{multline*}
@@
 \begin{multline*}
-\mass \frac{d }{dt}\adjointArg{n}{t} + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsromn ) \bigg]^T \stweightingMatArg{n} \basisspace  \adjointArg{n}{t}  \\
-= - \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)\bigg]^T \stweightingMatArg{n}\bigg( \mathbf{I} - \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \bigg)  \bigg(  \basisspace \genstateDotArg{n}{t}   -  \velocity(\veloargsromn) \bigg).
+\mass \frac{d }{dt}\adjointArg{}{t} + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsrom ) \bigg]^T \stweightingMatArg{} \basisspace  \adjointArg{}{t}  \\
+= - \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom)\bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg).
 \end{multline*}
@@
 \begin{align*}\label{eq:lspg_continuous_appendix}
-& \mass \frac{d}{dt}\genstateArg{n}{t}  -  \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) =  \mass \adjointArg{n}{t},\\
-&\mass \frac{d}{dt} \adjointArg{n}{t}  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \bigg]^T \stweightingMatArg{n} \basisspace \adjointArg{n}{t} = \\
-& -\basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \big] ^T \stweightingMatArg{n}  \bigg( \mathbf{I} -   \basisspace [\massArgnt{n}]^{-1} \basisspace^T   \stweightingMatArg{n} \bigg)\bigg( \basisspace \genstateDotArg{n}{t}   -   \velocity(\veloargsromn) \bigg) .
-\end{align*}
-\section{Evaluation of Gradients for Optimal Control Formulation}\label{appendix:optimal_control}
+& \mass \frac{d}{dt}\genstateArg{}{t}  -  \basisspace^T \stweightingMatArg{} \velocity(\veloargsrom) =  \mass \adjointArg{}{t},\\
+&\mass \frac{d}{dt} \adjointArg{}{t}  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace \adjointArg{}{t} = \\
+& -\basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \big] ^T \stweightingMatArg{}  \bigg( \mathbf{I} -   \basisspace [\massArgnt{}]^{-1} \basisspace^T   \stweightingMatArg{} \bigg)\bigg( \basisspace \genstateDotArg{}{t}   -   \velocity(\veloargsrom) \bigg) .
+\end{align*}
+\section{Evaluation of gradients for optimal control formulation}\label{appendix:optimal_control}
 When formulated as an optimal control problem of Lagrange type, the gradients of the Hamiltonian with respect to the state, controller, and costate 
@@
 \begin{align*}
-\hamiltonian^n \; &: \;  (\genstateyDiscreteArgnt{},\adjointDiscreteDumArgnt{},\controllerDiscreteDumArgnt{},t) \mapsto 
- \adjointDiscreteDumArgnt{T} \bigg[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \bigg] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t) \\
+\hamiltonian^n \; &: \;  (\genstateyDiscreteArgnt{},\adjointDiscreteDumArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \mapsto 
+ \adjointDiscreteDumArgnt{T} \bigg[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \bigg] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \\
 &: \; \RR{\romdim} \times \RR{\romdim} \times \RR{\romdim} \times [0,T] \rightarrow \RR{},
@@
 \begin{align*}
- \objectiveControlArg{} &:  (\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t)
-\mapsto \frac{1}{2} \bigg[ \basisspace \bigg(  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \bigg) -
+ \objectiveControlArg{} &:  (\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)
+\mapsto \frac{1}{2} \bigg[ \basisspace \bigg(  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \bigg) -
 \velocity(\veloargsromy) \bigg]^T
-\stweightingMatArg{n}  \\ & \hspace{1.5 in}\bigg[ \basisspace \bigg(
-[\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{}
+\stweightingMatArg{}  \\ & \hspace{1.5 in}\bigg[ \basisspace \bigg(
+[\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{}
 \bigg) - \velocity( \veloargsromy ) \bigg]
@@
 As described in Section~\ref{sec:optimal_control}, to derive the stationary conditions of the \methodAcronym\ objective function, we require evaluating the following gradients,
-$ \frac{\partial \hamiltonianArg{n}}{\partial \adjointDiscreteDumArgnt{}{}}, \;   \frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}}$, and $\frac{\partial \hamiltonianArg{n}}{\partial \controllerDiscreteDumArgnt{}{}}.$  Starting with $\frac{\partial \hamiltonianArg{n}}{\partial \adjointDiscreteDumArgnt{}{}}$, we have,
-$$
-\bigg[\frac{\partial \hamiltonianArg{n}}{\partial \adjointDiscreteDumArgnt{}{}}\bigg]^T = [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} .
-$$
-Next, we address $ \frac{\partial \hamiltonianArg{n}}{\partial \genstateyArgnt{}{}}$. 
-\begin{align*}
-  \frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}} &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
- &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy
-) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
- &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg( \velocity(\veloargsromy)\bigg) \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
- &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscreteArgnt{}{}}{\partial  \genstateyDiscreteArgnt{}{} } \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
- &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg].
-\end{align*}
-To evaluate $\frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg]$, we leverage the previous result~\eqref{eq:int_expand} and insert $\genstateyDiscreteDotArgnt{} =  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} .$ This leads to the expression for the expanded Lagrangian,
+$ \frac{\partial \hamiltonianArg{}}{\partial \adjointDiscreteDumArgnt{}{}}, \;   \frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}}$, and $\frac{\partial \hamiltonianArg{}}{\partial \controllerDiscreteDumArgnt{}{}}.$  Starting with $\frac{\partial \hamiltonianArg{}}{\partial \adjointDiscreteDumArgnt{}{}}$, we have,
+$$
+\bigg[\frac{\partial \hamiltonianArg{}}{\partial \adjointDiscreteDumArgnt{}{}}\bigg]^T = [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} .
+$$
+Next, we address $ \frac{\partial \hamiltonianArg{}}{\partial \genstateyArgnt{}{}}$. 
+\begin{align*}
+  \frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}} &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
+ &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy
+) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
+ &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg( \velocity(\veloargsromy)\bigg) \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
+ &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscreteArgnt{}{}}{\partial  \genstateyDiscreteArgnt{}{} } \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
+ &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg].
+\end{align*}
+To evaluate $\frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg]$, we leverage the previous result~\eqref{eq:int_expand} and insert $\genstateyDiscreteDotArgnt{} =  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t} .$ This leads to the expression for the expanded Lagrangian,
 \begin{multline*}
- \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) = \frac{1}{2}\big[\basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T  \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t}\big] \\  -  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big].
+ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) = \frac{1}{2}\big[\basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T  \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t}\big] \\  -  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
 \end{multline*}
@@
 $$
-\objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg)  = 
-\objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg)  + 
-\objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg)  + 
-\objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) ,$$
+\objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)  = 
+\objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)  + 
+\objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)  + 
+\objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) ,$$
 where
 \begin{align*}
-& \objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \equiv  \frac{1}{2}\big[\basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t}  \big]^T  \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy)   +  \basisspace [\massArgnt{n}]^{-1}\controllerDiscreteDumArg{}{t} \big],\\
-& \objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \defeq -  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{n}]^{-1} \controllerDiscreteDumArgnt{n}\big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] ,\\
-& \objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \defeq \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big] .
+& \objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) =  \frac{1}{2}\big[\basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{}{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy)   +  \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{}{} \big],\\
+& \objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) = -  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{}]^{-1} \controllerDiscreteDumArgnt{}\big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] ,\\
+& \objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) = \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] .
 \end{align*}
@@
 \frac{\partial  \objectiveControlArg{}_1 }{\partial \genstateyDiscreteArgnt{}{}} &= 
-\frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) +  \\
-& \qquad \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{n}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) + \\
-& \qquad \frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \controllerDiscreteDumArgnt{}\bigg)
+\frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) +  \\
+& \qquad \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) + \\
+& \qquad \frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \controllerDiscreteDumArgnt{}\bigg)
 ,\\ 
 &= 
-[\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace + 
-[\controllerDumArgnt{n}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \statey} \basisspace , \\ 
+[\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace + 
+[\controllerDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \statey} \basisspace , \\ 
 %&= 
@@
 &= 
-\bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \bigg)\basisspace^T  \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
+\bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \bigg)\basisspace^T  \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
 &= 
-[\genstateyDiscreteDotArgnt{} ]^T \basisspace^T  \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace , \\
+[\genstateyDiscreteDotArgnt{} ]^T \basisspace^T  \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace , \\
 &= 
-[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace .
+[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace .
 \end{align*}
@@
 \frac{\partial  \objectiveControlArg{}_2 }{\partial \genstateyDiscreteArgnt{}{}} &= 
--\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}   \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\&=
--\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}    [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T 
- \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  - 
-  [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace 
+-\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}   \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\&=
+-\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}    [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T 
+ \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  - 
+  [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace 
 \\
 &= 
-- 2[\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  
+- 2[\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  
  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace - 
- [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
-&= - \bigg[ 2[\velocity(\veloargsromy)]^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T  [\massArgnt{n}]^{-1} \bigg] \basisspace^T \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
-&= -\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \bigg) \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
+ [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
+&= - \bigg[ 2[\velocity(\veloargsromy)]^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T  [\massArgnt{}]^{-1} \bigg] \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
+&= -\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \bigg) \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
 \end{align*}
@@
 $$ \frac{\partial  \objectiveControlArg{}_3 }{\partial \genstateyDiscreteArgnt{}{}} =
-  [\velocity(\veloargsromy) ]^T  \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
+  [\velocity(\veloargsromy) ]^T  \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
 $$
@@
 \frac{\partial \objectiveControlArg{}}{\partial \genstateyDiscreteArgnt{}} &= 
-[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
-\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)]^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T  \bigg) \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace +
- [\velocity(\veloargsromy) ]^T  \stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
+[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
+\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)]^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T  \bigg) \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace +
+ [\velocity(\veloargsromy) ]^T  \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
 &= 
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{n}
+\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
+\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{}
 \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace \\ 
 &= 
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{n} \basisspace  [\massArgnt{n}]^{-1}\basisspace^T - \mathbf{I} \bigg)
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, 
+\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{} \basisspace  [\massArgnt{}]^{-1}\basisspace^T - \mathbf{I} \bigg)
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, 
 \end{align*}
@@
 \begin{equation*}
-\frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}} = 
-  \adjointDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big]  + 
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{n} \basisspace  [\massArgnt{n}]^{-1}\basisspace^T - \mathbf{I} \bigg)
-\stweightingMatArg{n} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
+\frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}} = 
+  \adjointDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big]  + 
+\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{} \basisspace  [\massArgnt{}]^{-1}\basisspace^T - \mathbf{I} \bigg)
+\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
 \end{equation*}
@@
 \begin{equation*}
-\bigg[ \frac{\partial \hamiltonianArg{n}}{\partial \genstateyDiscreteArgnt{}{}} \bigg]^T =  \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T \stweightingMatArg{n}  \basisspace [\massArgnt{n}]^{-1}  \adjointDiscreteDumArgnt{} +
+\bigg[ \frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}} \bigg]^T =  \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T \stweightingMatArg{}  \basisspace [\massArgnt{}]^{-1}  \adjointDiscreteDumArgnt{} +
  \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T 
-  \stweightingMatArg{n} \bigg(  \basisspace  [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n}- \mathbf{I} \bigg)
+  \stweightingMatArg{} \bigg(  \basisspace  [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{}- \mathbf{I} \bigg)
  \bigg( \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \bigg) 
 \end{equation*}
-Next, we evaluate $\frac{\partial \hamiltonianArg{n}}{\partial \controllerDumArgnt{}}$:
+Next, we evaluate $\frac{\partial \hamiltonianArg{}}{\partial \controllerDumArgnt{}}$:
 \begin{align}
-  \frac{\partial \hamiltonianArg{n}}{\partial \controllerDumArgnt{}{}} &= \frac{\partial}{\partial \controllerDumArgnt{}{}} \bigg[  \adjointDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t) \bigg] \\
- &= \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromy) + [\massArgnt{n}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg] \\
- &= [\adjointDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} + \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},t) \bigg].
+  \frac{\partial \hamiltonianArg{}}{\partial \controllerDumArgnt{}{}} &= \frac{\partial}{\partial \controllerDumArgnt{}{}} \bigg[  \adjointDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
+ &= \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
+ &= [\adjointDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} + \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg].
 \end{align}
-To evaluate $\frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDumArg) \bigg]$, we again 
+To evaluate $\frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg]$, we again 
 go term by term. Starting with the first term,
@@
 \frac{\partial  \objectiveControlArg{}_1 }{\partial \genstateyDiscreteArgnt{}{}} &= 
-\frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) + \\
-& \qquad \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{n}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg) + \\
-& \qquad \frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \controllerDiscreteDumArgnt{}\bigg) \\
-&= \bigg(  [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T                      
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big] \bigg)^T +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \\
-&= \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}.
+\frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) + \\
+& \qquad \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) + \\
+& \qquad \frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \controllerDiscreteDumArgnt{}\bigg) \\
+&= \bigg(  [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T                      
+\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg)^T +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \\
+&= \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}.
 \end{align*}
@@
 \frac{\partial  \objectiveControlArg{}_2 }{\partial \controllerDiscreteDumArgnt{}{}} &= 
--\frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}}   \big[ \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-\stweightingMatArg{n}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\
-&=- \bigg[ [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \velocity(\basisspace \genstateyDiscrete + \stateIntercept) \bigg]^T \\
-&= - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}. 
+-\frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}}   \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
+\stweightingMatArg{}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\
+&=- \bigg[ [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace \genstateyDiscrete + \stateIntercept) \bigg]^T \\
+&= - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}. 
 \end{align*}
@@
 \begin{align*}
-\frac{\partial \hamiltonianArg{n}}{\partial \controllerDiscreteDumArgnt{}} &= 
-[\adjointDiscreteDumArgnt{}]^T[\massArgnt{n}]^{-1}  +  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} + \\
-& \qquad   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}, \\
-& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace \bigg) [\massArgnt{n}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{n}]^{-1}\basisspace^T  \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1}  \\
-& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{n} \basisspace \bigg) [\massArgnt{n}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{n} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{n}]^{-1}  \\
-& =  [\adjointDiscreteDumArgnt{}]^T  [\massArgnt{n}]^{-1}  +  [\controllerDiscreteDumArgnt{}]^T   [\massArgnt{n}]^{-1}  
+\frac{\partial \hamiltonianArg{}}{\partial \controllerDiscreteDumArgnt{}} &= 
+[\adjointDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1}  +  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} + \\
+& \qquad   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}, \\
+& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace \bigg) [\massArgnt{}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  \\
+& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace \bigg) [\massArgnt{}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{}]^{-1}  \\
+& =  [\adjointDiscreteDumArgnt{}]^T  [\massArgnt{}]^{-1}  +  [\controllerDiscreteDumArgnt{}]^T   [\massArgnt{}]^{-1}  
 \end{align*}
@@
 \begin{equation*}
-\frac{\partial \hamiltonianArg{n}}{\partial \controllerDiscreteDumArgnt{}} = 
-  [\massArgnt{n}]^{-1} [\adjointDiscreteDumArgnt{} +  \controllerDiscreteDumArgnt{}]. 
+\frac{\partial \hamiltonianArg{}}{\partial \controllerDiscreteDumArgnt{}} = 
+  [\massArgnt{}]^{-1} [\adjointDiscreteDumArgnt{} +  \controllerDiscreteDumArgnt{}]. 
 \end{equation*}
-Thus, the conditions stated via the Pontryagin Maximum Principle are,
+Evaluating at $(\genstateArg{}{t},\genstateDotArg{}{t},t$), the gradients in the Pontryagin Maximum Principle yield
 \begin{align*}%\label{eq:hamiltonian_sysa}
-&\frac{d}{dt}\genstateArg{n}{t} =  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\veloargsromn) + [\massArgnt{n}]^{-1}\controllerArg{n}{t}\\ 
-&\frac{d }{dt} \adjointArg{n}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n}  \basisspace [\massArgnt{n}]^{-1}  \adjointArg{n}{t} = 
- - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n} 
-\bigg(  \basisspace  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n} - \mathbf{I} \bigg)
- \bigg( \basisspace \genstateDotArg{n}{t} - \velocity(\basisspace \genstateArg{n}{t} + \stateIntercept) \bigg) \\ 
-& \adjointArg{n}{t}= -\controllerArg{n}{t}.
-\end{align*}
-Equivalently,
+&\frac{d}{dt}\genstateArg{}{t} =  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsrom) + [\massArgnt{}]^{-1}\controllerArg{}{t}\\ 
+&\frac{d }{dt} \adjointArg{}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{}  \basisspace [\massArgnt{}]^{-1}  \adjointArg{}{t} = 
+ - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{} 
+\bigg(  \basisspace  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{} - \mathbf{I} \bigg)
+ \bigg( \basisspace \genstateDotArg{}{t} - \velocity(\basisspace \genstateArg{}{t} + \stateIntercept) \bigg) \\ 
+& \adjointArg{}{t}= -\controllerArg{}{t}.
+\end{align*}
+This can be written equivalently as
 \begin{align*}%\label{eq:hamiltonian_sys}
-&\frac{d}{dt}\genstateArg{n}{t} =  [\massArgnt{n}]^{-1}\basisspace^T \stweightingMatArg{n}\velocity(\basisspace \genstateArg{n}{t} + \stateIntercept) + [\massArgnt{n}]^{-1}\controllerArg{n}{t} \\ 
-&\frac{d}{dt} \controllerArg{n}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n}  \basisspace [\massArgnt{n}]^{-1}  \controllerArg{n}{t} = 
- - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{n} 
-\bigg( \mathbf{I} -   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T \bigg)
-\stweightingMatArg{n} \bigg( \basisspace \genstateDotArg{n}{t} - \velocity(\basisspace \genstateArg{n}{t} + \stateIntercept) \bigg) 
+&\frac{d}{dt}\genstateArg{}{t} =  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\basisspace \genstateArg{}{t} + \stateIntercept) + [\massArgnt{}]^{-1}\controllerArg{}{t} \\ 
+&\frac{d}{dt} \controllerArg{}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{}  \basisspace [\massArgnt{}]^{-1}  \controllerArg{}{t} = 
+ - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{} 
+\bigg( \mathbf{I} -   \basisspace  [\massArgnt{}]^{-1}\basisspace^T \bigg)
+\stweightingMatArg{} \bigg( \basisspace \genstateDotArg{}{t} - \velocity(\basisspace \genstateArg{}{t} + \stateIntercept) \bigg). 
 \end{align*}

### 2019-10-09 main.tex
update, taking a break

@@
 Existing space--time approaches differ from the classic LSPG and Galerkin approaches in (1) the definition of the space--time trial subspace and (2) the definition of the residual minimization problem.
-First, space--time approaches leverage a space--time trial
-subspace that associates with a space--time projection operator; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with a spatial projection operator only.
+First, space--time approaches leverage a space--time trial basis that characterizes both the spatial \textit{and} temporal dependence of the state, classic "spatial" model reduction approaches such as LSPG and Galerkin leverage only a spatial trial basis that characterizes the spatial dependence of the state.
+%First, space--time approaches leverage a space--time trial 
+%subspace that associates with a space--time projection operator; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with a spatial projection operator only.
 Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time-step instance (LSPG).
@@
 Finally, \methodAcronym\ is formulated for two kinds of space--time trial
-subspaces: one that associates with spatial projection only (as employed by traditional spatial-projection-based methods), and one that associates with space--time projection (as employed by space--time methods).
+subspaces: one that associates with spatial dimension reduction (as employed by traditional spatial-projection-based methods), and one that associates with space--time dimension reduction (as employed by space--time methods).
 The above attributes allow the \methodAcronym\ approach to be viewed as a
@@
 present work distinguishes itself from Ref.~\cite{constantine_strom} in that
-(1) this work also considers trial subspaces characterized by spatial projection only, and (2) we
+(1) this work also considers trial subspaces characterized by spatial dimension reduction only, and (2) we
 minimize the time-continuous FOM residual over arbitrary time
@@
 arbitrary time windows, and (3) this work also considers trial subspaces
-associated with spatial projection only.
+associated with spatial dimension reduction only.
 
@@
 \caption{Relationship diagram for the \methodAcronym\ approach for model
-	reduction. \KTC{Inconsistent capitalization. I recommend using (1)
-	``Limiting case'', (2) Space--time LSPG. Also, can we make this more specific? In particular, can
+	reduction. \KTC{Also, can we make this more specific? In particular, can
 	we write specific equations (e.g., ``$W = T$ in Eq.~(10)'') for the limiting
-	cases?}} 
+	cases?} \EP{I tried, it becomes very crowded. I think with the description of trial spaces its pretty good now}} 
 \label{fig:flowchart} 
@@
 %		temporal dimensions of the FOM. 
-\item Support of two space--time trial subspaces: one that associates with spatial projection only and 
-    one that associates with space--time projection.
+\item Support of two space--time trial subspaces: one that associates with spatial dimension reduction and 
+    one that associates with spatial and temporal dimension reduction.
 The former case is of particular
@@
 \item Error analysis of the \methodAcronym\ approach using
-	trial subspaces associated with spatial projection.
+	trial subspaces associated with spatial dimension reduction.
 	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
@@
 \item Numerical experiments for 
-	trial subspaces associated with spatial projection only, which demonstrate two key findings:
+	trial subspaces associated with spatial dimenion reduction, which demonstrate two key findings:
 \begin{itemize}
@@
 	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
-		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection operator} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
+		model (\spatialAcronym)}.
+These trial subspaces are characterized by a \textit{spatial projection operator}, associate with a basis that represents the spatial dependence of the state, and are employed in classic model reduction approaches, e.g., Galerkin and LSPG.
+%These
 %	trial subspaces are commonly used
@@
 		model (\spaceTimeAcronym)}.
-These trial subspaces are characterized by a \textit{space--time projection operator} and are employed in space--time 
+These trial subspaces are characterized by a \textit{space--time projection operator}, associate with a basis that represents the spatial and temporal dependence of the state, and are employed in space--time
 model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
@@
 $$
-Thus we have,
+Thus we have
 \begin{align*}
@@
 \bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{}
-\frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace \\ 
+\frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\ 
 &= 
@@
 \end{equation*}
-Or, equivalently,
+Equivalently we write
 \begin{equation*}
@@
   \stweightingMatArg{} \bigg(  \basisspace  [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{}- \mathbf{I} \bigg)
- \bigg( \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \bigg) 
+ \bigg( \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \bigg). 
 \end{equation*}
 Next, we evaluate $\frac{\partial \hamiltonianArg{}}{\partial \controllerDumArgnt{}}$:
-\begin{align}
+\begin{align*}
   \frac{\partial \hamiltonianArg{}}{\partial \controllerDumArgnt{}{}} &= \frac{\partial}{\partial \controllerDumArgnt{}{}} \bigg[  \adjointDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
@@
  &= [\adjointDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} + \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg].
-\end{align}
+\end{align*}
 To evaluate $\frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg]$, we again 
@@
 
-%\section*{References}
+\section*{References}
 \bibliographystyle{siam}

### 2019-10-09 section4/spatial_numerical.tex
latest
cdots in 35
above remark 4.4

@@
 \begin{align}
-&\timeWindowArg{n}{0} \le \ldots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
+&\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
 \timeWindowArg{n}{0} = \timeStartArg{n},\;
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $tol$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
+\Input{tolerance, $\epsilon$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
 factor, $\fbsmDecay$} 
@@
 $i \leftarrow 1$ \Comment{Set counter}\\
-\While{$tol \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
+\While{$\epsilon \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
 \small{
@@
 \begin{enumerate}
-\item Jacobian-free Derivative Approximation: A non-intrusive way to evaluate the Jacobian transpose in~\eqref{eq:lspg_adjoint} is to recognize that all terms including the Jacobian transpose are left multiplied by the transpose of the trial basis.
+\item \textit{Jacobian-free approximation}: A non-intrusive way to evaluate the Jacobian transpose in~\eqref{eq:lspg_adjoint} is to recognize that all terms including the Jacobian transpose are left multiplied by the transpose of the trial basis.
 One can make the manipulation,
@@
 
-\item Automatic Differentiation: A more intrusive, but potentially more efficient, method of computing the action of the Jacobian transpose on a vector is through automatic differentiation (AD).
+\item \textit{Automatic differentiation}: A more intrusive, but potentially more efficient, method of computing the action of the Jacobian transpose on a vector is through automatic differentiation (AD).
 AD methods comprise a class of techniques that can be used to numerically evaluate derivatives of functions (e.g., Jacobians, vector-Jacobian products) by recursively applying the chain rule.

### 2019-10-09 section4/spatial_numerical.tex
update fbsm algo

@@
 %Compute: $[\jacobianSlabArg{n}]^T \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{compute RHS for normal equations} \\
-$\Delta \genstatecollocMatSlabArg{n} \solveAlgo [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
+ Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
 $\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
@@
 guess for costate}\\ 
-$\genstateArgnt{n}_1 \solveAlgo \basisspaceTArg{n} \stweightingMatArg{n}
-\basisspaceArg{n} \genstateDotArgnt{n}_1  -  \basisspaceTArg{n} \stweightingMatArg{n}
-\velocity(\veloargsromntArg{1}) =  \controllerArgnt{n}$ 
+$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n}
+\basisspaceArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
+\velocity(\veloargsromArg{1}) =  \controllerArg{n}{t}$ 
 % Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
 % obtain $\genstate_0^{n}(t)$ 
-\Comment{Solve~\eqref{eq:lspg_continuous} with initial guess to
-controller}\\ 
+\Comment{Solve~\eqref{eq:lspg_continuous}}\\ 
 %Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
@@
 \begin{multline*}
-\adjointArgnt{n} \underset{\text{solve}}{\leftarrow} 
-\frac{d}{dt} \adjoint^n  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjoint^n = \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspace [\massArg{n}]^{-1} \basisspaceTArg{n} \bigg)  \stweightingMatArg{n} \bigg( \basisspaceArg{n} \dot{\genstate}_i^n   -   \velocity( \basisspaceArg{n} \genstate_i^n  +\stateInterceptArg{n},t) \bigg) \bigg]
+\text{ Compute } \adjointArgnt{n} \text{ satisfying }
+\adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg]
 \end{multline*} }
-\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate state} \\
+\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate} \\
 %Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
 $\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
-$i \leftarrow i+1$ \Comment{Update Counter}\\
-$\genstateArgnt{n}_i \solveAlgo \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i + \stateInterceptArg{n},t) =  \controllerArgnt{n} $
-\Comment{Solve~\ref{eq:lspg_continuous} with updated guess to costate}\\
+$i \leftarrow i+1$ \Comment{Update counter}\\
+$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) =  \controllerArg{n}{t} $
+\Comment{Solve~\eqref{eq:lspg_continuous}}\\
 %$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
-\uIf{ $\objectiveArg{n}({\genstate_i^n}) \le \objectiveArg{n}({\genstate_{i-1}^n})$}
+\uIf{ $\objectiveArg{n}({\basisspaceArg{n}\genstate_i^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}}) \le \objectiveArg{n}({\basisspaceArg{n}\genstate_{i-1}^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}})$}
 {

### 2019-10-10 cavity.tex
updated figs and language in first numerical experiment

@@
 \subsection{Cavity flow}
-The next numerical example considers hyper-reduced ROMs of a viscous, compressible flow in a two-dimensional cavity. 
+The next numerical example considers collocated ROMs\footnote{Collocation is a form of hyper-reduction which requires sampling the full-order model at only select grid points.} of a viscous, compressible flow in a two-dimensional cavity.
 The flow is described by the two-dimensional compressible Navier-Stokes equations,
@@
 \qquad \mathbf{F}_{v_1} = \begin{Bmatrix} 0 \\ \tau_{11} \\ \tau_{12}  \\ u_j \tau_{j1} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_1}  \end{Bmatrix}, 
-\qquad \mathbf{F}_{v_2} = \begin{Bmatrix} 0 \\ \tau_{21} \\ \tau_{22}  \\ u_j \tau_{j2} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_2}  \end{Bmatrix}.
+\qquad \mathbf{F}_{v_2} = \begin{Bmatrix} 0 \\ \tau_{21} \\ \tau_{22}  \\ u_j \tau_{j2} + c_p \frac{\mu}{\text{Pr}} \frac{\partial T}{\partial x_2}  \end{Bmatrix},
 $$
-We assume a Newtonian fluid, which leads to a viscous stress tensor of the form,
+where $\mu \in \RRplus$ is the viscosity, $\text{Pr} = 0.72$ is the Prandtl number, and $c_p \in \RRplus$ the heat capacity ratio.
+We assume a Newtonian fluid, which leads to a viscous stress tensor of the form
 \begin{equation*}
@@
 \end{equation*}
-where,
+where
 \begin{equation*}
@@
 \end{equation*}
-The Navier-Stokes equations are closed with a constitutive relationship for a calorically perfect gas,
+The Navier--Stokes equations are closed with a constitutive relationship for a calorically perfect gas,
 $$p = (\gamma - 1)( \rho E - \frac{1}{2} \rho u_1^2 - \frac{1}{2} \rho u_2^2 \big),$$
@@
 of the form,
-$$\frac{d \state}{dt} = \mass_{\text{DG}} \velocity_{\text{DG}} (\state),$$
-where $\mass_{\text{DG}} \in \RR{N \times N}$ is the DG mass matrix and $\velocity_{\text{DG}}: \RR{N} \rightarrow \RR{N}$ is the DG velocity operator containing 
-surface and volume integrals. By the definition of the FOM~\eqref{eq:FOM}, the dynamical system velocity is $\velocity = \mass_{\text{DG}} \velocity_{\text{DG}}$. 
+$$\frac{d \state}{dt} = \mass_{\text{DG}}^{-1} \velocity_{\text{DG}} (\state),$$
+where $\mass_{\text{DG}} \in \RR{N \times N}$ is the (block diagonal) DG mass matrix and $\velocity_{\text{DG}}: \RR{N} \rightarrow \RR{N}$ is the DG velocity operator containing 
+surface and volume integrals. By the definition of the FOM~\eqref{eq:FOM}, the dynamical system velocity is $\velocity = \mass_{\text{DG}}^{-1} \velocity_{\text{DG}}$. 
 Figure~\ref{fig:cav_mesh} shows the computational mesh. The 
-DG method leverages the Rusanov flux at the cell interfaces~\cite{rusanov} and uses the first form of Bassi and Rebay~\cite{br1} for the viscous fluxes. Time integration 
-is performed via a third-order strong stability preserving Runge-Kutta method~\cite{ssp_rk3} with a time-step size of $\Delta t = 0.001$.
+DG method uses the Rusanov flux at the cell interfaces~\cite{rusanov} and uses the first form of Bassi and Rebay~\cite{br1} for the viscous fluxes. Time integration 
+is performed via a third-order strong stability preserving Runge-Kutta method~\cite{ssp_rk3} with a time step of $\Delta t = 0.001$.
 
-The reduced-order models leverage POD to construct the trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to construct the weighting 
-matrix $\stweightingMatArg{n}$. The process used to construct the initial conditions and trial subspaces for the ROMs are as follows:
+The reduced-order models leverage POD to construct the \spatialAcronym\ trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to select the sampling points and, as a result the weighting
+matrix $\stweightingMatArg{}$; we use a constant weighting matrix across all windows, e.g., $\stweightingMatArg{n} \equiv \stweightingMatArg{}$.
+The process used to construct the initial conditions and trial subspaces for the ROMs is as follows:
 \begin{enumerate}
@@
 \begin{itemize}
-\item \textit{Galerkin ROM with collocation}: The Galerkin ROM with collocation is obtained through Galerkin projection of the FOM. Unless 
+\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation. Unless 
 otherwise noted, the Galerkin ROM is evolved in time with the Crank--Nicolson time scheme at a time step of $\Delta t =0.1$.
 
-\item \textit{LSPG ROM with collocation}: The LSPG ROM with collocation is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
+\item \textit{LSPG ROM with collocation}: We consider a collocated LSPG ROM, which is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
 discretization. Unless otherwise noted, the time step for the LSPG ROM is $\Delta t  = 0.1$. As in the previous example, 
-the Gauss--Newton method is used to solve the non-linear least-squares problem arising at each time instance. The Gauss--Newton 
-algorithm is deemed converged when the gradient norm is less than $1e-4$. All Jacobians are again computed via automatic differentiation.
+the Gauss--Newton method is used to solve the non-linear least-squares problem arising at each time instance. We deem the Gauss--Newton 
+algorithm converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation.
 
-\item \textit{\methodAcronymROMs\ with collocation}: \methodAcronymROMs\ are solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time-step size of 
-$\Delta t = 0.1$. The Gauss--Newton algorithm is deemed converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation. Uniform quadrature 
+\item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time step of 
+$\Delta t = 0.1$. We deem the Gauss--Newton algorithm converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation. Uniform quadrature 
 weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.
@@
 \includegraphics[trim={0cm 14cm 0cm 14cm},clip,width=1.\linewidth]{figs/cavity/grid.png}
-\caption{Computational mesh}
+\caption{Computational mesh employed in cavity flow simulations.}
 \label{fig:cav_mesh}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.65\linewidth]{figs/cavity/hyper_grid.png}
-\caption{Close up of computational mesh with highlighted collocation cells} 
+\caption{Close up of computational mesh with highlighted collocation cells.} 
 \label{fig:cav_sampmesh}

### 2019-10-10 numerical_experiments.tex
updated figs and language in first numerical experiment

@@
 In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs.
-Additionally, we assess the impact of the window size, time-step size, and time scheme on \methodAcronym.
-It is noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs.
-Future state prediction and/or interpolation to new flow conditions is not considered as these problems introduce confounding factors (e.g., accuracy of the basis).
+The purpose of the numerical experiments is to assess the impact on minimizing the residual over an arbitrarily sized time window on the solution accuracy.
+We additionally assess the impact of the time step, and time scheme on \methodAcronym.
+In both experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$.
+It is also noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs.
+Future state prediction and prediction at new parameter instances are not considered as these problems introduce factors that confound the solution accuracy with the solution methodology (e.g., accuracy of the basis).
 
@@
 
-\subsubsection{Description of FOM and generation of trial space}
+\subsubsection{Description of FOM and generation of \spatialAcronym\ trial subspace}\label{sec:sod_fom}
 The 1D compressible Euler equations are solved using a finite volume method.
@@
 The \spatialAcronym\ trial subspace is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $.
-The resulting trial subspace comprises $99.99\%$ of the snapshot matrix energy.
+The resulting trial subspace corresponds to an energy criterion of $99.99\%$.
 %  The solution is saved every other time-step, resulting in 500 solutions snapshots.
@@
 \subsubsection{Description of reduced-order models}
-Reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approaches are considered. No hyper-reduction is considered 
-in this example, i.e., $\stweightingMatArg{} = \mathbf{I}$. Details on the implementation of the 
+We consider reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approach. No hyper-reduction is considered 
+in this example, i.e., $\stweightingMatArg{n} = \mathbf{I}$, $n=1,\ldots,\nslabs$. Details on the implementation of the 
 different reduced-order models is as follows:
 \begin{itemize}
-\item \textbf{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
-a constant time-step size of $ \Delta t = 0.002$.
+\item \textit{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
+a constant time step of $ \Delta t = 0.002$.
 
-\item \textbf{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time-scheme as previously described.
-Unless noted otherwise, the LSPG ROM uses a constant time-step size of $\Delta t = 0.002$.
-The nonlinear least-squares problem arising at each time-step instance is solved via the Gauss--Newton method.
+\item \textit{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time scheme as previously described.
+Unless noted otherwise, the LSPG ROM uses a constant time step of $\Delta t = 0.002$.
+The nonlinear least-squares problem arising at each time instance is solved via the Gauss--Newton method.
 All Jacobians are computed via automatic differentiation~\cite{adolc}.
 The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$.
-\item \textbf{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
+\item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
 \begin{itemize}
-\item Direct Method: \methodAcronymROMs\ solved via the direct method are implemented for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time-step size of $\Delta t = 0.0005$.
+\item Direct method: We consider \methodAcronymROMs\ solved via the direct method for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time step of $\Delta t = 0.0005$.
 The nonlinear least-squares problem
@@
 
-\item Indirect Method: Two \methodAcronymROMs\ solved via the indirect method are implemented.
-The first uses the same CN discretization at at time-step size of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time-step size of $\Delta t = 0.0005$.
+\item Indirect method: We consider two \methodAcronymROMs\ solved via the indirect method.
+The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$.
 The coupled two-point boundary
@@
 \subsubsection{Numerical results}
-The impact of the window size on \methodAcronymROMs\ is first assessed. A set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
-$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$ are considered.
-The standard Galerkin and LSPG ROMs are additionally considered.
-Results are shown for the direct method with the CN time-discretization scheme; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section.
-First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = \{0.5,1.0\}$. Figure~\ref{fig:sod_xt} shows 
+The impact of the window size on \methodAcronymROMs\ is first assessed. We consider a set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
+$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$.
+We additionally consider the standard Galerkin and LSPG ROMs.
+We first show results for \methodAcronymROMs\ using the direct method with CN time discretization; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section.
+First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
 $x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, it is observed that the LSPG and Galerkin ROMs accurately characterize 
-the system: they correctly track the shock location, expansion waves, etc. Both predictions, however, are observed to by highly oscillatory. These oscillations are 
-not physical and can lead to numerical blowup; e.g., due to negative pressure. The \methodAcronymROMs\ produce less oscillatory solutions than both 
-the Galerkin and LSPG ROMs. Critically, it is seen that, as the window size over which the residual is minimized grows, the solution becomes less oscillatory. 
-When the residual is minimized over the entire space--time domain, the solution displays no oscillations.
+the system: they correctly track the shock location, expansion waves, etc. We observe both predictions, however, to be highly oscillatory. These oscillations are 
+not physical and can lead to numerical blowup; e.g., due to negative pressure. We observe the \methodAcronymROMs\ to produce less oscillatory solutions than both 
+the Galerkin and LSPG ROMs.
+Critically, we see that the solution becomes less oscillatory as the window size over which the residual is minimized grows.
+The solution displays no oscillations when the residual is minimized over the entire space--time domain.
 
@@
 
-Next, Figure~\ref{fig:sod_error} shows integrated state errors and the objective function (i.e., the residual norm) over $t \in [0,1]$ for the various  ROMs. Most notably, it is observed 
-that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the integrated $\elltwo$  
-error. Increasing the window size does, however, lead to a monotonic decrease in the total integrated residual. It is additionally noted that, although the integrated
+Next, Figure~\ref{fig:sod_error} shows space--time (i.e., integrated) state errors and the objective function~\eqref{eq:obj} (i.e., the space--time residual norm) over $t \in [0,1]$ for the various ROMs.
+Most notably, we observe
+that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the space--time $\elltwo$  
+error. Increasing the window size does lead to a monotonic decrease in the space--time residual, however. We additionally note that, although the space--time
 $\elltwo$ error of 
-the projected FOM solution is significantly lower than that of the various ROM solutions, the integrated residual norm of the projected FOM solution is \textit{higher} 
+the projected FOM solution is significantly lower than that of the various ROM solutions, the space--time residual norm of the projected FOM solution is \textit{higher} 
 than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$ norm, it does not satisfy the governing equations as well the ROM solutions.
@@
 Figure~\ref{fig:sod_error_methods} 
-shows the same integrated $\elltwo$ state errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
-the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, it is observed that, although some 
-deviations are observed, the \methodAcronym\ method is relatively insensitive to the solution technique (direct vs indirect) and underlying discretization scheme.
-In particular, ROMs using the second order explicit AB2 scheme with a time-step size of $\Delta t = 0.0005$ provide similar results to the 
-ROMs using the second-order CN scheme at a time-step size of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal integrated 
+shows the same space--time $\elltwo$ state errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
+the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, we observe observed that the \methodAcronym\ method 
+is relatively insensitive to the solution technique (direct vs indirect) and underlying discretization scheme, although some minor differences are observed.
+In particular, ROMs using the second order explicit AB2 scheme with a time step of $\Delta t = 0.0005$ provide similar results to the 
+ROMs using the second-order CN scheme at a time step of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal space--time
 $\elltwo$ state error occurs when the window size is $\DeltaSlabArg{} = 0.1$ and the residual decreases monotonically as the window size grows. 
@@
 as the window size is increased. For indirect methods, this increase in cost is due to the fact that, as the window size 
-grows, more iterations of the FBSM are required for convergence. For direct methods, the increase in cost is due to (1.) the cost 
-associated with forming and solving the normal equations at each Gauss--Newton iteration and (2.) the increased number of 
+grows, more iterations of the FBSM are required for convergence. For direct methods, the increase in cost is due to (1) the cost 
+associated with forming and solving the normal equations at each Gauss--Newton iteration and (2) the increased number of 
 Gauss--Newton iterations required for convergence. The cost of the FBSM method is observed to increase more rapidly 
@@
 utilizing the CN time discretization. This is despite the fact that AB2 is evolved at a time step of $\Delta t = 0.0005$, while CN uses $\Delta t = 0.002$; thus 
-AB2 contains 4 times more temporal degrees of freedom than CN.  Finally, it is emphasized that the 
+AB2 contains 4 times more temporal degrees of freedom than CN.  Finally, we emphasize that the 
 results presented here are presented for standard algorithms (e.g., Guass--Newton and the FBSM). As mentioned in 
-Remarks~\ref{remark:gaussnewton} and~\ref{remark:fbsm}, it is expected the computational cost 
-of both the indirect methods and direct methods can be decreased through the use and/or development of 
+Remarks~\ref{remark:gaussnewton} and~\ref{remark:fbsm}, we expect the computational cost 
+of both the indirect and direct methods can be decreased through the use and/or development of 
 algorithms tailored to the minimization problem.
@@
 
-Finally, we study the impact of the time-step size on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
-window size of $\Delta T = 0.1$, with time-step sizes of $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
+Finally, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
+window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
 $100$, $50$, $20$, $10$, $2$, and $1$ time-steps per window). We additionally 
-consider LSPG ROMs leveraging the same set of time steps. Figure~\ref{fig:time_step_study} shows the integrated $\elltwo$ error 
-of the various ROMs as well as the integrated residual norm. We observe that both the $\elltwo$ state error and integrated residual 
-of the \methodAcronymROMs\ decrease and converge as the time-step size decreases. This is in direct contrast to the LSPG ROMs, in where both the $\elltwo$ 
-error and integrated residual norm increase as the time-step size decreases. This result 
+consider LSPG ROMs leveraging the same set of time steps.
+To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 1e-4$.
+Figure~\ref{fig:time_step_study} shows the space--time $\elltwo$ error and residual norm
+of the various ROMs. We observe that both the space--time $\elltwo$ error and residual norm 
+of the \methodAcronymROMs\ decrease and converge as the time step decreases. This is in direct contrast to the LSPG ROMs, in where both the $\elltwo$ 
+space--time error and residual norm increase as the time step decreases. This result 
 demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
-shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that, for all time-step sizes considered, the \methodAcronym-ROMs 
+shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that, for all time steps considered, the \methodAcronym-ROMs 
 are less than 6x the cost LSPG; despite the window sizes comprising up to 100 time steps. 
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
-\caption{Integrated $\elltwo$ error of the \methodAcronymROM.}
+\caption{Space--time $\elltwo$ error of the \methodAcronymROM.}
 \label{fig:sod_error_methods_a}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
-\caption{Integrated residual norm of the \methodAcronymROM.} 
+\caption{Space--time residual norm of the \methodAcronymROM.} 
 \label{fig:sod_error_methods_b}
 \end{subfigure}
-\caption{Integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs} 
+\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.} 
 \label{fig:time_step_study}
@@
 \end{subfigure}
-\caption{Comparison of wall-clock times of the \methodAcronymROMs\ to LSPG ROMs as a function of time-step size.} 
+\caption{Comparison of wall-clock times of the \methodAcronymROMs\ to LSPG ROMs as a function of the time step. WLS ROMs have a fixed window size of $\Delta T = 0.1$.} 
 \label{fig:walltime_dtvar}
@@
 \begin{enumerate}
-\item Increasing the window size over which the residual is minimized lead to more physically relevant solutions.
-Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ lead to less oscillatory solutions.
-\item Increasing the window size over which the residual is minimized does not necessary lead to a more accurate solution in the $\elltwo$ sense.
-We observed that minimizing the residual over an intermediary window size led to the most accurate results as measured by the $\elltwo$ norm.
-\item The \methodAcronym\ displays time-step convergence: both the integrated $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time-step size decreased.
+\item Increasing the window size over which the residual is minimized led to more physically relevant solutions.
+Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ led to less oscillatory solutions.
+\item Increasing the window size over which the residual is minimized does not necessary lead to a lower space--time state error as measured by the $\elltwo$ norm.
+We observed that minimizing the residual over an intermediary window size led to the lowest space--time state error in the $\elltwo$ norm.
+\item The \methodAcronym\ displays time-step convergence: both the space--time $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time step decreased.
 This is in contrast to LSPG.
-\item For all examples considered, in where the windows comprised up to 2000 time-steps, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
+\item For all examples considered, in where the windows comprised up to 2000 time instances, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
 direct method was observed to be, in general, slightly more efficient and robust than the indirect method.

### 2019-10-10 cavity.tex
through conclusions

@@
 $\Delta t = 0.1$. We deem the Gauss--Newton algorithm converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation. Uniform quadrature 
-weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.
+weights are used for evaluating the integral in the minimization problem~\eqref{eq:obj_gen_slab}.
 
@@
 We first assess the performance of \methodAcronymROMs\ with varying window sizes. To this end, we consider \methodAcronymROMs\ with uniform window 
-sizes of $\Delta T^n \equiv \Delta T = 0.2,0.5,1.0,$ and $2.0$, along with the Galerkin and LSPG ROMs. Results are first considered for basis \#1 as described in 
+sizes of $\Delta T^n \equiv \Delta T = 0.2,0.5,1.0,$ and $2.0$, along with the Galerkin and LSPG ROMs. We first consider results for basis \#1 as described in 
 Table~\ref{tab:rom_basis_details}. For all ROMs, we evolve 
@@
 First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$ state error of the various reduced-order models.
-Both the hyper-reduced Galerkin and LSPG ROMs blow up/fail to converge within the first several time units.
+Both the collocated Galerkin and LSPG ROMs blow up/fail to converge within the first several time units.
 The \methodAcronymROM\ minimizing the residual over a window of size $\Delta T = 0.2$ also fails to converge.
@@
 \end{center}
-\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various hyper-reduced ROMs to the full-order model solution.}
+\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various collocated ROMs to the full-order model solution.}
 \label{fig:cav_results1}
@@
 
-Figure~\ref{fig:cav_results2} shows the integrated error and objective function of the stable \methodAcronymROMs. It is seen that growing the window size over 
+Figure~\ref{fig:cav_results2} shows the space--time error and objective function of the stable \methodAcronymROMs. It is seen that growing the window size over 
 which the residual is minimized leads to a lower space--time residual, but not necessarily a lower $\elltwo$ error.
@@
 Next, Figure~\ref{fig:cav_wallclock}
-shows the wall-clock times of the \methodAcronymROMs\ for $t \in [0,10]$ as compared to the LSPG ROM\footnote{Note that LSPG blew up at $t \approx 16.0$, so we focus on the first ten time units}.
-As expected, increasing the time-slab size again leads to an increase in computational cost.
-Minimizing the residual over a time-slab comprising 20 time-steps leads to a 2.5x increase in cost over LSPG.
+shows the wall-clock times of the \methodAcronymROMs\ for $t \in [0,10]$ as compared to the LSPG ROM\footnote{Note that LSPG blew up at $t \approx 16.0$, so we focus on the first ten time units.}.
+As expected, increasing $\Delta T$ again leads to an increase in computational cost; minimizing the residual over a window comprising 20 time instances yields a 2.5x increase in cost over LSPG.
 
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window_basis2.pdf}
-\caption{Integrated $\elltwo$ error}
+\caption{Space--time $\elltwo$ error}
 \label{fig:cav_results2a}
@@
 \end{center}
-\caption{Integrated error (left) and objective function (right) as a function of window size.}
+\caption{Space--time error (left) and objective function (right) as a function of window size.}
 \label{fig:cav_results2}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.49\linewidth]{figs/cavity/walltime_vs_window_compare_basis2.pdf}
-\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM}
+\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM.}
 \label{fig:cav_wallclock}
@@
 \methodAcronymROMs\ again yield improved performance: \methodAcronymROMs\ minimizing the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are less than $5\%$.
-The \methodAcronymROMs\ with $\Delta T \ge 0.5$ leveraging the enriched basis yield more accurate results than \methodAcronymROMs\ using the coarse basis.
+The \methodAcronymROMs\ with $\Delta T \ge 0.5$ and basis \#2 yield more accurate results than \methodAcronymROMs\ with basis \#1.
 
@@
 \end{center}
-\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various hyper-reduced ROMs to the full-order model solution.}
+\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various collocated ROMs to the full-order model solution.}
 \label{fig:cav_results1_basis1}
@@
 Increasing the window size again leads to an increase in computational cost.
-Minimizing the residual over a window comprising 20 time-steps leads to a 2.5x increase in cost over LSPG.
+Minimizing the residual over a window comprising 20 time instances yields a 2.5x increase in cost over LSPG.
 This increase in computational cost is particularly reasonable in this example as LSPG yielded an unstable solution.

### 2019-10-10 main.tex
updated figure notation in cavity example
through conclusions

@@
 \section{Conclusions}\label{sec:conclude}
-This paper introduced a windowed least-squares reduced-order modeling (\methodAcronym) formulation for dynamical systems.
+This paper introduced the windowed least-squares (\methodAcronym) \approachKwd\ for model reductio\  of dynamical systems.
 The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial subspace over a series of time windows.
@@
 
-Two differing techniques to solve the least-squares minimization problems were considered: direct (discretize then minimize) and indirect (minimize then discretize). 
+We considered two differing techniques to solve the least-squares minimization problems arising from the \methodAcronym\ \approachKwd: direct (discretize-then-optimize) and indirect (optimize-then-discretize).
 The case of \spatialAcronym\ trial subspaces proved particularly interesting: it was shown that indirect methods comprise solving a coupled two-point 
@@
 
-Numerical experiments of the compressible Euler and Navier-Stokes equations demonstrated the utility in the proposed approach.
-The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded smoother and more physically relevant solutions.
-Further, increasing the window size over which the residual was minimized led to a monotonic decrease in the total space--time residual.
+Numerical experiments of the compressible Euler and Navier--Stokes equations demonstrated the utility in the proposed approach.
+The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals.
 Increasing the window size over which the residual was minimized, however, did not necessarily decrease the $\elltwo$ solution error; this was observed to occur over an intermediary window size.
-It was
-additionally shown that the \methodAcronymROMs\ overcome the time-step sensitivity challenges that LSPG is subject to.
-The second numerical experiment, which examined hyper-reduced ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow.
+We additionally observed that the
+\methodAcronym\ \approachKwd\ overcomes the time-step sensitivity challenges that LSPG is subject to.
+The second numerical experiment, which examined collocated ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow.
 In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up.
@@
 The principle challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to, in general, a higher computational cost.
-This increased cost is due to the fact that forming and solving the nonlinear least-squares problem becomes more expensive and difficult as the window size grows.
-While numerical experiments demonstrated that this increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.
+In the context of the direct solution approach, this increased cost is due to the fact that forming and solving the least-squares problem becomes more expensive as the window size grows.
+In the context of indirect methods, this increased cost is a result of the increased number of iterations of the forward backward sweep method.
+While numerical experiments demonstrated that the increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.
 
@@
 \end{equation}
-where $\integrand: \RR{M} \otimes \timeSpaceArg{} \times \RR{M} \otimes \timeSpaceArg{} \times[0,T] \rightarrow  \RRplus$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$.
+where $\integrand: \RR{M}  \times \RR{M}  \times[0,T] \rightarrow  \RRplus$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$.
 We now introduce the function $\statez: [0,T] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(0) = \statez_0$.

### 2019-10-11 section3/general_form.tex
comments

--- 2019-10-08 section3/general_form.tex
@@
 outlines the proposed \methodNameLower\ 
-(\methodAcronym) \approachKwd. In contrast to (1) the Galerkin approach, 
+(\methodAcronym) \approachKwd. In contrast to (1) Galerkin projection, 
 which minimizes the (time-continuous) FOM ODE residual at a time instance, 
-(2) the LSPG approach, 
+(2) LSPG projection, 
 which minimizes
 the (time-discrete) FOM O$\Delta$E residual 
-over a time step, and (3) the ST-LSPG approach,  
+over a time step, and (3) ST-LSPG projection, 
 which minimizes
 the  FOM O$\Delta$E residual 
-over the entire time interval, the proposed \methodAcronym\ method sequentially minimizes the 
-time-continuous FOM ODE residual over a series of arbitrarily defined
+over the entire time interval, the proposed \methodAcronym\ \approachKwd\ sequentially minimizes the 
+FOM ODE residual over arbitrarily defined
 \textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
@@
 for a general space--time trial subspace. We then examine the \spatialAcronym\ 
-trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
-associated with the associated residual-minimization problems.
+trial subspace in this context, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
+associated with the residual-minimization problems.
  %The proposed method can be formulated equivalently from
@@
 $\timeStartArg{n+1} = \timeEndArg{n}$,
-$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts this partitioning.
+$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts such a partitioning.
 \begin{figure} 
@@
 \caption{Partitioning of the time
-domain into time windows.} 
+	domain into time windows. \KTC{Improve this by making the time interval
+	finite (i.e., $[0,T]$) and introduce a partitioning. You should also match the font size with the rest of
+	the text. See Fig 1 of our paper ``Data-driven time parallelism via
+	forecasting", for example.}} 
 \label{fig:slab_fig} 
@@
 $t\in[\timeStartArg{n},\timeEndArg{n}]$, which is enforced to reside in 
-the $n$th space--time trial subspace such that
+the $n$th space--time trial subspace 
+$\stspaceArg{n}$
+such that
 \begin{equation}
@@
 \end{equation}
-where $\stspaceArg{n}$ is the space--time trial subspace over the $n$th window and $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
-$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} = \{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$)
+where  $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
+$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} =
+\{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$).
 %and
@@
 $$ 
+\KTC{This above doesn't make sense.}
 To outline \methodAcronym, we now define the objective functional over the $n$th window:

### 2019-10-11 section3/general_form.tex
fix eq 13
comment

@@
 For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
-$$\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n} \subset \stspaceArg{n}, \qquad n=1,\ldots,\nslabs. 
-%, \qquad 
-%\approxstateArg{n}{\timeEndArg{n}} \in \stspaceBoundEndArg{n} \subset \stspaceArg{n}.
-$$ 
-\KTC{This above doesn't make sense.}
-To outline \methodAcronym, we now define the objective functional over the $n$th window:
+\begin{equation}
+	\stspaceBoundStartArg{n}\defeq\mathrm{span}(\{\statey(\timeStartArg{n})\,|\, \statey\in
+	\stspaceArg{n}\})\subseteq\RR{\fomdim}, \qquad n=1,\ldots,\nslabs. 
+\end{equation}
+such that $\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n}$.
+To outline \methodAcronym, we now define the objective functional over the
+$n$th window as
 \begin{equation}\label{eq:obj}
@@
 where $\projectorArg{n} : \RR{\fomdim} \rightarrow \stspaceBoundStartArg{n}$ 
-is a (spatial) projector onto the start of the $n$th trial space (e.g., $\elltwo$).
+is a (spatial) projection operator onto the start of the $n$th trial space
+(e.g., $\elltwo$-orthogonal projection operator).
 
@@
 We leave further investigation into other approaches, such as nonlinear
-	manifolds~\cite{leeCarlberg}, as a subject for future work.
+	trial manifolds~\cite{leeCarlberg}, as a subject for future work.
 %A variety of trial spaces have been examined in the model-order reduction community.

### 2019-10-11 section4/numerical_methods.tex
fixed en dahshes

--- 2019-10-08 section4/numerical_methods.tex
@@
 \begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A direct approach leveraging linear multistep methods to directly solve the optimization problem~\eqref{eq:tclsrm}.
-\item \textit{Indirect methods (optimize-then-discretize)}: An indirect approach leveraging the forward-backward sweep algorithm to solve the Euler--Lagrange equations~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}.
+\item \textit{Indirect methods (optimize-then-discretize)}: An indirect approach leveraging the forward--backward sweep algorithm to solve the Euler--Lagrange equations~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}.
 
@@
 of additional approaches.
-Section~\ref{sec:direct}  outlines direct solution approaches based on linear
-multistep methods, while Section~\ref{sec:indirect} outlines the indirect
-solution approach.
+Section~\ref{sec:direct} outlines a direct solution approach based on linear
+multistep methods, while Section~\ref{sec:indirect} outlines an indirect
+solution approach based on the forward--backward sweep algorithm.
 

### 2019-10-11 section4/spatial_numerical.tex
fixed en dahshes

@@
 Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time step, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that
-define the specific type of multistep scheme. Examples of linear multistep schemes include Adams-Bashforth schemes and  
-Adams-Moulton schemes. 
+define the specific type of multistep scheme. Examples of linear multistep schemes include Adams--Bashforth schemes and  
+Adams--Moulton schemes. 
 \begin{remark}\label{remark:LMS}
@@
 techniques include shooting methods, multiple shooting
-methods~\cite{multiple_shooting}, and the forward-backward sweep
+methods~\cite{multiple_shooting}, and the forward--backward sweep
 method~\cite{fbs} (FBSM).  This work explores using the FBSM.
@@
 
-\subsubsection{Forward-backward sweep method (FBSM)}\label{sec:FBSM}
+\subsubsection{Forward--backward sweep method (FBSM)}\label{sec:FBSM}
 %\subsubsection{Solution Strategy} Solving Eq.~\ref{eq:lspg_continuous} is
@@
 The FBSM is a simple iterative method for solving the coupled two-point boundary value problem. For large time windows, however, the FBSM may require many 
-forward-backward iterations for convergence. More sophisticated solution techniques, such as a multiple FBSM method or multiple shooting methods, promise 
+forward--backward iterations for convergence. More sophisticated solution techniques, such as a multiple FBSM method or multiple shooting methods, promise 
 to reduce this cost. Analyses of additional solution techniques will be a subject of future work.

### 2019-10-11 section4/spacetime_numerical.tex
fixed error

@@
 Numerically solving the minimization problem requires the introduction of a quadrature rule for 
-discretization of the integral. We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ N )$ collocation points over the $n$th window, 
+discretization of the integral.
+We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ N )$ \EP{Just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window,
 $$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$

### 2019-10-11 section4/spatial_numerical.tex
fixed error

@@
 \subsubsection{Linear multistep schemes}
-\KTC{HERE}
+\KTC{HERE} \EP{This section has the double superscript. We could make one of these a subscript, but then need to change the algorithms a bit as 
+I use subscripts to denote iterations. Either is fine}
 Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time-instances.  
@@
 \end{equation*}
-Now, define the weighted space--time residual over the entire window as
+We now define the weighted space--time residual over the entire window as
 \begin{equation*}
 \residLMSSlabArg{n} : (\genstatecollocMatySlabArg{};\genstateDiscreteArg{n,0}) \mapsto \begin{bmatrix}
- \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n,1} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
+ \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
 \vdots \\
- \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n,\nstepsArg{n}} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
+ \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
 \end{bmatrix},

### 2019-10-12 main.tex
app c

@@
 \end{equation}
-where $\integrand: \RR{M}  \times \RR{M}  \times[0,T] \rightarrow  \RRplus$ (for arbitrary $M$) with $\integrand: (\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto \integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$.
-We now introduce the function $\statez: [0,T] \rightarrow \RR{M}$ that is implicitly defined as an extremum of~\eqref{eq:el_gen} subject to the boundary condition $\statez(0) = \statez_0$.
+where $\integrand: \RR{M}  \times \RR{M}  \times[0,T] \rightarrow  \RRplus$
+	(for arbitrary $M$) with $\integrand:
+	(\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto
+	\integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$. We now introduce
+	the function $\statez: [0,T] \rightarrow \RR{M}$ that is implicitly defined
+	as an extremum of~\eqref{eq:el_gen} subject to the boundary condition
+	$\statez(0) = \statez_0$. \KTC{This previous sentence isn't correct. The
+	functional defined here has two arguments; the extremum can't be a single
+	argument, as it's a pair of arguments. You need to connect these by saying
+	they associate with the same time-dependent funciton.} 
 We introduce an arbitrary function $\variation : [0,T] \rightarrow \RR{M}$ with the boundary condition $\variationArg{0} = \bz$ and define a variation from the extremal by,

### 2019-10-12 main.tex
app c

@@
 \section{Derivation of the Euler--Lagrange equations}\label{appendix:eulerlagrange}
+	\KTC{In this section, $a$ and $b$ are sometimes used as time limits, while
+	$0$ and $T$ are used elsewhere. Please go through carefully and make this
+	consistent. I recommend using $a$ and $b$ because this is used within time
+	windows.}
+	\KTC{In this section, "extremum" is used a lot. You are just finding
+	stationary points, which could be local minima, local maxima, or saddle
+	points. Please avoid using "extemum" for this reason.} 
 This section details the derivation of the Euler--Lagrange equations. To this end, we consider the generic functional of the form,
@@
 	functional defined here has two arguments; the extremum can't be a single
-	argument, as it's a pair of arguments. You need to connect these by saying
-	they associate with the same time-dependent funciton.} 
-We introduce an arbitrary function $\variation : [0,T] \rightarrow \RR{M}$ with the boundary condition $\variationArg{0} = \bz$ and define a variation from the extremal by,
+	argument, as it's a pair of arguments. You need to connect these by defining
+	a new function that maps a time-dependent function to this functional
+	evaluated at the time-dependent function and its derivative..} 
+We introduce an arbitrary function $\variation : [0,T] \rightarrow \RR{M}$
+with the boundary condition $\variationArg{0} = \bz$ and define a variation
+from the extremum by
 \begin{align*}
@@
 %$$
-The objective is to now find $\overline{\statez}$ that makes $\mathcal{J}_{\delta}$ stationary.
-This can be done by differentiating with respect to $\delta$, and setting the result to zero,% Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
-$$
+The objective is to now find $\overline{\statez}$ that makes
+$\mathcal{J}_{\delta}$ stationary. This can be done by differentiating with
+respect to $\delta$ and setting the result to zero, i.e., % Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
+\begin{equation}\label{eq:stationaryOne}
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)=  0.
-$$
+\end{equation}
 %or equivalently,
@@
 We integrate the second term by parts,
-\begin{multline*}
-\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
-=  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon) ,T \big) \variation(T)  - \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(0,\epsilon),\statezDotBar(0,\epsilon) ,0\big) \variation(0) .
-\end{multline*}
-Noting that $\variationArg{0} = \bz$,
-\begin{multline*}
+\begin{align}\label{eq:gradient}
+	\begin{split}
 %\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg] dt 
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-=  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \bigg) \variation(t) \bigg]dt + \\ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon) ,T \big) \variation(T) .
-\end{multline*}
-Setting equal to zero to find the stationary points,
-$$
+	=&  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }
+	\big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big)
+	\variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial
+	\stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon)
+	,t \big) \bigg) \variation(t) \bigg]dt + \\ &\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon) ,T \big) \variation(T) . 
+	\end{split}
+\end{align}
+where we have used $\variationArg{0} = \bz$.
+Substituting Eq.~\eqref{eq:gradient} in the stationarity conditions~\eqref{eq:stationaryOne} yields
+\begin{equation}\label{eq:statOne}
   \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon),T \big) \variation(T)  = 0.
-$$
-By definition, we know $\statez(\cdot) \equiv \statezBar(\cdot;0)$ is a stationary point. Thus, setting $\epsilon = 0$,
- $$
+\end{equation}
+By construction, $\statez(\cdot) \equiv \statezBar(\cdot;0)$ is a stationary
+point; thus, setting $\epsilon = 0$ in Eq.~\eqref{eq:statOne} yields
+\begin{equation}
   \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(T),\statezDot(T) ,T\big) \variation(T)  = 0.
-$$
+\end{equation}
 %$$  \int_a^b \bigg[\frac{\partial \integrand  }{\partial \overline{\genstate}} \variation  - \frac{d}{dt}\frac{\partial \integrand}{\partial {\overline{\genstateDot}}}\variation \bigg]_{\epsilon = 0}dt + \bigg[\bigg[ \frac{\partial \integrand}{\partial {\overline{\genstateDot}}} \variation \bigg]_a^b\bigg]_{\epsilon = 0} = 0.
@@
 %$$ \int_a^b \bigg(\frac{\partial \integrand  }{\partial \genstate} \variation  - \frac{\partial \integrand}{\partial {\genstateDot}}\variation \bigg)dt +  \frac{\partial \integrand}{\partial {\genstateDot}} \variation \bigg|_{t=b} = 0.$$
-As $\variation$ is an arbitrary function, in order for the equality to hold we must have,
+As $\variation$ is an arbitrary function, this equality requires
 %\begin{equation}\label{eq:euler_lagrange_appendix}
@@
 %\end{equation}
-\begin{equation}\label{eq:euler_lagrange_appendix}
-\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz
-\end{equation}
-subject to,
-$$
+\begin{gather}\label{eq:euler_lagrange_appendix}
+\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete }
+	\big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg(
+	\frac{\partial \integrand}{\partial \stateyDotDiscrete }
+	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz  \\
 \statez(a) = \statez_a, \qquad \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \variation(b) \bigg]^T = \bz.
+\end{gather}
 %\genstate(a) = \genstate_a, \qquad \bigg[ \frac{\partial \integrand}{\partial {\genstateDot}} \bigg]^T_{t=b} = \bz.
-$$
-Equation~\eqref{eq:euler_lagrange_appendix} is known as the Euler--Lagrange equation.
-It states that, if $\statez$ is an extremal of $\mathcal{J}$, then it must satisfy~\eqref{eq:euler_lagrange_appendix}.
-It is emphasized that~\eqref{eq:euler_lagrange_appendix} is a necessary condition on $\statez$ to make $\mathcal{J}$ stationary, but it is not a sufficient condition.
-It is additionally noted that~\eqref{eq:euler_lagrange_appendix} provides a stationary point of $\mathcal{J}$, but the resulting stationary point can be either a maximum or a minimum.
+\KTC{I removed "subject to" as this is only used in an optimization-problem
+statement. Also, I don't understand the $z(a) = z_a$ conditions, or how that
+derives obviously from above. Can you clarify? Also note the inconsistent use
+of $a$, $b$, and $0$, $T$.}
+Equation~\eqref{eq:euler_lagrange_appendix} is known as the Euler--Lagrange
+equation. It states that, if $\statez$ is an extremum of $\mathcal{J}$
+\KTC{Again, note that $\mathcal J$ has two arguments, so it can't havec a
+single vector servec as an extremum. Fix needed}, then it must
+satisfy~\eqref{eq:euler_lagrange_appendix}. It is emphasized
+that~\eqref{eq:euler_lagrange_appendix} is a necessary condition on $\statez$
+to make $\mathcal{J}$ stationary, but it is not a sufficient condition. It is
+additionally noted that~\eqref{eq:euler_lagrange_appendix} provides a
+stationary point of $\mathcal{J}$, but the resulting stationary point can be
+either a maximum or a minimum. \KTC{See above. Strictly, these could be local
+minima, local maxima, or saddle points.}
 

### 2019-10-12 main.tex
address comments and appendix 3

@@
 \section{Derivation of the Euler--Lagrange equations}\label{appendix:eulerlagrange}
-	\KTC{In this section, $a$ and $b$ are sometimes used as time limits, while
-	$0$ and $T$ are used elsewhere. Please go through carefully and make this
-	consistent. I recommend using $a$ and $b$ because this is used within time
-	windows.}
-	\KTC{In this section, "extremum" is used a lot. You are just finding
-	stationary points, which could be local minima, local maxima, or saddle
-	points. Please avoid using "extemum" for this reason.} 
 This section details the derivation of the Euler--Lagrange equations. To this end, we consider the generic functional of the form,
 \begin{equation}\label{eq:el_gen}
-\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_0^T \integrand(\statey(t),\stateyDot(t),t)  dt,
-\end{equation}
-where $\integrand: \RR{M}  \times \RR{M}  \times[0,T] \rightarrow  \RRplus$
+\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_a^b \integrand(\statey(t),\stateyDot(t),t)  dt,
+\end{equation}
+where $\integrand: \RR{M}  \times \RR{M}  \times[a,b] \rightarrow  \RRplus$
 	(for arbitrary $M$) with $\integrand:
@@
 	\integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$. We now introduce
-	the function $\statez: [0,T] \rightarrow \RR{M}$ that is implicitly defined
-	as an extremum of~\eqref{eq:el_gen} subject to the boundary condition
-	$\statez(0) = \statez_0$. \KTC{This previous sentence isn't correct. The
+	the function $\statez: [a,b] \rightarrow \RR{M}$ with $\statez: \timeDummy \mapsto \statez(\timeDummy)$ along with $\statezDot \equiv d \statez /d\timeDummy$.
+We define this function to be a stationary point of~\eqref{eq:el_gen} (with $\statez$ being the first argument and $\statezDot$ being the second argument) subject to the boundary condition
+	$\statez(a) = \statez_a$. \KTC{This previous sentence isn't correct. The
 	functional defined here has two arguments; the extremum can't be a single
@@
 	a new function that maps a time-dependent function to this functional
-	evaluated at the time-dependent function and its derivative..} 
-We introduce an arbitrary function $\variation : [0,T] \rightarrow \RR{M}$
-with the boundary condition $\variationArg{0} = \bz$ and define a variation
-from the extremum by
+	evaluated at the time-dependent function and its derivative..}\EP{ Check this, I think this is more precise now.} 
+We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$
+with the boundary condition $\variationArg{a} = \bz$ and define a variation
+from the stationary point by
 \begin{align*}
@@
 \end{align*}
-We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{0} = \bz$.
+We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{a} = \bz$.
 We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)= 
-\int_0^T \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt .
+\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt .
 $$
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-= \int_0^T \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variationDot(t) \bigg]dt.
+= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variationDot(t) \bigg]dt.
 $$
@@
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-	=&  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }
+	=&  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }
 	\big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big)
@@
 	\stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon)
-	,t \big) \bigg) \variation(t) \bigg]dt + \\ &\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon) ,T \big) \variation(T) . 
+	,t \big) \bigg) \variation(t) \bigg]dt + \\ &\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) , 
 	\end{split}
 \end{align}
-where we have used $\variationArg{0} = \bz$.
+where we have used $\variationArg{a} = \bz$.
 Substituting Eq.~\eqref{eq:gradient} in the stationarity conditions~\eqref{eq:stationaryOne} yields
 \begin{equation}\label{eq:statOne}
-  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(T,\epsilon),\statezDotBar(T,\epsilon),T \big) \variation(T)  = 0.
-\end{equation}
-By construction, $\statez(\cdot) \equiv \statezBar(\cdot;0)$ is a stationary
+  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon),b \big) \variation(b)  = 0.
+\end{equation}
+By construction, $\statez(\cdot) \equiv \statezBar(\cdot;0)$ and $\statezDot(\cdot) \equiv \statezDotBar(\cdot;0)$, comprise a stationary
 point; thus, setting $\epsilon = 0$ in Eq.~\eqref{eq:statOne} yields
 \begin{equation}
-  \int_0^T  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(T),\statezDot(T) ,T\big) \variation(T)  = 0.
+  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b) ,b\big) \variation(b)  = 0.
 \end{equation}
@@
 	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz  \\
-\statez(a) = \statez_a, \qquad \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \variation(b) \bigg]^T = \bz.
+\statez(a) = \statez_a, \qquad \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \bigg]^T = \bz.
 \end{gather}
@@
 derives obviously from above. Can you clarify? Also note the inconsistent use
-of $a$, $b$, and $0$, $T$.}
+of $a$, $b$, and $0$, $T$.} \EP{Yes, sorry about the [a,b], I changed this recently and obviously missed things. As for the BC, 
+the requirement that $z(a)=z_a$ comes in first from the implicit definition of $z$, which is required to satisfy this BC. Then, in defining the variation, 
+we explicitly force the variation to be zero at the IC such that all possible functions satisfy the BC. We don't, however, make any such requirement on the end point 
+condition, $\eta(b)$ can be anything.
+This yields the endpoint boundary condition on the gradient (i.e., second term in Eq 56.).
+If we had a fixed final boundary condition, then the second term in Eq.
+56 would be replaced
+with this end point condition.} 
 Equation~\eqref{eq:euler_lagrange_appendix} is known as the Euler--Lagrange
-equation. It states that, if $\statez$ is an extremum of $\mathcal{J}$
-\KTC{Again, note that $\mathcal J$ has two arguments, so it can't havec a
-single vector servec as an extremum. Fix needed}, then it must
+equation. It states that, for ($\statez$,$\statezDot$) to define a stationary point of $\mathcal{J}$, then they must 
 satisfy~\eqref{eq:euler_lagrange_appendix}. It is emphasized
-that~\eqref{eq:euler_lagrange_appendix} is a necessary condition on $\statez$
+that~\eqref{eq:euler_lagrange_appendix} is a necessary condition on ($\statez,\statezDot$)
 to make $\mathcal{J}$ stationary, but it is not a sufficient condition. It is
 additionally noted that~\eqref{eq:euler_lagrange_appendix} provides a
-stationary point of $\mathcal{J}$, but the resulting stationary point can be
-either a maximum or a minimum. \KTC{See above. Strictly, these could be local
-minima, local maxima, or saddle points.}
+stationary point of $\mathcal{J}$, but the resulting stationary point could be a local minima, local maxima, or saddle point.
 

### 2019-10-12 cavity.tex
fixing punctuation

@@
 resulting in $36000$ unknowns for each conserved variable. Spatial discretization via the discontinuous Galerkin method yields a dynamical system 
-of the form,
+of the form
 $$\frac{d \state}{dt} = \mass_{\text{DG}}^{-1} \velocity_{\text{DG}} (\state),$$

### 2019-10-12 numerical_experiments.tex
through conclusions
fixed en dahshes
fixing punctuation

@@
 $T = 1$ the final time. 
-The problem setup is given by the initial conditions,
+The problem setup is given by the initial conditions
 
@@
 The coupled two-point boundary
-problem is solved via the Forward Backward Sweep Method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. 
+problem is solved via the forward--backward sweep method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. 
 \end{itemize}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_compare.pdf}
-\caption{Integrated $\elltwo$ error of various ROMs.}
+\caption{Space--time $\elltwo$ error of various ROMs.}
 \label{fig:sod_error_a}

### 2019-10-12 section4/spacetime_numerical.tex
fixed definition of objective function below 41
fixing punctuation

--- 2019-10-08 section4/spacetime_numerical.tex
@@
 It is noted that the boundary conditions are automatically satisfied through the definition of the $\spaceTimeAcronym$\ trial subspace.
-The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss-Newton method.
+The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss--Newton method.
 
@@
 \subsection{\spaceTimeAcronym: indirect solution approach}
-As opposed to the direct approach, the indirect approach directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime},
+As opposed to the direct approach, the indirect approach directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime} and sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
 \begin{equation}\label{eq:obj_gen_slab2}
@@
 Numerically solving the minimization problem requires the introduction of a quadrature rule for 
-discretization of the integral. We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ N )$ collocation points over the $n$th window, 
+discretization of the integral.
+We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ \EP{Originally this was divided by $N$, just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window
 $$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$
-Leveraging these collocation points, the indirect method for \methodAcronym\ with a \spaceTimeAcronym\ trial subspace yields,
+Leveraging these collocation points, the indirect method for \methodAcronym\ with an \spaceTimeAcronym\ trial subspace computes the generalized coordinates 
+$\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
 \begin{equation}\label{eq:obj_gen_slab2} 
 \begin{split}
-&\underset{\stgenstatey}{\text{minimize }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n}  \bigg) , %\\
+&\underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize }} \objectiveDiscreteSTArg{n} \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n}  \bigg) , %\\
 %&\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
@@
 %\end{equation}
-where the discrete objective function is given by, 
-$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
-\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T
+where the discrete objective function is given by
+$$\objectiveDiscreteSTArg{n}: \statey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
+\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg]^T 
 \stweightingMatArg{n} 
-\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
+\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg].
 $$
-The least-squares problem can again be solved, e.g., via the Gauss-Newton method. 
+%$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
+%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T
+%\stweightingMatArg{n} 
+%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
+%$$
+The least-squares problem can again be solved, e.g., via the Gauss--Newton method. 
 %The stationary points are given by,

### 2019-10-12 section4/spatial_numerical.tex
through updates of Sec 3
fixing punctuation

@@
 \subsubsection{Linear multistep schemes}
-\KTC{HERE} \EP{This section has the double superscript. We could make one of these a subscript, but then need to change the algorithms a bit as 
+\EP{This section has the double superscript. We could make one of these a subscript, but then need to change the algorithms a bit as 
 I use subscripts to denote iterations. Either is fine}
@@
 \end{align*}
-where $\stateFOMDiscreteArg{n,i} \approx \stateFOMArg{n}{\timeWindowArg{n}{i}} \in \RR{\fomdim}$ is the discrete approximation to the FOM state at the $i$th time instance over the $n$th window and $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time instance of the $n$th window,
+where $\stateFOMDiscreteArg{n,i} \approx \stateFOMArg{n}{\timeWindowArg{n}{i}} \in \RR{\fomdim}$ is the discrete approximation to the FOM state at the $i$th time instance over the $n$th window and $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time instance of the $n$th window and is defined as
 \begin{align*}
@@
 \end{align*}
-for arbitrary $p$ and $m$. The stacked vector of generalized coordinates over the $n$th time window are then defined as, 
+for arbitrary $p$ and $m$. The stacked vector of generalized coordinates over the $n$th time window are then defined as
 \begin{equation*}
@@
 In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the time window residual Jacobian will (almost) be a block-diagonal
-sparse matrix with the sparsity pattern,
+sparse matrix with the sparsity pattern
 \begin{equation*}
@@
 $$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T.$$
-This manipulation allows one to compute the action of the Jacobian on each column of $\basisspaceArg{n}$ by, e.g., a finite difference approximation,
+This manipulation allows one to compute the action of the Jacobian on each column of $\basisspaceArg{n}$ by, e.g., the finite difference approximation
 $$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$

### 2019-10-14 main.tex
minor changes to 3.3

@@
 Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.
-To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand,
+To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand
 \begin{align*}\label{eq:integrand_apx}

### 2019-10-14 section3/space_time.tex
through updates of Sec 3
fixing punctuation
minor changes to 3.3

@@
 \subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
+\KTC{HERE}
 %\KTC{Error in first equation. This is the approximate solution not the real
@@
 Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
-For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
-At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
+For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary condition at $t=0$ and enforces continuity between windows.
+At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as
  \begin{equation}\label{eq:stapprox}
@@
 \begin{split}
-&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} ).% \\ 
+&\underset{\stgenstateyArg{} \in \RR{\stdimArg{n}}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{} + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} ).% \\ 
 %      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =0

### 2019-10-14 section3/tclsrm_formulation.tex
fixing punctuation
minor notation change in 4.2.2 for lmsWidthArg

@@
 \end{equation*}
-where,
+where
 \begin{align*}
@@
 \end{align*} 
-subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler--Lagrange equations expose the fact that, for \spatialAcronym\
-trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
+subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}.
+\methodAcronym\ with \spatialAcronym\ trial subspaces can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a
 controller that enforces the residual minimization property. 

### 2019-10-14 section4/numerical_methods.tex
minor notation change in 4.2.2 for lmsWidthArg

@@
 the infinite dimensional problem into a finite-dimensional nonlinear
-programming (optimization) problem. Direct approaches thus ``discretize-then-optimize."
+programming (optimization) problem. Direct approaches thus ``discretize-then-optimize".
 Indirect methods, on the other hand, compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first order optimality conditions of the objective functional).

### 2019-10-14 section4/spatial_numerical.tex
minor notation change in 4.2.2 for lmsWidthArg

@@
 ``transcribing" the infinite dimensional optimization problem into a finite
-dimensional one by discretizing the state and objective function in time.
+dimensional one by discretizing the state and objective functional in time.
 The minimization problem is then reformulated as a (non)linear programming, or
@@
 (2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
-To achieve this agenda, first define a discrete time grid 
+To this end, we start by defining a discrete time grid 
 over each window by introducing
@@
 \begin{align*}
-&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i},\ldots,\stateFOMDiscreteArg{n,i-k^n(i)}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
-&\stateArgnt{n,0} = \begin{cases}
+&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i},\ldots,\stateFOMDiscreteArg{n,i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
+&\stateFOMDiscreteArg{n,0} = \begin{cases}
 \stateFOMDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
@@
 \end{align*}
-where $\stateFOMDiscreteArg{n,i} \approx \stateFOMArg{n}{\timeWindowArg{n}{i}} \in \RR{\fomdim}$ is the discrete approximation to the FOM state at the $i$th time instance over the $n$th window and $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time instance of the $n$th window and is defined as
+where $\stateFOMDiscreteArg{n,i} \approx \stateFOMArg{}{\timeWindowArg{n}{i}} \in \RR{\fomdim}$ is the discrete approximation to the FOM state at the $i$th time instance over the $n$th window and $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time instance of the $n$th window and is defined as
 \begin{align*}
@@
 \end{align*} 
-Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time step, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that
+Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time step, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme at the $i$th time instance over the $n$th window, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that
 define the specific type of multistep scheme. Examples of linear multistep schemes include Adams--Bashforth schemes and  
@@
 \objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-k^n(i)})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-k^n(i)}), \\
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
 &\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}+1} \rightarrow

### 2019-10-14 main.tex
thru 3

@@
 proper orthogonal decomposition (POD)~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}. 
-Thus, at any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\ trial subspace approximate the state as
+Thus, at any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\
+trial subspace approximate the FOM ODE solution as
 \begin{equation}\label{eq:affine_trialspace}
@@
 Thus, at any time instance $t\in[0,T]$, ROMs that employ the
-\spaceTimeAcronym\ trial subspace approximate the state as
+\spaceTimeAcronym\ trial subspace approximate the FOM ODE solution as
 \begin{equation}\label{eq:stapprox1}

### 2019-10-14 section3/space_time.tex
thru 3

@@
 \subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
-\KTC{HERE}
 %\KTC{Error in first equation. This is the approximate solution not the real
@@
 \end{equation}
-Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
-For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary condition at $t=0$ and enforces continuity between windows.
+Here $\stbasisArgnt{n} \in \RR{\fomdim \times
+\stdimArg{n}}\otimes\timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix {function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
+To enforce the initial condition and ensure solution continuity across time windows, we set $\stateInterceptSTArg{1} = \stateFOMIC$ and
+$\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for
+$n=2,\ldots,\nslabs$.
 At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as
@@
 \end{comment}
-Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm} implies that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
+Setting $\stspace^n \leftarrow \stspaceSTArg{n}$ in the
+\methodAcronym\ minimization problem~\eqref{eq:tclsrm} implies that
+\methodAcronym\ with \spaceTimeAcronym\ trial subspaces sequentially computes
+solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
 \begin{align}\label{eq:obj_gen_slab_spacetime}
@@
 \subsubsection{Stationary conditions} 
-The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$.
-Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
+The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial
+	subspaces is as follows: generalized coordinates for \spaceTimeAcronym\ trial subspaces 
+	comprise a vector in $\RR{\stdimArg{n}}$, while generalized coordinates for
+	the \spatialAcronym\ trial subspaces comprise a
+	time-dependent vector in
+	$\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$. 
+Thus, in the \spaceTimeAcronym\ case, the optimization problem is no longer
+	minimizing a functional with respect to a (time-dependent) function, but now is minimizing a function with 
 respect to vector.  As such, the first-order optimality conditions can be
 	derived using standard calculus. Differentiating the objective with respect
-	to $\stgenstatey$ and setting equal to zero yields 
+	to the generalized coordinates and setting the result equal to zero yields 
 \begin{equation}\label{eq:st_stationary}
@@
 %n=1. \end{cases} \end{equation}
-For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of equations, as opposed to a system of differential equations.
+Thus, for \spaceTimeAcronym\ trial subspaces, the stationary conditions
+	comprise an algebraic system of equations, as opposed to a system of
+	differential equations as with \spatialAcronym\ trial subspaces.

### 2019-10-14 section3/tclsrm_formulation.tex
thru 3

@@
 \subsection{\methodAcronym\ summary}
-This section outlined the \methodAcronym\ method for model reduction.
-In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows; i.e.,
+This section outlined the \methodAcronym\ method for model reduction. In
+summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE
+residual within the range of a space--time trial over time
+windows subject to boundary conditions associated with continuity across time
+windoes; i.e.,  
 %\methodAcronym\  
@@
 \methodAcronym\ sequentially computes approximate solutions
-$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that satisfy
-\begin{equation*}
-\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
-\mathcal{J}^n(\stateyArgnt{}),
-\end{equation*}
-where
-\begin{align*}
-\mathcal{J}^n &\vcentcolon \statey \mapsto
-\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
-- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
-\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
-&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
-\RRplus,
-\end{align*}
-subject to the boundary conditions.
-For \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
-\begin{align*}\
-&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
-\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
- &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
-\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
- \adjointArg{n}{t} = \nonumber \\ 
-&-\basisspaceTArg{n} \big[
-\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
-\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
- \big( \basisspace \genstateDotArg{n}{t}  -
-\velocity(\veloargsromn) \big),  &
-%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
-%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
-%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
-%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
-\end{align*} 
-subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}.
+$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that comprise
+solutions to the optimization problem \eqref{eq:tclsrm}.
+\KTC{I removed repeated equations here and just referred to their previous
+definitions. This will cut down on the text.}
+%\begin{equation*}
+%\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
+%\mathcal{J}^n(\stateyArgnt{}),
+%\end{equation*}
+%where
+%\begin{align*}
+%\mathcal{J}^n &\vcentcolon \statey \mapsto
+%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
+%- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
+%\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
+%&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
+%\RRplus,
+%\end{align*}
+%subject to the boundary conditions.
+
+For \spatialAcronym\ trial subspaces, the stationary conditions for the
+residual minimization problem can be derived via the Euler--Lagrange equations
+and yield the system \eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} over the $n$th window for $t \in
+[\timeStartArg{n},\timeEndArg{n}]$.
+%\begin{align*}\
+%&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
+%\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
+% &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
+%\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
+% \adjointArg{n}{t} = \nonumber \\ 
+%&-\basisspaceTArg{n} \big[
+%\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
+%\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
+% \big( \basisspace \genstateDotArg{n}{t}  -
+%\velocity(\veloargsromn) \big),  &
+%%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
+%%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
+%%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
+%%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
+%\end{align*} 
+%subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. 
 \methodAcronym\ with \spatialAcronym\ trial subspaces can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
 controller that enforces the residual minimization property. 
-Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces.
-For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.
+Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\
+trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the
+generalized coordinates are no longer functions, the stationary conditions
+correspond to the system of algebraic equations \eqref{eq:st_stationary}.

### 2019-10-14 section4/numerical_methods.tex
thru 3

@@
 \section{Numerical solution techniques}\label{sec:numerical_techniques}
+\KTC{HERE}
 The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 

### 2019-10-14 section4/numerical_methods.tex
thru 4.1

--- 2019-10-14 section4/numerical_methods.tex
@@
 \section{Numerical solution techniques}\label{sec:numerical_techniques}
-The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
-\methodAcronym\ displays commonalities with optimal
-control problems, and as such a large existing body of work can be leveraged. In this section we 
-outline a variety of such techniques. We focus primarily on 
-%\KTC{earlier we kept
-%saying 
-%direct (i.e.,
-%	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
-%	Let's be consistent when we re-introduce it here.}
-direct (discretize-the-optimize) and indirect (optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
+Given the formulation of the \methodAcronym\ method presented in Section
+\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
+Due to the commonalities that the \methodAcronym\ method exhibits with
+optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
+state this.} the derivation of these techniques can leverage a large existing
+body of work. In this section, we outline a variety of such techniques. We
+focus primarily on 
+direct (i.e., discretize-the-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces.
@@
 Numerical techniques to
-solve optimal control problems can be classified as either
+solve optimal-control problems can be classified as either
 \textit{direct} or \textit{indirect}
@@
 the infinite dimensional problem into a finite-dimensional nonlinear
-programming (optimization) problem. Direct approaches thus ``discretize-then-optimize".
-Indirect methods, on the other hand, compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first order optimality conditions of the objective functional).
+programming (optimization) problem. Direct approaches thus
+``discretize-then-optimize''.
+In contrast, indirect methods compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first-order optimality conditions of the objective functional).
 Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
-``indirectly". A variety of discretization/solution techniques are possible for both direct and indirect methods. Collocation methods,
-finite elements, spectral methods, shooting methods, etc., are all examples of
+``indirectly". A variety of both discretization/solution \KTC{why are these
+lumped together? Could we just say ``discretization'', ``solution'', or
+``discretization and solutin''?} techniques are possible
+for both direct and indirect methods. Collocation methods,
+finite-element methods, spectral methods, and shooting methods are all examples of
 plausible solution techniques.
@@
 In the present context, we investigate both direct and indirect methods to solve \methodAcronym\ with \spatialAcronym\ trial
-subspaces. We consider: 
-\begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A direct approach leveraging linear multistep methods to directly solve the optimization problem~\eqref{eq:tclsrm}.
-\item \textit{Indirect methods (optimize-then-discretize)}: An indirect approach leveraging the forward--backward sweep algorithm to solve the Euler--Lagrange equations~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}.
-
+subspaces. In particular, we consider:
+\begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A direct approach leveraging linear multistep methods to directly solve the optimization problem~\eqref{eq:obj_gen_slab}.
+Section~\ref{sec:direct} outlines this approach.
+\item \textit{Indirect methods (optimize-then-discretize)}: An indirect
+	approach leveraging the forward--backward sweep algorithm to solve the
+		Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}.
+Section~\ref{sec:indirect} outlines this technique.
 \end{itemize} 
-We emphasize that a variety of other approaches exist. Future work will comprise a detailed exploration
-of additional approaches.
-Section~\ref{sec:direct} outlines a direct solution approach based on linear
-multistep methods, while Section~\ref{sec:indirect} outlines an indirect
-solution approach based on the forward--backward sweep algorithm.
+We note that a variety of other approaches exist, and their investigation
+comprises the subject of future work.
 

### 2019-10-14 section3/space_time.tex
minor fixes in response to comments

@@
 Thus, for \spaceTimeAcronym\ trial subspaces, the stationary conditions
-	comprise an algebraic system of equations, as opposed to a system of
+	comprise a system of algebraic equations, as opposed to a system of
 	differential equations as with \spatialAcronym\ trial subspaces.

### 2019-10-14 section3/tclsrm_formulation.tex
minor fixes in response to comments

@@
 \subsection{\methodAcronym\ summary}
-This section outlined the \methodAcronym\ method for model reduction. In
+This section outlined the \methodAcronym\ \approachKwd\ for model reduction. In
 summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE
-residual within the range of a space--time trial over time
-windows subject to boundary conditions associated with continuity across time
-windoes; i.e.,  
+residual within the range of a space--time trial subspace over time
+windows; i.e.,  
 %\methodAcronym\  
@@
 solutions to the optimization problem \eqref{eq:tclsrm}.
-\KTC{I removed repeated equations here and just referred to their previous
-definitions. This will cut down on the text.}
+%\KTC{I removed repeated equations here and just referred to their previous
+%definitions. This will cut down on the text.}
 %\begin{equation*}
@@
 %subject to the boundary conditions. 
-
 For \spatialAcronym\ trial subspaces, the stationary conditions for the

### 2019-10-14 section4/numerical_methods.tex
minor fixes in response to comments

@@
 \section{Numerical solution techniques}\label{sec:numerical_techniques}
-Given the formulation of the \methodAcronym\ method presented in Section
-\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
-Due to the commonalities that the \methodAcronym\ method exhibits with
-optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
-state this.} the derivation of these techniques can leverage a large existing
-body of work. In this section, we outline a variety of such techniques. We
-focus primarily on 
+We now consider numerical-solution techniques for the \methodAcronym\ \approachKwd.
+We focus primarily on 
 direct (i.e., discretize-the-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
-
+%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
+%Given the formulation of the \methodAcronym\ method presented in Section
+%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
+%Due to the commonalities that the \methodAcronym\ method exhibits with
+%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
+%state this.} the derivation of these techniques can leverage a large existing
+%body of work. In this section, we outline a variety of such techniques. 
 \subsection{\spatialAcronym: direct and indirect methods}
+\methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
 Numerical techniques to
@@
 Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
-``indirectly". A variety of both discretization/solution \KTC{why are these
-lumped together? Could we just say ``discretization'', ``solution'', or
-``discretization and solutin''?} techniques are possible
+``indirectly". A variety of both discretization and solution techniques are possible
 for both direct and indirect methods. Collocation methods,

### 2019-10-15 section4/spacetime_numerical.tex
minor changes to space--time solution methods

@@
 \begin{remark}
-For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and a uniform quadrature weights are used, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
+For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and uniform quadrature weights are used, and the weighting matrix $\stweightingMatArg{n} = \mathbf{I}$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG with $\lspgWeightingSTArg{\cdot} = \mathbf{I}$.
+For equivalence conditions in the case where $\lspgWeightingSTArg{\cdot} \ne \mathbf{I}$, the objective function~\eqref{eq:obj_gen_lms_final_st} can be modified to enable weighted space--time norm.
 \end{remark}

### 2019-10-15 section4/spatial_numerical.tex
thru 4.1
minor changes to space--time solution methods

@@
 \subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
+\KTC{HERE}
 \KTC{Does the time grid definted in this section have any relationship to the
@@
 
-Direct approaches solve~\eqref{eq:tclsrm} by
+Direct approaches solve the optimization problem~\eqref{eq:tclsrm} by
 ``transcribing" the infinite dimensional optimization problem into a finite
@@
 I use subscripts to denote iterations. Either is fine}
-Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time-instances.  
+Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  
 %Discretizing the FOM ODE~\eqref{eq:FOM} with a linear $k$-step scheme at the $i$th time-instance over the $n$th time slab yields,
@@
 where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. 
-\methodAcronym\ sequentially computes the approximate solutions $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, $n=1,\ldots,\nslabs$, as the solution to the minimization problems
+\methodAcronym\ sequentially computes the approximate solutions $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, $n=1,\ldots,\nslabs$, as the solution to the minimization problem
 \begin{align*}
@@
 \begin{remark}
-For the limiting case where the window size is equivalent to the time-step size, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
-quadrature weights are used, \methodAcronym\ recovers LSPG.
+For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
+quadrature weights are used, \methodAcronym\ with \spatialAcronym\ trial subspaces solved via the direct approach recovers LSPG.
 \end{remark} 
@@
 \objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
-\equiv
+=
 \bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg].
@@
 Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method.
-The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the time window residual-Jacobian and form the normal equations, and (3) solve the normal equations and update the state.
+The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and form the normal equations, and (3) solve the normal equations and update the state.
 
-The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the time window residual-Jacobians.
-In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the time window residual Jacobian will (almost) be a block-diagonal
+The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the Jacobian of the residual over the time window.
+In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will an almost block-diagonal
 sparse matrix with the sparsity pattern
@@
 \end{equation*}
-This sparsity pattern can be leveraged to assemble the time window residual-Jacobian from local residual Jacobians. 
+This sparsity pattern can be leveraged to assemble the Jacobian of the residual over the time window from Jacobians of the residual at single time instances. 
 Further, the sparsity pattern can be 
@@
 convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
-contains three parameters: the damping factor $\rho$, the growth factor
-$\fbsmGrowth$, and the decay factor $\fbsmDecay$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
+contains three parameters: the damping factor $\rho \le 1$, the growth factor
+$\fbsmGrowth \ge 1$, and the decay factor $\fbsmDecay \ge 1$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
 closer $\rho$ is to unity, the faster the resulting algorithm will converge.
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $\epsilon$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
-factor, $\fbsmDecay$} 
+\Input{tolerance, $\epsilon$; damping factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
+factor, $\fbsmDecay \ge 1$} 
 \Output{Stationary point, $\genstateArgnt{n}$ }
@@
 \subsubsection{Numerical considerations for the forward and backward problems}
-The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
+The FBSM requires solving the systems~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
 The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
-an efficient method for computing action of the Jacobian transpose on a vector.
+an efficient method for computing the action of the Jacobian transpose on a vector.
 

### 2019-10-15 section4/spacetime_numerical.tex
latest

@@
 %\end{equation}
-It is noted that the boundary conditions are automatically satisfied through the definition of the $\spaceTimeAcronym$\ trial subspace.
+It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
 The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss--Newton method.
@@
 %n=1. \end{cases} \end{equation} 
-
 \subsection{\spaceTimeAcronym\: summary}
-\spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window.
+\spaceTimeAcronym\ trial subspaces yield a series of space--time systems of algebraic equations over each window.
 As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces,

### 2019-10-15 section3/general_form.tex
updated figures

@@
 \caption{Partitioning of the time
-	domain into time windows. \KTC{Improve this by making the time interval
-	finite (i.e., $[0,T]$) and introduce a partitioning. You should also match the font size with the rest of
-	the text. See Fig 1 of our paper ``Data-driven time parallelism via
-	forecasting", for example.}} 
+	domain into time windows.} 
 \label{fig:slab_fig} 

### 2019-10-15 figs/make_grid.py

@@ -6,10 +6,10 @@ from matplotlib import rc
 rc('text', usetex=True)
 rc('text.latex', preamble='\usepackage{amsmath},\usepackage{amssymb}')
 import os
-axis_font = {'size':16,'family':'serif'}
+axis_font = {'size':12,'family':'serif'}
 mpl.rc('font',family='serif')
 rc('text', usetex=True)
-axis_font = {'size':'14'}
+axis_font = {'size':'12'}
 close("all")
 import matplotlib.patches as patches
 fig,ax = plt.subplots(1)
@@ -35,4 +35,4 @@ ax.axis('off')
 #ax.set_xticks([])
 #ax.set_yticks([])
 savefig('time_grid.pdf')
-show()
+#show()

### 2019-10-16 section4/numerical_methods.tex
small changes

@@
 subspaces. In particular, we consider:
-\begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A direct approach leveraging linear multistep methods to directly solve the optimization problem~\eqref{eq:obj_gen_slab}.
+\begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A
+			direct approach that leverages linear multistep methods to directly solve the optimization problem~\eqref{eq:obj_gen_slab}.
 Section~\ref{sec:direct} outlines this approach.
 \item \textit{Indirect methods (optimize-then-discretize)}: An indirect
-	approach leveraging the forward--backward sweep algorithm to solve the
+	approach that leverages the forward--backward sweep algorithm to solve the
 		Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}.

### 2019-10-16 section4/spatial_numerical.tex
updated figures
small changes

@@
-\subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
-\KTC{HERE}
-\KTC{Does the time grid definted in this section have any relationship to the
-time grid used for time discretization? Haven't gotten past HERE below, but
-wanted to check (maybe it's made clear later).}
-\EP{This time grid becomes the time discretization over each window. At this point in the paper, 
-there is no notion of a time grid for \methodAcronym}
+\subsection{\spatialAcronym\ trial subspaces: direct solution approach}\label{sec:direct}
 
-Direct approaches solve the optimization problem~\eqref{eq:tclsrm} by
+Direct approaches solve optimization problem~\eqref{eq:tclsrm} by
 ``transcribing" the infinite dimensional optimization problem into a finite
@@
 
-In the present context, the most straightforward direct solution approach consists of the 
-following steps: (1) numerically discretize the FOM ODE (and hence the \textit{integrand} of~\eqref{eq:obj_gen_slab}) and 
+In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
+following steps: (1) numerically discretize the FOM ODE (and hence the
+\textit{integrand} of the objective function in problem~\eqref{eq:obj_gen_slab}) and 
 (2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
-To this end, we start by defining a discrete time grid 
-over each window by introducing
-$\nstepsArg{n}+1$ strictly increasing time instances described by
-$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with
+To this end, we define a time grids
+$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$,
+$n=1,\ldots,\nslabs$ that
+satisfy 
+$\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
+ = \timeEndArg{n}$.
 %\KTC{use cdots, not
@@
 %aren't identically the same:}
-\begin{align}
-&\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
-\timeWindowArg{n}{0} = \timeStartArg{n},\;
-\timeWindowArg{n}{\nstepsArg{n}} = \timeEndArg{n} . 
-%\timeWindowArg{n}{0} = \timeStartArg{n}, \timeWindowArg{n}{\nstepsArg{n}} =
-%\timeEndArg{n}
-\end{align}
-Figure~\ref{fig:slab_fig2} depicts this discretization.
+Figure~\ref{fig:slab_fig2} depicts such a discretization.
 We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
-other time-stepping methods (e.g., Runge--Kutta) follows closely. 
+other time-integration methods (e.g., Runge--Kutta) follows closely. 
 \begin{figure} 
 \begin{centering} 
-\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
-	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 3$ for all $n$.} 
+\includegraphics[trim={0.0cm 4.5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
+	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 2$ for all $n$.} 
 \label{fig:slab_fig2} 
@@
 \subsubsection{Linear multistep schemes}
-\EP{This section has the double superscript. We could make one of these a subscript, but then need to change the algorithms a bit as 
-I use subscripts to denote iterations. Either is fine}
+\KTC{HERE} I use subscripts to denote iterations. Either is fine}
 Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  
@@
 
-\subsection{\spatialAcronym: indirect solution approach}\label{sec:indirect}
+\subsection{\spatialAcronym\ trial subspaces: indirect solution approach}\label{sec:indirect}
 In contrast to the direct approach,

### 2019-10-16 section4/spatial_numerical.tex
bug fix
comment

@@
 \subsubsection{Linear multistep schemes}
-\KTC{HERE} I use subscripts to denote iterations. Either is fine}
 Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  
@@
 %\end{equation}
-Employing a linear multistep method for temporal discretization of the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
+Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
+\KTC{Note: the notation here is not consistent with what was defined earlier
+for a discrete residual. See email.}
 \begin{align*}

### 2019-10-16 section4/spatial_numerical.tex
add linear/nonlinear comment

@@
 \end{align*}
-where $\stateFOMDiscreteArg{n,i} \approx \stateFOMArg{}{\timeWindowArg{n}{i}} \in \RR{\fomdim}$ is the discrete approximation to the FOM state at the $i$th time instance over the $n$th window and $\residLMSArg{n,i}$ is the discrete linear multistep residual over the $i$th time instance of the $n$th window and is defined as
+where $\stateFOMDiscreteArg{n,i} (\approx
+\stateFOMArg{}{\timeWindowArg{n}{i}})
+\in \RR{\fomdim}$ and $\residLMSArg{n,i}$ denotes the FOM O$\Delta$E
+residual over the $i$th time instance of the $n$th window 
+defined as
 \begin{align*}
@@
 \end{align*} 
-Additionally, $\Delta t^{n,i} = \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$ is the time step, $\lmsWidthArg{n}{i}$ is the width of the multistep scheme at the $i$th time instance over the $n$th window, and $\alpha^{n,i}_j$ and $\beta^{n,i}_j$ are coefficients that
-define the specific type of multistep scheme. Examples of linear multistep schemes include Adams--Bashforth schemes and  
-Adams--Moulton schemes. 
+Here, $\Delta t^{n,i} \defeq \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$
+denotes the time step, $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window, and
+$\alpha^{n,i}_j,\beta^{n,i}_j\in\RR{}$ denote coefficients that 
+define the specific type of multistep scheme. 
 \begin{remark}\label{remark:LMS}
-For notational simplicity, we assume the linear multistep method employs time instances from only within the current time window.
-In general, the linear multistep method at the $n$th window could be designed to employ multiple time instances from previous time windows.
+For notational simplicity, we assume the linear multistep method employs time
+	instances from only within the current time window. In general, the linear
+	multistep method at the $n$th window could be designed to employ multiple
+	time instances from previous time windows. \KTC{What are the implications of
+	this? I suggest removing this and adding a simple condition, maybe with a
+	footnote.}
 \end{remark}
@@
 where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights.
-\methodAcronym\ sequentially computes the approximate solutions $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, $n=1,\ldots,\nslabs$, as the solution to the minimization problem
+\KTC{Don't these weights need to be non-negative to guarantee the integral is
+non-negative? If so, then specify $\quadWeightsLMSScalarArg{n}{i} \in\RR{}_+$}
+\methodAcronym\ sequentially computes the solutions
+$\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
+$n=1,\ldots,\nslabs$ that satisfy
+	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
+	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
+	this should be the discrete solution, not the exact solution evaluated at
+	the final time.}
 \begin{align*}
-&\underset{\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}} \in \trialspace}{\text{minimize } }
+	&\underset{(\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}}{\text{minimize } }
 \objectiveArgLMS{n} (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \\
@@
 %\approxstateIC & n=1. \end{cases} \end{align*}
-Defining $\genstateDiscreteArg{n,i} \in \RR{\romdim}$ to be the discrete approximation to the generalized coordinates at the $i$th time instance over the $n$th window, 
-the minimization problem can be written equivalently as
+	Equivalently, WLS sequentially computes the generalized
+	coordinates
+	$\genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\timeWindowArg{n}{i}}$,
+	$n=1,\ldots,\nslabs$ 
+with $\genstateDiscreteArg{n,i}(\approx
+	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdim}$
+	that satisfy
 \begin{equation}\label{eq:obj_gen_lms_final}
 \begin{split}
-& \underset{\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}}}{\text{minimize } }
+	&
+	\underset{(\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdim}\otimes \RR{\lmsWidthArg{n}{i}+1}}{\text{minimize } }
 \objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}) \\ 
@@
 %\end{equation}
-The optimization problem takes the form of a \textit{weighted (non)linear least-squares problem}, 
-in where the weights correspond the quadrature weights.  
-Before proceeding, it is worth mentioning that~\eqref{eq:obj_gen_lms_final} corresponds to a least-squares problem due
-to the reduction in \textit{spatial} complexity, as opposed to the reduction
-in temporal complexity (the same  number of temporal constraints and temporal 
-degrees of freedom have been introduced).
+The optimization problem takes the form of a \textit{weighted least-squares
+	problem}. \KTC{I got rid of (non)linear. If it's either linear or nonlinear,
+	then omit it as an adjective, as this is no longer descriptive. }
+We emphasize that optimization problem~\eqref{eq:obj_gen_lms_final} associates
+	with an \spatialAcronym\ trial subspace characterized by a reduction in
+	spatial complexity, but no reduction in temporal complexity.
 
@@
 For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
-quadrature weights are used, \methodAcronym\ with \spatialAcronym\ trial subspaces solved via the direct approach recovers LSPG.
+quadrature weights are used, \methodAcronym\ with \spatialAcronym\ trial
+	subspaces solved via the direct approach recovers LSPG. \KTC{Not exactly. To
+	make this LSPG as presented earlier, also require the time grid here to
+	align with the time grid used for time integration earlier. Also, this
+	formulation requires allowing the stencil to go back to time steps at
+	previous windows.} 
 \end{remark} 
 \subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
+	\KTC{If the dynamics are linear, doesn't this just yield a linear
+	least-squares problem that can be solved directly? We need to present this
+	more clearly.}
 Discretization through linear multistep methods (as well as other techniques) 

### 2019-10-16 section4/spatial_numerical.tex
comment

@@
 
-To describe the Gauss--Newton method, we first define a ``vectorization" function that takes a tuple of time-local solutions at each time instance and outputs 
-a concatenated vector, 
+To describe the Gauss--Newton method, we first define a ``vectorization" function 
 \begin{align*}
  \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
-&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm},
+&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm}.
 \end{align*}
-for arbitrary $p$ and $m$. The stacked vector of generalized coordinates over the $n$th time window are then defined as
+The vectorized generalized coordinates over the $n$th time window are then
+defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
 \begin{equation*}
@@
 We now define the weighted space--time residual over the entire window as
+\KTC{To make this valid for seeping back into the previous time window, the
+second argument could be omitted here (okay because there is already a
+superscript on the residual) and arbitrary previous solutions could be
+included in the first residual here.}
 \begin{equation*}
@@
 where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
-It is worth noting that, by design,
+It is worth noting that, by design, \KTC{This is incorrect. Missing a 1/2
+factor}
 \begin{equation*}
@@
 In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will an almost block-diagonal
-sparse matrix with the sparsity pattern
+sparse matrix with the sparsity pattern \KTC{number of residual arguments
+inconsistent with above}
 \begin{equation*}
@@
 Further, the sparsity pattern can be 
-leveraged to efficiently compute the Jacobian matrix-matrix product in the normal equations.
+leveraged \KTC{split infinitive:} to efficiently compute the Jacobian
+\KTC{shoudl be an en-dash:} matrix-matrix product in the normal equations.
+\KTC{We should never be solving the normal equations. This squares the
+condition number!}
 It is also worth noting that the 
-normal equations will also (almost) be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
+normal equations will also (almost) \KTC{What does ``almost'' block diagonal
+mean? This is imprecise} be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
 

### 2019-10-16 section4/spatial_numerical.tex
grammar

@@
 \begin{remark}\label{remark:gaussnewton}\textit{(Acceleration of the Gauss--Newton Solve)}\\
-The principle cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
-reducing this computational burden exist; Jacobian-free Newton-Krylov methods~\cite{jfnk}, Broyden's method~\cite{broyden} (as explored in Ref.~\cite{carlberg_thesis}, Appendix A), and frozen Jacobian approximations are several such examples.
-Further, the space--time formulation exposes an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM.
+	The \KTC{this should be `principal', not `principle':} principle cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
+reducing this computational burden exist; Jacobian-free Newton-Krylov
+	methods~\cite{jfnk}, Broyden's method~\cite{broyden} (as explored in
+	Ref.~\cite{carlberg_thesis}, Appendix A), and frozen Jacobian approximations
+	are several such examples.
+\KTC{Expose used twice in this sentence} Further, the space--time formulation exposes an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM.
 The investigation of these additional, potentially more efficient, solution algorithms will be a topic of future work.

### 2019-10-16 section4/spatial_numerical.tex
workin on section 4.2.1

@@
 \begin{align*}
-&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i},\ldots,\stateFOMDiscreteArg{n,i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
+&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteArg{n,i-1},\ldots,\stateFOMDiscreteArg{n,i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
 &\stateFOMDiscreteArg{n,0} = \begin{cases}
@@
 \begin{align*}
-\residArg{n,i} &: (\stateyDiscreteArgnt{i},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
+\residArg{n,i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
                &: \RR{\fomdim} \otimes \RR{\lmsWidthArg{n}{i}+1} \rightarrow \RR{\fomdim}. 
@@
 Here, $\Delta t^{n,i} \defeq \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$
-denotes the time step, $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window, and
+denotes the time step, $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window and
 $\alpha^{n,i}_j,\beta^{n,i}_j\in\RR{}$ denote coefficients that 
 define the specific type of multistep scheme.
-\begin{remark}\label{remark:LMS}
-For notational simplicity, we assume the linear multistep method employs time
-	instances from only within the current time window. In general, the linear
-	multistep method at the $n$th window could be designed to employ multiple
-	time instances from previous time windows. \KTC{What are the implications of
-	this? I suggest removing this and adding a simple condition, maybe with a
-	footnote.}
-\end{remark}
+To handle the case where the linear multistep method employs time instances from the previous window(s), we enforce the condition on the indices $(n,i) = (n-1,\nstepsArg{n-1}+i)$ for $i<0$.
+%\begin{remark}\label{remark:LMS}
+%For notational simplicity, we assume the linear multistep method employs time
+%	instances from only within the current time window. In general, the linear
+%	multistep method at the $n$th window could be designed to employ multiple
+%	time instances from previous time windows. \KTC{What are the implications of
+%	this? I suggest removing this and adding a simple condition, maybe with a
+%	footnote.}
+%\end{remark}
 %Employing a linear multistep method for temporal discretization of the FOM ODE leads to the FOM O$\Delta$E over the $n$th slab,
@@
 %\end{align*} 
-Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
+\EP{In the process of reworking this slightly, need to take care for the case where we are employing multiple time steps 
+from the previous window. Still cleaning up}
+\methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
+$\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
+$n=1,\ldots,\nslabs$ that satisfy
+%	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
+%	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
+%	this should be the discrete solution, not the exact solution evaluated at
+%	the final time.} \EP{You are correct, I think there is an error now in the minimize term though.
+It was $\in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}$, it should be $\trialspace\otimes \RR{\nstepsArg{n}+1}$}
+\begin{align*}
+	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}{i}}}{\text{minimize } }
+\objectiveArgLMS{n} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}). 
+%%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
+%%\begin{cases} \basisspaceArg{n} \spatialICDiscrete + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
+%\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} 
+\end{align*}
+%Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
+%\EP{Had to change the definition of the objective function to reflect stepping back into previous window}
+In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
 \begin{equation}\label{eq:obj_lms}
 \begin{split} 
-\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
-&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}+1} \rightarrow
+\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i};\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i};\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
+&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
 \RR{}_+, 
@@
 \end{equation}
+where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights and, for $i \le 0$ in $\objectiveArgLMS{n}$, we take
+\begin{equation}\label{eq:wls_spatial_direct_args}
+\stateyDiscreteArgnt{i}  \equiv 
+\begin{cases}
+\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
+\basisspaceArg{n} \basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 1.
+\end{cases}
+\end{equation}
+Note that the boundary conditions are automatically satisfied by Eq.~\eqref{eq:wls_spatial_direct_args}.
+
 %where 
@@
 %is the discrete residual resulting from the linear multistep scheme and 
-where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights.
-\KTC{Don't these weights need to be non-negative to guarantee the integral is
-non-negative? If so, then specify $\quadWeightsLMSScalarArg{n}{i} \in\RR{}_+$}
-\methodAcronym\ sequentially computes the solutions
-$\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
-$n=1,\ldots,\nslabs$ that satisfy
-	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
-	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
-	this should be the discrete solution, not the exact solution evaluated at
-	the final time.}
-\begin{align*}
-	&\underset{(\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}}{\text{minimize } }
-\objectiveArgLMS{n} (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \\
-& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
-\begin{cases} \basisspaceArg{n} \spatialIC + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
-\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} \end{align*}
 %\methodAcronym\ solves the minimization problem over each time window, 
@@
 %\approxstateIC & n=1. \end{cases} \end{align*}
-	Equivalently, WLS sequentially computes the generalized
+	Equivalently, \methodAcronym\ sequentially computes the generalized
 	coordinates
-	$\genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\timeWindowArg{n}{i}}$,
+	$\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\timeWindowArg{n}{i}}$,
 	$n=1,\ldots,\nslabs$ 
@@
 	&
-	\underset{(\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdim}\otimes \RR{\lmsWidthArg{n}{i}+1}}{\text{minimize } }
-\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}) \\ 
-& \text{subject to} \hspace{0.2 in}
-\genstateyDiscreteArg{0} =
-\begin{cases} \spatialIC & n = 2,\ldots,\nslabs \\
-\genstateICOne& n=1. \end{cases} 
+	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdim}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
+\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
+%& \text{subject to} \hspace{0.2 in}
+%\genstateyDiscreteArg{0} =
+%\begin{cases} \genspatialICDiscrete  & n = 2,\ldots,\nslabs \\
+%\genstateICOne& n=1. \end{cases} 
 \end{split}
@@
 \begin{remark}
-For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
-quadrature weights are used, \methodAcronym\ with \spatialAcronym\ trial
+For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, uniform 
+quadrature weights are used, and uniform bases and weighting matrices are used, 
+\methodAcronym\ with \spatialAcronym\ trial
 	subspaces solved via the direct approach recovers LSPG. \KTC{Not exactly. To
@@
 	formulation requires allowing the stencil to go back to time steps at
-	previous windows.} 
+	previous windows.} \EP{I agree with the second comment, to be 100\% precise we need to go back into other time windows.
+Clearly you can do this if you would like, and I think the comment clarifies this.
+Hopefully can avoid this now with the condition you mentioned.
+I think the first comment is not required.
+The LSPG ``approach" minimizes the residual at a time instance; it isn't tied to the description we gave earlier.}
 \end{remark} 
@@
 defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
+\EP{The notation for space--time in 2.4 uses bars. I don't want to confuse these generalized coordinates with those, they conceptually two different things} 
 \begin{equation*}

### 2019-10-16 section4/spatial_numerical.tex
changes to 4.2

@@
 including collocation approaches, spectral  methods, and genetic algorithms.  
-
+\EP{Make sure 4.2.1 is settled on before this}
 In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
@@
 Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
-\KTC{Note: the notation here is not consistent with what was defined earlier
-for a discrete residual. See email.}
+\footnote{
+To maintain notational simplicity in this section, and handle the case where the linear multistep method employs time instances from the previous window(s), for superscripts $n=1,\ldots,\nslabs$ and $i<0$, we set the superscripts to be $n =n-1, i = \nstepsArg{n-1}+i$
+% $\stateFOMDiscreteLMSArg{n}{i} = \stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i}$, along with $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$.
+} 
 \begin{align*}
-&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteArg{n,i-1},\ldots,\stateFOMDiscreteArg{n,i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
-&\stateFOMDiscreteArg{n,0} = \begin{cases}
-\stateFOMDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
+&\residLMSWArg{n}{i} (\stateFOMDiscreteLMSArg{n}{i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
+&\stateFOMDiscreteLMSArg{n}{0} = \begin{cases}
+\stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
 \stateFOMIC & n=1, \end{cases}
 \end{align*}
-where $\stateFOMDiscreteArg{n,i} (\approx
+where $\stateFOMDiscreteLMSArg{n}{i} (\approx
 \stateFOMArg{}{\timeWindowArg{n}{i}})
-\in \RR{\fomdim}$ and $\residLMSArg{n,i}$ denotes the FOM O$\Delta$E
+\in \RR{\fomdim}$ and $\residLMSWArg{n}{i}$ denotes the FOM O$\Delta$E
 residual over the $i$th time instance of the $n$th window  
@@
 \begin{align*}
-\residArg{n,i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
+\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
                &: \RR{\fomdim} \otimes \RR{\lmsWidthArg{n}{i}+1} \rightarrow \RR{\fomdim}. 
@@
 define the specific type of multistep scheme. 
-To handle the case where the linear multistep method employs time instances from the previous window(s), we enforce the condition on the indices $(n,i) = (n-1,\nstepsArg{n-1}+i)$ for $i<0$.
 %\begin{remark}\label{remark:LMS}
@@
 %\end{align*}
-\EP{In the process of reworking this slightly, need to take care for the case where we are employing multiple time steps 
-from the previous window. Still cleaning up}
+
 \methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
@@
 In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
-\begin{equation}\label{eq:obj_lms}
+\begin{equation*}%\label{eq:obj_lms}
 \begin{split} 
-\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i};\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i};\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
+\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,1},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i}}), \\
 &\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
@@
 \end{split}
-\end{equation}
-where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights and, for $i \le 0$ in $\objectiveArgLMS{n}$, we take
+\end{equation*}
+where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights and, for $i \le 0$, we set
 \begin{equation}\label{eq:wls_spatial_direct_args}
-\stateyDiscreteArgnt{i}  \equiv 
+\stateyDiscreteArgnt{n,i}  \equiv 
 \begin{cases}
@@
 	&
-	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdim}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
+	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
 \objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
@@
 The optimization problem takes the form of a \textit{weighted least-squares
-	problem}. \KTC{I got rid of (non)linear. If it's either linear or nonlinear,
-	then omit it as an adjective, as this is no longer descriptive. }
+	problem}. 
 We emphasize that optimization problem~\eqref{eq:obj_gen_lms_final} associates
@@
 For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, uniform 
-quadrature weights are used, and uniform bases and weighting matrices are used, 
+quadrature weights are used, and uniform bases and weighting matrices are used (i.e., $\basisspaceArg{n} \equiv \basisspace, \stweightingMatArg{n} \equiv \stweightingMat, n=1,\ldots,\nslabs$), 
 \methodAcronym\ with \spatialAcronym\ trial
@@
 Discretization through linear multistep methods (as well as other techniques) 
-results in a (non)linear least-squares problem.
-A variety of algorithms exist for solving least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
-The numerical experiments presented in this work consider the Gauss--Newton method, and as such we outline this approach here.
+results in a least-squares problem. In particular, for the case where the full-order model~\eqref{eq:FOM} is nonlinear, the direct approach yields a nonlinear least-squares problem. 
+A variety of algorithms exist for solving nonlinear least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.
+The numerical experiments presented in this work comprise nonlinear systems and are solved via the Gauss--Newton method, and as such we outline this approach here.
 
@@
 \begin{equation*}
-\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{};\genstateDiscreteArg{n,0}) \mapsto \begin{bmatrix}
- \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
+\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{}) \mapsto \begin{bmatrix}
+ \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}, \ldots , \basisspaceArg{n} \genstateDiscreteArg{n,1 - \lmsWidthArg{n}{1}} + \stateInterceptArg{n}) \\
 \vdots \\
- \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
+ \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
 \end{bmatrix},
@@
 In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will an almost block-diagonal
-sparse matrix with the sparsity pattern \KTC{number of residual arguments
-inconsistent with above}
+sparse matrix with the sparsity pattern
 \begin{equation*}
@@
 This sparsity pattern can be leveraged to assemble the Jacobian of the residual over the time window from Jacobians of the residual at single time instances. 
-Further, the sparsity pattern can be 
-leveraged \KTC{split infinitive:} to efficiently compute the Jacobian
-\KTC{shoudl be an en-dash:} matrix-matrix product in the normal equations.
+Further, for solution techniques based on the normal equations, one can leverage the sparsity pattern 
+to efficiently compute the Jacobian matrix--matrix product.
 \KTC{We should never be solving the normal equations. This squares the
@@
 It is also worth noting that the 
-normal equations will also (almost) \KTC{What does ``almost'' block diagonal
-mean? This is imprecise} be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
+normal equations will be comprise a banded block system; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
 
 \begin{remark}\label{remark:gaussnewton}\textit{(Acceleration of the Gauss--Newton Solve)}\\
-	The \KTC{this should be `principal', not `principle':} principle cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
+	The principal cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
 reducing this computational burden exist; Jacobian-free Newton-Krylov
@@
 	are several such examples.
-\KTC{Expose used twice in this sentence} Further, the space--time formulation exposes an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM.
+Further, the space--time formulation introduces an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM.
 The investigation of these additional, potentially more efficient, solution algorithms will be a topic of future work.
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $\epsilon$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$; initial condition $\genstateDiscreteArg{n,0}$}
+\Input{tolerance, $\epsilon$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$}%; initial condition $\genstateDiscreteArg{n,0}$}
 \Output{Solution to least squares problem, $\genstatecollocMatSlabArg{n}$} 
@@
 %}
-$\mathbf{r} \leftarrow \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k;\genstateDiscreteArg{n,0})$ \Comment{Compute weighted residual over window} \\
+$\mathbf{r} \leftarrow \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{Compute weighted residual over window} \\
 $\mathbf{J} \leftarrow  
@@
 %Compute: $[\jacobianSlabArg{n}]^T \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{compute RHS for normal equations} \\
- Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
-$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
-\If{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \epsilon$ }{
+\uIf{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \epsilon$ }{
 {\text{converged} $\leftarrow$ \text{true}}  \Comment{Check and set convergence based on gradient norm} \\
 Return: $\genstatecollocMatSlabArg{n} = \genstatecollocMatSlabArg{n}_{k+1} $ \Comment{Return converged solution}\\
+}
+\Else{
+ Compute  $\Delta  \genstatecollocMatSlabArg{n} $ that minimizes $\norm{ \mathbf{J} \Delta \genstatecollocMatSlabArg{n} +\mathbf{r}}^2$ \Comment{Solve the linear least-squares problem} \\
+% Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
+$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
 }
@@
 backwards problems. As described in  
-the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time instances~\eqref{eq:timegrid1} over each time window.
+the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time instances over each time window.
 Linear multistep methods then use this time-grid for temporal discretization of~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}.

### 2019-10-16 section4/spatial_numerical.tex
working on 4.2.1

@@
 including collocation approaches, spectral  methods, and genetic algorithms.  
-\EP{Make sure 4.2.1 is settled on before this}
 In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
@@
 (2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
-To this end, we define a time grids
+To this end, we define time grids
 $\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$,
@@
 $\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
- = \timeEndArg{n}$.
+ = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$.
 %\KTC{use cdots, not
@@
 %\end{equation}
-Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
-\footnote{
+Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window\footnote{
 To maintain notational simplicity in this section, and handle the case where the linear multistep method employs time instances from the previous window(s), for superscripts $n=1,\ldots,\nslabs$ and $i<0$, we set the superscripts to be $n =n-1, i = \nstepsArg{n-1}+i$
 % $\stateFOMDiscreteLMSArg{n}{i} = \stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i}$, along with $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$.
-} 
+}, 
 \begin{align*}
 &\residLMSWArg{n}{i} (\stateFOMDiscreteLMSArg{n}{i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
-&\stateFOMDiscreteLMSArg{n}{0} = \begin{cases}
-\stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
-\stateFOMIC & n=1, \end{cases}
+&\stateFOMDiscreteLMSArg{n}{i} = 
+\begin{cases}
+\stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i} & n = 2,\ldots,\nslabs, \; i \le 0 \\
+\stateFOMIC & n=1, \; i=0,\end{cases}
 \end{align*}
@@
 \objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,1},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i}}), \\
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i}}), \\
 &\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
@@
 \end{equation*}
-where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights and, for $i \le 0$, we set
+where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights and, for $i \le 0$
 \begin{equation}\label{eq:wls_spatial_direct_args}
@@
 \methodAcronym\ with \spatialAcronym\ trial
-	subspaces solved via the direct approach recovers LSPG. \KTC{Not exactly. To
+	subspaces solved via the direct approach recovers the LSPG approach. \KTC{Not exactly. To
 	make this LSPG as presented earlier, also require the time grid here to
@@
 \residLMSSlabArg{n} : (\genstatecollocMatySlabArg{}) \mapsto \begin{bmatrix}
- \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}, \ldots , \basisspaceArg{n} \genstateDiscreteArg{n,1 - \lmsWidthArg{n}{1}} + \stateInterceptArg{n}) \\
+ \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; \ldots , \stateyDiscreteArgnt{n,1 - \lmsWidthArg{n}{1} } ) \\
 \vdots \\

### 2019-10-16 section4/spatial_numerical.tex
through section 4.2.1

@@
 $\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
- = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$.
+ = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$ and $n=2,\ldots,\nslabs$.
 %\KTC{use cdots, not
@@
 %\end{equation}
-Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window\footnote{
-To maintain notational simplicity in this section, and handle the case where the linear multistep method employs time instances from the previous window(s), for superscripts $n=1,\ldots,\nslabs$ and $i<0$, we set the superscripts to be $n =n-1, i = \nstepsArg{n-1}+i$
-% $\stateFOMDiscreteLMSArg{n}{i} = \stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i}$, along with $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$.
-}, 
+Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
+%\footnote{
+%To maintain notational simplicity in this section, and handle the case where the linear multistep method employs time instances from the previous window(s), for superscripts $n=1,\ldots,\nslabs$ and $i<0$, we set the superscripts to be $n =n-1, i = \nstepsArg{n-1}+i$
+%% $\stateFOMDiscreteLMSArg{n}{i} = \stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i}$, along with $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$.
+%}, 
 \begin{align*}
@@
 Here, $\Delta t^{n,i} \defeq \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$
-denotes the time step, $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window and
+denotes the time step, $\lmsWidthArg{n}{i}$ with $\lmsWidthArg{1}{1} = 1$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window  and
 $\alpha^{n,i}_j,\beta^{n,i}_j\in\RR{}$ denote coefficients that 
@@
 \begin{align*}
-	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}{i}}}{\text{minimize } }
+	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
 \objectiveArgLMS{n} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}). 
@@
 \end{equation*}
-where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights and, for $i \le 0$
+where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights, $\stateyDiscreteArgnt{n,i}$ for $i>0$ are the arguments, and for $i \le 0$
+\EP{I think this is a little unclear, could include ICs in the definition of the objective function and split the arguments with a ;  Thoughts?} 
 \begin{equation}\label{eq:wls_spatial_direct_args}
@@
 \begin{cases}
-\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
-\basisspaceArg{n} \basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 1.
+\projectorSpatialArg{n}(\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i})  & n = 2,\ldots,\nslabs,  \\
+\projectorSpatialArg{1}(\stateFOMIC)  & n = 1  \\
+%\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
+%\basisspaceArg{n} \basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 1.
 \end{cases}
 \end{equation}
+with 
+$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
 Note that the boundary conditions are automatically satisfied by Eq.~\eqref{eq:wls_spatial_direct_args}.
@@
 \begin{remark}
-For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, uniform 
-quadrature weights are used, and uniform bases and weighting matrices are used (i.e., $\basisspaceArg{n} \equiv \basisspace, \stweightingMatArg{n} \equiv \stweightingMat, n=1,\ldots,\nslabs$), 
+For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform 
+quadrature weights are used, a uniform basis is employed (i.e., $\basisspaceArg{n} = \basisspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
+$\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
+$\approxstateDiscreteArg{n,1}  = \approxstateLSPG^n$, $n=1,\ldots,\nslabs$ and  
 \methodAcronym\ with \spatialAcronym\ trial
@@
 	formulation requires allowing the stencil to go back to time steps at
-	previous windows.} \EP{I agree with the second comment, to be 100\% precise we need to go back into other time windows.
-Clearly you can do this if you would like, and I think the comment clarifies this.
-Hopefully can avoid this now with the condition you mentioned.
-I think the first comment is not required.
-The LSPG ``approach" minimizes the residual at a time instance; it isn't tied to the description we gave earlier.}
+	previous windows.}  
 \end{remark} 
@@
 \residLMSSlabArg{n} : (\genstatecollocMatySlabArg{}) \mapsto \begin{bmatrix}
- \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; \ldots , \stateyDiscreteArgnt{n,1 - \lmsWidthArg{n}{1} } ) \\
+ \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; 
+%\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
+\projectorSpatialArg{n}( \approxstateDiscreteArg{n,0}),
+\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{n,1 - \lmsWidthArg{n}{1} } )  ) \\
 \vdots \\
- \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
+ \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots,\projectorSpatialArg{n}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} )) \\
 \end{bmatrix},
@@
 where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
-It is worth noting that, by design, \KTC{This is incorrect. Missing a 1/2
-factor}
+It is worth noting that, by design, 
 \begin{equation*}
-\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
+\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
 =
-\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg].
+\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
 \end{equation*} 
@@
 	are several such examples.
-Further, the space--time formulation introduces an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM.
+Further, the space--time formulation introduces an extra dimension for parallelization that can be used to accelerate the wall-clock time of the ROM.
 The investigation of these additional, potentially more efficient, solution algorithms will be a topic of future work.

### 2019-10-16 section4/spatial_numerical.tex
through 4.2

@@
 \subsubsection{Linear multistep schemes}
-Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  
+Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ (with $\lmsWidthArg{1}{1} = 1$) denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window.
 %Discretizing the FOM ODE~\eqref{eq:FOM} with a linear $k$-step scheme at the $i$th time-instance over the $n$th time slab yields,
@@
 Here, $\Delta t^{n,i} \defeq \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$
-denotes the time step, $\lmsWidthArg{n}{i}$ with $\lmsWidthArg{1}{1} = 1$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window  and
+denotes the time step, and
 $\alpha^{n,i}_j,\beta^{n,i}_j\in\RR{}$ denote coefficients that 
@@
 \begin{equation*}
-\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{}) \mapsto \begin{bmatrix}
+\residLMSSlabArg{n} : \genstatecollocMatySlabArg{} \mapsto \begin{bmatrix}
  \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; 

### 2019-10-16 section4/spatial_numerical.tex
fix error

@@
 	coordinates
-	$\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\timeWindowArg{n}{i}}$,
+	$\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}$,
 	$n=1,\ldots,\nslabs$ 

### 2019-10-16 section4/spatial_numerical.tex
used mass in algo 2

@@
 \begin{remark}
-For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform 
+For the limiting case where $\nstepsArg{n} = 1$, $n=1,\ldots,\nslabs$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform
 quadrature weights are used, a uniform basis is employed (i.e., $\basisspaceArg{n} = \basisspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
@@
 It is also worth noting that the 
-normal equations will be comprise a banded block system; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
+normal equations will comprise a banded block system, this can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
 
@@
 guess for costate}\\ 
-$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n}
-\basisspaceArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
-\velocity(\veloargsromArg{1}) =  \controllerArg{n}{t}$ 
+$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying }  \massArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
+\velocity(\veloargsromArg{1}) =  \massArg{n} \controllerArg{n}{t}$ 
 % Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
@@
 \text{ Compute } \adjointArgnt{n} \text{ satisfying }
-\adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg]
+\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg]
 \end{multline*} }
@@
 $i \leftarrow i+1$ \Comment{Update counter}\\
-$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) =  \controllerArg{n}{t} $
+$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \massArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) = \massArg{n} \controllerArg{n}{t} $
 \Comment{Solve~\eqref{eq:lspg_continuous}}\\

### 2019-10-16 section4/spacetime_numerical.tex
latest

@@
 
+\EP{I will rework this section depending on 4.2\\
 Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
@@
 \end{remark}
-
+}
 \subsection{\spaceTimeAcronym: indirect solution approach}
@@
 discretization of the integral.
-We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ \EP{Originally this was divided by $N$, just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window
-$$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$
+Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ \EP{Originally this was divided by $N$, just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
+%with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$
 Leveraging these collocation points, the indirect method for \methodAcronym\ with an \spaceTimeAcronym\ trial subspace computes the generalized coordinates 
@@
 %$$
-The least-squares problem can again be solved, e.g., via the Gauss--Newton method. 
 %The stationary points are given by,
@@
 a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
-space--time trial subspace yields a series of dense (non)linear systems to be solved over each window.
+space--time trial subspace yields a series of dense systems to be solved over each window.

### 2019-10-16 section4/spatial_numerical.tex
latest

@@
 To this end, we define time grids
-$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$,
+$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$ \EP{Shouldn't this be subset? We can't recover the continuous time interval},
 $n=1,\ldots,\nslabs$ that

### 2019-10-16 section4/spatial_numerical.tex
minor changes

@@
 For the limiting case where $\nstepsArg{n} = 1$, $n=1,\ldots,\nslabs$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform
-quadrature weights are used, a uniform basis is employed (i.e., $\basisspaceArg{n} = \basisspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
+quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
 $\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 

### 2019-10-17 section3/space_time.tex
changed notation in 4.2

@@
 \end{equation}
+\EP{Is it wrong to say $\stbasisArg{n}{\cdot} \in \RR{\fomdim \times \stdimArg{n}}$ with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy) \ldots$? I don't particularly like
+$\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}}\otimes\timeSpaceArg{n}$; as the reader I think that $\stbasisArgnt{n}$ can take on all temporal functions.}
 Here $\stbasisArgnt{n} \in \RR{\fomdim \times

### 2019-10-17 section4/numerical_methods.tex
changed notation in 4.2

@@
 We focus primarily on 
-direct (i.e., discretize-the-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
+direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 

### 2019-10-17 section4/spatial_numerical.tex
changed notation in 4.2

@@
 $\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
- = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$ and $n=2,\ldots,\nslabs$.
-%\KTC{use cdots, not
+ = \timeEndArg{n}$. %\KTC{use cdots, not
 %hdots or ldots. Don't use $\defeq$ for the first and last times; these are
@@
 Figure~\ref{fig:slab_fig2} depicts such a discretization.
+For the purposes of indexing between different windows, we additionally define a function:
+$$\indexMapper: (n,i) \mapsto 
+\begin{cases}
+1,0 & n=1, \; i = 0,\\
+n,i & n \ge 1, \; i > 0, \\
+\indexMapper(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
+\end{cases}$$
 We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
@@
 \subsubsection{Linear multistep schemes}
-Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ (with $\lmsWidthArg{1}{1} = 1$) denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window.
+Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window.
 %Discretizing the FOM ODE~\eqref{eq:FOM} with a linear $k$-step scheme at the $i$th time-instance over the $n$th time slab yields,
@@
 \begin{align*}
-&\residLMSWArg{n}{i} (\stateFOMDiscreteLMSArg{n}{i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
-&\stateFOMDiscreteLMSArg{n}{i} = 
-\begin{cases}
-\stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i} & n = 2,\ldots,\nslabs, \; i \le 0 \\
-\stateFOMIC & n=1, \; i=0,\end{cases}
+&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
+&\stateFOMDiscreteArg{1,0} = \stateFOMIC. 
+%\begin{cases}
+%\stateFOMDiscreteLMSArg{n-1}{0} & n = 2,\ldots,\nslabs \\
+%\stateFOMIC & n=1, \end{cases}
 \end{align*}
-where $\stateFOMDiscreteLMSArg{n}{i} (\approx
+where $\stateFOMDiscreteArg{n,i} (\approx
 \stateFOMArg{}{\timeWindowArg{n}{i}})
@@
 \begin{align*}
-\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
+\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindow^{\indexMapper(n,i-j)}) \\
                &: \RR{\fomdim} \otimes \RR{\lmsWidthArg{n}{i}+1} \rightarrow \RR{\fomdim}. 
@@
 $\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
-$n=1,\ldots,\nslabs$ that satisfy
+$n=1,\ldots,\nslabs$, that satisfy
 %	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
@@
 \begin{cases}
-\projectorSpatialArg{n}(\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i})  & n = 2,\ldots,\nslabs,  \\
-\projectorSpatialArg{1}(\stateFOMIC)  & n = 1  \\
+\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,i)})  & n = 2,\ldots,\nslabs,  \\
+\projectorSpatialArg{1}(\stateFOMIC)  & n = 1, \\
 %\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
@@
 \end{equation}
-with 
+where $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector and is defined as 
 $$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
@@
 with $\genstateDiscreteArg{n,i}(\approx
-	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdim}$
+	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdimArg{n}}$
 	that satisfy
@@
 \methodAcronym\ with \spatialAcronym\ trial
-	subspaces solved via the direct approach recovers the LSPG approach. \KTC{Not exactly. To
-	make this LSPG as presented earlier, also require the time grid here to
-	align with the time grid used for time integration earlier. Also, this
-	formulation requires allowing the stencil to go back to time steps at
-	previous windows.}  
+	subspaces solved via the direct approach recovers the LSPG approach. 
 \end{remark} 
@@
 %\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
-\projectorSpatialArg{n}( \approxstateDiscreteArg{n,0}),
-\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{n,1 - \lmsWidthArg{n}{1} } )  ) \\
+\projectorSpatialArg{n}( \approxstateDiscreteArg{\indexMapper(n,0)}),
+\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{\indexMapper(n,1 - \lmsWidthArg{n}{1}) } )  ) \\
 \vdots \\
@@
 \end{equation*}
-where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
+where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} )$ and, for $i \le 0$,
+\begin{equation}\label{eq:wls_genspatial_direct_args}
+\genstateyDiscreteArgnt{n,i}  \equiv 
+\begin{cases}
+\genstateDiscreteArg{\indexMapper(n,i)}  & n = 2,\ldots,\nslabs,  \\
+\genstateICOne  & n = 1. 
+\end{cases}
+\end{equation}
 It is worth noting that, by design, 

### 2019-10-17 section4/spacetime_numerical.tex
section 4.5

@@
 \subsection{\spaceTimeAcronym: direct solution approach}
-The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods.
-
+The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E.
+For brevity, we again focus on linear multistep methods.
+For notational simplicity, we define an index mapping function that is
+equivalent to $\indexMapper$, but outputs only the first argument: 
+$$\indexMappern: (n,i) \mapsto 
+\begin{cases}
+1 & n=1, \; i = 0,\\
+n & n \ge 1, \; i > 0, \\
+\indexMappern(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
+\end{cases}$$
 %The definition of the fully discrete objective functions~\eqref{eq:obj_lms} as outlined in Section~\ref{sec:direct}, can be leveraged for this purpose. For simplicity, we only 
 %outline the case for linear multistep schemes. 
-
-\EP{I will rework this section depending on 4.2\\
-Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
+Assuming $\text{Rank}(\stweightingMatOneArg{n})\nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging an \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
  \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 & \underset{\stgenstatey}{\text{minimize } }
-\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}).%  \\
+\objectiveArgLMS{n}_{\text{D-ST}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}),%  \\
 %&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
@@
 \end{equation}
+where the space--time objective function for the direct method is
+$$
+\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots,
+ \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}})
+$$
+It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
+The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
+\begin{remark}
+For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1} = \lspgWeightingST = \mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
+\end{remark}
+
+\begin{remark}
+To enable equivalence in the case for general $\lspgWeightingSTArg{\cdot}$, the weighting matrix $\stweightingMatArg{1}$ needs to be a time-dependent matrix function such that the objective function~\eqref{eq:obj_gen_lms_final_st} permits a modified space--time norm.
+This is not pursued here.
+\end{remark}
+
+%\EP{I will rework this section depending on 4.2\\
+%Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
 % \begin{equation}\label{eq:obj_gen_lms_final_st}
 %\begin{split}
-%&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
-%\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
-%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
-%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
-%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
-%n=1. \end{cases}\end{split} 
+%& \underset{\stgenstatey}{\text{minimize } }
+%\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}).%  \\
+%%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
+%%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%%n=1. \end{cases}
+%\end{split} 
 %\end{equation}
-It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
+%% \begin{equation}\label{eq:obj_gen_lms_final_st}
+%%\begin{split}
+%%&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
+%%\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
+%%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
+%%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
+%%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
+%%n=1. \end{cases}\end{split} 
+%%\end{equation}
+%It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
 The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss--Newton method.
-
-\begin{remark}
-For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and uniform quadrature weights are used, and the weighting matrix $\stweightingMatArg{n} = \mathbf{I}$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG with $\lspgWeightingSTArg{\cdot} = \mathbf{I}$.
+%
+%\begin{remark}
+%For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and uniform quadrature weights are used, and the weighting matrix $\stweightingMatArg{n} = \mathbf{I}$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG with $\lspgWeightingSTArg{\cdot} = \mathbf{I}$.
 For equivalence conditions in the case where $\lspgWeightingSTArg{\cdot} \ne \mathbf{I}$, the objective function~\eqref{eq:obj_gen_lms_final_st} can be modified to enable weighted space--time norm.
-\end{remark}
-}
+%\end{remark}
+%}
 \subsection{\spaceTimeAcronym: indirect solution approach}
@@
 discretization of the integral.
-Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ \EP{Originally this was divided by $N$, just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
+Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ collocation points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
 %with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$

### 2019-10-17 section4/spatial_numerical.tex
section 4.5

@@
 %\end{align*} 
+Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
+%In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
+\begin{equation*}%\label{eq:obj_lms}
+\begin{split} 
+\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{-\lmsWidthArg{n}{1}+1}) \mapsto 
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
+&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
+\RR{}_+, 
+\end{split}
+\end{equation*}
+where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights.
 
@@
 	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}). 
+\objectiveArgLMS{n}_s (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
 %%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
@@
 \end{align*}
-%Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
-%\EP{Had to change the definition of the objective function to reflect stepping back into previous window}
-In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
-\begin{equation*}%\label{eq:obj_lms}
-\begin{split} 
-\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,1},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i}}), \\
-&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
-\RR{}_+, 
-\end{split}
-\end{equation*}
-where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights, $\stateyDiscreteArgnt{n,i}$ for $i>0$ are the arguments, and for $i \le 0$
-\EP{I think this is a little unclear, could include ICs in the definition of the objective function and split the arguments with a ;  Thoughts?} 
-\begin{equation}\label{eq:wls_spatial_direct_args}
-\stateyDiscreteArgnt{n,i}  \equiv 
-\begin{cases}
-\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,i)})  & n = 2,\ldots,\nslabs,  \\
-\projectorSpatialArg{1}(\stateFOMIC)  & n = 1, \\
-%\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
-%\basisspaceArg{n} \basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 1.
-\end{cases}
-\end{equation}
-where $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector and is defined as 
+where
+$$\objectiveArgLMS{n}_{\text{s}} :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}+1)})),
+$$
+and $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
 $$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
-Note that the boundary conditions are automatically satisfied by Eq.~\eqref{eq:wls_spatial_direct_args}.
-
+We additionally set $\approxstateDiscreteArg{1,0} = \projectorSpatialArg{n}(\stateFOMIC).
+$
 %where 
@@
 %\approxstateIC & n=1. \end{cases} \end{align*}
+
 Equivalently, \methodAcronym\ sequentially computes the generalized
@@
 	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
+\objectiveArgLMS{n}_s (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
 %& \text{subject to} \hspace{0.2 in}
@@
 \begin{equation*}
-\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
+\objectiveArgLMS{n}_s \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
 =

### 2019-10-17 section4/spacetime_numerical.tex
working on space--time section

@@
 We associate direct methods as those
-that work with the O$\Delta$E and indirect methods with those that work directly with the continuous ODE.
+that work with the O$\Delta$E and indirect methods with those that work directly with the ODE.
 
@@
 For notational simplicity, we define an index mapping function that is
-equivalent to $\indexMapper$, but outputs only the first argument: 
+equivalent to the mapping function $\indexMapper$, but outputs only the first argument: 
 $$\indexMappern: (n,i) \mapsto 
@@
 $$
-\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots,
- \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}})
+\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots,
+ \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
 $$
@@
 \begin{remark}
-For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1} = \lspgWeightingST = \mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
+For the limiting case where one window comprises entire domain (i.e., $\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1} = \lspgWeightingST = \mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
 \end{remark}

### 2019-10-17 section4/spatial_numerical.tex
working on space--time section

@@
 \begin{split} 
-\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{-\lmsWidthArg{n}{1}+1}) \mapsto 
+\objectiveArgLMS{n}_{\text{D}} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{-\lmsWidthArg{n}{1}+1}) \mapsto 
 \frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
-&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
+&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n} + \lmsWidthArg{n}{1}} \rightarrow
 \RR{}_+, 
@@
 	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_s (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
+\objectiveArgLMS{n}_{\text{D-S}} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
 %%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
@@
 where
-$$\objectiveArgLMS{n}_{\text{s}} :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}+1)})),
+$$\objectiveArgLMS{n}_{\text{D-S}} :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}+1)})),
 $$
@@
 	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_s (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
+\objectiveArgLMS{n}_{\text{D-S}} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
 %& \text{subject to} \hspace{0.2 in}
@@
 \begin{equation*}
-\objectiveArgLMS{n}_s \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
+\objectiveArgLMS{n}_{\text{D-S}} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg)
 =

### 2019-10-17 section4/spatial_numerical.tex
sml

--- 2019-10-17 section4/spatial_numerical.tex
@@
 including collocation approaches, spectral  methods, and genetic algorithms.
+
 In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
@@
 (2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
-To this end, we define time grids
-$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$ \EP{Shouldn't this be subset? We can't recover the continuous time interval},
+To this end, we define a time grids
+$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$,
 $n=1,\ldots,\nslabs$ that
@@
 $\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
- = \timeEndArg{n}$. %\KTC{use cdots, not
+ = \timeEndArg{n}$.
+%\KTC{use cdots, not
 %hdots or ldots. Don't use $\defeq$ for the first and last times; these are
@@
 Figure~\ref{fig:slab_fig2} depicts such a discretization.
-For the purposes of indexing between different windows, we additionally define a function:
-$$\indexMapper: (n,i) \mapsto 
-\begin{cases}
-1,0 & n=1, \; i = 0,\\
-n,i & n \ge 1, \; i > 0, \\
-\indexMapper(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
-\end{cases}$$
 We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
@@
 \subsubsection{Linear multistep schemes}
-Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window.
+Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  
 %Discretizing the FOM ODE~\eqref{eq:FOM} with a linear $k$-step scheme at the $i$th time-instance over the $n$th time slab yields,
@@
 Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
-%\footnote{
-%To maintain notational simplicity in this section, and handle the case where the linear multistep method employs time instances from the previous window(s), for superscripts $n=1,\ldots,\nslabs$ and $i<0$, we set the superscripts to be $n =n-1, i = \nstepsArg{n-1}+i$
-%% $\stateFOMDiscreteLMSArg{n}{i} = \stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i}$, along with $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$.
-%}, 
+\KTC{Note: the notation here is not consistent with what was defined earlier
+for a discrete residual. See email.}
 \begin{align*}
-&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
-&\stateFOMDiscreteArg{1,0} = \stateFOMIC. 
-%\begin{cases}
-%\stateFOMDiscreteLMSArg{n-1}{0} & n = 2,\ldots,\nslabs \\
-%\stateFOMIC & n=1, \end{cases}
+&\residLMSArg{n,i} (\stateFOMDiscreteArg{n,i},\ldots,\stateFOMDiscreteArg{n,i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
+&\stateFOMDiscreteArg{n,0} = \begin{cases}
+\stateFOMDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs, \\
+\stateFOMIC & n=1, \end{cases}
 \end{align*}
@@
 \stateFOMArg{}{\timeWindowArg{n}{i}})
-\in \RR{\fomdim}$ and $\residLMSWArg{n}{i}$ denotes the FOM O$\Delta$E
+\in \RR{\fomdim}$ and $\residLMSArg{n,i}$ denotes the FOM O$\Delta$E
 residual over the $i$th time instance of the $n$th window 
@@
 \begin{align*}
-\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindow^{\indexMapper(n,i-j)}) \\
+\residArg{n,i} &: (\stateyDiscreteArgnt{i},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\
                &: \RR{\fomdim} \otimes \RR{\lmsWidthArg{n}{i}+1} \rightarrow \RR{\fomdim}. 
@@
 Here, $\Delta t^{n,i} \defeq \timeWindowArg{n}{i} - \timeWindowArg{n}{i-1}$
-denotes the time step, and
+denotes the time step, $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window, and
 $\alpha^{n,i}_j,\beta^{n,i}_j\in\RR{}$ denote coefficients that 
 define the specific type of multistep scheme. 
-%\begin{remark}\label{remark:LMS}
-%For notational simplicity, we assume the linear multistep method employs time
-%	instances from only within the current time window. In general, the linear
-%	multistep method at the $n$th window could be designed to employ multiple
-%	time instances from previous time windows. \KTC{What are the implications of
-%	this? I suggest removing this and adding a simple condition, maybe with a
-%	footnote.}
-%\end{remark}
+\begin{remark}\label{remark:LMS}
+For notational simplicity, we assume the linear multistep method employs time
+	instances from only within the current time window. In general, the linear
+	multistep method at the $n$th window could be designed to employ multiple
+	time instances from previous time windows. \KTC{What are the implications of
+	this? I suggest removing this and adding a simple condition, maybe with a
+	footnote.}
+\end{remark}
 %Employing a linear multistep method for temporal discretization of the FOM ODE leads to the FOM O$\Delta$E over the $n$th slab,
@@
 Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
-%In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
-\begin{equation*}%\label{eq:obj_lms}
+\begin{equation}\label{eq:obj_lms}
 \begin{split} 
-\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{-\lmsWidthArg{n}{1}+1}) \mapsto 
-\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
-&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
+\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto
+\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residArg{n,i}(\stateyDiscreteArgnt{i},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
+&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}+1} \rightarrow
 \RR{}_+, 
 \end{split}
-\end{equation*}
-where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights.
-
-\methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
-$\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
-$n=1,\ldots,\nslabs$, that satisfy
-%	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
-%	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
-%	this should be the discrete solution, not the exact solution evaluated at
-%	the final time.} \EP{You are correct, I think there is an error now in the minimize term though.
-It was $\in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}$, it should be $\trialspace\otimes \RR{\nstepsArg{n}+1}$}
-\begin{align*}
-	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_s (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
-%%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
-%%\begin{cases} \basisspaceArg{n} \spatialICDiscrete + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
-%\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} 
-\end{align*}
-where
-$$\objectiveArgLMS{n}_{\text{s}} :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}+1)})),
-$$
-and $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
-$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
-We additionally set $\approxstateDiscreteArg{1,0} = \projectorSpatialArg{n}(\stateFOMIC).
-$
+\end{equation}
 %where 
@@
 %is the discrete residual resulting from the linear multistep scheme and 
+where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights.
+\KTC{Don't these weights need to be non-negative to guarantee the integral is
+non-negative? If so, then specify $\quadWeightsLMSScalarArg{n}{i} \in\RR{}_+$}
+\methodAcronym\ sequentially computes the solutions
+$\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
+$n=1,\ldots,\nslabs$ that satisfy
+	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
+	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
+	this should be the discrete solution, not the exact solution evaluated at
+	the final time.}
+\begin{align*}
+	&\underset{(\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}}{\text{minimize } }
+\objectiveArgLMS{n} (\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \\
+& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
+\begin{cases} \basisspaceArg{n} \spatialIC + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
+\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} \end{align*}
 %\methodAcronym\ solves the minimization problem over each time window, 
@@
 %\approxstateIC & n=1. \end{cases} \end{align*}
-
-Equivalently, \methodAcronym\ sequentially computes the generalized
+	Equivalently, WLS sequentially computes the generalized
 	coordinates
-	$\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}$,
+	$\genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\timeWindowArg{n}{i}}$,
 	$n=1,\ldots,\nslabs$ 
 with $\genstateDiscreteArg{n,i}(\approx
-	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdimArg{n}}$
+	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdim}$
 	that satisfy
@@
 	&
-	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_s (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
-%& \text{subject to} \hspace{0.2 in}
-%\genstateyDiscreteArg{0} =
-%\begin{cases} \genspatialICDiscrete  & n = 2,\ldots,\nslabs \\
-%\genstateICOne& n=1. \end{cases} 
+	\underset{(\genstateyDiscreteArg{0},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdim}\otimes \RR{\lmsWidthArg{n}{i}+1}}{\text{minimize } }
+\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}) \\ 
+& \text{subject to} \hspace{0.2 in}
+\genstateyDiscreteArg{0} =
+\begin{cases} \spatialIC & n = 2,\ldots,\nslabs \\
+\genstateICOne& n=1. \end{cases} 
 \end{split}
@@
 The optimization problem takes the form of a \textit{weighted least-squares
-	problem}. 
+	problem}. \KTC{I got rid of (non)linear. If it's either linear or nonlinear,
+	then omit it as an adjective, as this is no longer descriptive. }
 We emphasize that optimization problem~\eqref{eq:obj_gen_lms_final} associates
@@
 \begin{remark}
-For the limiting case where $\nstepsArg{n} = 1$, $n=1,\ldots,\nslabs$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform
-quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
-$\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
-$\approxstateDiscreteArg{n,1}  = \approxstateLSPG^n$, $n=1,\ldots,\nslabs$ and  
-\methodAcronym\ with \spatialAcronym\ trial
-	subspaces solved via the direct approach recovers the LSPG approach. 
+For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
+quadrature weights are used, \methodAcronym\ with \spatialAcronym\ trial
+	subspaces solved via the direct approach recovers LSPG. \KTC{Not exactly. To
+	make this LSPG as presented earlier, also require the time grid here to
+	align with the time grid used for time integration earlier. Also, this
+	formulation requires allowing the stencil to go back to time steps at
+	previous windows.} 
 \end{remark} 
 \subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
+	\KTC{I recommend rewriting this section with several important things in
+	mind. We can discuss.}
 	\KTC{If the dynamics are linear, doesn't this just yield a linear
@@
 Discretization through linear multistep methods (as well as other techniques) 
-results in a least-squares problem. In particular, for the case where the full-order model~\eqref{eq:FOM} is nonlinear, the direct approach yields a nonlinear least-squares problem. 
-A variety of algorithms exist for solving nonlinear least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.
-The numerical experiments presented in this work comprise nonlinear systems and are solved via the Gauss--Newton method, and as such we outline this approach here.
+results in a (non)linear least-squares problem.
+A variety of algorithms exist for solving least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
+The numerical experiments presented in this work consider the Gauss--Newton method, and as such we outline this approach here.
 
@@
 defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
-\EP{The notation for space--time in 2.4 uses bars. I don't want to confuse these generalized coordinates with those, they conceptually two different things} 
 \begin{equation*}
@@
 \begin{equation*}
-\residLMSSlabArg{n} : \genstatecollocMatySlabArg{} \mapsto \begin{bmatrix}
- \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; 
-%\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
-\projectorSpatialArg{n}( \approxstateDiscreteArg{\indexMapper(n,0)}),
-\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{\indexMapper(n,1 - \lmsWidthArg{n}{1}) } )  ) \\
+\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{};\genstateDiscreteArg{n,0}) \mapsto \begin{bmatrix}
+ \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
 \vdots \\
- \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots,\projectorSpatialArg{n}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} )) \\
+ \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
 \end{bmatrix},
 \end{equation*}
-where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} )$ and, for $i \le 0$,
-\begin{equation}\label{eq:wls_genspatial_direct_args}
-\genstateyDiscreteArgnt{n,i}  \equiv 
-\begin{cases}
-\genstateDiscreteArg{\indexMapper(n,i)}  & n = 2,\ldots,\nslabs,  \\
-\genstateICOne  & n = 1. 
-\end{cases}
-\end{equation}
-It is worth noting that, by design, 
+where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
+It is worth noting that, by design, \KTC{This is incorrect. Missing a 1/2
+factor}
 \begin{equation*}
-\objectiveArgLMS{n}_s \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
+\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
 =
-\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
+\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg].
 \end{equation*} 
@@
 This sparsity pattern can be leveraged to assemble the Jacobian of the residual over the time window from Jacobians of the residual at single time instances. 
-Further, for solution techniques based on the normal equations, one can leverage the sparsity pattern 
-to efficiently compute the Jacobian matrix--matrix product.
+Further, the sparsity pattern can be 
+leveraged \KTC{split infinitive:} to efficiently compute the Jacobian
+\KTC{shoudl be an en-dash:} matrix-matrix product in the normal equations.
 \KTC{We should never be solving the normal equations. This squares the
@@
 It is also worth noting that the 
-normal equations will comprise a banded block system, this can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
+normal equations will also (almost) \KTC{What does ``almost'' block diagonal
+mean? This is imprecise} be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
 
 \begin{remark}\label{remark:gaussnewton}\textit{(Acceleration of the Gauss--Newton Solve)}\\
-	The principal cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
+	The \KTC{this should be `principal', not `principle':} principle cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
 reducing this computational burden exist; Jacobian-free Newton-Krylov
@@
 	are several such examples.
-Further, the space--time formulation introduces an extra dimension for parallelization that can be used to accelerate the wall-clock time of the ROM.
+\KTC{Expose used twice in this sentence} Further, the space--time formulation exposes an extra dimension for parallelization that can be exposed to accelerate the wall-clock time of the ROM.
 The investigation of these additional, potentially more efficient, solution algorithms will be a topic of future work.
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $\epsilon$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$}%; initial condition $\genstateDiscreteArg{n,0}$}
+\Input{tolerance, $\epsilon$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$; initial condition $\genstateDiscreteArg{n,0}$}
 \Output{Solution to least squares problem, $\genstatecollocMatSlabArg{n}$} 
@@
 $\text{converged} \leftarrow \text{false}$ \Comment{Set convergence checker} \\
+$\genstatecollocMatSlabArg{n}_0 \leftarrow \unroll(\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
 $k \leftarrow 0$ \Comment{Set counter}\\
-$\genstatecollocMatSlabArg{n}_k \leftarrow \unroll(\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
-\While{\text{converged} == \text{false}}
+\While{\text{not converged}}
 {
@@
 %Compute: $[\jacobianSlabArg{n}]^T \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{compute RHS for normal equations} \\
-\uIf{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \epsilon$ }{
+ Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
+$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
+\If{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \epsilon$ }{
 {\text{converged} $\leftarrow$ \text{true}}  \Comment{Check and set convergence based on gradient norm} \\
 Return: $\genstatecollocMatSlabArg{n} = \genstatecollocMatSlabArg{n}_{k+1} $ \Comment{Return converged solution}\\
-}
-\Else{
- Compute  $\Delta  \genstatecollocMatSlabArg{n} $ that minimizes $\norm{ \mathbf{J} \Delta \genstatecollocMatSlabArg{n} +\mathbf{r}}^2$ \Comment{Solve the linear least-squares problem} \\
-% Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
-$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
 }
@@
 \subsection{\spatialAcronym\ trial subspaces: indirect solution approach}\label{sec:indirect}
+\KTC{This is a fantastic insight and a really great contribution!}
 In contrast to the direct approach,
 indirect methods ``indirectly" solve the minimization
-problem~\eqref{eq:tclsrm} by solving the Euler--Lagrange equations. The
-system given by the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_adjoint} comprises a coupled two-point boundary value
-problem. A variety of techniques have
-been devised to solve two-point boundary value problems of this type. These
-techniques include shooting methods, multiple shooting
+problem~\eqref{eq:tclsrm} by solving the Euler--Lagrange
+equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_adjoint} associated with
+stationarity. The
+system associated with the Euler--Lagrange equations comprises a coupled two-point boundary value
+problem. Several techniques have
+been devised to solve such problems, 
+including shooting methods, multiple shooting
 methods~\cite{multiple_shooting}, and the forward--backward sweep
@@
 guess for costate}\\ 
-$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying }  \massArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
-\velocity(\veloargsromArg{1}) =  \massArg{n} \controllerArg{n}{t}$ 
+$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n}
+\basisspaceArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
+\velocity(\veloargsromArg{1}) =  \controllerArg{n}{t}$ 
 % Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
@@
 \text{ Compute } \adjointArgnt{n} \text{ satisfying }
-\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg]
+\adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg]
 \end{multline*} }
@@
 $i \leftarrow i+1$ \Comment{Update counter}\\
-$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \massArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) = \massArg{n} \controllerArg{n}{t} $
+$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) =  \controllerArg{n}{t} $
 \Comment{Solve~\eqref{eq:lspg_continuous}}\\
@@
 backwards problems. As described in  
-the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time instances over each time window.
+the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time instances~\eqref{eq:timegrid1} over each time window.
 Linear multistep methods then use this time-grid for temporal discretization of~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}.

### 2019-10-17 main.tex

@@
 \label{alg:pod}
-Input: Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$ \; 
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$ \; 
 %\newline
-Output: POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
+\textbf{Output:} POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
 %\newline
@@
 \label{alg:qdeim}
-Input: Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$ \; 
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$ \; 
 %\newline
-Output: Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneTArg{} \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
+\textbf{Output:} Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneTArg{} \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
 %\newline

### 2019-10-17 section4/spatial_numerical.tex

--- 2019-10-17 section4/spatial_numerical.tex
@@
 \subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
+	\KTC{I recommend rewriting this section with several important things in
+	mind. We can discuss.}
 	\KTC{If the dynamics are linear, doesn't this just yield a linear

### 2019-10-17 main.tex
minor changes

@@
 Because Eq.~\eqref{eq:g_truncation} is overdetermined, a solution may not
-exist. Typically, either the \textit{Galerkin} or \textit{least-squares
-Petrov--Galerkin} approach is employed to reduce the number of equations
+exist. Typically, either \textit{Galerkin} or \textit{least-squares
+Petrov--Galerkin} projection is employed to reduce the number of equations
 such that a unique solution exists. We now describe these two methods.
@@
 \begin{equation*}
-\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,\ldots,\ntimeSteps,
+\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace + \stateIntercept }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,\ldots,\ntimeSteps,
 \end{equation*}

### 2019-10-17 section3/general_form.tex
changed notation in 4.2
minor changes

@@
 \end{equation}
-where  $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
+where  $\timeSpaceArg{n}$ denotes the set of (sufficiently smooth) real-valued functions acting on
 $[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} =
@@
 	\stspaceBoundStartArg{n}\defeq\mathrm{span}(\{\statey(\timeStartArg{n})\,|\, \statey\in
-	\stspaceArg{n}\})\subseteq\RR{\fomdim}, \qquad n=1,\ldots,\nslabs. 
+	\stspaceArg{n}\})\subseteq\RR{\fomdim}, \qquad n=1,\ldots,\nslabs, 
 \end{equation}
@@
 - \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
-\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
+\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t, \\
 %&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
@@
 			&\text{subject to } \;  \statey(\timeStartArg{n}) =
-\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs \\
+\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs, \\
 \mathbb{P}^n( \stateFOMIC)& n=1, \end{cases} 

### 2019-10-17 section3/space_time.tex
minor changes

@@
 Here $\stbasisArgnt{n} \in \RR{\fomdim \times
-\stdimArg{n}}\otimes\timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix {function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation.
+\stdimArg{n}}\otimes\timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix {function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$, $n=1,\ldots,\nslabs$ provides the affine transformation.
 To enforce the initial condition and ensure solution continuity across time windows, we set $\stateInterceptSTArg{1} = \stateFOMIC$ and
@@
 	minimizing a functional with respect to a (time-dependent) function, but now is minimizing a function with 
-respect to vector.  As such, the first-order optimality conditions can be
+respect to a vector.  As such, the first-order optimality conditions can be
 	derived using standard calculus. Differentiating the objective with respect

### 2019-10-17 section4/numerical_methods.tex
minor changes

--- 2019-10-16 section4/numerical_methods.tex
@@
 We focus primarily on 
-direct (i.e., discretize-the-optimize) and indirect (i.e., optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
+direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize) methods for \spatialAcronym\ trial subspaces, 
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 

### 2019-10-17 section4/spatial_numerical.tex
minor changes
latest

@@
 \subsubsection{Linear multistep schemes}
-Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window.
+Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance of the $n$th window.
 %Discretizing the FOM ODE~\eqref{eq:FOM} with a linear $k$-step scheme at the $i$th time-instance over the $n$th time slab yields,
@@
 \begin{align*}
-&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
-&\stateFOMDiscreteArg{1,0} = \stateFOMIC. 
+&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, 
 %\begin{cases}
@@
 \end{align*}
-where $\stateFOMDiscreteArg{n,i} (\approx
+along with the initial condition $\stateFOMDiscreteArg{1,0} = \stateFOMIC$.
+In the above, $\stateFOMDiscreteArg{n,i} (\approx
 \stateFOMArg{}{\timeWindowArg{n}{i}})
@@
 \begin{align*}
-\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindow^{\indexMapper(n,i-j)}) \\
+\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindow^{\indexMapper(n,i-j)}), \\
                &: \RR{\fomdim} \otimes \RR{\lmsWidthArg{n}{i}+1} \rightarrow \RR{\fomdim}. 
@@
 $\alpha^{n,i}_j,\beta^{n,i}_j\in\RR{}$ denote coefficients that 
-define the specific type of multistep scheme. 
+define the specific type of multistep scheme at the $i$th time instance of the $n$th window. 
 %\begin{remark}\label{remark:LMS}
@@
 \begin{align*}
-	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
+	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in ( \trialspace + \stateInterceptArg{n}) \otimes \RR{\nstepsArg{n}}}{\text{minimize } }
 \objectiveArgLMS{n}_{\text{D-S}} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
@@
 where
-$$\objectiveArgLMS{n}_{\text{D-S}} :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}+1)})),
-$$
+\begin{equation*}
+\begin{split}
+\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)})), \\
+&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow \RR{}_+,
+\end{split}
+\end{equation*}
 and $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
@@
 
-Equivalently, \methodAcronym\ sequentially computes the generalized
+Equivalently, \methodAcronym\ with the direct approach and a linear multistep method sequentially computes the generalized
 	coordinates
@@
 For the limiting case where $\nstepsArg{n} = 1$, $n=1,\ldots,\nslabs$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform
-quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
+quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace$ and $\stateInterceptArg{n} = \stateInterceptArg{}, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
 $\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
@@
 Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method.
-The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and form the normal equations, and (3) solve the normal equations and update the state.
+The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and (3) solve the linear least-squares problem and update the guess.
 
 The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the Jacobian of the residual over the time window.
-In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will an almost block-diagonal
+In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will yield an almost block-diagonal
 sparse matrix with the sparsity pattern
@@
 to efficiently compute the Jacobian matrix--matrix product.
-\KTC{We should never be solving the normal equations. This squares the
-condition number!}
 It is also worth noting that the 

### 2019-10-17 section4/spacetime_numerical.tex
latest
small changes

@@
 dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
-dimensional representation of the objective function~\eqref{eq:obj_gen_slab}.
+dimensional representation of the objective functional~\eqref{eq:obj_gen_slab}.
 In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE.
@@
 The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E.
-For brevity, we again focus on linear multistep methods.
+For brevity, we again focus on linear multistep methods and leverage the time grids introduced in Section~\ref{sec:direct}.
 For notational simplicity, we define an index mapping function that is
@@
 %outline the case for linear multistep schemes. 
-Assuming $\text{Rank}(\stweightingMatOneArg{n})\nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging an \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
+Assuming $\text{Rank}(\stweightingMatOneArg{n})\nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ with the direct approach and an \spaceTimeAcronym\ trial subspace sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
  \begin{equation}\label{eq:obj_gen_lms_final_st}
 \begin{split}
-& \underset{\stgenstatey}{\text{minimize } }
+& \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize } }
 \objectiveArgLMS{n}_{\text{D-ST}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}),%  \\
@@
 \begin{split}
- & \underset{\stgenstatey}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} \bigg) .%\\  
+ & \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} \bigg) .%\\  
 %&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
@@
 discretization of the integral.
-Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ collocation points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
+Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
 %with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$
-Leveraging these collocation points, the indirect method for \methodAcronym\ with an \spaceTimeAcronym\ trial subspace computes the generalized coordinates 
+Leveraging these quadrature points, \methodAcronym\ with the indirect method and an \spaceTimeAcronym\ trial subspace computes the generalized coordinates 
 $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
@@
 where the discrete objective function is given by
-$$\objectiveDiscreteSTArg{n}: \statey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
+$$\objectiveDiscreteSTArg{n}: \statey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \zeta^{n,i} 
 \bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg]^T 
 \stweightingMatArg{n} 
-\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg].
+\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg],
 $$
+where $\zeta^{n,i} \in \RRplus$, $i=1,\ldots,\ncollocSTArg{n}$ are quadrature weights over the $n$th window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.
 %$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 

### 2019-10-17 section4/spatial_numerical.tex
small changes

@@
 convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
-contains three parameters: the damping factor $\rho \le 1$, the growth factor
-$\fbsmGrowth \ge 1$, and the decay factor $\fbsmDecay \ge 1$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
-closer $\rho$ is to unity, the faster the resulting algorithm will converge.
+contains three parameters: the relaxation factor $\rho \le 1$, the growth factor
+$\fbsmGrowth \ge 1$, and the decay factor $\fbsmDecay \ge 1$. The relaxation factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
+closer $\rho$ is to unity, the faster the algorithm will converge.
 For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
 A proper value of $\rho$ can be
-obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the damping factor
+obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the relaxation factor
 according to the objective. Convergence properties of the FBSM method are
@@
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
-\Input{tolerance, $\epsilon$; damping factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
+\Input{tolerance, $\epsilon$; relaxation factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
 factor, $\fbsmDecay \ge 1$} 
@@
 {
-$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
+$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the relaxation factor}\\
 }
 \Else{
-$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
+$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the relaxation factor}\\ 
 $\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
@@
 The second ingredient, namely an efficient method for computing the action of the Jacobian transpose on a vector, can be challenging. 
-In the case one does not have access to the full-order model Jacobian (or it is too costly to compute), the evaluation of this term can be both challenging and expensive.
+In the case one does not have access to the full-order model Jacobian (or it is too costly to compute), the evaluation of this term can be challenging and expensive.
 Here, two methods are discussed that can be used to evaluate the action of the Jacobian transpose on a vector as seen in~\eqref{eq:lspg_adjoint}:
@@
 $$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$
-The Jacobian transpose can be formed in $K+1$ evaluations of the velocity.
+The Jacobian transpose can be formed in $\romdimArg{n}+1$ evaluations of the velocity.
 For cases where the reduced-order model is low dimensional, this approach is feasible.

### 2019-10-17 numerical_experiments.tex
minor changes

@@
 In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs.
-The purpose of the numerical experiments is to assess the impact on minimizing the residual over an arbitrarily sized time window on the solution accuracy.
-We additionally assess the impact of the time step, and time scheme on \methodAcronym.
+The purpose of the numerical experiments is to assess the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy.
+We additionally assess the impact of the time step and time scheme on \methodAcronym.
 In both experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$.
@@
 The nonlinear least-squares problem arising at each time instance is solved via the Gauss--Newton method.
+The linear least-squares problems arising at each Gauss--Newton iteration are solved via the normal equations.
 All Jacobians are computed via automatic differentiation~\cite{adolc}.
-The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$.
+The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $10^{-4}$.
 \item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
@@
 arising over each window is solved via the Gauss--Newton method.
-All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$.
+The linear least-squares problems arising at each Gauss--Newton iteration is solved via the normal equations.
+All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $10^{-4}$ (i.e., we use the parameter $\epsilon = 10^{-4}$ in Algorithm~\ref{alg:colloc_gn}).
 Critically, it is noted that the Jacobian
@@
 The coupled two-point boundary
-problem is solved via the forward--backward sweep method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. 
+problem is solved via the forward--backward sweep method.
+The action of the Jacobian transpose on vectors is computed via automatic differentiation.
+We use parameters $\epsilon = 10^{-6}$, $\fbsmGrowth = 1.1$, and $\fbsmDecay=2$ in Algorithm~\ref{alg:st_iter}.
 \end{itemize}

### 2019-10-17 cavity.tex
latest

@@
 
-The reduced-order models leverage POD to construct the \spatialAcronym\ trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to select the sampling points and, as a result the weighting
-matrix $\stweightingMatArg{}$; we use a constant weighting matrix across all windows, e.g., $\stweightingMatArg{n} \equiv \stweightingMatArg{}$.
+The reduced-order models leverage POD to construct the \spatialAcronym\ trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to select the sampling points (and as a result the weighting
+matrix $\stweightingMatArg{}$); we use a constant weighting matrix across all windows, e.g., $\stweightingMatArg{n} \equiv \stweightingMatArg{}$.
 The process used to construct the initial conditions and trial subspaces for the ROMs is as follows:
@@
 \item Evolve the FOM for $t \in [0,400]$.
-\item Reset the time-coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 100$ and $K = \{136,193\}$ over $t \in [0,100]$ to construct two trial subspaces comprising $\approx 95\%$ and $\approx 97\%$ of the snapshot energy, respectively.
+\item Reset the time coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 100$ and $K = \{136,193\}$ over $t \in [0,100]$ to construct two trial subspaces comprising $\approx 95\%$ and $\approx 97\%$ of the snapshot energy, respectively.
 \item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 100$, $n_s = 129$ to obtain the sampling point matrix of dimension $\stweightingMatOneArg{} \in \{0,1\}^{1603 \times \fomdim }$. 
@@
 \subsubsection{Description of reduced-order models}
-Hyper-reduced ROMs based on the Galerkin, collocated Least--Squares Petrov--Galerkin, and \methodAcronym\ approaches are 
-again considered. Details on the implementation of the methods is as follows:
+We consider collocated ROMs based on the Galerkin, least-squares Petrov--Galerkin, and \methodAcronym\ approaches. Details on the implementation of the methods is as follows:
 \begin{itemize}
-\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation. Unless 
-otherwise noted, the Galerkin ROM is evolved in time with the Crank--Nicolson time scheme at a time step of $\Delta t =0.1$.
+\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation. 
+The Galerkin ROM is evolved in time with the CN time scheme at a time step of $\Delta t =0.1$.
 
@@
 discretization. Unless otherwise noted, the time step for the LSPG ROM is $\Delta t  = 0.1$. As in the previous example, 
-the Gauss--Newton method is used to solve the non-linear least-squares problem arising at each time instance. We deem the Gauss--Newton 
+the Gauss--Newton method is used to solve the nonlinear least-squares problem arising at each time instance. The linear least-squares problems 
+arising at each Gauss--Newton iteration are solved via the normal equations. We deem the Gauss--Newton 
 algorithm converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation.
@@
 \item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time step of 
-$\Delta t = 0.1$. We deem the Gauss--Newton algorithm converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation. Uniform quadrature 
+$\Delta t = 0.1$. We deem the Gauss--Newton algorithm converged when the gradient norm is less than $1e-4$. The linear least-squares problems
+arising at each Gauss--Newton iteration are solved via the normal equations. All Jacobians are computed via automatic differentiation. Uniform quadrature 
 weights are used for evaluating the integral in the minimization problem~\eqref{eq:obj_gen_slab}.

### 2019-10-17 numerical_experiments.tex
latest

@@
 \subsubsection{Numerical results}
-The impact of the window size on \methodAcronymROMs\ is first assessed. We consider a set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
+We first assess the impact of the window size on the performance of the \methodAcronymROMs.  We consider a set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
 $\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$.
@@
 First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
-$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, it is observed that the LSPG and Galerkin ROMs accurately characterize 
+$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, we observe that the LSPG and Galerkin ROMs accurately characterize 
 the system: they correctly track the shock location, expansion waves, etc. We observe both predictions, however, to be highly oscillatory. These oscillations are 
@@
 500 time instances) only  
-cost approximately one order of magnitude more than the case where the residual is minimized over a single time-step (i.e., LSPG). Also of interest is the 
+cost approximately one order of magnitude more than the case where the residual is minimized over a single time step (i.e., LSPG). Also of interest is the 
 fact that the direct method utilizing the AB2 time-discretization scheme costs only slightly more per a given window size than the direct method 
@@
 of both the indirect and direct methods can be decreased through the use and/or development of 
-algorithms tailored to the minimization problem.
+algorithms tailored to the windowed minimization problem.
 
@@
 window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
-$100$, $50$, $20$, $10$, $2$, and $1$ time-steps per window). We additionally 
+$100$, $50$, $20$, $10$, $2$, and $1$ time instances per window). We additionally 
 consider LSPG ROMs leveraging the same set of time steps.
-To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 1e-4$.
+To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$.
 Figure~\ref{fig:time_step_study} shows the space--time $\elltwo$ error and residual norm
@@
 demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
-shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that, for all time steps considered, the \methodAcronym-ROMs 
-are less than 6x the cost LSPG; despite the window sizes comprising up to 100 time steps. 
+shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that for all time steps considered, the \methodAcronymROMs\ 
+are less than 6x the cost of LSPG, despite the window sizes comprising up to 100 time instances. 
 \begin{figure}
@@
 We observed that minimizing the residual over an intermediary window size led to the lowest space--time state error in the $\elltwo$ norm.
-\item The \methodAcronym\ displays time-step convergence: both the space--time $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time step decreased.
+\item \methodAcronym\ displays time-step convergence: both the space--time $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time step decreased.
 This is in contrast to LSPG.

### 2019-10-18 section4/spatial_numerical.tex
bug fixes

@@
 Direct approaches solve optimization problem~\eqref{eq:tclsrm} by
-``transcribing" the infinite dimensional optimization problem into a finite
-dimensional one by discretizing the state and objective functional in time.
-The minimization problem is then reformulated as a (non)linear programming, or
-optimization, problem. A variety of direct solution approaches exist, 
+``transcribing" the infinite-dimensional optimization problem into a finite-dimensional one by discretizing the state and objective functional in time.
+The minimization problem is then reformulated as a (non)linear 
+optimization problem. A variety of direct solution approaches exist, 
 including collocation approaches, spectral  methods, and genetic algorithms.  
@@
 \textit{integrand} of the objective function in problem~\eqref{eq:obj_gen_slab}) and 
-(2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
+(2) select a numerical quadrature rule to evaluate the \textit{integral}
+defining the objective function in~\eqref{eq:obj_gen_slab}.
 To this end, we define time grids
-$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$ \EP{Shouldn't this be subset? We can't recover the continuous time interval},
+$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subset[\timeStartArg{n},\timeEndArg{n}]$
+\EP{Shouldn't this be subset? We can't recover the continuous time interval}
+\KTC{Yes},
 $n=1,\ldots,\nslabs$ that
@@
 \begin{cases}
-1,0 & n=1, \; i = 0,\\
-n,i & n \ge 1, \; i > 0, \\
+	(n,i) & n \geq 1, \; i \geq 0, \\
 \indexMapper(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
 \end{cases}$$
-We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
+\KTC{Note: I added parenthesis here because a function can't really map to a
+pair of indices unless they are defined as a collection or tuple. Should be
+fine.}
+We now outline the direct solution approach for linear-multistep schemes; the formulation for
 other time-integration methods (e.g., Runge--Kutta) follows closely. 
@@
 %\end{equation}
-Employing a linear multistep method to discretize the FOM ODE yields the FOM O$\Delta$E over the $n$th window,
+Employing such a method to discretize the FOM ODE yields the sequence of FOM
+O$\Delta$Es defined over the $n$th time window
 %\footnote{
@@
 \begin{align*}
-&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, 
+&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}
 %\begin{cases}
@@
 along with the initial condition $\stateFOMDiscreteArg{1,0} = \stateFOMIC$.
-In the above, $\stateFOMDiscreteArg{n,i} (\approx
+Here, $\stateFOMDiscreteArg{n,i} (\approx
 \stateFOMArg{}{\timeWindowArg{n}{i}})
@@
 \begin{split} 
-\objectiveArgLMS{n}_{\text{D}} &\vcentcolon (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{-\lmsWidthArg{n}{1}+1}) \mapsto 
+\objectiveArgLMS{n}_{\text{D}} &\vcentcolon
+	(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}};
+	\stateyDiscreteArgnt{0},\ldots,\stateyDiscreteArgnt{-\lmsWidthArg{n}{1}+1})
+	\mapsto 
 \frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots, \stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}), \\
@@
 where
+\KTC{This next equation: I don't think you should project the previous
+solutions.}
 \begin{equation*}
@@
 \end{remark} 
-\subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
-	\KTC{I recommend rewriting this section with several important things in
-	mind. We can discuss.}
-	\KTC{If the dynamics are linear, doesn't this just yield a linear
-	least-squares problem that can be solved directly? We need to present this
-	more clearly.}
+\subsubsection{Solution to the least-squares problem through the Gauss--Newton
+	method}
 Discretization through linear multistep methods (as well as other techniques) 

### 2019-10-18 main.tex
minor change
removed projectors

@@
 	addresses key deficiencies in existing model-reduction
-	techniques, e.g., the dependence of (space--time) LSPG projection on the
+	techniques, e.g., the dependence of LSPG and space--time LSPG projection on the
 	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
@@
 %\newline
-Steps:
+\textbf{Steps:}
 \begin{enumerate}

### 2019-10-18 section4/spatial_numerical.tex
removed projectors

--- 2019-10-17 section4/spatial_numerical.tex
@@
 \begin{split}
-\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)})), \\
+\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}), \\
 &\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow \RR{}_+,
@@
 \end{equation*}
-and $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
-$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
-We additionally set $\approxstateDiscreteArg{1,0} = \projectorSpatialArg{n}(\stateFOMIC).
-$
+with the initial condition $\approxstateDiscreteArg{1,0} = \basisspaceArg{n}\basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n}$.
+%We additionally set $\approxstateDiscreteArg{1,0} = \projectorSpatialArg{n}(\stateFOMIC)$, where
+%$\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
+%$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
+
 %where 
@@
 %\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
-\projectorSpatialArg{n}( \approxstateDiscreteArg{\indexMapper(n,0)}),
-\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{\indexMapper(n,1 - \lmsWidthArg{n}{1}) } )  ) \\
+ \approxstateDiscreteArg{\indexMapper(n,0)},
+\ldots , \approxstateDiscreteArg{\indexMapper(n,1 - \lmsWidthArg{n}{1}) }   ) \\
 \vdots \\
- \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots,\projectorSpatialArg{n}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} )) \\
+ \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots, \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} ) \\
 \end{bmatrix},

### 2019-10-18 section3/space_time.tex

@@
 Thus, in the \spaceTimeAcronym\ case, the optimization problem is no longer
-	minimizing a functional with respect to a (time-dependent) function, but now is minimizing a function with 
+	minimizing a functional with respect to a (time-dependent) function, but is minimizing a function with 
 respect to a vector.  As such, the first-order optimality conditions can be
-	derived using standard calculus. Differentiating the objective with respect
+	derived using standard calculus. Differentiating the objective function with respect
 	to the generalized coordinates and setting the result equal to zero yields 

### 2019-10-18 section4/numerical_methods.tex

@@
 \methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
-Numerical techniques to
-solve optimal-control problems can be classified as either
+Numerical solution techniques for
+optimal-control problems can be classified as either
 \textit{direct} or \textit{indirect}
 methods~\cite{conway_optimalcontrolreview}.
-Rather than working with the first order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
-numerically discretizing the system and objective functional and ``transcribing"
-the infinite dimensional problem into a finite-dimensional nonlinear
-programming (optimization) problem. Direct approaches thus
-``discretize-then-optimize''.
-In contrast, indirect methods compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first-order optimality conditions of the objective functional).
-Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
-``indirectly". A variety of both discretization and solution techniques are possible
+Rather than working with the first-order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
+numerically discretizing the system \KTC{What does it mean to discretize the
+system?} and objective functional and ``transcribing"
+the infinite-dimensional problem into a finite-dimensional 
+problem. Direct approaches thus
+``discretize then optimize''. \KTC{Note: only use hyphens when
+discretize-then-optimize is used as an adjective.}
+In contrast, indirect methods compute solutions to the Euler--Lagrange
+equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the
+first-order optimality conditions). \KTC{Note: they are not first-order
+optimality conditions of the objective functional only; they are first-order
+optimality conditions of the optimization problem} Thus, indirect methods
+``optimize then discretize," and solve the optimization problem
+``indirectly.'' A variety of both discretization and solution techniques are possible
 for both direct and indirect methods. Collocation methods,
 finite-element methods, spectral methods, and shooting methods are all examples of
-plausible solution techniques.
+possible solution techniques.
 

### 2019-10-18 main.tex
small

--- 2019-10-17 main.tex
@@
 While the space--time Galerkin method~\cite{benner_st} is another alternative, it does not
-associate with any residual minimization principle, and thus we do not discuss
+associate with any residual-minimization principle, and thus we do not discuss
 it further.
@@
 
-The principle challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to, in general, a higher computational cost.
+The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to, in general, a higher computational cost.
 In the context of the direct solution approach, this increased cost is due to the fact that forming and solving the least-squares problem becomes more expensive as the window size grows.

### 2019-10-18 section4/spatial_numerical.tex
small

--- 2019-10-18 section4/spatial_numerical.tex
@@
 quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace$ and $\stateInterceptArg{n} = \stateInterceptArg{}, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
-$\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
+$\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the
+	time instances correspond to the time grid used for time discretization $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
 $\approxstateDiscreteArg{n,1}  = \approxstateLSPG^n$, $n=1,\ldots,\nslabs$ and  
@@
 	method}
-Discretization through linear multistep methods (as well as other techniques) 
-results in a least-squares problem. In particular, for the case where the full-order model~\eqref{eq:FOM} is nonlinear, the direct approach yields a nonlinear least-squares problem. 
-A variety of algorithms exist for solving nonlinear least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.
-The numerical experiments presented in this work comprise nonlinear systems and are solved via the Gauss--Newton method, and as such we outline this approach here.
+	Problem \eqref{eq:obj_gen_lms_final} corresponds to a discrete least-squares
+	problem, which is nonlinear if the full-order-model velocity $\velocity$ is nonlinear in
+	its first argument. 
+A variety of algorithms exist for solving nonlinear least-squares problems,
+	including the Gauss--Newton method, and the Levenberg--Marquardt method.
+	\KTC{Removed `trust-region approaches', because Levenberg--Marquardt is a
+	trust-region approach. Also added en-dashed.} The numerical experiments
+	presented in this work consider nonlinear dynamical systems and are solved
+	via the Gauss--Newton method; as such, we outline this approach here.
 
-To describe the Gauss--Newton method, we first define a ``vectorization" function 
+Defining a ``vectorization" function 
 \begin{align*}
  \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
-&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm}.
+&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm},
 \end{align*}
-The vectorized generalized coordinates over the $n$th time window are then
+the vectorized generalized coordinates over the $n$th time window are 
 defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
@@
 
-\item \textit{Automatic differentiation}: A more intrusive, but potentially more efficient, method of computing the action of the Jacobian transpose on a vector is through automatic differentiation (AD).
-AD methods comprise a class of techniques that can be used to numerically evaluate derivatives of functions (e.g., Jacobians, vector-Jacobian products) by recursively applying the chain rule.
-The numerical examples presented later in this work leverage AD.
-The principle drawback of AD is that AD methods are intrusive and may not be suitable for, e.g., legacy codes.
+\item \textit{Automatic differentiation}: A more intrusive, but potentially
+	more efficient, method of computing the action of the Jacobian transpose on
+		a vector is through automatic differentiation (AD). AD methods comprise a
+		class of techniques that can be used to numerically evaluate derivatives
+		of functions (e.g., Jacobians, vector-Jacobian products) by recursively
+		applying the chain rule. The numerical examples presented later in this
+		work leverage AD. The principal drawback of AD is that AD methods are intrusive and may not be suitable for, e.g., legacy codes.  
 \end{enumerate}

### 2019-10-18 section4/spatial_numerical.tex
fix 1/2

--- 2019-10-18 section4/spatial_numerical.tex
@@
 To this end, we define time grids
-$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subset[\timeStartArg{n},\timeEndArg{n}]$
-\EP{Shouldn't this be subset? We can't recover the continuous time interval}
-\KTC{Yes},
+$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subset[\timeStartArg{n},\timeEndArg{n}]$,
 $n=1,\ldots,\nslabs$ that

### 2019-10-18 section4/spatial_numerical.tex
update

--- 2019-10-18 section4/spatial_numerical.tex
@@
 the vectorized generalized coordinates over the $n$th time window are 
-defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
-\EP{The notation for space--time in 2.4 uses bars. I don't want to confuse these generalized coordinates with those, they conceptually two different things} 
+defined as 
 \begin{equation*}
@@
 We now define the weighted space--time residual over the entire window as
-\KTC{To make this valid for seeping back into the previous time window, the
-second argument could be omitted here (okay because there is already a
-superscript on the residual) and arbitrary previous solutions could be
-included in the first residual here.}
 \begin{equation*}
@@
 \end{equation*}
-where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} )$ and, for $i \le 0$,
+where $\genstatecollocMatySlabArg{} \equiv \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} )$ and, for $i \le 0$,
 \begin{equation}\label{eq:wls_genspatial_direct_args}
@@
 \end{equation}
-It is worth noting that, by design, 
+We note that
 \begin{equation*}
@@
 \end{equation*} 
-Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method.
-The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and (3) solve the linear least-squares problem and update the guess.
+Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard
+Gauss--Newton method. Each Gauss--Newton iteration consists of three fundamental steps: (1)
+compute the FOM O$\Delta$E residual given the current guess, (2) compute the
+Jacobian of the residual over the time window, and (3) solve the linear least-squares problem and update the guess.
 
-The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the Jacobian of the residual over the time window.
-In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will yield an almost block-diagonal
-sparse matrix with the sparsity pattern
-\begin{equation*}
-\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}(\genstatecollocMatSlabArg{n})  =
- \begin{bmatrix*}[l]
+The practical implementation of the Gauss--Newton algorithm requires an
+efficient method for computing the Jacobian of the residual over the time
+window
+$\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}$.
+For this purpose, we can leverage the fact that this Jacobian is a block lower
+triangular matrix with the following sparsity pattern (for $\lmsWidthArg{n}{i} = 1$,
+$i=1,\ldots,\nstepsArg{n}$):  \KTC{make sure the column widths are correct.
+Maybe add dimensions for each block?}
+\begin{equation*} \begin{bmatrix*}[l]
 \matshapeb & \\%[-5pt]
@@
 \end{equation*}
-This sparsity pattern can be leveraged to assemble the Jacobian of the residual over the time window from Jacobians of the residual at single time instances. 
-Further, for solution techniques based on the normal equations, one can leverage the sparsity pattern 
-to efficiently compute the Jacobian matrix--matrix product.
-It is also worth noting that the 
-normal equations will comprise a banded block system, this can be leveraged to speed up the linear solves at each Gauss--Newton iteration.
+In particular, solution techniques can leverage this structure, e.g., to
+efficiently compute Jacobian--vector products.
+Another consequence of this sparsity pattern is that the normal equations
+arising at each Gauss--Newton iteration 
+comprise a banded block system that can also be exploited.
 
-\begin{remark}\label{remark:gaussnewton}\textit{(Acceleration of the Gauss--Newton Solve)}\\
+\begin{remark}\label{remark:gaussnewton}\textit{(Acceleration of the
+	Gauss--Newton Method)}\\
 	The principal cost of a Gauss--Newton method is often the formation of the Jacobian matrix. A variety of techniques aimed at 
-reducing this computational burden exist; Jacobian-free Newton-Krylov
+reducing this computational burden exist; Jacobian-free Newton--Krylov
 	methods~\cite{jfnk}, Broyden's method~\cite{broyden} (as explored in
 	Ref.~\cite{carlberg_thesis}, Appendix A), and frozen Jacobian approximations
-	are several such examples.
-Further, the space--time formulation introduces an extra dimension for parallelization that can be used to accelerate the wall-clock time of the ROM.
-The investigation of these additional, potentially more efficient, solution algorithms will be a topic of future work.
+	are several such examples. Further, the space--time formulation introduces
+	an extra dimension for parallelization that can be exploited to furture
+	reduce the wall time. The investigation of these additional, potentially
+	more efficient, solution algorithms is a topic of future work. 
 \end{remark}

### 2019-10-18 section4/spatial_numerical.tex
added line search and removed comments we agreee on

@@
 where
-\KTC{This next equation: I don't think you should project the previous
-solutions.}
 \begin{equation*}
@@
 	including the Gauss--Newton method, and the Levenberg--Marquardt method.
-	\KTC{Removed `trust-region approaches', because Levenberg--Marquardt is a
-	trust-region approach. Also added en-dashed.} The numerical experiments
+ The numerical experiments
 	presented in this work consider nonlinear dynamical systems and are solved
@@
 triangular matrix with the following sparsity pattern (for $\lmsWidthArg{n}{i} = 1$,
-$i=1,\ldots,\nstepsArg{n}$):  \KTC{make sure the column widths are correct.
-Maybe add dimensions for each block?}
+$i=1,\ldots,\nstepsArg{n}$): 
 \begin{equation*} \begin{bmatrix*}[l]
-\matshapeb & \\%[-5pt]
+\matshapea & \\%[-5pt]
  \matshapea & \matshapea & \\%[-5pt]
@@
  & &  & \matshapea &  \matshapea 
-\end{bmatrix*}.
+\end{bmatrix*},
 % \begin{bmatrix*}[c]
@@
 \end{equation*}
+where each block comprises an $\fomdim \times \romdimArg{n}$ dense matrix.
 In particular, solution techniques can leverage this structure, e.g., to
@@
 % Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
+$\alpha \leftarrow \text{linesearch}(\Delta  \genstatecollocMatSlabArg{n} ) $ \Comment{Compute $\alpha$ based on a line search, or set to 1.}\\
 $\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\

### 2019-10-18 section4/spatial_numerical.tex
up

--- 2019-10-18 section4/spatial_numerical.tex
@@
 \subsection{\spatialAcronym\ trial subspaces: indirect solution approach}\label{sec:indirect}
-\KTC{This is a fantastic insight and a really great contribution!}
 In contrast to the direct approach,
@@
 equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_adjoint} associated with
-stationarity. The
-system associated with the Euler--Lagrange equations comprises a coupled two-point boundary value
+stationarity. This
+system comprises a coupled two-point boundary value
 problem. Several techniques have
@@
 
-The FBSM is an iterative approach that can be used to
-solve the two-point boundary value problem. The general process of the FBSM is
-as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
-time using an initial guess to the costate to obtain an initial guess to
-the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
-approximation to the generalized coordinates. This gives a new estimate to the
-costate, which is then used to again solve the forward problem to obtain
-a new estimate of the generalized coordinates. This process is continued until
-convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
+Until convergence, the FBSM alternates between solving the
+system~\eqref{eq:lspg_continuous} \textit{forward} in time given a fixed value
+of the costate, and solving the adjoint equation~\eqref{eq:lspg_adjoint}
+\textit{backward} in time given a fixed value for the generalized coordinates.
+Typically, the system~\eqref{eq:lspg_continuous} is solved first given an
+initial guess for the costate.
+%
+%consists of the following steps. First, the
+%system~\eqref{eq:lspg_continuous} is (numerically) solved \textit{forward} in
+%time using an initial guess of the costate to obtain an initial solution for
+%the generalized coordinates. Next, given this initial solution for the
+%generalized coordinates, the adjoint equation~\eqref{eq:lspg_adjoint} is
+%solved \textit{backward} in time for an updated costate solution.
+%approximation to the generalized coordinates. This gives a new estimate to the
+%costate, which is then used to again solve the forward problem to obtain
+%a new estimate of the generalized coordinates. This process is continued until
+%convergence. 
+Algorithm~\ref{alg:st_iter} outlines the algorithm, which 
 contains three parameters: the relaxation factor $\rho \le 1$, the growth factor
@@
 closer $\rho$ is to unity, the faster the algorithm will converge.
-For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
-A proper value of $\rho$ can be
-obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the relaxation factor
+For large window sizes, however, a large a value of $\rho$ can lead to an unstable iterative process. 
+In practice, a line search is used to compute an acceptable value for the
+relaxation factor $\rho$. The line search presented in Algorithm~\ref{alg:st_iter} adapts the relaxation factor
 according to the objective. Convergence properties of the FBSM method are
-presented in Ref.~\cite{McAsey2012ConvergenceOT}. It is shown that, for a small 
-enough value of $\rho$, the algorithm will converge.
+presented in Ref.~\cite{McAsey2012ConvergenceOT}, which shows that the
+algorithm will converge for a sufficiently small 
+value of $\rho$.
 

### 2019-10-18 section4/spacetime_numerical.tex
small
fix 1/2

@@
 \begin{cases}
-1 & n=1, \; i = 0,\\
 n & n \ge 1, \; i > 0, \\
@@
 \begin{remark}
-For the limiting case where one window comprises entire domain (i.e., $\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1} = \lspgWeightingST = \mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
+For the limiting case where one window comprises entire domain (i.e., $\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1}$ and $\lspgWeightingST$ comprise the identity matrix $\mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
 \end{remark}
@@
 where the discrete objective function is given by
-$$\objectiveDiscreteSTArg{n}: \statey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \zeta^{n,i} 
+$$\objectiveDiscreteSTArg{n}: \statey \mapsto \frac{1}{2}\sum_{i=1}^{\ncollocSTArg{n}} \zeta^{n,i} 
 \bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg]^T 

### 2019-10-18 section4/spatial_numerical.tex
index in 14

@@
 \begin{cases}
-	(n,i) & n \geq 1, \; i \geq 0, \\
+	(n,i) & n = 1, \; i = 0, \\
+	(n,i) & n \geq 1, \; i > 0, \\
 \indexMapper(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
 \end{cases}$$
-\KTC{Note: I added parenthesis here because a function can't really map to a
-pair of indices unless they are defined as a collection or tuple. Should be
-fine.}
 We now outline the direct solution approach for linear-multistep schemes; the formulation for

### 2019-10-18 section4/spatial_numerical.tex
update

--- 2019-10-18 section4/spatial_numerical.tex
@@
 %\end{align*}
-\subsubsection{Linear multistep schemes}
+\subsubsection{Linear multistep schemes}\label{sec:lmm}
 Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance of the $n$th window.
@@
 window
-$\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}$.
+${\partial \residLMSSlabArg{n}}/{\partial \genstatecollocMatySlabArg{}}$.
 For this purpose, we can leverage the fact that this Jacobian is a block lower
@@
 \end{algorithm}
-\subsubsection{Numerical considerations for the forward and backward problems}
-The FBSM requires solving the systems~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
+\subsubsection{Considerations for the numerically solving the forward and
+backward systems}
+The FBSM requires solving the forward~\eqref{eq:lspg_continuous} and
+backward~\eqref{eq:lspg_adjoint} systems, both of which are defined at the time-continuous level. 
 The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
-an efficient method for computing the action of the Jacobian transpose on a vector.
+an efficient method for computing terms involving the transpose of the
+Jacobian of the velocity.
 
-This work examines the use of linear multistep methods for the purpose of temporal discretization of the forward and 
-backwards problems. As described in  
-the direct solution approach, temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time instances over each time window.
-Linear multistep methods then use this time-grid for temporal discretization of~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}.
+This work employs linear multistep schemes for time discretization of the
+forward and backward problems. As described in Section \ref{sec:lmm},
+temporal discretization is achieved by introducing $\nstepsArg{n} + 1$ time
+instances over each time window.
 
-The second ingredient, namely an efficient method for computing the action of the Jacobian transpose on a vector, can be challenging. 
-In the case one does not have access to the full-order model Jacobian (or it is too costly to compute), the evaluation of this term can be challenging and expensive.
-Here, two methods are discussed that can be used to evaluate the action of the Jacobian transpose on a vector as seen in~\eqref{eq:lspg_adjoint}:
+The second ingredient, namely devising an efficient method for computing terms
+involving the transpose of the Jacobain of the velocity,
+can be challenging if one
+does not have explicit access to this Jacobian or it is too costly to
+compute. We discuss two methods that can be used to evaluate such terms
+that appear in the forward system~\eqref{eq:lspg_adjoint}:
 \begin{enumerate}
-\item \textit{Jacobian-free approximation}: A non-intrusive way to evaluate the Jacobian transpose in~\eqref{eq:lspg_adjoint} is to recognize that all terms including the Jacobian transpose are left multiplied by the transpose of the trial basis.
-One can make the manipulation,
-$$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T.$$
-This manipulation allows one to compute the action of the Jacobian on each column of $\basisspaceArg{n}$ by, e.g., the finite difference approximation
-$$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$
-The Jacobian transpose can be formed in $\romdimArg{n}+1$ evaluations of the velocity.
-For cases where the reduced-order model is low dimensional, this approach is feasible.
-For higher dimensional reduced-order models, this approach can be prohibitively expensive.
+\item \textit{Jacobian-free approximation}: A non-intrusive way to evaluate
+	these terms is to recognize that all terms including the transpose of the
+		Jacobian of the velocity are left multiplied by the transpose of the
+		spatial trial basis; this can be manipulated as
+$$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}
+		(\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial
+		\stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T,$$
+		which exposes the ability to approximate rows of this matrix via
+		finite differences, e.g., via forward differences as
+$$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)
+		\basisvec_i^n \approx \frac{1}{\epsilon}\bigg(
+		\velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} +
+		\epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg),\quad
+		i=1,\ldots,\romdimArg{n},$$
+which requires $\romdimArg{n}+1$ evaluations of the velocity.
 
-\item \textit{Automatic differentiation}: A more intrusive, but potentially
-	more efficient, method of computing the action of the Jacobian transpose on
-		a vector is through automatic differentiation (AD). AD methods comprise a
-		class of techniques that can be used to numerically evaluate derivatives
-		of functions (e.g., Jacobians, vector-Jacobian products) by recursively
-		applying the chain rule. The numerical examples presented later in this
-		work leverage AD. The principal drawback of AD is that AD methods are intrusive and may not be suitable for, e.g., legacy codes.  
+\item \textit{Automatic differentiation}: A more intrusive, but exact
+	method for computing these terms is
+		through automatic differentiation (AD), which evaluate derivatives of
+		functions (e.g., Jacobians, vector-Jacobian products) in a numerically
+		exact manner by recursively applying the chain rule. The numerical
+		examples presented later in this work leverage AD. The principal drawback
+		of this approach is their intrusiveness, which may prevent them from
+		practical application, e.g., in legacy codes.
+		\KTC{EP: verify this paragraph.}
 \end{enumerate}
 
+\KTC{HERE}
 \begin{remark}\label{remark:fbsm}(Acceleration of Indirect Methods)

### 2019-10-18 main.tex

@@
 	subspace, we consider two different solution techniques: direct (i.e.,
-	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
+	discretize then optimize) and indirect (i.e., optimize then discretize).
 	Numerical experiments conducted using trial subspaces characterized by spatial

### 2019-10-18 main.tex
minor additions in intro

@@
 Petrov--Galerkin projection applied to the time-continuous
-full-order-model residual, and (2) the LSPG and Galerkin projection are equivalent
+full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
 in the limit as the time step goes to zero (i.e., Galerkin projection 
@@
 Existing space--time approaches differ from the classic LSPG and Galerkin approaches in (1) the definition of the space--time trial subspace and (2) the definition of the residual minimization problem.
-First, space--time approaches leverage a space--time trial basis that characterizes both the spatial \textit{and} temporal dependence of the state, classic "spatial" model reduction approaches such as LSPG and Galerkin leverage only a spatial trial basis that characterizes the spatial dependence of the state.
+First, space--time approaches leverage a space--time trial basis that characterizes both the spatial \textit{and} temporal dependence of the state, classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage only a spatial trial basis that characterizes the spatial dependence of the state.
 %First, space--time approaches leverage a space--time trial 
 %subspace that associates with a space--time projection operator; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with a spatial projection operator only.
-Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time-step instance (LSPG).
-The result of these differences is that space--time approaches yield an algebraic system of equations defined
-over all space and time, whose solution comprises a \text{fixed} vector of
-(space--time) generalized coordinates; on the other hand, spatial projection only approaches 
-generally associates with a system of ODEs whose solution
-comprises a \textit{time-dependent} vector of (spatial) generalized
+Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time step (LSPG).
+The result of these differences is that space--time approaches yield a system of algebraic equations defined
+over all space and time, whose solution comprises a vector of
+(space--time) generalized coordinates; on the other hand, spatial-projection-only approaches 
+generally associate with systems of ODEs whose solutions
+comprise \textit{time-dependent} vectors of (spatial) generalized
 coordinates.
@@
 Space--time residual minimization approaches minimize the FOM residual over all of space and time and, as a result, yield models that are equipped with a notion of space--time optimality and \textit{a priori} error bounds that grow more slowly
-in time.
+in time. Further, space--time approaches reduce both the spatial and temporal dimensions of the full-order-model, and thus promise cost savings over spatial-projection-only approaches.
 % and the resulting \textit{trajectory} exhibits a notion of
@@
 %\EP{Come back to this, this also has to do with residual minimization}
-However, these techniques also suffer from several
+However, space--time techniques also suffer from several
 limitations. First, the computational cost of solving the algebraic system arising from space--time

### 2019-10-18 section4/spatial_numerical.tex
changed inputs for algo 2

@@
 \Input{tolerance, $\epsilon$; relaxation factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
-factor, $\fbsmDecay \ge 1$} 
+factor, $\fbsmDecay \ge 1$; initial guess for state $\genstate^n_0$; initial guess for costate $\controllerArgnt{n}$} 
 \Output{Stationary point, $\genstateArgnt{n}$ }
 \textbf{Online Steps:}\\ 
-$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
-$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
-guess for costate}\\ 
+%$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
+%$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
+%guess for costate}\\ 
 $\text{Compute } \genstateArgnt{n}_1 \text{ satisfying }  \massArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}

### 2019-10-18 section4/spacetime_numerical.tex
respond to comment

@@
 \begin{cases}
+n & n = 1, \; i = 0, \\
 n & n \ge 1, \; i > 0, \\
@@
 discretization of the integral.
-Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
+Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subset [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
 %with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$

### 2019-10-18 section4/spatial_numerical.tex
respond to comment

@@
 % Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
-$\alpha \leftarrow \text{linesearch}(\Delta  \genstatecollocMatSlabArg{n} ) $ \Comment{Compute $\alpha$ based on a line search, or set to 1.}\\
-$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
+$\alpha \leftarrow \text{linesearch}(\Delta  \genstatecollocMatSlabArg{n} ,\genstatecollocMatSlabArg{n}_k ) $ \Comment{Compute $\alpha$ based on a line search, or set to 1}\\
+$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \alpha \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
 }
@@
 		practical application, e.g., in legacy codes.
-		\KTC{EP: verify this paragraph.}
+		\KTC{EP: verify this paragraph.} \EP{This looks good to me}
 \end{enumerate}

### 2019-10-18 section4/spacetime_numerical.tex
latest

@@
 \subsection{\spaceTimeAcronym\ trial spaces: direct and indirect methods}
-We now briefly consider \spaceTimeAcronym\ trial subspaces. 
+We now consider \spaceTimeAcronym\ trial subspaces. 
 For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
@@
 dimensional representation of the objective functional~\eqref{eq:obj_gen_slab}.
-In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE.
+We describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE.
 We associate direct methods as those
@@
 The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E.
-For brevity, we again focus on linear multistep methods and leverage the time grids introduced in Section~\ref{sec:direct}.
+For brevity, we focus on linear multistep methods and leverage the time grids introduced in Section~\ref{sec:direct}.
 For notational simplicity, we define an index mapping function that is
@@
 $$
-It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
+The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
 The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.

### 2019-10-18 section3/space_time.tex
problems in 4.5

@@
 \EP{Is it wrong to say $\stbasisArg{n}{\cdot} \in \RR{\fomdim \times \stdimArg{n}}$ with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy) \ldots$? I don't particularly like
-$\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}}\otimes\timeSpaceArg{n}$; as the reader I think that $\stbasisArgnt{n}$ can take on all temporal functions.}
+$\stbasisArgnt{n} \in \RR{\fomdim \times
+\stdimArg{n}}\otimes\timeSpaceArg{n}$; as the reader I think that
+$\stbasisArgnt{n}$ can take on all temporal functions.} \KTC{I'm not
+understanding why this is confusing. We previously said that a time dependent
+$n$-vector is $\RR{n}\otimes \timeSpaceArg{n}$. This is the same exact thing
+except for a time-dependent matrix.}
 Here $\stbasisArgnt{n} \in \RR{\fomdim \times

### 2019-10-18 section4/spacetime_numerical.tex
problems in 4.5

@@
-\subsection{\spaceTimeAcronym\ trial spaces: direct and indirect methods}
-We now consider \spaceTimeAcronym\ trial subspaces. 
-For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
-dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
-dimensional representation of the objective functional~\eqref{eq:obj_gen_slab}.
-We describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE.
-We associate direct methods as those
-that work with the O$\Delta$E and indirect methods with those that work directly with the ODE.
+\subsection{\spaceTimeAcronym\ trial subspaces: direct and indirect methods}
+We now consider \spaceTimeAcronym\ trial subspaces.  Because the optimization
+variables (i.e., the generalized coordinates) in this case are already finite
+dimensional, solvers for this type of trial subspace need only develop a
+finite-dimensional representation of the objective
+functional in problem~\eqref{eq:obj_gen_slab_spacetime}. We describe two
+techniques for this purpose: a direct method 
+that operates on the FOM O$\Delta$E and an indirect method that operates on the FOM ODE.
 
-\subsection{\spaceTimeAcronym: direct solution approach}
-The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E.
-For brevity, we focus on linear multistep methods and leverage the time grids introduced in Section~\ref{sec:direct}.
-For notational simplicity, we define an index mapping function that is
-equivalent to the mapping function $\indexMapper$, but outputs only the first argument: 
+\subsection{\spaceTimeAcronym\ trial subspaces: direct solution approach}
+\KTC{Fix this section:
+\begin{enumerate} 
+\item remove $\theta^\star$
+\item replace $J_{D-ST}^n$ in (38) with $J_D^n$. You can do this because of Eq
+	(33)
+\item remove definition of $J_{D-ST}$.
+ \end{enumerate}
+}
+The direct solution technique seeks to minimize the fully discrete objective
+function associated with the FOM O$\Delta$E. For brevity, we focus on linear
+multistep methods and leverage the time grids introduced in
+Section~\ref{sec:direct}. For notational simplicity, we define an
+index-mapping function that is equivalent to the mapping function
+$\indexMapper$, but outputs only the first argument: 
 $$\indexMappern: (n,i) \mapsto 
@@
 %outline the case for linear multistep schemes. 
-Assuming $\text{Rank}(\stweightingMatOneArg{n})\nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ with the direct approach and an \spaceTimeAcronym\ trial subspace sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
+\methodAcronym\ with an \spaceTimeAcronym\ trial subspace and the direct approach sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
  \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 \end{equation}
-where the space--time objective function for the direct method is
+where the space--time objective function for the direct method is defined as
 $$
@@
 %}
-\subsection{\spaceTimeAcronym: indirect solution approach}
+\subsection{\spaceTimeAcronym\ trial subspaces: indirect solution approach}
 As opposed to the direct approach, the indirect approach directly minimizes the continuous objective function~\eqref{eq:obj_gen_slab_spacetime} and sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
@@
 discretization of the integral.
-Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subset [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
+Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subset [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
 %with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$
@@
 %n=1. \end{cases} \end{equation} 
-\subsection{\spaceTimeAcronym\: summary}
+\subsection{\spaceTimeAcronym\ trial subspaces: summary}
 \spaceTimeAcronym\ trial subspaces yield a series of space--time systems of algebraic equations over each window.

### 2019-10-18 section4/spatial_numerical.tex
problems in 4.5

@@
 \end{equation*}
+\KTC{On the RHS, shouldn't this be $\objectiveArgLMS{n}_\text{D}$? I couldn't
+find the definition of $\objectiveArgLMS{n}$ anywhere.}
 with the initial condition $\approxstateDiscreteArg{1,0} = \basisspaceArg{n}\basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n}$.
@@
 		practical application, e.g., in legacy codes.
-		\KTC{EP: verify this paragraph.} \EP{This looks good to me}
 \end{enumerate}
 
-\KTC{HERE}
 \begin{remark}\label{remark:fbsm}(Acceleration of Indirect Methods)
 The FBSM is a simple iterative method for solving the coupled two-point boundary value problem. For large time windows, however, the FBSM may require many 
-forward--backward iterations for convergence. More sophisticated solution techniques, such as a multiple FBSM method or multiple shooting methods, promise 
-to reduce this cost. Analyses of additional solution techniques will be a subject of future work.
+forward--backward iterations for convergence. More sophisticated solution
+	techniques, such as a multiple FBSM method or multiple shooting methods, can 
+reduce this cost in principle. Analyzing additional solution techniques is
+	the subject of future work.
 \end{remark}

### 2019-10-18 section4/spacetime_numerical.tex
to analysis

@@
 The direct solution technique seeks to minimize the fully discrete objective
-function associated with the FOM O$\Delta$E. For brevity, we focus on linear
-multistep methods and leverage the time grids introduced in
+function associated with the FOM O$\Delta$E. We consider linear
+multistep methods and leverage the discretization introduced in
 Section~\ref{sec:direct}. For notational simplicity, we define an
@@
 The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
+\KTC{I would add a remark that highlights the similarities between the direct
+approach for the two kinds of subspaces. They minimize the exact same
+objective function, which is interesting.}
 \begin{remark}
-For the limiting case where one window comprises entire domain (i.e., $\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1}$ and $\lspgWeightingST$ comprise the identity matrix $\mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
+For the limiting case where one window comprises entire domain (i.e.,
+	$\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial
+	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices
+	$\stweightingMatArg{1}$ and $\lspgWeightingST$ comprise the identity matrix
+	$\mathbf{I}$ \KTC{Don't we just need simply
+	$W_{ST}=\text{diag}(W^1)$ (with $W^1$ the half-factor of $A^1$)?. I think this
+	would be more general.}, and $\nstepsArg{1} = N_t$ time instances are employed that
+	satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then
+	$\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\
+	with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG. 
 \end{remark}
@@
 \begin{remark}
-To enable equivalence in the case for general $\lspgWeightingSTArg{\cdot}$, the weighting matrix $\stweightingMatArg{1}$ needs to be a time-dependent matrix function such that the objective function~\eqref{eq:obj_gen_lms_final_st} permits a modified space--time norm.
-This is not pursued here.
+To enable equivalence in the case for a general ST-LSPG weighting matrix
+	$\lspgWeightingSTArg{\cdot}$, the weighting matrix $\stweightingMatArg{1}$
+	must be time dependent matrix-valued, which associates the objective function 
+	\eqref{eq:obj_gen_lms_final_st} with a modified space--time
+	norm. For notational simplicity, we do not consider this case in the current
+	manuscript.
 \end{remark}
@@
 discretization of the integral.
-Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subset [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
+To this end, we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subset [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$.
 %with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$
@@
 \stweightingMatArg{n} 
-\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg],
+\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg]
 $$
-where $\zeta^{n,i} \in \RRplus$, $i=1,\ldots,\ncollocSTArg{n}$ are quadrature weights over the $n$th window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.
+\KTC{Can you add the domain and codomain of the function to be consistent with
+earlier functionals?} and $\zeta^{n,i} \in \RRplus$, $i=1,\ldots,\ncollocSTArg{n}$ are quadrature
+weights over the $n$th time window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.
 %$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 

### 2019-10-18 main.tex
update

@@
 model-reduction methods, and to provide a unifying framework from which
-exisiting methods can be assessed. Specifically, we look to overcome the
+existing methods can be assessed. Specifically, we look to overcome the
 complex dependence of LSPG on the time discretization,
@@
 \section{Conclusions}\label{sec:conclude}
-This paper introduced the windowed least-squares (\methodAcronym) \approachKwd\ for model reductio\  of dynamical systems.
+This paper introduced the windowed least-squares (\methodAcronym) \approachKwd\ for model reduction  of dynamical systems.
 The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial subspace over a series of time windows.

### 2019-10-18 section4/spacetime_numerical.tex
update

@@
 weights over the $n$th time window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.
+\KTC{It would be great to a comment about how this is equivalent to Qiqi Wang
+and Paul Constantine's method in the case of one time window and $\Pi^1$
+corresponding to full solution trajectories. They use collocation-based
+hyper-reduction at a few points in time (but for the full ODE residual at that
+time), which you could mention. This is a very cool connection I didn't
+realize before, and I think it's definitely worth mentioning!}
 %$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 

### 2019-10-18 section4/spacetime_numerical.tex
response

@@
 \begin{enumerate} 
-\item remove $\theta^\star$
+\item remove $\theta^\star$ \EP{I think we still need this for Eq. 38}
 \item replace $J_{D-ST}^n$ in (38) with $J_D^n$. You can do this because of Eq 
-	(33)
-\item remove definition of $J_{D-ST}$.
+	(33) \EP{Yes, done}
+\item remove definition of $J_{D-ST}$. \EP{Yes, done}
  \end{enumerate}
@@
 \methodAcronym\ with an \spaceTimeAcronym\ trial subspace and the direct approach sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
+\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. Why don't we need to worry about this?}
  \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 & \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_{\text{D-ST}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}),%  \\
+\objectiveArgLMS{n}_{\text{D}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n} ;\\
+& \hspace{2. in}  \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
+ \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).%  \\
 %&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
@@
 \end{equation}
-where the space--time objective function for the direct method is defined as
-$$
-\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots,
- \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
-$$
+%where the space--time objective function for the direct method is defined as
+%$$
+%\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots,
+% \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
+%$$
 The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace.
 The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
+\begin{remark}
+\methodAcronym\ with the direct solution approach and an \spaceTimeAcronym\ subspace comprises minimizing the same objective function as \methodAcronym\ with the direct solution approach and an \spatialAcronym\ trial subspace.
+\end{remark}
 \KTC{I would add a remark that highlights the similarities between the direct
 approach for the two kinds of subspaces. They minimize the exact same
-objective function, which is interesting.}
+objective function, which is interesting.} \EP{See if you agree with what I have}
 \begin{remark}
@@
 	$\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial
-	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices
-	$\stweightingMatArg{1}$ and $\lspgWeightingST$ comprise the identity matrix
-	$\mathbf{I}$ \KTC{Don't we just need simply
+	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \text{diag}(\stweightingMatOneArg{1})$
+	\KTC{Don't we just need simply
 	$W_{ST}=\text{diag}(W^1)$ (with $W^1$ the half-factor of $A^1$)?. I think this
-	would be more general.}, and $\nstepsArg{1} = N_t$ time instances are employed that
+	would be more general.}\EP{Yes, this is better. Is diag the right command though? It should be $W_{ST} = [W \; W \; W \; W]$}, and $\nstepsArg{1} = N_t$ time instances are employed that
 	satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then
@@
 where the discrete objective function is given by
-$$\objectiveDiscreteSTArg{n}: \statey \mapsto \frac{1}{2}\sum_{i=1}^{\ncollocSTArg{n}} \zeta^{n,i} 
+\begin{equation*}
+\begin{split}
+\objectiveDiscreteSTArg{n} &\vcentcolon \statey \mapsto \frac{1}{2}\sum_{i=1}^{\ncollocSTArg{n}} \zeta^{n,i} 
 \bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg]^T 
 \stweightingMatArg{n} 
-\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg]
-$$
-\KTC{Can you add the domain and codomain of the function to be consistent with
-earlier functionals?} and $\zeta^{n,i} \in \RRplus$, $i=1,\ldots,\ncollocSTArg{n}$ are quadrature
+\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg] \\
+& \vcentcolon \RR{\fomdim} \otimes \timeSpaceArg{n} \rightarrow \RRplus,
+\end{split}
+\end{equation*}
+and $\zeta^{n,i} \in \RRplus$, $i=1,\ldots,\ncollocSTArg{n}$ are quadrature
 weights over the $n$th time window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.

### 2019-10-18 main.tex
update language in proof1
clean up and conclusions

@@
 \section{Conclusions}\label{sec:conclude}
-This paper introduced the windowed least-squares (\methodAcronym) \approachKwd\ for model reduction  of dynamical systems.
-The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial subspace over a series of time windows.
-The formulation is applicable to both spatial and space--time dimension reduction trial subspaces.
-The Galerkin, Least-Squares Petrov--Galerkin (LSPG) and space--time LSPG methods were shown to be specific instances of the proposed method
-
-We considered two differing techniques to solve the least-squares minimization problems arising from the \methodAcronym\ \approachKwd: direct (discretize-then-optimize) and indirect (optimize-then-discretize).
-The case of \spatialAcronym\ trial subspaces proved particularly interesting: it was shown that indirect methods comprise solving a coupled two-point 
+This paper proposed the windowed least-squares (\methodAcronym) \approachKwd\ for model reduction  of dynamical systems. The approach sequentially minimizes the
+	time-continuous full-order-model residual within a low-dimensional space--time trial
+	subspace over time windows.
+The approach was formulated for two types of trial subspaces: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model.
+For each type of trial
+	subspace, we then outlined two different solution techniques: direct (i.e.,
+	discretize then optimize) and indirect (i.e., optimize then discretize). We showed that particular instances of the approach recover Galerkin,
+	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection.
+
+The case of \spatialAcronym\ trial subspaces is of particular interest in the \methodAcronym\ context: it was shown that indirect methods comprise solving a coupled two-point 
 Hamiltonian boundary value problem.
-The forward problem evolves the standard Galerkin ROM forced by a ``costate" variable, while a backward problem evolves the costate and is forced by the ROM residual.
+The forward system, which is forced by an auxiliary costate,
+		evolves the (spatial) generalized coordinates of the ROM in time. The
+		backward system, which is forced by the time-continuous FOM residual
+		evaluated about the ROM state, governs the dynamics of the costate.
 
@@
 The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals.
-Increasing the window size over which the residual was minimized, however, did not necessarily decrease the $\elltwo$ solution error; this was observed to occur over an intermediary window size.
+Increasing the window size over which the residual was minimized, however, did not necessarily decrease the solution error in the $\elltwo$-norm; this was observed to occur over an intermediary window size.
 We additionally observed that the
-\methodAcronym\ \approachKwd\ overcomes the time-step sensitivity challenges that LSPG is subject to.
+\methodAcronym\ \approachKwd\ overcomes the time-discretization sensitivity that LSPG is subject to.
 The second numerical experiment, which examined collocated ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow.
 In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up.
-Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower $\elltwo$ error.
+Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower error in the $\elltwo$-norm.
 
@@
 
-The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to, in general, a higher computational cost.
-In the context of the direct solution approach, this increased cost is due to the fact that forming and solving the least-squares problem becomes more expensive as the window size grows.
+The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost.
+In the context of the direct solution approach, this increased cost is due to the increased cost of forming and solving the least-squares problem associated with larger window sizes.
 In the context of indirect methods, this increased cost is a result of the increased number of iterations of the forward backward sweep method.
@@
 We define this function to be a stationary point of~\eqref{eq:el_gen} (with $\statez$ being the first argument and $\statezDot$ being the second argument) subject to the boundary condition
-	$\statez(a) = \statez_a$. \KTC{This previous sentence isn't correct. The
-	functional defined here has two arguments; the extremum can't be a single
-	argument, as it's a pair of arguments. You need to connect these by defining
-	a new function that maps a time-dependent function to this functional
-	evaluated at the time-dependent function and its derivative..}\EP{ Check this, I think this is more precise now.} 
-We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$
+	$\statez(a) = \statez_a$. 
+We additionally introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$
 with the boundary condition $\variationArg{a} = \bz$ and define a variation
@@
 We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{a} = \bz$.
-We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 
+We define a new function that is equivalent to the function~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 
 $\statezDotBar \equiv d\statezBar / d\timeDummy$ in the second argument,
@@
 We integrate the second term by parts,
-\begin{align}\label{eq:gradient}
-	\begin{split}
+\begin{multline}\label{eq:gradient}
+%	\begin{split}
 %\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg] dt 
 \frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-	=&  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }
+	=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }
 	\big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big)
@@
 	\stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon)
-	,t \big) \bigg) \variation(t) \bigg]dt + \\ &\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) , 
-	\end{split}
-\end{align}
+	,t \big) \bigg) \variation(t) \bigg]dt + \\ 
+\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) , 
+%	\end{split}
+\end{multline}
 where we have used $\variationArg{a} = \bz$.
-Substituting Eq.~\eqref{eq:gradient} in the stationarity conditions~\eqref{eq:stationaryOne} yields
+Substituting Eq.~\eqref{eq:gradient} in the stationarity condition~\eqref{eq:stationaryOne} yields
 \begin{equation}\label{eq:statOne}
@@
 point; thus, setting $\epsilon = 0$ in Eq.~\eqref{eq:statOne} yields
-\begin{equation}
+\begin{equation}\label{eq:euler_lagrange_analysis}
   \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b) ,b\big) \variation(b)  = 0.
@@
 %\end{equation}
-\begin{gather}\label{eq:euler_lagrange_appendix}
-\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete }
+\begin{equation}\label{eq:euler_lagrange_appendix}
+\begin{split}
+\bigg[ &\frac{\partial \integrand  }{\partial \stateyDiscrete }
 	\big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg(
 	\frac{\partial \integrand}{\partial \stateyDotDiscrete }
-	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz  \\
-\statez(a) = \statez_a, \qquad \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \bigg]^T = \bz.
-\end{gather}
+	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz , \\
+&\statez(a) = \statez_a, \hspace{0.65 in} \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \bigg]^T = \bz.
+\end{split}
+\end{equation}
 %\genstate(a) = \genstate_a, \qquad \bigg[ \frac{\partial \integrand}{\partial {\genstateDot}} \bigg]^T_{t=b} = \bz.
@@
 \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
-\stateIntercept,t) \big], \\ & 
+\stateIntercept,\timeDummy) \big], \\ & 
 \vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
@@
 \begin{align*}
-&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
-&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
-&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
+&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
+&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) = -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
+&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) = \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
 \end{align*}

### 2019-10-18 section4/spacetime_numerical.tex
clean up and conclusions

@@
 \item remove $\theta^\star$ \EP{I think we still need this for Eq. 38}
-\item replace $J_{D-ST}^n$ in (38) with $J_D^n$. You can do this because of Eq 
-	(33) \EP{Yes, done}
-\item remove definition of $J_{D-ST}$. \EP{Yes, done}
  \end{enumerate}
@@
 \methodAcronym\ with an \spaceTimeAcronym\ trial subspace and the direct approach sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
-\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. Why don't we need to worry about this?}
+\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. I think we need to mention this, I put in a note below. Let me know if I'm missing something}
  \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. 
-The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
+Assuming $\text{Rank}(\stweightingMatOneArg{n}) \nstepsArg{n} \ge \stdimArg{n}$, the minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
 \begin{remark}
@@
 \begin{remark}
-For the limiting case where one window comprises entire domain (i.e.,
+For the limiting case where one window comprises the entire domain (i.e.,
 	$\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial
-	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \text{diag}(\stweightingMatOneArg{1})$
+	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \begin{bmatrix} \stweightingMatOneArg{1} & \cdots & \stweightingMatOneArg{1}\end{bmatrix}$
 	\KTC{Don't we just need simply
 	$W_{ST}=\text{diag}(W^1)$ (with $W^1$ the half-factor of $A^1$)?. I think this
-	would be more general.}\EP{Yes, this is better. Is diag the right command though? It should be $W_{ST} = [W \; W \; W \; W]$}, and $\nstepsArg{1} = N_t$ time instances are employed that
+	would be more general.}\EP{Yes, this is better. Not sure if diag is right though, see if you agree.}, and $\nstepsArg{1} = N_t$ time instances are employed that
 	satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then
@@
 \stweightingMatArg{n} 
-\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg] \\
+\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg], \\
 & \vcentcolon \RR{\fomdim} \otimes \timeSpaceArg{n} \rightarrow \RRplus,
@@
 weights over the $n$th time window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.
+
+\begin{remark}
+\methodAcronym\ with \spaceTimeAcronym\ trial subspaces solved via the indirect approach naturally achieves ``collocation" in time as the full-order model residual needs to be queried at only the quadrature points.
+\end{remark}
+\begin{remark}
+For the limiting case where one window comprises the entire domain (i.e.,
+	$\DeltaSlabArg{1}\equiv T$), the trial
+	space is set to comprise the span of full solution trajectories, $\stspaceSTArg{1} = \text{span}\{ \stateFOMArgnt{}_i\}_{i=1}^{\stdimArg{1}}$ (where $\stateFOMArgnt{}_i$ are obtained, e.g.,  from training simulations at different parameter instances), and the weighting matrix is set to $\stweightingMatArg{1} = \mathbf{I}$, \methodAcronym\ with \spaceTimeAcronym\ trial subspaces and the indirect approach closely resembles the model reduction procedure proposed in~\cite{constantine_strom}; the approaches differ only in that Ref.~\cite{constantine_strom} places the constraint $\sum_{i=1}^{\stdimArg{1}} \stgenstateArg{1}_i = 1$.
+\end{remark}
+
 \KTC{It would be great to a comment about how this is equivalent to Qiqi Wang
@@
 realize before, and I think it's definitely worth mentioning!}
+\EP{See if you agree with what I have. I worked with the trial space instead of the basis because we prescribe certain structure to the basis (such as the affine transformation)}.
 %$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
@@
 As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces,
-a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
-space--time trial subspace yields a series of dense systems to be solved over each window.
+a detailed exposition of solution techniques for these systems is not pursued here.
+It is sufficient to say that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces yields a series of dense systems to be solved over each window.

### 2019-10-18 main.tex
latest

@@
 that such properties are difficult to assess quantitatively. As a result, it
-is difficult to develop reduced-order models that preserve important
+is challenging to develop reduced-order models that preserve important
 dynamical-system properties, which often results in methods that yield
@@
 efforts include residual-minimization approaches that equip the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
-space--time approaches that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
+space--time approaches that lead to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
 ``energy-based'' inner
@@
 Residual-minimization methods in model reduction compute the solution within a
-low-dimensional trial subspace that minimizes the full-order-model residual.
+low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods,~\cite{bochev_leastsquares,bochev_lsfem_book} which predate ROM residual-minimization approaches.}
 Researchers have developed such residual-minimizing model-reduction methods
@@
 numerical experiments have demonstrated that LSPG often yields more
-accurate and stable solutions than Galerkin
-~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
+accurate and stable solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
 The common intuitive explanation for this improved performance is that, by
@@
 Space--time residual minimization approaches minimize the FOM residual over all of space and time and, as a result, yield models that are equipped with a notion of space--time optimality and \textit{a priori} error bounds that grow more slowly
-in time. Further, space--time approaches reduce both the spatial and temporal dimensions of the full-order-model, and thus promise cost savings over spatial-projection-only approaches.
+in time. Further, space--time approaches reduce both the spatial and temporal dimensions of the full-order model, and thus promise cost savings over spatial-projection-only approaches.
 % and the resulting \textit{trajectory} exhibits a notion of
@@
 \caption{Relationship diagram for the \methodAcronym\ approach for model
-	reduction. \KTC{Also, can we make this more specific? In particular, can
-	we write specific equations (e.g., ``$W = T$ in Eq.~(10)'') for the limiting
-	cases?} \EP{I tried, it becomes very crowded. I think with the description of trial spaces its pretty good now}} 
+	reduction.} 
 \label{fig:flowchart} 
@@
 	model reduction. The approach sequentially minimizes the time-continuous
-		full-order model residual over arbitrary time windows.
+		full-order-model residual over arbitrary time windows.
 %\item Support of two space--time trial subspaces: one that reduces the spatial
@@
 	\item Derivation of two solution techniques:
-		\textit{discretize-then-optimize} (i.e., direct) and
-		\textit{optimize-then-discretize} (i.e., indirect). 
+		\textit{discretize then optimize} (i.e., direct) and
+		\textit{optimize then discretize} (i.e., indirect). 
 	\item Remarks and derivations of conditions under which the

### 2019-10-18 section3/space_time.tex
latest

--- 2019-10-18 section3/space_time.tex
@@
 \end{equation}
-\EP{Is it wrong to say $\stbasisArg{n}{\cdot} \in \RR{\fomdim \times \stdimArg{n}}$ with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy) \ldots$? I don't particularly like
-$\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}}\otimes\timeSpaceArg{n}$; as the reader I think that $\stbasisArgnt{n}$ can take on all temporal functions.}
 Here $\stbasisArgnt{n} \in \RR{\fomdim \times

### 2019-10-18 section4/numerical_methods.tex
latest

@@
 Rather than working with the first-order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
-numerically discretizing the system \KTC{What does it mean to discretize the
-system?} and objective functional and ``transcribing"
+numerically discretizing the objective functional and ``transcribing"
 the infinite-dimensional problem into a finite-dimensional 
 problem. Direct approaches thus
-``discretize then optimize''. \KTC{Note: only use hyphens when
-discretize-then-optimize is used as an adjective.}
+``discretize then optimize''. 
 In contrast, indirect methods compute solutions to the Euler--Lagrange
 equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the
-first-order optimality conditions). \KTC{Note: they are not first-order
-optimality conditions of the objective functional only; they are first-order
-optimality conditions of the optimization problem} Thus, indirect methods
+first-order optimality conditions). Thus, indirect methods
 ``optimize then discretize," and solve the optimization problem

### 2019-10-22 section4/spacetime_numerical.tex
through fixes

@@
 \subsection{\spaceTimeAcronym\ trial subspaces: direct solution approach}
-\KTC{Fix this section:
-\begin{enumerate} 
-\item remove $\theta^\star$ \EP{I think we still need this for Eq. 38}
- \end{enumerate}
-}
 The direct solution technique seeks to minimize the fully discrete objective
@@
 \methodAcronym\ with an \spaceTimeAcronym\ trial subspace and the direct approach sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
-\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. I think we need to mention this, I put in a note below. Let me know if I'm missing something}
  \begin{equation}\label{eq:obj_gen_lms_final_st}
@@
 \begin{remark}
-\methodAcronym\ with the direct solution approach and an \spaceTimeAcronym\ subspace comprises minimizing the same objective function as \methodAcronym\ with the direct solution approach and an \spatialAcronym\ trial subspace.
+	Comparing optimization problems \eqref{eq:obj_gen_lms_final} and
+	\eqref{eq:obj_gen_lms_final_st} reveals that
+\methodAcronym\ with the direct solution approach minimizes the same objective
+	function in the case of both \spaceTimeAcronym\ and \spatialAcronym\ trial
+	subspaces.
 \end{remark}
-\KTC{I would add a remark that highlights the similarities between the direct
-approach for the two kinds of subspaces. They minimize the exact same
-objective function, which is interesting.} \EP{See if you agree with what I have}
 \begin{remark}
@@
 	$\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial
-	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \begin{bmatrix} \stweightingMatOneArg{1} & \cdots & \stweightingMatOneArg{1}\end{bmatrix}$
-	\KTC{Don't we just need simply
-	$W_{ST}=\text{diag}(W^1)$ (with $W^1$ the half-factor of $A^1$)?. I think this
-	would be more general.}\EP{Yes, this is better. Not sure if diag is right though, see if you agree.}, and $\nstepsArg{1} = N_t$ time instances are employed that
+	subspace is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix
+	$\lspgWeightingST = \mathrm{diag}(\stweightingMatOneArg{1})$, and $\nstepsArg{1} = N_t$ time instances are employed that
 	satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then
@@
 	$\DeltaSlabArg{1}\equiv T$), the trial
-	space is set to comprise the span of full solution trajectories, $\stspaceSTArg{1} = \text{span}\{ \stateFOMArgnt{}_i\}_{i=1}^{\stdimArg{1}}$ (where $\stateFOMArgnt{}_i$ are obtained, e.g.,  from training simulations at different parameter instances), and the weighting matrix is set to $\stweightingMatArg{1} = \mathbf{I}$, \methodAcronym\ with \spaceTimeAcronym\ trial subspaces and the indirect approach closely resembles the model reduction procedure proposed in~\cite{constantine_strom}; the approaches differ only in that Ref.~\cite{constantine_strom} places the constraint $\sum_{i=1}^{\stdimArg{1}} \stgenstateArg{1}_i = 1$.
+	subspace is set to the span of full solution trajectories, $\stspaceSTArg{1}
+	= \text{span}\{ \stateFOMArgnt{}_i\}_{i=1}^{\stdimArg{1}}$ (where
+	$\stateFOMArgnt{}_i$ are obtained, e.g.,  from training simulations at
+	different parameter instances), and the weighting matrix is set to
+	$\stweightingMatArg{1} = \mathbf{I}$, \methodAcronym\ with
+	\spaceTimeAcronym\ trial subspaces and the indirect approach closely
+	resembles the model reduction procedure proposed in
+	Ref.~\cite{constantine_strom}; the approaches differ only in that
+	Ref.~\cite{constantine_strom} imposes the constraint
+	$\sum_{i=1}^{\stdimArg{1}} \stgenstateArg{1}_i = 1$ in the associated
+	minimization problem. 
 \end{remark}
-
-\KTC{It would be great to a comment about how this is equivalent to Qiqi Wang
-and Paul Constantine's method in the case of one time window and $\Pi^1$
-corresponding to full solution trajectories. They use collocation-based
-hyper-reduction at a few points in time (but for the full ODE residual at that
-time), which you could mention. This is a very cool connection I didn't
-realize before, and I think it's definitely worth mentioning!}
-\EP{See if you agree with what I have. I worked with the trial space instead of the basis because we prescribe certain structure to the basis (such as the affine transformation)}.
 %$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 

### 2019-10-23 numerical_experiments.tex
update

@@
 \subsubsection{Description of reduced-order models}
-We consider reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approach. No hyper-reduction is considered 
+We consider reduced-order models based on Galerkin projection, LSPG
+projection, and the \methodAcronym\ approach. No hyper-reduction is considered 
 in this example, i.e., $\stweightingMatArg{n} = \mathbf{I}$, $n=1,\ldots,\nslabs$. Details on the implementation of the 
@@
 
-\item \textit{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time scheme as previously described.
-Unless noted otherwise, the LSPG ROM uses a constant time step of $\Delta t = 0.002$.
-The nonlinear least-squares problem arising at each time instance is solved via the Gauss--Newton method.
-The linear least-squares problems arising at each Gauss--Newton iteration are solved via the normal equations.
-All Jacobians are computed via automatic differentiation~\cite{adolc}.
-The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $10^{-4}$.
-\item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
+\item \textit{LSPG ROM:} The LSPG ROM is built on top of the FOM
+	discretization leveraging the CN time scheme as previously described. Unless
+		noted otherwise, the LSPG ROM uses a constant time step of $\Delta t =
+		0.002$. The nonlinear least-squares problem arising at each time instance
+		is solved via the Gauss--Newton method. The linear least-squares problems
+		arising at each Gauss--Newton iteration are solved via the normal
+		equations. All Jacobians are computed via automatic
+		differentiation~\cite{adolc}. The Gauss--Newton algorithm is deemed
+		converged when the gradient norm is less than $10^{-4}$. 
+\item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the
+	direct and indirect methods with two different time-stepping schemes:
+		\KTC{are these time stepping schemes? earlier we referred to them as
+		solution techniques}
 \begin{itemize}
-\item Direct method: We consider \methodAcronymROMs\ solved via the direct method for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time step of $\Delta t = 0.0005$.
+	\item \textit{Direct method}: We consider \methodAcronymROMs\ solved via the direct method for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time step of $\Delta t = 0.0005$.
 The nonlinear least-squares problem
-arising over each window is solved via the Gauss--Newton method.
-The linear least-squares problems arising at each Gauss--Newton iteration is solved via the normal equations.
-All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $10^{-4}$ (i.e., we use the parameter $\epsilon = 10^{-4}$ in Algorithm~\ref{alg:colloc_gn}).
-Critically, it is noted that the Jacobian
+arising over each window is solved via the Gauss--Newton method. The linear
+		least-squares problems arising at each Gauss--Newton iteration is solved
+		via the normal equations. All Jacobians are again computed via automatic
+		differentiation, and the Gauss--Newton algorithm is deemed converged when
+		the gradient norm is less than $10^{-4}$ (i.e., we use the parameter $\epsilon = 10^{-4}$ in Algorithm~\ref{alg:colloc_gn}). Critically, it is noted that the Jacobian 
 matrix over a window is constructed by assembling local (dense) Jacobians.
@@
 
-\item Indirect method: We consider two \methodAcronymROMs\ solved via the indirect method.
+\item \textit{Indirect method}: We consider two \methodAcronymROMs\ solved via the indirect method.
 The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$.
@@
 \end{subfigure}
-\caption{$x-t$ diagrams for the density fields as predicted by the FOM, G-ROM, LSPG-ROM, and \methodAcronymROMs.}
+\caption{$x-t$ diagrams for the density fields as predicted by the FOM, G-ROM,
+	LSPG-ROM, and \methodAcronymROMs. \KTC{Xlabel should be `spatial variable
+	$\mathcal x$. Ylabel should be `time $t$'. I wouldn't call this an $x-t$
+	diagram; I'd call is a density-solution plot}}
 \label{fig:sod_xt}
@@
 that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the space--time $\elltwo$  
-error. Increasing the window size does lead to a monotonic decrease in the space--time residual, however. We additionally note that, although the space--time
+error. As expected, increasing the window size does lead to a monotonic decrease in the space--time residual, however. We additionally note that, although the space--time
 $\elltwo$ error of 

### 2019-10-23 cavity.tex
active voice

@@
 \end{equation}
-where $\mathbf{u}: [0,T] \times \Omega \rightarrow \RR{4}$ comprise the density, x and y momentum, and total energy.
+where $\mathbf{u}: [0,T] \times \Omega \rightarrow \RR{4}$ comprise the density, $\mathsf{x}_1$ and $\mathsf{x}_2$ momentum, and total energy.
 The terms $\mathbf{F}$ and $ \mathbf{F}_v$ are the inviscid and viscous fluxes, respectively.
@@
 \end{equation*}
-The Navier--Stokes equations are closed with a constitutive relationship for a calorically perfect gas,
+We close the Navier--Stokes equations with a constitutive relationship for a calorically perfect gas,
 $$p = (\gamma - 1)( \rho E - \frac{1}{2} \rho u_1^2 - \frac{1}{2} \rho u_2^2 \big),$$
@@
 
-Figure~\ref{fig:cav_fig} depicts the domain $\Omega$ and flow conditions. Free-stream boundary conditions 
-are used at the inlet, outlet, and top wall of the cavity. No-slip boundary conditions are enforced 
+Figure~\ref{fig:cav_fig} depicts the domain $\Omega$ and flow conditions.
+We note that the Reynolds number is defined as $\text{Re} = \rho \norm{\mathbf{v}_{\infty}} L / \mu$ with a characteristic length set at $L=1$ and $\mathbf{v} = [u_1 \; \; u_2]^T$, the speed of sound by $a_{\infty} = \sqrt{\gamma p_{\infty}/\rho}$, and $\infty$ subscripts refer to free-stream conditions.
+We employ free-stream boundary conditions at the inlet, outlet, and top wall of the cavity.
+We enforce no-slip boundary conditions
 on the bottom wall of the cavity.
@@
 \subsubsection{Description of FOM and generation of \spatialAcronym\ trial subspace}
-The full-order model comprises a discontinuous-Galerkin discretization. The discretization 
-is obtained by partitioning the domain into $100$ elements in the flow direction and $40$ elements 
-in the wall-normal direction. The solution is represented to third order over each element using tensor product polynomials of order $p=2$, 
+The full-order model comprises a discontinuous-Galerkin discretization. We obtain the discretization 
+by partitioning the domain into $100$ elements in the flow direction and $40$ elements 
+in the wall-normal direction. The discretization represents the solution to third order over each element using tensor product polynomials of order $p=2$, 
 resulting in $36000$ unknowns for each conserved variable. Spatial discretization via the discontinuous Galerkin method yields a dynamical system 
@@
 The reduced-order models leverage POD to construct the \spatialAcronym\ trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to select the sampling points (and as a result the weighting
-matrix $\stweightingMatArg{}$); we use a constant weighting matrix across all windows, e.g., $\stweightingMatArg{n} \equiv \stweightingMatArg{}$.
+matrix $\stweightingMatArg{}$); we use a constant weighting matrix across all windows, e.g., $\stweightingMatArg{n} \equiv \stweightingMatArg{}$, $n=1,\ldots,\nslabs$.
 The process used to construct the initial conditions and trial subspaces for the ROMs is as follows:
@@
 \begin{itemize}
-\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation. 
-The Galerkin ROM is evolved in time with the CN time scheme at a time step of $\Delta t =0.1$.
+\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation and evolve the ROM in time with the CN time scheme at a time step of $\Delta t =0.1$.
 
 \item \textit{LSPG ROM with collocation}: We consider a collocated LSPG ROM, which is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
-discretization. Unless otherwise noted, the time step for the LSPG ROM is $\Delta t  = 0.1$. As in the previous example, 
-the Gauss--Newton method is used to solve the nonlinear least-squares problem arising at each time instance. The linear least-squares problems 
-arising at each Gauss--Newton iteration are solved via the normal equations. We deem the Gauss--Newton 
-algorithm converged when the gradient norm is less than $1e-4$. All Jacobians are computed via automatic differentiation.
+discretization. We employ a time step of $\Delta t  = 0.1$. The implementation is the same as previously described.
 
 \item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time step of 
-$\Delta t = 0.1$. We deem the Gauss--Newton algorithm converged when the gradient norm is less than $1e-4$. The linear least-squares problems
-arising at each Gauss--Newton iteration are solved via the normal equations. All Jacobians are computed via automatic differentiation. Uniform quadrature 
-weights are used for evaluating the integral in the minimization problem~\eqref{eq:obj_gen_slab}.
-
+$\Delta t = 0.1$. The implementation is the same as previously described. 
 %\begin{itemize}
@@
 This comprises the same time interval used to construct the trial subspace.
-First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$ state error of the various reduced-order models.
+First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$ error of the various reduced-order models.
 Both the collocated Galerkin and LSPG ROMs blow up/fail to converge within the first several time units.
 The \methodAcronymROM\ minimizing the residual over a window of size $\Delta T = 0.2$ also fails to converge.
-The \methodAcronymROMs\ that minimize the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are again less than $10\%$.
-The most notable discrepancy between the \methodAcronymROM\ and FOM solutions is a slight phase difference.
+The \methodAcronymROMs\ that minimize the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are less than $10\%$.
+The most notable discrepancy between the \methodAcronymROM\ and FOM solutions is a phase difference.
 
@@
 
-Figure~\ref{fig:cav_results2} shows the space--time error and objective function of the stable \methodAcronymROMs. It is seen that growing the window size over 
+Figure~\ref{fig:cav_results2} shows the space--time error and objective function of the stable \methodAcronymROMs. We observe that growing the window size over 
 which the residual is minimized leads to a lower space--time residual, but not necessarily a lower $\elltwo$ error.
@@
 Next, Figure~\ref{fig:cav_wallclock}
-shows the wall-clock times of the \methodAcronymROMs\ for $t \in [0,10]$ as compared to the LSPG ROM\footnote{Note that LSPG blew up at $t \approx 16.0$, so we focus on the first ten time units.}.
+shows the wall-clock times of the \methodAcronymROMs\ for $t \in [0,10]$ as compared to the LSPG ROM\footnote{We note that LSPG failed to converge at $t \approx 16.0$, so we focus on the first ten time units.}.
 As expected, increasing $\Delta T$ again leads to an increase in computational cost; minimizing the residual over a window comprising 20 time instances yields a 2.5x increase in cost over LSPG.
@@
 \methodAcronymROMs\ again yield improved performance: \methodAcronymROMs\ minimizing the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are less than $5\%$.
-The \methodAcronymROMs\ with $\Delta T \ge 0.5$ and basis \#2 yield more accurate results than \methodAcronymROMs\ with basis \#1.
+The \methodAcronymROMs\ employing basis \#2 yield more accurate results than \methodAcronymROMs\ employing basis \#1.
 
@@
 Increasing the window size again leads to an increase in computational cost.
-Minimizing the residual over a window comprising 20 time instances yields a 2.5x increase in cost over LSPG.
+Minimizing the residual over a window comprising 20 time instances yields a 3x increase in cost over LSPG.
 This increase in computational cost is particularly reasonable in this example as LSPG yielded an unstable solution.

### 2019-10-23 main.tex
active voice

@@
 		model (\spatialAcronym)}.
-These trial subspaces are characterized by a \textit{spatial projection operator}, associate with a basis that represents the spatial dependence of the state, and are employed in classic model reduction approaches, e.g., Galerkin and LSPG.
+These trial subspaces are characterized by a spatial projection operator, associate with a basis that represents the spatial dependence of the state, and are employed in classic model reduction approaches, e.g., Galerkin and LSPG.
 %These
@@
 		model (\spaceTimeAcronym)}.
-These trial subspaces are characterized by a \textit{space--time projection operator}, associate with a basis that represents the spatial and temporal dependence of the state, and are employed in space--time
+These trial subspaces are characterized by a space--time projection operator, associate with a basis that represents the spatial and temporal dependence of the state, and are employed in space--time 
 model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
@@
 
-At each time instance on the time grid, the LSPG method substitutes the \spatialAcronym\ trial subspace approximation
+At each time instance on the time grid, LSPG substitutes the \spatialAcronym\ trial subspace approximation
 \eqref{eq:affine_trialspace} into the FOM O$\Delta$E~\eqref{eq:lms} and
@@
 	\stateFOMIC\otimes\onesFunction
-	\subseteq \RR{N} \otimes \timeSpace
+	\subseteq \RR{N} \otimes \timeSpace.
 \end{equation}
 %and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
-and $\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ with $\stbasis(0) =
+Here $\stbasis \in \RR{\fomdim \times \stdim} \otimes \timeSpace$, with $\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ and $\stbasis(0) =
 \boldsymbol 0$
@@
 %\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
-\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \end{bmatrix}  \\
+\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \end{bmatrix} , \\
 & \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t},
@@
 \residLMSArg{N_t}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg)  \\
-\end{bmatrix}  \\
+\end{bmatrix} , \\
 & \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
@@
 For each type of trial
-	subspace, we then outlined two different solution techniques: direct (i.e.,
+	subspace, we outlined two different solution techniques: direct (i.e.,
 	discretize then optimize) and indirect (i.e., optimize then discretize). We showed that particular instances of the approach recover Galerkin,
 	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection.  
-
 The case of \spatialAcronym\ trial subspaces is of particular interest in the \methodAcronym\ context: it was shown that indirect methods comprise solving a coupled two-point 
@@
 The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals.
-Increasing the window size over which the residual was minimized, however, did not necessarily decrease the solution error in the $\elltwo$-norm; this was observed to occur over an intermediary window size.
+Increasing the window size over which the residual was minimized, however, did not necessarily decrease the solution error in the $\elltwo$-norm; we observed this to occur over an intermediary window size.
 We additionally observed that the
@@
 %\adjoint^n(t) \defeq \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArg{n}{t}  -  \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) ,
-\adjointArg{}{\timeDummy}  : \timeDummy \mapsto \genstateDotArg{}{\timeDummy}  -  \mass^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace \genstateArg{}{t} + \stateIntercept, \timeDummy) ,
+\adjointArgnt{}  : \timeDummy \mapsto \genstateDotArg{}{\timeDummy}  -  \mass^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace \genstateArg{}{t} + \stateIntercept, \timeDummy) ,
 \end{equation*}
@@
 \end{multline*}
-We thus get the splitting,
+We thus get the splitting
 %Eq.~\ref{eq:clspg_2ord} can be split as,
@@
 When formulated as an optimal control problem of Lagrange type, the gradients of the Hamiltonian with respect to the state, controller, and costate 
-need to be evaluated. This section details this evaluation.
+need to be evaluated. This section details this evaluation. The case derivation is presented for the case with one window, for notational simplicity.
 
@@
 \begin{align*}
-\hamiltonian^n \; &: \;  (\genstateyDiscreteArgnt{},\adjointDiscreteDumArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \mapsto 
+\hamiltonian \; &: \;  (\genstateyDiscreteArgnt{},\adjointDiscreteDumArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \mapsto 
  \adjointDiscreteDumArgnt{T} \bigg[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \bigg] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \\

### 2019-10-23 numerical_experiments.tex
active voice

@@
 \section{Numerical experiments}\label{sec:numerical_experiments}
-This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity.
-In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs.
+We now analyze the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity.
+In each experiment, we compare \methodAcronymROMs\ to the Galerkin and LSPG ROMs.
 The purpose of the numerical experiments is to assess the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy.
@@
 In both experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$.
-It is also noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs.
-Future state prediction and prediction at new parameter instances are not considered as these problems introduce factors that confound the solution accuracy with the solution methodology (e.g., accuracy of the basis).
+We also note that both experiments are designed to test the \textit{reproductive} ability of the ROMs.
+We do not consider future state prediction and prediction at new parameter instances as these problems introduce factors that confound the solution accuracy with the solution methodology (e.g., accuracy of the basis).
 
 \subsection{Sod shock tube}
-We first consider reduced-order models of the Sod shock tube problem. The system is governed by the compressible Euler equations in one dimension, 
+We first consider reduced-order models of the Sod shock tube problem, which is governed by the compressible Euler equations in one dimension, 
 \begin{equation}\label{eq:euler_1D}
@@
 \end{equation}
-where $\boldsymbol u : \Omega \times [0,T] \rightarrow \RR{3}$ comprise the density, x-momentum, and total energy, $\mathsf{x} \in \Omega \defeq  [0,1]$ is the spatial domain, and 
+where $\boldsymbol u : \Omega \times [0,T] \rightarrow \RR{3}$ comprise the density, $\mathsf{x}$-momentum, and total energy, $\mathsf{x} \in \Omega \defeq  [0,1]$ is the spatial domain, and 
 $T = 1$ the final time. 
@@
 \subsubsection{Description of FOM and generation of \spatialAcronym\ trial subspace}\label{sec:sod_fom}
-The 1D compressible Euler equations are solved using a finite volume method.
-The domain is partitioned into 500 cells of uniform width.
-The finite volume method uses the Rusanov flux~\cite{rusanov} at the cell interfaces.
-The FOM uses the Crank--Nicolson (CN) scheme, which is a linear multistep method defined by the coefficients $\alpha_0 = 1,\alpha_1 = -1, \beta_0 = \beta_1 = 1/2$, for temporal integration.
-The FOM is evolved for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$.
-The \spatialAcronym\ trial subspace is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $.
+We solve the 1D compressible Euler equations with a finite volume method.
+We partition the domain into 500 cells of uniform width and employ the Rusanov flux~\cite{rusanov} at the cell interfaces.
+We employ the Crank--Nicolson (CN) scheme, which is a linear multistep method defined by the coefficients $\alpha_0 = 1,\alpha_1 = -1, \beta_0 = \beta_1 = 1/2$, for temporal integration.
+We evolve the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$.
+We construct the \spatialAcronym\ trial subspace by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $.
 The resulting trial subspace corresponds to an energy criterion of $99.99\%$.
@@
 \begin{itemize}
-\item \textit{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
-a constant time step of $ \Delta t = 0.002$.
+\item \textit{Galerkin ROM}: We obtain the Galerkin ROM through Galerkin projection of the FOM and evolve the Galerkin ROM in time with the CN time scheme at a constant time step of $ \Delta t = 0.002$.
 
-\item \textit{LSPG ROM:} The LSPG ROM is built on top of the FOM
+\item \textit{LSPG ROM:} We construct the LSPG ROM on top of the FOM
 	discretization leveraging the CN time scheme as previously described. Unless
-		noted otherwise, the LSPG ROM uses a constant time step of $\Delta t =
-		0.002$. The nonlinear least-squares problem arising at each time instance
-		is solved via the Gauss--Newton method. The linear least-squares problems
-		arising at each Gauss--Newton iteration are solved via the normal
-		equations. All Jacobians are computed via automatic
-		differentiation~\cite{adolc}. The Gauss--Newton algorithm is deemed
-		converged when the gradient norm is less than $10^{-4}$. 
-\item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the
-	direct and indirect methods with two different time-stepping schemes:
-		\KTC{are these time stepping schemes? earlier we referred to them as
-		solution techniques}
+		noted otherwise, we employ a constant time step size of $\Delta t =
+		0.002$ for the LSPG ROM. We solve the nonlinear least-squares problem arising at each time instance
+		via the Gauss--Newton method, and solve the linear least-squares problems
+		arising at each Gauss--Newton iteration via the normal
+		equations. We deem the Gauss--Newton iteration converged when the gradient norm is less than $10^{-4}$.  We compute all Jacobians via automatic
+		differentiation~\cite{adolc}. 
+\item \textit{\methodAcronymROM:} We consider \methodAcronymROMs\ solved via the
+	direct and indirect methods with two different solution techniques:
 \begin{itemize}
 	\item \textit{Direct method}: We consider \methodAcronymROMs\ solved via the direct method for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time step of $\Delta t = 0.0005$.
-The nonlinear least-squares problem
-arising over each window is solved via the Gauss--Newton method. The linear
-		least-squares problems arising at each Gauss--Newton iteration is solved
-		via the normal equations. All Jacobians are again computed via automatic
-		differentiation, and the Gauss--Newton algorithm is deemed converged when
-		the gradient norm is less than $10^{-4}$ (i.e., we use the parameter $\epsilon = 10^{-4}$ in Algorithm~\ref{alg:colloc_gn}). Critically, it is noted that the Jacobian 
-matrix over a window is constructed by assembling local (dense) Jacobians.
-The global Jacobian over a single window is then stored in a compressed sparse row format.
-Uniform quadrature weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.
+We solve the nonlinear least-squares problem
+arising over each window with the Gauss--Newton method, and solve the linear
+		least-squares problems arising at each Gauss--Newton iteration 
+		via the normal equations. We again compute all Jacobians via automatic
+		differentiation, and deem the Gauss--Newton algorithm converged when
+		the gradient norm is less than $10^{-4}$ (i.e., we use the parameter $\epsilon = 10^{-4}$ in Algorithm~\ref{alg:colloc_gn}). Critically, we note that we assemble the (sparse) Jacobian 
+matrix over a window by computing local (dense) Jacobians.
+We store the Jacobian matrix over a window in a compressed sparse row format.
+We employ uniform quadrature weights for evaluating the integral in~\eqref{eq:obj_gen_slab}.
 
@@
 The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$.
-The coupled two-point boundary
-problem is solved via the forward--backward sweep method.
-The action of the Jacobian transpose on vectors is computed via automatic differentiation.
+We solve the coupled two-point boundary
+problem via the forward--backward sweep method, and compute the action of the Jacobian transpose on vectors via automatic differentiation.
 We use parameters $\epsilon = 10^{-6}$, $\fbsmGrowth = 1.1$, and $\fbsmDecay=2$ in Algorithm~\ref{alg:st_iter}.
@@
 We first show results for \methodAcronymROMs\ using the direct method with CN time discretization; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section.
-First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
+First, Figure~\ref{fig:sod_density} presents the density solutions produced by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
 $x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, we observe that the LSPG and Galerkin ROMs accurately characterize 
 the system: they correctly track the shock location, expansion waves, etc. We observe both predictions, however, to be highly oscillatory. These oscillations are 
-not physical and can lead to numerical blowup; e.g., due to negative pressure. We observe the \methodAcronymROMs\ to produce less oscillatory solutions than both 
+not physical and can lead to numerical instabilities; e.g., due to negative pressure. We observe the \methodAcronymROMs\ to produce less oscillatory solutions than both 
 the Galerkin and LSPG ROMs.
@@
 
-Next, Figure~\ref{fig:sod_error} shows space--time (i.e., integrated) state errors and the objective function~\eqref{eq:obj} (i.e., the space--time residual norm) over $t \in [0,1]$ for the various ROMs.
+Next, Figure~\ref{fig:sod_error} shows space--time state errors and the objective function~\eqref{eq:obj} (i.e., the space--time residual norm) over $t \in [0,1]$ for the various ROMs.
 Most notably, we observe
-that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the space--time $\elltwo$  
-error. As expected, increasing the window size does lead to a monotonic decrease in the space--time residual, however. We additionally note that, although the space--time
-$\elltwo$ error of 
+that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the space--time error as measured in the $\elltwo$-norm (we refer to this as the $\elltwo$-error).
+As expected, increasing the window size does lead to a monotonic decrease in the space--time residual, however.
+We additionally note that, although the space--time
+$\elltwo$-error of 
 the projected FOM solution is significantly lower than that of the various ROM solutions, the space--time residual norm of the projected FOM solution is \textit{higher} 
-than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$ norm, it does not satisfy the governing equations as well the ROM solutions.
+than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$-norm, it does not satisfy the governing equations as well the ROM solutions.
 
@@
 Figure~\ref{fig:sod_error_methods} 
-shows the same space--time $\elltwo$ state errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
-the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, we observe observed that the \methodAcronym\ method 
+shows the same space--time $\elltwo$-errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
+the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, we observe that the \methodAcronym\ method 
 is relatively insensitive to the solution technique (direct vs indirect) and underlying discretization scheme, although some minor differences are observed.
 In particular, ROMs using the second order explicit AB2 scheme with a time step of $\Delta t = 0.0005$ provide similar results to the 
-ROMs using the second-order CN scheme at a time step of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal space--time
-$\elltwo$ state error occurs when the window size is $\DeltaSlabArg{} = 0.1$ and the residual decreases monotonically as the window size grows. 
+ROMs using the second-order CN scheme at a time step of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal 
+$\elltwo$-error occurs when the window size is $\DeltaSlabArg{} = 0.1$, and the residual decreases monotonically as the window size grows. 
 \begin{figure}
@@
 
-Next, Figure~\ref{fig:sod_walltime} provides the CPU wall-clock times for the various \methodAcronymROMs. The computational cost of all methods is seen to grow 
+Next, Figure~\ref{fig:sod_walltime} provides the CPU wall-clock times for the various \methodAcronymROMs. We observe that the computational cost of all methods grows
 as the window size is increased. For indirect methods, this increase in cost is due to the fact that, as the window size 
@@
 associated with forming and solving the normal equations at each Gauss--Newton iteration and (2) the increased number of 
-Gauss--Newton iterations required for convergence. The cost of the FBSM method is observed to increase more rapidly 
+Gauss--Newton iterations required for convergence. We observe the cost of the FBSM method to increase more rapidly 
 than the direct methods. Encouragingly, \methodAcronymROMs\ based on the CN discretization
@@
 AB2 contains 4 times more temporal degrees of freedom than CN.  Finally, we emphasize that the 
-results presented here are presented for standard algorithms (e.g., Guass--Newton and the FBSM). As mentioned in 
-Remarks~\ref{remark:gaussnewton} and~\ref{remark:fbsm}, we expect the computational cost 
-of both the indirect and direct methods can be decreased through the use and/or development of 
+results presented here are for standard algorithms (e.g., Gauss--Newton and the FBSM). As mentioned in 
+Remarks~\ref{remark:gaussnewton} and~\ref{remark:fbsm}, we expect that the computational cost 
+of both indirect and direct methods can be decreased through the use and/or development of 
 algorithms tailored to the windowed minimization problem.
@@
 To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$.
-Figure~\ref{fig:time_step_study} shows the space--time $\elltwo$ error and residual norm
-of the various ROMs. We observe that both the space--time $\elltwo$ error and residual norm 
-of the \methodAcronymROMs\ decrease and converge as the time step decreases. This is in direct contrast to the LSPG ROMs, in where both the $\elltwo$ 
-space--time error and residual norm increase as the time step decreases. This result 
+Figure~\ref{fig:time_step_study} shows the $\elltwo$-error and residual norm
+of the various ROMs. We observe that the $\elltwo$-error and residual norm 
+of the \methodAcronymROMs\ decrease and converge as the time step decreases.
+This is in direct contrast to the LSPG ROMs, in where the $\elltwo$-error and residual norm display a complex dependence on the time step.
+This result
 demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
 shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that for all time steps considered, the \methodAcronymROMs\ 
-are less than 6x the cost of LSPG, despite the window sizes comprising up to 100 time instances. 
+are less than 6x the cost of LSPG; this is despite the window sizes comprising up to 100 time instances. 
 \begin{figure}
@@
 Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ led to less oscillatory solutions.
-\item Increasing the window size over which the residual is minimized does not necessary lead to a lower space--time state error as measured by the $\elltwo$ norm.
-We observed that minimizing the residual over an intermediary window size led to the lowest space--time state error in the $\elltwo$ norm.
-\item \methodAcronym\ displays time-step convergence: both the space--time $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time step decreased.
+\item Increasing the window size over which the residual is minimized does not necessary lead to a lower space--time error as measured by the $\elltwo$ norm.
+We observed that minimizing the residual over an intermediary window size led to the lowest space--time error in the $\elltwo$ norm.
+\item \methodAcronym\ displays time-step convergence: both the space--time $\elltwo$ error and residual norm decreased and displayed time-step convergence as the time step decreased.
 This is in contrast to LSPG.
 \item For all examples considered, in where the windows comprised up to 2000 time instances, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
-direct method was observed to be, in general, slightly more efficient and robust than the indirect method.
+direct method was observed to be slightly more efficient and robust than the indirect method.
 \end{enumerate}

### 2019-10-23 section3/general_form.tex
active voice

@@
 %$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. 
-For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
+For notational purposes, we additionally define the spatial trial subspaces at the start of each window as
 \begin{equation}
@@
 
-The \methodAcronym\ framework sequentially computes approximate solutions
+The \methodAcronym\ \approachKwd\ sequentially computes approximate solutions
 $\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 

### 2019-10-23 section4/spatial_numerical.tex
response
active voice

@@
 $\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
-$n=1,\ldots,\nslabs$, that satisfy
+$n=1,\ldots,\nslabs$ that satisfy
 %	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
@@
 	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in ( \trialspace + \stateInterceptArg{n}) \otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_{\text{D-S}} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
+\objectiveArgLMS{n}_{\text{D}} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}} ; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}),
 %%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
@@
 \end{align*}
-where
-\begin{equation*}
-\begin{split}
-\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}), \\
-&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow \RR{}_+,
-\end{split}
-\end{equation*}
-\KTC{On the RHS, shouldn't this be $\objectiveArgLMS{n}_\text{D}$? I couldn't
-find the definition of $\objectiveArgLMS{n}$ anywhere.}
+%where
+%\begin{equation*}
+%\begin{split}
+%\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_D(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}), \\
+%&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow \RR{}_+,
+%\end{split}
+%\end{equation*}
 with the initial condition $\approxstateDiscreteArg{1,0} = \basisspaceArg{n}\basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n}$.
@@
 	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
-\objectiveArgLMS{n}_{\text{D-S}} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
+\objectiveArgLMS{n}_{\text{D}} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}).
 %& \text{subject to} \hspace{0.2 in}
@@
 $\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the
-	time instances correspond to the time grid used for time discretization $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
+	time instances satisfy $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
 $\approxstateDiscreteArg{n,1}  = \approxstateLSPG^n$, $n=1,\ldots,\nslabs$ and  
@@
 \begin{align*}
- \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
+ \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T , \\
 &\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm},
@@
 \begin{equation*}
-\objectiveArgLMS{n}_{\text{D-S}} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg)
+\objectiveArgLMS{n}_{\text{D}} \bigg( \approxstateDiscreteArg{n,1} ,\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}  ; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}  \bigg)
 =
@@
 		examples presented later in this work leverage AD. The principal drawback
-		of this approach is their intrusiveness, which may prevent them from
+		of this approach is its intrusiveness, which may prevent them from
 		practical application, e.g., in legacy codes.

### 2019-10-23 numerical_experiments.tex
updated fig 6

@@
 \end{subfigure}
-\caption{$x-t$ diagrams for the density fields as predicted by the FOM, G-ROM,
-	LSPG-ROM, and \methodAcronymROMs. \KTC{Xlabel should be `spatial variable
-	$\mathcal x$. Ylabel should be `time $t$'. I wouldn't call this an $x-t$
-	diagram; I'd call is a density-solution plot}}
+\caption{$\mathsf{x}-t$ diagrams for the density fields as predicted by the FOM, G-ROM,
+	LSPG-ROM, and \methodAcronymROMs. \EP{I am going to stick with $x-t$ diagram, I feel like this is the standard terminology}}
 \label{fig:sod_xt}

### 2019-10-23 cavity.tex
updating cav figs

@@
 $$
-where $\mu \in \RRplus$ is the viscosity, $\text{Pr} = 0.72$ is the Prandtl number, and $c_p \in \RRplus$ the heat capacity ratio.
+where $\mu \in \RRplus$ is the dynamic viscosity, $\text{Pr} = 0.72$ is the Prandtl number, and $c_p \in \RRplus$ the heat capacity ratio.
 We assume a Newtonian fluid, which leads to a viscous stress tensor of the form
@@
 Figure~\ref{fig:cav_fig} depicts the domain $\Omega$ and flow conditions.
-We note that the Reynolds number is defined as $\text{Re} = \rho \norm{\mathbf{v}_{\infty}} L / \mu$ with a characteristic length set at $L=1$ and $\mathbf{v} = [u_1 \; \; u_2]^T$, the speed of sound by $a_{\infty} = \sqrt{\gamma p_{\infty}/\rho}$, and $\infty$ subscripts refer to free-stream conditions.
+The Reynolds number is defined as $\text{Re} = \rho_{\infty} \norm{\mathbf{v}_{\infty}} L / \mu$ with a characteristic length set at $L=1$ and $\mathbf{v} = [u_1 \; \; u_2]^T$, the speed of sound is defined by $a_{\infty} = \sqrt{\gamma p_{\infty}/\rho_{\infty}}$, and $\infty$ subscripts refer to free-stream conditions.
 We employ free-stream boundary conditions at the inlet, outlet, and top wall of the cavity.
@@
 
-Figure~\ref{fig:cav_snapshots} shows vorticity fields for the FOM, LSPG ROM, and \methodAcronymROM\ with $\Delta T = 0.5$ for the time instance $t = 5.0$.
+Figure~\ref{fig:cav_snapshots} shows vorticity fields for the FOM, LSPG ROM, and \methodAcronymROMs\ with $\Delta T = 0.5,1.0$ for the time instance $t = 5.0$.
 LSPG is observed to exhibit artificial oscillations; the Gauss-Newton method fails to converge at $t \approx 16.0$.
@@
 
+%\begin{figure}
+%\begin{center}
+%\begin{subfigure}[t]{0.95\textwidth}
+%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_fom_t5_basis2.pdf}
+%\caption{FOM} 
+%\end{subfigure}
+%%\begin{subfigure}[t]{0.95\textwidth}
+%%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c2_t10.pdf}
+%%\caption{\methodAcronym, $\Delta T = 0.2$} 
+%%\end{subfigure}
+%\begin{subfigure}[t]{0.95\textwidth}
+%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_lspg_t5_basis2.pdf}
+%\caption{LSPG} 
+%\end{subfigure}
+%\begin{subfigure}[t]{0.95\textwidth}
+%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c5_t5_basis2.pdf}
+%\caption{\methodAcronym, $\Delta T = 0.5$} 
+%\end{subfigure}
+%%\begin{subfigure}[t]{0.95\textwidth}
+%%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c10_t10.pdf}
+%%\caption{\methodAcronym, $\Delta T = 2.0$} 
+%%\end{subfigure}
+%\caption{Field solutions for the vorticity at $t = 5.0$}
+%\label{fig:cav_snapshots}
+%\end{center}
+%\end{figure}
+
 \begin{figure}
 \begin{center}
-\begin{subfigure}[t]{0.95\textwidth}
-\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_fom_t5_basis2.pdf}
+%\begin{subfigure}[t]{0.85\textwidth}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/fom.png}
 \caption{FOM}
 \end{subfigure}
-%\begin{subfigure}[t]{0.95\textwidth}
-%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c2_t10.pdf}
-%\caption{\methodAcronym, $\Delta T = 0.2$} 
-%\end{subfigure}
-\begin{subfigure}[t]{0.95\textwidth}
-\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_lspg_t5_basis2.pdf}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/lspg.png}
 \caption{LSPG}
 \end{subfigure}
-\begin{subfigure}[t]{0.95\textwidth}
-\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c5_t5_basis2.pdf}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/wls_window5.png}
 \caption{\methodAcronym, $\Delta T = 0.5$}
 \end{subfigure}
-%\begin{subfigure}[t]{0.95\textwidth}
-%\includegraphics[trim={0cm 3.9cm 0cm 3.9cm},clip,width=1.\linewidth]{figs/cavity/u_c10_t10.pdf}
-%\caption{\methodAcronym, $\Delta T = 2.0$} 
-%\end{subfigure}
-\caption{Field solutions for the vorticity at $t = 5.0$}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/wls_window10.png}
+\caption{\methodAcronym, $\Delta T = 1.0$}
+\end{subfigure}
+\caption{Vorticity snapshots from the FOM and ROM simulations at $t=5.0$.} 
 \label{fig:cav_snapshots}

### 2019-10-23 cavity.tex
updated figures

@@
 This comprises the same time interval used to construct the trial subspace.
-First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$ error of the various reduced-order models.
+First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$-error of the various reduced-order models.
 Both the collocated Galerkin and LSPG ROMs blow up/fail to converge within the first several time units.
@@
 \includegraphics[trim={0cm 2.8cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/error_basis2.pdf}
-\caption{Normalized $\elltwo$ error}
+\caption{Normalized $\elltwo$-error}
 \label{fig:cav_results1b}
@@
 \end{center}
-\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various collocated ROMs to the full-order model solution.}
+\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$-errors (bottom) of various collocated ROMs to the full-order model solution.}
 \label{fig:cav_results1}
@@
 Figure~\ref{fig:cav_results2} shows the space--time error and objective function of the stable \methodAcronymROMs. We observe that growing the window size over 
-which the residual is minimized leads to a lower space--time residual, but not necessarily a lower $\elltwo$ error.
+which the residual is minimized leads to a lower space--time residual, but not necessarily a lower $\elltwo$-error.
 This result is consistent with the previous numerical example.
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window_basis2.pdf}
-\caption{Space--time $\elltwo$ error}
+\caption{Space--time $\elltwo$-error}
 \label{fig:cav_results2a}
@@
 \includegraphics[trim={0cm 2.5cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/error.pdf}
-\caption{Normalized $\elltwo$ error}
+\caption{Normalized $\elltwo$-error}
 \label{fig:cav_results1b_basis1}
@@
 \end{center}
-\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$ state errors (bottom) of various collocated ROMs to the full-order model solution.}
+\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$-errors (bottom) of various collocated ROMs to the full-order model solution.}
 \label{fig:cav_results1_basis1}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window.pdf}
-\caption{Integrated $\elltwo$ error}
+\caption{Integrated $\elltwo$-error}
 \label{fig:cav_results3a}

### 2019-10-23 numerical_experiments.tex
updated figures

@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_compare.pdf}
-\caption{Space--time $\elltwo$ error of various ROMs.}
+\caption{Space--time $\elltwo$-error of various ROMs.}
 \label{fig:sod_error_a}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window.pdf}
-\caption{Integrated $\elltwo$ error of the \methodAcronymROMs.}
+\caption{Integrated $\elltwo$-error of the \methodAcronymROMs.}
 \label{fig:sod_error_methods_a}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
-\caption{Space--time $\elltwo$ error of the \methodAcronymROM.}
+\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
 \label{fig:sod_error_methods_a}
@@
 Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ led to less oscillatory solutions.
-\item Increasing the window size over which the residual is minimized does not necessary lead to a lower space--time error as measured by the $\elltwo$ norm.
-We observed that minimizing the residual over an intermediary window size led to the lowest space--time error in the $\elltwo$ norm.
-\item \methodAcronym\ displays time-step convergence: both the space--time $\elltwo$ error and residual norm decreased and displayed time-step convergence as the time step decreased.
-This is in contrast to LSPG.
+\item Increasing the window size over which the residual is minimized does not necessary lead to a lower space--time error as measured by the $\elltwo$-norm.
+We observed that minimizing the residual over an intermediary window size led to the lowest space--time error in the $\elltwo$-norm.
+\item \methodAcronym\ displays time-step convergence: both the $\elltwo$-error and residual norm decreased and displayed time-step convergence as the time step decreased. This is in contrast to LSPG.
 \item For all examples considered, in where the windows comprised up to 2000 time instances, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 

### 2019-10-23 main.tex
updated proof

@@
 		evaluated at the FOM state projected onto the trial subspace. 
-\EP{Updated this, I don't show what you said}%As a result,
+%As a result,
 %		the error bound exhibits exponential growth in the number of time windows
@@
 %\genstate(a) = \genstate_a, \qquad \bigg[ \frac{\partial \integrand}{\partial {\genstateDot}} \bigg]^T_{t=b} = \bz.
-\KTC{I removed "subject to" as this is only used in an optimization-problem
-statement. Also, I don't understand the $z(a) = z_a$ conditions, or how that
-derives obviously from above. Can you clarify? Also note the inconsistent use
-of $a$, $b$, and $0$, $T$.} \EP{Yes, sorry about the [a,b], I changed this recently and obviously missed things. As for the BC, 
-the requirement that $z(a)=z_a$ comes in first from the implicit definition of $z$, which is required to satisfy this BC. Then, in defining the variation, 
-we explicitly force the variation to be zero at the IC such that all possible functions satisfy the BC. We don't, however, make any such requirement on the end point 
-condition, $\eta(b)$ can be anything.
+%\KTC{I removed "subject to" as this is only used in an optimization-problem
+%statement. Also, I don't understand the $z(a) = z_a$ conditions, or how that
+%derives obviously from above. Can you clarify? Also note the inconsistent use
+%of $a$, $b$, and $0$, $T$.} \EP{Yes, sorry about the [a,b], I changed this recently and obviously missed things. As for the BC, 
+%the requirement that $z(a)=z_a$ comes in first from the implicit definition of $z$, which is required to satisfy this BC. Then, in defining the variation, 
+%we explicitly force the variation to be zero at the IC such that all possible functions satisfy the BC. We don't, however, make any such requirement on the end point 
+%condition, $\eta(b)$ can be anything.
 This yields the endpoint boundary condition on the gradient (i.e., second term in Eq 56.).
@@
 56 would be replaced
-with this end point condition.} 
+%with this end point condition.} 
 Equation~\eqref{eq:euler_lagrange_appendix} is known as the Euler--Lagrange

### 2019-10-24 numerical_experiments.tex
minor grammar

@@
 \caption{$\mathsf{x}-t$ diagrams for the density fields as predicted by the FOM, G-ROM,
-	LSPG-ROM, and \methodAcronymROMs. \EP{I am going to stick with $x-t$ diagram, I feel like this is the standard terminology}}
+	LSPG-ROM, and \methodAcronymROMs.}
 \label{fig:sod_xt}

### 2019-10-24 main.tex
remove old files

@@
 Residual-minimization methods in model reduction compute the solution within a
-low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods,~\cite{bochev_leastsquares,bochev_lsfem_book} which predate ROM residual-minimization approaches.}
+low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods~\cite{bochev_leastsquares,bochev_lsfem_book}.}
 Researchers have developed such residual-minimizing model-reduction methods

### 2019-10-24 main.tex
small
spelling and cleanup

@@
 It is unclear if this is due to the time-discretization error associated with enlarging the time step, or rather if it is due to the size of the window the residual is being minimized over.
-% but this also has the effect of
-%reducing the time-dicretization error of the full-order model.
 As a
@@
 The former case is of particular
-		interest in the \methodAcronym\  contex, as the stationary conditions are
+		interest in the \methodAcronym\  context, as the stationary conditions are
 		derived via the Euler--Lagrange equations and comprise a coupled two-point
@@
 \item Numerical experiments for 
-	trial subspaces associated with spatial dimenion reduction, which demonstrate two key findings:
+	trial subspaces associated with spatial dimension reduction, which demonstrate two key findings:
 \begin{itemize}
@@
 & \stspaceS \defeq \trialspace \otimes \timeSpace +
-	\stateIntercept\otimes\onesFunction\subseteq\RR{N}\otimes\timeSpace
+	\stateIntercept\otimes\onesFunction\subseteq\RR{N}\otimes\timeSpace,
 \end{split}
@@
 subspace $\trialspace$.
-%To perform dimension reduction, 
-%To perform dimension reduction, the governing equations are projected by a ``test" basis,
-%\begin{equation}\label{eq:g_truncation}
-%\testbasis^T \basisspace \frac{d \genstate}{dt} = \testbasis^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
-%\end{equation}
-%The most common choice is to set the test space and trial space to be equivalent, which results in the Galerkin reduced-order model,
-%\begin{equation}\label{eq:g_truncation}
-%\frac{d \genstate}{dt} = \basisspace^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
-%\end{equation}
 \subsubsection{Least-squares Petrov--Galerkin projection}
@@
 \end{equation}
-where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
+with the initial condition $\stateFOMDiscreteArg{0} = \stateFOMIC$. In the above, $k^n$ denotes the number of steps employed by the scheme at the $n$th
 time instance and 
@@
 that define the linear multistep method at the $n$th time instance.
-%\begin{equation}\label{eq:coarse_implicit_euler_0}
-%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
-%\end{equation}
-%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
-%\begin{equation}\label{eq:coarse_implicit_euler_0}
-%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
-%\end{equation}
-%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.
 
@@
 
-%In the case of spatial dimension-reduction-only, \methodAcronym\ displays improved performance over LSPG at the expense of an increased computational cost.
-This increase in computational cost is due to the fact that forming and/or solving the nonlinear minimization problem becomes more difficult as the time-slab size grows.
-To minimize this increase in cost, we intend to pursue the development of new solution techniques tailored to \methodAcronym.
-
-The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost.
+In the case of \spatialAcronym\ trial subspaces, \methodAcronym\ displays improved performance over Galerkin and LSPG projection at the expense of an increased computational cost.
+This increase in computational cost is due to the fact that forming and/or solving the nonlinear minimization problem becomes more difficult as the time window grows.
+To minimize this increase in cost, future work will target the development of new solution techniques tailored to \methodAcronym.
+
+%The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost.
 In the context of the direct solution approach, this increased cost is due to the increased cost of forming and solving the least-squares problem associated with larger window sizes.
@@
 \section{Acknowledgments}
-The authors thank Yukiko Shimizo and Patrick Blonigan for numerous conversations from which this work benefited. 
+The authors thank Yukiko Shimizu and Patrick Blonigan for numerous conversations from which this work benefited. 
 E.\ Parish acknowledges an appointment to the Sandia National Laboratories'
@@
 
-\section*{References}
 \bibliographystyle{siam}

### 2019-10-24 section3/space_time.tex
spelling and cleanup

@@
 \subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
-%\KTC{Error in first equation. This is the approximate solution not the real
-%one. Use parallel language to hwo this was done in the last section for
-%S-reduction trial subspaces. Also make sure the basis is zero at the beginning
-%of the interval and the affine transformation is taken to be the solution at
-%the end of the last window. This will not be too trivial, as the solution on
-%the previous slab enters the trial subspace definition.}
-
-%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
 The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
@@
 where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
-\begin{comment}
-each window, we introduce the space--time trial subspace
-$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
-along with the space--time trial basis matrix \textit{function}, 
-\begin{align*}
-\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
- &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
-\end{align*}
-\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
-be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window.
-For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
-\KTC{Remove all commas before equations unless it makes sense gramatically.
-Add the $n$ superscript to $x_{ref}$}
-The state over each window is approximated by,
-\begin{equation}\label{eq:stapprox}
- \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
-\end{equation}
-%and the state is then approximated by,
-%\begin{equation}\label{eq:stapprox}
-% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
-%\end{equation}
-%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.
-
-\end{comment}
 Setting $\stspace^n \leftarrow \stspaceSTArg{n}$ in the

### 2019-10-24 section4/numerical_methods.tex
spelling and cleanup

@@
 and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
-%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
-%Given the formulation of the \methodAcronym\ method presented in Section
-%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
-%Due to the commonalities that the \methodAcronym\ method exhibits with
-%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
-%state this.} the derivation of these techniques can leverage a large existing
-%body of work. In this section, we outline a variety of such techniques. 
 \subsection{\spatialAcronym: direct and indirect methods}
@@
 subspaces. In particular, we consider:
-\begin{itemize} \item \textit{Direct methods (discretize-then-optimize)}: A
+\begin{itemize} \item \textit{Direct methods (discretize then optimize)}: A
 			direct approach that leverages linear multistep methods to directly solve the optimization problem~\eqref{eq:obj_gen_slab}.
 Section~\ref{sec:direct} outlines this approach.
-\item \textit{Indirect methods (optimize-then-discretize)}: An indirect
+\item \textit{Indirect methods (optimize then discretize)}: An indirect
 	approach that leverages the forward--backward sweep algorithm to solve the

### 2019-10-24 section4/spatial_numerical.tex
spelling and cleanup

@@
 	are several such examples. Further, the space--time formulation introduces
-	an extra dimension for parallelization that can be exploited to furture
+	an extra dimension for parallelization that can be exploited to future
 	reduce the wall time. The investigation of these additional, potentially
@@
 \end{remark}
-%\subsubsection{Jacobian-Free Implementation}
-%The dominant cost of the Gauss Newton algorithm is the computatation of the action of the Jacobian matrix on the trial basis. To alleviate this burden, the following Jacobian-free method is proposed:
 \begin{algorithm}
@@
  Compute  $\Delta  \genstatecollocMatSlabArg{n} $ that minimizes $\norm{ \mathbf{J} \Delta \genstatecollocMatSlabArg{n} +\mathbf{r}}^2$ \Comment{Solve the linear least-squares problem} \\
-% Compute  $\Delta  \genstatecollocMatSlabArg{n} $ satisfying $ [\mathbf{J}]^T \mathbf{J} \Delta \genstatecollocMatSlabArg{n}=  -\mathbf{J}^T\mathbf{r}$ \Comment{Solve the normal equations} \\
 $\alpha \leftarrow \text{linesearch}(\Delta  \genstatecollocMatSlabArg{n} ,\genstatecollocMatSlabArg{n}_k ) $ \Comment{Compute $\alpha$ based on a line search, or set to 1}\\
@@
 The second ingredient, namely devising an efficient method for computing terms
-involving the transpose of the Jacobain of the velocity,
+involving the transpose of the Jacobian of the velocity,
 can be challenging if one

### 2019-10-24 main.tex
minor language changes

@@
 
-In the case of \spatialAcronym\ trial subspaces, \methodAcronym\ displays improved performance over Galerkin and LSPG projection at the expense of an increased computational cost.
+%In the case of \spatialAcronym\ trial subspaces, \methodAcronym\ displays improved performance over Galerkin and LSPG projection at the expense of an increased computational cost.
 This increase in computational cost is due to the fact that forming and/or solving the nonlinear minimization problem becomes more difficult as the time window grows.
@@
 
-%The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost.
-In the context of the direct solution approach, this increased cost is due to the increased cost of forming and solving the least-squares problem associated with larger window sizes.
+The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost.
+In the context of the direct solution approach, this increased cost is due to the increased expense of forming and solving the least-squares problem associated with larger window sizes.
 In the context of indirect methods, this increased cost is a result of the increased number of iterations of the forward backward sweep method.
-While numerical experiments demonstrated that the increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.
+While numerical experiments demonstrated that the increase in computational cost is mild, future work will target the development of new solution techniques tailored for the \methodAcronym\ approach.
 

### 2019-11-04 pres/pres.tex
presentation

--- 2019-10-24 main.tex
@@
 \documentclass[3p,computermodern,10pt]{elsarticle}
-\input{packages}
-\input{commands}
-\input{commands_cwst}
+\input{../packages}
+\input{../commands}
+\input{../commands_cwst}
 %\input{title}
@@
 
-Simulating parameterized dynamical systems arises in many applications across
-science and engineering. In many contexts, executing a dynamical-system
-simulation at a single parameter instance---which entails the numerical
-integration of a system of ordinary differential equations (ODEs)---incurs an
-extremely large computational cost.  This occurs, for example, when the
-state-space dimension is large (e.g., due to fine spatial resolution when
-discretizing a partial differential equation) and/or when the number of time
-instances is large (e.g., due to time-step limitations incurred by stability
-or accuracy considerations).  When the application is time critical or many
-query in nature, analysts must replace such large-scale parameterized
-dynamical-system models (which we refer to as the full-order model) with a low-cost approximation that makes the
-application tractable.
+$$\state \approx \approxstate \in \stspace \subseteq \RR{N} \otimes \timeSpaceArg{} , \qquad  n = 1,\ldots,\nslabs$$
 
-Projection-based reduced-order models (ROMs) comprise one such approximation
-strategy. First, these techniques execute a computationally expensive
-\textit{offline} stage that computes a low-dimensional \textit{trial subspace} on
-which the dynamical-system state can be well approximated (e.g., by computing
-state ``snapshots'' at different time and parameter instances, by solving
-Lyapunov equations). Second, these methods execute an inexpensive
-\textit{online} stage during which they compute approximations to the
-dynamical-system trajectory that reside on this trial subspace via, e.g., 
-projection of the full-order model or residual minimization.
-
-Model reduction for linear-time-invariant systems (and other well structured
-dynamical systems) is quite mature
-\cite{wilcox_benner_rev,moore,roberts,GugercinIRKA}, as system-theoretic properties (e.g., controllability,
-observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
-quantified and accounted for; this often results in
-reduced-order models that inherit such important properties.  The primary
-challenge in
-developing reduced-order models for general nonlinear dynamical systems is
-that such properties are difficult to assess quantitatively. As a result, it
-is challenging to develop reduced-order models that preserve important
-dynamical-system properties, which often results in methods that yield
-trajectories that are
-inaccurate, unstable, or violate physical properties.  To address this, researchers have pursued several directions that
-aim to imbue reduced-order models for nonlinear
-dynamical systems with properties that can improve robustness and accuracy.
-These
-efforts include residual-minimization approaches that equip the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
-space--time approaches that lead to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
-``energy-based'' inner
-products that ensure non-increasing entropy in the ROM
-solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014,chan2019entropy};
-basis-adaptation methods that improve the ROM's accuracy \textit{a
-posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer,etter2019online}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
-methods that enforce conservation \cite{carlberg_conservative_rom} or
-(port-)Hamiltonian/Lagrangian structure
-\cite{LALL2003304,carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional} in the ROM; and subgrid-scale
-modeling methods that aim to improve accuracy by addressing the closure
-problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
-We note that these techniques are often not mutually exclusive.
-%Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
-Residual-minimizing and space--time approaches are the most relevant classes of
-methods for the current work and comprise the focus of the following review.
-
-Residual-minimization methods in model reduction compute the solution within a
-low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods~\cite{bochev_leastsquares,bochev_lsfem_book}.}
-Researchers have developed such residual-minimizing model-reduction methods
-for both static systems (i.e., systems without time-dependence)
-\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
-and dynamical
-systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
-In the latter category,
-Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
-formulated the residual minimization problem for dynamical systems by
-sequentially minimizing the \textit{time-discrete} full-order-model residual
-(i.e., the residual arising after applying time discretization) at each time
-instance on the time-discretization grid. This formulation is often referred
-to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
-Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
-formulation and examined its connections with Galerkin projection. Critically,
-this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
-Petrov--Galerkin projection applied to the time-continuous
-full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
-in the limit as the time step goes to zero (i.e., Galerkin projection 
-minimizes the time-instantaneous full-order-model residual). Numerous
-numerical experiments have demonstrated that LSPG often yields more
-accurate and stable solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
-The common intuitive explanation for this improved performance is that, by
-minimizing the full-order-model residual over a finite time window (rather
-than time instantaneously), LSPG computes solutions that are more
-accurate over a larger part of the trajectory as compared to Galerkin.
-
-However, LSPG has several notable shortcomings. First, LSPG
-exhibits a complex dependence on the time discretization. In
-particular, changing
-the time step ($\Delta t$) modifies both the
-time window over which LSPG minimizes the residual as well as the
-time-discretization error of the full-order model on which LSPG is based. 
-As LSPG and Galerkin projection are equivalent in the
-limit of $\Delta t \rightarrow 0$, the accuracy of LSPG approaches the (sometimes poor)
-accuracy of Galerkin as the time step shrinks.
-For too-large a time step the accuracy of LSPG also degrades.
-It is unclear if this is due to the time-discretization error associated with enlarging the time step, or rather if it is due to the size of the window the residual is being minimized over.
-As a
-consequence, LSPG often yields the smallest error for an intermediate value of the
-time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}); there
-is no known way to compute this optimal time step \textit{a priori}.
-Second, as the LSPG approach performs sequential residual minimization in
-time, its \textit{a posteriori} error bounds grow exponentially in time
-\cite{carlberg_lspg_v_galerkin}, and it is not equipped with any notion of
-optimality over the entire time domain of interest. As a result, LSPG is not
-equipped with \textit{a priori} guarantees of accuracy or stability, even for
-linear time-invariant systems~\cite{bui_thesis}.
-
-Researchers have pursued the development of space--time
-residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
-address the issues incurred by sequential residual minimization in time.
-Existing space--time approaches differ from the classic LSPG and Galerkin approaches in (1) the definition of the space--time trial subspace and (2) the definition of the residual minimization problem.
-First, space--time approaches leverage a space--time trial basis that characterizes both the spatial \textit{and} temporal dependence of the state, classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage only a spatial trial basis that characterizes the spatial dependence of the state.
-%First, space--time approaches leverage a space--time trial 
-%subspace that associates with a space--time projection operator; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with a spatial projection operator only.
-Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time step (LSPG).
-The result of these differences is that space--time approaches yield a system of algebraic equations defined
-over all space and time, whose solution comprises a vector of
-(space--time) generalized coordinates; on the other hand, spatial-projection-only approaches 
-generally associate with systems of ODEs whose solutions
-comprise \textit{time-dependent} vectors of (spatial) generalized
-coordinates.
-
-%Space--time residual-minimization approaches compute the space--time
-%trajectory within a low-dimensional space--time trial trial space that associates with space--time projection (spanned by
-%a small number of space--time basis vectors) that minimizes the full-order-model
-%residual over the entire time domain.
-%We note that methods leveraging a trial subspace that associates with space--time 
-%projection yields a fundamentally different model than methods that leverage a trial subspace that 
-%associates with spatial projection only; in particular space--time
-%projection associates with an algebraic system of equations defined
-%over all space and time, whose solution comprises a \text{fixed} vector of
-%(space--time) generalized coordinates; on the other hand, spatial projection only  
-%generally associates with a system of ODEs whose solution
-%comprises a \textit{time-dependent} vector of (spatial) generalized
-%coordinates.
-%We note that space--time residual minimization yields a fundamentlaly
-%different model than sequential residual minimization in time. In particular, space--time
-%residual minimization associates with an algebraic system of equations defined
-%over all space and time, whose solution comprises a \text{fixed} vector of
-%(space--time) generalized coordinates; on the other hand, sequential residual
-%minimization in time (often) associates with a system of ODEs whose solution
-%comprises a \textit{time-dependent} vector of (spatial) generalized
-%coordinates.
-
-Space--time residual minimization approaches minimize the FOM residual over all of space and time and, as a result, yield models that are equipped with a notion of space--time optimality and \textit{a priori} error bounds that grow more slowly
-in time. Further, space--time approaches reduce both the spatial and temporal dimensions of the full-order model, and thus promise cost savings over spatial-projection-only approaches.
-% and the resulting \textit{trajectory} exhibits a notion of
-%optimality; this implies that the solution exhibits a notion of optimality
-%over the entire time domain. 
-%\EP{Come back to this, this also has to do with residual minimization}
-However, space--time techniques also suffer from several
-limitations. First, the computational cost of solving the algebraic system arising from space--time
-approaches scales cubically with the number of space--time degrees of freedom;
-in contrast, the computational cost incurred by standard
-spatial-projection-based ROMs is linear in the number of temporal degrees of
-freedom, as the attendant solvers can leverage the lower-block triangular
-structure of the system arising from the sequential nature of time
-evolution. As a result, solving the algebraic systems arising from space--time
-projection is generally intractable without applying hyper-reduction in time
-\cite{choi_stlspg,constantine_strom}. Second, space--time residual
-minimization precludes future
-state prediction, as these methods employ space--time basis vectors defined over the
-entire time domain of interest, which must have been included in the training
-simulations.
-
-The objectives of this work are to overcome the shortcomings of existing
-residual-minimizing methods, and to provide a unifying framework from which
-existing methods can be assessed. In essence, the proposed \textit{\methodNameLower\
-(\methodAcronym)} \approachKwd\ sequentially minimizes the FOM
-residual over a sequence of arbitrarily defined
-time windows. The method is characterized by three notable aspects.  First, the method
-minimizes the \textit{time-continuous}  residual (i.e., that associated with
-the full-order model ODE). By adopting a time-continuous viewpoint, the
-formulation decouples the underlying temporal discretization scheme from the
-residual-minimization problem, thus addressing a key deficiency of both LSPG
-and space--time LSPG. Critically, time-continuous residual minimization also exposes
-two different solution methods:  a \textit{discretize-then-optimize}
-(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
-indirect) method. Second, the method sequentially minimizes
-the residual over arbitrarily defined \textit{time windows} rather than
-sequentially minimizing the residual over time steps (as in LSPG)
-or over the entire time domain (as in space--time residual-minimization
-methods). This equips the method with additionally flexibility that enables it
-to explore more fine-grained tradeoffs between computational cost and error.
-Finally, \methodAcronym\ is formulated for two kinds of space--time trial
-subspaces: one that associates with spatial dimension reduction (as employed by traditional spatial-projection-based methods), and one that associates with space--time dimension reduction (as employed by space--time methods).
-The above attributes allow the \methodAcronym\ approach to be viewed as a
-generalization of existing model-reduction methods, as 
-Galerkin, LSPG, and space--time LSPG projection correspond to specific
-instances of the formulation.
-Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
-provides a unifying framework from which these existing approaches can be derived.
-
-The \methodAcronym\ approach can be viewed as a hybrid space--time method and displays 
-commonalities with several related efforts. 
-First, Ref.~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
-and connected this formulation with optimal control; specifically it mentioned the
-optimize-then-discretize vs.\ discretize-then-optimize approaches. This work
-did not fully develop this approach, eschewing it 
-for 
-sequential residual minimization in time (i.e., LSPG). The present work thus
-formally develops and extends several of the concepts put forth
-in Ref.~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
-space--time residual minimization formulation for model interpolation.  The
-present work distinguishes itself from Ref.~\cite{constantine_strom} in that
-(1) this work also considers trial subspaces characterized by spatial dimension reduction only, and (2) we
-minimize the time-continuous FOM residual over arbitrary time
-windows. We note that,
-similar to the current work, Ref.~\cite{constantine_strom} employs
-minimization of the time-continuous FOM residual as a starting point; as such,
-this work shares some thematic similarities with
-Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
-space--time extension of LSPG projection that minimizes the time-discrete
-FOM residual over the entire time domain.  The present work distinguishes
-itself from Ref.~\cite{choi_stlspg} in that (1) this work minimizes the
-time-continuous FOM residual, (2) this work minimizes this residual over
-arbitrary time windows, and (3) this work also considers trial subspaces
-associated with spatial dimension reduction only.
-
-\begin{figure} 
-\begin{centering} 
-\includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
-\caption{Relationship diagram for the \methodAcronym\ approach for model
-	reduction.} 
-\label{fig:flowchart} 
-\end{centering} 
-\end{figure}
-
- In summary, specific contributions of this work include:
-\begin{enumerate}
-%\item The development of a time-continuous residual minimization framework for reduced-order models.
-\item The \methodNameLower\ (\methodAcronym) approach for dynamical-system
-	model reduction. The approach sequentially minimizes the time-continuous
-		full-order-model residual over arbitrary time windows.
-%\item Support of two space--time trial subspaces: one that reduces the spatial
-%	dimension of the FOM only, and one that reduces both the spatial and
-%		temporal dimensions of the FOM. 
-\item Support of two space--time trial subspaces: one that associates with spatial dimension reduction and 
-    one that associates with spatial and temporal dimension reduction.
-The former case is of particular
-		interest in the \methodAcronym\  context, as the stationary conditions are
-		derived via the Euler--Lagrange equations and comprise a coupled two-point
-		Hamiltonian boundary value problem containing a forward and backward
-		system. The forward system, which is forced by an auxiliary costate,
-		evolves the (spatial) generalized coordinates of the ROM in time. The
-		backward system, which is forced by the time-continuous FOM residual
-		evaluated about the ROM state, governs the dynamics of the costate. 
-	\item Derivation of two solution techniques:
-		\textit{discretize then optimize} (i.e., direct) and
-		\textit{optimize then discretize} (i.e., indirect). 
-	\item Remarks and derivations of conditions under which the
-		\methodAcronym\ approach recovers Galerkin, LSPG, and space--time LSPG.
-\item Error analysis of the \methodAcronym\ approach using
-	trial subspaces associated with spatial dimension reduction.
-	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
-		is bounded \textit{a priori} by a combination of the
-		error at the start of the window and the integrated FOM ODE residual
-		evaluated at the FOM state projected onto the trial subspace. 
-%As a result,
-%		the error bound exhibits exponential growth in the number of time windows
-%		(not total time instances), yielding smaller error-bound growth in thime than
-%		Galerkin or LSPG projection. \KTC{Check this}
-\item Numerical experiments for 
-	trial subspaces associated with spatial dimension reduction, which demonstrate two key findings:
-\begin{itemize}
-\item Minimizing the residual over a larger time window leads to more stable solutions 
-with lower space--time residuals norms.
-\item Minimizing the residual over a larger time window does not necessarily
-	lead to a more accurate trajectory (as measured in the space--time
-		$\elltwo$-norm of the solution). Instead, minimizing the residual over an
-		intermediate-sized time window leads to the smallest trajectory error.
-\end{itemize}
-\end{enumerate}
-
-The paper proceeds as follows: Section~\ref{sec:math}
-outlines the mathematical setting for the full-order model, along with Galerkin, LSPG, and
-space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
-\approachKwd. Section~\ref{sec:numerical_techniques}
-outlines numerical techniques for solving \methodAcronymROMs, including both direct and
-indirect methods. Section~\ref{sec:analysis} provides
-equivalence conditions and error analysis for \methodAcronymROMs.
-Section~\ref{sec:numerical_experiments} presents numerical experiments.
-Section~\ref{sec:conclude} provides conclusions and perspectives.
-We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
-x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
-matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
-\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
-calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
-function with respect to time with the $\dot\ $ operator.
-%Model reduction of nonlinear dynamical systems is a growing research area.
-%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems.
-Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step.
-LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
-%\begin{comment} 
-\section{Mathematical formulation}\label{sec:math}
-	We begin by providing the formulation for the full-order model,
-	followed by a description of standard model-reduction methods
-	classified according to the type of trial subspace they employ.
-	\subsection{Full-order model}\label{sec;FOM}
-We consider the full-order model to be a dynamical system expressed as a
-	system of ordinary differential equations (ODEs)
-\begin{equation}\label{eq:FOM}
- \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
-	\stateFOMIC,\qquad t \in [0,T],
-\end{equation}
-where  $\stateFOM: [0,T]
-	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
-	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
-	denotes the state implicitly defined as the solution to initial value
-	problem~\eqref{eq:FOM}, 
-$T \in \RRplus$ denotes the final time, 
- $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
-	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
-	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
-	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
-	nonlinear in its first argument. For subsequent exposition, we introduce
-	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
-	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
-	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
-	\timeSpace $.  We refer to the initial value problem defined in
-	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
-	although the problem of interest described in the introduction corresponds
-	to a parameterized dynamical system, we suppress dependence of the FOM ODE
-	\eqref{eq:FOM} on such parameters for notational convenience, as this work
-	focuses on devising a model-reduction approach applicable to a specific parameter instance.
-
-Directly solving the FOM ODE~\eqref{eq:FOM} is computationally expensive if
-	either the state-space dimension $\fomdim$ is large, or if the time-interval
-	length $T$ is large relative to the time step required to numerically
-	integrate Eq.~\eqref{eq:FOM}. For time-critical or many-query applications,
-	it is essential to replace the FOM ODE~\eqref{eq:FOM} with a strategy that
-	enables an approximate trajectory to be computed at lower computational
-	cost. Projection-based ROMs constitute one such promising approach.
-
-\subsection{Reduced-order models}
-Projection-based ROMs generate approximate solutions to the FOM
-	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
-	subspace. Two types of space--time trial subspaces are commonly used for
-	this purpose:\footnote{For both spatial and space--time ROMs of dynamical systems, all trial subspaces are, strictly speaking, space--time subspaces.} 
-\begin{enumerate} 
-	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
-		model (\spatialAcronym)}.
-These trial subspaces are characterized by a spatial projection operator, associate with a basis that represents the spatial dependence of the state, and are employed in classic model reduction approaches, e.g., Galerkin and LSPG.
-%These
-%	trial subspaces are commonly used
-%		in model-reduction methods based on spatial projection followed by time
-%		integration (e.g., the Galerkin and LSPG approaches).
-	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
-		model (\spaceTimeAcronym)}.
-These trial subspaces are characterized by a space--time projection operator, associate with a basis that represents the spatial and temporal dependence of the state, and are employed in space--time 
-model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
-% These trial subspaces are commonly employed in
-%		model-reduction methods based on space--time projection (e.g., space--time
-%		LSPG).
-\end{enumerate}
- We now describe these two types of space--time trial subspaces and their
-	application to the Galerkin, LSPG, and space--time LSPG approaches.
-
-\subsection{\spatialAcronym\ trial subspaces}
-At a given time instance 
-$t\in[0,T]$,
-\spatialAcronym\ trial subspaces approximate the FOM ODE solution
-	as $\approxstate(t)\approx\stateFOM(t)$, which is enforced to reside in an
-	affine spatial trial subspace of dimension $K\ll\fomdim$ such that
-	$\approxstate(t)\in
-	\stateIntercept+\trialspace
-\subseteq\RR{\fomdim}$, where $\dim(\trialspace) = K$
-and $\stateIntercept \in \mathbb{R}^{\fomdim}$ denotes the reference state, which
-	is often taken to be the initial condition (i.e., $\stateIntercept = \stateFOMIC$).
-Here, the trial subspace
-$\trialspace$ 
-is spanned by an orthogonal basis such that
-$ \trialspace= \Range{\basismat}$
-with 
-$ \basismat \equiv \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}
-	\in \RRStar{\romdim}{\fomdim}$, where $\RRStar{\romdim}{\fomdim}$ denotes the compact Stiefel manifold (i.e.,  $
-	\RRStar{\romdim}{\fomdim}\defeq
-	\{ \mathbf{X} \in \RR{\fomdim
-	\times \romdim}\, \big|\, \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$).
-The basis vectors $\basisvec_i$, $i=1,\ldots,K$ are typically constructed
-using state snapshots, e.g., via
-proper orthogonal decomposition (POD)~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}. 
-Thus, at any time instance $t\in[0,T]$, ROMs that employ the  \spatialAcronym\
-trial subspace approximate the FOM ODE solution as
-\begin{equation}\label{eq:affine_trialspace}
-\stateFOMArg{}{t} \approx \approxstate(t) = \basisspace \genstateArg{}{t} + \stateIntercept,
-\end{equation}
-where $\genstate \in \RR{\romdim} \otimes \timeSpace$ with
-$\genstate:\timeDummy\mapsto\genstate(\timeDummy)$
-denotes the generalized
-coordinates. From the space--time perspective, this is equivalent to approximating the
-	FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$ with 
-	$\approxstate\in \stspaceS$, where
-\begin{equation}\label{eq:spatial_subspace}
-\begin{split}
-& \stspaceS \defeq \trialspace \otimes \timeSpace +
-	\stateIntercept\otimes\onesFunction\subseteq\RR{N}\otimes\timeSpace,
-\end{split}
-\end{equation}
-with $\onesFunction\in\timeSpace$ defined as
-$\onesFunction:\timeDummy\mapsto 1$.
-
-Substituting the approximation~\eqref{eq:affine_trialspace} into the FOM ODE~\eqref{eq:FOM} and performing orthogonal
-$\elltwo$-projection of the initial condition onto the trial subspace yields
-the overdetermined system of ODEs
-\begin{equation}\label{eq:g_truncation}
-\basisspace \genstateDotArg{}{t} = \velocity(\basisspace
-\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC,
-	\qquad t \in [0,T],
-\end{equation}
-where $\genstateDot\equiv {d \genstate}/{d\tau}$.
-Because Eq.~\eqref{eq:g_truncation} is overdetermined, a solution may not
-exist. Typically, either \textit{Galerkin} or \textit{least-squares
-Petrov--Galerkin} projection is employed to reduce the number of equations
-such that a unique solution exists. We now describe these two methods.
-
-\subsubsection{Galerkin projection}
-The Galerkin approach reduces the number of equations in
-Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
-residual to the spatial trial subspace in the (semi-)inner product induced by the 
-positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneTArg{ }
-\stweightingMatOneArg{ }$ (commonly set to
-$\stweightingMatArg{}=\mathbf{I}$), i.e.,
-\begin{equation}\label{eq:g_truncation2}
-\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace
-\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
-\end{equation}
-where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
-denotes the $K\times K$ positive definite mass matrix.
-As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, the Galerkin approach
-can be viewed alternatively as a residual-minimization
-method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
-to
-\begin{equation}\label{eq:GalOptimal}
-\genstateGalerkinDotArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}
-	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
-	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
-	\|_{\stweightingMatArg{}}^2, \qquad \genstate(0) = \genstateIC,
-	\qquad t \in [0,T],
-%\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
-\end{equation}
-where $\|\mathbf{x}\|_{\stweightingMatArg{}}\equiv
-\sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
-Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
-FOM ODE residual evaluated at the state $\basisspace
-\genstateGalerkin(t) + \stateIntercept$ and time instance $t$ over the spatial trial
-subspace $\trialspace$.
-\subsubsection{Least-squares Petrov--Galerkin projection}
-Despite its time-instantaneous residual-minimization optimality property 
-\eqref{eq:GalOptimal}, the Galerkin approach can yield inaccurate solutions,
-particularly if the velocity is not self-adjoint or
-is nonlinear. 
-Least-squares Petrov--Galerkin (LSPG)
-~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
-was developed
-as an alternative method that exhibits several advantages over the Galerkin approach.
-Rather than minimize the (time-continuous) FOM ODE residual at a time instance (as in Galerkin), LSPG minimizes
-the (time-discrete) FOM O$\Delta$E residual 
-(i.e., the residual arising after applying time discretization to the FOM ODE)
-over a time step.
-We now describe the LSPG approach in the case of linear multistep methods;
-Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG for Runge--Kutta
-schemes.
-
-Without loss of generality, we introduce a uniform time
-grid characterized by time step $\Delta t$ and time instances
-$t^n = n\Delta
-t$, $n=0,\ldots,N_t$.
-Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
-with this time grid
-yields the FOM O$\Delta$E, which computes the sequence of discrete
-solutions
-$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
-as the implicit solution to the system of algebraic equations
-\begin{equation}\label{eq:lms}
-\residLMSArg{n}
-	(\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n})
-	= \bz,\qquad n=1,\ldots,N_t,
-\end{equation}
-with the initial condition $\stateFOMDiscreteArg{0} = \stateFOMIC$. In the above, $k^n$ denotes the number of steps employed by the scheme at the $n$th
-time instance and 
-$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual defined as
-\begin{align*}
-\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
-\\
-&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim}.
-\end{align*} 
-Here, $\alpha^n_j,\beta^n_j\in\RR{}$, $j=0,\ldots,k^n$ are coefficients
-that define the linear multistep method at the $n$th time instance.
-
-At each time instance on the time grid, LSPG substitutes the \spatialAcronym\ trial subspace approximation
-\eqref{eq:affine_trialspace} into the FOM O$\Delta$E~\eqref{eq:lms} and
-minimizes the residual, i.e., LSPG sequentially computes the solutions
-$\approxstateLSPG^n \approx \stateFOMDiscreteArg{n}$, $n=1,\ldots,
-\ntimeSteps$ that satisfy
-\begin{equation*}
-\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace + \stateIntercept }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,\ldots,\ntimeSteps,
-\end{equation*}
-where 
-$\lspgWeightingArg{\cdot} \in \RR{\nsamples \times \fomdim}$, with $\romdim
-\le \nsamples \le \fomdim$, is a weighting matrix that can be used, e.g., to
-enable hyper-reduction by requiring it to have a small number of nonzero
-columns.
-
-As described in the introduction, although numerical experiments have
-demonstrated that LSPG often yields more accurate and stable
-solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg},
-LSPG still suffers from several shortcomings. In particular, LSPG suffers from its complex
-dependence on the time discretization, exponentially growing error bounds, and
-lack of optimality for the trajectory defined over the entire time domain.
-
-\subsection{\spaceTimeAcronym\ trial spaces and space--time ROMs}
-
-Space--time projection methods that employ \spaceTimeAcronym\ trial
-spaces~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}
-aim to overcome the latter two shortcomings of LSPG. Because these methods employ \spaceTimeAcronym\ trial
-spaces, they reduce both the spatial and temporal dimensions of the full-order
-model; further, they yield error bounds that grow more slowly in time and
-their trajectories exhibit an optimality property over the entire time domain.
-
-\spaceTimeAcronym\ trial subspaces approximate the FOM ODE solution
-trajectory
-	$\stateFOM\in\RR{N}\otimes \timeSpace$ with an approximation that resides in an
-	affine space--time trial subspace of dimension $\stdim\ll\fomdim$, i.e., 
-	$\approxstate\in \stspaceST$ with $\dim(\stspaceST) =\stdim $, where
-\begin{equation}\label{eq:sttrialspace_def}
- \stspaceST \defeq 
-	\Range{\stbasis} + 
-	\stateFOMIC\otimes\onesFunction
-	\subseteq \RR{N} \otimes \timeSpace.
-\end{equation}
-%and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
-Here $\stbasis \in \RR{\fomdim \times \stdim} \otimes \timeSpace$, with $\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ and $\stbasis(0) =
-\boldsymbol 0$
-denotes the space--time trial basis. 
-Thus, at any time instance $t\in[0,T]$, ROMs that employ the
-\spaceTimeAcronym\ trial subspace approximate the FOM ODE solution as
-\begin{equation}\label{eq:stapprox1}
- \stateFOMArg{}{t} \approx \approxstateArg{}{t}  = \stbasisArg{}{t} \stgenstate + \stateFOMIC,
-\end{equation}
-where $\stgenstateArg{} \in \RR{\stdim}$ denotes the space--time generalized coordinates. 
-Critically, comparing the approximations arising from \spatialAcronym\ and
-\spaceTimeAcronym\ trial subspaces in Eqs.~\eqref{eq:affine_trialspace} and \eqref{eq:stapprox1}, respectively,
-highlights that the former approximation associates with time-dependent
-generalized coordinates, while the latter approximation associates with a
-time-dependent basis matrix.
-
-Substituting the approximation~\eqref{eq:stapprox1} into the FOM
-ODE~\eqref{eq:FOM} yields
-\begin{equation}\label{eq:st_fom}
-\stbasisDotArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate +
-	\stateFOMIC,t) , \qquad t \in [0,T],
-\end{equation}
-where 
-$\stbasisDot\equiv d\stbasis/d\timeDummy$. We note that the initial conditions
-are automatically satisfied from the definition of the \spaceTimeAcronym\
-trial subspace.
-
-Space--time methods reduce the number of equations in
-\eqref{eq:st_fom} to ensure a unique solution exists.
-We now outline one such method: space--time least-squares Petrov--Galerkin (ST-LSPG)
-~\cite{choi_stlspg}. 
-While the space--time Galerkin method~\cite{benner_st} is another alternative, it does not
-associate with any residual-minimization principle, and thus we do not discuss
-it further.
-\subsubsection{Space--time LSPG projection} 
-Analogously to LSPG, space--time LSPG~\cite{choi_stlspg}
-minimizes the (time-discrete) FOM O$\Delta$E residual, but does so
-using the \spaceTimeAcronym\ subspace and simultaneously minimizes this
-residual over all $N_t$ time instances.
-We first introduce the full space--time FOM O$\Delta$E residual for linear
-multistep methods as
-\begin{align*}
-\residST_{} & \vcentcolon (\stateyDiscreteArg{1}, \ldots
-	,\stateyDiscreteArg{N_t};\stateFOMIC) \mapsto \begin{bmatrix}
-\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
-%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
-\\ 
-%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
-\vdots \\
-%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
-\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \end{bmatrix} , \\
-& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t},
-\end{align*}
-and define the counterpart function acting on space--time generalized
-coordinates:
-\begin{align*}
-\genresidST_{\text{}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
-%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
-\residLMSArg{1}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg)
-\\ 
-\vdots \\
-\residLMSArg{N_t}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg)  \\
-\end{bmatrix} , \\
-& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
-\end{align*}
-ST-LSPG computes the space--time generalized coordinates that
-minimize the space--time FOM O$\Delta$E residual:
-\begin{equation}\label{eq:stlspg}
-	\stgenstate_\text{ST-LSPG} = \underset{ \stgenstatey\in\RR{\stdim} }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{}}(\stgenstatey;\stateFOMIC) ||_2^2, 
-\end{equation}
-where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim
-\le n_{st} \le N N_t$, is a space--time weighting matrix that can be chosen,
-e.g., to enable hyper-reduction.
-
-ST-LSPG overcomes two of the primary shortcomings of LSPG. In particular, it
-leads to error bounds that grow sub-quadratically in time rather than
-exponentially in time, and it generates entire trajectories that associate
-with an optimality property over the entire time domain \cite{choi_stlspg}.
-However, it is subject to several challenges. First, the computational cost of
-solving Eq.~\eqref{eq:stlspg} scales cubically with the number of space--time
-degrees of freedom $\stdim$. This cost is due to the fact that ST-LSPG yields 
-dense systems that do not expose any natural mechanism for exploiting
-the sequential nature of time evolution. Second, it is unclear how these
-methods can be employed for future state prediction, as the space--time trial
-basis $\stbasis$ must be defined over the entire time interval of interest
-$[0,T]$. Third, ST-LSPG is still strongly tied to the time discretization
-employed for the full-order model, as it minimizes the (time-discrete) FOM
-O$\Delta$E residual over all time instances.
-
-\subsection{Outstanding challenges}
-This work seeks to overcome the limitations of existing residual-minimizing
-model-reduction methods, and to provide a unifying framework from which
-existing methods can be assessed. Specifically, we look to overcome the
-complex dependence of LSPG on the time discretization,
-exponential time growth of the error bounds for Galerkin and LSPG,
-the cubic dependence of the computational cost of ST-LSPG on the number of
-degrees of freedom, and the lack of ability for ST-LSPG to perform prediction
-in time. We now describe the proposed \methodNameLower\ \approachKwd\ for this
-purpose. 
-%\end{comment}
-%\input{formulation_3}
-
-%\input{space_time}
-%\input{timecont_lspg}
-\input{section3/tclsrm_formulation}
-\input{section4/numerical_methods}
-\input{analysis}
-\input{numerical_experiments}
-
-\section{Conclusions}\label{sec:conclude}
-This paper proposed the windowed least-squares (\methodAcronym) \approachKwd\ for model reduction  of dynamical systems. The approach sequentially minimizes the
-	time-continuous full-order-model residual within a low-dimensional space--time trial
-	subspace over time windows.
-The approach was formulated for two types of trial subspaces: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model.
-For each type of trial
-	subspace, we outlined two different solution techniques: direct (i.e.,
-	discretize then optimize) and indirect (i.e., optimize then discretize). We showed that particular instances of the approach recover Galerkin,
-	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection.  
-The case of \spatialAcronym\ trial subspaces is of particular interest in the \methodAcronym\ context: it was shown that indirect methods comprise solving a coupled two-point 
-Hamiltonian boundary value problem.
-The forward system, which is forced by an auxiliary costate,
-		evolves the (spatial) generalized coordinates of the ROM in time. The
-		backward system, which is forced by the time-continuous FOM residual
-		evaluated about the ROM state, governs the dynamics of the costate.
-
-Numerical experiments of the compressible Euler and Navier--Stokes equations demonstrated the utility in the proposed approach.
-The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals.
-Increasing the window size over which the residual was minimized, however, did not necessarily decrease the solution error in the $\elltwo$-norm; we observed this to occur over an intermediary window size.
-We additionally observed that the
-\methodAcronym\ \approachKwd\ overcomes the time-discretization sensitivity that LSPG is subject to.
-The second numerical experiment, which examined collocated ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow.
-In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up.
-Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower error in the $\elltwo$-norm.
-
-In the case of \spatialAcronym\ trial subspaces, \methodAcronym\ displays improved performance over Galerkin and LSPG projection at the expense of an increased computational cost.
-This increase in computational cost is due to the fact that forming and/or solving the nonlinear minimization problem becomes more difficult as the time window grows.
-To minimize this increase in cost, future work will target the development of new solution techniques tailored to \methodAcronym.
-
-%The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost.
-In the context of the direct solution approach, this increased cost is due to the increased cost of forming and solving the least-squares problem associated with larger window sizes.
-In the context of indirect methods, this increased cost is a result of the increased number of iterations of the forward backward sweep method.
-While numerical experiments demonstrated that the increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.
-
-%The second subspace we consider arrises when the decoder, $\decoder$, is linear and time-varying, while the generalized coordinates are time-stationary.
-This is the case in space-time discretizations.
-In this case one has,
-%$$\state(t) \approx \approxstate(t)  =  \decoder(\genstate(t),t) = \basisst(t) \genstate + \state_0 $$
-%where $\genstate \in \mathbb{R}^{\romstdim}$ are the time-stationary (space-time) generalized coordinates and the time-continuous basis is given by $\basisst \in \mathbb{R}^{romdim \times \romstdim} \times [0,T]$.
-The Euler-Lagrange equations are given by,
-%\begin{equation}\label{eq:euler_lagrange_st}
-% \bigg[ \dot{\basisst}^T \stweightingMat^T  - \basisst^T \big[\frac{\partial \velocity}{\partial \state} \big]^T \stweightingMat^T \bigg]  \bigg( \dot{\basisst} \genstate -    \velocity \bigg) = \mathbf{0}, \qquad \state(t=0) = \state_0\end{equation}
-%Note that the end-point boundary condition in the Euler-Lagrange equations is automatically satisfied.
-
-%\subsubsection{Solution techniques}
-%Equation~\ref{eq:euler_lagrange_st} consists of a nonlinear algebraic system to be solved at each time-instance, $t$.
-To obtain solutions, we must employ a temporal discretization.
-Here, we propose using a variational discretization with tensor products, e.g.
-finite elements.
-Let,
-%$$\basisst \equiv \basisspace \otimes \basistime \subseteq \basisspace \otimes \mathcal{T},$$
-%where $\mathcal{T}$ is the finite elements space. Similarly, let
-%$$\genstate = \genstate_s \otimes \genstate_t$$
-
-\section{Acknowledgments}
-The authors thank Yukiko Shimizu and Patrick Blonigan for numerous conversations from which this work benefited. 
-E.\ Parish acknowledges an appointment to the Sandia National Laboratories'
-John von Neumann Postdoctoral Research Fellowship in Computational Science. 
-This work was partially sponsored by Sandia's Advanced Simulation and
-Computing (ASC) Verification and Validation (V\&V) Project/Task
-\#103723/05.30.02.  This paper describes objective technical results and
-analysis. Any subjective views or opinions that might be expressed in the
-paper do not necessarily represent the views of the U.S.\ Department of Energy
-or the United States Government.  Sandia National Laboratories is a
-multimission laboratory managed and operated by National Technology and
-Engineering Solutions of Sandia, LLC., a wholly owned subsidiary of Honeywell
-International, Inc., for the U.S.\  Department of Energy's National Nuclear
-Security Administration under contract DE-NA-0003525.
-
-\begin{appendices}
-\section{Proper orthogonal decomposition}
-Algorithm~\ref{alg:pod} presents the algorithm for computing the trial basis via proper orthogonal decomposition.
-\begin{algorithm}
-\caption{Algorithm for generating POD Basis.}
-\label{alg:pod}
-\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$ \; 
-%\newline
-\textbf{Output:} POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
-%\newline
-\textbf{Steps:}
-\begin{enumerate}
-    \item Solve FOM O$\Delta$E and collect solutions into snapshot matrix
-$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} - \stateInterceptArg{n} & \stateFOMDiscreteArg{N_{\text{skip}}} - \stateInterceptArg{n} & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} - \stateInterceptArg{n} \end{bmatrix}$$
-    \item Compute the thin singular value decomposition, $$\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
-    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
-    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_K \end{bmatrix}$
-\end{enumerate}
-
-\end{algorithm}
-
-\section{Selection of sampling points}
-To construct the sampling point matrix used for hyper-reduction in the second numerical experiment, we employ q-sampling~\cite{qdeim_drmac} and the sample 
-mesh concept~\cite{carlberg_gnat}. Algorithm~\ref{alg:qdeim} outlines the steps used in the second numerical experiment to compute the sampling points.
-
-\begin{algorithm}
-\caption{Algorithm for generating the sampling matrix through q-sampling.}
-\label{alg:qdeim}
-\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$ \; 
-%\newline
-\textbf{Output:} Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneTArg{} \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
-%\newline
-\textbf{Steps:}
-\begin{enumerate}
-    \item Solve FOM O$\Delta$E and collect velocity snapshots 
-$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \velocity (\stateFOMDiscreteArg{0})  & \velocity( \stateFOMDiscreteArg{N_{\text{skip}}}) & \cdots & \velocity( \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} ) \end{bmatrix}$$
-    \item Compute the thin singular value decomposition, $$\mathbf{F} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
-    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
-    \item Compute the QR factorization of  $\mathbf{U}^T$ with column pivoting,
-    \begin{equation*}
-        \mathbf{U}^T \mathbf{P}^* = \mathbf{QR}
-    \end{equation*}
-    with $\mathbf{P}^* \equiv \begin{bmatrix} \mathbf{p}_1 & \cdots & \mathbf{p}_{\text{floor}(N_t/N_{\text{skip}})} \end{bmatrix}$, $\mathbf{p}_i \in \{0,1\}^N$. 
-    \item Select the first $n_s$ columns of $\mathbf{P}^*$ to form the sampling point matrix $\stweightingMatOneTArg{} \in \{0,1\}^{N \times n_s}$. 
-    \item Augment the sampling point matrix, $\stweightingMatOneArg{}$, with additional columns such that all unknowns are computed at the mesh cells selected by Step 3.
-For the second numerical experiment, these additional unknowns correspond to each conserved variable and quadrature point in the selected cells.
-\end{enumerate}
-
-\end{algorithm}
-
-\section{Derivation of the Euler--Lagrange equations}\label{appendix:eulerlagrange}
-This section details the derivation of the Euler--Lagrange equations. To this end, we consider the generic functional of the form,
-\begin{equation}\label{eq:el_gen}
-\mathcal{J}: (\statey,\stateyDot) \mapsto  \int_a^b \integrand(\statey(t),\stateyDot(t),t)  dt,
-\end{equation}
-where $\integrand: \RR{M}  \times \RR{M}  \times[a,b] \rightarrow  \RRplus$
-	(for arbitrary $M$) with $\integrand:
-	(\stateyDiscrete,\stateyDotDiscrete,\timeDummy) \mapsto
-	\integrand(\stateyDiscrete,\stateyDotDiscrete,\timeDummy)$. We now introduce
-	the function $\statez: [a,b] \rightarrow \RR{M}$ with $\statez: \timeDummy \mapsto \statez(\timeDummy)$ along with $\statezDot \equiv d \statez /d\timeDummy$.
-We define this function to be a stationary point of~\eqref{eq:el_gen} (with $\statez$ being the first argument and $\statezDot$ being the second argument) subject to the boundary condition
-	$\statez(a) = \statez_a$. 
-We additionally introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$
-with the boundary condition $\variationArg{a} = \bz$ and define a variation
-from the stationary point by
-\begin{align*}
-\statezBar  &: (\tau,\delta) \mapsto \statez(\tau) + \delta \variation(\tau),\\
-&: [a,b] \times \RR{} \rightarrow \RR{M}.
-\end{align*}
-We note that $\statezBar$ satisfies the same boundary condition $\statez$ since $\variationArg{a} = \bz$.
-We define a new function that is equivalent to the function~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 
-$\statezDotBar \equiv d\statezBar / d\timeDummy$ in the second argument,
-$$
-\mathcal{J}_{\delta}: \delta \mapsto \int_a^b \integrand(\statezBar(t,\delta),\statezDotBar(t,\delta) ,t)  dt.
-$$
-%$$
-%\mathcal{J}(\epsilon) = \int_a^b \integrand(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t)  dt.
-%$$
-The objective is to now find $\overline{\statez}$ that makes
-$\mathcal{J}_{\delta}$ stationary. This can be done by differentiating with
-respect to $\delta$ and setting the result to zero, i.e., % Noting that the extermum of Eq.~\ref{eq:el_gen} occurs when $\state = \state^*$, and thus when $\epsilon = 0$, we have,
-\begin{equation}\label{eq:stationaryOne}
-\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)=  0.
-\end{equation}
-%or equivalently,
-%$$
-% \int_a^b \frac{d}{d \delta} \big(\integrand(\statezBar(t,\delta),\statezDotBar(t,\delta) ,t)\big)  dt = 0
-%$$
-%\int_a^b \frac{d}{d\epsilon} \bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big)\bigg] dt=  0.
-%$$
-%As $\statez$ is a stationary point of $\mathcal{J}$, it then directly follows that,
-%\begin{equation}\label{eq:el1_appendix}
-%\bigg[ \frac{d \mathcal{J}}{d \epsilon} (\statezBar(t;0),\statezDotBar(t;0) \bigg]= 0.
-%\end{equation}
-%This fact can be used to derive the Euler--Lagrange equations. Evaluating Eq.~\eqref{eq:el1_appendix},
-%$$ \int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) \big) \bigg] = 0. $$
-Using the chain rule,
-$$ 
-\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)= 
-\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t\big) \frac{\partial \statezBar }{\partial \delta}(t,\epsilon)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \frac{\partial \statezDotBar}{\partial \delta} (t,\epsilon)\bigg]dt .
-$$
-%$$\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \statezBar } \frac{\partial \statezBar }{\partial \epsilon}\big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) \big)  + \frac{\partial \integrand}{\partial \statezDotBar }\frac{\partial \statezDotBar}{\partial \epsilon} \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) \big)\bigg]dt = 0.
-$$
-%Setting $\epsilon = 0$,
-%$$\int_a^b \bigg[ \frac{\partial \integrand  }{\partial \statezBar } \frac{\partial \statezBar }{\partial \epsilon}\big(\statezBar(t,0),\statezDotBar(t,0) \big)  + \frac{\partial \integrand}{\partial \statezDotBar }\frac{\partial \statezDotBar}{\partial \epsilon} \big(\statezBar(t,0),\statezDotBar(t,0) \big)\bigg]dt = 0.
-$$
-Noting that
-$$\frac{\partial \statezBar}{\partial \delta} (t,\cdot)= \variation(t), \qquad \frac{\partial \statezDotBar }{\partial \delta}(t,\cdot)= \variationDot(t),$$
-where $\variationDot \equiv d \variation / d\timeDummy$, 
-we have
-$$
-%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon) ,t \big) \bigg] dt 
-\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-= \int_a^b \bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t)  + \frac{\partial \integrand}{\partial \stateyDotDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) ,t \big) \variationDot(t) \bigg]dt.
-$$
-%$$\int_a^b \bigg[\frac{\partial \integrand  }{\partial \overline{\genstate}} \variation + \frac{\partial \integrand}{\partial {\overline{\genstateDot}}}\variationDot \bigg]_{\epsilon = 0}dt = 0.$$
-We integrate the second term by parts,
-\begin{multline}\label{eq:gradient}
-%	\begin{split}
-%\int_a^b \frac{d}{d\epsilon}\bigg[ \integrand \big(\statezBar(t;\epsilon),\statezDotBar(t;\epsilon),t \big) \bigg] dt 
-\frac{d}{d\delta} \big( \mathcal{J}_{\delta} \big)(\epsilon)
-	=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }
-	\big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon) , t \big)
-	\variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial
-	\stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon)
-	,t \big) \bigg) \variation(t) \bigg]dt + \\ 
-\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) , 
-%	\end{split}
-\end{multline}
-where we have used $\variationArg{a} = \bz$.
-Substituting Eq.~\eqref{eq:gradient} in the stationarity condition~\eqref{eq:stationaryOne} yields
-\begin{equation}\label{eq:statOne}
-  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statezBar(t,\epsilon),\statezDotBar(t,\epsilon),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon),b \big) \variation(b)  = 0.
-\end{equation}
-By construction, $\statez(\cdot) \equiv \statezBar(\cdot;0)$ and $\statezDot(\cdot) \equiv \statezDotBar(\cdot;0)$, comprise a stationary
-point; thus, setting $\epsilon = 0$ in Eq.~\eqref{eq:statOne} yields
-\begin{equation}\label{eq:euler_lagrange_analysis}
-  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete } \big(\statez(t),\statezDot(t),t \big) \variation(t) - \frac{d}{dt}\bigg( \frac{\partial \integrand}{\partial \stateyDotDiscrete }  \big(\statez(t),\statezDot(t),t \big) \bigg) \variation(t) \bigg]dt + \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b) ,b\big) \variation(b)  = 0.
-\end{equation}
-%$$  \int_a^b \bigg[\frac{\partial \integrand  }{\partial \overline{\genstate}} \variation  - \frac{d}{dt}\frac{\partial \integrand}{\partial {\overline{\genstateDot}}}\variation \bigg]_{\epsilon = 0}dt + \bigg[\bigg[ \frac{\partial \integrand}{\partial {\overline{\genstateDot}}} \variation \bigg]_a^b\bigg]_{\epsilon = 0} = 0.
-$$
-%At $\epsilon = 0$ we have $\overline{\genstate} = \genstate$ and, as $\variationArg{a} = 0$, we obtain,
-%$$ \int_a^b \bigg(\frac{\partial \integrand  }{\partial \genstate} \variation  - \frac{\partial \integrand}{\partial {\genstateDot}}\variation \bigg)dt +  \frac{\partial \integrand}{\partial {\genstateDot}} \variation \bigg|_{t=b} = 0.$$
-As $\variation$ is an arbitrary function, this equality requires
-%\begin{equation}\label{eq:euler_lagrange_appendix}
-%\bigg[ \frac{\partial \integrand  }{\partial \genstate} \bigg]^T  - \bigg[\frac{\partial \integrand}{\partial {\genstateDot}}\bigg]^T = \boldsymbol 0, 
-%\end{equation}
-\begin{equation}\label{eq:euler_lagrange_appendix}
-\begin{split}
-\bigg[ &\frac{\partial \integrand  }{\partial \stateyDiscrete }
-	\big(\statez(t),\statezDot(t) ,t\big) \bigg]^T - \bigg[ \frac{d}{dt}\bigg(
-	\frac{\partial \integrand}{\partial \stateyDotDiscrete }
-	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz , \\
-&\statez(a) = \statez_a, \hspace{0.65 in} \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \bigg]^T = \bz.
-\end{split}
-\end{equation}
-%\genstate(a) = \genstate_a, \qquad \bigg[ \frac{\partial \integrand}{\partial {\genstateDot}} \bigg]^T_{t=b} = \bz.
-%\KTC{I removed "subject to" as this is only used in an optimization-problem
-%statement. Also, I don't understand the $z(a) = z_a$ conditions, or how that
-%derives obviously from above. Can you clarify? Also note the inconsistent use
-%of $a$, $b$, and $0$, $T$.} \EP{Yes, sorry about the [a,b], I changed this recently and obviously missed things. As for the BC, 
-%the requirement that $z(a)=z_a$ comes in first from the implicit definition of $z$, which is required to satisfy this BC. Then, in defining the variation, 
-%we explicitly force the variation to be zero at the IC such that all possible functions satisfy the BC. We don't, however, make any such requirement on the end point 
-%condition, $\eta(b)$ can be anything.
-This yields the endpoint boundary condition on the gradient (i.e., second term in Eq 56.).
-If we had a fixed final boundary condition, then the second term in Eq.
-56 would be replaced
-%with this end point condition.} 
-Equation~\eqref{eq:euler_lagrange_appendix} is known as the Euler--Lagrange
-equation. It states that, for ($\statez$,$\statezDot$) to define a stationary point of $\mathcal{J}$, then they must 
-satisfy~\eqref{eq:euler_lagrange_appendix}. It is emphasized
-that~\eqref{eq:euler_lagrange_appendix} is a necessary condition on ($\statez,\statezDot$)
-to make $\mathcal{J}$ stationary, but it is not a sufficient condition. It is
-additionally noted that~\eqref{eq:euler_lagrange_appendix} provides a
-stationary point of $\mathcal{J}$, but the resulting stationary point could be a local minima, local maxima, or saddle point.
-
-\section{Evaluation of gradients in the Euler--Lagrange equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
-We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces.
-Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.
-To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand
-\begin{align*}\label{eq:integrand_apx}
- \minintegrand & \vcentcolon
-(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,\timeDummy) \mapsto \frac{1}{2} \big[
-\basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
-+ \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[
-\basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
-\stateIntercept,\timeDummy) \big], \\ & 
-\vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
- \rightarrow \RR{} .  
-\end{align*}
-To evaluate the gradients, we first expand $\minintegrand$:
-\begin{align*}
- \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\
- &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
-\end{align*}
-Since $\stweightingMat$ is symmetric,
-\begin{equation}\label{eq:int_expand}
- \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
-\end{equation}
-For notational purposes, we write the above as,
-$$  \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy),$$
-where
-\begin{align*}
-&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
-&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) = -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
-&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) = \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
-\end{align*}
-Constructing the Euler--Lagrange equations for this functional $\minintegrand$ requires evaluating the derivatives $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot}$.
-We start by evaluating $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and go term by term.
-
-Starting with $\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot)$, we see,
-$$\frac{\partial \minintegrand_1 }{\partial \genstateyDiscrete} = \boldsymbol 0,$$
-where it is noted that $\mathcal{I}_1$ only depends on $\genstateyDiscreteDot$. Working with the second term:
-\begin{align*}
-\frac{\partial \minintegrand_2}{\partial \genstateyDiscrete}  &= -\frac{\partial}{\partial \genstateyDiscrete} \bigg( \big[ \basisspace \genstateyDiscreteDot \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] \bigg) \\
-&= - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \frac{\partial}{\partial \genstateyDiscreteArgnt{}} \bigg( \big[ \velocity(\veloargsromy) \big]  \bigg)\\
- &= -\big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}   \frac{\partial \velocity}{\partial \stateyDiscrete}\frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
-& = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace \\
-&= -[\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete}  \basisspace,
-\end{align*}
-where we have suppressed the arguments of the Jacobian for simplicity; e.g., formally 
-$$\frac{\partial \velocity}{\partial \stateyDiscrete} : (\statewDiscrete,\timeDummy) \mapsto \frac{ \partial \velocity }{ \partial \stateyDiscrete} (\statewDiscrete, \timeDummy).$$ 
-For $\minintegrand_3$,
-\begin{align}\label{eq:i3sol}
-\frac{\partial  \minintegrand_3}{\partial \genstateyDiscrete}  &= \frac{1}{2} \frac{\partial }{\partial \genstateyDiscrete} \bigg( \big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{}\big[ \velocity(\veloargsromy) \big] \bigg) \\
-&=  [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscrete}{\partial \genstateyDiscrete} \\
- &=   [\velocity(\veloargsromy) ]^T  \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
-\end{align}
-This gives the final expression,
-$$
- \frac{\partial \minintegrand}{\partial \boldsymbol \genstateyDiscrete} = - \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace +  [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
-$$%
-We now evaluate $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDotArgnt{}}$ and again go term by term. Starting with $\minintegrand_1$,
-\begin{align*}
-\frac{\partial \minintegrand_1}{\partial \genstateyDiscreteDotArgnt{}} &=
-\frac{1}{2} \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} }\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  \bigg) \\
-&= [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \basisspace.
-\end{align*}
-Now working with the second term:
-\begin{align*}
-\frac{\partial \minintegrand_2}{\partial \genstateyDiscreteDotArgnt{} } &=
-\frac{\partial}{\partial \genstateyDiscreteDotArgnt{}  } \bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \bigg)\\
- &= \frac{\partial}{\partial \genstateyDiscreteDotArgnt{} } \bigg( [\genstateyDiscreteDotArgnt{}]^T \bigg) \basisspace^T  \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] \\ 
- &= \bigg[  \basisspace^T  \stweightingMatArg{}  \velocity(\veloargsromy )   \bigg]^T \\
- &= [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace.
-\end{align*}
-Finally, for the last term,
-\begin{align*}
-\frac{\partial \minintegrand_3}{\partial \genstateyDiscreteDotArgnt{} } &= \boldsymbol 0,
-\end{align*}
-where it is noted that $\mathcal{I}_3$ only depends on $\genstateyDiscrete$. We thus have,
-$$\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot } =   [\genstateyDiscreteDotArgnt{}]^T \basisspace^T \stweightingMatArg{} \basisspace -  [\velocity(\veloargsromy ) ]^T \stweightingMatArg{} \basisspace.$$
-Combining all terms and evaluating at $(\genstateArg{}{t},\genstateDotArg{}{t},t)$,
-\begin{multline*}
- \frac{\partial \minintegrand}{\partial \genstateyDiscrete}(\genstateArg{}{t},\genstateDotArg{}{t},t)  - \frac{d}{dt} \bigg[ \frac{\partial \minintegrand}{\partial \genstateyDiscreteDotArgnt{}} (\genstateArg{}{t},\genstateDotArg{}{t},t) \bigg] =  - \big[ \basisspace \genstateDotArg{}{t}  \big]^T \stweightingMatArg{} \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \big] \basisspace + \\  [\velocity(\veloargsrom) ]^T \stweightingMatArg{} \big[ \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \big] \basisspace - \frac{d}{dt} \bigg[  [\genstateDotArg{}{t} ]^T  \basisspace^T \stweightingMatArg{} \basisspace - [\velocity(\veloargsrom) ]^T \stweightingMatArg{} \basisspace \bigg]  = \boldsymbol 0.
-\end{multline*}
-To put this in a more recognizable form, we can pull out the common factor in the first two terms,
-\begin{multline*}
-- \bigg( \big[ \basisspace \genstateDotArg{}{t} \big]^T  -  [\velocity(\veloargsrom)] ^T \bigg) \stweightingMatArg{}
- \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \basisspace -  \\
-\frac{d}{dt} \bigg[  [\genstateDotArg{}{t}]^T \basisspace^T \stweightingMatArg{} \basisspace -  
-[\velocity(\veloargsrom) ]^T \stweightingMatArg{} \basisspace  \bigg] = \boldsymbol 0.
-\end{multline*}
-Taking the transpose to put into the common column major format,
-$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom)  \big]^T \stweightingMatArg{} \bigg( \big[ \basisspace \genstateDotArg{}{t}\big]  -  \velocity(\veloargsrom)  \bigg) -  \frac{d}{dt} \bigg[  \basisspace^T \stweightingMatArg{} \basisspace \genstateDotArg{}{t}  - \basisspace^T \stweightingMatArg{} \velocity(\veloargsrom)   \bigg] = \bz.
- $$
-We now factor the second term,
-$$ -  \basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom)  \big]^T \stweightingMatArg{ } \bigg(  \basisspace \genstateDotArg{ }{t}  -  \velocity(\veloargsrom) \bigg) -  \basisspace^T \stweightingMatArg{ } \frac{d}{dt} \bigg[   \basisspace \genstateDotArg{}{t}  - \velocity(\veloargsrom) \bigg] = \bz.
-$$
-Gathering terms and multiplying by negative one, the final form of the Euler--Lagrange equations are obtained,
-\begin{equation}\label{eq:el_secondorder}
- \bigg[\basisspace^T \big[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \big]^T \stweightingMatArg{} + \basisspace^T \stweightingMatArg{} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{}{t} -  \velocity(\veloargsrom) \bigg) = \bz.
-\end{equation} 
-This is the \methodAcronym-ROM. Note that this is a second order equation and can be written as two separate first order equations. Defining the ``costate" as,
-\begin{equation*}
-%\adjoint^n(t) \defeq \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArg{n}{t}  -  \basisspace^T \stweightingMatArg{n} \velocity(\veloargsromn) ,
-\adjointArgnt{}  : \timeDummy \mapsto \genstateDotArg{}{\timeDummy}  -  \mass^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace \genstateArg{}{t} + \stateIntercept, \timeDummy) ,
-\end{equation*}
-we can manipulate Eq.~\eqref{eq:el_secondorder} as follows: %First, to enable hyper-reduction later on, we leverage the fact that $\stweightingMat^2 = \stweightingMat$,
-%$$
-% \bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn)\bigg]^T \stweightingMat^n \stweightingMat^n + \basisspace^T \stweightingMat^n \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}^n   -  \velocity(\veloargsromn) \bigg) = \bz.
-%$$
-First, we add and subtract the first term multiplied by $\basisspace [\massArgnt{n}]^{-1}\basisspace^T \stweightingMat$, 
-\begin{multline*} 
-\bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1} \basisspace^T\stweightingMatArg{} + \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  + \basisspace^T \stweightingMatArg{}  \frac{d}{dt} \bigg] \\
-\bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg) = \bz.
-\end{multline*}
-Pulling out the term multiplied by the positive portion of $\basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}$,
-\begin{multline*} 
-\bigg[\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1}  \basisspace^T  \stweightingMatArg{} \bigg)  + \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  \basisspace^T \stweightingMatArg{} +   \basisspace^T \stweightingMatArg{} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg) = \bz.
-\end{multline*}
-Splitting into two separate terms, 
-\begin{multline*}
-\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1}  \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg)  + \\
-\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} +   \basisspace^T \stweightingMatArg{} \frac{d}{dt} \bigg] \bigg(  \basisspace \genstateDotArg{}{t}  -  \velocity(\veloargsrom) \bigg) = \bz.
-\end{multline*}
-Pulling $\mass^{-1} \basisspace^T \stweightingMatArg{}$ inside the parenthesis on the second term,
-\begin{multline*}
-\basisspace^T \bigg[\frac{\partial \velocity}{\partial \state^n}(\veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg)  + \\
-\bigg[ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace  +   \mass \frac{d}{dt} \bigg] \bigg( \genstateDotArg{}{t}  - \mass^{-1} \basisspace^T \stweightingMatArg{}  \velocity(\veloargsrom) \bigg) = \bz.
-\end{multline*}
-By definition, the term inside the parenthesis of the second term is $\adjointArg{}{t}$,
-\begin{multline*}
-\basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}    -  \velocity(\veloargsrom) \bigg)  +   \\ \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace \adjointArg{}{t}  +  \massArgnt{} \frac{d}{dt} \adjointArg{}{t}= \bz.
-\end{multline*}
-Re-arranging,
-\begin{multline*}
-\mass \frac{d }{dt}\adjointArg{}{t} + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} ( \veloargsrom ) \bigg]^T \stweightingMatArg{} \basisspace  \adjointArg{}{t}  \\
-= - \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom)\bigg]^T \stweightingMatArg{}\bigg( \mathbf{I} - \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \bigg)  \bigg(  \basisspace \genstateDotArg{}{t}   -  \velocity(\veloargsrom) \bigg).
-\end{multline*}
-We thus get the splitting
-%Eq.~\ref{eq:clspg_2ord} can be split as,
-%\begin{equation}
-\begin{align*}\label{eq:lspg_continuous_appendix}
-& \mass \frac{d}{dt}\genstateArg{}{t}  -  \basisspace^T \stweightingMatArg{} \velocity(\veloargsrom) =  \mass \adjointArg{}{t},\\
-&\mass \frac{d}{dt} \adjointArg{}{t}  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsrom) \bigg]^T \stweightingMatArg{} \basisspace \adjointArg{}{t} = \\
-& -\basisspace^T \big[ \frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsrom) \big] ^T \stweightingMatArg{}  \bigg( \mathbf{I} -   \basisspace [\massArgnt{}]^{-1} \basisspace^T   \stweightingMatArg{} \bigg)\bigg( \basisspace \genstateDotArg{}{t}   -   \velocity(\veloargsrom) \bigg) .
-\end{align*}
-\section{Evaluation of gradients for optimal control formulation}\label{appendix:optimal_control}
-When formulated as an optimal control problem of Lagrange type, the gradients of the Hamiltonian with respect to the state, controller, and costate 
-need to be evaluated. This section details this evaluation. The case derivation is presented for the case with one window, for notational simplicity.
-
-The Pontryagin Maximum Principle leverages the following Hamiltonian,
-\begin{align*}
-\hamiltonian \; &: \;  (\genstateyDiscreteArgnt{},\adjointDiscreteDumArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \mapsto 
- \adjointDiscreteDumArgnt{T} \bigg[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \bigg] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \\
-&: \; \RR{\romdim} \times \RR{\romdim} \times \RR{\romdim} \times [0,T] \rightarrow \RR{},
-\end{align*} 
-where,
-\begin{align*}
- \objectiveControlArg{} &:  (\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)
-\mapsto \frac{1}{2} \bigg[ \basisspace \bigg(  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \bigg) -
-\velocity(\veloargsromy) \bigg]^T
-\stweightingMatArg{}  \\ & \hspace{1.5 in}\bigg[ \basisspace \bigg(
-[\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{}
-\bigg) - \velocity( \veloargsromy ) \bigg]
-, \nonumber \\ & : \RR{\romdim} \times \RR{\romdim} \times [0,T] \rightarrow \RR{}.
-\end{align*} 
-As described in Section~\ref{sec:optimal_control}, to derive the stationary conditions of the \methodAcronym\ objective function, we require evaluating the following gradients,
-$ \frac{\partial \hamiltonianArg{}}{\partial \adjointDiscreteDumArgnt{}{}}, \;   \frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}}$, and $\frac{\partial \hamiltonianArg{}}{\partial \controllerDiscreteDumArgnt{}{}}.$  Starting with $\frac{\partial \hamiltonianArg{}}{\partial \adjointDiscreteDumArgnt{}{}}$, we have,
-$$
-\bigg[\frac{\partial \hamiltonianArg{}}{\partial \adjointDiscreteDumArgnt{}{}}\bigg]^T = [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} .
-$$
-Next, we address $ \frac{\partial \hamiltonianArg{}}{\partial \genstateyArgnt{}{}}$. 
-\begin{align*}
-  \frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}} &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
- &= \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy
-) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
- &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}} \bigg( \velocity(\veloargsromy)\bigg) \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
- &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete} \frac{\partial \stateyDiscreteArgnt{}{}}{\partial  \genstateyDiscreteArgnt{}{} } \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
- &=   \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big] +  \frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg].
-\end{align*}
-To evaluate $\frac{\partial}{\partial \genstateyDiscreteArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg]$, we leverage the previous result~\eqref{eq:int_expand} and insert $\genstateyDiscreteDotArgnt{} =  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t} .$ This leads to the expression for the expanded Lagrangian,
-\begin{multline*}
- \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) = \frac{1}{2}\big[\basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T  \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy)   + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t}\big] \\  -  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArg{}{t} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
-\end{multline*}
-Again for notational purposes, we split this into three terms,
-$$
-\objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)  = 
-\objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)  + 
-\objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy)  + 
-\objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) ,$$
-where
-\begin{align*}
-& \objectiveControlArg{}_1(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) =  \frac{1}{2}\big[\basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{}{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy)   +  \basisspace [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{}{} \big],\\
-& \objectiveControlArg{}_2(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) = -  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy)  + \basisspace [\massArgnt{}]^{-1} \controllerDiscreteDumArgnt{}\big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] ,\\
-& \objectiveControlArg{}_3(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) = \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] .
-\end{align*}
-Evaluating the first term,
-\begin{align*}
-\frac{\partial  \objectiveControlArg{}_1 }{\partial \genstateyDiscreteArgnt{}{}} &= 
-\frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) +  \\
-& \qquad \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) + \\
-& \qquad \frac{1}{2} \frac{\partial }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \controllerDiscreteDumArgnt{}\bigg)
-,\\ 
-&= 
-[\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace + 
-[\controllerDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \statey} \basisspace , \\ 
-%&= 
-%[\velocity(\basisspace \genstateyArgnt{} + \stateIntercept)]^T \stweightingMatArg{n} \basisspace [\massArgnt{n}]^{-1} \basisspace^T \stweightingMatArg{n}  
-% \frac{\partial \velocity}{\partial \statey} \basisspace + 
-%[\controllerArgnt{n}]^T [\massArgnt{n}]^{-1} \stweightingMatArg{n}   \basisspace  [\massArgnt{n}]^{-1}\basisspace^T
-%\stweightingMatArg{n} \frac{\partial \velocity}{\partial \statey} \basisspace , \\  
-&= 
-\bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \bigg)\basisspace^T  \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
-&= 
-[\genstateyDiscreteDotArgnt{} ]^T \basisspace^T  \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace , \\
-&= 
-[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace .
-\end{align*}
-Now evaluating the second term,
-\begin{align*}
-\frac{\partial  \objectiveControlArg{}_2 }{\partial \genstateyDiscreteArgnt{}{}} &= 
--\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}   \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\&=
--\frac{\partial  }{\partial \genstateyDiscreteArgnt{}{}}    [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T 
- \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  - 
-  [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace 
-\\
-&= 
-- 2[\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  
- \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace - 
- [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
-&= - \bigg[ 2[\velocity(\veloargsromy)]^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  + [\controllerDiscreteDumArgnt{}]^T  [\massArgnt{}]^{-1} \bigg] \basisspace^T \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace ,\\
-&= -\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \bigg) \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
-\end{align*}
-Next, we can use the result~\eqref{eq:i3sol} to have,
-$$ \frac{\partial  \objectiveControlArg{}_3 }{\partial \genstateyDiscreteArgnt{}{}} =
-  [\velocity(\veloargsromy) ]^T  \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
-$$
-Thus we have
-\begin{align*}
-\frac{\partial \objectiveControlArg{}}{\partial \genstateyDiscreteArgnt{}} &= 
-[\basisspace \genstateyDiscreteDotArgnt{} ]^T   \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
-\bigg( \big[ \basisspace \genstateyDiscreteDotArgnt{} ]^T  + \velocity(\veloargsromy)]^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T  \bigg) \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace +
- [\velocity(\veloargsromy) ]^T  \stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\
-&= 
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{}   \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace -
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)  \stweightingMatArg{}
-\frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, \\ 
-&= 
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{} \basisspace  [\massArgnt{}]^{-1}\basisspace^T - \mathbf{I} \bigg)
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace, 
-\end{align*}
-such that,
-\begin{equation*}
-\frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}} = 
-  \adjointDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}  \frac{\partial \velocity}{\partial \stateyDiscrete}\basisspace  \big]  + 
-\bigg( [\basisspace \genstateyDiscreteDotArgnt{} ]^T - \velocity(\veloargsromy)^T \bigg)   \bigg( \stweightingMatArg{} \basisspace  [\massArgnt{}]^{-1}\basisspace^T - \mathbf{I} \bigg)
-\stweightingMatArg{} \frac{\partial \velocity}{\partial \stateyDiscrete} \basisspace.
-\end{equation*}
-Equivalently we write
-\begin{equation*}
-\bigg[ \frac{\partial \hamiltonianArg{}}{\partial \genstateyDiscreteArgnt{}{}} \bigg]^T =  \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T \stweightingMatArg{}  \basisspace [\massArgnt{}]^{-1}  \adjointDiscreteDumArgnt{} +
- \basisspace^T \big[  \frac{\partial \velocity}{\partial \stateyDiscrete} \big]^T 
-  \stweightingMatArg{} \bigg(  \basisspace  [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{}- \mathbf{I} \bigg)
- \bigg( \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \bigg). 
-\end{equation*}
-Next, we evaluate $\frac{\partial \hamiltonianArg{}}{\partial \controllerDumArgnt{}}$:
-\begin{align*}
-  \frac{\partial \hamiltonianArg{}}{\partial \controllerDumArgnt{}{}} &= \frac{\partial}{\partial \controllerDumArgnt{}{}} \bigg[  \adjointDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDumArgnt{} \big] +  \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
- &= \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}} \bigg[  \adjointDiscreteDumArgnt{T} \big[  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsromy) + [\massArgnt{}]^{-1}\controllerDiscreteDumArgnt{} \big] \bigg]+  \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg] \\
- &= [\adjointDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} + \frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg].
-\end{align*}
-To evaluate $\frac{\partial}{\partial \controllerDiscreteDumArgnt{}{}}  \bigg[ \objectiveControlArg{}(\genstateyDiscreteArgnt{},\controllerDiscreteDumArgnt{},\timeDummy) \bigg]$, we again 
-go term by term. Starting with the first term,
-\begin{align*}
-\frac{\partial  \objectiveControlArg{}_1 }{\partial \genstateyDiscreteArgnt{}{}} &= 
-\frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}} \bigg( [\velocity(\veloargsromy)]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) + \\
-& \qquad \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg) + \\
-& \qquad \frac{1}{2} \frac{\partial }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \controllerDiscreteDumArgnt{}\bigg) \\
-&= \bigg(  [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{}  \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T                      
-\stweightingMatArg{}  \velocity(\veloargsromy) \big] \bigg)^T +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} \\
-&= \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} +   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}.
-\end{align*}
-Moving on to the second term,
-\begin{align*}
-\frac{\partial  \objectiveControlArg{}_2 }{\partial \controllerDiscreteDumArgnt{}{}} &= 
--\frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}}   \big[ \basisspace  [\massArgnt{}]^{-1}\basisspace^T
-\stweightingMatArg{}  \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  - \frac{\partial  }{\partial \controllerDiscreteDumArgnt{}{}} \bigg( [\controllerDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \bigg)  \\
-&=- \bigg[ [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace \genstateyDiscrete + \stateIntercept) \bigg]^T \\
-&= - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}. 
-\end{align*}
-For the third term we have simply,
-$$ \frac{\partial  \objectiveControlArg{}_3 }{\partial \controllerDiscreteDumArgnt{}{}} = \bz .$$
-Thus,
-\begin{align*}
-\frac{\partial \hamiltonianArg{}}{\partial \controllerDiscreteDumArgnt{}} &= 
-[\adjointDiscreteDumArgnt{}]^T[\massArgnt{}]^{-1}  +  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} + \\
-& \qquad   [\controllerDiscreteDumArgnt{}]^T [\massArgnt{}]^{-1} \basisspace^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1} - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}, \\
-& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace \bigg) [\massArgnt{}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{}]^{-1}\basisspace^T  \stweightingMatArg{} \basisspace [\massArgnt{}]^{-1}  \\
-& = \bigg( [\adjointDiscreteDumArgnt{}]^T - [\velocity(\veloargsromy) ]^T \stweightingMatArg{} \basisspace \bigg) [\massArgnt{}]^{-1}  + \bigg(  \big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{} \basisspace + [\controllerDiscreteDumArgnt{}]^T   \bigg) [\massArgnt{}]^{-1}  \\
-& =  [\adjointDiscreteDumArgnt{}]^T  [\massArgnt{}]^{-1}  +  [\controllerDiscreteDumArgnt{}]^T   [\massArgnt{}]^{-1}  
-\end{align*}
-Or, equivalently,
-\begin{equation*}
-\frac{\partial \hamiltonianArg{}}{\partial \controllerDiscreteDumArgnt{}} = 
-  [\massArgnt{}]^{-1} [\adjointDiscreteDumArgnt{} +  \controllerDiscreteDumArgnt{}]. 
-\end{equation*}
-Evaluating at $(\genstateArg{}{t},\genstateDotArg{}{t},t$), the gradients in the Pontryagin Maximum Principle yield
-\begin{align*}%\label{eq:hamiltonian_sysa}
-&\frac{d}{dt}\genstateArg{}{t} =  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\veloargsrom) + [\massArgnt{}]^{-1}\controllerArg{}{t}\\ 
-&\frac{d }{dt} \adjointArg{}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{}  \basisspace [\massArgnt{}]^{-1}  \adjointArg{}{t} = 
- - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{} 
-\bigg(  \basisspace  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{} - \mathbf{I} \bigg)
- \bigg( \basisspace \genstateDotArg{}{t} - \velocity(\basisspace \genstateArg{}{t} + \stateIntercept) \bigg) \\ 
-& \adjointArg{}{t}= -\controllerArg{}{t}.
-\end{align*}
-This can be written equivalently as
-\begin{align*}%\label{eq:hamiltonian_sys}
-&\frac{d}{dt}\genstateArg{}{t} =  [\massArgnt{}]^{-1}\basisspace^T \stweightingMatArg{}\velocity(\basisspace \genstateArg{}{t} + \stateIntercept) + [\massArgnt{}]^{-1}\controllerArg{}{t} \\ 
-&\frac{d}{dt} \controllerArg{}{t}  + \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{}  \basisspace [\massArgnt{}]^{-1}  \controllerArg{}{t} = 
- - \basisspace^T \big[  \frac{\partial \velocity}{\partial \statey} \big]^T \stweightingMatArg{} 
-\bigg( \mathbf{I} -   \basisspace  [\massArgnt{}]^{-1}\basisspace^T \bigg)
-\stweightingMatArg{} \bigg( \basisspace \genstateDotArg{}{t} - \velocity(\basisspace \genstateArg{}{t} + \stateIntercept) \bigg). 
-\end{align*}
-
-\begin{comment}
-
-====================
-To do this, first write the above as,
-$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T + \basisspace \basisspace^T \bigg)  \stweightingMat^T + \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0.
-$$
-Re-arranging gives,
-$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg]  +  \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T  \basisspace   +  \frac{d}{dt} \bigg] \bigg( \basisspace^T \stweightingMat \basisspace \dot{\genstate}   -  \basisspace^T \stweightingMat  \velocity \bigg) = 0.
-$$
-Defining the adjoint as,
-\begin{equation}
-\basisspace^T \stweightingMat \basisspace \frac{d \genstate}{dt}  -  \basisspace^T \stweightingMat \velocity =  \boldsymbol \lambda , \qquad \genstate(t=0) = \genstate_0,
-%\basisspace \frac{d \genstate}{dt}  -   \velocity =  \adjoint  , \qquad \genstate(t=0) = \genstate_0
-\end{equation}
-we then get the auxiliary adjoint equation,
-\begin{equation}
-%\basisspace^T \stweightingMat^T \frac{d}{dt} \adjoint  + \basisspace^T \frac{\partial \velocity}{\partial \genstate}^T \stweightingMat^T \adjoint = 0, \qquad \basisspace^T \stweightingMat \adjoint(t=T) = 0
- \frac{d}{dt} \adjoint  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \state} \bigg]^T \basisspace \adjoint = -\bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg] , \qquad \adjoint(t=T) = 0.
-\end{equation}
-\end{comment}
-\begin{comment}
-=================================
-
-%We start $ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate}$. We go term by term and note the useful link: \url{https://en.wikipedia.org/wiki/Matrix_calculus#Layout_conventions}
-%First term
-\begin{enumerate}
-\item First term: Only depends on $\dot{\genstate}$, term is zero
-\item Second term: $$\frac{\partial}{\partial \genstate} \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)=  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \frac{\partial}{\partial \genstate} \bigg( \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg)$$
-$$ = \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat   \frac{\partial \boldsymbol f(\state)}{\partial \state}\frac{\partial \state}{\partial \genstate} $$
-$$ =  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace $$
-%$${\color{red} \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big] \bigg) = \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \bigg) \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
-$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
-$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \basisspace.$$
-%$${\color{red} ?  = \big[ \frac{\partial \velocity}{\partial \state}\frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
-%$${\color{red} ?  = \big[ \basisspace \frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
-\item Third term: 
-$$\frac{1}{2} \frac{\partial }{\partial \genstate} \bigg( \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)  =  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
-$$ =   \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisspace$$
-\end{enumerate}
-This gives us,
-$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} = - \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace +  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace $$
-
-Now we work on $\frac{\partial \minintegrand}{\partial \dot{ \genstate } }$:
-\begin{enumerate}
-\item First term:
-$$\frac{1}{2} \frac{\partial}{\partial \dot{\genstate}}\bigg( \big[\dot{\genstate}^T  \basisspace^T \big]  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  \bigg) =  \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace$$
-\item Second term:
-$$\frac{\partial}{\partial \dot{\genstate} } \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg) = \frac{\partial}{\partial \dot{\genstate} } \bigg( \dot{\genstate}^T \bigg) \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  $$
-$$ = \bigg[  \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg]^T $$
-$$ = \velocity^T \stweightingMat^T \basisspace$$
-\end{enumerate}
-This gives,
-$$\frac{\partial \minintegrand}{\partial \dot{ \genstate } } =   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace -  \velocity^T \stweightingMat^T \basisspace.$$
-%\end{comment}
-
-\begin{comment}
-Let $\stweightingMat = \mathbf{I}$,
-$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} - \frac{d}{dt} \bigg[ \frac{\partial \boldsymbol F}{\partial \dot{\genstate}} \bigg] =   \bigg( -\big[ \basisspace \dot{\genstate} \big]^T  +  \velocity^T \bigg) \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T  - \velocity^T \basisspace \bigg] $$
-Transposing,
-$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[   \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
-$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[ \basisspace^T \basisspace  \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
-$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
-$$ =    -\basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( \basisspace \dot{\genstate}  -  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
-$$ = -\bigg[ \basisspace^T \frac{d}{dt} + \basisspace^T \big[ \frac{\partial \velocity}{\partial \state} \big]^T  \bigg] \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$ 
-Adjoint style
-$$ =   \basisspace^T  \frac{d }{dt}\lambda  +   \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \lambda  $$
-$$ \basisspace  \dot{\genstate}  -\velocity = \lambda$$
-%Gathering terms
-
-\end{comment}
-%\begin{comment}
-
-\begin{comment}
-
-\subsection{Time-varying spacetime subspace}
-We have $\state(t)= \basisst(t) \genstate + \state_0$.
-Note that, in this case, the generalized coordinates are stationary in time, $\dot{\genstate} = \mathbf{0}$.
-Thus we only need to compute $\frac{\partial \minintegrand}{\partial \genstate}$.
-Going term by term:
-
-\begin{enumerate}
-\item First term: 
-$$\frac{1}{2} \frac{\partial}{\partial \genstate} \bigg(\big[\frac{\partial \decoder}{\partial t}  \big]^T \stweightingMat  \big[\frac{\partial \decoder}{\partial t} \big]\bigg) = 
-\frac{\partial}{\partial \genstate} \bigg(\big[\frac{\partial \basisst}{\partial t} \genstate \big]^T \stweightingMat  \big[\frac{\partial \basisst}{\partial t} \genstate \big]\bigg) $$
-$$
-= \genstate^T \dot{\basisst}^T \stweightingMat \dot{ \basisst }
-$$
-\item Second term: $$\frac{\partial}{\partial \genstate} \bigg( \big[ \frac{\partial \decoder}{\partial t} \big]^T \stweightingMat \big[ \velocity(\decoder,t;\param) \big] \bigg)= \frac{\partial}{\partial \genstate} \bigg( \big[ \frac{\partial \basisst}{\partial t} \genstate  \big]^T \stweightingMat  \big[ \velocity(\decoder,t;\param) \big]  \bigg)$$
-$$ = \big[ \frac{\partial \basisst}{\partial t} \genstate  \big]^T \stweightingMat \frac{\partial}{\partial \genstate}\bigg( \big[ \velocity(\decoder,t;\param) \big] \bigg) +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T \frac{\partial}{\partial \genstate} \bigg( \big[ \frac{\partial \basisst}{\partial t} \genstate  \big] \bigg)$$
-$$ = \big[ \dot{\basisst} \genstate  \big]^T \stweightingMat \frac{\partial \velocity}{\partial \decoder} \frac{\partial \decoder}{\partial \genstate} +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T  \dot{ \basisst}$$
-$$ = \big[ \dot{\basisst} \genstate  \big]^T \stweightingMat \frac{\partial \velocity}{\partial \decoder} \basisst +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T  \dot{ \basisst}$$
-\item Third term: 
-$$\frac{1}{2} \frac{\partial }{\partial \genstate} \bigg( \big[\velocity(\decoder,t;\param) \big]^T \stweightingMat \big[ \velocity(\decoder,t;\param) \big] \bigg)  =  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
-$$ =   \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisst$$
-Putting it all together we get,
-$$ \genstate^T \dot{\basisst}^T \stweightingMat \dot{\basisst} - \bigg(  \big[ \dot{\basisst} \genstate  \big]^T \stweightingMat \frac{\partial \velocity}{\partial \decoder} \basisst +  \big[ \velocity(\decoder,t;\param) \big]^T \stweightingMat^T  \dot{ \basisst} \bigg) + \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisst = \mathbf{0}$$
-Transposing, we obtain the Euler--Lagrange equations for a space-time basis
-$$ \basisst^T \stweightingMat^T \dot{\basisst} \genstate - \bigg( \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \dot{\basisst} \genstate + \dot{\basisst}^T \stweightingMat \velocity  \bigg) + \basisst^T  \big[ \frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \velocity  = \mathbf{0}, \qquad \state(t=0) = \state_0.$$
-Note that the additional boundary condition is automatically satisfied.
-$$ \dot{\basisst}^T \stweightingMat^T \bigg( \dot{\basisst} \genstate -    \velocity \bigg)   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \dot{\basisst} \genstate  + \basisst^T  \big[ \frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \velocity  = \mathbf{0}, \qquad \state(t=0) = \state_0.$$
-$$ \dot{\basisst}^T \stweightingMat^T \bigg( \dot{\basisst} \genstate -    \velocity \bigg)   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \bigg( \dot{\basisst} \genstate  -  \velocity \bigg) = \mathbf{0} $$
-$$ \bigg[ \dot{\basisst}^T \stweightingMat^T  - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \bigg]  \bigg( \dot{\basisst} \genstate -    \velocity \bigg) = \mathbf{0}$$
-
-Note that we can write this as two systems. To put it in the best form, write the above as,
-$$ \dot{\basisst}^T \bigg( \mathbf{I} -  \basisst \basisst^T  +  \basisst \basisst^T \bigg)  \stweightingMat^T \bigg( \dot{\basisst} \genstate -    \velocity \bigg)   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \bigg( \mathbf{I} -  \basisst \basisst^T  +  \basisst \basisst^T \bigg)\stweightingMat^T \bigg( \dot{\basisst} \genstate  -  \velocity \bigg) = \mathbf{0} $$
-Defining,
-$$  \basisst^T \stweightingMat^T \dot{\basisst} \genstate  -  \basisst^T \stweightingMat^T \velocity = \boldsymbol \lambda,$$
-we have,
-$$ \bigg[ \dot{\basisst}^T   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \bigg] \basisst \adjoint  = -\bigg[\dot{\basisst}^T   - \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \bigg]  \bigg( \mathbf{I} -  \basisst \basisst^T  \bigg)\stweightingMat^T \bigg( \dot{\basisst} \genstate  -  \velocity \bigg)$$
-
-%The above can be written as two systems,
-%$$  \dot{\basisst}^T \stweightingMat^T \dot{\basisst} \genstate - \dot{\basisst}^T \stweightingMat \velocity  = \boldsymbol \lambda $$
-%$$  \basisst^T \big[\frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \dot{\basisst} \genstate  - \basisst^T  \big[ \frac{\partial \velocity}{\partial \decoder} \big]^T \stweightingMat^T \velocity  = \lambda$$
-
-\end{enumerate}
-\end{comment}
-
-\begin{comment}
-\textbf{Linear subspace with full temporal support}:
-For a linear subspace with full temporal support, we have $\state = \basisspace \genstate + \state_0$
-$ \frac{\partial \state}{\partial \genstate} = \basisspace.$
-We have,
-$$\minintegrand = \frac{1}{2} \big[ \basisspace \dot{\genstate} \big]^T  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  -  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]
-  + \frac{1}{2} \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]$$
-Write this as,
-$$\minintegrand = \minintegrand_1 + \minintegrand_2 + \minintegrand_3 $$
-
-%other form
-%$${\color{red} = \big[ \basisspace \dot{\genstate} \big]^T  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  - 2 \big[ \velocity(\state(\genstate),t;\param) \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big] + \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] }$$
-
-First we compute $ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate}$. We go term by term and note the useful link: \url{https://en.wikipedia.org/wiki/Matrix_calculus#Layout_conventions}
-%First term
-\begin{enumerate}
-\item First term: Only depends on $\dot{\genstate}$, term is zero
-\item Second term: $$\frac{\partial}{\partial \genstate} \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)=  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \frac{\partial}{\partial \genstate} \bigg( \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg)$$
-$$ = \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat   \frac{\partial \boldsymbol f(\state)}{\partial \state}\frac{\partial \state}{\partial \genstate} $$
-$$ =  \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace $$
-%$${\color{red} \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big] \bigg) = \frac{\partial}{\partial \genstate} \bigg(\big[ \velocity(\state(\genstate),t;\param) \big]^T \bigg) \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
-$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
-$$ = \dot{\genstate}^T \basisspace^T \stweightingMat^T \frac{\partial \velocity}{\partial \state} \basisspace.$$
-%$${\color{red} ?  = \big[ \frac{\partial \velocity}{\partial \state}\frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
-%$${\color{red} ?  = \big[ \basisspace \frac{\partial \state}{\partial \genstate} \big]^T \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  }  $$
-\item Third term: 
-$$\frac{1}{2} \frac{\partial }{\partial \genstate} \bigg( \big[\velocity(\state(\genstate),t;\param) \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big] \bigg)  =  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \frac{\partial \state}{\partial \genstate}.$$
-$$ =   \velocity^T \stweightingMat\frac{\partial \velocity}{\partial \state} \basisspace$$
-\end{enumerate}
-This gives us,
-$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} = - \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace +  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace $$
-Now we work on $\frac{\partial \minintegrand}{\partial \dot{ \genstate } }$:
-\begin{enumerate}
-\item First term:
-$$\frac{1}{2} \frac{\partial}{\partial \dot{\genstate}}\bigg( \big[\dot{\genstate}^T  \basisspace^T \big]  \stweightingMat  \big[ \basisspace \dot{\genstate} \big]  \bigg) =  \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace$$
-\item Second term:
-$$\frac{\partial}{\partial \dot{\genstate} } \bigg( \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg) = \frac{\partial}{\partial \dot{\genstate} } \bigg( \dot{\genstate}^T \bigg) \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  $$
-$$ = \bigg[  \basisspace^T  \stweightingMat \big[ \velocity(\state(\genstate),t;\param) \big]  \bigg]^T $$
-$$ = \velocity^T \stweightingMat^T \basisspace$$
-\end{enumerate}
-This gives,
-$$\frac{\partial \minintegrand}{\partial \dot{ \genstate } } =   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace -  \velocity^T \stweightingMat^T \basisspace.$$
-Putting it all together, the Euler--Lagrange equation reads,
-$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} - \frac{d}{dt} \bigg[ \frac{\partial \minintegrand}{\partial \dot{\genstate}} \bigg] =  - \big[ \basisspace \dot{\genstate} \big]^T \stweightingMat  \frac{\partial \boldsymbol f(\state)}{\partial \state}\basisspace +  \velocity^T \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace - \velocity^T \stweightingMat^T \basisspace \bigg]  = 0$$
-Factoring the first term,
-$$ - \bigg( \big[ \basisspace \dot{\genstate} \big]^T  -  \velocity^T \bigg) \stweightingMat \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T \basisspace^T \stweightingMat \basisspace -  \velocity^T \stweightingMat^T \basisspace  \bigg] = 0$$
-Taking the transpose,
-$$ -  \basisspace^T \frac{\partial \velocity}{\partial \state}^T \stweightingMat^T \bigg( \big[ \basisspace \dot{\genstate} \big]  -  \velocity \bigg) -  \frac{d}{dt} \bigg[  \basisspace^T \stweightingMat^T \basisspace \dot{\genstate} - \basisspace^T \stweightingMat \velocity  \bigg] = 0.
- $$
-Noting that $\stweightingMat = \stweightingMat^T$, we factor the second term,
-$$ -  \basisspace^T \frac{\partial \velocity}{\partial \state}^T \stweightingMat^T \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) -  \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg[   \basisspace \dot{\genstate} - \velocity  \bigg] = 0.
-$$
-Gathering terms and multiplying by negative one,
-$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \stweightingMat^T + \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0.
-$$
-%Equivalently,
-%$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T  \basisspace \basisspace^T \stweightingMat^T + \basisspace^T  \basisspace \basisspace^T \stweightingMat^T  \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0.
-$$
-%Pulling in the $\stweightingMat^T \basisspace^T$,
-%$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \basisspace  +  \frac{d}{dt} \bigg] \bigg(  \basisspace^T \stweightingMat \basisspace \dot{\genstate}   -  \basisspace^T \stweightingMat \velocity \bigg) = 0.
-$$
-
-This is the time-continuous LSPG ROM. Note that this is a second order equation and can be written as two separate first order equations. To do this, first write the above as,
-$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T + \basisspace \basisspace^T \bigg)  \stweightingMat^T + \basisspace^T \stweightingMat^T \frac{d}{dt} \bigg] \bigg(  \basisspace \dot{\genstate}   -  \velocity \bigg) = 0.
-$$
-Re-arranging gives,
-$$ \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg]  +  \bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T  \basisspace   +  \frac{d}{dt} \bigg] \bigg( \basisspace^T \stweightingMat \basisspace \dot{\genstate}   -  \basisspace^T \stweightingMat  \velocity \bigg) = 0.
-$$
-Defining the adjoint as,
-\begin{equation}
-\basisspace^T \stweightingMat \basisspace \frac{d \genstate}{dt}  -  \basisspace^T \stweightingMat \velocity =  \boldsymbol \lambda , \qquad \genstate(t=0) = \genstate_0,
-%\basisspace \frac{d \genstate}{dt}  -   \velocity =  \adjoint  , \qquad \genstate(t=0) = \genstate_0
-\end{equation}
-we then get the auxiliary adjoint equation,
-\begin{equation}
-%\basisspace^T \stweightingMat^T \frac{d}{dt} \adjoint  + \basisspace^T \frac{\partial \velocity}{\partial \genstate}^T \stweightingMat^T \adjoint = 0, \qquad \basisspace^T \stweightingMat \adjoint(t=T) = 0
- \frac{d}{dt} \adjoint  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \state} \bigg]^T \basisspace \adjoint = -\bigg[\basisspace^T \frac{\partial \velocity}{\partial \state}^T \bigg( \mathbf{I} -   \basisspace \basisspace^T \bigg)  \stweightingMat^T \bigg( \basisspace \dot{\genstate}   -   \velocity \bigg) \bigg] , \qquad \adjoint(t=T) = 0.
-\end{equation}
-
-\end{comment}
-
-\begin{comment}
-Let $\stweightingMat = \mathbf{I}$,
-$$ \frac{\partial \minintegrand}{\partial \boldsymbol \genstate} - \frac{d}{dt} \bigg[ \frac{\partial \boldsymbol F}{\partial \dot{\genstate}} \bigg] =   \bigg( -\big[ \basisspace \dot{\genstate} \big]^T  +  \velocity^T \bigg) \frac{\partial \velocity}{\partial \state} \basisspace - \frac{d}{dt} \bigg[   \dot{\genstate}^T  - \velocity^T \basisspace \bigg] $$
-Transposing,
-$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[   \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
-$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)- \frac{d}{dt} \bigg[ \basisspace^T \basisspace  \dot{\genstate}  -\basisspace^T  \velocity \bigg] $$
-$$ =    \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( -\big[ \basisspace \dot{\genstate} \big]  +  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
-$$ =    -\basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \bigg( \basisspace \dot{\genstate}  -  \velocity \bigg)-  \basisspace^T  \frac{d}{dt} \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$
-$$ = -\bigg[ \basisspace^T \frac{d}{dt} + \basisspace^T \big[ \frac{\partial \velocity}{\partial \state} \big]^T  \bigg] \bigg[\basisspace  \dot{\genstate}  -\velocity \bigg] $$ 
-Adjoint style
-$$ =   \basisspace^T  \frac{d }{dt}\lambda  +   \basisspace^T  \big[ \frac{\partial \velocity}{\partial \state} \big]^T   \lambda  $$
-$$ \basisspace  \dot{\genstate}  -\velocity = \lambda$$
-%Gathering terms
-\end{comment}
-
-\end{appendices}
-
-\bibliographystyle{siam}
-\bibliography{refs}
 \end{document}

### 2020-08-25 numerical_experiments.tex
minor language changes
working on revision

@@
 First, Figure~\ref{fig:sod_density} presents the density solutions produced by the various ROMs at $t = 0.5$ and $1.0$. Figure~\ref{fig:sod_xt} shows 
-$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, we observe that the LSPG and Galerkin ROMs accurately characterize 
+$\mathsf{x}-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, we observe that the LSPG and Galerkin ROMs accurately characterize 
 the system: they correctly track the shock location, expansion waves, etc. We observe both predictions, however, to be highly oscillatory. These oscillations are 
@@
 
-Finally, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
+Next, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
 window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
@@
 
+\subsubsection{Results as a function of basis dimension}
+Lastly, we assess the impact of the basis dimension on the performance of the various reduced-order models.
+We again consider S-reduction trial subspaces.
+We obtain the trial subspaces by evolving the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$ with the CN scheme.
+We execute Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}} = 2$, $\stateIntercept = \state_0$\footnote{We observed negative pressures in the projected initial conditions for trial subspaces of dimension $K=20,30,40$ if we used the offset $\stateIntercept = \bz$.
+Thus, for our convergence study we employed $\stateIntercept = \state_0$ so-as to avoid negative pressures in the projected initial conditions.}, and $\romdim = 10,20,30,40,50$.
+We investigate reduced-order models based on Galerkin projection, LSPG projection, and the WLS approach.
+All methods leverage the CN time scheme with $\Delta t = 0.002$.
+We present results for windows of size $\Delta T^n \equiv \Delta T = 0.002,0.004,0.02,0.04,0.10$ for the WLS approach.
+All WLS ROMs are solved via the direct approach described previously.
+
+Figure~\ref{fig:convergence_study} shows the integrated $\ell^2$-error of the various ROM solutions, as well as the optimal solution as measured in the $\ell^2$-norm.
+In Figure~\ref{fig:sod_error_converge} we first observe that for a basis of dimension $\romdim = 10$, the Galerkin, LSPG, and WLS $\Delta T = 0.002$, $\Delta T = 0.004$ ROMs do not converge.
+Upon investigation, this lack of convergence occurred due to the presence of negative pressure.
+We next observe that the $\ell^2$-error monotonically decreases as the basis dimension increases for all remaining ROM solutions with the exception of the Galerkin, LSPG, and WLS $\Delta T = 0.002$, $\Delta T = 0.004$ at dimension $\romdim = 30$.
+Overall, we observe that larger window sizes yield lower errors when the basis dimension is small.
+For larger basis dimensions we observe all ROMs to yield accurate solutions, and a larger window size does not always yield a lower error.
+In Figure~\ref{fig:sod_residual_converge}, we next observe that the residual monotonically decreases as the basis dimension grows.
+This result is supported, although not guaranteed, by the error analysis presented in Section~\ref{sec:analysis} and Ref.~\cite{carlberg_lspg_v_galerkin}.
+Finally, we observe that larger window sizes yields lower integrated residuals.
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[width=1.\linewidth]{figs/sod/error_converge.pdf}
+\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
+\label{fig:sod_error_converge}
+\end{subfigure}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[width=1.\linewidth]{figs/sod/residual_converge.pdf}
+\caption{Space--time residual norm of the \methodAcronymROM.} 
+\label{fig:sod_residual_converge}
+\end{subfigure}
+\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
+\label{fig:convergence_study}
+\end{center}
+\end{figure}
+
 \subsubsection{Summary of numerical results for the Sod shock tube}

### 2020-10-09 response_to_reviewers/response.tex
latest updates

%\documentclass{article}
\documentclass[3p,computermodern,10pt]{elsarticle}
\input{../packages}
\input{../commands}
\input{../commands_cwst}
%\input{title}
%\begin{frontmatter}
%\title{Continuous Least-Squares Petrov-Galerkin for Model Reduction}
%\author{Eric J. Parish}
\begin{document}
\begin{frontmatter}

%\title{Model Reduction via \methodName}
%\title{Windowed least--squares reduced-order models for dynamical systems}
\title{Windowed least-squares model reduction for dynamical systems}
%\title{The windowed least-squares framework for model reduction of dynamical systems}

\author[a]{Eric J. Parish and Kevin T. Carlberg}
%\ead{ejparis@sandia.gov}

\address[a]{Sandia National Laboratories,  Livermore, CA}
\end{frontmatter}

%\maketitle
We would like to thank both of the reviewers for their helpful comments. We have done our best to address all concerns, and we feel that the comments have improved the quality of the manuscript. We have primarily improved our numerical results section, where we have (1) included studies of convergence with basis dimension, (2) included a parameterized example, and (3) increased the difficulty of our cavity flow experiment. In addition, we have worked on our analyses to try and address several of the concerns raised by the reviewers. Please see our responses below, in red.
\section{Reviewer 2}
The article proposes a windowed least-squares approach for model reduction of dynamical systems. The methodology is proposed as a generalization of existing approaches such as the least square Petrov-Galerkin (LSPG) method and the space-time LSPG method. The proposed methodology is clearly presented and related with existing and previously proposed approaches. The WLS methods is numerically implemented using two different approaches a direct one, which first discretizes then optimizes and an indirect one which first optimizes then discretizes. The article proposes also a theoretical analysis of the proposed methods. The paper is well written and easy to follows. All the methods are presented with a great level of details and results are convincing therefore I recommend publication after the raised issues are addressed.

\subsection{Major comments}
My biggest concern is about the numerical results section that in my opinion should be improved
\begin{enumerate}
\item Both numerical results should also show a convergence analysis of the methodology with respect to the dimension of the trial subspace used for the projection. Is the methodology sensitive with respect to the dimension of the trial subspace? Does the error decreases monotonically as I increase the number of reduced basis functions?

{\color{red} We have added a new sections to both of our first numerical experiment to address this question. We show that the behavior of the method is as one would expect: increasing the basis dimension yields both lower errors and lower residuals. We observe that the error monotonically decreases for larger window sizes. For smaller window sizes we do not see monotonic convergence of the error; this result has been seen in the community for, e.g., the Galerkin method and LSPG. We observe similar behavior with the residuals.}

\item In order to have a better understanding of the time saving of the implemented ROM, in figure 8 the authors should report also the wall-clock times of the FOM. Also figure 10 should have a comparison also with the FOM wall-clock times.
{\color{red} We have added CPU timings to our second numerical experiment to address this concern. In our first numerical experiment, which serves as a reproducible benchmark, we do not employ hyper-reduction, and therefore do not obtain computational savings. Our reasoning for not including CPU timings in the first draft of the manuscript was that we felt that the timing with respect to the least-squares Petrov--Galerkin method was the relevant metric to report. The reason for this is that the speedup obtained by the ROM for nonlinear problems is highly dependent on the sample mesh, the associated stencil, and the computational kernels involved in evaluating residuals on the sample mesh. As LSPG has been shown to yield significant computational speedups in previous work (e.g.,~\cite{carlberg_lspg_v_galerkin}), and WLS with S-reduction subspaces is a windowed extension of the LSPG approach, we felt that comparing the speed of the proposed method to LSPG is the relevant timing. However, we do agree that comparison to the FOM walltime is a metric of interest, and therefore we have worked hard to integrate it into the second example.}

\item I think that both numerical examples are rather simple to test the effectiveness of the method in real world situations. Moreover the numerical examples are used only to the test the reproductive ability of the ROM. I think that the authors should show the effectiveness of the methods on a more complex example which involves parametric dependency or time extrapolation.

{\color{red} To address this concern, we have (1) increased the difficulty of the second numerical experiment by raising the mach number such that is supersonic, thus adding a complex shock boundary layer interaction to the problem, and (3) added a new numerical experiment with varying parameters. In both cases, we observe WLS to yield improvements over the LSPG approach.}

\end{enumerate}

\subsection{Minor comments}

\begin{itemize}
\item page 25 second line, $\Delta T = .002 \rightarrow  \Delta T = 0.002$
{\color{red} Thank you, good catch!}

\item Could the authors explain what is the meaning of the integrated residual norm for the l2
projection? This is not clear to me.

{\color{red} We have rephrased this as the ``time-integrated residual norm". By integrated norm, we were referring to the Euclidean $\ell^2$-norm of the state integrated in time, as denoted in, e.g., Figure 6a. In the previous manuscript, we also had an inconsistency of using both the term ``space--time" error norm and "integrated" error norm to refer to the same thing; we have corrected this in the revision by replacing space--time error norm with time-integrated norm.}

\item In figure 6, the metric used to measure the error is not so informative, probably a relative error would be more informative.

{\color{red} Thank you, we have updated figures with relative errors.}
\end{itemize}

\section{Reviewer 3}
The authors presented a space-time model reduction method that aims to generalize/unify a number of different model reduction techniques and improve some shortcomings (exponential error growth in time, strong dependence on time discretization, non-intuitive relationship between error and time step size, etc). The proposed approach minimizes the time-continuous residual (integrated) over some time window in a space-only or space-time trial space. Two different approaches are proposed for solving the resulting WLS optimization problem: direct and indirect; both approaches perform similarly in numerical experiments. Numerical experiments show improved performance of WLS over the Galerkin and minimum-residual methods (called LSPG in this paper) and extensive testing is done to determine the performance of the method with respect to time step and time window. The proposed method is interesting and novel; however, I am recommending major revisions as there are
important aspects of the analysis and numerical experiments that should be addressed.

\subsection{Minor comments}
\begin{enumerate}
\item The "windowing" approach proposed in this work is very similar or identical to the concept of a space-time slab. There is no need to change the name of the method, but the similarity should be mentioned at least in the introduction and relevant literature cited.

{\color{red} We agree with the reviewer in this similarity (in fact, we used the terminology time-slab in our early drafts of the manuscript) , and have added several relevant citations. We are glad to add additional ones if the author has any recommendations. As to the difference between our approach and a time-slab, our perspective is as follows: While the term time-slab can entail many things, from our perspective, typically a space--time slab is viewed as a discretization technique. For example, in a space--time DG method, we might break the space--time domain down into space--time slabs, then represent the spatial and temporal dependence of the solution with polynomials on each space--time slab. In our setup, we break the space--time domain down into windows, and then rather than discretizing in time, we reformulate the problem as a least-squares problem on each window. We are then flexible to use any time discretization technique on that window, so perhaps a window may comprise various space--time slabs in the traditional sense. However, the two concepts are very closely tied and its easy to make a counter argument to our perspective! We have added the following to the manuscript, at the end of page 2 in or related work sections:

``Lastly, we note that the windowing approach proposed here displays commonalities with space–time methods employing space–time slabs, e.g., Refs.[35, 54, 36, 34]. In these approaches, the space–time domain is typically decomposed into a series of space–time slabs. A system of algebraic equations is then formed over each slab via, e.g., the discontinuous Galerkin method, and the space–time method operates by sequentially marching through these space–time slabs. From this perspective, the proposed windowing approach is best viewed as a space–time method in where each window comprises multiple time-slabs. The time-slabs within each window experience a two-way coupling to each other, by virtue of the residual minimization statement being defined over the entire window, and thus one can no longer sequentially march over the time-slabs."
} 
\item It should be pointed out that the argument in the middle of Page 22 (the equation before (5.5)) assumes the WLS method finds the *global* minimizer (which is not guaranteed because the optimization problems are non-convex).

{\color{red} We have added a footnote clarifying this, thank you!}
\end{enumerate}
Major comments
\begin{enumerate}
\item A confusing aspect of the paper that should be addressed is the "space" terminology. There is no concept of a spatial domain here since the starting point is a system of ODEs (2.1); there is only a concept of the state and how it evolves in time. If these ODEs come from the discretization of PDEs, then there is an equivalence between the ODE state and the solution of the PDE at points in space (which is the case in the numerical examples). If the authors believe the method is useful beyond semi-discrete PDEs, then the terminology should be adjusted or at least clarified. If not, then the paper should indicate this method only makes sense for semi-discrete PDEs.

{\color{red} This is a good point, and one that was on our mind while preparing the manuscript. We have added the following clarifying footnote to Page 6 after the concept of "spatial" model reduction is first introduced:

\textit{In this work, we refer to the “state-space” dimension of an ordinary differential equation as the “spatial” dimension. We use this notation due to the common application of projection-based ROMs to semi-discretized partial differential equations. We emphasize, however, that the presented approach is applicable to ordinary differential equations that do not associate with some type of spatial discretization, i.e., as in molecular dynamics.}

We have not modified the terminology in the paper, however. We feel that it is best to not do this as 1) previous work on which this builds uses the notion of space in a similar fashion (i.e., the space--time LSPG approach), and 2) we feel that referring to the ``spatial" dimension as, i.e., the ``state-space" dimension becomes cumbersome and slightly confusing. We are glad to adapt a new notation at the editors request, however.}

 \item  Assumption A4 is also very strong. What would happen if a similar assumption was introduced into the analysis of the Galerkin or LSPG methods? To provide a "fair" comparison between the error bound for the WLS method and Galerkin/LSPG methods (to support the claim the WLS method overcomes the exponential error growth) the authors are attempting to improve upon, the same (or very similar) assumptions should be made in both cases.

{\color{red} In our revision, we have removed this assumption by slightly modifying our analysis. We now place a bound on the in-plane error, and the constant multiplying the residual term now has an additional factor of two. The main take away of  the analysis remains the same.}

\item While I appreciate the author's analysis, Assumptions A3 is unreasonable unless somehow justified. Are there any relevant dynamical systems whose residual function is inverse Lipschitz continuous? If so, are any of these semi-discrete PDEs? It seems unlikely even for simpler PDEs as inverse Lipschitz continuity would guarantee that PDE residual bounds the PDE error. The analysis would be much more meaningful without this assumption; obviously it would change the theorem because the last term of (5.3) would be different. At minimum, a detailed remark should be included regarding the relevance of such an assumption.

{\color{red} We agree that the assumption that the residual is inverse Lipshitz continuous in unreasonable. In our analysis, however, we make the assumption that the integrated residual is inverse Lipshitz continuous for any two trajectories starting from the same initial condition. We do not believe that we explained this very well, and have added a comment on this.

We believe that this is \textit{not} an unreasonable assumption, as it is essentially stating that the ODE is well-posed. Note that, for the ODE to be well-posed, only one trajectory can give a zero residual. Thus, for a well-posed ODE 
$$ \int_0^T \| \boldsymbol r(\boldsymbol x,t) - \boldsymbol r(\boldsymbol y,t) \|_2 dt > 0 $$
for all $\boldsymbol x, \boldsymbol y \in \{ \mathbb{R}^N \otimes \mathcal{T} | \boldsymbol x(0) = \boldsymbol y(0), \boldsymbol x \ne \boldsymbol y$\}. From this, the inverse Lipshitz assumption 
$$ \int_0^T \| \boldsymbol r(\boldsymbol x,t) - \boldsymbol r(\boldsymbol y,t) \|_2 dt \ge \alpha \int_0^T \| \boldsymbol x(t) - \boldsymbol y(t) \|_2 dt$$
is reasonable. Additionally, it is worth noting that in Ref.~\cite{choi_stlspg}, the time-discrete space--time residual is shown to be inverse Lipshitz continuous under several assumptions. While there is without a doubt a difference between the present formulation and that provided in~\cite{choi_stlspg}, we feel this supports our argument.

We are in complete agreement, however, with the fact that computing these Lipshitz constants is not practical. Doing so would be extremely difficult, if not impossible. We have added a remark to our analysis about this. 
}

\item I appreciate the authors extensive numerical experiments to study the performance/cost of the method, particularly with dt and the time window; however, there is some important information missing. For Section 6.1, the authors freeze the basis size and perform dt and dT sweeps, but provide no insight on how the various methods perform (accuracy and cost) as the basis size increases. Since the WLS approach is more expensive than Galerkin/LSPG for a given basis size, this basis size sweep will provide critical information regarding the relative performance (accuracy and cost) of the methods. To avoid running an exponential number of cases, the optimal dt and dT determined from the previous studies can be used. Similarly for Section 6.2: two basis sizes were considered but the energy only varies from 95\% to 97\%; many papers have shown higher energy (99\% or more) is required for complex problems like this, which should be done for a "fair" comparison to Galerkin/LSPG.

{\color{red} We appreciate these comments. In an effort to address them, we have added new sections to both our numerical experiments where we assess the performance of the method as a function of basis dimension. For the first numerical experiment, we do this for numerous window sizes, while for the second experiment we do it for a single window size. We additionally report CPU timings for all of these different permutations. With respect to the final comment on basis dimension, we had a typo in reporting the energy content, thank you for pointing this out. This has been corrected in the revised manuscript, and all bases considered contain $99.99\%$ of the energy.}

\item Neither of the examples vary parameters (purely reproducing the solution). While this was explicitly stated in the paper as not a priority, it significantly lessens relevance of the paper. I strongly recommend including a parametric example.
{\color{red} To address this concern, we added a new numerical experiment with varying parameters. We observe WLS to yield improvements over the LSPG approach.}

\end{enumerate}
\bibliographystyle{siam}
\bibliography{../refs}

\end{document}

### 2020-10-09 cavity.tex
latest updates

@@
-\subsection{Cavity flow}
-The next numerical example considers collocated ROMs\footnote{Collocation is a form of hyper-reduction which requires sampling the full-order model at only select grid points.} of a viscous, compressible flow in a two-dimensional cavity.
+\subsection{Supersonic cavity flow}
+The next numerical example considers collocated ROMs\footnote{Collocation is a form of hyper-reduction which requires sampling the full-order model at only select grid points.} of a viscous, supersonic compressible flow in a two-dimensional cavity.
 The flow is described by the two-dimensional compressible Navier-Stokes equations,
@@
 $$
-where $\mu \in \RRplus$ is the dynamic viscosity, $\text{Pr} = 0.72$ is the Prandtl number, and $c_p \in \RRplus$ the heat capacity ratio.
+where $\mu \in \RRplus$ is the dynamic viscosity, $\text{Pr} = 0.72$ is the Prandtl number, $T : \Omega \rightarrow \RRplus$ the temperature, and $c_p \in \RRplus$ the heat capacity ratio.
 We assume a Newtonian fluid, which leads to a viscous stress tensor of the form
@@
 The full-order model comprises a discontinuous-Galerkin discretization. We obtain the discretization 
-by partitioning the domain into $100$ elements in the flow direction and $40$ elements 
+by partitioning the domain into $200$ elements in the flow direction and $80$ elements 
 in the wall-normal direction. The discretization represents the solution to third order over each element using tensor product polynomials of order $p=2$, 
-resulting in $36000$ unknowns for each conserved variable. Spatial discretization via the discontinuous Galerkin method yields a dynamical system 
+resulting in $144000$ unknowns for each conserved variable. Spatial discretization via the discontinuous Galerkin method yields a dynamical system 
 of the form
-$$\frac{d \state}{dt} = \mass_{\text{DG}}^{-1} \velocity_{\text{DG}} (\state),$$
+$$  \mass_{\text{DG}}\frac{d \state}{dt} = \velocity_{\text{DG}} (\state),$$
 where $\mass_{\text{DG}} \in \RR{N \times N}$ is the (block diagonal) DG mass matrix and $\velocity_{\text{DG}}: \RR{N} \rightarrow \RR{N}$ is the DG velocity operator containing 
@@
 Figure~\ref{fig:cav_mesh} shows the computational mesh. The 
-DG method uses the Rusanov flux at the cell interfaces~\cite{rusanov} and uses the first form of Bassi and Rebay~\cite{br1} for the viscous fluxes. Time integration 
-is performed via a third-order strong stability preserving Runge-Kutta method~\cite{ssp_rk3} with a time step of $\Delta t = 0.001$.
+DG method uses the Rusanov flux at the cell interfaces~\cite{rusanov} and uses an interior penalty method for the viscous fluxes. Time integration 
+is performed via the Crank Nicolson method  with a time step of $\Delta t = 5 \times 10^{-3}.$\footnote{We note that the Newton solver in our code failed to converge for simulations using Crank Nicolson at a higher time step.} The nonlinear system arising at each time step is solved via a Jacobian-free general minimized residual (GMRES) algorithm~\cite{gmres,jfnk}.
 
-The reduced-order models leverage POD to construct the \spatialAcronym\ trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM velocity to select the sampling points (and as a result the weighting
+The reduced-order models leverage POD to construct the \spatialAcronym\ trial basis and use q-sampling~\cite{qdeim_drmac} based on snapshots of the FOM solution to select the sampling points (and as a result the weighting
 matrix $\stweightingMatArg{}$); we use a constant weighting matrix across all windows, e.g., $\stweightingMatArg{n} \equiv \stweightingMatArg{}$, $n=1,\ldots,\nslabs$.
-The process used to construct the initial conditions and trial subspaces for the ROMs is as follows:
+The process used to construct the initial conditions, trial subspaces, and weighting matrices for the ROMs is as follows:
 \begin{enumerate}
 \item Initialize the FOM with uniform free-stream conditions.
-\item Evolve the FOM for $t \in [0,400]$.
-\item Reset the time coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 100$ and $K = \{136,193\}$ over $t \in [0,100]$ to construct two trial subspaces comprising $\approx 95\%$ and $\approx 97\%$ of the snapshot energy, respectively.
-\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 100$, $n_s = 129$ to obtain the sampling point matrix of dimension $\stweightingMatOneArg{} \in \{0,1\}^{1603 \times \fomdim }$. 
+\item Evolve the FOM until a statistically steady state has been reached.
+\item Reset the time coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 8$, $K = \{50,100,156,200\}$, and $\podIP =  \mass_{\text{DG}}$ over $t \in [0,50]$ to construct the four trial subspaces tabulated in Table~\ref{tab:rom_basis_details}.
+\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 8$, $n_s = 250$ to obtain the sampling point matrix of dimension $\stweightingMatOneArg{} \in \{0,1\}^{10170 \times \fomdim }$. 
 \end{enumerate}
 Figure~\ref{fig:cav_sampmesh} shows the resulting sample mesh used in the ROM simulations.
-To depict the nature of the flow, Figure~\ref{fig:fom_sols_cav} shows snapshots of the vorticity field generated by the FOM for several time instances used in training.
+To depict the nature of the flow, Figure~\ref{fig:fom_sols_cav} shows snapshots of the vorticity and density gradient fields generated by the FOM for several time instances used in training.
 \begin{table}[]
@@
 \hline
-%1    & $136$ & $95.0\%$ & $580$ \\
-%2    & $193$ & $97.0\%$ & $580$ \\
-1    & $136$ & $95.0\%$ & $1603$ \\
-2    & $193$ & $97.0\%$ & $1603$ \\
+1    & $50$ &  $99.99738\%$ & $10170$ \\
+2    & $100$ & $99.99956\%$ & $10170$ \\
+3    & $156$ & $99.99990\%$ & $10170$ \\
+4    & $200$ & $99.99999\%$ & $10170$ \\
 \hline
@@
 \subsubsection{Description of reduced-order models}
-We consider collocated ROMs based on the Galerkin, least-squares Petrov--Galerkin, and \methodAcronym\ approaches. Details on the implementation of the methods is as follows:
+We consider collocated ROMs based on the least-squares Petrov--Galerkin and \methodAcronym\ approaches. Details on the implementation of the methods is as follows:
 \begin{itemize}
-\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation and evolve the ROM in time with the CN time scheme at a time step of $\Delta t =0.1$.
+%\item \textit{Galerkin ROM with collocation}: We consider a Galerkin ROM with collocation and evolve the ROM in time with the CN time scheme at a time step of $\Delta t =0.1$.
 
-\item \textit{LSPG ROM with collocation}: We consider a collocated LSPG ROM, which is built on top of the FOM discretization using the Crank--Nicolson scheme for temporal 
-discretization. We employ a time step of $\Delta t  = 0.1$. The implementation is the same as previously described.
+\item \textit{LSPG ROM with collocation}: We consider a collocated LSPG ROM, which is built on top of the FOM discretization using the CN scheme for temporal 
+discretization. We employ a time step of $\Delta t  = 0.05$. The implementation is the same as previously described.
 
-\item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the Crank--Nicolson time-discretization with a time step of 
-$\Delta t = 0.1$. The implementation is the same as previously described. 
+\item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the CN time discretization with a time step of 
+$\Delta t = 0.05$. The implementation is the same as previously described. 
 %\begin{itemize}
@@
 \begin{center}
-%\begin{subfigure}[t]{0.85\textwidth}
-\includegraphics[trim={2cm 7cm 4cm 6cm},clip,width=0.95\linewidth]{cav_fig.pdf}
-%\end{subfigure}
+\begin{subfigure}[t]{0.85\textwidth}
+\includegraphics[trim={2cm 5.5cm 1cm 7cm},clip,width=0.98\linewidth]{figs/cavity_new/cav_geom.pdf}
+\end{subfigure}
 \caption{Figure depicting geometry and flow conditions of the cavity flow problem.} 
@@
 \begin{center}
-\includegraphics[trim={0cm 14cm 0cm 14cm},clip,width=1.\linewidth]{figs/cavity/grid.png}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={2cm 2cm 4cm 2cm},clip,width=1.\linewidth]{figs/cavity_new/grid.png}
+\caption{Full computational mesh}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/sampleMesh.png}
+\caption{Close up of computational mesh with highlighted collocation cells.} 
+\label{fig:cav_sampmesh}
+\end{subfigure}
 \caption{Computational mesh employed in cavity flow simulations.}
@@
 
-\begin{figure} 
-\begin{center}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.65\linewidth]{figs/cavity/hyper_grid.png}
-\caption{Close up of computational mesh with highlighted collocation cells.} 
-\label{fig:cav_sampmesh}
-\end{center}
-\end{figure}
+%\begin{figure} 
+%\begin{center}
+%\caption{Close up of computational mesh with highlighted collocation cells.} 
+%
+%\end{center}
+%\end{figure}
 
@@
 \begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0000.png}
+\includegraphics[trim={15cm 7cm 16cm 20cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_anim0000.png}
 \caption{$t=0.0$}
 \end{subfigure}
+%\begin{subfigure}[t]{0.24\textwidth}
+%\includegraphics[trim={12cm 7cm 16cm 16cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_anim0010.png}
+%\caption{$t=2.0$}
+%\end{subfigure}
+%\begin{subfigure}[t]{0.24\textwidth}
+%\includegraphics[trim={12cm 7cm 16cm 16cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_anim0020.png}
+%\caption{$t=4.0$}
+%\end{subfigure}
 \begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0040.png}
-\caption{$t=4.0$}
+\includegraphics[trim={15cm 7cm 16cm 20cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_anim0030.png}
+\caption{$t=6.0$}
 \end{subfigure}
+%\begin{subfigure}[t]{0.85\textwidth}
 \begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0080.png}
-\caption{$t=8.0$}
+\includegraphics[trim={15cm 7cm 16cm 20cm},clip,width=1.0\linewidth]{figs/cavity_new/dense_anim0000.png}
+\caption{$t=0.0$}
 \end{subfigure}
+%\begin{subfigure}[t]{0.24\textwidth}
+%\includegraphics[trim={12cm 7cm 16cm 16cm},clip,width=1.0\linewidth]{figs/cavity_new/dense_anim0010.png}
+%\caption{$t=2.0$}
+%\end{subfigure}
+%\begin{subfigure}[t]{0.24\textwidth}
+%\includegraphics[trim={12cm 7cm 16cm 16cm},clip,width=1.0\linewidth]{figs/cavity_new/dense_anim0020.png}
+%\caption{$t=4.0$}
+%\end{subfigure}
 \begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0120.png}
-\caption{$t=12.0$}
+\includegraphics[trim={15cm 7cm 16cm 20cm},clip,width=1.0\linewidth]{figs/cavity_new/dense_anim0030.png}
+\caption{$t=6.0$}
 \end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0160.png}
-\caption{$t=16.0$}
-\end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={18cm 16cm 18cm 15cm},clip,width=1.0\linewidth]{figs/cavity/animate/sol0200.png}
-\caption{$t=20.0$}
-\end{subfigure}
-\caption{Vorticity snapshots from the FOM simulation at various time instances.} 
+\caption{Vorticity (top) and density gradient snapsots (bottom) from the FOM simulation at various time instances.} 
 \label{fig:fom_sols_cav}
@@
 
-\subsubsection{Numerical results}
-We first assess the performance of \methodAcronymROMs\ with varying window sizes. To this end, we consider \methodAcronymROMs\ with uniform window 
-sizes of $\Delta T^n \equiv \Delta T = 0.2,0.5,1.0,$ and $2.0$, along with the Galerkin and LSPG ROMs. We first consider results for basis \#1 as described in 
+\subsubsection{Numerical results for varying window size}
+We first assess the performance of \methodAcronymROMs\ with varying window sizes for a fixed basis size. To this end, we consider \methodAcronymROMs\ with uniform window 
+sizes of $\Delta T^n \equiv \Delta T = 0.1,0.25,0.5,$ and $1.0$, along with the LSPG ROM. We first consider results for basis \#3 as described in 
 Table~\ref{tab:rom_basis_details}. For all ROMs, we evolve 
-the solution for $t \in [0,100]$.
+the solution for $t \in [0,50]$.
 This comprises the same time interval used to construct the trial subspace.
 First, Figure~\ref{fig:cav_results1a} depicts the evolution of the pressure at the bottom wall in the midpoint of the computational domain, while Figure~\ref{fig:cav_results1b} depicts the evolution of the normalized $\elltwo$-error of the various reduced-order models.
-Both the collocated Galerkin and LSPG ROMs blow up/fail to converge within the first several time units.
-The \methodAcronymROM\ minimizing the residual over a window of size $\Delta T = 0.2$ also fails to converge.
-The \methodAcronymROMs\ that minimize the residual over window sizes of $\Delta T \ge 0.5$ are seen to all be stable and accurate; the pressure response is well characterized and the normalized state errors are less than $10\%$.
-The most notable discrepancy between the \methodAcronymROM\ and FOM solutions is a phase difference.
+We first observe that the collocated LSPG ROM fails to yield an accurate prediction for the pressure signal.
+We next observe the collocated WLS ROMs to yield more accurate solutions than LSPG.
+The pressure signal is well characterized by all WLS ROMs, and the $\ell^2$-errors are significantly lower.
+Figure~\ref{fig:cav_snapshots} shows vorticity and density gradient fields for the FOM, LSPG ROM, and \methodAcronymROMs\ with $\Delta T = 0.25,1.0$ for the final time instance $t = 50.0$.
+LSPG is observed to exhibit artificial oscillations, and poorly captures the flow state at the final time instane.
+The \methodAcronymROMs\ at $\Delta T = 0.25,1.0$ yield substantial improvements: both are able to capture the important features of the flow, including the structure of the shock at the end of the ramp, and remain accurate for the entire time interval.
 
@@
 
-\begin{subfigure}[t]{0.95\textwidth}
-\includegraphics[trim={0cm 3.0cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/pressure_basis2.pdf}
+\begin{subfigure}[t]{0.495\textwidth}
+\includegraphics[trim={0cm 00.0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/p_vs_t.pdf}
 \caption{Pressure} 
@@
 \end{subfigure}
-
-\begin{subfigure}[t]{0.95\textwidth}
-\includegraphics[trim={0cm 2.8cm 0cm 3cm},clip,width=1.\linewidth]{figs/cavity/error_basis2.pdf}
+\begin{subfigure}[t]{0.495\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/error_vs_t.pdf}
 \caption{Normalized $\elltwo$-error}
@@
 
-Figure~\ref{fig:cav_results2} shows the space--time error and objective function of the stable \methodAcronymROMs. We observe that growing the window size over 
-which the residual is minimized leads to a lower space--time residual, but not necessarily a lower $\elltwo$-error.
+\begin{figure}
+\begin{center}
+%\begin{subfigure}[t]{0.85\textwidth}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_fom_t50.png}
+\caption{FOM}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_wls_numStepsInWindow_1_t50.png}
+\caption{LSPG}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_wls_numStepsInWindow_5_t50.png}
+\caption{\methodAcronym, $\Delta T = 0.25$}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/vort_wls_numStepsInWindow_20_t50.png}
+\caption{\methodAcronym, $\Delta T = 1.0$}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/dens_fom_t50.png}
+\caption{FOM}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/dens_wls_numStepsInWindow_1_t50.png}
+\caption{LSPG}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/dens_wls_numStepsInWindow_5_t50.png}
+\caption{\methodAcronym, $\Delta T = 0.25$}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={10cm 3cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity_new/dens_wls_numStepsInWindow_20_t50.png}
+\caption{\methodAcronym, $\Delta T = 1.0$}
+\end{subfigure}
+\caption{Vorticity (a-d) and density gradient (e-h) snapshots from the FOM and ROM simulations at $t=50.0$.} 
+\label{fig:cav_snapshots}
+\end{center}
+\end{figure}
+
+Next, Figure~\ref{fig:cav_results2} shows the time-integrated $\elltwo$-errors and residual norms of the WLS ROMs as a function of the window size. We observe that growing the window size over 
+which the residual is minimized leads to a lower time-integrated residual, but not necessarily a lower $\elltwo$-error.
 This result is consistent with the previous numerical example.
+We find that, with respect to LSPG, WLS ROMs with window sizes of $\Delta T = 0.1, 0.25, 0.5,$  and $1.0$ are able to drive down the $\elltwo$-errors by approximately a factor of $3, 5.5, 4.5$, and $4$, respectively.
+Similarly, for the residual, WLS ROMs with window sizes of $\Delta T = 0.1, 0.25, 0.5,$  and $1.0$ are able to drive down the residual by approximately a factor of $13,57,65$, and $70$, respectively.
+
 Next, Figure~\ref{fig:cav_wallclock}
-shows the wall-clock times of the \methodAcronymROMs\ for $t \in [0,10]$ as compared to the LSPG ROM\footnote{We note that LSPG failed to converge at $t \approx 16.0$, so we focus on the first ten time units.}.
-As expected, increasing $\Delta T$ again leads to an increase in computational cost; minimizing the residual over a window comprising 20 time instances yields a 2.5x increase in cost over LSPG.
+shows the wall-clock times\footnote{We computed wall-clock times over the time inteval $t \in [0,1]$} of the \methodAcronymROMs\ as compared to (1) the LSPG ROM (Figure~\ref{fig:cav_wallclock_lspg}) and (2) the FOM (Figure~\ref{fig:cav_wallclock_fom}).
+As expected, increasing $\Delta T$ again leads to an increase in computational cost; minimizing the residual over a window comprising 20 time instances yields a 4x increase in cost over LSPG.
+With respect to the FOM, we find that WLS ROMs with window sizes of $\Delta T = 0.1, 0.25, 0.5,$  and $1.0$ are approximately $25$x, $16$x, $12$x, and $8$x faster, repectively.
+The LSPG ROM is approximately 30x faster than the FOM.
+We make several remarks about these CPU times.
+First, it must be noted that, due to a decreased stiffness in the system, we are able to run the ROMs at a time step that is 10x greater than the FOM.
+Second, we note that in our DG code, the vast majority of CPU time in the WLS ROMs is spent computing the Jacobians.
+For WLS ROMs emplying a sample mesh of size $n_s=10170$ with $K=156$ basis vectors and a window size of $\Delta T = 0.025$, it takes approximately $2$(s) to compute the residual over one time window; it takes $500$(s) to compute the Jacobian via automatic differentiation.
+Our full-order model solver uses a Jacobian-free GMRES solver, and thus is not subject to this bottleneck.
+While this is in part due to the fact that high-order DG methods have dense block Jacobians (each element has 45 DOFs associated with it for a $p=2$ discretization of the Navier--Stokes equations), we expect that significant computational speedups can be obtained by employing, e.g., analytic Jacobians, matrix-free Jacobian transpose methods~\cite{doi:10.2514/6.2016-0833}.
+This will be a topic of future work.
+Nonetheless, the present example demonstrates some of the practical challenges that exist for least-squares ROMs such as WLS and LSPG.
 
@@
 \begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window_basis2.pdf}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/walltime_vs_windowSize.pdf}
+\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM.}
+\label{fig:cav_wallclock_lspg}
+\end{subfigure}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/rel_walltime_vs_windowSize.pdf}
+\caption{Wall-clock times of \methodAcronymROMs\ with respect to the full-order model.}
+\label{fig:cav_wallclock_fom}
+\end{subfigure}
+\caption{Relative wall-clock times of \methodAcronymROMs.}
+\label{fig:cav_wallclock}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/error_vs_windowSize.pdf}
 \caption{Space--time $\elltwo$-error}
@@
 \begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/objective_vs_window_basis2.pdf}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/residual_vs_windowSize.pdf}
 \caption{Objective function} 
@@
 
+\subsubsection{Numerical results as a function of basis dimension}
+We next consider the impact of the basis dimension on the ROM performance. We consider WLS ROMs with a uniform window
+size of $\Delta T^n \equiv \Delta T = 0.25$, along with LSPG ROMs. For both WLS and LSPG ROMs, we consider results for bases of 
+dimension $K=50,100,156$, and $200$, as described in Table~\ref{tab:rom_basis_details}. Figure~\ref{fig:cav_results_basisdim}
+presents results for the time-integrated $\elltwo$-error and residual norm as a function of basis dimension.
+Figure~\ref{fig:cav_results_pressure} presents the $\elltwo$-error in the pressure at the midpoint of the bottom wall, as well as the pressure signal in the same location.
+We observe that, for WLS, increasing the basis dimension monotinically decreases the $\elltwo$-error, residual norm, and $\elltwo$-error in the pressure.
+Increasing the basis dimension does not yield monotonically decreasing error metrics for LSPG.
+We find that LSPG yields a stable, although inaccurate, prediction at $K=50$, but then yields an unstable prediction at $K=100$.
+Increasing the basis dimension to $K=156$ and $K=200$ results in a stable, but still inaccurate solution.
+We note that not a single LSPG ROM yields a qualitatively accurate pressure signal for the entire time domain, while the WLS ROM is able to achieve this with a basis dimension of $K=100$.
 \begin{figure}
 \begin{center}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=0.49\linewidth]{figs/cavity/walltime_vs_window_compare_basis2.pdf}
-\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM.}
-\label{fig:cav_wallclock}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/error_vs_basisSize.pdf}
+\caption{Space--time $\elltwo$-error}
+\label{fig:cav_results_basisdima}
+\end{subfigure}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/residual_vs_basisSize.pdf}
+\caption{Objective function} 
+\label{fig:cav_results_basisdimb}
+\end{subfigure}
 \end{center}
+\caption{Space--time error (left) and objective function (right) as a function of window size.}
+\label{fig:cav_results_basisdim}
 \end{figure}
 
-Figure~\ref{fig:cav_snapshots} shows vorticity fields for the FOM, LSPG ROM, and \methodAcronymROMs\ with $\Delta T = 0.5,1.0$ for the time instance $t = 5.0$.
-LSPG is observed to exhibit artificial oscillations; the Gauss-Newton method fails to converge at $t \approx 16.0$.
-The \methodAcronymROM\ at $\Delta T = 0.5$ is able to capture the important features of the flow, including the points of flow separation at the start and end of the ramp, and remains stable for the entire time interval.
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/perror_vs_basisSize.pdf}
+\caption{Objective function} 
+\label{fig:cav_results_pressurea}
+\end{subfigure}
+\begin{subfigure}[t]{0.54\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/p_vs_basisSize.pdf}
+\caption{Space--time $\elltwo$-error}
+\label{fig:cav_results_pressureb}
+\end{subfigure}
+\end{center}
+\caption{Space--time error (left) and objective function (right) as a function of window size.}
+\label{fig:cav_results_pressure}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-%\begin{subfigure}[t]{0.85\textwidth}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/fom.png}
-\caption{FOM}
-\end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/lspg.png}
-\caption{LSPG}
-\end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/wls_window5.png}
-\caption{\methodAcronym, $\Delta T = 0.5$}
-\end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics[trim={10cm 7cm 10cm 12cm},clip,width=1.0\linewidth]{figs/cavity/wls_window10.png}
-\caption{\methodAcronym, $\Delta T = 1.0$}
-\end{subfigure}
-\caption{Vorticity snapshots from the FOM and ROM simulations at $t=5.0$.} 
-\label{fig:cav_snapshots}
-\end{center}
-\end{figure}
+\begin{comment}
 
@@
 
-Finally, Figure~\ref{fig:cav_wallclock} shows the wall-clock times of the \methodAcronymROMs\ as compared to the LSPG ROMs for basis \#2.
+Finally, Figure~\ref{fig:cav_wallclock} shows the wall-clock times for $t \in [0,4]$ of the \methodAcronymROMs\ as compared to the LSPG ROMs for basis \#2.
 Increasing the window size again leads to an increase in computational cost.
 Minimizing the residual over a window comprising 20 time instances yields a 3x increase in cost over LSPG.
-This increase in computational cost is particularly reasonable in this example as LSPG yielded an unstable solution.
 \begin{figure}
@@
 
-%\begin{figure}
-%\begin{center}
-%\begin{subfigure}[t]{0.45\textwidth}
-%\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/error_vs_window.pdf}
-%\caption{Integrated $L^2$ error}
-%\label{fig:cav_results2a_basis1}
-%\end{subfigure}
-%\begin{subfigure}[t]{0.45\textwidth}
-%\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity/objective_vs_window.pdf}
-%\caption{Objective function} 
-%\label{fig:cav_results2b_basis1}
-%\end{subfigure}
-%\end{center}
-%\caption{Integrated error (left) and objective function (right) as a function of window size.}
-%\label{fig:cav_results2_basis1}
-%\end{figure}
+\end{comment}
+
+\begin{comment}
+
+\subsection{Supersonic cavity flow with varying Reynolds number and time-local bases}
+Lastly, we consider the supersonic cavity flow presented above with a varying Reynolds number over the domain $\text{Re} \in \mathsf{U}(5000,15000)$.
+We again consider colllocated reduced-order models based on the LSPG and WLS approaches leveraging \spatialAcronym\ trial subspaces.
+The ROMs use the same collocation points generated in the previous example.
+As opposed to the previous example, however, here we consider time-local subspaces.
+The process used to construct these subspaces is as follows:
+\begin{enumerate}
+\item Initialize the FOM with uniform free-stream conditions.
+\item Evolve the FOM for $t \in [0,500]$ at $\text{Re} = 10000$. (Same as step 2 described previously).
+\item Reset the time coordinate, $0 \leftarrow t$, and set parameter samples to be $\text{Re} = 5000, 7500, 10000,12500,15000$.
+\item Execute Algorithm~\ref{alg:pod_tlp} ten different times --- with the inputs descibed in Table~\ref{tab:} --- to construct 10 trial subspaces: one for $t \in [0,5)$, one for $t \in [5,10)$, ..., one for $t \in [45,50]$.
+We denote these trial susbpaces as $\spaceTrialSpace_{\text{S1}}, \ldots, \spaceTrialSpace_{\text{S10}}.$
+\end{enumerate}
+
+To demonstrate the impact of Reynolds number on the flow, Figure~\ref{fig:cav_param_dens} shows FOM solutions for the Re=$5000$ and $15000$ cases at $t=50$.
+Similarly, Figure~\ref{cav_param_ptrain} presents the pressure signal at the midpoint of the bottom wall for all training simulations.
+We observe that increasing the Reynolds number leads to an increase in unsteadiness in the shock location, increases the frequency of, e.g., the pressure signal, and in general leads to more fine-scale features in the flow.
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={4cm 8cm 0cm 4cm},clip,width=1.\linewidth]{figs/cavity_new/re5000_vort_t50.png}
+\caption{Re=$5000$}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={4cm 8cm 0cm 4cm},clip,width=1.\linewidth]{figs/cavity_new/re15000_vort_t50.png}
+\caption{Re=$15000$}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={4cm 8cm 0cm 4cm},clip,width=1.\linewidth]{figs/cavity_new/re5000_dens_t50.png}
+\caption{Re=$5000$}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={4cm 8cm 0cm 4cm},clip,width=1.\linewidth]{figs/cavity_new/re15000_dens_t50.png}
+\caption{Re=$15000$}
+\end{subfigure}
+\end{center}
+\caption{Density gradient fields at $t=50$ for the Re=$5000$ (left) and Re=$15000$ (right) training runs.}
+\label{fig:cav_param_dens}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/pfom_param.pdf}
+\end{subfigure}
+\end{center}
+\caption{Pressure signal at the midpoint of the bottom wall for all training simulations}
+\label{fig:cav_param_ptrain}
+\end{figure}
+\end{comment}

### 2020-10-09 main.tex
latest updates

@@
 
-Researchers have pursued the development of space--time
-residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
+Researchers have pursued the development of space--time residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
 address the issues incurred by sequential residual minimization in time.
@@
 this work shares some thematic similarities with
-Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
+Ref.~\cite{constantine_strom}.  Additionally, Ref.~\cite{choi_stlspg} develops a
 space--time extension of LSPG projection that minimizes the time-discrete
@@
 arbitrary time windows, and (3) this work also considers trial subspaces
-associated with spatial dimension reduction only.
+associated with spatial dimension reduction only. Lastly, we note that the windowing approach proposed here  
+displays commonalities with space--time methods employing space--time slabs, e.g., Refs.\cite{HUGHES1988339,PIRONNEAU1992117,HUGHES1996217,HUGHES1989173}. 
+In these approaches, the space--time domain is typically decomposed into a series of space--time slabs. A system of algebraic equations is then formed over each slab 
+via, e.g., the discontinuous Galerkin method, and the space--time method operates by sequentially marching through these space--time slabs.
+From this perspective, the proposed windowing approach is best viewed as a space--time method in where each window comprises multiple time-slabs.
+The time-slabs within each window experience a two-way coupling to each other, by virtue of the residual minimization statement being defined over the entire window, and thus one can no longer sequentially march over the time-slabs.
 
@@
 	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
-		model (\spatialAcronym)}.
+		model (\spatialAcronym)}\footnote{In this work, we refer to the ``state-space" dimension of an ordinary differential equation as the ``spatial" dimension.
+We use this notation due to the common application of projection-based ROMs to semi-discretized partial differential equations.
+We emphasize, however, that the presented approach is applicable to ordinary differential equations that do not associate with some type of spatial discretization, i.e., as in molecular dynamics.}
+.
 These trial subspaces are characterized by a spatial projection operator, associate with a basis that represents the spatial dependence of the state, and are employed in classic model reduction approaches, e.g., Galerkin and LSPG.
@@
 \begin{algorithm}
-\caption{Algorithm for generating POD Basis.}
+\caption{Algorithm for generating POD basis.}
 \label{alg:pod}
-\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$ \; 
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$; weighting matrix, $\mathbf{P}$ \; 
 %\newline
-\textbf{Output:} POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
+\textbf{Output:} POD Basis $\basisspace \in \{ \mathbf{X} \in \RR{\fomdim \times \romdim} | \mathbf{X}^T\mathbf{P} \mathbf{X} = \mathbf{I} \} $ \;
 %\newline
@@
 $$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} - \stateInterceptArg{n} & \stateFOMDiscreteArg{N_{\text{skip}}} - \stateInterceptArg{n} & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} - \stateInterceptArg{n} \end{bmatrix}$$
-    \item Compute the thin singular value decomposition, $$\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
-    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
-    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_K \end{bmatrix}$
+    \item Compute the thin singular value decomposition, $$ \sqrt{\mathbf{P}}^T\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
+    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$ and $\sqrt{\mathbf{P}}$ is the Cholesky decomposition of $\mathbf{P}$, i.e., $\sqrt{\mathbf{P}} \sqrt{\mathbf{P}}^T = \mathbf{P}$
+    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} [\sqrt{\mathbf{P}}^T]^{-1}  \mathbf{u}_1 & \cdots & [\sqrt{\mathbf{P}}^T]^{-1} \mathbf{u}_K \end{bmatrix}$
 \end{enumerate}
@@
 \end{algorithm}
+
+%In the final numerical experiment, time-local trial bases are constructed for a set of varying input parameters. Algorithm~\ref{alg:pod_tlp} describes the construction of the POD bases for this case.
+%\begin{algorithm}
+%\caption{Algorithm for generating time-local POD basis over multiple parameter instances}
+%\label{alg:pod_tlp}
+%\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; starting time-instance, $N_s$; number of instances $N_{i}$; intercept $\stateInterceptArg{}$; parameter instances $\param_{1} , \ldots, \param_{n_{\text{train}}}$; basis dimension $K$ \\
+%%\newline
+%\textbf{Output:} POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
+%%\newline
+%\textbf{Steps:}
+%\begin{enumerate}
+%    \item Solve FOM O$\Delta$E for $\param_1, \ldots, \param_{n_{\text{train}}}$ and collect solutions into snapshot matrix
+%\small{$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteParamArg{N_s}{\param_1} - \stateInterceptArg{n} & \cdots & \stateFOMDiscreteParamArg{N_s + N_i N_{\text{skip}}}{\param_1}- \stateInterceptArg{n} & \cdots & \stateFOMDiscreteParamArg{N_s + N_i N_{\text{skip}}}{\param_{n_{\text{train}}}}- \stateInterceptArg{n} \end{bmatrix}$$}
+%    \item Compute the thin singular value decomposition, $$\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
+%    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
+%    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_K \end{bmatrix}$
+%\end{enumerate}
+%\end{algorithm}
 
@@
     \item Solve FOM O$\Delta$E and collect velocity snapshots 
-$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \velocity (\stateFOMDiscreteArg{0})  & \velocity( \stateFOMDiscreteArg{N_{\text{skip}}}) & \cdots & \velocity( \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} ) \end{bmatrix}$$
+%$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \velocity (\stateFOMDiscreteArg{0})  & \velocity( \stateFOMDiscreteArg{N_{\text{skip}}}) & \cdots & \velocity( \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} ) \end{bmatrix}$$
+$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} & \stateFOMDiscreteArg{N_{\text{skip}}}  & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} \end{bmatrix}$$
     \item Compute the thin singular value decomposition, $$\mathbf{F} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
@@
     \item Augment the sampling point matrix, $\stweightingMatOneArg{}$, with additional columns such that all unknowns are computed at the mesh cells selected by Step 3.
-For the second numerical experiment, these additional unknowns correspond to each conserved variable and quadrature point in the selected cells.
+For the second numerical experiment, these additional unknowns correspond to each conserved variable and degree of freedom in the selected cells.
 \end{enumerate}

### 2020-10-09 numerical_experiments.tex
latest updates

@@
 We evolve the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$.
-We construct the \spatialAcronym\ trial subspace by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $.
+We construct the \spatialAcronym\ trial subspace by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $, and $\mathbf{P} = \mathbf{I}$.
 The resulting trial subspace corresponds to an energy criterion of $99.99\%$.
@@
 We first assess the impact of the window size on the performance of the \methodAcronymROMs.  We consider a set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
-$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$.
+$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  0.002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$.
 We additionally consider the standard Galerkin and LSPG ROMs.
@@
 
-Next, Figure~\ref{fig:sod_error} shows space--time state errors and the objective function~\eqref{eq:obj} (i.e., the space--time residual norm) over $t \in [0,1]$ for the various ROMs.
+Next, Figure~\ref{fig:sod_error} shows time-integrated $\ell^2$-norm of the state errors and the objective function~\eqref{eq:obj} (i.e., the time-integrated residual norm) over $t \in [0,1]$ for the various ROMs.
 Most notably, we observe
-that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the space--time error as measured in the $\elltwo$-norm (we refer to this as the $\elltwo$-error).
-As expected, increasing the window size does lead to a monotonic decrease in the space--time residual, however.
-We additionally note that, although the space--time
-$\elltwo$-error of 
-the projected FOM solution is significantly lower than that of the various ROM solutions, the space--time residual norm of the projected FOM solution is \textit{higher} 
+that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the error as measured in the time-integrated $\elltwo$-norm (we refer to this as the $\elltwo$-error).
+As expected, increasing the window size does lead to a monotonic decrease in the time-integrated residual norm, however.
+We additionally note that, although the time-integrated $\elltwo$-error of
+the projected FOM solution is significantly lower than that of the various ROM solutions, the time-integrated residual norm of the projected FOM solution is \textit{higher} 
 than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$-norm, it does not satisfy the governing equations as well the ROM solutions.
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_compare.pdf}
-\caption{Space--time $\elltwo$-error of various ROMs.}
+\caption{Relative time-integrated $\elltwo$-error of various ROMs.}
 \label{fig:sod_error_a}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_compare.pdf}
-\caption{Integrated residual norm of various ROMs.} 
+\caption{Time-integrated residual norm of various ROMs.} 
 \label{fig:sod_error_b}
 \end{subfigure}
-\caption{Integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.  Note that the Galerkin and LSPG ROMs do not depend on the window size.} 
+\caption{Time-integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.  Note that the Galerkin and LSPG ROMs do not depend on the window size.} 
 \label{fig:sod_error}
@@
 Figure~\ref{fig:sod_error_methods} 
-shows the same space--time $\elltwo$-errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
+shows the same time-integrated $\elltwo$-errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
 the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, we observe that the \methodAcronym\ method 
@@
 In particular, ROMs using the second order explicit AB2 scheme with a time step of $\Delta t = 0.0005$ provide similar results to the 
-ROMs using the second-order CN scheme at a time step of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal 
-$\elltwo$-error occurs when the window size is $\DeltaSlabArg{} = 0.1$, and the residual decreases monotonically as the window size grows. 
+ROMs using the second-order CN scheme at a time step of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal time-integrated 
+$\elltwo$-error occurs when the window size is $\DeltaSlabArg{} = 0.1$, and the time-integrated residual norm decreases monotonically as the window size grows. 
 \begin{figure}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window.pdf}
-\caption{Integrated $\elltwo$-error of the \methodAcronymROMs.}
+\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROMs.}
 \label{fig:sod_error_methods_a}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window.pdf}
-\caption{Integrated residual norm of the \methodAcronymROMs.} 
+\caption{Time-integrated residual norm of the \methodAcronymROMs.} 
 \label{fig:sod_error_methods_b}
 \end{subfigure}
-\caption{Integrated performance metrics of various \methodAcronymROMs.} 
+\caption{Time-integrated performance metrics of various \methodAcronymROMs.} 
 \label{fig:sod_error_methods}
@@
 To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$.
-Figure~\ref{fig:time_step_study} shows the $\elltwo$-error and residual norm
-of the various ROMs. We observe that the $\elltwo$-error and residual norm 
+Figure~\ref{fig:time_step_study} shows the time-integrated $\elltwo$-error and residual norm
+of the various ROMs. We observe that the time-integrated $\elltwo$-error and residual norm 
 of the \methodAcronymROMs\ decrease and converge as the time step decreases.
-This is in direct contrast to the LSPG ROMs, in where the $\elltwo$-error and residual norm display a complex dependence on the time step.
+This is in direct contrast to the LSPG ROMs, in where the time-integrated $\elltwo$-error and residual norm display a complex dependence on the time step.
 This result
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
-\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
+\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
 \label{fig:sod_error_methods_a}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
-\caption{Space--time residual norm of the \methodAcronymROM.} 
+\caption{Time-integrated residual norm of the \methodAcronymROM.} 
 \label{fig:sod_error_methods_b}
@@
 
-Figure~\ref{fig:convergence_study} shows the integrated $\ell^2$-error of the various ROM solutions, as well as the optimal solution as measured in the $\ell^2$-norm.
-In Figure~\ref{fig:sod_error_converge} we first observe that for a basis of dimension $\romdim = 10$, the Galerkin, LSPG, and WLS $\Delta T = 0.002$, $\Delta T = 0.004$ ROMs do not converge.
+Figure~\ref{fig:convergence_study} shows the time-integrated $\ell^2$-error of the various ROM solutions, as well as that of the optimal solution as measured in the $\ell^2$-norm (i.e., the orthogonal projection of the FOM solution onto the trial subspace).
+In Figure~\ref{fig:sod_error_converge} we first observe that for a basis of dimension $\romdim = 10$, the Galerkin, LSPG, and WLS ($\Delta T = 0.002$, $\Delta T = 0.004$) ROMs do not converge.
 Upon investigation, this lack of convergence occurred due to the presence of negative pressure.
-We next observe that the $\ell^2$-error monotonically decreases as the basis dimension increases for all remaining ROM solutions with the exception of the Galerkin, LSPG, and WLS $\Delta T = 0.002$, $\Delta T = 0.004$ at dimension $\romdim = 30$.
+We next observe that the time-integrated $\ell^2$-error monotonically decreases as the basis dimension increases for all remaining ROM solutions with the exception of the Galerkin, LSPG, and WLS ($\Delta T = 0.002$, $\Delta T = 0.004$) ROMs at the dimension $\romdim = 30$.
 Overall, we observe that larger window sizes yield lower errors when the basis dimension is small.
 For larger basis dimensions we observe all ROMs to yield accurate solutions, and a larger window size does not always yield a lower error.
-In Figure~\ref{fig:sod_residual_converge}, we next observe that the residual monotonically decreases as the basis dimension grows.
-This result is supported, although not guaranteed, by the error analysis presented in Section~\ref{sec:analysis} and Ref.~\cite{carlberg_lspg_v_galerkin}.
-Finally, we observe that larger window sizes yields lower integrated residuals.
+In Figure~\ref{fig:sod_residual_converge}, we next observe that the time-integrated residual norm monotonically decreases as the basis dimension grows.
+This result is supported, although not guaranteed, by the error analysis presented in Section~\ref{sec:analysis} (and Ref.~\cite{carlberg_lspg_v_galerkin} for LSPG).
+Finally, we observe that larger window sizes yield lower time-integrated residuals.
 
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_converge.pdf}
-\caption{Space--time $\elltwo$-error of the \methodAcronymROM.}
+\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
 \label{fig:sod_error_converge}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/residual_converge.pdf}
-\caption{Space--time residual norm of the \methodAcronymROM.} 
+\caption{Time-integrated residual norm of the \methodAcronymROM.} 
 \label{fig:sod_residual_converge}
+\end{subfigure}
+\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
+\label{fig:convergence_study}
+\end{center}
+\end{figure}
+
+Next, Figure~\ref{fig:convergence_study_pareto} presents Pareto plots for the time-integrated $\elltwo$-error and residual as a function of wall-clock time for ROMs with varying basis dimensions and window sizes.
+We observe intermediate window sizes (i.e., $\Delta T = 0.02,0.04$) to be Pareto-optimal for the error (i.e., lowest CPU time for a given error tolerance) for all runs with the exception of the $K=50$ case, in which case window sizes of $\Delta T = 0.002$ (i.e., LSPG) and $\Delta T = 0.004$ are Pareto-optimal.
+We next observe that intermediate window sizes are Pareto-optimal for the integrated residual norm for all cases, with the exception of runs using $\Delta T = 0.02, 0.04$, and $K=20$.
+Summing across the Pareto plots for the time-integrated error and residual, we find that --- for the window sizes and bases considered --- $\Delta T = 0.002$ (i.e., LSPG) is Pareto optimal three times, $\Delta T = 0.004$ one time, $\Delta T = 0.02$ 2 times, $\Delta T = 0.04$ 6 times, and $\Delta T = 0.2$ two times.
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[width=1.\linewidth]{figs/sod/pareto_error.pdf}
+\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
+\label{fig:sod_error_converge}
+\end{subfigure}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[width=1.\linewidth]{figs/sod/pareto_resids.pdf}
+\caption{Time-integrated residual norm of the \methodAcronymROM.} 
+\label{fig:sod_residual_converge}
+\end{subfigure}
+\caption{Pareto fronts for the time-integrate $\elltwo$-error (left) and time-integrated residual norm (right) as a function of wall-clock time.} 
+\label{fig:convergence_study_pareto}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.45\textwidth}
+\includegraphics[width=1.\linewidth]{figs/sod/converge_walltimes.pdf}
+\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
+\label{fig:sod_error_converge}
 \end{subfigure}

### 2020-10-12 response_to_reviewers/response.tex
latest

@@
 
-{\color{red} We have added a new sections to both of our first numerical experiment to address this question.
+{\color{red} We have added a new sections to both of our numerical experiments to address this question.
+We additionally examine convergence in the new example we added.
 We show that the behavior of the method is as one would expect: increasing the basis dimension yields both lower errors and lower residuals.
@@
 Also figure 10 should have a comparison also with the FOM wall-clock times.
-{\color{red} We have added CPU timings to our second numerical experiment to address this concern.
+{\color{red} We have added CPU timings to our second and third numerical experiments to address this concern.
 In our first numerical experiment, which serves as a reproducible benchmark, we do not employ hyper-reduction, and therefore do not obtain computational savings.
@@
 As LSPG has been shown to yield significant computational speedups in previous work (e.g.,~\cite{carlberg_lspg_v_galerkin}), and WLS with S-reduction subspaces is a windowed extension of the LSPG approach, we felt that comparing the speed of the proposed method to LSPG is the relevant timing.
-However, we do agree that comparison to the FOM walltime is a metric of interest, and therefore we have worked hard to integrate it into the second example.}
+However, we do agree that comparison to the FOM walltime is a metric of interest, and therefore we have worked hard to integrate it into the second and third examples.
+For the problems considered, WLS gives between a 2x and a 20x speedup over the FOM, depending on the basis dimension and window size.
+We also investigate the performance of freezing the Jacobian in the Gauss--Newton algorithm on the solution speed.}
 
@@
 
-{\color{red} To address this concern, we have (1) increased the difficulty of the second numerical experiment by raising the mach number such that is supersonic, thus adding a complex shock boundary layer interaction to the problem, and (3) added a new numerical experiment with varying parameters.
-In both cases, we observe WLS to yield improvements over the LSPG approach.}
+{\color{red} To address this concern, we have (1) increased the difficulty of the second numerical experiment by raising the Mach number such that is supersonic, thus adding a complex shock boundary layer interaction to the problem, and (2) added a new numerical experiment, which starts at 6.3, with varying parameters.
+With respect to the new parameterized experiment, in addition to presenting various ROMs, we demonstrate that a linear surrogate model, in where we build a multivariate linear model for each degree of freedom of the dynamical system, fails to yield an adequate response.
+We appreciate the reviewers remarks, as we feel that the new numerical results provide stronger evidence of the utility of the approach.}
 
@@
 \item page 25 second line, $\Delta T = .002 \rightarrow  \Delta T = 0.002$
+
 {\color{red} Thank you, good catch!}
@@
 
-{\color{red} We have rephrased this as the ``time-integrated residual norm".
-By integrated norm, we were referring to the Euclidean $\ell^2$-norm of the state integrated in time, as denoted in, e.g., Figure 6a.
+{\color{red} Thank you for drawing this to our attention.
+To clarify, we have (1) rephrased this as the ``time-integrated residual norm" and (2) added the clarifying statement" ``(i.e., the time-integrated norm of the residual evaluated at the $\elltwo$-orthogonal projection of the FOM solution onto the trial subspace)".
 In the previous manuscript, we also had an inconsistency of using both the term ``space--time" error norm and "integrated" error norm to refer to the same thing; we have corrected this in the revision by replacing space--time error norm with time-integrated norm.}
@@
 However, the two concepts are very closely tied and its easy to make a counter argument to our perspective!
-We have added the following to the manuscript, at the end of page 2 in or related work sections:
+We have added the following to the manuscript, at the end of page 2 in or related work section:
 
@@
 
-{\color{red} In our revision, we have removed this assumption by slightly modifying our analysis.
+{\color{red} We appreciate this criticism, and agree with it.
+In our revision, we have removed this assumption by slightly modifying our analysis.
 We now place a bound on the in-plane error, and the constant multiplying the residual term now has an additional factor of two.
-The main take away of  the analysis remains the same.}
+The main take away of  the analysis remains the same, in that the error grows recursively with respect to the number of windows, opposed to the number of time steps.}
 
@@
 {\color{red} We appreciate these comments.
-In an effort to address them, we have added new sections to both our numerical experiments where we assess the performance of the method as a function of basis dimension.
+In an effort to address them, we have added new sections to all of our numerical experiments where we assess the performance of the method as a function of basis dimension.
 For the first numerical experiment, we do this for numerous window sizes, while for the second experiment we do it for a single window size.
+We do it for numerous window sizes for our third experiment.
 We additionally report CPU timings for all of these different permutations.
@@
 I strongly recommend including a parametric example.
-{\color{red} To address this concern, we added a new numerical experiment with varying parameters. We observe WLS to yield improvements over the LSPG approach.}
+{\color{red} To address this concern, we added a new numerical experiment with varying parameters.
+In this new experiment, we explore the performance of LSPG and WLS for prediction at a parameter outside the training set.
+We study the impact of window size and basis dimension on the results.
+In addition to presenting various ROMs, we demonstrate that a linear surrogate model, in where we build a multivariate linear model for each degree of freedom of the dynamical system, fails to yield an adequate response.
+We observe WLS to yield improvements over the LSPG approach.}
 

### 2020-10-12 cavity.tex
latest

@@
 \item Reset the time coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 8$, $K = \{50,100,156,200\}$, and $\podIP =  \mass_{\text{DG}}$ over $t \in [0,50]$ to construct the four trial subspaces tabulated in Table~\ref{tab:rom_basis_details}.
-\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 8$, $n_s = 250$ to obtain the sampling point matrix of dimension $\stweightingMatOneArg{} \in \{0,1\}^{10170 \times \fomdim }$. 
+\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 8$, $n_s = 250$ to obtain the sampling point matrix of dimension $\mathbf{S} \in \{0,1\}^{10170 \times \fomdim }$.
+\item Set $\stweightingMatOneArg{} = \mathbf{S} \mass_{\text{DG}}$. 
 \end{enumerate}
@@
 \item \textit{LSPG ROM with collocation}: We consider a collocated LSPG ROM, which is built on top of the FOM discretization using the CN scheme for temporal 
-discretization. We employ a time step of $\Delta t  = 0.05$. The implementation is the same as previously described.
+discretization. We employ a time step of $\Delta t  = 0.05$. The implementation is the same as previously described in Section~\ref{sec:sod_rom_description}.
 
 \item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the CN time discretization with a time step of 
-$\Delta t = 0.05$. The implementation is the same as previously described. 
-%\begin{itemize}
-%\item Direct Method: \methodAcronym ROMs solved via the direct method are again implemented with the same Crank--Nicolson time discretization with a nominal 
-%time-step size of $\Delta t = 0.1$. 
-%\end{itemize}
+$\Delta t = 0.05$. The implementation is the same as previously described in Section~\ref{sec:sod_rom_description}. 
 \end{itemize}
@@
 Next, Figure~\ref{fig:cav_wallclock}
-shows the wall-clock times\footnote{We computed wall-clock times over the time inteval $t \in [0,1]$} of the \methodAcronymROMs\ as compared to (1) the LSPG ROM (Figure~\ref{fig:cav_wallclock_lspg}) and (2) the FOM (Figure~\ref{fig:cav_wallclock_fom}).
+shows the wall-clock times\footnote{We computed wall-clock times over the time interval $t \in [0,1]$} of the \methodAcronymROMs\ as compared to (1) the LSPG ROM (Figure~\ref{fig:cav_wallclock_lspg}) and (2) the FOM (Figure~\ref{fig:cav_wallclock_fom}).
 As expected, increasing $\Delta T$ again leads to an increase in computational cost; minimizing the residual over a window comprising 20 time instances yields a 4x increase in cost over LSPG.
-With respect to the FOM, we find that WLS ROMs with window sizes of $\Delta T = 0.1, 0.25, 0.5,$  and $1.0$ are approximately $25$x, $16$x, $12$x, and $8$x faster, repectively.
+With respect to the FOM, we find that WLS ROMs with window sizes of $\Delta T = 0.1, 0.25, 0.5,$  and $1.0$ are approximately $25$x, $16$x, $12$x, and $8$x faster, respectively.
 The LSPG ROM is approximately 30x faster than the FOM.
@@
 Second, we note that in our DG code, the vast majority of CPU time in the WLS ROMs is spent computing the Jacobians.
-For WLS ROMs emplying a sample mesh of size $n_s=10170$ with $K=156$ basis vectors and a window size of $\Delta T = 0.025$, it takes approximately $2$(s) to compute the residual over one time window; it takes $500$(s) to compute the Jacobian via automatic differentiation.
+For WLS ROMs employing a sample mesh of size $n_s=10170$ with $K=156$ basis vectors and a window size of $\Delta T = 0.025$, it takes approximately $2$(s) to compute the residual over one time window; it takes $500$(s) to compute the Jacobian via automatic differentiation.
 Our full-order model solver uses a Jacobian-free GMRES solver, and thus is not subject to this bottleneck.
-While this is in part due to the fact that high-order DG methods have dense block Jacobians (each element has 45 DOFs associated with it for a $p=2$ discretization of the Navier--Stokes equations), we expect that significant computational speedups can be obtained by employing, e.g., analytic Jacobians, matrix-free Jacobian transpose methods~\cite{doi:10.2514/6.2016-0833}.
-This will be a topic of future work.
+While this is in part due to the fact that high-order DG methods have dense block Jacobians (each element has 45 DOFs associated with it for a $p=2$ discretization of the Navier--Stokes equations), we expect that significant computational speedups can be obtained by employing, e.g., frozen Jacobians, analytic Jacobians, matrix-free Jacobian transpose methods~\cite{doi:10.2514/6.2016-0833}.
+We briefly investigate the use of frozen Jacobians in the following numerical experiment and Appendix~\ref{appendix:gnupdate}, where we show that employing a Gauss--Newton algorithm with frozen Jacobians can yield substantial speedups.
 Nonetheless, the present example demonstrates some of the practical challenges that exist for least-squares ROMs such as WLS and LSPG.

### 2020-10-12 main.tex
latest

@@
 
-Numerical experiments of the compressible Euler and Navier--Stokes equations demonstrated the utility in the proposed approach.
+Across three numerical experiments conducted on the compressible Euler equations, compressible Navier--Stokes equations, and parameterized shallow water equations, we observed WLS with S-reduction trial subspaces to yield improvements over the Galerkin and LSPG approaches.
 The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals.
@@
 \methodAcronym\ \approachKwd\ overcomes the time-discretization sensitivity that LSPG is subject to.
-The second numerical experiment, which examined collocated ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow.
-In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up.
+The second numerical experiment, which examined collocated ROMs of a supersonic compressible cavity flow featuring a two-dimensional shock boundary layer interaction, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow.
+In this experiment, \methodAcronymROMs\ yielded accurate predictions at a $10$x reduction in cost as compared to the FOM.
+The WLS ROMs additionally displayed convergence with respect to an increasing basis dimension.
+LSPG ROMs, on the other hand, failed to produce an accurate prediction for the quantity-of-interest for every basis dimension considered.
+Additionally, LSPG did not display convergence with increasing basis dimension.
 Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower error in the $\elltwo$-norm.
+Lastly, the final numerical experiment demonstrated the utility of \methodAcronym\ on a parameterized system of equations.
+When predicting at a novel parameter instance, we observed WLS to again yield accurate solutions at a $10$x reduction in computational cost.
+For the same training set, the LSPG ROMs, as well as a surrogate model based on multivariate linear regression, failed to yield accurate solutions.
 
@@
 \begin{appendices}
+\section{Gauss--Newton method with frozen Jacobians}\label{appendix:gnupdate}
+One of the primary costs incurred by both the windowed least-squares approach (solved via the direct method), and the least-squares Petrov--Galerkin approach, is the computation of the Jacobian in the Gauss--Newton algorithm.
+For instance, in the second numerical example, for WLS ROMs employing a sample mesh of size $10170$ with $K=156$ basis vectors and a window size of $\Delta T = 0.025$ ($=5\Delta T$), computing the Jacobian over a time window via automatic differentiation incurs a computational cost that is $250$x greater than computing the residual over the window.
+In the third numerical example, for a sample mesh of size $1752$, a basis size of dimension $K = 83$, and a window size of $\Delta T = 2.0$ ($=100\Delta T$), computing the Jacobian over a time window via automatic differentiation incurs a computational cost that is $50$x greater than computing the residual.
+
+We expect that the computational cost associated with computing the Jacobian can be reduced by employing, for example, frozen Jacobians, analytic Jacobians, matrix-free Jacobian transpose methods~\cite{doi:10.2514/6.2016-0833}.
+While we leave a full investigation into these methods as a topic for future work, here we examine the use of a Gauss--Newton method employing frozen Jacobians.
+The main thesis behind a Gauss--Newton method employing frozen Jacobians is that approximate gradients may suffice for convergence of the nonlinear least-squares problem.
+The algorithm operates similarly to a standard Gauss--Newton method, but differs in that the Jacobian matrix is only updated ever $n_{\text{update}}$ iterations.
+Algorithm~\ref{alg:colloc_gn_frozen} outlines the algorithm.
+
+\begin{algorithm}
+\caption{\spatialAcronym\ trial subspace: algorithm for the direct solution technique with the Gauss--Newton method with frozen Jacobians and a linear multistep method over the $n$th window}
+\label{alg:colloc_gn_frozen}
+\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
+\Input{tolerance, $\epsilon$; initial guess, $\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0}$; global iteration number, $m$; Jacobian update frequency, $n_{\text{update}}$; initial guess to residual-Jacobian $\boldsymbol J$}%; initial condition $\genstateDiscreteArg{n,0}$}
+\Output{Solution to least squares problem, $\genstatecollocMatSlabArg{n}$} 
+\textbf{Online Steps}: \\
+$\text{converged} \leftarrow \text{false}$ \Comment{Set convergence checker} \\
+$\genstatecollocMatSlabArg{n}_0 \leftarrow \unroll(\genstateGuessDiscreteArg{n,1}{0},\ldots,\genstateGuessDiscreteArg{n,\nstepsArg{n}}{0})$ \Comment{Assemble generalized coordinates over window} \\
+$k \leftarrow 0$ \Comment{Set counter}\\
+\While{\text{not converged}}
+{
+%\For{$i=1,\hdots,\nstepsArg{n}$}{
+%  \For{$j=1,\hdots,\ncollocArg{n}{i}$}{
+%Compute: $\approxstateArgnt{n,i}  =  \basisspace  \genstateDiscreteArg{n,i} + \stateIntercept$ 
+%\Comment{Compute state} \\
+%Compute: $\velocity(\basisspace \genstateDiscreteArg{n,i} + \stateIntercept ) $ \Comment{Compute velocity at each time-step instance}\\
+%Compute: $\residLMSArg{n,i}(\basisspace \genstateDiscreteArg{n,i} + \stateIntercept, \ldots , \basisspace \genstateDiscreteArg{n,i - k^n(i)} + \stateIntercept) $  \Comment{Compute residual} \\
+%}
+$\mathbf{r} \leftarrow \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{Compute weighted residual over window} \\
+\uIf{ $m \% n_{\text{update}} == 0$}{
+$\mathbf{J} \leftarrow  
+\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}(\genstatecollocMatSlabArg{n}_k)$ 
+\Comment{Compute weighted residual-Jacobian over window} \\
+}
+%Compute: $[\jacobianSlabArg{n}]^T \jacobianSlabArg{n}$ \Comment{Compute system matrix for the normal equations} \\
+%Compute: $[\jacobianSlabArg{n}]^T \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}_k)$ \Comment{compute RHS for normal equations} \\
+\uIf{ $\norm{ \mathbf{J}^T\mathbf{r}  } \le \epsilon$ }{
+{\text{converged} $\leftarrow$ \text{true}}  \Comment{Check and set convergence based on gradient norm} \\
+Return: $\genstatecollocMatSlabArg{n} = \genstatecollocMatSlabArg{n}_{k+1} $ \Comment{Return converged solution}\\
+}
+\Else{
+ Compute  $\Delta  \genstatecollocMatSlabArg{n} $ that minimizes $\norm{ \mathbf{J} \Delta \genstatecollocMatSlabArg{n} +\mathbf{r}}^2$ \Comment{Solve the linear least-squares problem} \\
+$\alpha \leftarrow \text{linesearch}(\Delta  \genstatecollocMatSlabArg{n} ,\genstatecollocMatSlabArg{n}_k ) $ \Comment{Compute $\alpha$ based on a line search, or set to 1}\\
+$\genstatecollocMatSlabArg{n}_{k+1} \leftarrow \genstatecollocMatSlabArg{n}_k + \alpha \Delta \genstatecollocMatSlabArg{n}$ \Comment{Update guess to the state} \\
+}
+$k\leftarrow k+1$ \\
+$m\leftarrow m+1$
+}
+\end{algorithm}
+
+The final numerical experiment considered in this work presents results for ROMs employing the Gauss--Newton method with frozen Jacobians.
+Here, we investigate the sensitivity of the ROMs to the parameter controlling the Jacobian update frequency, $n_{\text{update}}$.
+To this end we consider the case outlined in Section~\ref{sec:swe_results}, in where WLS ROMs of varying window sizes and a basis of dimension $K=83$ are used to solve the shallow water equations for the parameter instance $(\mu_1, \mu_2) = (7.5,0.16)$.
+Figure~\ref{fig:rom_metrics_swe_updatefreq} presents performance metrics for the various WLS ROMs, solved via the Gauss--Newton method with frozen Jacobians, as a function of the window size.
+We note that, in Algorithm~\ref{alg:colloc_gn_frozen}, we use the Jacobian over the previous window as our initial guess to the Jacobian.
+Results are shown for $n_{\text{update}} = 1,2,5,10,20$.
+We observe that, for all instances of $n_{\text{update}}$, WLS ROMs solved via Gauss--Newton with frozen Jacobians yield results that are similar to those of the WLS ROMs solved via the standard Gauss--Newton algorithm.
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_error_updateFreq_K83.pdf}
+\caption{Relative time-integrated $\elltwo$-error.}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_residual_updateFreq_K83.pdf}
+\caption{Time-integrated residual norm.}
+\end{subfigure}
+\caption{Time-integrated performance metrics of the LSPG and WLS ROMs as a function of window size.
+We examine ROMs solved via the Gauss--Newton method with frozen Jacobians for different update frequencies, $n_{\text{update}}$}
+\label{fig:rom_metrics_swe_updatefreq}
+\end{center}
+\end{figure}
+
+Next, Figure~\ref{fig:rom_walltimes_swe_updatefreq} reports the relative wall times of the various WLS ROMs with respect to the FOM.
+We make several interesting observations.
+We first observe that, for the values of $n_{\text{update}}$ considered, higher values of $n_{\text{update}}$ yield lower wall times for WLS with smaller window sizes.
+For WLS with larger window sizes, larger values of $n_{\text{update}}$ yield lower wall times until $n_{\text{update}}$ reaches a critical value.
+After this value has been exceeded, we do not observe a further reduction in cost.
+For example, WLS with $\Delta T = 10.0$ incurs the same computational cost for $n_{\text{update}} = 5$, as it does for $n_{\text{update}} = 10$ and $n_{\text{update}} = 20$.
+Similarly, WLS with $\Delta T = 5.0$ incurs the same computational cost for $n_{\text{update}}= 10$ as it does for $n_{\text{update}} = 20$.
+Across all cases, employing a frozen Jacobian algorithm leads to between a $2$x and $4$x decrease in computational cost.
+This speedup is obtained without compromising the solution accuracy, as shown in Figure~\ref{fig:rom_metrics_swe_updatefreq}.
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.49\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_walltime_updateFreq_K83.pdf}
+\caption{Relative time-integrated $\elltwo$-error.}
+\end{subfigure}
+%\begin{subfigure}[t]{0.49\textwidth}
+%\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_walltimeLspg2_updateFreq_K83.pdf}
+%\caption{Time-integrated residual norm.}
+%\end{subfigure}
+\caption{Relative wall times of WLS ROMs solved via the Gauss--Newton method with frozen Jacobians. Results are presented for various Jacobian update frequencies, $n_{\text{update}}$} 
+\label{fig:rom_walltimes_swe_updatefreq}
+\end{center}
+\end{figure}
+
 \section{Proper orthogonal decomposition}
 Algorithm~\ref{alg:pod} presents the algorithm for computing the trial basis via proper orthogonal decomposition.
+In the final numerical experiment, trial bases are constructed for a set of varying input parameters. Algorithm~\ref{alg:pod_tlp} describes the construction of the POD bases for this case.
 \begin{algorithm}
@@
 \label{alg:pod}
-\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$; weighting matrix, $\mathbf{P}$ \; 
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; basis dimension $K$; weighting matrix, $\mathbf{T}$ \; 
 %\newline
-\textbf{Output:} POD Basis $\basisspace \in \{ \mathbf{X} \in \RR{\fomdim \times \romdim} | \mathbf{X}^T\mathbf{P} \mathbf{X} = \mathbf{I} \} $ \;
+\textbf{Output:} POD Basis $\basisspace \in \{ \mathbf{X} \in \RR{\fomdim \times \romdim} | \mathbf{X}^T\mathbf{T} \mathbf{X} = \mathbf{I} \} $ \;
 %\newline
@@
     \item Solve FOM O$\Delta$E and collect solutions into snapshot matrix
-$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} - \stateInterceptArg{n} & \stateFOMDiscreteArg{N_{\text{skip}}} - \stateInterceptArg{n} & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} - \stateInterceptArg{n} \end{bmatrix}$$
-    \item Compute the thin singular value decomposition, $$ \sqrt{\mathbf{P}}^T\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
-    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$ and $\sqrt{\mathbf{P}}$ is the Cholesky decomposition of $\mathbf{P}$, i.e., $\sqrt{\mathbf{P}} \sqrt{\mathbf{P}}^T = \mathbf{P}$
-    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} [\sqrt{\mathbf{P}}^T]^{-1}  \mathbf{u}_1 & \cdots & [\sqrt{\mathbf{P}}^T]^{-1} \mathbf{u}_K \end{bmatrix}$
+$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} - \stateInterceptArg{} & \stateFOMDiscreteArg{N_{\text{skip}}} - \stateInterceptArg{} & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} - \stateInterceptArg{} \end{bmatrix}$$
+    \item Compute the thin singular value decomposition, $$ \sqrt{\mathbf{T}}^T\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
+    where $\sqrt{\mathbf{T}}$ is the Cholesky decomposition of $\mathbf{T}$, i.e., $\sqrt{\mathbf{T}} \sqrt{\mathbf{T}}^T = \mathbf{T}$
+    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} [\sqrt{\mathbf{T}}^T]^{-1}  \mathbf{u}_1 & \cdots & [\sqrt{\mathbf{T}}^T]^{-1} \mathbf{u}_K \end{bmatrix}$
 \end{enumerate}
-
 \end{algorithm}
-
-%In the final numerical experiment, time-local trial bases are constructed for a set of varying input parameters. Algorithm~\ref{alg:pod_tlp} describes the construction of the POD bases for this case.
-%\begin{algorithm}
-%\caption{Algorithm for generating time-local POD basis over multiple parameter instances}
-%\label{alg:pod_tlp}
-%\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; starting time-instance, $N_s$; number of instances $N_{i}$; intercept $\stateInterceptArg{}$; parameter instances $\param_{1} , \ldots, \param_{n_{\text{train}}}$; basis dimension $K$ \\
-%%\newline
-%\textbf{Output:} POD Basis $\basisspace \in \mathbb{V}_{K}(\RR{N})$ \;
-%%\newline
-%\textbf{Steps:}
-%\begin{enumerate}
-%    \item Solve FOM O$\Delta$E for $\param_1, \ldots, \param_{n_{\text{train}}}$ and collect solutions into snapshot matrix
-%\small{$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteParamArg{N_s}{\param_1} - \stateInterceptArg{n} & \cdots & \stateFOMDiscreteParamArg{N_s + N_i N_{\text{skip}}}{\param_1}- \stateInterceptArg{n} & \cdots & \stateFOMDiscreteParamArg{N_s + N_i N_{\text{skip}}}{\param_{n_{\text{train}}}}- \stateInterceptArg{n} \end{bmatrix}$$}
-%    \item Compute the thin singular value decomposition, $$\mathbf{S} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
-%    where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
-%    \item Truncate left singular vectors and to form the basis, $\basisspace \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_K \end{bmatrix}$
-%\end{enumerate}
-%\end{algorithm}
+\begin{algorithm}
+\caption{Algorithm for generating POD basis over multiple parameter instances}
+\label{alg:pod_tlp}
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$; intercept $\stateInterceptArg{}$; parameter training set $\mathcal{D}_{\text{train}}$; basis dimension $K$; weighting matrix, $\mathbf{T}$ \;
+%\newline
+\textbf{Output:} POD Basis $\basisspace \in \{ \mathbf{X} \in \RR{\fomdim \times \romdim} | \mathbf{X}^T\mathbf{T} \mathbf{X} = \mathbf{I} \} $ \;
+%\newline
+\textbf{Steps:}
+\begin{enumerate}
+    \item Execute Algorithm~\ref{alg:pod}, with the following modification to step 1: Solve the FOM O$\Delta$E for $\params \in \mathcal{D}_{\text{train}}$ and collect solutions into snapshot matrix
+\small{$$\mathbf{S}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteParamArg{0}{\param_1} - \stateInterceptArg{} & \cdots & \stateFOMDiscreteParamArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}}{\param_1}- \stateInterceptArg{} & \cdots & \stateFOMDiscreteParamArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}}{\param_{n_{\text{train}}}}- \stateInterceptArg{} \end{bmatrix}$$}
+\end{enumerate}
+\end{algorithm}
 
@@
 To construct the sampling point matrix used for hyper-reduction in the second numerical experiment, we employ q-sampling~\cite{qdeim_drmac} and the sample 
-mesh concept~\cite{carlberg_gnat}. Algorithm~\ref{alg:qdeim} outlines the steps used in the second numerical experiment to compute the sampling points.
+mesh concept~\cite{carlberg_gnat}.
+Algorithm~\ref{alg:qdeim} outlines the steps used in the second numerical experiment to compute the sampling points.
+In the final numerical experiment, trial bases are constructed for a set of varying input parameters.
+Algorithm~\ref{alg:qdeim_tlp} outlines the steps used for this case.
 
@@
 \label{alg:qdeim}
-\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$ \; 
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$, weighting  matrix, $\mathbf{T}$\; 
 %\newline
@@
 \begin{enumerate}
-    \item Solve FOM O$\Delta$E and collect velocity snapshots 
+    \item Solve FOM O$\Delta$E and collect solution snapshots 
 %$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \velocity (\stateFOMDiscreteArg{0})  & \velocity( \stateFOMDiscreteArg{N_{\text{skip}}}) & \cdots & \velocity( \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} ) \end{bmatrix}$$
 $$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteArg{0} & \stateFOMDiscreteArg{N_{\text{skip}}}  & \cdots & \stateFOMDiscreteArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}} \end{bmatrix}$$
-    \item Compute the thin singular value decomposition, $$\mathbf{F} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
+    \item Compute the thin singular value decomposition, $$ \sqrt{\mathbf{T}}^T \mathbf{F} (N_{\text{skip}}) = \mathbf{U \Sigma Z}^T,$$
     where $\mathbf{U} \equiv \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}})}\end{bmatrix}$
-    \item Compute the QR factorization of  $\mathbf{U}^T$ with column pivoting,
+    \item Set  $\basisspace \equiv \begin{bmatrix} [\sqrt{\mathbf{T}}^T]^{-1}  \mathbf{u}_1 & \cdots & [\sqrt{\mathbf{T}}^T]^{-1} \mathbf{u}_{\text{floor}(N_t/N_{\text{skip}}}\end{bmatrix}$
+    \item Compute the QR factorization of  $\mathbf{\basisspace}^T$ with column pivoting,
     \begin{equation*}
-        \mathbf{U}^T \mathbf{P}^* = \mathbf{QR}
+        \mathbf{\basisspace}^T \mathbf{P}^* = \mathbf{QR}
     \end{equation*}
@@
     \item Augment the sampling point matrix, $\stweightingMatOneArg{}$, with additional columns such that all unknowns are computed at the mesh cells selected by Step 3. 
-For the second numerical experiment, these additional unknowns correspond to each conserved variable and degree of freedom in the selected cells.
+\end{enumerate}
+
+\end{algorithm}
+\begin{algorithm}
+\caption{Algorithm for generating the sampling matrix through q-sampling.}
+\label{alg:qdeim_tlp}
+\textbf{Input:} Number of time-steps between snapshots $N_{\text{skip}}$, number of primal sampling points, $n_s$, parameter training set $\mathcal{D}_{\text{train}}$, weighting  matrix, $\mathbf{T}$\;
+%\newline
+\textbf{Output:} Weighting matrix $\stweightingMatArg{} \equiv \stweightingMatOneTArg{} \stweightingMatOneArg{} \in \{0,1\}^{N \times N}$ \;
+%\newline
+\textbf{Steps:}
+\begin{enumerate}
+    \item Execute Algorithm~\ref{alg:qdeim}, with the following modification to step 1: Solve the FOM O$\Delta$E for $\params \in \mathcal{D}_{\text{train}}$ and collect solutions into snapshot matrix
+\small{$$\mathbf{F}(N_{\text{skip}}) \defeq \begin{bmatrix} \stateFOMDiscreteParamArg{0}{\param_1}  & \cdots & \stateFOMDiscreteParamArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}}{\param_1} & \cdots & \stateFOMDiscreteParamArg{\text{floor}(N_t/N_{\text{skip}})N_{\text{skip}}}{\param_{n_{\text{train}}}} \end{bmatrix}$$}
+
 \end{enumerate}

### 2020-10-12 numerical_experiments.tex
latest

@@
 \section{Numerical experiments}\label{sec:numerical_experiments}
-We now analyze the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity.
-In each experiment, we compare \methodAcronymROMs\ to the Galerkin and LSPG ROMs.
-The purpose of the numerical experiments is to assess the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy.
-We additionally assess the impact of the time step and time scheme on \methodAcronym.
-In both experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$.
-We also note that both experiments are designed to test the \textit{reproductive} ability of the ROMs.
-We do not consider future state prediction and prediction at new parameter instances as these problems introduce factors that confound the solution accuracy with the solution methodology (e.g., accuracy of the basis).
+We now analyze the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on three benchmark problems: (1) a reproductive ROM for the Sod shock tube problem (2) a reproductive ROM for a supersonic compressible flow in a cavity, and (3) a predictive ROM for the parameterized shallow water equations.
+In each experiment, we compare \methodAcronymROMs\ to either the LSPG, or the Galerkin and LSPG, ROMs.
+In the experiments, we assess (1) the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy, (2) the impact of the time step and time scheme on \methodAcronym, (3) the impact of the window size on the convergence properties of the ROMs, and (4) the impact of the window size on the cost of the ROMs.
+In all experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$.
+We note that the first two experiments are designed to test the \textit{reproductive} ability of the ROMs, while the third experiment tests the predictive capabilities of the ROM.
 
@@
 We evolve the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$.
-We construct the \spatialAcronym\ trial subspace by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $, and $\mathbf{P} = \mathbf{I}$.
+We construct the \spatialAcronym\ trial subspace by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $, and $\mathbf{T} = \mathbf{I}$.
 The resulting trial subspace corresponds to an energy criterion of $99.99\%$.
@@
 
-\subsubsection{Description of reduced-order models}
+\subsubsection{Description of reduced-order models}\label{sec:sod_rom_description}
 We consider reduced-order models based on Galerkin projection, LSPG
@@
 We additionally note that, although the time-integrated $\elltwo$-error of
-the projected FOM solution is significantly lower than that of the various ROM solutions, the time-integrated residual norm of the projected FOM solution is \textit{higher} 
+the projected FOM solution is significantly lower than that of the various ROM solutions, the time-integrated residual norm of the projected FOM solution (i.e., the time-integrated norm of the residual evaluated at the $\elltwo$-orthogonal projection of the FOM solution onto the trial subspace) is \textit{higher}
 than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$-norm, it does not satisfy the governing equations as well the ROM solutions.
@@
 \includegraphics[width=1.\linewidth]{figs/sod/error_converge.pdf}
-\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
+\caption{Relative time-integrated $\elltwo$-error.}
 \label{fig:sod_error_converge}
@@
 \includegraphics[width=1.\linewidth]{figs/sod/residual_converge.pdf}
-\caption{Time-integrated residual norm of the \methodAcronymROM.} 
+\caption{Time-integrated residual norm.} 
 \label{fig:sod_residual_converge}
@@
 \input{cavity}
+\input{shallow_water}

### 2020-10-12 section4/spatial_numerical.tex
latest

@@
 	an extra dimension for parallelization that can be exploited to future
-	reduce the wall time. The investigation of these additional, potentially
-	more efficient, solution algorithms is a topic of future work. 
+	reduce the wall time. While we leave a detailed investigation of these additional, potentially
+	more efficient, solution algorithms as a topic of future work, we briefly examine the use of frozen Jacobians in Section~\ref{sec:swe} and Appendix~\ref{appendix:gnupdate}. 
 \end{remark}

### 2020-10-12 shallow_water.tex
latest

\subsection{Parameterized shallow water equations}\label{sec:swe}
The final example considers reduced-order models of the shallow water equations with varying parameters. The governing system of PDEs for the shallow water equations is 
\begin{equation}\label{eq:swes}
\begin{split}
&\frac{\partial h}{\partial t} + \frac{\partial}{\partial \spatialvar }(  h u) + \frac{\partial}{\partial \spatialvarTwo }( h v) = 0\\
&\frac{\partial h u}{\partial t} + \frac{\partial}{\partial \spatialvar } (h u^2 + \frac{1}{2} \parameter_1 h^2) + \frac{\partial}{\partial \spatialvarTwo }( h u v) = 0\\
&\frac{\partial h v}{\partial t} + \frac{\partial}{\partial \spatialvar } (h u v) + \frac{\partial}{\partial \spatialvarTwo }( h v^2 +  \frac{1}{2} \parameter_1 h^2) = 0
\end{split}
\end{equation}
on the domain $(\spatialvar ,\spatialvarTwo ) \in \physDomain = [-5,5] \times [-5,5]$, $t \in [0,T]$
with $T=10$, and 
where $h,u,v: \physDomain \times [0,T] \rightarrow \RR{}$, with $h$ denoting the height
of the water surface, $u$ denoting the $\spatialvar$-velocity, and $v$
denoting the $\spatialvarTwo$-velocity.
The parameter $\parameter_1 \in \RR{+}$ is the ``gravity'' parameter.  In this example, we consider a set of parameterized initial conditions given by 
$$h(\spatialvar
,\spatialvarTwo ,0) = 1 + \parameter_2 e^{ -(\spatialvar  - 1)^2 -
(\spatialvarTwo  - 1)^2 }, \quad u(\spatialvar ,\spatialvarTwo ,0) =
v(\spatialvar ,\spatialvarTwo ,0) = 0,\quad (\spatialvar ,\spatialvarTwo ) \in
\physDomain,$$ which corresponds to a Gaussian pulse centered at $(1,1)$ with
a magnitude of $\parameter_2$. We consider a parameter domain
$(\parameter_1,\parameter_2)\in\parameterdomain = [3,9]\times [0.05,2]$.

\subsubsection{Description of FOM and generation of \spatialAcronym\ trial subspaces}
The full-order model again comprises a discontinuous Galerkin discretization of the system~\eqref{eq:swes}. We obtain the discretization 
by partitioning the domain into 128 elements of equal size in both the $\spatialvar$ and $\spatialvarTwo$ directions. Over each element, 
the solution is represented to second order via tensor product polynomials of order $p=1$. The resulting FOM dimension is $\fomdim = 196608$. The Rusanov flux is employed at the cell interfaces. Similarly to the previous section, spatial discretization via the discontinuous Galerkin method yields a dynamical system 
of the form
$$  \mass_{\text{DG}}\frac{d \state}{dt} = \velocity_{\text{DG}} (\state).$$
Time integration 
is again performed via the CN method with a time step of $\Delta t = 0.01$. The nonlinear system arising at each time step is solved via a Jacobian-free 
GMRES algorithm. We employ slip-wall boundary conditions at the edge of the domain.

The reduced-order models again leverage POD to construct the \spatialAcronym\ trial basis with q-sampling based on snapshots of the FOM solution to select the sampling points. Once again, a uniform basis and sampling matrix is employed across all windows. The process used to construct the trial subspaces and weighting matrices for the ROMs is as follows:
\begin{enumerate}
\item Set the parameter-grid to be $\mathcal{D}_{\text{train}}  \{ 3,6,7 \} \times \{0.05,0.125,0.2\}$. 
\item Execute Algorithm~\ref{alg:pod_tlp} with inputs $N_{\text{skip}} = 2$, $\stateInterceptArg{} = \bz$, $\mathcal{D}_{\text{train}}$, $\romdim = \{ 83,143 ,223 \}$, and $\mathbf{T} = \mathbf{M}_{\text{DG}}$ to construct POD bases capturing $99.9999\%, \; 99.99999\%,$ and $99.999999\%$ of the energy, respectively.
\item Execute Algorithm~\ref{alg:qdeim_tlp} with inputs $\nskip = 2$, $\nsamples = 150$, $\mathcal{D}_{\text{train}}$ and $\mathbf{T} = \mathbf{M}_{\text{DG}}$ to construct a sample mesh of dimension $\mathbf{S} \in \{0,1\}^{1752\times \fomdim }$.
 \item Set $\stweightingMatOneArg{} = \mathbf{S} \mass_{\text{DG}}$. 
\end{enumerate}  
To depict the nature of the problem, Figure~\ref{fig:fom_sols_swe} presents surface plots for the surface height, $h$, at various time instances for the training parameter instance $\mu_1 = 9.0,\mu_2 = 0.125$.
\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_t0.png}
\caption{$t=0.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_t1.png}
\caption{$t=1.0$}
\end{subfigure}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_t3.png}
\caption{$t=3.0$}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_t10.png}
\caption{$t=10.0$}
\end{subfigure}
\caption{Surface plots of the surface height, $h$, for the training parameter instance $\mu_1 = 9.0,\mu_2 = 0.125$.} 
\label{fig:fom_sols_swe}
\end{center}
\end{figure}

\subsubsection{Description of reduced-order models}\label{sec:swe_describe_roms}
We consider collocated ROMs based on the least-squares Petrov--Galerkin and \methodAcronym\ approaches. Details on the implementation of the methods is as follows:
\begin{itemize}

\item \textit{LSPG ROM with collocation}: We consider collocated LSPG ROMs, which are built on top of the FOM discretization using the CN scheme for temporal 
discretization. We employ a time step of $\Delta t  = 0.02$, which corresponds to the frequency at which snapshots were collected. 
The implementation is the same as previously described in Section~\ref{sec:sod_rom_description}, with the following exception: in the Gauss--Newton iteration used to solve the nonlinear least-squares problem arising at each time instance, a frozen Jacobian algorithm is employed, in where the Jacobian is updated every $n_{\text{update}}=10$ iterations. Appendix~\ref{appendix:gnupdate} details the impact of the update frequency of the Jacobian on the numerical results. The algorithm for the Gauss--Newton method with frozen Jacobians is presented in Algorithm~\ref{alg:colloc_gn_frozen} in Appendix~\ref{appendix:gnupdate}. We note here that the Gauss--Newton method with frozen Jacobians yields similar solutions to the standard Gauss--Newton approach, but at a much reduced computational cost.

\item \textit{\methodAcronymROMs\ with collocation}: We consider \methodAcronymROMs\ solved via the direct method. The ROMs use the CN time discretization with a time step of 
$\Delta t = 0.02$. The implementation is the same as previously described in Section~\ref{sec:sod_rom_description}, with the exception that the ROMs employ the Gauss--Newton method with frozen Jacobians detailed above. 
%\begin{itemize}
%\item Direct Method: \methodAcronym ROMs solved via the direct method are again implemented with the same Crank--Nicolson time discretization with a nominal 
%time-step size of $\Delta t = 0.1$. 
%\end{itemize}
\end{itemize}
As an additional benchmark, we consider a time-local multivariate linear regression surrogate model. For each time instance, this surrogate operates by constructing a linear mapping from the input parameters, $(\mu_1 ,\mu_2)$, to the field solution at each grid point. Thus, the surrogate model is given as
\begin{equation}\label{eq:lin_surrogate}
\approxstateLR^n = {\approxstateLRIntercept}^n + \mu_0 \mathbf{c_1}^n + \mu_1 \mathbf{c_2}^n ,
\end{equation}
where $\approxstateLRIntercept^n \in \RR{\fomdim}$, $\mathbf{c_1}^n \in \RR{\fomdim}$, $\mathbf{c_2}^n \in \RR{\fomdim}$, $n=0,\ldots,N_t$, with $\fomdim = 196608$, are the model parameters. 
The surrogate model parameters are fits via \textsf{scikit-learn}~\cite{scikit-learn}. The same training data used for the ROM is used for the surrogate model.

\subsubsection{Numerical results at novel parameter instance}\label{sec:swe_results}
We investigate the performance of the reduced-order models at the novel parameter instance $(\mu_1, \mu_2) = (7.5,0.16)$. We emphasize that this parameter instance is not in the training set. We consider a set of WLS ROMs that minimize the residuals over $\DeltaSlabArg{n} \equiv \DeltaSlabArg{} = 0.04$, $0.1$, $0.2$, $0.5$, $1.0$, $2.0$, $5.0$, $10.0$, along with LSPG and the linear surrogate model~\eqref{eq:lin_surrogate}. First, Figure~\ref{fig:rom_sols_swe1} presents results for WLS ROMs at $\Delta T = 0.01,0.5,2.0,$ and $10.0$, as well as for the LSPG ROM and the linear regression surrogate model. Figure~\ref{fig:rom_sols_swe1h} depicts the evolution of the surface height at $(x,y) = (-2.4834,-2.4834)$ as a function of time, while Figure~\ref{fig:rom_sols_swe1e} depicts the relative $\elltwo$-error as a function of time. We first observe that the linear regression surrogate model performs significantly worse than all ROMs; this result demonstrates the complex dynamics of the problem and highlights the utility of ROMs. Next, we observe that LSPG yields poor solutions for later time instances. At $t=10$, LSPG only yields a marginally better prediction than the linear surrogate model. We next observe the WLS ROMs to yield improved solutions as compared to LSPG. We observe that the evolution of the surface height as a function of time is better characterized; specifically we once again observe that growing the window size yields smoother solutions. Next, Figure~\ref{fig:rom_hsols_swe} presents surface plots of the surface height, $h$, for the FOM and various ROMs at $t = 10.0$. We observe the LSPG solution to be contaminated by spurious oscillations and, as a result, it fails to accurately characterize the FOM. On the other hand, we observe the WLS ROMs to yield smooth solutions that accurately characterize the FOM.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_h_vs_t_K83.pdf}
\caption{Surface height at $(x,y) = (-2.4834,-2.4834)$ as a function of time.}
\label{fig:rom_sols_swe1h}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_error_vs_t_K83.pdf}
\caption{Relative error as a function of time.}
\label{fig:rom_sols_swe1e}
\end{subfigure}
\caption{Surface height (left) and relative error (right) as a function of time.} 
\label{fig:rom_sols_swe1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_fom_t50.png}
\caption{FOM}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_lspg_t50.png}
\caption{LSPG}
\end{subfigure}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_wls_DT05_t50.png}
\caption{WLS ($\Delta T = 0.5$)}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={4cm 0cm 2cm 2cm},clip,width=1.0\linewidth]{figs/swe/h_wls_DT20_t50.png}
\caption{WLS ($\Delta T=2.0$)}
\end{subfigure}
\caption{Surface plots of the surface height, $h$, at $t=10.0$ for the novel parameter instance $\mu_1 = 7.5,\mu_2 = 0.16$ as predicted by the FOM and various ROMs.} 
\label{fig:rom_hsols_swe}
\end{center}
\end{figure}

Next, Figure~\ref{fig:rom_swe_metrics} depicts the impact of the window size on the relative error and residual norms. The time-integrated $\elltwo$-error of the linear surrogate model is additionally shown for reference. We make the same observations as before: increasing the window size leads to a monotonic decrease in the time-integrated residual norm, but not in the time-integrated error norm. We observe that the time integrated $\elltwo$-error is lowest for an intermeidate window size. At $\Delta T = 1$, the time-integrated $\elltwo$-error of the WLS ROM is approximately half that of LSPG. We additionally note that all ROMs are significantly more accurate than the linear surrogate model. Similarly, WLS with $\Delta T = 10.0$ yields a time-integrated residual norm that is almost three orders of magnitude lower than LSPG.

\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_error_K83.pdf}
\caption{Relative time-integrated $\elltwo$-error.}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_residual_K83.pdf}
\caption{Time-integrated residual norm.}
\end{subfigure}
%\begin{subfigure}[t]{0.85\textwidth}
\caption{Time-integrated performance metrics of the LSPG and WLS ROMs as a function of window size} 
\label{fig:rom_swe_metrics}
\end{center}
\end{figure}

Next, Figure~\ref{fig:rom_swe_timings} presents timings for the various ROMs as a function of the window size. We note that, as described in Section~\ref{sec:swe_describe_roms}, all ROMs employ a Gauss--Newton solve with a Jacobian update frequency of $n_{\text{update}} = 10$; the sensitity of the results to this parameter are detailed in Appendix~\ref{appendix:gnupdate}. We again obsere that the cost of WLS grows with window size: WLS with $\Delta T = 10.0$ ($=500\Delta t$) incurs a 5x increase in computational cost with repsect to LSPG. With respect to the FOM, we observe LSPG to result in a $20$x speedup, while WLS with the largest window size $\Delta T = 10.0$ yields a $4$x speedup.  
\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_walltimeLspg_updateFreq_K83.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_windowSize_vs_walltimeFom_updateFreq_K83.pdf}
\end{subfigure}
%\begin{subfigure}[t]{0.85\textwidth}
\caption{Wall clock times as a function of window size relative to LSPG (left) and the FOM (right).} 
\label{fig:rom_swe_timings}
\end{center}
\end{figure}

We next assess the impact of the basis dimension on the results. Figure~\ref{fig:rom_swe_converge} summarizes the performance of WLS and LSPG ROMs as a function of basis dimension. We first observe that LSPG, and WLS with smaller window sizes, do not display monotonic convergence with increasing basis dimension. 
We expect that this lack of convergence is because of the fact that, as the basis dimension increases, the ROM trial space can represent sharper gradients. Without proper stabilization, this can yield spurious oscillations, as previously observed in Figure~\ref{fig:rom_hsols_swe}. As the window size increases, the ROM is more stable and, thus, increasing the basis dimension yields improved solutions. 
For window sizes of $\Delta T \ge 0.5$, we observe that increasing the basis dimension leads to a monotonic decrease in both the time-integrated $\elltwo$-error and residual norm.

\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_converge_error.pdf}
\caption{Relative time-inegrated $\elltwo$-error.}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_converge_resid.pdf}
\caption{Time integrated residual norm.}
\end{subfigure}
%\begin{subfigure}[t]{0.85\textwidth}
\caption{Performance metrics of the LSPG and WLS ROMs as a function of basis dimension.} 
\label{fig:rom_swe_converge}
\end{center}
\end{figure}

Lastly, Figure~\ref{fig:rom_swe_pareto} presents Pareto plots for the time-integrated $\elltwo$-error and residual as a function of wall-clock times for ROMs with basis dimensions $K = 83,143,223$ and window sizes $\Delta T = 0.04, 0.1,0.2,0.5,1.0,2.0,$ and $5.0$, along with LSPG. We again observe ROMs employing an intermediate window size to be Pareto-optimal for the time-integrated $\elltwo$-error. It is interesting to note that the error vs. wall time curves for WLS ($\Delta T = 0.01$, $0.02$, $0.05$, $1.0$ ,$2.0$) fall almost on top of each other. Lastly, we observe that larger window sizes with lower basis dimensions are Pareto optimal for the residual. For instance, we observe that WLS with $\Delta T = 5.0$, $K=83$ yields a lower residual for a given wall time than WLS with $\Delta T = 2.0$, $K=143$. 
\begin{figure}
\begin{center}
%\begin{subfigure}[t]{0.85\textwidth}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_converge_error_pareto.pdf}
\caption{Relative time-inegrated $\elltwo$-error.}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_converge_resid_pareto.pdf}
\caption{Time integrated residual norm.}
\end{subfigure}
%\begin{subfigure}[t]{0.85\textwidth}
\caption{Performance metrics of the LSPG and WLS ROMs as a function of basis dimension.} 
\label{fig:rom_swe_pareto}
\end{center}
\end{figure}

### 2020-10-13 cavity.tex
made minor edits

@@
 
+\begin{figure}
+\begin{center}
+\begin{subfigure}[t]{0.85\textwidth}
+\includegraphics[trim={2cm 5.5cm 1cm 7cm},clip,width=0.98\linewidth]{figs/cavity_new/cav_geom.pdf}
+\end{subfigure}
+\caption{Figure depicting geometry and flow conditions of the cavity flow problem.} 
+\label{fig:cav_fig}
+\end{center}
+\end{figure}
+
 \subsubsection{Description of FOM and generation of \spatialAcronym\ trial subspace}
@@
 by partitioning the domain into $200$ elements in the flow direction and $80$ elements 
-in the wall-normal direction. The discretization represents the solution to third order over each element using tensor product polynomials of order $p=2$, 
+in the wall-normal direction. The solution is then represented to third order over each element using tensor product polynomials of order $p=2$, 
 resulting in $144000$ unknowns for each conserved variable. Spatial discretization via the discontinuous Galerkin method yields a dynamical system 
@@
 \item Initialize the FOM with uniform free-stream conditions.
-\item Evolve the FOM until a statistically steady state has been reached.
+\item Evolve the FOM for $t \in [0,500]$.
 \item Reset the time coordinate, $0 \leftarrow t$, and execute Algorithm~\ref{alg:pod} with $\stateInterceptArg{} = \bz, N_{\text{skip}} = 8$, $K = \{50,100,156,200\}$, and $\podIP =  \mass_{\text{DG}}$ over $t \in [0,50]$ to construct the four trial subspaces tabulated in Table~\ref{tab:rom_basis_details}.
-\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 8$, $n_s = 250$ to obtain the sampling point matrix of dimension $\mathbf{S} \in \{0,1\}^{10170 \times \fomdim }$.
+\item Execute Algorithm~\ref{alg:qdeim} with $N_{\text{skip}} = 8$, $n_s = 250$, and $\mathbf{T} = \mathbf{M}_{\text{DG}}$ to obtain the sampling point matrix of dimension $\mathbf{S} \in \{0,1\}^{10170 \times \fomdim }$.
 \item Set $\stweightingMatOneArg{} = \mathbf{S} \mass_{\text{DG}}$. 
@@
 \end{itemize}
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}[t]{0.85\textwidth}
-\includegraphics[trim={2cm 5.5cm 1cm 7cm},clip,width=0.98\linewidth]{figs/cavity_new/cav_geom.pdf}
-\end{subfigure}
-\caption{Figure depicting geometry and flow conditions of the cavity flow problem.} 
-\label{fig:cav_fig}
-\end{center}
-\end{figure}
 
@@
 We first observe that the collocated LSPG ROM fails to yield an accurate prediction for the pressure signal.
-We next observe the collocated WLS ROMs to yield more accurate solutions than LSPG.
+We observe the collocated WLS ROMs to yield more accurate solutions than LSPG.
 The pressure signal is well characterized by all WLS ROMs, and the $\ell^2$-errors are significantly lower.
@@
 \end{center}
-\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (top) and normalized $\elltwo$-errors (bottom) of various collocated ROMs to the full-order model solution.}
+\caption{Comparison of the pressure profiles obtained at the midpoint of the bottom wall (left) and normalized $\elltwo$-errors (right) of various collocated ROMs to the full-order model solution.}
 \label{fig:cav_results1}
@@
 
-Next, Figure~\ref{fig:cav_results2} shows the time-integrated $\elltwo$-errors and residual norms of the WLS ROMs as a function of the window size. We observe that growing the window size over 
+Figure~\ref{fig:cav_results2} shows the time-integrated $\elltwo$-errors and residual norms of the WLS ROMs as a function of the window size. We observe that growing the window size over 
 which the residual is minimized leads to a lower time-integrated residual, but not necessarily a lower $\elltwo$-error.
@@
 
-Next, Figure~\ref{fig:cav_wallclock}
-shows the wall-clock times\footnote{We computed wall-clock times over the time interval $t \in [0,1]$} of the \methodAcronymROMs\ as compared to (1) the LSPG ROM (Figure~\ref{fig:cav_wallclock_lspg}) and (2) the FOM (Figure~\ref{fig:cav_wallclock_fom}).
+Figure~\ref{fig:cav_wallclock}
+shows the wall-clock times\footnote{We computed wall-clock times over the time interval $t \in [0,1]$} of the \methodAcronymROMs\ as compared to (1) the LSPG ROM and (2) the FOM.
 As expected, increasing $\Delta T$ again leads to an increase in computational cost; minimizing the residual over a window comprising 20 time instances yields a 4x increase in cost over LSPG.
@@
 First, it must be noted that, due to a decreased stiffness in the system, we are able to run the ROMs at a time step that is 10x greater than the FOM.
-Second, we note that in our DG code, the vast majority of CPU time in the WLS ROMs is spent computing the Jacobians.
-For WLS ROMs employing a sample mesh of size $n_s=10170$ with $K=156$ basis vectors and a window size of $\Delta T = 0.025$, it takes approximately $2$(s) to compute the residual over one time window; it takes $500$(s) to compute the Jacobian via automatic differentiation.
+Second, we note that in our DG code, the majority of CPU time in the WLS ROMs is spent computing the Jacobians.
+For WLS ROMs employing a sample mesh of size $10170$ with $K=156$ basis vectors and a window size of $\Delta T = 0.025$, it takes approximately $2$ seconds to compute the residual over one time window; it takes $500$ seconds to compute the Jacobian via automatic differentiation.
 Our full-order model solver uses a Jacobian-free GMRES solver, and thus is not subject to this bottleneck.
@@
 \begin{center}
-\begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/walltime_vs_windowSize.pdf}
-\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM.}
-\label{fig:cav_wallclock_lspg}
+\begin{subfigure}[t]{0.47\textwidth}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/error_vs_windowSize.pdf}
+%\caption{Space--time $\elltwo$-error}
+%\label{fig:cav_results2a}
 \end{subfigure}
 \begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/rel_walltime_vs_windowSize.pdf}
-\caption{Wall-clock times of \methodAcronymROMs\ with respect to the full-order model.}
-\label{fig:cav_wallclock_fom}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/residual_vs_windowSize.pdf}
+%\caption{Objective function} 
+%\label{fig:cav_results2b}
 \end{subfigure}
-\caption{Relative wall-clock times of \methodAcronymROMs.}
-\label{fig:cav_wallclock}
 \end{center}
+\caption{Time-integrated $\elltwo$-error (left) and objective function (right) as a function of window size.}
+\label{fig:cav_results2}
 \end{figure}
@@
 \begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/error_vs_windowSize.pdf}
-\caption{Space--time $\elltwo$-error}
-\label{fig:cav_results2a}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/walltime_vs_windowSize.pdf}
+%\caption{Wall-clock times of \methodAcronymROMs\ with respect to the LSPG ROM.}
+%\label{fig:cav_wallclock_lspg}
 \end{subfigure}
 \begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/residual_vs_windowSize.pdf}
-\caption{Objective function} 
-\label{fig:cav_results2b}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/cavity_new/rel_walltime_vs_windowSize.pdf}
+%\caption{Wall-clock times of \methodAcronymROMs\ with respect to the full-order model.}
+%\label{fig:cav_wallclock_fom}
 \end{subfigure}
+\caption{Relative wall-clock times of \methodAcronymROMs\ with respect to LSPG (left) and the FOM (right).}
+\label{fig:cav_wallclock}
 \end{center}
-\caption{Space--time error (left) and objective function (right) as a function of window size.}
-\label{fig:cav_results2}
 \end{figure}
@@
 Figure~\ref{fig:cav_results_pressure} presents the $\elltwo$-error in the pressure at the midpoint of the bottom wall, as well as the pressure signal in the same location.
-We observe that, for WLS, increasing the basis dimension monotinically decreases the $\elltwo$-error, residual norm, and $\elltwo$-error in the pressure.
+We observe that, for WLS, increasing the basis dimension monotonically decreases the $\elltwo$-error, residual norm, and $\elltwo$-error in the pressure.
 Increasing the basis dimension does not yield monotonically decreasing error metrics for LSPG.
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/error_vs_basisSize.pdf}
-\caption{Space--time $\elltwo$-error}
-\label{fig:cav_results_basisdima}
+%\caption{Time-integrated $\elltwo$-error}
+%\label{fig:cav_results_basisdima}
 \end{subfigure}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/residual_vs_basisSize.pdf}
-\caption{Objective function} 
-\label{fig:cav_results_basisdimb}
+%\caption{Objective function} 
+%\label{fig:cav_results_basisdimb}
 \end{subfigure}
 \end{center}
-\caption{Space--time error (left) and objective function (right) as a function of window size.}
+\caption{Time-integrated $\elltwo$-error (left) and objective function (right) as a function of basis dimension.}
 \label{fig:cav_results_basisdim}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/perror_vs_basisSize.pdf}
-\caption{Objective function} 
-\label{fig:cav_results_pressurea}
+%\caption{Objective function} 
+%\label{fig:cav_results_pressurea}
 \end{subfigure}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.\linewidth]{figs/cavity_new/p_vs_basisSize.pdf}
-\caption{Space--time $\elltwo$-error}
-\label{fig:cav_results_pressureb}
+%\caption{Space--time $\elltwo$-error}
+%\label{fig:cav_results_pressureb}
 \end{subfigure}
 \end{center}
-\caption{Space--time error (left) and objective function (right) as a function of window size.}
+\caption{Time-integrated pressure error (left) and pressure signal (right) as a function of basis dimension.}
 \label{fig:cav_results_pressure}

### 2020-10-13 main.tex
made minor edits

@@
 \begin{abstract}
-This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
+This work proposes a \methodNameLower\ (\methodAcronym) approach for model reduction
 	of dynamical systems. The proposed approach sequentially minimizes the
@@
 	addition, the approach
-	addresses key deficiencies in existing model-reduction
+	addresses key deficiencies in existing model reduction
 	techniques, e.g., the dependence of LSPG and space--time LSPG projection on the
@@
 low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods~\cite{bochev_leastsquares,bochev_lsfem_book}.}
-Researchers have developed such residual-minimizing model-reduction methods
+Researchers have developed such residual-minimizing model reduction methods
 for both static systems (i.e., systems without time-dependence)
@@
 The above attributes allow the \methodAcronym\ approach to be viewed as a
-generalization of existing model-reduction methods, as 
+generalization of existing model reduction methods, as 
 Galerkin, LSPG, and space--time LSPG projection correspond to specific
@@
 	We begin by providing the formulation for the full-order model,
-	followed by a description of standard model-reduction methods
+	followed by a description of standard model reduction methods
 	classified according to the type of trial subspace they employ.
@@
 	\eqref{eq:FOM} on such parameters for notational convenience, as this work
-	focuses on devising a model-reduction approach applicable to a specific parameter instance.
+	focuses on devising a model reduction approach applicable to a specific parameter instance.
 
@@
 This work seeks to overcome the limitations of existing residual-minimizing
-model-reduction methods, and to provide a unifying framework from which
+model reduction methods, and to provide a unifying framework from which
 existing methods can be assessed. Specifically, we look to overcome the
@@
 \section{Acknowledgments}
-The authors thank Yukiko Shimizu and Patrick Blonigan for numerous conversations from which this work benefited. 
+The authors thank Yukiko Shimizu, Patrick Blonigan, and Francesco Rizzi for numerous conversations from which this work benefited. 
 E.\ Parish acknowledges an appointment to the Sandia National Laboratories'

### 2020-10-13 numerical_experiments.tex
made minor edits

@@
 In each experiment, we compare \methodAcronymROMs\ to either the LSPG, or the Galerkin and LSPG, ROMs.
-In the experiments, we assess (1) the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy, (2) the impact of the time step and time scheme on \methodAcronym, (3) the impact of the window size on the convergence properties of the ROMs, and (4) the impact of the window size on the cost of the ROMs.
+In the experiments, we assess (1) the impact of minimizing the residual over different sized time windows on the solution accuracy, (2) the impact of the time step and time scheme on \methodAcronym, (3) the impact of the window size on the convergence properties of the ROMs, and (4) the impact of the window size on the cost of the ROMs.
 In all experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$.
 We note that the first two experiments are designed to test the \textit{reproductive} ability of the ROMs, while the third experiment tests the predictive capabilities of the ROM.
+The following numerical experiments are all executed on a machine with a 2.7 GHz 12-Core Intel Xeon E5 processor.
+All codes are written efficiently in Python, and we report timings for runs executed in serial.
 
@@
 \item \textit{Indirect method}: We consider two \methodAcronymROMs\ solved via the indirect method.
-The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$.
+The first uses the same CN discretization at a time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$.
 We solve the coupled two-point boundary
@@
 \caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
-\label{fig:sod_error_methods_a}
+\label{fig:sod_time_step_a}
 \end{subfigure}
@@
 \caption{Time-integrated residual norm of the \methodAcronymROM.} 
-\label{fig:sod_error_methods_b}
+\label{fig:sod_time_step_b}
 \end{subfigure}
@@
 All methods leverage the CN time scheme with $\Delta t = 0.002$.
-We present results for windows of size $\Delta T^n \equiv \Delta T = 0.002,0.004,0.02,0.04,0.10$ for the WLS approach.
+We present results for windows of size $\Delta T^n \equiv \Delta T = 0.002,0.004,0.02,0.04,0.20$ for the WLS approach.
 All WLS ROMs are solved via the direct approach described previously.
@@
 \caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
-\label{fig:sod_error_converge}
+\label{fig:sod_error_converge_pareto}
 \end{subfigure}
@@
 \caption{Time-integrated residual norm of the \methodAcronymROM.} 
-\label{fig:sod_residual_converge}
+\label{fig:sod_residual_converge_pareto}
 \end{subfigure}
@@
 
-\begin{figure}
-\begin{center}
-\begin{subfigure}[t]{0.45\textwidth}
-\includegraphics[width=1.\linewidth]{figs/sod/converge_walltimes.pdf}
-\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
-\label{fig:sod_error_converge}
-\end{subfigure}
-\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
-\label{fig:convergence_study}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\begin{subfigure}[t]{0.45\textwidth}
+%\includegraphics[width=1.\linewidth]{figs/sod/converge_walltimes.pdf}
+%\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
+%\label{fig:sod_error_converge}
+%\end{subfigure}
+%\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
+%\label{fig:convergence_study}
+%\end{center}
+%\end{figure}
 

### 2020-10-13 shallow_water.tex
made minor edits

@@
 a magnitude of $\parameter_2$. We consider a parameter domain
-$(\parameter_1,\parameter_2)\in\parameterdomain = [3,9]\times [0.05,2]$.
+$(\parameter_1,\parameter_2)\in\parameterdomain = [3,9]\times [0.05,0.2]$.
 
@@
 \begin{enumerate}
-\item Set the parameter-grid to be $\mathcal{D}_{\text{train}}  \{ 3,6,7 \} \times \{0.05,0.125,0.2\}$. 
+\item Set the parameter-grid to be $\mathcal{D}_{\text{train}}  \{ 3,6,9 \} \times \{0.05,0.125,0.2\}$. 
 \item Execute Algorithm~\ref{alg:pod_tlp} with inputs $N_{\text{skip}} = 2$, $\stateInterceptArg{} = \bz$, $\mathcal{D}_{\text{train}}$, $\romdim = \{ 83,143 ,223 \}$, and $\mathbf{T} = \mathbf{M}_{\text{DG}}$ to construct POD bases capturing $99.9999\%, \; 99.99999\%,$ and $99.999999\%$ of the energy, respectively.
@@
 \begin{equation}\label{eq:lin_surrogate}
-\approxstateLR^n = {\approxstateLRIntercept}^n + \mu_0 \mathbf{c_1}^n + \mu_1 \mathbf{c_2}^n ,
+\approxstateLR^n = {\approxstateLRIntercept}^n + \mu_1 \mathbf{c_1}^n + \mu_2 \mathbf{c_2}^n ,
 \end{equation}
@@
 Figure~\ref{fig:rom_sols_swe1h} depicts the evolution of the surface height at $(x,y) = (-2.4834,-2.4834)$ as a function of time, while Figure~\ref{fig:rom_sols_swe1e} depicts the relative $\elltwo$-error as a function of time.
-We first observe that the linear regression surrogate model performs significantly worse than all ROMs; this result demonstrates the complex dynamics of the problem and highlights the utility of ROMs.
+We first observe that the linear regression surrogate model performs significantly worse than all ROMs; this result demonstrates the complex dynamics of the problem.
 Next, we observe that LSPG yields poor solutions for later time instances.
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_h_vs_t_K83.pdf}
-\caption{Surface height at $(x,y) = (-2.4834,-2.4834)$ as a function of time.}
+\caption{Surface height.}
 \label{fig:rom_sols_swe1h}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_error_vs_t_K83.pdf}
-\caption{Relative error as a function of time.}
+\caption{Relative error.}
 \label{fig:rom_sols_swe1e}
 \end{subfigure}
-\caption{Surface height (left) and relative error (right) as a function of time.} 
+\caption{Surface height at $(x,y) = (-2.4834,-2.4834)$ (left) and relative error (right) as a function of time.} 
 \label{fig:rom_sols_swe1}
@@
 
-Next, Figure~\ref{fig:rom_swe_metrics} depicts the impact of the window size on the relative error and residual norms.
+Figure~\ref{fig:rom_swe_metrics} depicts the impact of the window size on the relative error and residual norms.
 The time-integrated $\elltwo$-error of the linear surrogate model is additionally shown for reference.
 We make the same observations as before: increasing the window size leads to a monotonic decrease in the time-integrated residual norm, but not in the time-integrated error norm.
-We observe that the time integrated $\elltwo$-error is lowest for an intermeidate window size.
+We observe that the time integrated $\elltwo$-error is lowest for an intermediate window size.
 At $\Delta T = 1$, the time-integrated $\elltwo$-error of the WLS ROM is approximately half that of LSPG.
@@
 %\begin{subfigure}[t]{0.85\textwidth}
-\caption{Time-integrated performance metrics of the LSPG and WLS ROMs as a function of window size} 
+\caption{Time-integrated error metrics of the LSPG and WLS ROMs as a function of window size} 
 \label{fig:rom_swe_metrics}
@@
 
-Next, Figure~\ref{fig:rom_swe_timings} presents timings for the various ROMs as a function of the window size.
+Figure~\ref{fig:rom_swe_timings} presents timings for the various ROMs as a function of the window size.
 We note that, as described in Section~\ref{sec:swe_describe_roms}, all ROMs employ a Gauss--Newton solve with a Jacobian update frequency of $n_{\text{update}} = 10$; the sensitity of the results to this parameter are detailed in Appendix~\ref{appendix:gnupdate}.
-We again obsere that the cost of WLS grows with window size: WLS with $\Delta T = 10.0$ ($=500\Delta t$) incurs a 5x increase in computational cost with repsect to LSPG.
+We again observe that the cost of WLS grows with window size: WLS with $\Delta T = 10.0$ ($=500\Delta t$) incurs a 5x increase in computational cost with respect to LSPG.
 With respect to the FOM, we observe LSPG to result in a $20$x speedup, while WLS with the largest window size $\Delta T = 10.0$ yields a $4$x speedup.
@@
 We first observe that LSPG, and WLS with smaller window sizes, do not display monotonic convergence with increasing basis dimension.
-We expect that this lack of convergence is because of the fact that, as the basis dimension increases, the ROM trial space can represent sharper gradients.
+We expect that this lack of convergence is due to the fact that, as the basis dimension increases, the ROM trial space can represent sharper gradients.
 Without proper stabilization, this can yield spurious oscillations, as previously observed in Figure~\ref{fig:rom_hsols_swe}.
@@
 %\begin{subfigure}[t]{0.85\textwidth}
-\caption{Performance metrics of the LSPG and WLS ROMs as a function of basis dimension.} 
+\caption{Time-integrated error metrics of the LSPG and WLS ROMs as a function of basis dimension.} 
 \label{fig:rom_swe_converge}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_converge_error_pareto.pdf}
-\caption{Relative time-inegrated $\elltwo$-error.}
+%\caption{Relative time-inegrated $\elltwo$-error.}
 \end{subfigure}
@@
 \includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=1.0\linewidth]{figs/swe/swe_converge_resid_pareto.pdf}
-\caption{Time integrated residual norm.}
+%\caption{Time integrated residual norm.}
 \end{subfigure}
 %\begin{subfigure}[t]{0.85\textwidth}
-\caption{Performance metrics of the LSPG and WLS ROMs as a function of basis dimension.} 
+\caption{Pareto fronts for the time-integrated $\elltwo$-error (left) and time-integrated residual (right) for LSPG and WLS ROMs.} 
 \label{fig:rom_swe_pareto}
